### 2023-04-18 thesis/sections/2-background.tex
basic structure

\section{Background and Related Work}
\label{sec:background}
% subsections are \subsection{title}
%% subssubsections are \subsubsection{title}
%% numbering will work automatically

\cmtDK[inline]{7 pages}

\subsection{Dataset}
\cmtDK[inline]{\\
    - CLEVR \\
    ---- 3D objects \\
    ---- simple objects
}
\subsection{Language Games}
\cmtDK[inline]{\\
    - EGG \\
    - Resnet \\
    - CLEVR baselines \\
}

### 2023-04-18 thesis/sections/3-exp-setup.tex
basic structure

\section{Experimental Setup}
\label{sec:exp-setup}

\cmtDK[inline]{15 pages}
\cmtDK[inline]{
    - what is ambiguity? \\
    - control over dataset \\
    - target object/target group/non-target group: \\
    ---- 1. how many objects? \\
    ---- 2. how many shared attributes? \\
    ---- 3. which shared attributes? \\
    ---- 4. level of ambiguity \\
    - as close to CLEVR as possible \\
    ---- image size \\
    ---- number (10.000) \\
    ---- between 3 to 10 objects ??? \\
    - unambigous dataset: \\
    ---- target group = non-target group \\
    ---- all attributes shared, except for the color \\
    ---- \ \  --> describable with one 'word'
}

### 2023-04-18 thesis/sections/4-results.tex
basic structure

\section{Results and Discussion}
\label{sec:results}

\cmtDK[inline]{15 pages}

### 2023-04-18 source/preexperiments/data_readers.py
save outputs

@@ -42,14 +42,16 @@ class ClassifierDataset(Dataset):
         selected_scenes = random.sample(scenes, max_number_samples)
 
         for scene_file in selected_scenes:
-            with open(os.path.join(scenes_json_dir, scene_file), 'r', encoding='utf-8') as f:
+            with open(
+                os.path.join(scenes_json_dir, scene_file), "r", encoding="utf-8"
+            ) as f:
                 scene = json.load(f)
 
-            image = Image.open(image_path + scene['image_filename']).convert('RGB')
+            image = Image.open(image_path + scene["image_filename"]).convert("RGB")
 
             bounding_boxes = self._get_bounding_boxes(image, scene)
 
-            target_object = scene['groups']['target'][0]
+            target_object = scene["groups"]["target"][0]
             enumerated = list(enumerate(bounding_boxes))
             random.shuffle(enumerated)
 
@@ -64,17 +66,22 @@ class ClassifierDataset(Dataset):
         BOUNDING_BOX_SIZE = image.size[0] / 5
 
         object_bounding_boxes = []
-        for obj in scene['objects']:
-            x_center, y_center, _ = obj['pixel_coords']
-            bounding_box = image.crop((
-                x_center - BOUNDING_BOX_SIZE/2,
-                y_center - BOUNDING_BOX_SIZE/2,
-                x_center + BOUNDING_BOX_SIZE/2,
-                y_center + BOUNDING_BOX_SIZE/2
-            ))
+        for obj in scene["objects"]:
+            x_center, y_center, _ = obj["pixel_coords"]
+            bounding_box = image.crop(
+                (
+                    x_center - BOUNDING_BOX_SIZE / 2,
+                    y_center - BOUNDING_BOX_SIZE / 2,
+                    x_center + BOUNDING_BOX_SIZE / 2,
+                    y_center + BOUNDING_BOX_SIZE / 2,
+                )
+            )
             object_bounding_boxes.append(preprocess(bounding_box))
 
-        object_bounding_boxes.extend([torch.zeros_like(object_bounding_boxes[0])] * (10 - len(object_bounding_boxes)))
+        object_bounding_boxes.extend(
+            [torch.zeros_like(object_bounding_boxes[0])]
+            * (10 - len(object_bounding_boxes))
+        )
 
         return object_bounding_boxes
 
@@ -96,14 +103,18 @@ class AttentionDataset(Dataset):
         selected_scenes = random.sample(scenes, max_number_samples)
 
[…]

### 2023-04-18 source/preexperiments/train.py

@@ -2,14 +2,24 @@ import argparse
 import sys
 
 import torch
-from classification_models import (ResnetAttentionAttributeClassifier,
-                                   ResnetAttentionClassifier)
+from classification_models import (
+    ResnetAttentionAttributeClassifier,
+    ResnetAttentionClassifier,
+)
 from data_readers import AttentionAttributeDataset, AttentionDataset
-from torch import nn, optim
+from torch import Tensor, nn, optim
 from torch.utils.data import DataLoader, random_split
 from torcheval.metrics import BinaryAccuracy, Mean, MulticlassAccuracy
 
-if __name__ == '__main__':
+
+def save_to_csv(data, file):
+    with open(file, "w", encoding="utf-8") as f:
+        f.write("image_id,x,y\n")
+        for image_id, pixels in data:
+            f.write(f"{image_id},{pixels[0]},{pixels[1]}\n")
+
+
+if __name__ == "__main__":
     parser = argparse.ArgumentParser()
     parser.add_argument(
         "--scene_json_dir", type=str, default=None, help="Path to the scene json dir"
@@ -20,29 +30,27 @@ if __name__ == '__main__':
     parser.add_argument(
         "--max_samples", type=int, default=None, help="max samples to load"
     )
-    parser.add_argument(
-        "--epochs", type=int, default=None, help="number of epochs"
-    )
-    parser.add_argument(
-        "--device", type=str, default=None, help="cpu or cuda"
-    )
-    parser.add_argument(
-        "--batch_size", type=int, default=32, help="batch size"
-    )
+    parser.add_argument("--epochs", type=int, default=None, help="number of epochs")
+    parser.add_argument("--device", type=str, default=None, help="cpu or cuda")
+    parser.add_argument("--batch_size", type=int, default=32, help="batch size")
     args = parser.parse_args()
 
-    if args.device == 'cpu':
-        device = torch.device('cpu')
-    elif args.device == 'cuda':
-        device = torch.device('cuda:0')
+    if args.device == "cpu":
+        device = torch.device("cpu")
+    elif args.device == "cuda":
+        device = torch.device("cuda:0")
     else:
-        raise AttributeError('Device must be cpu or cuda')
+        raise AttributeError("Device must be cpu or cuda")
 
-    dataset = AttentionAttributeDataset(args.scene_json_dir, args.image_dir, args.max_samples)
+    dataset = AttentionAttributeDataset(
+        args.scene_json_dir, args.image_dir, args.max_samples
+    )
 
[…]

### 2023-06-07 source/language_games/data_readers.py
restruturing, setuptools, referential game

@@ -1,120 +1,0 @@
-import itertools
-import json
-import os
-import random
-from dataclasses import dataclass
-
-import torch
-from PIL import Image
-from torch.utils.data import Dataset
-from torchvision.models import ResNet50_Weights
-
-
-@dataclass
-class ReferentialGameSample:
-    image_1_id: str
-    image_2_id: str
-    image_1: torch.Tensor
-    image_2: torch.Tensor
-
-    # target
-    target_image: torch.Tensor
-
-
-class ReferentialGameDataset(Dataset):
-    def __init__(
-        self,
-        scenes_json_dir,
-        image_path="",
-        max_number_samples="",
-        preprocess=ResNet50_Weights.DEFAULT.transforms(),
-    ) -> None:
-        super().__init__()
-
-        self.samples: list[ReferentialGameSample] = []
-
-        scenes = os.listdir(scenes_json_dir)
-        print("loading scenes...")
-        loaded_scenes = []
-        for scene_file in scenes:
-            with open(
-                os.path.join(scenes_json_dir, scene_file), "r", encoding="utf-8"
-            ) as f:
-                loaded_scenes.append(json.load(f))
-
-        print("creating combinations...")
-        scene_combinations = list(
-            itertools.product(range(len(loaded_scenes)), repeat=2)
-        )
-        print("shuffling...")
-        random.shuffle(scene_combinations)
-
-        print("looping...")
-
-        for scene_1_index, scene_2_index in scene_combinations:
-            if len(self.samples) == max_number_samples:
-                break
-
-            scene_1 = loaded_scenes[scene_1_index]
-            scene_2 = loaded_scenes[scene_2_index]
-
-            target_object_1_index = scene_1["groups"]["target"][0]
-            target_object_2_index = scene_2["groups"]["target"][0]
-
-            target_object_1 = scene_1["objects"][target_object_1_index]
-            target_object_2 = scene_2["objects"][target_object_2_index]
-
-            # no matching attribute
-            if (
-                target_object_1["color"] == target_object_2["color"]
-                or target_object_1["shape"] == target_object_2["shape"]
-                or target_object_1["size"] == target_object_2["size"]
-            ):
-                continue
-
-            image_1 = Image.open(image_path + scene_1["image_filename"]).convert("RGB")
-            image_2 = Image.open(image_path + scene_2["image_filename"]).convert("RGB")
-            target_image = random.randint(0, 1)
-
-            self.samples.append(
[…]

### 2023-06-07 source/language_games/play.py

@@ -1,176 +1,0 @@
-import argparse
-import sys
-
-import egg.core as core
-import torch.nn.functional as F
-from data_readers import ReferentialGameDataset
-from models import ReferentialGameReceiver, ReferentialGameSender
-from torch.utils.data import DataLoader, random_split
-
-
-def classification_loss(
-    _sender_input,
-    _message,
-    _receiver_input,
-    receiver_output,
-    labels,
-    _aux_input,
-):
-    # in the discriminative case, accuracy is computed by comparing the index with highest score in Receiver output (a distribution of unnormalized
-    # probabilities over target poisitions) and the corresponding label read from input, indicating the ground-truth position of the target
-    acc = (receiver_output.argmax(dim=1) == labels).detach().float()
-    # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
-    loss = F.cross_entropy(receiver_output, labels, reduction="none")
-    return loss, {"acc": acc}
-
-
-def get_params(params):
-    parser = argparse.ArgumentParser()
-
-    # -- DATASET --
-    parser.add_argument(
-        "--scene_json_dir", type=str, default=None, help="Path to the scene json dir"
-    )
-    parser.add_argument(
-        "--image_dir", type=str, default=None, help="Path to the scene image dir"
-    )
-    parser.add_argument(
-        "--max_samples", type=int, default=None, help="max samples to load"
-    )
-
-    parser.add_argument(
-        "--validation_batch_size",
-        type=int,
-        default=0,
-        help="Batch size when processing validation data, whereas training data batch_size is controlled by batch_size (default: same as training data batch size)",
-    )
-    # arguments concerning the training method
-    parser.add_argument(
-        "--mode",
-        type=str,
-        default="rf",
-        help="Selects whether Reinforce or Gumbel-Softmax relaxation is used for training {rf, gs} (default: rf)",
-    )
-    parser.add_argument(
-        "--temperature",
-        type=float,
-        default=1.0,
-        help="GS temperature for the sender, only relevant in Gumbel-Softmax (gs) mode (default: 1.0)",
-    )
-    parser.add_argument(
-        "--sender_entropy_coeff",
-        type=float,
-        default=1e-1,
[…]

### 2023-06-07 source/mlt/language_games/data_readers.py

@@ -1,0 +1,189 @@
+import itertools
+import json
+import os
+import pickle
+import random
+from collections import defaultdict
+from dataclasses import dataclass
+
+import h5py
+import torch
+from mlt.preexperiments.data_readers import CaptionGeneratorDataset
+from PIL import Image
+from torch.utils.data import Dataset
+from torchvision.models import ResNet50_Weights
+
+
+@dataclass
+class ReferentialGameSample:
+    image_1_id: str
+    image_2_id: str
+    image_1: torch.Tensor
+    image_2: torch.Tensor
+
+    # target
+    target_image: torch.Tensor
+
+
+class LazaridouReferentialGameDataset(Dataset):
+    def __init__(self, feature_file_path, label_file_path, max_number_samples) -> None:
+        super().__init__()
+
+        self.samples: list[ReferentialGameSample] = []
+
+        fc = h5py.File(feature_file_path, "r")
+
+        data = torch.FloatTensor(list(fc["dataset_1"]))
+
+        with open(label_file_path, "rb") as f:
+            labels = pickle.load(f)
+
+        zipped = list(enumerate(zip(data, labels)))
+        concept_dict = defaultdict(list)
+        for image_id, (data, label) in zipped:
+            concept_dict[label].append((image_id, data))
+
+        for _ in range(max_number_samples):
+            concepts = random.sample(concept_dict.keys(), 2)
+
+            image_1_id, image_1 = random.choice(concept_dict[concepts[0]])
+
+            image_2_id, image_2 = random.choice(concept_dict[concepts[1]])
+
+            target_image = random.randint(0, 1)
+
+            self.samples.append(
+                ReferentialGameSample(
+                    image_1_id=image_1_id,
+                    image_2_id=image_2_id,
+                    image_1=image_1,
+                    image_2=image_2,
+                    target_image=target_image,
+                )
+            )
+
+    def __getitem__(self, index):
+        sample = self.samples[index]
+        sender_input = torch.stack((sample.image_1, sample.image_2))
+
+        receiver_input = [sample.image_2]
+        receiver_input.insert(sample.target_image, sample.image_1)
+        receiver_input = torch.stack(receiver_input)
+
+        target = sample.target_image
+
+        return (sender_input, target, receiver_input)
+
+    def __len__(self) -> int:
+        return len(self.samples)
+
+
+class BoundingBoxReferentialGameDataset(Dataset):
+    def __init__(
+        self,
+        scenes_json_dir,
+        image_path="",
+        max_number_samples=0,
[…]

### 2023-06-07 source/mlt/language_games/models.py

@@ -1,0 +1,111 @@
+import torch
+from torch import nn
+
+
+class ReferentialGameSender(nn.Module):
+    def __init__(self, hidden_size, embedding_dimension=256, vocab_size=100) -> None:
+        super().__init__()
+        self.image_encoder = nn.LazyLinear(embedding_dimension, bias=False)
+
+        self.conv2 = nn.LazyConv2d(
+            hidden_size,
+            kernel_size=(2, 1),
+            stride=(2, 1),
+            bias=False,
+        )
+
+        self.sigmoid = nn.Sigmoid()
+
+        self.conv3 = nn.LazyConv2d(
+            1, kernel_size=(hidden_size, 1), stride=(hidden_size, 1), bias=False
+        )
+        self.lin4 = nn.LazyLinear(vocab_size, bias=False)
+
+        self.softmax = nn.Softmax(dim=1)
+
+    def forward(self, x, _aux_input):
+        image_1 = self.image_encoder(x[:, 0])
+        image_2 = self.image_encoder(x[:, 1])
+
+        concatenated = torch.stack((image_1, image_2), dim=1)
+        conv = self.conv2(concatenated.unsqueeze(dim=1))
+
+        conv = conv.transpose(1, 2)
+        conv = self.conv3(conv)
+        squeezed = conv.squeeze()
+
+        fc = self.lin4(squeezed)
+
+        return self.softmax(fc)
+
+
+class ReferentialGameReceiver(nn.Module):
+    def __init__(self, hidden_size, embedding_dimension=256) -> None:
+        super().__init__()
+        self.image_encoder = nn.LazyLinear(embedding_dimension, bias=False)
+
+        self.linear_message = nn.LazyLinear(embedding_dimension, bias=False)
+
+        self.softmax = nn.Softmax(dim=1)
+
+    def forward(self, message, x, _aux_input):
+        image_1 = self.image_encoder(x[:, 0])
+
+        image_2 = self.image_encoder(x[:, 1])
+
+        message = self.linear_message(message)
+        dot_product_1 = torch.sum(message * image_1, dim=1)
+        dot_product_2 = torch.sum(message * image_2, dim=1)
+
+        output = torch.stack((dot_product_1, dot_product_2), dim=1)
+
+        return self.softmax(output)
+
+
+class MaskedDaleSender(nn.Module):
+    def __init__(self, image_encoder, masked_image_encoder) -> None:
+        super().__init__()
+        self.image_encoder = image_encoder
+        self.masked_image_encoder = masked_image_encoder
+
[…]

### 2023-06-07 source/mlt/language_games/play.py

@@ -1,0 +1,182 @@
+import argparse
+import sys
+
+import egg.core as core
+import torch.nn.functional as F
+from mlt.language_games.data_readers import LazaridouReferentialGameDataset
+from mlt.language_games.models import ReferentialGameReceiver, ReferentialGameSender
+from torch.utils.data import DataLoader, random_split
+
+
+def classification_loss(
+    _sender_input,
+    _message,
+    _receiver_input,
+    receiver_output,
+    labels,
+    _aux_input,
+):
+    # in the discriminative case, accuracy is computed by comparing the index with highest score in Receiver output (a distribution of unnormalized
+    # probabilities over target poisitions) and the corresponding label read from input, indicating the ground-truth position of the target
+    acc = (receiver_output.argmax(dim=1) == labels).detach().float()
+    # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
+    loss = F.cross_entropy(receiver_output, labels, reduction="none")
+    return loss, {"acc": acc}
+
+
+def get_params(params):
+    parser = argparse.ArgumentParser()
+
+    # -- DATASET --
+    parser.add_argument(
+        "--feature_file_path", type=str, default=None, help="Path to the scene json dir"
+    )
+    parser.add_argument(
+        "--label_file_path", type=str, default=None, help="Path to the scene image dir"
+    )
+    parser.add_argument(
+        "--max_samples", type=int, default=None, help="max samples to load"
+    )
+
+    parser.add_argument(
+        "--validation_batch_size",
+        type=int,
+        default=0,
+        help="Batch size when processing validation data, whereas training data batch_size is controlled by batch_size (default: same as training data batch size)",
+    )
+    # arguments concerning the training method
+    parser.add_argument(
+        "--mode",
+        type=str,
+        default="rf",
+        help="Selects whether Reinforce or Gumbel-Softmax relaxation is used for training {rf, gs} (default: rf)",
+    )
+    parser.add_argument(
+        "--temperature",
+        type=float,
+        default=1.0,
+        help="GS temperature for the sender, only relevant in Gumbel-Softmax (gs) mode (default: 1.0)",
+    )
+    parser.add_argument(
+        "--sender_entropy_coeff",
+        type=float,
+        default=1e-1,
[…]

### 2023-06-07 source/mlt/language_games/test.py

@@ -1,0 +1,17 @@
+import sys
+
+import egg
+import torch
+import torchvision.transforms as transforms
+from PIL import Image
+
+interaction = torch.load(sys.argv[1])
+
+transform = transforms.ToPILImage()
+
+for index, sender_input in enumerate(interaction.sender_input):
+    sender_input -= sender_input.min()
+    sender_input /= sender_input.max()
+
+    image: Image.Image = transform(sender_input)
+    image.save(f"out/images/interaction_{index}.jpg", format="JPEG")

### 2023-06-07 source/mlt/preexperiments/data_readers.py

@@ -1,0 +1,569 @@
+import itertools
+import json
+import os
+import random
+from abc import ABC, abstractmethod
+from dataclasses import dataclass
+from enum import Enum
+
+import torch
+from mlt.preexperiments.image_loader import ImageLoader
+from mlt.preexperiments.models import FeatureExtractor
+from PIL import Image
+from torch.utils.data import Dataset
+from torchvision import transforms
+from torchvision.models import ResNet101_Weights
+
+
+class Attribute(Enum):
+    @classmethod
+    def names(cls):
+        return list(map(lambda a: a.name, cls))
+
+
+class Shape(Attribute):
+    CUBE = 0
+    SPHERE = 1
+    CYLINDER = 2
+
+
+class Color(Attribute):
+    GRAY = 0
+    RED = 1
+    BLUE = 2
+    GREEN = 3
+    BROWN = 4
+    PURPLE = 5
+    CYAN = 6
+    YELLOW = 7
+
+
+class Size(Attribute):
+    SMALL = 0
+    LARGE = 1
+
+
+class PreprocessScratch:
+    def __init__(self, image_size):
+        self.transform = transforms.Compose(
+            [
+                transforms.Resize(image_size),
+                transforms.PILToTensor(),
+                transforms.ConvertImageDtype(torch.float),
+            ]
+        )
+        ratio = 1.5
+        self.resize_size = (image_size * ratio, image_size)
+        self.crop_size = None
+
+    def __call__(self, image):
+        return self.transform(image)
+
+
+class BoundingBoxClassifierDataset(Dataset):
+    """
+    Input:
+     - bounding boxes of all objects
+
+    Ouput:
+     - index of target bounding box
+    """
+
+    def __init__(
+        self,
+        scenes_json_dir,
+        image_path,
+        max_number_samples,
+        feature_extractor: FeatureExtractor = None,
+        preprocess=ResNet101_Weights.DEFAULT.transforms(),
+        device=torch.device("cpu"),
+    ) -> None:
+        super().__init__()
+
+        if feature_extractor is not None:
+            feature_extractor = feature_extractor.to(device)
+            feature_extractor.eval()
+
+        self.samples = []
+
+        scenes = os.listdir(scenes_json_dir)
+        print("sampling scenes...")
+        selected_scenes = random.sample(scenes, max_number_samples)
+
+        for scene_index, scene_file in enumerate(selected_scenes):
+            if scene_index % 50 == 0:
+                print(f"processing scene {scene_index}...", end="\r")
+
+            with open(
[…]

### 2023-06-07 source/mlt/preexperiments/feature_extractors.py

@@ -1,0 +1,231 @@
+import argparse
+import os
+from abc import ABC, abstractmethod
+
+import h5py
+import numpy as np
+import torch
+from PIL import Image
+from torch import nn
+from torchvision.models import ResNet101_Weights, VGG19_Weights, resnet101, vgg19
+
+
+class FeatureExtractor(ABC, nn.Module):
+    @property
+    @abstractmethod
+    def feature_shape(self):
+        ...
+
+
+class DummyFeatureExtractor(FeatureExtractor):
+    """Feature extractor that ouputs input unchanged
+
+    Use this in a model if the input to the model are already extraced features (e.g. from .h5 files
+    generated by this script) and not the images themselves.
+    """
+
+    @property
+    def feature_shape(self):
+        return (0,)
+
+    def forward(self, data):
+        return data
+
+
+class ResnetFeatureExtractor(FeatureExtractor):
+    def __init__(
+        self, pretrained=True, fine_tune=False, number_blocks=4, avgpool=True, fc=True
+    ) -> None:
+        super().__init__()
+
+        if pretrained:
+            resnet = resnet101(weights=ResNet101_Weights.IMAGENET1K_V2)
+        else:
+            resnet = resnet101()
+
+        if number_blocks > 4 or number_blocks < 1:
+            raise AttributeError("number of blocks need to be between 1 and 3")
+
+        self.resnet = nn.Sequential(*list(resnet.children())[:5])
+        self._feature_shape = (256, 56, 56)
+
+        if number_blocks > 1:
+            self.resnet.append(resnet.layer2)
+            self._feature_shape = (512, 28, 28)
+
+        if number_blocks > 2:
+            self.resnet.append(resnet.layer3)
+            self._feature_shape = (1024, 14, 14)
+
+        if number_blocks > 3:
+            self.resnet.append(resnet.layer4)
+            self._feature_shape = (2048, 7, 7)
+
+        if avgpool:
+            self.resnet.append(resnet.avgpool)
+            self._feature_shape = (2048, 1, 1)
+
+        if fc:
+            self.resnet.append(nn.Flatten())
+            self.resnet.append(resnet.fc)
+            self._feature_shape = (1000,)
+
+        if not fine_tune:
+            for param in self.resnet.parameters():
+                param.requires_grad = False
+            self.resnet.eval()
+
+    @property
+    def feature_shape(self):
+        return self._feature_shape
+
+    def forward(self, data):
+        return self.resnet(data)
+
+
[…]

### 2023-06-07 source/mlt/preexperiments/image_loader.py

@@ -1,0 +1,42 @@
+import os
+from abc import ABC, abstractmethod
+
+import h5py
+from PIL import Image
+
+
+class ImageLoader(ABC):
+    @abstractmethod
+    def get_image(self, image_id):
+        ...
+
+
+class ClevrImageLoader(ImageLoader):
+    def __init__(self, image_dir, preprocess) -> None:
+        self.image_dir = image_dir
+        self.preprocess = preprocess
+
+    def get_image(self, image_id):
+        image = Image.open(os.path.join(self.image_dir, image_id + ".png")).convert(
+            "RGB"
+        )
+        return image, self.preprocess(image), image.size
+
+
+class FeatureImageLoader(ImageLoader):
+    def __init__(self, feature_file, image_dir) -> None:
+        self.image_dir = image_dir
+
+        with h5py.File(feature_file, "r") as f:
+            feature_data_set = f["features"]
+            self.image_size = feature_data_set.attrs["image_size"]
+            self.features = list(feature_data_set)
+
+    def get_image(self, image_id):
+        image_index = int(image_id[-6:])
+
+        image = Image.open(os.path.join(self.image_dir, image_id + ".png")).convert(
+            "RGB"
+        )
+
+        return image, self.features[image_index], self.image_size

### 2023-06-07 source/mlt/preexperiments/models.py

@@ -1,0 +1,394 @@
+import torch
+from mlt.preexperiments.feature_extractors import FeatureExtractor
+from torch import nn
+
+
+class BoundingBoxClassifier(nn.Module):
+    """
+    Output:
+     - classified bounding box (10 dimensions)
+
+    Input:
+     - bounding boxes of objects
+    """
+
+    def __init__(self, feature_extractor: FeatureExtractor) -> None:
+        super().__init__()
+        self.feature_extractor = feature_extractor
+
+        self.classifier = nn.Sequential(
+            nn.Flatten(),
+            nn.LazyLinear(4096),
+            nn.ReLU(),
+            nn.Linear(4096, 10),
+            nn.Softmax(dim=1),
+        )
+
+    def forward(self, data):
+        data = data.permute(1, 0, 2, 3, 4)
+
+        if self.inputs_are_feature:
+            stacked = data
+        else:
+            after_feature_extractor = []
+            for bounding_box in data:
+                after_feature_extractor.append(self.feature_extractor(bounding_box))
+            stacked = torch.stack(after_feature_extractor)
+            stacked = stacked.permute(1, 0, 2, 3, 4)
+
+        classified = self.classifier(stacked)
+
+        return classified
+
+
+class CoordinatePredictor(nn.Module):
+    """
+    Output:
+     - x and y coordinates of target object
+
+    Input:
+     - image
+    """
+
+    def __init__(self, feature_extractor: FeatureExtractor) -> None:
+        super().__init__()
+        self.feature_extractor = feature_extractor
+
+        self.classifier = nn.Sequential(
+            nn.Flatten(),
+            nn.LazyLinear(1024),
+            nn.ReLU(),
+            nn.Linear(1024, 2),
+        )
+
+    def forward(self, data):
+        image, *_ = data
+
+        extracted_features = self.feature_extractor(image)
+        classified = self.classifier(extracted_features)
+
+        return classified
+
+
+class AttributeCoordinatePredictor(nn.Module):
+    """
+    Output:
+     - x and y coordinates of target object
+
+    Input:
+     - image
+     - attributes (shape, size, color)
+    """
+
+    def __init__(
+        self,
+        number_colors,
+        number_shapes,
+        number_sizes,
+        feature_extractor: FeatureExtractor,
+    ) -> None:
+        super().__init__()
+        self.feature_extractor = feature_extractor
+
[…]

### 2023-06-07 source/mlt/preexperiments/save.py

@@ -1,0 +1,103 @@
+import os
+from time import gmtime, strftime
+
+import torch
+
+
+class ModelSaver:
+    def __init__(self, out_dir, model_name, output_processor) -> None:
+        folder_name = f'{strftime("%Y-%m-%d_%H-%M-%S", gmtime())}_{model_name}'
+        self.directory = os.path.join(out_dir, folder_name)
+        if not os.path.exists(self.directory):
+            os.makedirs(self.directory)
+
+        self.output_processor = output_processor
+
+    def save_model(self, model, file_name):
+        torch.save(
+            model.state_dict(),
+            os.path.join(self.directory, file_name),
+        )
+
+    def save_log(self, log, file_name):
+        with open(os.path.join(self.directory, file_name), "w", encoding="utf-8") as f:
+            f.writelines(log)
+
+    def save_output(self, output, file_name):
+        processed_output = self.output_processor.process(output)
+        self._save_to_csv(
+            processed_output,
+            os.path.join(self.directory, file_name),
+        )
+
+    def _save_to_csv(self, data, file):
+        with open(file, "w", encoding="utf-8") as f:
+            for line in data:
+                f.write(",".join(line) + "\n")
+
+
+class StandardOutputProcessor:
+    def __init__(self, output_fields, dataset) -> None:
+        self.output_fields = output_fields
+        self.dataset = dataset
+
+    def process(self, outputs):
+        return [self.output_fields, *outputs]
+
+
+class BoundingBoxOutputProcessor(StandardOutputProcessor):
+    def process(self, outputs):
+        processed_output = [self.output_fields]
+
+        for image_id, output, ground_truth in outputs:
+            # test data
+            if output.dim() == 0:
+                index = output
+
+            # train data
+            else:
+                index = torch.max(output, dim=0).indices
+
+            processed_output.append((image_id, str(int(index)), str(int(ground_truth))))
+
+        return processed_output
+
+
+class PixelOutputProcessor(StandardOutputProcessor):
+    def process(self, outputs):
+        processed_output = [self.output_fields]
+        for image_id, pixels, ground_truth in outputs:
+            string_pixels = [str(p) for p in pixels.tolist()]
+            string_ground_truths = [str(p) for p in ground_truth.tolist()]
+            processed_output.append((image_id, *string_pixels, *string_ground_truths))
+
+        return processed_output
+
+
+class CaptionOutputProcessor(StandardOutputProcessor):
+    def process(self, outputs):
[…]

### 2023-06-07 source/mlt/preexperiments/test.py

@@ -1,0 +1,112 @@
+from abc import abstractmethod
+
+import torch
+from torcheval.metrics import (
+    BinaryAccuracy,
+    Mean,
+    MulticlassAccuracy,
+    MultilabelAccuracy,
+)
+
+
+class Tester:
+    @abstractmethod
+    def test(self, model, test_loader, device):
+        pass
+
+
+class CoordinatePredictorTester(Tester):
+    def test(self, model, test_loader, device):
+        model.eval()
+        accuracy = BinaryAccuracy(device=device)
+        mean = Mean(device=device)
+
+        test_outputs = []
+        for model_input, ground_truth, image_id in test_loader:
+            model_input = [t.to(device) for t in model_input]
+            ground_truth = ground_truth.to(device)
+
+            output = model(model_input).detach()
+            test_outputs.extend(zip(image_id, output, ground_truth))
+
+            distances = torch.diagonal(torch.cdist(output, ground_truth.float()))
+            mean.update(distances)
+
+            positives = torch.where(distances < 20, distances, 0)
+            accuracy.update(positives, torch.ones_like(positives))
+
+        return {
+            "accuracy": f"{accuracy.compute():.2f}",
+            "mean test loss": f"{mean.compute():.2f}",
+        }, test_outputs
+
+
+class BoundingBoxClassifierTester(Tester):
+    def test(self, model, test_loader, device):
+        model.eval()
+        accuracy = MulticlassAccuracy(device=device)
+
+        test_outputs = []
+        for model_input, ground_truth, image_id in test_loader:
+            model_input = model_input.to(device)
+            ground_truth = ground_truth.to(device)
+            output = model(model_input).detach()
+            max_indices = torch.max(output, dim=1)[1]
+            test_outputs.extend(zip(image_id, max_indices, ground_truth))
+
+            accuracy.update(max_indices, ground_truth)
+
+        return {
+            "accuracy": f"{accuracy.compute():.2f}",
+        }, test_outputs
+
+
+class CaptionGeneratorTester(Tester):
+    def test(self, model, test_loader, device):
+        model.eval()
+        accuracy = BinaryAccuracy(device=device)
+        hamming_accuracy = MulticlassAccuracy(device=device)
+        non_target_accuracy = BinaryAccuracy(device=device)
+
+        test_outputs = []
+        for model_input, ground_truth, image_id in test_loader:
+            image, _, non_target_captions, masked_image, *_ = model_input
+
+            image = image.to(device)
+            masked_image = masked_image.to(device)
[…]

### 2023-06-07 source/mlt/preexperiments/train.py

@@ -1,0 +1,393 @@
+import argparse
+from dataclasses import dataclass
+from typing import Callable
+
+import torch
+from mlt.preexperiments.data_readers import (
+    BasicImageMasker,
+    BoundingBoxClassifierDataset,
+    CaptionGeneratorDataset,
+    Color,
+    CoordinatePredictorDataset,
+    DaleCaptionAttributeEncoder,
+    OneHotAttributeEncoder,
+    PreprocessScratch,
+    Shape,
+    Size,
+)
+from mlt.preexperiments.feature_extractors import (
+    DummyFeatureExtractor,
+    ResnetFeatureExtractor,
+)
+from mlt.preexperiments.image_loader import ClevrImageLoader, FeatureImageLoader
+from mlt.preexperiments.models import (
+    AttributeCoordinatePredictor,
+    AttributeLocationCoordinatePredictor,
+    BoundingBoxClassifier,
+    CaptionDecoder,
+    CaptionGenerator,
+    CoordinatePredictor,
+    DaleAttributeCoordinatePredictor,
+    ImageEncoder,
+    MaskedCaptionGenerator,
+    MaskedCoordinatePredictor,
+)
+from mlt.preexperiments.save import (
+    BoundingBoxOutputProcessor,
+    CaptionOutputProcessor,
+    ModelSaver,
+    PixelOutputProcessor,
+    StandardOutputProcessor,
+)
+from mlt.preexperiments.test import (
+    BoundingBoxClassifierTester,
+    CaptionGeneratorTester,
+    CoordinatePredictorTester,
+    Tester,
+)
+from torch import nn, optim
+from torch.nn import Module
+from torch.utils.data import DataLoader, Dataset, random_split
+from torcheval.metrics import Mean
+from torchvision.models import ResNet101_Weights
+
+
+def pixel_loss(model_output, ground_truth):
+    loss = torch.diagonal(torch.cdist(model_output, ground_truth.float()))
+
+    return torch.mean(loss)
+
+
+@dataclass
+class ModelDefinition:
+    dataset: Dataset
+    dataset_args: dict
+    preprocess: Callable
+    model: Module
+    model_args: dict
+    loss_function: Callable
+    tester: Tester
+    output_processor: StandardOutputProcessor
+    output_processor_args: dict
+
+
+models = {
+    "coordinate_predictor": ModelDefinition(
+        dataset=CoordinatePredictorDataset,
+        dataset_args={},
[…]

### 2023-06-07 source/preexperiments/data_readers.py

@@ -1,564 +1,0 @@
-import itertools
-import json
-import os
-import random
-from abc import ABC, abstractmethod
-from dataclasses import dataclass
-from enum import Enum
-
-import torch
-from image_loader import ImageLoader
-from models import FeatureExtractor
-from PIL import Image
-from torch.utils.data import Dataset
-from torchvision import transforms
-from torchvision.models import ResNet101_Weights
-
-
-class Attribute(Enum):
-    @classmethod
-    def names(cls):
-        return list(map(lambda a: a.name, cls))
-
-
-class Shape(Attribute):
-    CUBE = 0
-    SPHERE = 1
-    CYLINDER = 2
-
-
-class Color(Attribute):
-    GRAY = 0
-    RED = 1
-    BLUE = 2
-    GREEN = 3
-    BROWN = 4
-    PURPLE = 5
-    CYAN = 6
-    YELLOW = 7
-
-
-class Size(Attribute):
-    SMALL = 0
-    LARGE = 1
-
-
-class PreprocessScratch:
-    def __init__(self, image_size):
-        self.transform = transforms.Compose(
-            [
-                transforms.Resize(image_size),
-                transforms.PILToTensor(),
-                transforms.ConvertImageDtype(torch.float),
-            ]
-        )
-        ratio = 1.5
-        self.resize_size = (image_size * ratio, image_size)
-        self.crop_size = None
-
-    def __call__(self, image):
-        return self.transform(image)
-
-
-class BoundingBoxClassifierDataset(Dataset):
-    """
-    Input:
-     - bounding boxes of all objects
-
-    Ouput:
-     - index of target bounding box
-    """
-
-    def __init__(
-        self,
-        scenes_json_dir,
-        image_path,
-        max_number_samples,
-        feature_extractor: FeatureExtractor = None,
-        preprocess=ResNet101_Weights.DEFAULT.transforms(),
-        device=torch.device("cpu"),
-    ) -> None:
-        super().__init__()
-
-        if feature_extractor is not None:
-            feature_extractor = feature_extractor.to(device)
-            feature_extractor.eval()
-
-        self.samples = []
-
-        scenes = os.listdir(scenes_json_dir)
-        print("sampling scenes...")
-        selected_scenes = random.sample(scenes, max_number_samples)
-
-        for scene_index, scene_file in enumerate(selected_scenes):
-            if scene_index % 50 == 0:
-                print(f"processing scene {scene_index}...", end="\r")
-
-            with open(
[…]

### 2023-06-07 source/preexperiments/feature_extractors.py

@@ -1,231 +1,0 @@
-import argparse
-import os
-from abc import ABC, abstractmethod
-
-import h5py
-import numpy as np
-import torch
-from PIL import Image
-from torch import nn
-from torchvision.models import ResNet101_Weights, VGG19_Weights, resnet101, vgg19
-
-
-class FeatureExtractor(ABC, nn.Module):
-    @property
-    @abstractmethod
-    def feature_shape(self):
-        ...
-
-
-class DummyFeatureExtractor(FeatureExtractor):
-    """Feature extractor that ouputs input unchanged
-
-    Use this in a model if the input to the model are already extraced features (e.g. from .h5 files
-    generated by this script) and not the images themselves.
-    """
-
-    @property
-    def feature_shape(self):
-        return (0,)
-
-    def forward(self, data):
-        return data
-
-
-class ResnetFeatureExtractor(FeatureExtractor):
-    def __init__(
-        self, pretrained=True, fine_tune=False, number_blocks=4, avgpool=True, fc=True
-    ) -> None:
-        super().__init__()
-
-        if pretrained:
-            resnet = resnet101(weights=ResNet101_Weights.IMAGENET1K_V2)
-        else:
-            resnet = resnet101()
-
-        if number_blocks > 4 or number_blocks < 1:
-            raise AttributeError("number of blocks need to be between 1 and 3")
-
-        self.resnet = nn.Sequential(*list(resnet.children())[:5])
-        self._feature_shape = (256, 56, 56)
-
-        if number_blocks > 1:
-            self.resnet.append(resnet.layer2)
-            self._feature_shape = (512, 28, 28)
-
-        if number_blocks > 2:
-            self.resnet.append(resnet.layer3)
-            self._feature_shape = (1024, 14, 14)
-
-        if number_blocks > 3:
-            self.resnet.append(resnet.layer4)
-            self._feature_shape = (2048, 7, 7)
-
-        if avgpool:
-            self.resnet.append(resnet.avgpool)
-            self._feature_shape = (2048, 1, 1)
-
-        if fc:
-            self.resnet.append(nn.Flatten())
-            self.resnet.append(resnet.fc)
-            self._feature_shape = (1000,)
-
-        if not fine_tune:
-            for param in self.resnet.parameters():
-                param.requires_grad = False
-            self.resnet.eval()
-
-    @property
-    def feature_shape(self):
-        return self._feature_shape
-
-    def forward(self, data):
-        return self.resnet(data)
-
-
[…]

### 2023-06-07 source/preexperiments/models.py

@@ -1,396 +1,0 @@
-import copy
-
-import torch
-from feature_extractors import FeatureExtractor
-from torch import nn
-
-
-class BoundingBoxClassifier(nn.Module):
-    """
-    Output:
-     - classified bounding box (10 dimensions)
-
-    Input:
-     - bounding boxes of objects
-    """
-
-    def __init__(self, feature_extractor: FeatureExtractor) -> None:
-        super().__init__()
-        self.feature_extractor = feature_extractor
-
-        self.classifier = nn.Sequential(
-            nn.Flatten(),
-            nn.LazyLinear(4096),
-            nn.ReLU(),
-            nn.Linear(4096, 10),
-            nn.Softmax(dim=1),
-        )
-
-    def forward(self, data):
-        data = data.permute(1, 0, 2, 3, 4)
-
-        if self.inputs_are_feature:
-            stacked = data
-        else:
-            after_feature_extractor = []
-            for bounding_box in data:
-                after_feature_extractor.append(self.feature_extractor(bounding_box))
-            stacked = torch.stack(after_feature_extractor)
-            stacked = stacked.permute(1, 0, 2, 3, 4)
-
-        classified = self.classifier(stacked)
-
-        return classified
-
-
-class CoordinatePredictor(nn.Module):
-    """
-    Output:
-     - x and y coordinates of target object
-
-    Input:
-     - image
-    """
-
-    def __init__(self, feature_extractor: FeatureExtractor) -> None:
-        super().__init__()
-        self.feature_extractor = feature_extractor
-
-        self.classifier = nn.Sequential(
-            nn.Flatten(),
-            nn.LazyLinear(1024),
-            nn.ReLU(),
-            nn.Linear(1024, 2),
-        )
-
-    def forward(self, data):
-        image, *_ = data
-
-        extracted_features = self.feature_extractor(image)
-        classified = self.classifier(extracted_features)
-
-        return classified
-
-
-class AttributeCoordinatePredictor(nn.Module):
-    """
-    Output:
-     - x and y coordinates of target object
-
-    Input:
-     - image
-     - attributes (shape, size, color)
-    """
-
-    def __init__(
-        self,
-        number_colors,
-        number_shapes,
-        number_sizes,
-        feature_extractor: FeatureExtractor,
-    ) -> None:
-        super().__init__()
-        self.feature_extractor = feature_extractor
-
[…]

### 2023-06-07 source/preexperiments/object_detection.py

@@ -1,39 +1,0 @@
-import sys
-
-import torch
-from PIL import Image
-from transformers import DetrForObjectDetection, DetrImageProcessor
-
-file_name = sys.argv[1]
-number_of_boxes = int(sys.argv[2])
-
-image = Image.open(file_name).convert("RGB")
-
-processor = DetrImageProcessor.from_pretrained("facebook/detr-resnet-101")
-model = DetrForObjectDetection.from_pretrained("facebook/detr-resnet-101")
-
-inputs = processor(images=image, return_tensors="pt")
-outputs = model(**inputs)
-
-# convert outputs (bounding boxes and class logits) to COCO API
-# let's only keep detections with score > 0.9
-target_sizes = torch.tensor([image.size[::-1]])
-results = processor.post_process_object_detection(
-    outputs, target_sizes=target_sizes, threshold=0
-)[0]
-
-print(f"number of boxes: {number_of_boxes}")
-
-sorted_zipped = sorted(
-    zip(results["scores"], results["labels"], results["boxes"]),
-    key=lambda x: x[0],
-    reverse=True,
-)
-for index, (score, label, box) in enumerate(sorted_zipped):
-    if index == number_of_boxes:
-        break
-    box = [round(i, 2) for i in box.tolist()]
-    print(
-        f"Detected {model.config.id2label[label.item()]} with confidence "
-        f"{round(score.item(), 6)} at location {box}"
-    )

### 2023-06-07 source/preexperiments/save.py

@@ -1,103 +1,0 @@
-import os
-from time import gmtime, strftime
-
-import torch
-
-
-class ModelSaver:
-    def __init__(self, out_dir, model_name, output_processor) -> None:
-        folder_name = f'{strftime("%Y-%m-%d_%H-%M-%S", gmtime())}_{model_name}'
-        self.directory = os.path.join(out_dir, folder_name)
-        if not os.path.exists(self.directory):
-            os.makedirs(self.directory)
-
-        self.output_processor = output_processor
-
-    def save_model(self, model, file_name):
-        torch.save(
-            model.state_dict(),
-            os.path.join(self.directory, file_name),
-        )
-
-    def save_log(self, log, file_name):
-        with open(os.path.join(self.directory, file_name), "w", encoding="utf-8") as f:
-            f.writelines(log)
-
-    def save_output(self, output, file_name):
-        processed_output = self.output_processor.process(output)
-        self._save_to_csv(
-            processed_output,
-            os.path.join(self.directory, file_name),
-        )
-
-    def _save_to_csv(self, data, file):
-        with open(file, "w", encoding="utf-8") as f:
-            for line in data:
-                f.write(",".join(line) + "\n")
-
-
-class StandardOutputProcessor:
-    def __init__(self, output_fields, dataset) -> None:
-        self.output_fields = output_fields
-        self.dataset = dataset
-
-    def process(self, outputs):
-        return [self.output_fields, *outputs]
-
-
-class BoundingBoxOutputProcessor(StandardOutputProcessor):
-    def process(self, outputs):
-        processed_output = [self.output_fields]
-
-        for image_id, output, ground_truth in outputs:
-            # test data
-            if output.dim() == 0:
-                index = output
-
-            # train data
-            else:
-                index = torch.max(output, dim=0).indices
-
-            processed_output.append((image_id, str(int(index)), str(int(ground_truth))))
-
-        return processed_output
-
-
-class PixelOutputProcessor(StandardOutputProcessor):
-    def process(self, outputs):
-        processed_output = [self.output_fields]
-        for image_id, pixels, ground_truth in outputs:
-            string_pixels = [str(p) for p in pixels.tolist()]
-            string_ground_truths = [str(p) for p in ground_truth.tolist()]
-            processed_output.append((image_id, *string_pixels, *string_ground_truths))
-
-        return processed_output
-
-
-class CaptionOutputProcessor(StandardOutputProcessor):
-    def process(self, outputs):
[…]

### 2023-06-07 source/preexperiments/test.py

@@ -1,112 +1,0 @@
-from abc import abstractmethod
-
-import torch
-from torcheval.metrics import (
-    BinaryAccuracy,
-    Mean,
-    MulticlassAccuracy,
-    MultilabelAccuracy,
-)
-
-
-class Tester:
-    @abstractmethod
-    def test(self, model, test_loader, device):
-        pass
-
-
-class CoordinatePredictorTester(Tester):
-    def test(self, model, test_loader, device):
-        model.eval()
-        accuracy = BinaryAccuracy(device=device)
-        mean = Mean(device=device)
-
-        test_outputs = []
-        for model_input, ground_truth, image_id in test_loader:
-            model_input = [t.to(device) for t in model_input]
-            ground_truth = ground_truth.to(device)
-
-            output = model(model_input).detach()
-            test_outputs.extend(zip(image_id, output, ground_truth))
-
-            distances = torch.diagonal(torch.cdist(output, ground_truth.float()))
-            mean.update(distances)
-
-            positives = torch.where(distances < 20, distances, 0)
-            accuracy.update(positives, torch.ones_like(positives))
-
-        return {
-            "accuracy": f"{accuracy.compute():.2f}",
-            "mean test loss": f"{mean.compute():.2f}",
-        }, test_outputs
-
-
-class BoundingBoxClassifierTester(Tester):
-    def test(self, model, test_loader, device):
-        model.eval()
-        accuracy = MulticlassAccuracy(device=device)
-
-        test_outputs = []
-        for model_input, ground_truth, image_id in test_loader:
-            model_input = model_input.to(device)
-            ground_truth = ground_truth.to(device)
-            output = model(model_input).detach()
-            max_indices = torch.max(output, dim=1)[1]
-            test_outputs.extend(zip(image_id, max_indices, ground_truth))
-
-            accuracy.update(max_indices, ground_truth)
-
-        return {
-            "accuracy": f"{accuracy.compute():.2f}",
-        }, test_outputs
-
-
-class CaptionGeneratorTester(Tester):
-    def test(self, model, test_loader, device):
-        model.eval()
-        accuracy = BinaryAccuracy(device=device)
-        hamming_accuracy = MulticlassAccuracy(device=device)
-        non_target_accuracy = BinaryAccuracy(device=device)
-
-        test_outputs = []
-        for model_input, ground_truth, image_id in test_loader:
-            image, _, non_target_captions, masked_image, *_ = model_input
-
-            image = image.to(device)
-            masked_image = masked_image.to(device)
[…]

### 2023-06-07 source/preexperiments/train.py

@@ -1,387 +1,0 @@
-import argparse
-from dataclasses import dataclass
-from test import (
-    BoundingBoxClassifierTester,
-    CaptionGeneratorTester,
-    CoordinatePredictorTester,
-    Tester,
-)
-from typing import Callable
-
-import torch
-from data_readers import (
-    BasicImageMasker,
-    BoundingBoxClassifierDataset,
-    CaptionGeneratorDataset,
-    Color,
-    CoordinatePredictorDataset,
-    DaleCaptionAttributeEncoder,
-    OneHotAttributeEncoder,
-    PreprocessScratch,
-    Shape,
-    Size,
-)
-from feature_extractors import DummyFeatureExtractor, ResnetFeatureExtractor
-from image_loader import ClevrImageLoader, FeatureImageLoader
-from models import (
-    AttributeCoordinatePredictor,
-    AttributeLocationCoordinatePredictor,
-    BoundingBoxClassifier,
-    CaptionDecoder,
-    CaptionGenerator,
-    CoordinatePredictor,
-    DaleAttributeCoordinatePredictor,
-    ImageEncoder,
-    MaskedCaptionGenerator,
-    MaskedCoordinatePredictor,
-)
-from save import (
-    BoundingBoxOutputProcessor,
-    CaptionOutputProcessor,
-    ModelSaver,
-    PixelOutputProcessor,
-    StandardOutputProcessor,
-)
-from torch import nn, optim
-from torch.nn import Module
-from torch.utils.data import DataLoader, Dataset, random_split
-from torcheval.metrics import Mean
-from torchvision.models import ResNet101_Weights
-
-
-def pixel_loss(model_output, ground_truth):
-    loss = torch.diagonal(torch.cdist(model_output, ground_truth.float()))
-
-    return torch.mean(loss)
-
-
-@dataclass
-class ModelDefinition:
-    dataset: Dataset
-    dataset_args: dict
-    preprocess: Callable
-    model: Module
-    model_args: dict
-    loss_function: Callable
-    tester: Tester
-    output_processor: StandardOutputProcessor
-    output_processor_args: dict
-
-
-models = {
-    "coordinate_predictor": ModelDefinition(
-        dataset=CoordinatePredictorDataset,
-        dataset_args={},
-        preprocess=ResNet101_Weights.IMAGENET1K_V2.transforms(),
-        model=CoordinatePredictor,
-        model_args={"feature_extractor": DummyFeatureExtractor()},
[…]

### 2023-06-22 thesis/sections/3-exp-setup.tex
restructure
tesis description of dataset

@@
 \cmtDK[inline]{15 pages}
-\cmtDK[inline]{
+\cmtDK[inline]{\\
     - what is ambiguity? \\
@@
 }
+In this section I will describe which experiments were conducted to answer the \cmtDK{?}{questions} from the previous sections as well as the setup of these experiments.
+This will be done in two parts.
+The first part contains the creation of the datasets that will be used in the experiments.
+In the second part I will go deeper into discussing each experiment, including the necessity of the experiment, the specifities of the dataset, as well as the final architecture of the models.
+
+\subsection{Creation of the dataset}
+The basis for my dataset is the CLEVR dataset (\cite{Johnson2016}).
+This dataset includes 3D-generated images depicting scenes with different kind of objects.
+Each of these objects has different combinations attributes, such as \emph{shape}, \emph{color}, \emph{size} and \emph{material}.
+The possible values of these attributes are listed in Table \ref{tab:clevr-attributes}.
+Three to ten objects are randomly placed into the scene and assigned with random attributes.
+To enhance realism and reduce ambiguity, objects do not intersect, have a certain distance from each other, and are at least partially visible.
+Figure \ref{fig:clevr-example} shows an example of a generated image in the CLEVR dataset.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cccc}
+        \toprule
+        \textbf{ shape } & \textbf{ color } & \textbf{ size } & \textbf{ material } \\
+        cube             & gray             & small           & rubber              \\
+        sphere           & red              & large           & metal               \\
+        cylinder         & blue                                                     \\
+                         & green                                                    \\
+                         & brown                                                    \\
+                         & purple                                                   \\
+                         & cyan                                                     \\
+                         & yellow                                                   \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attributes of objects in the CLEVR dataset}
+    \label{tab:clevr-attributes}
+\end{table}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.8\linewidth]{figures/CLEVR_example.png}
+    \caption{Example of a generated image in the CLEVR dataset}
+    \label{fig:clevr-example}
+\end{figure}
+
+Furthermore, the dataset contains information about each scene.
+This includes the selected attributes for each object as well as the exact position of the centers of all the objects, both 3D-coordinates in the 3D scene and 2D-coordinates in the final rendered image.
+In addition, simple spatial relations (in front of, behind, left, right) between the objects are calculated and stored.
+These are simply based on the 3D-coordinates of the objects from the perspective of the camera position.
+
+\cmtDK[inline]{ambiguity}
+
+This research investigates how agents communicate about the relations of objects seen in images.
+For that reason the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
+Following, I extended the source code to generated images for the CLEVR dataset.\footnote{The extended source code can be found on GitHub: \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}} To simplify the generation, I only focused on the attributes \emph{shape}, \emph{size} and \emph{color}.
+The \emph{material} is always the same for all objects in a generated image.
+There were three main extensions to the code:
+
+First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
+The target object is the main object in the scene that should be identified and communicated by the agents.
+All other objects and their relations are based on this target object.
+The target group contains similar objects to the target object.
+These are objects that the agents need to discriminate the target object from.
+Finally, the distractors are objects that add noise to the scene and should make it more complex and are expected to teach the agents more precise descriptions.
+The number of the objects in both group can be controlled.
+
+In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
+The relation is defined as \textbf{how many} attributes of the target object are identical with objects in the target group and distractors respectively.
+For example the target object is a \emph{small red cube}.
+If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere} but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
+The number of shared attributes can also be set to a range to make the discrimination task more challenging.
+
+Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
+For example, it can be defined to have the same size for objects in the target group, but have randomly selected different shapes and color.
+This allows a very controlled generation of relations between the objects in the scene.
+Figure \ref{fig:clevr-extended_example} shows one generated image with this extended source code.
+Here, the target object is the large purple cylinder.
+The target group contains four objects that share zero to a maximum of two attributes.
+It is not controlled, which attributes are shared (they are selected randomly).
+There are no distractor objects.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}
+    \caption{Example of a generated image with the extended code}
+    \label{fig:clevr-extended_example}
+\end{figure}
+
+For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
+The size of the generated images is 480x320 pixels.
+10.000 images are created for each of the datasets.
+Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance and are partially visible.
+
+\subsection{Description of experiments}

### 2023-06-24 thesis/sections/3-exp-setup.tex
thesis

@@
     - what is ambiguity? \\
-    - control over dataset \\
-    - target object/target group/non-target group: \\
-    ---- 1. how many objects? \\
-    ---- 2. how many shared attributes? \\
-    ---- 3. which shared attributes? \\
-    ---- 4. level of ambiguity \\
-    - as close to CLEVR as possible \\
-    ---- image size \\
-    ---- number (10.000) \\
-    ---- between 3 to 10 objects ??? \\
     - unambigous dataset: \\
@@
 \subsection{Description of experiments}
+Language games are very complex setups for machine learning models.
+The models need to solve multiple tasks at the same time to solve the overall problem.
+For instance in a simple setup of a game two agents are involved.
+The first agent, the sender is shown a scene with objects and needs to communicate on target object to the other agent, the receiver.
+The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
+In this case, the sender first needs to learn to encode the scene as well as the information about the target object into its own \cmtDK{?}{space}.
+In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
+The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene.
+And finally it needs to learn how to identify the target object with this information.
+
+For this reason I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.
+This will give me a very detailed overview, where the models struggle to learn and in which ways they can be improved.
+Mainly, I separated the tasks into language games with two agents (final experiments) and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone (pre-experiments).
+With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
+
+\subsubsection{Pre-experiments without language games}
+\paragraph{coordinate predictor}
+\cmtDK[inline]{\\
+    - pretrained ResNet/VGG\\
+    - scratch \\
+    - encoding of attributes (one-hot) \\
+    - encoding of attributes (dale) \\
+    - encoding of locations \\
+    - masking}
+\paragraph{bounding box classifier}
+
+\paragraph{caption generator}
+\cmtDK[inline]{\\
+    - dale \\
+    - masked}
+\subsubsection{Language games}

### 2023-06-26 thesis/sections/3-exp-setup.tex
thesis

@@
 \subsubsection{Pre-experiments without language games}
+The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
+Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge.
+Here, I structured the experiments into three levels.
+In the first level, the model are trained to learn the position of objects in the image and attend to specific regions of the image.
+In the second level, the models are trained to differentiate objects in the image from each other.
+The last level adds language to the experiment, more specifically the models learn to caption and describe objects in and image.
+These combined experiments should lay the basis for how to build up the agents in the language games.
+
 \paragraph{coordinate predictor}
+\cmtDK[inline]{talk about why coordinates are necessary for language games}
 \cmtDK[inline]{\\
@@
     - masking}
+
+For these experiments, the model is trained to predict the location of an object for a given image.
+In the simplest setup, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+Here the image is first passed through a \emph{feature extractor}.
+These feature extractors are a deep neural model that are designed to extract important information from images.
+Here, I chose ResNet (\cite{He2016}) and VGG (\cite{Simonyan2015}) as a basis.
+Both of those models can be used with pretrained weights, trained on classifying images from big datasets.
+Since these pretrained weights are based on a very different task than in this thesis, I experimented with different architectures of these feature extractors.
+Namely, I used both VGG19 and ResNet101 with random initializations and pretrained on the classifying task.
+Furthermore, I removed and/or replaced the last layers of either network.
+The idea behind this was, that \cmtDK{cite}{the models learn more general features in the first layers, while the last layers are fine-tuned more specifically to this task.}
+
 \paragraph{bounding box classifier}
@@
     - masked}
+
 \subsubsection{Language games}

### 2023-06-27 paper/paper.tex
acl template for paper

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage[review]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{Instructions for *ACL Proceedings}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\}

\begin{document}
\maketitle
\begin{abstract}
    This document is a supplement to the general instructions for *ACL authors. It contains instructions for using the \LaTeX{} style files for ACL conferences.
    The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
    These instructions should be used both for papers submitted for review and for final versions of accepted papers.
\end{abstract}

\section{Introduction}

These instructions are for authors submitting papers to *ACL conferences using \LaTeX. They are not self-contained. All authors must follow the general instructions for *ACL proceedings,\footnote{\url{http://acl-org.github.io/ACLPUB/formatting.html}} and this document contains additional instructions for the \LaTeX{} style files.

The templates include the \LaTeX{} source of this document (\texttt{acl.tex}),
the \LaTeX{} style file used to format it (\texttt{acl.sty}),
an ACL bibliography style (\texttt{acl\_natbib.bst}),
an example bibliography (\texttt{custom.bib}),
and the bibliography for the ACL Anthology (\texttt{anthology.bib}).

\section{Engines}

To produce a PDF file, pdf\LaTeX{} is strongly recommended (over original \LaTeX{} plus dvips+ps2pdf or dvipdf). Xe\LaTeX{} also produces PDF files, and is especially suitable for text in non-Latin scripts.

\section{Preamble}

The first line of the file must be
\begin{quote}
    \begin{verbatim}
\documentclass[11pt]{article}
\end{verbatim}
\end{quote}

To load the style file in the review version:
\begin{quote}
    \begin{verbatim}
\usepackage[review]{acl}
\end{verbatim}
\end{quote}
For the final version, omit the \verb|review| option:
\begin{quote}
    \begin{verbatim}
\usepackage{acl}
\end{verbatim}
\end{quote}

To use Times Roman, put the following in the preamble:
\begin{quote}
    \begin{verbatim}
\usepackage{times}
\end{verbatim}
\end{quote}
(Alternatives like txfonts or newtx are also acceptable.)

Please see the \LaTeX{} source of this document for comments on other packages that may be useful.

Set the title and author using \verb|\title| and \verb|\author|. Within the author list, format multiple authors using \verb|\and| and \verb|\And| and \verb|\AND|; please see the \LaTeX{} source for examples.

By default, the box containing the title and author names is set to the minimum of 5 cm. If you need more space, include the following in the preamble:
\begin{quote}
    \begin{verbatim}
\setlength\titlebox{<dim>}
\end{verbatim}
\end{quote}
where \verb|<dim>| is replaced with a length. Do not set this length smaller than 5 cm.

\section{Document Body}

\subsection{Footnotes}

Footnotes are inserted with the \verb|\footnote| command.\footnote{This is a footnote.}

\subsection{Tables and figures}

See Table~\ref{tab:accents} for an example of a table and its caption.
\textbf{Do not override the default caption sizes.}

\begin{table}
    \centering
    \begin{tabular}{lc}
        \hline
        \textbf{Command} & \textbf{Output} \\
        \hline
        \verb|{\"a}|     & {\"a}           \\
        \verb|{\^e}|     & {\^e}           \\
        \verb|{\`i}|     & {\`i}           \\
        \verb|{\.I}|     & {\.I}           \\
        \verb|{\o}|      & {\o}            \\
        \verb|{\'u}|     & {\'u}           \\
        \verb|{\aa}|     & {\aa}           \\\hline
    \end{tabular}
    \begin{tabular}{lc}
        \hline
        \textbf{Command} & \textbf{Output} \\
        \hline
        \verb|{\c c}|    & {\c c}          \\
        \verb|{\u g}|    & {\u g}          \\
        \verb|{\l}|      & {\l}            \\
        \verb|{\~n}|     & {\~n}           \\
        \verb|{\H o}|    & {\H o}          \\
        \verb|{\v r}|    & {\v r}          \\
        \verb|{\ss}|     & {\ss}           \\
        \hline
    \end{tabular}
    \caption{Example commands for accented characters, to be used in, \emph{e.g.}, Bib\TeX{} entries.}
    \label{tab:accents}
\end{table}

\subsection{Hyperlinks}

Users of older versions of \LaTeX{} may encounter the following error during compilation:
\begin{quote}
    \tt\verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.
\end{quote}
This happens when pdf\LaTeX{} is used and a citation splits across a page boundary. The best way to fix this is to upgrade \LaTeX{} to 2018-12-01 or later.

\subsection{Citations}

\begin{table*}
    \centering
    \begin{tabular}{lll}
        \hline
        \textbf{Output}           & \textbf{natbib command} & \textbf{Old ACL-style command} \\
        \hline
        \citep{Gusfield:97}       & \verb|\citep|           & \verb|\cite|                   \\
        \citealp{Gusfield:97}     & \verb|\citealp|         & no equivalent                  \\
        \citet{Gusfield:97}       & \verb|\citet|           & \verb|\newcite|                \\
        \citeyearpar{Gusfield:97} & \verb|\citeyearpar|     & \verb|\shortcite|              \\
        \hline
    \end{tabular}
    \caption{\label{citation-guide}
        Citation commands supported by the style file.
        The style is based on the natbib package and supports all natbib citation commands.
        It also supports commands defined in previous ACL style files for compatibility.
    }
\end{table*}

Table~\ref{citation-guide} shows the syntax supported by the style files.
We encourage you to use the natbib styles.
You can use the command \verb|\citet| (cite in text) to get ``author (year)'' citations, like this citation to a paper by \citet{Gusfield:97}.
You can use the command \verb|\citep| (cite in parentheses) to get ``(author, year)'' citations \citep{Gusfield:97}.
You can use the command \verb|\citealp| (alternative cite without parentheses) to get ``author, year'' citations, which is useful for using citations within parentheses (e.g. \citealp{Gusfield:97}).

\subsection{References}

\nocite{Ando2005,andrew2007scalable,rasooli-tetrault-2015}

The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
If your own bib file is named \texttt{custom.bib}, then placing the following before any appendices in your \LaTeX{} file will generate the references section for you:
\begin{quote}
    \begin{verbatim}
\bibliography{custom}
\end{verbatim}
\end{quote}

You can obtain the complete ACL Anthology as a Bib\TeX{} file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
To include both the Anthology and your own .bib file, use the following instead of the above.
\begin{quote}
    \begin{verbatim}
\bibliography{anthology,custom}
\end{verbatim}
\end{quote}

Please see Section~\ref{sec:bibtex} for information on preparing Bib\TeX{} files.

\subsection{Appendices}

Use \verb|\appendix| before any appendix section to switch the section numbering over to letters. See Appendix~\ref{sec:appendix} for an example.

\section{Bib\TeX{} Files}
\label{sec:bibtex}

Unicode cannot be used in Bib\TeX{} entries, and some ways of typing special characters can disrupt Bib\TeX's alphabetization. The recommended way of typing special characters is shown in Table~\ref{tab:accents}.

Please ensure that Bib\TeX{} records contain DOIs or URLs when possible, and for all the ACL materials that you reference.
Use the \verb|doi| field for DOIs and the \verb|url| field for URLs.
If a Bib\TeX{} entry has a URL or DOI field, the paper title in the references section will appear as a hyperlink to the paper, using the hyperref \LaTeX{} package.

\section*{Acknowledgements}

This document has been adapted
by Steven Bethard, Ryan Cotterell and Rui Yan
from the instructions for earlier ACL and NAACL proceedings, including those for
ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
NAACL 2019 by Stephanie Lukin and Alla Roskovskaya,
ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu,
NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
ACL 2017 by Dan Gildea and Min-Yen Kan,
NAACL 2017 by Margaret Mitchell,
ACL 2012 by Maggie Li and Michael White,
ACL 2010 by Jing-Shin Chang and Philipp Koehn,
ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui,
ACL 2005 by Hwee Tou Ng and Kemal Oflazer,
ACL 2002 by Eugene Charniak and Dekang Lin,
and earlier ACL and EACL formats written by several people, including
John Chen, Henry S. Thompson and Donald Walker.
Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.

% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}

\appendix

\section{Example Appendix}
\label{sec:appendix}

This is an appendix.

\end{document}

### 2023-06-27 paper/paper.tex
adapted template

@@
 
-\author{First Author \\
+\author{Dominik Künkele \\
   Affiliation / Address line 1 \\
@@
   Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
+  \texttt{dominik.kuenkele@outlook.com} \\\And
   Second Author \\
@@
 \begin{abstract}
-    This document is a supplement to the general instructions for *ACL authors. It contains instructions for using the \LaTeX{} style files for ACL conferences.
-    The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
-    These instructions should be used both for papers submitted for review and for final versions of accepted papers.
+
 \end{abstract}
@@
 
-These instructions are for authors submitting papers to *ACL conferences using \LaTeX.
-They are not self-contained.
-All authors must follow the general instructions for *ACL proceedings,\footnote{\url{http://acl-org.github.io/ACLPUB/formatting.html}} and this document contains additional instructions for the \LaTeX{} style files.
+\section{Background and previous work}
 
-The templates include the \LaTeX{} source of this document (\texttt{acl.tex}),
-the \LaTeX{} style file used to format it (\texttt{acl.sty}),
-an ACL bibliography style (\texttt{acl\_natbib.bst}),
-an example bibliography (\texttt{custom.bib}),
-and the bibliography for the ACL Anthology (\texttt{anthology.bib}).
+\section{Materials and methods}
 
-\section{Engines}
+\section{Results}
 
-To produce a PDF file, pdf\LaTeX{} is strongly recommended (over original \LaTeX{} plus dvips+ps2pdf or dvipdf).
-Xe\LaTeX{} also produces PDF files, and is especially suitable for text in non-Latin scripts.
+\section{Discussion}
 
-\section{Preamble}
-
-The first line of the file must be
-\begin{quote}
-    \begin{verbatim}
-\documentclass[11pt]{article}
-\end{verbatim}
-\end{quote}
-
-To load the style file in the review version:
-\begin{quote}
-    \begin{verbatim}
-\usepackage[review]{acl}
-\end{verbatim}
-\end{quote}
-For the final version, omit the \verb|review| option:
-\begin{quote}
-    \begin{verbatim}
-\usepackage{acl}
-\end{verbatim}
-\end{quote}
-
-To use Times Roman, put the following in the preamble:
-\begin{quote}
-    \begin{verbatim}
-\usepackage{times}
-\end{verbatim}
-\end{quote}
-(Alternatives like txfonts or newtx are also acceptable.)
-
-Please see the \LaTeX{} source of this document for comments on other packages that may be useful.
-
-Set the title and author using \verb|\title| and \verb|\author|.
-Within the author list, format multiple authors using \verb|\and| and \verb|\And| and \verb|\AND|; please see the \LaTeX{} source for examples.
-
-By default, the box containing the title and author names is set to the minimum of 5 cm. If you need more space, include the following in the preamble:
-\begin{quote}
-    \begin{verbatim}
-\setlength\titlebox{<dim>}
-\end{verbatim}
-\end{quote}
-where \verb|<dim>| is replaced with a length. Do not set this length smaller than 5 cm.
-
-\section{Document Body}
-
-\subsection{Footnotes}
-
-Footnotes are inserted with the \verb|\footnote| command.\footnote{This is a footnote.}
-
-\subsection{Tables and figures}
-
-See Table~\ref{tab:accents} for an example of a table and its caption.
-\textbf{Do not override the default caption sizes.}
-
-\begin{table}
-    \centering
-    \begin{tabular}{lc}
-        \hline
-        \textbf{Command} & \textbf{Output} \\
-        \hline
-        \verb|{\"a}|     & {\"a}           \\
-        \verb|{\^e}|     & {\^e}           \\
-        \verb|{\`i}|     & {\`i}           \\
-        \verb|{\.I}|     & {\.I}           \\
-        \verb|{\o}|      & {\o}            \\
-        \verb|{\'u}|     & {\'u}           \\
-        \verb|{\aa}|     & {\aa}           \\\hline
-    \end{tabular}
-    \begin{tabular}{lc}
-        \hline
-        \textbf{Command} & \textbf{Output} \\
-        \hline
-        \verb|{\c c}|    & {\c c}          \\
-        \verb|{\u g}|    & {\u g}          \\
-        \verb|{\l}|      & {\l}            \\
-        \verb|{\~n}|     & {\~n}           \\
-        \verb|{\H o}|    & {\H o}          \\
-        \verb|{\v r}|    & {\v r}          \\
-        \verb|{\ss}|     & {\ss}           \\
-        \hline
-    \end{tabular}
-    \caption{Example commands for accented characters, to be used in, \emph{e.g.}, Bib\TeX{} entries.}
-    \label{tab:accents}
-\end{table}
-
-\subsection{Hyperlinks}
-
-Users of older versions of \LaTeX{} may encounter the following error during compilation:
-\begin{quote}
-    \tt\verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.
-\end{quote}
-This happens when pdf\LaTeX{} is used and a citation splits across a page boundary. The best way to fix this is to upgrade \LaTeX{} to 2018-12-01 or later.
-
-\subsection{Citations}
-
-\begin{table*}
-    \centering
-    \begin{tabular}{lll}
-        \hline
-        \textbf{Output}           & \textbf{natbib command} & \textbf{Old ACL-style command} \\
-        \hline
-        \citep{Gusfield:97}       & \verb|\citep|           & \verb|\cite|                   \\
-        \citealp{Gusfield:97}     & \verb|\citealp|         & no equivalent                  \\
-        \citet{Gusfield:97}       & \verb|\citet|           & \verb|\newcite|                \\
-        \citeyearpar{Gusfield:97} & \verb|\citeyearpar|     & \verb|\shortcite|              \\
-        \hline
-    \end{tabular}
-    \caption{\label{citation-guide}
-        Citation commands supported by the style file.
-        The style is based on the natbib package and supports all natbib citation commands.
-        It also supports commands defined in previous ACL style files for compatibility.
-    }
-\end{table*}
-
-Table~\ref{citation-guide} shows the syntax supported by the style files.
-We encourage you to use the natbib styles.
-You can use the command \verb|\citet| (cite in text) to get ``author (year)'' citations, like this citation to a paper by \citet{Gusfield:97}.
-You can use the command \verb|\citep| (cite in parentheses) to get ``(author, year)'' citations \citep{Gusfield:97}.
-You can use the command \verb|\citealp| (alternative cite without parentheses) to get ``author, year'' citations, which is useful for using citations within parentheses (e.g. \citealp{Gusfield:97}).
-
-\subsection{References}
-
-\nocite{Ando2005,andrew2007scalable,rasooli-tetrault-2015}
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
-If your own bib file is named \texttt{custom.bib}, then placing the following before any appendices in your \LaTeX{} file will generate the references section for you:
-\begin{quote}
-    \begin{verbatim}
-\bibliography{custom}
-\end{verbatim}
-\end{quote}
-
-You can obtain the complete ACL Anthology as a Bib\TeX{} file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
-To include both the Anthology and your own .bib file, use the following instead of the above.
-\begin{quote}
-    \begin{verbatim}
-\bibliography{anthology,custom}
-\end{verbatim}
-\end{quote}
-
-Please see Section~\ref{sec:bibtex} for information on preparing Bib\TeX{} files.
-
-\subsection{Appendices}
-
-Use \verb|\appendix| before any appendix section to switch the section numbering over to letters. See Appendix~\ref{sec:appendix} for an example.
-
-\section{Bib\TeX{} Files}
-\label{sec:bibtex}
-
-Unicode cannot be used in Bib\TeX{} entries, and some ways of typing special characters can disrupt Bib\TeX's alphabetization.
-The recommended way of typing special characters is shown in Table~\ref{tab:accents}.
-
-Please ensure that Bib\TeX{} records contain DOIs or URLs when possible, and for all the ACL materials that you reference.
-Use the \verb|doi| field for DOIs and the \verb|url| field for URLs.
-If a Bib\TeX{} entry has a URL or DOI field, the paper title in the references section will appear as a hyperlink to the paper, using the hyperref \LaTeX{} package.
+\section{Conclusions and further work}
 
@@
 
-This document has been adapted
-by Steven Bethard, Ryan Cotterell and Rui Yan
-from the instructions for earlier ACL and NAACL proceedings, including those for
-ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
-NAACL 2019 by Stephanie Lukin and Alla Roskovskaya,
-ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu,
-NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
-Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
-ACL 2017 by Dan Gildea and Min-Yen Kan,
-NAACL 2017 by Margaret Mitchell,
-ACL 2012 by Maggie Li and Michael White,
-ACL 2010 by Jing-Shin Chang and Philipp Koehn,
-ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui,
-ACL 2005 by Hwee Tou Ng and Kemal Oflazer,
-ACL 2002 by Eugene Charniak and Dekang Lin,
-and earlier ACL and EACL formats written by several people, including
-John Chen, Henry S. Thompson and Donald Walker.
-Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.
-
 % Entries for the entire Anthology, followed by custom entries
-\bibliography{anthology,custom}
+%\bibliography{anthology,custom}
 
@@
 \section{Example Appendix}
-\label{sec:appendix}
-
-This is an appendix.
 

### 2023-06-30 paper/paper.tex
paper

@@
 % This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
-\pdfoutput=1
+%\pdfoutput=1
 % In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.
@@
 \section{Introduction}
-
+TODO:
+\begin{itemize}
+  \item how
+\end{itemize}
 \section{Background and previous work}
@@
 \section{Materials and methods}
+TODO:
+\begin{itemize}
+  \item creation of dataset (CLEVR)
+        \begin{itemize}
+          \item multiple 'real' objects in scene
+          \item 3 attributes (color, size, shape) differentiate objects
+          \item using 'dale' setup to uniquely identify target object
+        \end{itemize}
+  \item building a language game using EGG
+  \item based on feature extractors ResNet/VGG
+  \item
+  \item setup of discriminating game of objects in image
+  \item message encoder/decoder is auto-encoder
+  \item
+\end{itemize}
 
 \section{Results}
+TODO:
+\begin{itemize}
+  \item DaleTwo:
+        \begin{itemize}
+          \item small hidden/embedding dims, small vocab -> high accuracy
+          \item high hidden/embedding dims, small vocav -> low accuracy
+        \end{itemize}
+  \item DaleFive:
+        \begin{itemize}
+          \item small hidden/embedding dims, small vocab -> low accuracy
+          \item small hidden/embedding dims, bigger vocab -> higher accuracy
+        \end{itemize}
+  \item ... test different hidden/embedding dims
+  \item ... test 3/4 objects
+\end{itemize}
 
 \section{Discussion}
+TODO:
+\begin{itemize}
+  \item maybe calculation of loss (multiplicating instead of summing loss per token), unlikely, since sequence length short -> shouldn't result in big differences
+  \item reducing dims of image better the increasing dims of message, increasing dims is not learnable for models
+  \item Vocab:
+        \begin{itemize}
+          \item vocabulary could describe attributes of target image (non-discriminative) or describe only differences (discriminative)
+          \item in second case, two images is a far easier task than five images. Hence, much lower accuracy
+        \end{itemize}
+\end{itemize}
 

### 2023-07-06 paper/paper.tex
paper

@@
 % Remove the "review" option to generate the final version.
-\usepackage[review]{acl}
+\usepackage{acl}
 
@@
 
+\usepackage{graphicx}
+
 \title{Instructions for *ACL Proceedings}
@@
   \texttt{dominik.kuenkele@outlook.com} \\\And
-  Second Author \\
+  Simon Dobnik \\
   Affiliation / Address line 1 \\
@@
   Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
+  \texttt{simon.dobnik@gu.se} \\}
 
@@
 \begin{itemize}
-  \item how
+  \item how can models reidentify attributes (and objects) from learned encodings in an artificial language \textbf{(done)}
 \end{itemize}
+Language games with a sender and receiver offer an opportunity on to look at how neural models condense seen information into a distinct representation of symbols.
+The receiver needs to interpret the meaning of these symbols and combine it with other information.
+This paper studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
+Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
+
 \section{Background and previous work}
@@
 \begin{itemize}
-  \item creation of dataset (CLEVR)
+  \item creation of dataset (CLEVR) \textbf{(done)}
         \begin{itemize}
-          \item multiple 'real' objects in scene
-          \item 3 attributes (color, size, shape) differentiate objects
-          \item using 'dale' setup to uniquely identify target object
+          \item multiple 'real' objects in scene \textbf{(done)}
+          \item 3 attributes (color, size, shape) differentiate objects \textbf{(done)}
+          \item using 'dale' setup to uniquely identify target object \textbf{(done)}
+          \item extracting bounding boxes \textbf{(done)}
         \end{itemize}
-  \item building a language game using EGG
-  \item based on feature extractors ResNet/VGG
-  \item
-  \item setup of discriminating game of objects in image
+  \item building a language game using EGG \textbf{(done)}
+  \item based on feature extractors ResNet \textbf{(done)}
+  \item setup of discriminating game of objects in image \textbf{(done)}
   \item message encoder/decoder is auto-encoder
-  \item
 \end{itemize}
+
+\subsection{Dataset}
+The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{The extended source code can be found on GitHub: \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}} This dataset includes 3D-generated images depicting scenes with different kinds of objects.
+Each of these objects has different combinations of attributes, such as \emph{shape}, \emph{color} and \emph{size}.
+The objects are randomly placed into the scene and assigned with random attributes.
+Figure \ref{fig:clevr-extended_example} shows an example of a generated image of this CLEVR dataset.
+The dataset also contains ground truth information about each image, including attributes and locations for all objects.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}
+  \caption{Example of a generated image with the extended code}
+  \label{fig:clevr-extended_example}
+\end{figure}
+
+The extended version of the dataset divides the objects into \emph{target object} and \emph{distractors}.
+The target object is the main object in the scene that should be identified and communicated by the agents.
+The distractors contain similar objects to the target object.
+The number of shared attributes is defined by rules researched in \citet{Dale1995}.
+The target object is therefore always identifiable using either only the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
+
+For this research, two datasets are created.
+The \emph{DaleTwo} dataset contains one target object and one distractor, while the \emph{DaleFive} dataset contains one target object and four distractors.
+Both datasets contain 10.000 images.
+For each image fixed size bounding boxes around each object are extracted.
+These bounding boxes are the input for the sender and receiver in the language game.
+
+\subsection{Language game}
+The language game was developed and run in the EGG framework \citep{Kharitonov2019}.
+The setup of the game is based on the game described in the paper by \citet{Lazaridou2016}.
+The same multiple images are passed to a sender and a receiver.
+The sender must communicate the target image to the receiver with a message, who needs to identify this target image.
+Instead of distinguishing two images from different concepts, in this experiment the agents need to distinguish two objects with shared attributes.
+Both sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} in \citet{Lazaridou2016}.
+One central difference is the production of the message.
+Their paper focuses on the classification of a concept for the input image and therefore produces only one-symbol messages that should correspond to these concepts.
+This research focuses on the identification of attributes for the objects and their combination, which is why our models produce a sequence of symbols for the message (which may correspond to the attributes).
+This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) (see Figure \ref*{todo}).
+
+The initial state of the encoder LSTM is the image, passed through ResNet101 and a following linear layer that reduces the dimensions.
+The sequence is then created through Gumbel-Softmax relaxation \citep{Jang2016}.
+
+The receiver on the other side takes the sequence as input for its decoder LSTM.
+The hidden state is randomly initialized.
+After each step of the LSTM, the receiver calculates the dot product between the hidden state and its own image encoding (calculated as the sender's image encoding).
+The receiver then 'points' to one of the images, while applying the softmax function over the results.
+The loss is calculated using the NLL-loss.
+Following, the losses for each token in the message sequence are summed up, and all weights of the receiver as well as the sender are updated, based on this summed loss.
+
+There are five variables in the experiments that are adjusted: (1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
+
+The results will be evaluated using the accuracy if the receiver could identify the target object.
+A random guess corresponds to 50\% in the \emph{DaleTwo} dataset and 20\% in the \emph{DaleFive} dataset.
 
@@
 \begin{itemize}
-  \item DaleTwo:
+  \item DaleTwo: \textbf{(done)}
         \begin{itemize}
           \item small hidden/embedding dims, small vocab -> high accuracy
-          \item high hidden/embedding dims, small vocav -> low accuracy
+          \item high hidden/embedding dims, small vocab -> low accuracy
         \end{itemize}
-  \item DaleFive:
+  \item DaleFive: \textbf{(done)}
         \begin{itemize}
@@
 \end{itemize}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|ccccc|c}
+    \hline
+    \textbf{Dataset} & $h_{s}$ & $e_{s}$ & $h_{r}$ & $e_{r}$ & $|V|$ & \textbf{Acc.} \\
+    \hline
+    DaleTwo          & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
+    DaleTwo          & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
+    DaleFive         & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
+    DaleFive         & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
+    DaleFive         & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
+    \hline
+  \end{tabular}
+  \caption{Result of the experiments with different hidden sizes, embedding sizes and vocabulary sizes}
+  \label{tab:results}
+\end{table}
+
+The results of the experiments are summarized in Table \ref{tab:results}.
+For the \emph{DaleTwo} dataset it can be clearly seen that very small embedding and hidden sizes are beneficial for identifying the correct object.
+The receiver identifies indeed almost every sample correctly with all sizes of 10.
+When the hidden and embedding sizes are increased, the guesses by the receiver are random.
+Interestingly, a vocab size of 10 is enough to communicate a meaningful message for the \emph{DaleTwo} dataset.
+
+The result change, when using the \emph{DateFive} dataset with four distractors.
+With low hidden, embedding and vocab size, the agents barely pass the random baseline with 23\%.
+Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\%.
+This accuracy is still far lower than the 95\% with the \emph{DaleTwo} dataset.
 
@@
 % Entries for the entire Anthology, followed by custom entries
-%\bibliography{anthology,custom}
+\bibliography{anthology,custom}
 

### 2023-07-08 paper/paper.tex
Introdcution and comments

@@
 \section{Introduction}
-TODO:
-\begin{itemize}
-  \item how can models reidentify attributes (and objects) from learned encodings in an artificial language \textbf{(done)}
-\end{itemize}
-Language games with a sender and receiver offer an opportunity on to look at how neural models condense seen information into a distinct representation of symbols.
-The receiver needs to interpret the meaning of these symbols and combine it with other information.
-This paper studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
+% TODO:
+% \begin{itemize}
+%   \item how can models reidentify attributes (and objects) from learned encodings in an artificial language \textbf{(done)}
+%   \end{itemize}
+
+Agents interact with the physical world through their action and perception and with other agents through language.
+Their sensors and actuators (if they are artificial agents but the same also holds for natural agents) allow them to sample the world and their their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
+On the other hand language that is used to communicta ewith other agents is based on representations that are composed of a limited set of discrete symbols.
+How can both domains and representations arising from these interaction be combined? How are ranges of measurements exprseed in a continuous domain mapped to discrete linguistic labels?
+How is ambiguity and underpecification resolved?
+Ho can agents achive this through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?
+
+% SD 2023-07-08 14:11:44 +0200: A tip for writing LateX on Git: write each sentence in a separate line, this way any edits and conflicts will be much easier to spot and edit that comparing entire paragraphs.
+
+% \section{Background and previous work}
+
+Language games with a sender and a receiver \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa} offer an opportunity to examine how neural models condense seen information into a distinct representation of symbols through linguistic interaction.
+A sender describes a perceptual situation that can also be seen by a receiver starting with a random label while
+the receiver attempts to interpret the meaning of these symbols and combine it with other information.
+It then sends feedback to the sender about its interpretation.
+Interaction is optimised based on the communicative success of the sender and the receiver.
+Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa,Kelleher:2020aa}.
+Communication is successful and grounding converges because learning is constrained by joint attention and the agents are playing following the rule of the communicative games.
+
+In this this paper we explore how agents based on artificial neural networks learn referential grounding of entitities in images of 3-dimensional scenes where one agent is discribing the entities and the other agent learns to interpret the reference of symbols in the scene either by identifying the object in the scene as a bounding box or location of the object based on object attributes such as shape, colour and size.
+% studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
 Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
+The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that that can be referred to and discriminated based on the attributes such as shape, colour and size whereby the discrimination is based on different overalaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects aither as one of the potential bounding boxes or location of the target.
 
-\section{Background and previous work}
+% SD 2023-07-08 15:05:37 +0200: This hasn't been studied and implemented by lazaridou and others in the EGG framework, right?
 
-\section{Materials and methods}
-TODO:
-\begin{itemize}
-  \item creation of dataset (CLEVR) \textbf{(done)}
-        \begin{itemize}
-          \item multiple 'real' objects in scene \textbf{(done)}
-          \item 3 attributes (color, size, shape) differentiate objects \textbf{(done)}
-          \item using 'dale' setup to uniquely identify target object \textbf{(done)}
-          \item extracting bounding boxes \textbf{(done)}
-        \end{itemize}
-  \item building a language game using EGG \textbf{(done)}
-  \item based on feature extractors ResNet \textbf{(done)}
-  \item setup of discriminating game of objects in image \textbf{(done)}
-  \item message encoder/decoder is auto-encoder
-\end{itemize}
+% SD 2023-07-08 15:15:46 +0200: Just saw that we are only reporting on object from bounding box identification and not location, right?
 
-\subsection{Dataset}
-The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{The extended source code can be found on GitHub: \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}} This dataset includes 3D-generated images depicting scenes with different kinds of objects.
+% \section{Materials and methods}
+% TODO:
+% \begin{itemize}
+%   \item creation of dataset (CLEVR) \textbf{(done)}
+%         \begin{itemize}
+%           \item multiple 'real' objects in scene \textbf{(done)}
+%           \item 3 attributes (color, size, shape) differentiate objects \textbf{(done)}
+%           \item using 'dale' setup to uniquely identify target object \textbf{(done)}
+%           \item extracting bounding boxes \textbf{(done)}
+%         \end{itemize}
+%   \item building a language game using EGG \textbf{(done)}
+%   \item based on feature extractors ResNet \textbf{(done)}
+%   \item setup of discriminating game of objects in image \textbf{(done)}
+%   \item message encoder/decoder is auto-encoder
+% \end{itemize}
+
+\section{Our dataset: CLEVR-DaleTwo and DaleFive}
+
+% SD 2023-07-08 14:58:04 +0200: Rephrase this paragraph to continue from the previous section and emphasise our extension and contribution to the dataset. For our experiments we extend...
+
+The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{%The extended source code can be found on GitHub:
+  \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+This dataset includes 3D-generated images depicting scenes with different kinds of objects.
 Each of these objects has different combinations of attributes, such as \emph{shape}, \emph{color} and \emph{size}.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}
-  \caption{Example of a generated image with the extended code}
+  \frame{\includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}}
+  \caption{Example of a generated image with the extended code} % SD 2023-07-08 15:00:20 +0200: What dataset: Dale-2?
   \label{fig:clevr-extended_example}
@@
 
+% SD 2023-07-08 15:23:38 +0200: Describe a bit more how DaleTwo and DaleFive differ, in particular how many attributes are shared and how between the objects in the DaleFive?
+
 \subsection{Language game}
@@
 This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) (see Figure \ref*{todo}).
+
+% SD 2023-07-08 15:09:29 +0200: We have now already said some of this in the introduction.
+In the introduction we can describe the general setup of language games and here we say explicitly what is our configuration and emphasise in particular in what ways it is different from the previously implemented games.
+Including a list of technical details of the models is great.
+When restructuign sentences, copy those that fit the previous section there (and comment one out if necessary) and keep the rest here.
+It requires a bit of strategic thinking where to mention things in such a short paper so that we only say it once and save space.
 
@@
 
+% SD 2023-07-08 15:17:45 +0200: The hidden size is related to the vocabulary size. Both have to be optimally selected for the models to converge.
+
 The result change, when using the \emph{DateFive} dataset with four distractors.
@@
 This accuracy is still far lower than the 95\% with the \emph{DaleTwo} dataset.
+
+% SD 2023-07-08 15:18:24 +0200: So we are not using just two objects to discriminate with as stated in the introduction and the previous section.
+
+% SD 2023-07-08 15:19:05 +0200: Why is the accuarcy lower?
+One of the reason of course is that there is more uncertanty and the networks have to learn how different combinations of labels that will be up to n!
+relate to individual perceptual features that are encoded in the linguistic categories.
+Since we discrimnate objects based on properties that are also distinguished in human cognition (colour, size, shape) we expect that the vocabulary onto which the agents will converge will reflect these catgories anf therefore be close to human vocabulary.
+We will imnvestiaget the overlap between the vocabularies in our ongoing and future work.
 
@@
 \end{document}
+%%% Local Variables: 
+%%% coding: utf-8
+%%% mode: latex
+%%% mode: flyspell
+%%% ispell-local-dictionary: "british"
+%%% TeX-engine: default
+%%% TeX-master: t
+%%% TeX-PDF-mode: t
+%%% TeX-command-extra-options: "-shell-escape"
+%%% End:

### 2023-07-08 paper/paper.tex
merge

@@
 
-\title{Instructions for *ACL Proceedings}
+\title{Discriminating 3D-objects using }
 
@@
 
+Two interesting conclusions can be drawn.
+First, the hidden as well as the embedding sizes need not be bigger than the vocabulary size.
+Even though that means that the image encodings of 4096 or even higher dimensions need to be compressed to fit the vocabulary size, this will still help to produce and interpret more meaningful messages.
+The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
+from low $h_r$ to high $e_r$).
+
+The second conclusion that can be drawn looks at the difference between the two datasets.
+Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one object.
+Still the difference in the necessary vocabulary size is striking.
+There are 48 possible objects.
+Still a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
+This hints to the fact, that the agents don't describe the complete target object, but only rely on discriminating attributes between the objects.
+This gets more complex when the agents need to discriminate between five objects.
+A bigger vocabulary offers a bigger space
+
 \section{Conclusions and further work}

### 2023-07-08 paper/paper.tex
merges

@@
 Agents interact with the physical world through their action and perception and with other agents through language.
-Their sensors and actuators (if they are artificial agents but the same also holds for natural agents) allow them to sample the world and their their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
-On the other hand language that is used to communicta ewith other agents is based on representations that are composed of a limited set of discrete symbols.
-How can both domains and representations arising from these interaction be combined? How are ranges of measurements exprseed in a continuous domain mapped to discrete linguistic labels?
-How is ambiguity and underpecification resolved?
-Ho can agents achive this through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?
+Their sensors and actuators (if they are artificial agents, but the same also holds for natural agents) allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
+On the other hand language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete symbols.
+How can both domains and representations arising from these interactions be combined? How are ranges of measurements expressed in a continuous domain mapped to discrete linguistic labels?
+How is ambiguity and underspecification resolved?
+How can agents achieve this through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?
 
@@
 
-% \section{Background and previous work}
+\section{Background and previous work}
 
@@
 It then sends feedback to the sender about its interpretation.
-Interaction is optimised based on the communicative success of the sender and the receiver.
-Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa,Kelleher:2020aa}.
+Interaction is optimized based on the communicative success of the sender and the receiver.
+Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa, Kelleher2020}.
 Communication is successful and grounding converges because learning is constrained by joint attention and the agents are playing following the rule of the communicative games.
 
-In this this paper we explore how agents based on artificial neural networks learn referential grounding of entitities in images of 3-dimensional scenes where one agent is discribing the entities and the other agent learns to interpret the reference of symbols in the scene either by identifying the object in the scene as a bounding box or location of the object based on object attributes such as shape, colour and size.
+In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes where one agent is describing the entities and the other agent learns to interpret the reference of symbols in the scene either by identifying the object in the scene as a bounding box or location of the object based on object attributes such as shape, color and size.
 % studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
 Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
-The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that that can be referred to and discriminated based on the attributes such as shape, colour and size whereby the discrimination is based on different overalaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects aither as one of the potential bounding boxes or location of the target.
+The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as shape, color and size whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects either as one of the potential bounding boxes or location of the target.
 
@@
 
-\section{Our dataset: CLEVR-DaleTwo and DaleFive}
+\section{Our dataset: CLEVR-Dale-2 and Dale-5}
 
@@
 
-The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{%The extended source code can be found on GitHub:
-  \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
 This dataset includes 3D-generated images depicting scenes with different kinds of objects.
@@
   \frame{\includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}}
-  \caption{Example of a generated image with the extended code} % SD 2023-07-08 15:00:20 +0200: What dataset: Dale-2?
+  \caption{Example of a generated image in the Dale-5 dataset with the extended code}
   \label{fig:clevr-extended_example}
@@
 For this research, two datasets are created.
-The \emph{DaleTwo} dataset contains one target object and one distractor, while the \emph{DaleFive} dataset contains one target object and four distractors.
+The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
 Both datasets contain 10.000 images.
@@
 
-% SD 2023-07-08 15:23:38 +0200: Describe a bit more how DaleTwo and DaleFive differ, in particular how many attributes are shared and how between the objects in the DaleFive?
+% SD 2023-07-08 15:23:38 +0200: Describe a bit more how Dale-2 and Dale-5 differ, in particular how many attributes are shared and how between the objects in the Dale-5?
 
-\subsection{Language game}
+\section{Language game}
 The language game was developed and run in the EGG framework \citep{Kharitonov2019}.
@@
 
-There are five variables in the experiments that are adjusted: (1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
+There are five variables in the experiments that are adjusted:
+(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
 
 The results will be evaluated using the accuracy if the receiver could identify the target object.
-A random guess corresponds to 50\% in the \emph{DaleTwo} dataset and 20\% in the \emph{DaleFive} dataset.
+A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
 
@@
 \begin{itemize}
-  \item DaleTwo: \textbf{(done)}
+  \item Dale-2: \textbf{(done)}
         \begin{itemize}
@@
         \end{itemize}
-  \item DaleFive: \textbf{(done)}
+  \item Dale-5: \textbf{(done)}
         \begin{itemize}
@@
     \hline
-    DaleTwo          & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
-    DaleTwo          & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
-    DaleFive         & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
-    DaleFive         & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
-    DaleFive         & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
+    Dale-2           & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
+    Dale-2           & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
+    Dale-5           & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
+    Dale-5           & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
+    Dale-5           & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
     \hline
@@
 The results of the experiments are summarized in Table \ref{tab:results}.
-For the \emph{DaleTwo} dataset it can be clearly seen that very small embedding and hidden sizes are beneficial for identifying the correct object.
+For the \emph{Dale-2} dataset it can be clearly seen that very small embedding and hidden sizes are beneficial for identifying the correct object.
 The receiver identifies indeed almost every sample correctly with all sizes of 10.
 When the hidden and embedding sizes are increased, the guesses by the receiver are random.
-Interestingly, a vocab size of 10 is enough to communicate a meaningful message for the \emph{DaleTwo} dataset.
+Interestingly, a vocab size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.
 
@@
 
-The result change, when using the \emph{DateFive} dataset with four distractors.
+The result change, when using the \emph{Date-5} dataset with four distractors.
 With low hidden, embedding and vocab size, the agents barely pass the random baseline with 23\%.
 Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\%.
-This accuracy is still far lower than the 95\% with the \emph{DaleTwo} dataset.
+This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.
 
@@
 Even though that means that the image encodings of 4096 or even higher dimensions need to be compressed to fit the vocabulary size, this will still help to produce and interpret more meaningful messages.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$).
+The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g. from low $h_r$ to high $e_r$).
 

### 2023-07-09 paper/paper.tex
paper

@@
 % Remove the "review" option to generate the final version.
-\usepackage{acl}
+\usepackage[review]{acl}
 
@@
 
-\section{Background and previous work}
+% \section{Background and previous work}
 
@@
 
-In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes where one agent is describing the entities and the other agent learns to interpret the reference of symbols in the scene either by identifying the object in the scene as a bounding box or location of the object based on object attributes such as shape, color and size.
+In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes.
+One agent is describing the entities, more specifically bounding boxes of objects and the other agent learns to interpret the reference of symbols by identifying and referring to one of the bounding boxes based on object attributes such as shape, color and size.
 % studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
 Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
-The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as shape, color and size whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects either as one of the potential bounding boxes or location of the target.
+The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as \emph{shape}, \emph{color} and \emph{size} whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects as one of the potential bounding boxes.
 
 % SD 2023-07-08 15:05:37 +0200: This hasn't been studied and implemented by lazaridou and others in the EGG framework, right?
+% DK 2023-07-09: yes, the main addition in my view is the focus on controlles attributes and relations of objects as opposed to discriminating images from different concepts
 
 % SD 2023-07-08 15:15:46 +0200: Just saw that we are only reporting on object from bounding box identification and not location, right?
+% DK 2023-07-09: exactly. we only looked at the locations without language games, and hopefully also in the thesis
 
@@
 
-\section{Our dataset: CLEVR-Dale-2 and Dale-5}
+\section{Our datasets: CLEVR-Dale-2 and Dale-5}
 
@@
 
-The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-This dataset includes 3D-generated images depicting scenes with different kinds of objects.
-Each of these objects has different combinations of attributes, such as \emph{shape}, \emph{color} and \emph{size}.
-The objects are randomly placed into the scene and assigned with random attributes.
-Figure \ref{fig:clevr-extended_example} shows an example of a generated image of this CLEVR dataset.
-The dataset also contains ground truth information about each image, including attributes and locations for all objects.
+For these experiments, we extend the CLEVR dataset, by dividing the objects into a \emph{target object} and \emph{distractors}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+The target object is the main object in the scene that should be identified and communicated by the agents.
+The distractors contain similar objects to the target object.
+The number of shared attributes between these groups is defined by rules researched in \citet{Dale1995}.
+The target object is always unique and at least one attribute is different from the distractors.
+First, a random target object is created.
+In the next step, the distractors are created at a random location in the scene.
+Each of the distractors can share a maximum of two attributes with the target object.
+There is no restriction on the relation between distractors, namely it is possible to have multiple identical distractors in one image.
+The target object is therefore always identifiable using either only the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
+
+For this research, two datasets are created.
+The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
+The way, how the distractors are created is the same.
+Both datasets contain 10.000 images.
+For each image fixed size bounding boxes around each object are extracted.
+These bounding boxes are the input for the sender and receiver in the language game.
+
+Figure \ref{fig:clevr-extended_example} shows an example of a generated image of the Dale-5 dataset.
+Both datasets also contain ground truth information about each image, including attributes and locations for all objects.
 
@@
   \centering
-  \frame{\includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}}
+  \frame{\includegraphics[width=.6\linewidth]{figures/CLEVR_extended_example.png}}
   \caption{Example of a generated image in the Dale-5 dataset with the extended code}
@@
 
-The extended version of the dataset divides the objects into \emph{target object} and \emph{distractors}.
-The target object is the main object in the scene that should be identified and communicated by the agents.
-The distractors contain similar objects to the target object.
-The number of shared attributes is defined by rules researched in \citet{Dale1995}.
-The target object is therefore always identifiable using either only the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
-
-For this research, two datasets are created.
-The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
-Both datasets contain 10.000 images.
-For each image fixed size bounding boxes around each object are extracted.
-These bounding boxes are the input for the sender and receiver in the language game.
-
 % SD 2023-07-08 15:23:38 +0200: Describe a bit more how Dale-2 and Dale-5 differ, in particular how many attributes are shared and how between the objects in the Dale-5?
+% DK 2023-07-09: Dale-2 and Dale-5 only differ in the number of distractors. The rest is the same. I described in the previous paragrapgh a bit more detailed, how the distractors are created.
 
@@
 The language game was developed and run in the EGG framework \citep{Kharitonov2019}.
-The setup of the game is based on the game described in the paper by \citet{Lazaridou2016}.
-The same multiple images are passed to a sender and a receiver.
-The sender must communicate the target image to the receiver with a message, who needs to identify this target image.
-Instead of distinguishing two images from different concepts, in this experiment the agents need to distinguish two objects with shared attributes.
+\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
 Both sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} in \citet{Lazaridou2016}.
 One central difference is the production of the message.
-Their paper focuses on the classification of a concept for the input image and therefore produces only one-symbol messages that should correspond to these concepts.
-This research focuses on the identification of attributes for the objects and their combination, which is why our models produce a sequence of symbols for the message (which may correspond to the attributes).
-This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) (see Figure \ref*{todo}).
+This research focuses on the identification of attributes for the objects and their combination, which is why our models produce a sequence of symbols for the message (which may correspond to the attributes) instead of a single symbol (which corresponded to different concepts of the images).
+This is done using an encoder LSTM (sender) and a decoder LSTM (receiver).
 
@@
 
-The initial state of the encoder LSTM is the image, passed through ResNet101 and a following linear layer that reduces the dimensions.
+The target object is always the first image that is passed to the sender.
+The order of the objects for the receiver is random.
+Figure \ref*{fig:architectures} shows the architectures of the sender and the receiver.
+For the sender, the images are passed through \emph{ResNet101} \citep{He2016} and a following linear layer that reduces the dimensions to an embedding size $e_s$.
+All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
+This is then used as the initial state of the encoder LSTM.
 The sequence is then created through Gumbel-Softmax relaxation \citep{Jang2016}.
 
-The receiver on the other side takes the sequence as input for its decoder LSTM.
-The hidden state is randomly initialized.
-After each step of the LSTM, the receiver calculates the dot product between the hidden state and its own image encoding (calculated as the sender's image encoding).
-The receiver then 'points' to one of the images, while applying the softmax function over the results.
+The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
+The sequence received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
+After each step of the LSTM, the receiver calculates the dot product between the hidden state and all of its image encodings.
+The receiver then 'points' to one of the images, by applying the softmax function over the results of the dot products.
 The loss is calculated using the NLL-loss.
-Following, the losses for each token in the message sequence are summed up, and all weights of the receiver as well as the sender are updated, based on this summed loss.
+Following, the losses for all steps are summed up, and all weights of the receiver as well as the sender are updated, based on this summed loss.
+
+\begin{figure}[h]
+  \centering
+  \frame{\includegraphics[width=\linewidth]{figures/architecture.png}}
+  \caption{Architectures of the sender and receiver}
+  \label{fig:architectures}
+\end{figure}
 
@@
 \section{Results}
-TODO:
-\begin{itemize}
-  \item Dale-2: \textbf{(done)}
-        \begin{itemize}
-          \item small hidden/embedding dims, small vocab -> high accuracy
-          \item high hidden/embedding dims, small vocab -> low accuracy
-        \end{itemize}
-  \item Dale-5: \textbf{(done)}
-        \begin{itemize}
-          \item small hidden/embedding dims, small vocab -> low accuracy
-          \item small hidden/embedding dims, bigger vocab -> higher accuracy
-        \end{itemize}
-  \item ... test different hidden/embedding dims
-  \item ... test 3/4 objects
-\end{itemize}
+% TODO:
+% \begin{itemize}
+%   \item Dale-2: \textbf{(done)}
+%         \begin{itemize}
+%           \item small hidden/embedding dims, small vocab -> high accuracy
+%           \item high hidden/embedding dims, small vocab -> low accuracy
+%         \end{itemize}
+%   \item Dale-5: \textbf{(done)}
+%         \begin{itemize}
+%           \item small hidden/embedding dims, small vocab -> low accuracy
+%           \item small hidden/embedding dims, bigger vocab -> higher accuracy
+%         \end{itemize}
+%   \item ... test different hidden/embedding dims
+%   \item ... test 3/4 objects
+% \end{itemize}
 
@@
 The results of the experiments are summarized in Table \ref{tab:results}.
-For the \emph{Dale-2} dataset it can be clearly seen that very small embedding and hidden sizes are beneficial for identifying the correct object.
+For the \emph{Dale-2} dataset it can be clearly seen that an embedding and hidden size that are as high as the vocabulary size are beneficial for identifying the correct object.
 The receiver identifies indeed almost every sample correctly with all sizes of 10.
 When the hidden and embedding sizes are increased, the guesses by the receiver are random.
-Interestingly, a vocab size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.
+Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.
 
@@
 
-The result change, when using the \emph{Date-5} dataset with four distractors.
+The results change, when using the \emph{Date-5} dataset with four distractors.
 With low hidden, embedding and vocab size, the agents barely pass the random baseline with 23\%.
@@
 % SD 2023-07-08 15:18:24 +0200: So we are not using just two objects to discriminate with as stated in the introduction and the previous section.
+% DK 2023-07-09: only in the Dale-2 dataset. The Dale-5 dataset discriminates 5 objects.
 
@@
 We will imnvestiaget the overlap between the vocabularies in our ongoing and future work.
+% DK 2023-07-09: I added it to the Discussion section
 
 \section{Discussion}
-TODO:
-\begin{itemize}
-  \item maybe calculation of loss (multiplicating instead of summing loss per token), unlikely, since sequence length short -> shouldn't result in big differences
-  \item reducing dims of image better the increasing dims of message, increasing dims is not learnable for models
-  \item Vocab:
-        \begin{itemize}
-          \item vocabulary could describe attributes of target image (non-discriminative) or describe only differences (discriminative)
-          \item in second case, two images is a far easier task than five images. Hence, much lower accuracy
-        \end{itemize}
-\end{itemize}
+% TODO:
+% \begin{itemize}
+%   \item maybe calculation of loss (multiplicating instead of summing loss per token), unlikely, since sequence length short -> shouldn't result in big differences
+%   \item reducing dims of image better the increasing dims of message, increasing dims is not learnable for models
+%   \item Vocab:
+%         \begin{itemize}
+%           \item vocabulary could describe attributes of target image (non-discriminative) or describe only differences (discriminative)
+%           \item in second case, two images is a far easier task than five images. Hence, much lower accuracy
+%         \end{itemize}
+% \end{itemize}
 
+% DK 2023-07-09: is the follwoing paragraph necessary?
+The findings are quite obviuous in the first look, but would need more detailed investigation, why this is happening.
+This is not the focus of this paper and maybe we can use the space more for other findings.
 Two interesting conclusions can be drawn.
-First, the hidden as well as the embedding sizes need not be bigger than the vocabulary size.
-Even though that means that the image encodings of 4096 or even higher dimensions need to be compressed to fit the vocabulary size, this will still help to produce and interpret more meaningful messages.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g. from low $h_r$ to high $e_r$).
+First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
+This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
+The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
+from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector.
 
-The second conclusion that can be drawn looks at the difference between the two datasets.
+The second conclusion that can be drawn looks at the differences between the two datasets.
 Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one object.
-Still the difference in the necessary vocabulary size is striking.
-There are 48 possible objects.
-Still a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact, that the agents don't describe the complete target object, but only rely on discriminating attributes between the objects.
-This gets more complex when the agents need to discriminate between five objects.
-A bigger vocabulary offers a bigger space
+Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape) we expect that the vocabulary onto which the agents will converge will reflect these categories and therefore be close to human vocabulary.
+There are 48 possible combinations of attributes.
+Still for Dale-2 a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
+This hints to the fact, that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
+The need for a more detailed description of discriminative attributes is higher, when more distractors are involved.
+Therefore, the models need to learn more combinations of symbols to attest to this higher level of detail and especially how to relate them to features in the images.
 
 \section{Conclusions and further work}
+In this research, we introduced an extended version of the CLEVR dataset that offers a high control over the relations between generated objects.
+This should help to study, how models can learn these relations and refer to them as well as the objects.
+In our research we used this dataset to study, how agents can learn to communicate and refer to these objects in a language game setup.
+We found that a bigger number of distractors requires a higher vocabulary size.
+This hints to the fact that the agents learned to communicate discriminative attributes between the images.
+
+In future research, we want to investigate deeper the emerged language and the new vocabulary.
+Especially, we want to test if our hypothesis that this emerged language uses similar categories as human language holds true and how these words are combined to form complete messages.
 

### 2023-07-10 paper/paper.tex
appendix

@@
 \title{Discriminating 3D-objects using }
-
 % Author information can be set in various styles:
@@
 
-Figure \ref{fig:clevr-extended_example} shows an example of a generated image of the Dale-5 dataset.
+Appendix \ref{sec:extended-datasets} shows examples of generated images with our extended code.
 Both datasets also contain ground truth information about each image, including attributes and locations for all objects.
-
-\begin{figure}[h]
-  \centering
-  \frame{\includegraphics[width=.6\linewidth]{figures/CLEVR_extended_example.png}}
-  \caption{Example of a generated image in the Dale-5 dataset with the extended code}
-  \label{fig:clevr-extended_example}
-\end{figure}
 
@@
 The order of the objects for the receiver is random.
-Figure \ref*{fig:architectures} shows the architectures of the sender and the receiver.
+The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
 For the sender, the images are passed through \emph{ResNet101} \citep{He2016} and a following linear layer that reduces the dimensions to an embedding size $e_s$.
@@
 Following, the losses for all steps are summed up, and all weights of the receiver as well as the sender are updated, based on this summed loss.
-
-\begin{figure}[h]
-  \centering
-  \frame{\includegraphics[width=\linewidth]{figures/architecture.png}}
-  \caption{Architectures of the sender and receiver}
-  \label{fig:architectures}
-\end{figure}
 
@@
 
-\begin{table}
+\begin{table}[b]
   \centering
@@
 This is not the focus of this paper and maybe we can use the space more for other findings.
-Two interesting conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
+
+% Two interesting conclusions can be drawn.
+% First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
+% This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
+% The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
 from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector.
 
-The second conclusion that can be drawn looks at the differences between the two datasets.
+% The second conclusion that can be drawn looks at the differences between the two datasets.
 Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one object.
@@
 
-\section*{Acknowledgements}
-
 % Entries for the entire Anthology, followed by custom entries
@@
 
-\section{Example Appendix}
+\section{Extended CLEVR datasets}
+\label{sec:extended-datasets}
+\begin{figure}[h]
+  \centering
+  \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-FIVE_example.png}}
+  \caption{Example of a generated image in the Dale-5 dataset with the extended code}
+  \label{fig:clevr-dale-5}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-TWO_example.png}}
+  \caption{Example of a generated image in the Dale-2 dataset with the extended code}
+  \label{fig:clevr-dale-2}
+\end{figure}
+
+\section{Setup of the language game}
+\label{sec:setup-language-game}
+
+\begin{figure}[h!]
+  \centering
+  \frame{\includegraphics[width=\linewidth]{figures/architecture.png}}
+  \caption{Architectures of the sender and receiver}
+  \label{fig:architectures}
+\end{figure}
 

### 2023-07-10 paper/paper.tex
appendix

@@
   \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-FIVE_example.png}}
-  \caption{Example of a generated image in the Dale-5 dataset with the extended code}
+  \caption{Example of a generated image in the Dale-5 dataset}
   \label{fig:clevr-dale-5}
 \end{figure}
+The target object in Figure \ref{fig:clevr-dale-5} is the purple cylinder.
+It shares the same color and size with the purple sphere, the same size with the two cubes and no attribute with the turquoise sphere.
+It can be uniquely identified as the \emph{cylinder}.
 
@@
   \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-TWO_example.png}}
-  \caption{Example of a generated image in the Dale-2 dataset with the extended code}
+  \caption{Example of a generated image in the Dale-2 dataset}
   \label{fig:clevr-dale-2}
 \end{figure}
+
+In Figure \ref*{fig:clevr-dale-2}, the small red cube is the target object.
+Since all attributes except for the size are shared with the distractor, all three attributes are necessary, to identify it following \citet{Dale1995}'s rules, namely the \emph{small red cube}.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}
   \centering

### 2023-07-10 paper/paper.tex
Sections 1-3

@@
 % Remove the "review" option to generate the final version.
-\usepackage[review]{acl}
+% \usepackage[review]{acl}
+\usepackage{acl}
 
@@
 
-\title{Discriminating 3D-objects using }
+\title{Referring as a collaborative process: learning to ground \\ language through language games} % Discriminating 3D-objects using
 % Author information can be set in various styles:
@@
 
-\author{Dominik Künkele \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{dominik.kuenkele@outlook.com} \\\And
-  Simon Dobnik \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{simon.dobnik@gu.se} \\}
+\author{Dominik Künkele$^{1}$ \and Simon Dobnik$^{1,2}$ \\
+        Department of Philosophy, Linguistics and Theory of Science$^{1}$ \\ 
+        Centre for Linguistic Theory and Studies in Probability (CLASP)$^{2}$ \\
+        University of Gothenburg, Sweden \\
+        \texttt{dominik.kuenkele@outlook.com} \and \texttt{simon.dobnik@gu.se}}
+
+% \author{Dominik Künkele \\
+%   Affiliation / Address line 1 \\
+%   Affiliation / Address line 2 \\
+%   Affiliation / Address line 3 \\
+%   \texttt{dominik.kuenkele@outlook.com} \\\And
+%   Simon Dobnik \\
+%   Affiliation / Address line 1 \\
+%   Affiliation / Address line 2 \\
+%   Affiliation / Address line 3 \\
+%   \texttt{simon.dobnik@gu.se} \\}
 
@@
 \begin{abstract}
-
+  How do artificial agents based on neural networks coordinate on a new language through referential games over 3-dimensional scenes?
+  We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referrential grounding learned by the agents.
 \end{abstract}
@@
 Agents interact with the physical world through their action and perception and with other agents through language.
-Their sensors and actuators (if they are artificial agents, but the same also holds for natural agents) allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
-On the other hand language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete symbols.
+Their sensors and actuators % (if they are artificial agents, but the same also holds for natural agents)
+allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
+On the other hand language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete and arbitrarily chosen symbols.
 How can both domains and representations arising from these interactions be combined? How are ranges of measurements expressed in a continuous domain mapped to discrete linguistic labels?
@@
 
-Language games with a sender and a receiver \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa} offer an opportunity to examine how neural models condense seen information into a distinct representation of symbols through linguistic interaction.
-A sender describes a perceptual situation that can also be seen by a receiver starting with a random label while
-the receiver attempts to interpret the meaning of these symbols and combine it with other information.
-It then sends feedback to the sender about its interpretation.
-Interaction is optimized based on the communicative success of the sender and the receiver.
-Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa, Kelleher2020}.
-Communication is successful and grounding converges because learning is constrained by joint attention and the agents are playing following the rule of the communicative games.
+% Language games with a sender and a receiver \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa} offer an opportunity to examine how neural models condense seen information into a distinct representation of symbols through linguistic interaction.
+% A sender describes a perceptual situation that can also be seen by a receiver starting with a random label while
+% the receiver attempts to interpret the meaning of these symbols and combine it with other information.
+% It then sends feedback to the sender about its interpretation.
+% Interaction is optimized based on the communicative success of the sender and the receiver.
+% Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa, Kelleher2020}.
+% Communication is successful and grounding converges because learning is constrained by joint attention and the agents are playing following the rule of the communicative games.
 
-In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes.
-One agent is describing the entities, more specifically bounding boxes of objects and the other agent learns to interpret the reference of symbols by identifying and referring to one of the bounding boxes based on object attributes such as shape, color and size.
+In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes through language games \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa}.
+One agent is describing the entities represented as features within bounding boxes of objects, inventing new vocabulary as neccessary, and the other agent learns to interpret the reference of symbols by identifying % and referring to
+one ofthe bounding boxes based on object attributes such as shape, color and size.
 % studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
 Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
-The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as \emph{shape}, \emph{color} and \emph{size} whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects as one of the potential bounding boxes.
+Both agents learn through the success of interaction.
+The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016} is that we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as \emph{shape}, \emph{color} and \emph{size} whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects.
+% and (ii) we implement a focused referring to objects as one of the potential bounding boxes.
 
@@
 
-\section{Our datasets: CLEVR-Dale-2 and Dale-5}
+\section{CLEVR-Dale-2 and Dale-5}
 
@@
 
-For these experiments, we extend the CLEVR dataset, by dividing the objects into a \emph{target object} and \emph{distractors}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-The target object is the main object in the scene that should be identified and communicated by the agents.
-The distractors contain similar objects to the target object.
-The number of shared attributes between these groups is defined by rules researched in \citet{Dale1995}.
+% For these experiments,
+We extend the CLEVR dataset \citep{Johnson2016}, by dividing the objects into a \emph{target object} and \emph{distractors}
+% The target object is the main object in the scene that should be identified and communicated by the agents.
+% The distractors contain similar objects to the target object.
+and control for the number of shared attributes between these groups as % is defined by rules researched in
+in the GRE algorithim in \citep{Dale1995}.
 The target object is always unique and at least one attribute is different from the distractors.
-First, a random target object is created.
-In the next step, the distractors are created at a random location in the scene.
+% First, a random target object is created.
+% In the next step, the distractors are created at a random location in the scene.
 Each of the distractors can share a maximum of two attributes with the target object.
-There is no restriction on the relation between distractors, namely it is possible to have multiple identical distractors in one image.
-The target object is therefore always identifiable using either only the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
-
-For this research, two datasets are created.
+There is no restriction on the relation between distractors, hence it is possible to have multiple identical distractors in one image.
+Given the ranking of features in the original GRE algorithm the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
+For each image fixed size bounding boxes around the centre-point of each object are extracted.
+% These bounding boxes are the input for the sender and receiver in the language game.
+%
+% For this research, two datasets are created.
 The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
-The way, how the distractors are created is the same.
+% The way, how the distractors are created is the same.
 Both datasets contain 10.000 images.
-For each image fixed size bounding boxes around each object are extracted.
-These bounding boxes are the input for the sender and receiver in the language game.
-
-Appendix \ref{sec:extended-datasets} shows examples of generated images with our extended code.
-Both datasets also contain ground truth information about each image, including attributes and locations for all objects.
+Examples are shown in 
+Appendix \ref{sec:extended-datasets}. % shows examples of generated images with our extended code.
+% Both datasets also contain ground truth information about each image, including attributes and locations for all objects.
 
@@
 
-\section{Language game}
-The language game was developed and run in the EGG framework \citep{Kharitonov2019}.
-\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-Both sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} in \citet{Lazaridou2016}.
+\section{Language games}
+The language games were developed and run in the EGG framework \citep{Kharitonov2019}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
+Both our sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} of \citep{Lazaridou2016} as shown in Appendix \ref*{sec:setup-language-game}.
 One central difference is the production of the message.
-This research focuses on the identification of attributes for the objects and their combination, which is why our models produce a sequence of symbols for the message (which may correspond to the attributes) instead of a single symbol (which corresponded to different concepts of the images).
-This is done using an encoder LSTM (sender) and a decoder LSTM (receiver).
-
+% This research focuses on the identification of attributes for the objects and their combination, which is why
+As we focus on sequences of referring expressions made-up of different attributes, our models produce sequences of symbols for the message % (which may correspond to the attributes)
+instead of a single symbol referring to an image. % (which corresponded to different concepts of the images).
+This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) to encode language descriptions. 
+%
 % SD 2023-07-08 15:09:29 +0200: We have now already said some of this in the introduction.
@@
 It requires a bit of strategic thinking where to mention things in such a short paper so that we only say it once and save space.
-
-The target object is always the first image that is passed to the sender.
-The order of the objects for the receiver is random.
-The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
+%
+%
+% The target object is always the first image that is passed to the sender
+% The order of the objects for the receiver is random.
+Another difference is that both sender and receiver receive visual input as segmented objects rather than two images.
+The order of objects is random, except that the first order for the sender is always the target object to be referred to.
+% The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
 For the sender, the images are passed through \emph{ResNet101} \citep{He2016} and a following linear layer that reduces the dimensions to an embedding size $e_s$.
@@
 The sequence is then created through Gumbel-Softmax relaxation \citep{Jang2016}.
-
+%
 The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
@@
 
+\section{Experiments and results}
+
 There are five variables in the experiments that are adjusted:
@@
 
-\section{Results}
 % TODO:

### 2023-07-10 paper/paper.tex
typos, corrections

@@
   How do artificial agents based on neural networks coordinate on a new language through referential games over 3-dimensional scenes?
-  We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referrential grounding learned by the agents.
+  We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
 \end{abstract}
@@
 In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes through language games \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa}.
-One agent is describing the entities represented as features within bounding boxes of objects, inventing new vocabulary as neccessary, and the other agent learns to interpret the reference of symbols by identifying % and referring to
-one ofthe bounding boxes based on object attributes such as shape, color and size.
+One agent is describing the entities represented as features within bounding boxes of objects, inventing new vocabulary as necessary, and the other agent learns to interpret the reference of symbols by identifying % and referring to
+one of the bounding boxes based on object attributes such as shape, color and size.
 % studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
@@
 and control for the number of shared attributes between these groups as % is defined by rules researched in
-in the GRE algorithim in \citep{Dale1995}.
+in the GRE algorithm in \citep{Dale1995}.
 The target object is always unique and at least one attribute is different from the distractors.
@@
 Given the ranking of features in the original GRE algorithm the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
-For each image fixed size bounding boxes around the centre-point of each object are extracted.
+For each image fixed size bounding boxes around the center point of each object are extracted.
 % These bounding boxes are the input for the sender and receiver in the language game.
@@
 Another difference is that both sender and receiver receive visual input as segmented objects rather than two images.
-The order of objects is random, except that the first order for the sender is always the target object to be referred to.
+The order of objects is random, except that the first object for the sender is always the target object to be referred to.
 % The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
@@
 
-The results will be evaluated using the accuracy if the receiver could identify the target object.
+The results are evaluated using the accuracy if the receiver could identify the target object.
 A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.

### 2023-07-10 paper/paper.tex
The remaining sections

@@
 In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes through language games \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa}.
+% Clark:1986aa
 One agent is describing the entities represented as features within bounding boxes of objects, inventing new vocabulary as necessary, and the other agent learns to interpret the reference of symbols by identifying % and referring to
-one of the bounding boxes based on object attributes such as shape, color and size.
+one of the bounding boxes based on object attributes such as shape, colour and size.
 % studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
@@
 Both agents learn through the success of interaction.
-The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016} is that we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as \emph{shape}, \emph{color} and \emph{size} whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects.
+The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016} is that we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as \emph{shape}, \emph{colour} and \emph{size} whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects.
 % and (ii) we implement a focused referring to objects as one of the potential bounding boxes.
@@
 There is no restriction on the relation between distractors, hence it is possible to have multiple identical distractors in one image.
-Given the ranking of features in the original GRE algorithm the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{color} (2) or the \emph{shape}, \emph{color} and \emph{size} (3).
-For each image fixed size bounding boxes around the center point of each object are extracted.
+Given the ranking of features in the original GRE algorithm the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{colour} (2) or the \emph{shape}, \emph{colour} and \emph{size} (3).
+For each image fixed size bounding boxes around the centre-point of each object are extracted.
 % These bounding boxes are the input for the sender and receiver in the language game.
@@
 Another difference is that both sender and receiver receive visual input as segmented objects rather than two images.
-The order of objects is random, except that the first object for the sender is always the target object to be referred to.
+The order of objects is random, except that the first order for the sender is always the target object to be referred to.
 % The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
@@
 The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
-The sequence received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
+The sequence received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialised.
 After each step of the LSTM, the receiver calculates the dot product between the hidden state and all of its image encodings.
-The receiver then 'points' to one of the images, by applying the softmax function over the results of the dot products.
+The receiver then `points' to one of the images, by applying the softmax function over the results of the dot products.
 The loss is calculated using the NLL-loss.
@@
 (1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
-
-The results are evaluated using the accuracy if the receiver could identify the target object.
+%
+% The results are evaluated using accuracy if the receiver can identify the target object.
+Table \ref{tab:results} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object. 
 A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
@@
   \end{tabular}
-  \caption{Result of the experiments with different hidden sizes, embedding sizes and vocabulary sizes}
+  \caption{Results: $h$ are different hidden sizes, $e$ embedding sizes and $|V|$ vocabulary sizes.}
   \label{tab:results}
@@
 
-The results of the experiments are summarized in Table \ref{tab:results}.
+% The results of the experiments are summarized in Table \ref{tab:results}.
 For the \emph{Dale-2} dataset it can be clearly seen that an embedding and hidden size that are as high as the vocabulary size are beneficial for identifying the correct object.
-The receiver identifies indeed almost every sample correctly with all sizes of 10.
+The receiver identifies almost every sample correctly with all sizes of 10.
 When the hidden and embedding sizes are increased, the guesses by the receiver are random.
 Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.
-
+%
 % SD 2023-07-08 15:17:45 +0200: The hidden size is related to the vocabulary size. Both have to be optimally selected for the models to converge.
-
-The results change, when using the \emph{Date-5} dataset with four distractors.
-With low hidden, embedding and vocab size, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\%.
-This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.
+%
+% The results change, when using the \emph{Date-5} dataset with four distractors.
+\emph{Date-5} with four distractor objects and with low hidden, embedding and vocabulary size, the agents barely pass the random baseline with 23\%.
+Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\% which is still considerably lower than 95\% of the \emph{Dale-2} dataset.
+% This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.
 
@@
 
-\section{Discussion}
+\section{Discussion and future work}
 % TODO:
@@
 Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one object.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape) we expect that the vocabulary onto which the agents will converge will reflect these categories and therefore be close to human vocabulary.
+Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape) we expect that the vocabulary onto which the agents converge reflects these categories and therefore be close to human vocabulary.
 There are 48 possible combinations of attributes.
@@
 Therefore, the models need to learn more combinations of symbols to attest to this higher level of detail and especially how to relate them to features in the images.
+% SD 2023-07-10 18:04:24 +0200: It would be interesting to compare then the length of messages between Dale-2 and Dale-5 that becomes used.
 
-\section{Conclusions and further work}
-In this research, we introduced an extended version of the CLEVR dataset that offers a high control over the relations between generated objects.
-This should help to study, how models can learn these relations and refer to them as well as the objects.
-In our research we used this dataset to study, how agents can learn to communicate and refer to these objects in a language game setup.
-We found that a bigger number of distractors requires a higher vocabulary size.
-This hints to the fact that the agents learned to communicate discriminative attributes between the images.
+% \section{Conclusions and further work}
+% In this research, we introduced an extended version of the CLEVR dataset that offers a high control over the relations between generated objects.
+% This should help to study, how models can learn these relations and refer to them as well as the objects.
+% In our research we used this dataset to study, how agents can learn to communicate and refer to these objects in a language game setup.
+% We found that a bigger number of distractors requires a higher vocabulary size.
+% This hints to the fact that the agents learned to communicate discriminative attributes between the images.
 
-In future research, we want to investigate deeper the emerged language and the new vocabulary.
-Especially, we want to test if our hypothesis that this emerged language uses similar categories as human language holds true and how these words are combined to form complete messages.
-
+In our ongoing work we are investigating deeper the emerged language and the new vocabulary,
+in particular whether
+this uses similar categories as human language and how its words are combined to form complete messages.
+In our future work we will also extend learning to relations between entities and the features required to capture them.
 % Entries for the entire Anthology, followed by custom entries
@@
 The target object in Figure \ref{fig:clevr-dale-5} is the purple cylinder.
-It shares the same color and size with the purple sphere, the same size with the two cubes and no attribute with the turquoise sphere.
+It shares the same colour and size with the purple sphere, the same size with the two cubes and no attribute with the turquoise sphere.
 It can be uniquely identified as the \emph{cylinder}.

### 2023-07-10 paper/paper.tex

@@
   \centering
+  \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-TWO_example.png}}
+  \caption{An example from the Dale-2 dataset}
+  \label{fig:clevr-dale-2}
+\end{figure}
+
+In Figure \ref*{fig:clevr-dale-2}, the small red cube is the target object.
+Since all attributes except for the size are shared with the distractor, all three attributes are necessary, to identify it following \citet{Dale1995}'s rules, namely the \emph{small red cube}.
+
+\begin{figure}[h]
+  \centering
   \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-FIVE_example.png}}
-  \caption{Example of a generated image in the Dale-5 dataset}
+  \caption{An example from the Dale-5 dataset}
   \label{fig:clevr-dale-5}
@@
 
-\begin{figure}[h]
-  \centering
-  \frame{\includegraphics[width=\linewidth]{figures/CLEVR_DALE-TWO_example.png}}
-  \caption{Example of a generated image in the Dale-2 dataset}
-  \label{fig:clevr-dale-2}
-\end{figure}
-
-In Figure \ref*{fig:clevr-dale-2}, the small red cube is the target object.
-Since all attributes except for the size are shared with the distractor, all three attributes are necessary, to identify it following \citet{Dale1995}'s rules, namely the \emph{small red cube}.
-
 \section{Setup of the language game}
 \label{sec:setup-language-game}
-
-\begin{figure}
+%\begin{figure}
   \centering
   \frame{\includegraphics[width=\linewidth]{figures/architecture.png}}
-  \caption{Architectures of the sender and receiver}
-  \label{fig:architectures}
-\end{figure}
+  % \caption{Architectures of the sender and receiver}
+%  \label{fig:architectures}
+% \end{figure}
 

### 2023-07-10 paper/paper.tex
corrections

@@
   We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
+  % complicated
 \end{abstract}
@@
 
-Agents interact with the physical world through their action and perception and with other agents through language.
+Agents interact with the physical world through their action and perception, and with other agents through language.
 Their sensors and actuators % (if they are artificial agents, but the same also holds for natural agents)
-allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
-On the other hand language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete and arbitrarily chosen symbols.
-How can both domains and representations arising from these interactions be combined? How are ranges of measurements expressed in a continuous domain mapped to discrete linguistic labels?
+allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and other measures which can be measured with a high degree of accuracy.
+On the other hand, the language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete and arbitrarily chosen symbols.
+How can both domains and representations arising from these interactions be combined? How are the ranges of measurements expressed in a continuous domain mapped to discrete linguistic labels?
 How is ambiguity and underspecification resolved?
@@
 % Clark:1986aa
-One agent is describing the entities represented as features within bounding boxes of objects, inventing new vocabulary as necessary, and the other agent learns to interpret the reference of symbols by identifying % and referring to
+One agent is describing the entities represented as features within bounding boxes of objects, inventing new vocabulary as necessary.
+The other agent learns to interpret the reference of symbols by identifying % and referring to
 one of the bounding boxes based on object attributes such as shape, colour and size.
 % studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object.
-Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
+Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, colour, and size.
 Both agents learn through the success of interaction.
-The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016} is that we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as \emph{shape}, \emph{colour} and \emph{size} whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects.
+The novelty of our work, compared with the previous work with this setup \citep{Kharitonov2019,Lazaridou2016}, consists the extension of the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects.
+These can be referred to based on attributes such as \emph{shape}, \emph{colour} and \emph{size} and discriminated based on different overlaps of these attributes between the target and the distractor objects.
 % and (ii) we implement a focused referring to objects as one of the potential bounding boxes.
@@
 %           \item multiple 'real' objects in scene \textbf{(done)}
-%           \item 3 attributes (color, size, shape) differentiate objects \textbf{(done)}
+%           \item 3 attributes (colour, size, shape) differentiate objects \textbf{(done)}
 %           \item using 'dale' setup to uniquely identify target object \textbf{(done)}
@@
 % For these experiments,
-We extend the CLEVR dataset \citep{Johnson2016}, by dividing the objects into a \emph{target object} and \emph{distractors}
+We extend the CLEVR dataset \citep{Johnson2016}, by dividing the objects into one \emph{target object} and \emph{distractors}
 % The target object is the main object in the scene that should be identified and communicated by the agents.
 % The distractors contain similar objects to the target object.
-and control for the number of shared attributes between these groups as % is defined by rules researched in
+and by controlling for the number of shared attributes between these groups as % is defined by rules researched in
 in the GRE algorithm in \citep{Dale1995}.
-The target object is always unique and at least one attribute is different from the distractors.
+The target object is always unique, because at least one attribute is different from the distractors.
 % First, a random target object is created.
 % In the next step, the distractors are created at a random location in the scene.
-Each of the distractors can share a maximum of two attributes with the target object.
+Each distractor can share a maximum of two attributes with the target object.
 There is no restriction on the relation between distractors, hence it is possible to have multiple identical distractors in one image.
-Given the ranking of features in the original GRE algorithm the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{colour} (2) or the \emph{shape}, \emph{colour} and \emph{size} (3).
-For each image fixed size bounding boxes around the centre-point of each object are extracted.
+Given the ranking of features in the original GRE algorithm, the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{colour} (2) or the \emph{shape}, \emph{colour} and \emph{size} (3).
+For each image, fixed-size bounding boxes are extracted around the centre-point of each object.
 % These bounding boxes are the input for the sender and receiver in the language game.
@@
 Another difference is that both sender and receiver receive visual input as segmented objects rather than two images.
-The order of objects is random, except that the first object for the sender is always the target object to be referred to.
+The order of the objects is random, except that the first object for the sender is always the target object to be referred to.
 % The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
@@
 The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
-The sequence received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialised.
+The sequence received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
 After each step of the LSTM, the receiver calculates the dot product between the hidden state and all of its image encodings.
@@
 Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one object.
-Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape) we expect that the vocabulary onto which the agents converge reflects these categories and therefore be close to human vocabulary.
+Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and therefore be close to human vocabulary.
 There are 48 possible combinations of attributes.
-Still for Dale-2 a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
+Still, for Dale-2 a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
 This hints to the fact, that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
@@
 % SD 2023-07-10 18:04:24 +0200: It would be interesting to compare then the length of messages between Dale-2 and Dale-5 that becomes used.
+% DK 2023-07-10: Yes, especially with a more detailed analysis of the vocabulary.
+Right now, the average is always 2 symbols, for both Dale-2 and Dale-5 as well as for all different vocabulary sizes.
+So far, I didn't check if there are sometimes longer messages and sometimes shorter messages.
 
@@
 in particular whether
-this uses similar categories as human language and how its words are combined to form complete messages.
+it uses similar categories as human language and how its words are combined to form complete messages.
 In our future work we will also extend learning to relations between entities and the features required to capture them.

### 2023-07-10 paper/paper.tex
corrections

@@
 \begin{abstract}
-  How do artificial agents based on neural networks coordinate on a new language through referential games over 3-dimensional scenes?
+  How do artificial agents based on neural networks coordinate on a new language through referential games over 3-d scenes?
   We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
@@
 
-Agents interact with the physical world through their action and perception, and with other agents through language.
+Agents interact with the physical world through their actions and perception, and with other agents through language.
 Their sensors and actuators % (if they are artificial agents, but the same also holds for natural agents)
-allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and other measures which can be measured with a high degree of accuracy.
+allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
 On the other hand, the language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete and arbitrarily chosen symbols.
@@
 How is ambiguity and underspecification resolved?
-How can agents achieve this through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?
+How can agents achieve it through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?
 
@@
 The language games were developed and run in the EGG framework \citep{Kharitonov2019}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-Both our sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} of \citep{Lazaridou2016} as shown in Appendix \ref*{sec:setup-language-game}.
+Both our sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} of \citep{Lazaridou2016}, as shown in Appendix \ref*{sec:setup-language-game}.
 One central difference is the production of the message.
 % This research focuses on the identification of attributes for the objects and their combination, which is why
-As we focus on sequences of referring expressions made-up of different attributes, our models produce sequences of symbols for the message % (which may correspond to the attributes)
-instead of a single symbol referring to an image. % (which corresponded to different concepts of the images).
-This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) to encode language descriptions.
+As we focus on sequences of referring expressions, made-up of different attributes, our models produce sequences of symbols for the message % (which may correspond to the attributes)
+instead of a single symbol to refer to an image. % (which corresponded to different concepts of the images).
+This is done by using an encoder LSTM (sender) and a decoder LSTM (receiver) to encode language descriptions.
 %
@@
 % The order of the objects for the receiver is random.
-Another difference is that both sender and receiver receive visual input as segmented objects rather than two images.
+Another difference is that both sender and receiver receive visual input as segmented objects rather than as two images.
 The order of the objects is random, except that the first object for the sender is always the target object to be referred to.
@@
 This is then used as the initial state of the encoder LSTM.
-The sequence is then created through Gumbel-Softmax relaxation \citep{Jang2016}.
+After, the sequence is created through Gumbel-Softmax relaxation \citep{Jang2016}.
 %
@@
 
-\begin{table}[b]
+\begin{table}
   \centering
@@
 % The results of the experiments are summarized in Table \ref{tab:results}.
-For the \emph{Dale-2} dataset it can be clearly seen that an embedding and hidden size that are as high as the vocabulary size are beneficial for identifying the correct object.
+For the \emph{Dale-2} dataset it can be clearly seen that an embedding size and hidden size that are as high as the vocabulary size are beneficial for identifying the correct object.
 The receiver identifies almost every sample correctly with all sizes of 10.
@@
 % The results change, when using the \emph{Date-5} dataset with four distractors.
-\emph{Date-5} with four distractor objects and with low hidden, embedding and vocabulary size, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\% which is still considerably lower than 95\% of the \emph{Dale-2} dataset.
+Using \emph{Dale-5} with four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
+Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\% which is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
 % This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.
@@
 % The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one object.
-Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and therefore be close to human vocabulary.
+Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
+Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
 There are 48 possible combinations of attributes.
-Still, for Dale-2 a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact, that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher, when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols to attest to this higher level of detail and especially how to relate them to features in the images.
+Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
+This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
+The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
+Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
 % SD 2023-07-10 18:04:24 +0200: It would be interesting to compare then the length of messages between Dale-2 and Dale-5 that becomes used.
@@
 it uses similar categories as human language and how its words are combined to form complete messages.
-In our future work we will also extend learning to relations between entities and the features required to capture them.
+In future work we will also extend the learning to the relations between entities and the features required to capture them.
 % Entries for the entire Anthology, followed by custom entries

### 2023-07-10 paper/paper.tex
typos

@@
 % For these experiments,
-We extend the CLEVR dataset \citep{Johnson2016}, by dividing the objects into one \emph{target object} and \emph{distractors}
+We extend the CLEVR dataset \citep{Johnson2016} by dividing the objects into one \emph{target object} and \emph{distractors}
 % The target object is the main object in the scene that should be identified and communicated by the agents.
@@
 The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
-The sequence received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
+The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
 After each step of the LSTM, the receiver calculates the dot product between the hidden state and all of its image encodings.
-The receiver then `points' to one of the images, by applying the softmax function over the results of the dot products.
+The receiver then `points' to one of the images by applying the softmax function over the results of the dot products.
 The loss is calculated using the NLL-loss.
-Following, the losses for all steps are summed up, and all weights of the receiver as well as the sender are updated, based on this summed loss.
+Following, the losses for all steps are summed up, and all weights of the receiver as well as the sender are updated based on this summed loss.
 

### 2023-07-11 thesis/sections/3-exp-setup.tex
git friendly latex

@@
 For that reason the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generated images for the CLEVR dataset.\footnote{The extended source code can be found on GitHub: \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}} To simplify the generation, I only focused on the attributes \emph{shape}, \emph{size} and \emph{color}.
+Following, I extended the source code to generated images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+To simplify the generation, I only focused on the attributes \emph{shape}, \emph{size} and \emph{color}.
 The \emph{material} is always the same for all objects in a generated image.

### 2023-07-14 thesis/sections/3-exp-setup.tex
describe coordinate predictors

--- 2023-06-24 thesis/sections/3-exp-setup.tex
@@
 \cmtDK[inline]{15 pages}
-\cmtDK[inline]{\\
-    - what is ambiguity? \\
-    - unambigous dataset: \\
-    ---- target group = non-target group \\
-    ---- all attributes shared, except for the color \\
-    ---- \ \  --> describable with one 'word'
-}
+
 In this section I will describe which experiments were conducted to answer the \cmtDK{?}{questions} from the previous sections as well as the setup of these experiments.
@@
 \subsection{Creation of the dataset}
-The basis for my dataset is the CLEVR dataset (\cite{Johnson2016}).
-This dataset includes 3D-generated images depicting scenes with different kind of objects.
+The basis for my datasets is the CLEVR dataset (\cite{Johnson2016}).
+This dataset includes 3D-generated images depicting scenes with different kinds of objects.
 Each of these objects has different combinations attributes, such as \emph{shape}, \emph{color}, \emph{size} and \emph{material}.
@@
 Furthermore, the dataset contains information about each scene.
-This includes the selected attributes for each object as well as the exact position of the centers of all the objects, both 3D-coordinates in the 3D scene and 2D-coordinates in the final rendered image.
+This includes all selected attributes for each object as well as the exact position of the centers of all the objects, both 3D-coordinates in the 3D scene and 2D-coordinates in the final rendered image.
 In addition, simple spatial relations (in front of, behind, left, right) between the objects are calculated and stored.
-These are simply based on the 3D-coordinates of the objects from the perspective of the camera position.
+These are simply based on the 3D-coordinates of the objects in relation to the position of the camera.
 
@@
 This research investigates how agents communicate about the relations of objects seen in images.
-For that reason the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generated images for the CLEVR dataset.\footnote{The extended source code can be found on GitHub: \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}} To simplify the generation, I only focused on the attributes \emph{shape}, \emph{size} and \emph{color}.
+For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
+Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+To simplify the generation and the later learning of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
 The \emph{material} is always the same for all objects in a generated image.
@@
 These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex and are expected to teach the agents more precise descriptions.
+Finally, the distractors are objects that add noise to the scene and should make it more complex. They expected to teach the agents more precise descriptions of the target object.
 The number of the objects in both group can be controlled.
@@
 In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with objects in the target group and distractors respectively.
+The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
 For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere} but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
+If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
 The number of shared attributes can also be set to a range to make the discrimination task more challenging.
@@
 Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have randomly selected different shapes and color.
-This allows a very controlled generation of relations between the objects in the scene.
+For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
+This allows for a very controlled generation of relations between the objects in the scene.
 Figure \ref{fig:clevr-extended_example} shows one generated image with this extended source code.
@@
 It is not controlled, which attributes are shared (they are selected randomly).
+The large purple cylinder shares the same color and size with the large purple spherem, the same size with both cubes and no attribute with the small turquoise sphere.
 There are no distractor objects.
@@
 10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance and are partially visible.
+Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are partially visible from the camera.
 
@@
 The models need to solve multiple tasks at the same time to solve the overall problem.
-For instance in a simple setup of a game two agents are involved.
-The first agent, the sender is shown a scene with objects and needs to communicate on target object to the other agent, the receiver.
+For instance, in a simple setup of a game two agents are involved.
+The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
 The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene as well as the information about the target object into its own \cmtDK{?}{space}.
+In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own \cmtDK{?}{space}.
 In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene.
+The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
 And finally it needs to learn how to identify the target object with this information.
 
-For this reason I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.
+For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.
 This will give me a very detailed overview, where the models struggle to learn and in which ways they can be improved.
@@
 
+\subsubsection{Feature extractors}
+To extract features from the images, I make use of different deep neural network architectures that are developed for this task.
+First, I use the \cmtDK{why ResNet/VGG?} VGG19 (\cite{Simonyan2015}) which is an architecture based on many convolutional layers.
+After the convolutional layers, the data is passed first through an average pooling layer which outputs 512x7x7 dimensions.
+Next follow three linear layers with \emph{ReLU} non-linearities in between.
+After flattening the input, these classification layers output 4069, 4069 and 1000 dimensions respectively.
+
+Secondly, I include the ResNet-101 (\cite{He2016}) that tries to overcome problems of very deep networks, using \cmtDK{needs to be explained deeper?}{residual blocks}.
+There are four blocks that output 256x56x56, 512x28x28, 1024x14x14 and 2048x1x1 dimensions.
+A following average pooling layer outputs 2048x1x1 dimensions as well.
+The final linear layer reduces the flattened data to 1000 dimensions, corresponding to the ImageNet classes.
+
+Both architectures are available pretrained on an image classification task on the ImageNet dataset.
+Since the task in this research is very different from a classification, multiple different adaptions of these architectures are compared.
+\cmtDK{deeper discussion about different layers of these networks}
+Table \ref*{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
+In the later chapters, it will be referred to these adaptions using the name in the table.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{rlc}
+        \toprule
+                            & \textbf{description}                                            & \textbf{ output dimensions } \\\midrule
+        \textbf{VGG-0}      & contains only the convolutional layers                          & 512x7x7                      \\
+        \textbf{VGG-avg}    & contains an additional average pooling layer                    & 512x7x7                      \\
+        \textbf{VGG-cls1}   & \makecell[cl]{contains an additional one classification layer,                                 \\ including its non-linearity} & 4069                         \\
+        \textbf{VGG-cls2}   & \makecell[cl]{contains another additional classification layer,                                \\ including its non-linearity} & 4069                         \\
+        \textbf{VGG-cls3}   & the original VGG19 architecture                                 & 1000                         \\\midrule
+        \textbf{ResNet-1}   & contains one residual block                                     & 256x56x56                    \\
+        \textbf{ResNet-2}   & contains two residual blocks                                    & 512x28x28                    \\
+        \textbf{ResNet-3}   & contains three residual blocks                                  & 1024x14x14                   \\
+        \textbf{ResNet-4}   & contains four residual blocks                                   & 2048x1x1                     \\
+        \textbf{ResNet-avg} & contains an additional average pooling layer                    & 2048x1x1                     \\
+        \textbf{ResNet-cls} & the original ResNet-101 architecture                            & 1000                         \\
+        \bottomrule
+    \end{tabular}
+    \caption{Different adaptions of VGG19 and ResNet-101 used in this research}
+    \label{tab:feature-extractor-archs}
+\end{table}
+
+Furthermore, I experimented with both the pretrained models as well as with the architectures trained from scratch with a random initialization for the weights.
+This reason for this was to test if the success of an experiment was actually making use of the pretrained knowledge incorporated in the models.
+If that was not the case, the agents were likely not using image features, but instead relying on some other underlying patterns to solve the task.
+Basically, this approach works as an indicator to determine the actual success of the agents aside from measures as the accuracy or precision.
+
 \subsubsection{Pre-experiments without language games}
-\paragraph{coordinate predictor}
-\cmtDK[inline]{\\
-    - pretrained ResNet/VGG\\
-    - scratch \\
-    - encoding of attributes (one-hot) \\
-    - encoding of attributes (dale) \\
-    - encoding of locations \\
-    - masking}
-\paragraph{bounding box classifier}
+The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
+Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge.
+Here, I structured the experiments into three levels.
+In the first level, the model are trained to learn the position of objects in the image and attend to specific regions of the image.
+In the second level, the models are trained to differentiate objects in the scene from each other.
+The last level adds language to the experiment, more specifically the models learn to caption and describe objects in the image.
+These combined experiments should lay the basis for how to build up the agents in the language games.
 
-\paragraph{caption generator}
+\paragraph{Level 1 - coordinate predictor:}
+
+This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
+As described before, the sender should communicate an object in the image and the receiver needs to identify it.
+The challenge lies in how the receiver refers to the identified object.
+There are multiple possibilities, how it can be done.
+One of them could be to describe the target object with human language, using the attributes.
+The main goal however is to let the language of the agents emerge as natural as possible.
+Including human knowledge into the task would bias also the emerged language towards attributes and words, humans are using.
+For this reason, the final task of the receiver will be to 'point' to the target object.
+Since also positional information of the objects should be part of the discrimination, the image will not be separated into several bounding boxes around the objects, but instead the whole image is the basis.
+The models are therefore tasked to predict the center coordinates of the target object.
+With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
+
+To achieve this goal, multiple setups of models are tested.
+In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+Here the image is first passed through one of the \emph{feature extractors}.
+Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity between.
+These reduce the dimensions first to 1024 and finally to 2.
+
+To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
+This distance is learned to be minimized.
+By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target object.
+
+With this simple setup, the model is in theory able to focus on an object in the image.
+The problem arises as soon as multiple objects are present in the image.
+There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
+Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize. \cmtDK{go deeper, why}
+Therefore, the models need to receive more information.
+Here, I try out four different ways to encode the target object.
+
+In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
+There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
+The dimensions of these vectors can either be zero or one, depending on the attributes of the target object.
+These three encodings of the attributes are then concatenated.
+The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
+The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
+
+In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
+This should help the model to identify all possible options to chose from, when predicting the target object.
+All the center coordinates are simply extracted and shuffled.
+Since there are varying numbers of objects in the image, this vector of variable length is padded to 10 objects, resulting in a vector with a length of 20.
+The padding objects have zeros for both coordinates, which no other real object can have.
+For this model, I also made use of a more complex way to encode the image.
+Here, I based the approach on code that implements baselines for the CLEVR dataset (\cite{Johnson2017}). \cmtDK{why?}
+The image is first passed through the feature extractor.
+Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
+After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
+The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
+This is then passed again through a final linear layer to predict x- and y-coordinates.
+
+The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \cite{Dale1995}.
+This opposes the idea described before, to share as few human knowledge as possible with the model.
+Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
+If the model is able to match parts of the image with human words it would show that the model learned this attribute.
+If the model in a next step can learn this over the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
+This insight could help then to design another model that makes use of these learnings without human language.
+
+Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
+In other words, the object is described unambiguously using the lowest number of words.
+The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
+This order defines, which attributes can be left, while still identifying the object uniquely.
+For our research, we rely on the following order from most important to least important: shape, color, size
+Given for example the scene from \ref*{fig:clevr-extended_example} with the target object being the \emph{big purple cylinder}.
+Using all three attributes, this description identifies the object perfectly and uniquely.
+Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
+This can be taken even one step further by removing also the \emph{color}.
+Describing it as the \emph{cylinder} still doesn't describe any other object, since the target object is the only cylinder in the scene.
+
+For the experiments, each every image is captioned, by a description of the target object using the described algorithm.
+To use it in the model, the caption needs to be padded to an equal length, in this case to a length of 3, which is the maximum number of attributes that can be used.
+For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
+
+The model is made up of three parts:
+The first part extracts features from the image.
+Here, the setup is similar to the previous model.
+The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
+A \emph{ReLU} function is applied, after both convolutional layers.
+As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
+In the second part, the caption is encoded, using an LSTM.
+Here, the learned embedding of each token is parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
+The third part is again the predictor of the coordinates of the target object.
+Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
+The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
+Between each linear layer, the \emph{ReLU} function is applied.
+This architecture is also inspired by the baselines described in \cite{Johnson2017}.
+
+The forth method, to encode the target object utilizes \textbf{masking} of the image.
+For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
+The side of the square is always two fifth of the image width, in this case 192 pixels.
+This size always includes the whole object if it is large or small, while not being cut of if the target object is too close to the border of the image.
+The square is filled in white, while the rest of the image is filled in black.
+This approach has the advantage of providing as few as possible human bias to the model.
+While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
+It therefore can only rely on its own extracted visual features when looking at masked images.\cmtDK{can just rely on masked image when predicting coords, more useful for captioning or language game}
+
+The model is presented with both the original image and the masked image.
+Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
+The resulting vectors are concatenated and passes to another linear layer that predicts the coordinates.
+
+\paragraph{Level 2 - bounding box classifier:}
+
+\paragraph{Level 3 - caption generator:}
 \cmtDK[inline]{\\
     - dale \\
+    - - different paddings (back, front, reversed) \\
     - masked}
+
 \subsubsection{Language games}

### 2023-07-14 thesis/sections/3-exp-setup.tex
describe bounding-box classifier, caption generator

@@
 For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.
+\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
 This will give me a very detailed overview, where the models struggle to learn and in which ways they can be improved.
@@
 In the second part, the caption is encoded, using an LSTM.
-Here, the learned embedding of each token is parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
+Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
 The third part is again the predictor of the coordinates of the target object.
@@
 Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passes to another linear layer that predicts the coordinates.
+The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
+
+\cmtDK{test accuracies}
 
@@
 
+One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
+Feature extractors like VGG or ResNet are trained on extracting visual information.
+In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
+To address this challenge, I restructured the task in this experiment.
+Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
+This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
+To do this, fixed-size bounding boxes are extracted around each of the object in the image.
+As for the masking, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
+The bounding boxes of all objects are padded to a number of 10 bounding boxes, the maximum number of objects in an image with a matrix of zeros.
+These bounding boxes are shuffled.
+
+The model passes each of the bounding boxes through a feature extractor.
+The resulting vectors are concatenated, which basically is a concatenation of the extracted features of each object.
+This long vector is passed through a linear layer with 4096 with a subsequent \emph{ReLU} function.
+Afterwards, a linear layer reduces the dimensions to 10, corresponding to the 10 bounding boxes in the input.
+With a final \emph{softmax} function, the model then points to one of the bounding boxes as the target bounding box.
+The loss is calculated using cross entropy.
+
 \paragraph{Level 3 - caption generator:}
-\cmtDK[inline]{\\
-    - dale \\
-    - - different paddings (back, front, reversed) \\
-    - masked}
+
+As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
+It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
+This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
+There are some minor additions concerning the padding of the caption.
+As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
+
+However, there are three different ways how the padding is applied.
+First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
+A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
+More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
+The second last token if existing describes the color, while the third last token if existing describes the size.
+As soon as this sequence is padded at the end, these slots suddenly disappear.
+A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
+Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
+For this reason, the second method of padding the caption is prepending the caption with padding tokens.
+By this, the positions of the slots are preserved and if not specified just filled with a padding token.
+The last variation concerns the order of producing each token.
+When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
+This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
+Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
+Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
+Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
+
+Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
+Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
+
+I compare two different models against each other.
+The first model just takes the image as the input.
+During training, also the ground truth caption is passed to the model.
+The image is processed as in the section before.
+First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
+After each convolutional layer the \emph{ReLU} function is applied.
+The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
+The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
+During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
+No teacher forcing is applied.
+The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
+The loss is calculated using cross entropy.
+When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
+Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
+
+Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
+As in the section above, a masked version of the image is created and passed to the model.
+The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
+Both encodings are concatenated and then used as the initial hidden state for the LSTM.
+This should point the model to which object to describe and discriminate from the other objects.
+
+To test the success of the model, three measures are calculated.
+The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
+This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
+However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
+
+It could be the case that the model predicted the correct shape, but wrong color.
+Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
+For instance, consider the scene in Figure \ref*{fig:clevr-extended_example}.
+The correct caption is \emph{cylinder}.
+If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
+The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
+To mitigate this, I also include a \textbf{word-by-word accuracy}.
+This measure calculates the accuracy not on sentence level, but on word level.
+The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
+yield 0\%.
+This can give a better understanding of the errors the model makes.
+
+With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
+This is basically an inverted accuracy score; the lower the score, the better the model fares.
+For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
+If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
+If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
+Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
 

### 2023-07-15 thesis/sections/3-exp-setup.tex
describe datasets

@@
 
-\cmtDK[inline]{ambiguity}
+\cmtDK{ambiguity}
+\cmtDK{referring expressions}
 
@@
 This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-extended_example} shows one generated image with this extended source code.
+Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
 Here, the target object is the large purple cylinder.
@@
 It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple spherem, the same size with both cubes and no attribute with the small turquoise sphere.
+The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
 There are no distractor objects.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}
-    \caption{Example of a generated image with the extended code}
-    \label{fig:clevr-extended_example}
-\end{figure}
 
@@
 Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are partially visible from the camera.
+
+\begin{figure}[h]
+    \centering
+    \subfigure['CLEVR single', \emph{large yellow sphere}]{
+        \includegraphics[width=0.42\linewidth]{figures/CLEVR_single.png}
+        \label{fig:clevr-single}
+    }
+    \subfigure['CLEVR unambigous', \emph{large green sphere}]{
+        \includegraphics[width=0.42\linewidth]{figures/CLEVR_unambiguous.png}
+        \label{fig:clevr-unambiguous}
+    }
+    \subfigure['CLEVR color', \emph{small brown cylinder}]{
+        \includegraphics[width=0.42\linewidth]{figures/CLEVR_color.png}
+        \label{fig:clevr-color}
+    }
+    \subfigure['Dale-2', \emph{small green cylinder}]{
+        \includegraphics[width=0.42\linewidth]{figures/CLEVR_dale-2.png}
+        \label{fig:clevr-dale-2}
+    }
+    \subfigure['Dale-5', \emph{large purple cylinder}]{
+        \includegraphics[width=0.42\linewidth]{figures/CLEVR_dale-5.png}
+        \label{fig:clevr-dale-5}
+    }
+    \caption{Example images of each dataset, with the target object specified}
+    \label{fig:clevr-examples}
+\end{figure}
+
+\subsubsection{CLEVR single}
+The simplest dataset is called 'CLEVR single'.
+This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
+Each scene in the dataset contains only one single object, the target object.
+There are neither objects in the target nor in the distractor group.
+All attributes are assigned randomly to the target object.
+The differences across the whole dataset are the locations and rotations of the objects.
+With this dataset, neural models can focus on only the features, as well as the locations of this single object.
+There are no objects that distract the model from extracting features from the target object.
+This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
+Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
+
+\subsubsection{CLEVR unambiguous}
+The second dataset that is created is called 'CLEVR unambiguous'.
+The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
+For this reason, there exist only two groups in the scene, the target object and distractors.
+The distractor group can contain in between 6 and 9 objects.
+The target object doesn't share any attributes with the distractors.
+Namely, if the target is for instance green, no other object in the scene is green as well.
+This also applies to the other attributes.
+When referring to this target object, it is therefore enough to only describe one of the attributes.
+
+Consider the scene in Figure \ref{fig:clevr-unambiguous}.
+The target object can be completely described as the \emph{large green sphere}.
+On the other hand, also descriptions like the \emph{green object}, the \emph{large object} or the \emph{sphere} would be unambiguously referring to the target object.
+
+\subsubsection{CLEVR color}
+This task could on the other hand be difficult to learn for neural models.
+Although the target object is unique in the scene with resect to the distractors, the description isn't.
+Considering all unordered combinations of the three attributes and a composite referring expression, there are $\binom{3}{3} + \binom{3}{2} + \binom{3}{1} = 7$ possible descriptions of the target object.
+This can propose a difficulty, when the models try to converge towards one representation.
+For this reason, I created the 'CLEVR color' dataset.
+It has a similar setup as the previous dataset, containing between 6 and 9 distractors.
+But instead of the target object sharing no attribute with the distractors, they share exactly two attributes.
+Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
+The distractors have always the same size and shape as the target object, but the color is always different.
+The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
+
+As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
+The number of possible descriptions is lowered to three, since the color \emph{brown} needs to be part of it.
+Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
+
+\subsubsection{CLEVR Dale datasets}
+Both of the above described datasets are very restrictive on the disambiguating relations between the target object and the rest of the objects.
+Either no attribute or only one attribute is shared.
+More realistically would be if objects can share a variable number of attributes.
+In real situations, there is no restriction at all, how objects or things relate to each other.
+Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
+This emergence of referring attributes and their combination is studied deeper in this work.
+
+For this, I created a dataset that allows almost any relation between a target object and the distractors.
+The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \cite{Dale1995}.
+This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
+Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
+In other words, the object is described unambiguously using the lowest number of words.
+For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
+This ensures the uniqueness of the target object.
+On the other side, it is not controlled which attributes are shared.
+These are assigned randomly.
+There is again no control over the relations between distractors, which means that distractors can appear multiple times.
+
+Two datasets, following these rules are created.
+The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
+Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
+The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
+
+These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
+More specifically how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
 
@@
 \cmtDK{deeper discussion about different layers of these networks}
-Table \ref*{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
+Table \ref{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
 In the later chapters, it will be referred to these adaptions using the name in the table.
@@
 For our research, we rely on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref*{fig:clevr-extended_example} with the target object being the \emph{big purple cylinder}.
+Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
 Using all three attributes, this description identifies the object perfectly and uniquely.
@@
 Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref*{fig:clevr-extended_example}.
+For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
 The correct caption is \emph{cylinder}.

### 2023-07-18 thesis/sections/3-exp-setup.tex
language games + discrimination games

@@
 The first part contains the creation of the datasets that will be used in the experiments.
-In the second part I will go deeper into discussing each experiment, including the necessity of the experiment, the specifities of the dataset, as well as the final architecture of the models.
+In the second part I will go deeper into discussing each experiment, including the necessity of the experiment, which datasets the model is trained on, as well as the final architecture of the models.
 
 \subsection{Creation of the dataset}
-The basis for my datasets is the CLEVR dataset (\cite{Johnson2016}).
+The basis for my datasets is the CLEVR dataset \citep{Johnson2016}.
 This dataset includes 3D-generated images depicting scenes with different kinds of objects.
@@
     }
-    \subfigure['CLEVR unambigous', \emph{large green sphere}]{
+    \subfigure['CLEVR unambiguous', \emph{large green sphere}]{
         \includegraphics[width=0.42\linewidth]{figures/CLEVR_unambiguous.png}
@@
 For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \cite{Dale1995}.
+The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
 This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
@@
 More specifically how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
+
+\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
 
@@
 To extract features from the images, I make use of different deep neural network architectures that are developed for this task.
-First, I use the \cmtDK{why ResNet/VGG?} VGG19 (\cite{Simonyan2015}) which is an architecture based on many convolutional layers.
+First, I use the \cmtDK{why ResNet/VGG?} VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
 After the convolutional layers, the data is passed first through an average pooling layer which outputs 512x7x7 dimensions.
@@
 
-Secondly, I include the ResNet-101 (\cite{He2016}) that tries to overcome problems of very deep networks, using \cmtDK{needs to be explained deeper?}{residual blocks}.
+Secondly, I include the ResNet-101 \citep{He2016} that tries to overcome problems of very deep networks, using \cmtDK{needs to be explained deeper?}{residual blocks}.
 There are four blocks that output 256x56x56, 512x28x28, 1024x14x14 and 2048x1x1 dimensions.
@@
 
-\paragraph{Level 1 - coordinate predictor:}
+\paragraph{Level 1 - Coordinate predictor:}
 
@@
 For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset (\cite{Johnson2017}). \cmtDK{why?}
+Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}. \cmtDK{why?}
 The image is first passed through the feature extractor.
@@
 
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \cite{Dale1995}.
+The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
 This opposes the idea described before, to share as few human knowledge as possible with the model.
@@
 Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \cite{Johnson2017}.
+This architecture is also inspired by the baselines described in \citet{Johnson2017}.
 
@@
 
-\paragraph{Level 2 - bounding box classifier:}
+The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
+The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
+Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
+Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
+In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
+Here, the models should make use of the additional given information about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
+'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
+Latter task is assumed to be significantly harder.
+
+\paragraph{Level 2 - Bounding box classifier:}
+\cmtDK{better discriminators?}
 
@@
 
-\paragraph{Level 3 - caption generator:}
+The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
+The 'CLEVR unambiguous' and 'CLEVR color' datasets are directly comparable to each other for the similar setup of their creation.
+Especially interesting is the effect of the number of shared attributes.
+Does it help the model if the target object is unique in every attribute or does it help if there is only one specific attribute that is shared?
+A similar assumption can be made for the 'Dale' datasets.
+How big is the effect of increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
+
+\paragraph{Level 3 - Caption generator:}
 
@@
 
+The caption generator models are trained on both 'Dale' datasets.
+Again, each of these datasets increases the complexity of the description.
+While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
+Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
+
+\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
+
 \subsubsection{Language games}
+The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.
+
+The language games in this research have an asymmetric setup.
+One agent, the sender is shown some information and needs to generate a message.
+This message is received by the second agent, the receiver.
+The receiver needs to parse this message and combine it with the information they are given as well.
+The receiver then makes a prediction, which is compared to the ground truth.
+The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
+All weights of the agents are adapted in the same way.
+\cmtDK{Gumbel-Softmax vs REINFORCE}
+\cmtDK{better in theoretical background?}
+\cmtDK{vocabulary?}
+
+All language games are set up and run in the EGG framework \citep{Kharitonov2019}, developed in PyTorch.
+This framework consists of heavy configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
+The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
+Furthermore, runs of games can be saved to analyze the used messages of the agents and how they change over the duration of the learning.
+
+The framework is hereby setup in three levels.
+Part of the lowest level are the \emph{agents} themselves.
+The agents are neural models that need to be developed from scratch and define how, the agents process their input and combine it with the message (in case of the receiver) and what they output.
+The second level are \emph{wrappers} that take care of generating and parsing the message.
+The sender wrapper uses the output of the sender agent, to produce a message.
+The receiver on the other hand parses the message received by the sender and passes the result as an additional input to the receiver agent.
+The third level, the \emph{game} links all described parts together.
+It provides the agents with the input and passes the message from the sender to the receiver.
+Furthermore, it uses the output of the receiver and calculates the loss, which is then the basis for the adaption of the weights for both wrappers and agents.
+
+For the following language games, the sender will always produce a sequence of symbols as a message, which the receiver will parse.
+Gumbel-Softmax relaxation is applied to produce discrete symbols.
+This is done using two LSTMs, an encoder LSTM in the sender wrapper and a decoder LSTM in the receiver wrapper.
+The output of the sender agent is used as the initial hidden state for the encoder LSTM.
+This LSTM is then producing symbols until it generates an end-of-sequence symbol.
+This sequence is then passed to the receiver wrapper with its decoder LSTM.
+Its hidden state is initialized randomly.
+The received message sequence is processed symbol by symbol.
+After each time, a symbol is processed by the LSTM, the resulting new hidden state is passed to the receiver agent as the parsed message.
+The receiver agent is combining it with its representation of the image input and predicting an output.
+The \emph{game} is afterwards calculating a loss for each of these outputs separately.
+These losses are summed up to a total loss that is used to adapt the weights in both agents as well as in both LSTMs.
+
+First, I will discuss \emph{discrimination games}, because they have the simplest setup.
+Furthermore, other language games that research this topic use a very similar setup.
+In the next step, I will look into caption generators that are set up as a language game.
+Here, the sender describes the scene, while the receiver needs to generate a caption.
+In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
+
+\paragraph{Discrimination games:}
+In a discrimination game, the agents are presented with two or more images, one of these being the target image.
+The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
+The receiver then needs to decide based on the message, which of the images is the target image.
+
+The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2016}.
+The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
+One central difference is the production of the message.
+The main goal of their language game was the identification of the concept that and image was related to.
+Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
+Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
+It is therefore assumed that the sender will communicate these discriminative attributes.
+For that reason, the sender is allowed to generate sequences as a message.
+
+Both 'Dale' datasets are used to train the agents in this mode.
+Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
+As in \citet{Lazaridou2016}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
+The sender is assumed to learn which of the input images is the target images by this approach.
+The receiver receives the images in completely shuffled order.
+
+For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
+All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
+This is then used as the initial state of the encoder LSTM in the sender wrapper.
+
+The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
+The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
+As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
+This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
+This allows the calculation of the dot product between it and the image representation.
+If the message describes an object well, the resulting dot product should be higher.
+The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
+The loss is calculated using the NLL-loss.
+
+During the experiments, five variables are adjusted to compare their effects:
+(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
+
+\paragraph{Caption generators:}
+
+\paragraph{Coordinate predictors:}

### 2023-07-25 thesis/sections/3-exp-setup.tex
results

@@
 
-\subsubsection{CLEVR single}
+\subsubsection*{CLEVR single}
 The simplest dataset is called 'CLEVR single'.
@@
 
-\subsubsection{CLEVR unambiguous}
+\subsubsection*{CLEVR unambiguous}
 The second dataset that is created is called 'CLEVR unambiguous'.
@@
 
-\subsubsection{CLEVR color}
+\subsubsection*{CLEVR color}
 This task could on the other hand be difficult to learn for neural models.
@@
 
-\subsubsection{CLEVR Dale datasets}
+\subsubsection*{CLEVR Dale datasets}
 Both of the above described datasets are very restrictive on the disambiguating relations between the target object and the rest of the objects.
@@
 
-\cmtDK{test accuracies}
+The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
+This distance needs to be minimized.
+The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
+Since this score only takes the average of all predictions into account it can't evaluate every prediction individually.
+If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
+It doesn't reflect this change.
+For that reason, I also introduced an accuracy score.
+For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
+If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
+These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
+This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
+A high accuracy may indicate that the model could identify this specific object better.
 
@@
 \paragraph{Level 2 - Bounding box classifier:}
-\cmtDK{better discriminators?}
+\cmtDK{better to call it discriminators?}
 

### 2023-07-25 thesis/sections/4-results.tex
results

@@
 \cmtDK[inline]{15 pages}
+\subsection{Pre-experiments without language games}
+\subsubsection{Coordinate predictor}
+%     \item visualizations of predictions
+%     \item difference between layers of feature extractors (especially compare to baseline learned from scratch)
+%     \item masking is helping very much (especially for Dale-2)
+%     \item visualizations show that model only focuses on center of the image
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
+    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
+    \label{fig:coordinate-predictor_loss}
+\end{figure}
+
+Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
+The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ significantly with other feature extractors.
+As can be seen, the success between the different datasets differ significantly.
+The more objects are present in an image, the worse the model performs.
+The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
+This prediction even though not perfectly on the center point is always on the object.
+Opposed to that, using the Dale-2 dataset with two objects, the mean distance is from the first epoch between 37 and 38 pixels.
+With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
+
+An interesting observation is the difference of the mean distances between training and testing data.
+The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
+This points to the fact that the model is not generalizing the task and learning abstract patterns that can be applied to unseen data, but instead learning the training data by heart.
+That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
+Applying a higher dropout didn't have an impact on the results.
+
+This behavior is indeed not very surprising.
+First, these results show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
+Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
+Secondly, more than one objects present in an image confuse the model, and it is not able to consistently point to one of them.
+This can have multiple reasons, for instance that models lacks the ability to separate objects in the extracted features.
+But even given, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
+The guess is then more or less random.
+This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
+For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
+This task on the other hand seems very difficult to learn.
+In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
+
+When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
+One factor that now has a much higher impact is the feature extractor that is used.
+Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
+The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
+First, a big difference can be seen between the datasets.
+The model doesn't converge on the correct coordinates for the Dale-5 dataset at all, looking at the test data.
+All resulting mean distances are in a similar range around 50 pixels.
+Using the Dale-2 dataset, the behavior is a little different.
+All ResNet extractors with four residual blocks and optionally additional average pooling and classifier layers reach similar scores as the experiments before without any one-hot encodings.
+Interestingly, using less residual layers, the model doesn't converge at all and the mean distances jump up and down between the epochs.
+This effect also applies when using less residual blocks.
+Using the VGG, both using VGG-avg and VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
+\cmtDK{compare to scratch}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{rcccc}
+        \toprule
+                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
+                            & train                               & test                                & train          & test           \\\midrule
+        \textbf{VGG-0}      & 30,26                               & 46,2                                & \textbf{10,81} & 48,33          \\
+        \textbf{VGG-avg}    & \textbf{9,96}                       & 36,9                                & 32,17          & 54,40          \\
+        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
+        \textbf{VGG-cls2}   & 34,95                               & \textbf{36,32}                      & 47,63          & \textbf{46,38} \\
+        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & 46,77          \\\midrule
+        \textbf{ResNet-3}   & 92,07                               & 91,12                               & 11,52          & 48,52          \\
+        \textbf{ResNet-4}   & \textbf{8,09}                       & \textbf{36,78}                      & \textbf{9,05}  & 46,84          \\
+        \textbf{ResNet-avg} & 29,21                               & 38,94                               & 37,10          & 48,71          \\
+        \textbf{ResNet-cls} & 33,29                               & 38,83                               & 41,57          & \textbf{46,57} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
+    \label{tab:feature-extractor-mean-distances}
+\end{table}
+\cmtDK{wrong numbers}
+Secondly, the training loss now looks also different.
+In almost no cases, the models converge to a very low mean distance, meaning a high precision in their predictions, as they did in the experiment before.
+There are only exceptions when ResNet-4 is used or respectively VGG-avg for Dale-2 and VGG-0 for Dale-5.
+In other words, that means that the models are again not able to generalize, but in specific cases to learn the patterns in the train data by heart.
+This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
+Especially high layers, like the classifier layers and in the case of ResNet also the average pooling layers seem to already encode to task specific knowledge that was learned during the pre-training of these deep neural models.
+This information can't be utilized for the tasks in this research and earlier layers with more abstract information need to be used.
+The experiments in the following sections are set up using these lower layers as ResNet-3, ResNet-4, VGG-0 and VGG-avg.
+
+Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
+In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
+This hypothesis doesn't hold.
+All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
+The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
+The model is still not able to understand, which object is the target object.
+For that reason, a better representation of the target object is necessary.
+
+In a next step, information about the attributes is included using the \emph{GRE-algorithm} from \citet{Dale1995}.
+\cmtDK{change}
+
+\begin{figure}[h]
+    \centering
+    \subfigure['Dale-2', train split]{
+        \includegraphics[width=1\linewidth]{figures/visualization_dale-2_train.png}
+        \label{fig:visualizations_dale-2_train}
+    }
+    \subfigure['Dale-2', test split]{
+        \includegraphics[width=1\linewidth]{figures/visualization_dale-2_test.png}
+        \label{fig:visualizations_dale-2_test}
+    }
+    \subfigure['Dale-5', train split]{
+        \includegraphics[width=1\linewidth]{figures/visualization_dale-5_train.png}
+        \label{fig:visualizations_dale-5_train}
+    }
+    \subfigure['Dale-5', test split]{
+        \includegraphics[width=1\linewidth]{figures/visualization_dale-5_test.png}
+        \label{fig:visualizations_dale-5_test}
+    }
+    \caption{Visualization of the models' predictions in the 'Dale' datasets}
+    \label{fig:visualizations_dale}
+\end{figure}
+
+An interesting pattern appears when doing a qualitative analysis of the models' predictions.
+Here, I visualized the predicted coordinates compared to the ground truth coordinates.
+Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
+The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
+As can be seen, the prediction are very precise.
+Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
+Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are places as green circles.
+The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
+As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
+
+The results look very different for the test split.
+As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
+For all the images, the predictions don't lie on any object.
+In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
+These findings align with the mean distance scores, described in the sections before.
+However, it seems that the model's predictions are all towards the center of the image.
+This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
+Again, the green circles form the shape of rhombus.
+In contrast, the predictions in red almost all cluster in the center of the image.
+They form roughly the shape of a smaller rhombus.
+This behavior can be observed for all datasets and architectures of the model.
+Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
+Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
+Also the combined visualization shows the same clustering of predictions in the center of the scene, but pattern of the smaller rhombus is better visible.
+
+\begin{figure}[h]
+    \centering
+    \subfigure['Dale-2', train split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
+        \label{fig:visualizations_dale-2_complete_train}
+    }
+    \subfigure['Dale-2', test split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
+        \label{fig:visualizations_dale-2_complete_test}
+    }
+    \subfigure['Dale-5', train split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
+        \label{fig:visualizations_dale-5_complete_train}
+    }
+    \subfigure['Dale-5', test split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
+        \label{fig:visualizations_dale-5_complete_test}
+    }
+    \caption{Visualization of the models' predictions in the Dale datasets}
+    \label{fig:visualizations_dale_complete}
+\end{figure}
+
+These results allow two conclusions.
+First, the models are biased to predict
+
+Second, even though the models are biased towards the center of the image, the predictions are still often leaning in the direction of the target object.
+This indicates that it is still able to
+
+\subsubsection{Bounding Box classifier}
+
+\subsubsection{Caption generator}
+In the next paragraphs, the results of the caption generator will be discussed.
+Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
+ResNet-3 produced the best results, when it was used as a feature extractor.
+The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{rccc}
+        \toprule
+                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
+        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
+        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
+        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
+        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the caption generator, based on ResNet-3}
+    \label{tab:results_caption-generator}
+\end{table}
+
+The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
+This means that the models produce almost always correct descriptions of an object in the image.
+Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
+Features of both objects are therefore extracted perfectly and related to the vocabulary.
+As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
+With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
+This changes, when also the masked image is presented to the model.
+Then, the accuracy for the target object is with 72\% well above chance.
+Still, the wrong descriptions describe almost every time the distractor object.
+This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
+
+When the model is trained on the 'Dale-5' dataset, the results are much worse.
+Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
+With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
+Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
+The accuracy stays at the same value.
+In contrast, the non-target accuracy is increased by 5\% points.
+The model is more likely to identify a wrong object.
+Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
+This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
+For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
+After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
+This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
+
+The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
+When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
+The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
+
+There are two main possible explanations for the big difference between the two datasets.
+First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
+In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
+The second explanation lies in the used GRE-algorithm.
+When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
+For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
+Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
+With four distractors the algorithm is much more likely to produce longer captions.
+These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
+
+\subsection{Language games}
+\subsubsection{Discrimination games}
+There are five variables in the experiments that are adjusted:
+(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
+
+Table \ref{tab:results} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
+A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{c|ccccc|c}
+        \hline
+        \textbf{Dataset} & $h_{s}$ & $e_{s}$ & $h_{r}$ & $e_{r}$ & $|V|$ & \textbf{Acc.} \\
+        \hline
+        Dale-2           & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
+        Dale-2           & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
+        Dale-5           & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
+        Dale-5           & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
+        Dale-5           & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
+        \hline
+    \end{tabular}
+    \caption{Results: $h$ are different hidden sizes, $e$ embedding sizes and $|V|$ vocabulary sizes.}
+    \label{tab:results}
+\end{table}
+
+For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
+A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
+The receiver identifies almost every sample correctly when all sizes are 10.
+When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
+Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
+
+The results change, when using the \emph{Dale-5} dataset with four distractors.
+With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
+Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
+This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
+
+\cmtDK{calculation of loss?}
+
+Two conclusions can be drawn.
+First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
+This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
+The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
+from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
+
+The second conclusion that can be drawn looks at the differences between the two datasets.
+Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
+Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
+There are 48 possible combinations of attributes.
+Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
+This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
+The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
+Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
+
+\cmtDK{look at language}
+\cmtDK{test different sizes}
+\cmtDK{similarity to bounding box classifier}
+
+\subsubsection{Caption generators}
+
+\subsubsection{Coordinate predictors}

### 2023-07-25 source/mlt/language_games/data_readers.py
captioning game

@@ -1,4 +1,3 @@
-import itertools
 import json
 import os
 import pickle
@@ -11,10 +10,11 @@ from typing import Iterator
 import h5py
 import torch
 from mlt.image_loader import ImageLoader
-from mlt.preexperiments.data_readers import CaptionGeneratorDataset
-from PIL import Image
+from mlt.preexperiments.data_readers import (
+    CaptionGeneratorDataset,
+    CaptionGeneratorSample,
+)
 from torch.utils.data import DataLoader, Dataset, Subset
-from torchvision.models import ResNet101_Weights
 
 
 @dataclass
@@ -30,7 +30,7 @@ class ReferentialGameSample:
 
 class GameBatchIterator(Iterator, ABC):
     @abstractmethod
-    def __init__(self, loader, batch_size, n_batches, seed):
+    def __init__(self, loader, batch_size, n_batches, train_mode, seed):
         ...
 
 
@@ -40,6 +40,7 @@ class GameLoader(DataLoader):
         iterator: GameBatchIterator,
         batch_size,
         batches_per_epoch,
+        train_mode,
         *args,
         seed=None,
         **kwargs,
@@ -47,6 +48,7 @@ class GameLoader(DataLoader):
         self.iterator = iterator
         self.batches_per_epoch = batches_per_epoch
         self.seed = seed
+        self.train_mode = train_mode
 
         # batch_size is part of standard DataLoader arguments
         kwargs["batch_size"] = batch_size
@@ -61,12 +63,13 @@ class GameLoader(DataLoader):
             self,
             batch_size=self.batch_size,
             n_batches=self.batches_per_epoch,
+            train_mode=self.train_mode,
             seed=seed,
         )
 
 
 class LazaridouReferentialGameDataset(Dataset):
-    def __init__(self, data_root_dir, *args, **kwargs) -> None:
+    def __init__(self, data_root_dir, *_args, **_kwargs) -> None:
         super().__init__()
 
         feature_file_path = os.path.join(
@@ -99,11 +102,12 @@ class LazaridouReferentialGameDataset(Dataset):
 
 
 class LazaridouReferentialGameBatchIterator(GameBatchIterator):
-    def __init__(self, loader, batch_size, n_batches, seed) -> None:
+    def __init__(self, loader, batch_size, n_batches, train_mode, seed) -> None:
         self.loader = loader
         self.batch_size = batch_size
         self.n_batches = n_batches
         self.batches_generated = 0
[…]

### 2023-07-25 source/mlt/language_games/models.py

@@ -1,9 +1,10 @@
 import torch
+from mlt.preexperiments.data_readers import CaptionGeneratorDataset
 from torch import nn
 
 
 class ReferentialGameSender(nn.Module):
-    def __init__(self, hidden_size, embedding_dimension=256) -> None:
+    def __init__(self, hidden_size, *_args, embedding_dimension=256, **_kwargs) -> None:
         super().__init__()
         self.image_encoder = nn.Sequential(
             nn.Flatten(), nn.LazyLinear(embedding_dimension, bias=False)
@@ -23,7 +24,7 @@ class ReferentialGameSender(nn.Module):
 
 
 class ReferentialGameReceiver(nn.Module):
-    def __init__(self, embedding_dimension=256) -> None:
+    def __init__(self, *_args, embedding_dimension=256, **_kwargs) -> None:
         super().__init__()
         self.image_encoder = nn.Sequential(
             nn.Flatten(), nn.LazyLinear(embedding_dimension, bias=False)
@@ -46,51 +47,83 @@ class ReferentialGameReceiver(nn.Module):
         return self.softmax(output)
 
 
-class MaskedDaleSender(nn.Module):
-    def __init__(self, image_encoder, masked_image_encoder) -> None:
+class CaptionGeneratorSender(nn.Module):
+    def __init__(
+        self, image_encoder, masked_image_encoder, hidden_size, *_args, **_kwargs
+    ) -> None:
         super().__init__()
         self.image_encoder = image_encoder
         self.masked_image_encoder = masked_image_encoder
 
-    def forward(self, x, _aux_input):
-        image, masked_image = x
+        self.linear = nn.LazyLinear(hidden_size)
 
-        encoded_image = self.image_encoder(image).unsqueeze(dim=0)
-        encoded_masked_image = self.masked_image_encoder(masked_image).unsqueeze(dim=0)
-        concatenated = torch.cat((encoded_image, encoded_masked_image), dim=2)
+    def forward(self, x, aux_input):
+        image = x
+        masked_image = torch.stack([sample["masked_image"] for sample in aux_input])
 
-        return concatenated
+        encoded_image = self.image_encoder(image)
+        encoded_masked_image = self.masked_image_encoder(masked_image)
+
+        concatenated = torch.cat((encoded_image, encoded_masked_image), dim=1)
+
+        linear = self.linear(concatenated)
+        return linear
 
 
-class DaleReceiver(nn.Module):
[…]

### 2023-07-25 source/mlt/language_games/play.py

@@ -7,17 +7,30 @@ from typing import Callable
 
 import egg.core as core
 import torch.nn.functional as F
+from mlt.feature_extractors import DummyFeatureExtractor
 from mlt.image_loader import FeatureImageLoader, ImageLoader
 from mlt.language_games.callbacks import LogSaver
 from mlt.language_games.data_readers import (
+    CaptionGeneratorGameBatchIterator,
+    CaptionGeneratorGameDataset,
+    DaleReferentialGameBatchIterator,
     DaleReferentialGameDataset,
-    DaleReferentialGameGameBatchIterator,
     GameBatchIterator,
     GameLoader,
     LazaridouReferentialGameBatchIterator,
     LazaridouReferentialGameDataset,
 )
-from mlt.language_games.models import ReferentialGameReceiver, ReferentialGameSender
+from mlt.language_games.models import (
+    CaptionGeneratorReceiver,
+    CaptionGeneratorSender,
+    ReferentialGameReceiver,
+    ReferentialGameSender,
+)
+from mlt.preexperiments.data_readers import (
+    BasicImageMasker,
+    DaleCaptionAttributeEncoder,
+)
+from mlt.preexperiments.models import CaptionDecoder, ImageEncoder
 from torch.nn import Module
 from torch.utils.data import Dataset, random_split
 
@@ -35,6 +48,20 @@ def classification_loss(
     acc = (receiver_output.argmax(dim=1) == labels).detach().float()
     # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
     loss = F.nll_loss(receiver_output, labels, reduction="none")
+    return loss, {"acc": acc}
+
+
+def captioning_loss(
+    _sender_input,
+    _message,
+    _receiver_input,
+    receiver_output,
+    labels,
+    _aux_input,
+):
+    acc = (receiver_output.argmax(dim=1) == labels).detach().float()
+    # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
+    loss = F.cross_entropy(receiver_output, labels)
     return loss, {"acc": acc}
 
 
@@ -70,12 +97,61 @@ models = {
         dataset_args={},
         split_dataset=True,
         image_loader=FeatureImageLoader,
-        iterator=DaleReferentialGameGameBatchIterator,
+        iterator=DaleReferentialGameBatchIterator,
         sender=ReferentialGameSender,
         sender_args={},
         receiver=ReferentialGameReceiver,
[…]

### 2023-07-25 source/mlt/preexperiments/data_readers.py

@@ -9,7 +9,6 @@ from enum import Enum
 import torch
 from mlt.image_loader import ImageLoader
 from mlt.preexperiments.models import FeatureExtractor
-from PIL import Image
 from torch.utils.data import Dataset
 from torchvision import transforms
 from torchvision.models import ResNet101_Weights
@@ -488,8 +487,10 @@ class CaptionGeneratorDataset(Dataset):
         image_loader: ImageLoader,
         max_number_samples,
         captioner: Captioner,
+        *_args,
         image_masker: ImageMasker = None,
         preprocess=ResNet101_Weights.DEFAULT.transforms(),
+        **_kwargs,
     ) -> None:
         super().__init__()
         self.captioner = captioner

### 2023-07-25 source/mlt/language_games/models.py
captioning game

@@ -104,6 +104,8 @@ class CaptionGeneratorReceiver(nn.Module):
             lstm_states = linear, linear
             predicted, lstm_states = self.caption_decoder(captions[:, :-1], lstm_states)
 
+            print(predicted.shape)
+
             return predicted.permute(0, 2, 1)
 
         else:

### 2023-07-25 source/mlt/language_games/play.py

@@ -59,10 +59,8 @@ def captioning_loss(
     labels,
     _aux_input,
 ):
-    acc = (receiver_output.argmax(dim=1) == labels).detach().float()
-    # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
     loss = F.cross_entropy(receiver_output, labels)
-    return loss, {"acc": acc}
+    return loss, {}
 
 
 @dataclass

### 2023-07-25 source/mlt/language_games/play.py
captioning game

@@ -60,7 +60,7 @@ def captioning_loss(
     _aux_input,
 ):
     loss = F.cross_entropy(receiver_output, labels)
-    return loss, {}
+    return loss, {"acc": 0}
 
 
 @dataclass

### 2023-07-25 source/mlt/language_games/data_readers.py
captioning game

@@ -304,24 +304,26 @@ class CaptionGeneratorGameBatchIterator(GameBatchIterator):
         sender_inputs = []
         targets = []
         receiver_inputs = []
-        aux_inputs = []
+        masked_images = []
+        captions = []
+        train_modes = []
 
         for sample in samples:
             sender_inputs.append(sample.image)
             targets.append(sample.caption[1:])
 
             receiver_inputs.append(sample.image)
-            aux_inputs.append(
-                {
-                    "masked_image": sample.masked_image,
-                    "caption": sample.caption,
-                    "train_mode": self.train_mode,
-                }
-            )
+            masked_images.append(sample.masked_image)
+            captions.append(sample.caption)
+            train_modes.append(self.train_mode)
 
         return (
             torch.stack(sender_inputs),
             torch.stack(targets),
             torch.stack(receiver_inputs),
-            aux_inputs,
+            {
+                "masked_image": torch.stack(masked_images),
+                "caption": torch.stack(captions),
+                "train_mode": train_modes,
+            },
         )

### 2023-07-25 source/mlt/language_games/models.py

@@ -59,7 +59,7 @@ class CaptionGeneratorSender(nn.Module):
 
     def forward(self, x, aux_input):
         image = x
-        masked_image = torch.stack([sample["masked_image"] for sample in aux_input])
+        masked_image = aux_input["masked_image"]
 
         encoded_image = self.image_encoder(image)
         encoded_masked_image = self.masked_image_encoder(masked_image)
@@ -89,11 +89,11 @@ class CaptionGeneratorReceiver(nn.Module):
 
     def forward(self, message, x, aux_input):
         image = x
-        masked_image = torch.stack([sample["masked_image"] for sample in aux_input])
-        captions = torch.stack([sample["caption"] for sample in aux_input])
+        masked_image = aux_input["masked_image"]
+        captions = aux_input["caption"]
         device = image.device
 
-        if aux_input[0]["train_mode"]:
+        if aux_input["train_mode"][0]:
             encoded_image = self.image_encoder(image)
             encoded_masked_image = self.masked_image_encoder(masked_image)
 
@@ -103,8 +103,6 @@ class CaptionGeneratorReceiver(nn.Module):
             linear = self.linear(concatenated)
             lstm_states = linear, linear
             predicted, lstm_states = self.caption_decoder(captions[:, :-1], lstm_states)
-
-            print(predicted.shape)
 
             return predicted.permute(0, 2, 1)
 

### 2023-07-25 source/mlt/language_games/data_readers.py
captioning game

@@ -324,6 +324,6 @@ class CaptionGeneratorGameBatchIterator(GameBatchIterator):
             {
                 "masked_image": torch.stack(masked_images),
                 "caption": torch.stack(captions),
-                "train_mode": train_modes,
+                "train_mode": torch.tensor(train_modes),
             },
         )

### 2023-07-25 source/mlt/language_games/models.py
captioning game

@@ -126,4 +126,6 @@ class CaptionGeneratorReceiver(nn.Module):
                 word = torch.max(predicted_word_layer, dim=2).indices
                 predicted.append(predicted_word_layer)
 
+            print(torch.stack(predicted).shape)
+
             return torch.stack(predicted).permute(0, 2, 1)

### 2023-07-25 source/mlt/language_games/models.py
captioning game

@@ -104,6 +104,7 @@ class CaptionGeneratorReceiver(nn.Module):
             lstm_states = linear, linear
             predicted, lstm_states = self.caption_decoder(captions[:, :-1], lstm_states)
 
+            print(predicted)
             return predicted.permute(0, 2, 1)
 
         else:
@@ -124,8 +125,8 @@ class CaptionGeneratorReceiver(nn.Module):
                     word, lstm_states
                 )
                 word = torch.max(predicted_word_layer, dim=2).indices
-                predicted.append(predicted_word_layer)
+                predicted.append(predicted_word_layer.squeeze(dim=1))
 
             print(torch.stack(predicted).shape)
 
-            return torch.stack(predicted).permute(0, 2, 1)
+            return torch.stack(predicted).permute(1, 0, 2)

### 2023-07-25 source/mlt/language_games/models.py
captioning game

@@ -104,9 +104,6 @@ class CaptionGeneratorReceiver(nn.Module):
             lstm_states = linear, linear
             predicted, lstm_states = self.caption_decoder(captions[:, :-1], lstm_states)
 
-            print(predicted)
-            return predicted.permute(0, 2, 1)
-
         else:
             encoded_image = self.image_encoder(image)
             encoded_masked_image = self.masked_image_encoder(masked_image)
@@ -127,6 +124,4 @@ class CaptionGeneratorReceiver(nn.Module):
                 word = torch.max(predicted_word_layer, dim=2).indices
                 predicted.append(predicted_word_layer.squeeze(dim=1))
 
-            print(torch.stack(predicted).shape)
-
-            return torch.stack(predicted).permute(1, 0, 2)
+            return torch.stack(predicted).permute(1, 2, 0)

### 2023-07-25 source/mlt/language_games/models.py
captioning game

@@ -104,6 +104,8 @@ class CaptionGeneratorReceiver(nn.Module):
             lstm_states = linear, linear
             predicted, lstm_states = self.caption_decoder(captions[:, :-1], lstm_states)
 
+            return predicted.permute(0, 2, 1)
+
         else:
             encoded_image = self.image_encoder(image)
             encoded_masked_image = self.masked_image_encoder(masked_image)

### 2023-07-25 source/mlt/language_games/data_readers.py
captioning game

@@ -306,6 +306,7 @@ class CaptionGeneratorGameBatchIterator(GameBatchIterator):
         receiver_inputs = []
         masked_images = []
         captions = []
+        non_target_captions = []
         train_modes = []
 
         for sample in samples:
@@ -315,6 +316,7 @@ class CaptionGeneratorGameBatchIterator(GameBatchIterator):
             receiver_inputs.append(sample.image)
             masked_images.append(sample.masked_image)
             captions.append(sample.caption)
+            non_target_captions.append(sample.non_target_captions)
             train_modes.append(self.train_mode)
 
         return (
@@ -324,6 +326,7 @@ class CaptionGeneratorGameBatchIterator(GameBatchIterator):
             {
                 "masked_image": torch.stack(masked_images),
                 "caption": torch.stack(captions),
+                "non_target_captions": torch.stack(non_target_captions),
                 "train_mode": torch.tensor(train_modes),
             },
         )

### 2023-07-25 source/mlt/language_games/play.py

@@ -6,7 +6,6 @@ from time import gmtime, strftime
 from typing import Callable
 
 import egg.core as core
-import torch.nn.functional as F
 from mlt.feature_extractors import DummyFeatureExtractor
 from mlt.image_loader import FeatureImageLoader, ImageLoader
 from mlt.language_games.callbacks import LogSaver
@@ -33,34 +32,6 @@ from mlt.preexperiments.data_readers import (
 from mlt.preexperiments.models import CaptionDecoder, ImageEncoder
 from torch.nn import Module
 from torch.utils.data import Dataset, random_split
-
-
-def classification_loss(
-    _sender_input,
-    _message,
-    _receiver_input,
-    receiver_output,
-    labels,
-    _aux_input,
-):
-    # in the discriminative case, accuracy is computed by comparing the index with highest score in Receiver output (a distribution of unnormalized
-    # probabilities over target poisitions) and the corresponding label read from input, indicating the ground-truth position of the target
-    acc = (receiver_output.argmax(dim=1) == labels).detach().float()
-    # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
-    loss = F.nll_loss(receiver_output, labels, reduction="none")
-    return loss, {"acc": acc}
-
-
-def captioning_loss(
-    _sender_input,
-    _message,
-    _receiver_input,
-    receiver_output,
-    labels,
-    _aux_input,
-):
-    loss = F.cross_entropy(receiver_output, labels)
-    return loss, {"acc": 0}
 
 
 @dataclass

### 2023-07-25 source/mlt/language_games/test.py

@@ -1,0 +1,67 @@
+from cProfile import label
+
+import torch
+import torch.nn.functional as F
+from torcheval.metrics import BinaryAccuracy, Mean, MulticlassAccuracy
+
+
+def classification_loss(
+    _sender_input,
+    _message,
+    _receiver_input,
+    receiver_output,
+    labels,
+    _aux_input,
+):
+    # in the discriminative case, accuracy is computed by comparing the index with highest score in Receiver output (a distribution of unnormalized
+    # probabilities over target poisitions) and the corresponding label read from input, indicating the ground-truth position of the target
+    acc = (receiver_output.argmax(dim=1) == labels).detach().float()
+    # similarly, the loss computes cross-entropy between the Receiver-produced target-position probability distribution and the labels
+    loss = F.nll_loss(receiver_output, labels, reduction="none")
+    return loss, {"acc": acc}
+
+
+def captioning_loss(
+    _sender_input,
+    _message,
+    _receiver_input,
+    receiver_output,
+    labels,
+    aux_input,
+):
+    device = receiver_output.device
+    accuracy = BinaryAccuracy(device=device)
+    hamming_accuracy = MulticlassAccuracy(device=device)
+    non_target_accuracy = BinaryAccuracy(device=device)
+
+    predicted_captions = receiver_output.max(dim=1).indices
+    for sample, output_sample in zip(
+        aux_input["non_target_captions"], predicted_captions
+    ):
+        for caption in sample:
+            described_non_target_object = torch.tensor(False)
+            if torch.equal(output_sample, caption):
+                described_non_target_object = torch.tensor(True)
+                break
+        non_target_accuracy.update(
+            described_non_target_object.unsqueeze(dim=0),
+            torch.tensor(True).unsqueeze(dim=0),
+        )
+
+    for output_sample, ground_truth_sample in zip(predicted_captions, labels):
+        hit = torch.equal(output_sample, ground_truth_sample)
+
+        accuracy.update(
+            torch.tensor(hit).unsqueeze(dim=0),
+            torch.tensor(True).unsqueeze(dim=0),
+        )
+
+    hamming_accuracy.update(predicted_captions.flatten(), labels.flatten())
+
+    loss = F.cross_entropy(receiver_output, labels)
+
+    return loss, {
+        "accuracy": f"{accuracy.compute():.2f}",
+        "hamming_accuracy": f"{hamming_accuracy.compute():.2f}",
+        "non_target_accuracy": f"{non_target_accuracy.compute():.2f}",
+    }

### 2023-07-25 source/mlt/language_games/play.py
captioning game

@@ -25,6 +25,7 @@ from mlt.language_games.models import (
     ReferentialGameReceiver,
     ReferentialGameSender,
 )
+from mlt.language_games.test import captioning_loss, classification_loss
 from mlt.preexperiments.data_readers import (
     BasicImageMasker,
     DaleCaptionAttributeEncoder,

### 2023-07-25 source/mlt/language_games/test.py
captioning game

@@ -61,7 +61,7 @@ def captioning_loss(
     loss = F.cross_entropy(receiver_output, labels)
 
     return loss, {
-        "accuracy": f"{accuracy.compute():.2f}",
-        "hamming_accuracy": f"{hamming_accuracy.compute():.2f}",
-        "non_target_accuracy": f"{non_target_accuracy.compute():.2f}",
+        "accuracy": accuracy.compute(),
+        "hamming_accuracy": hamming_accuracy.compute(),
+        "non_target_accuracy": non_target_accuracy.compute(),
     }

### 2023-07-25 source/mlt/language_games/test.py
captioning game

@@ -61,7 +61,7 @@ def captioning_loss(
     loss = F.cross_entropy(receiver_output, labels)
 
     return loss, {
-        "accuracy": accuracy.compute(),
-        "hamming_accuracy": hamming_accuracy.compute(),
-        "non_target_accuracy": non_target_accuracy.compute(),
+        "accuracy": accuracy.compute().detach().float(),
+        "hamming_accuracy": hamming_accuracy.compute().detach().float(),
+        "non_target_accuracy": non_target_accuracy.compute().detach().float(),
     }

### 2023-07-25 source/mlt/language_games/test.py
captioning game

@@ -61,7 +61,7 @@ def captioning_loss(
     loss = F.cross_entropy(receiver_output, labels)
 
     return loss, {
-        "accuracy": accuracy.compute().detach().float(),
-        "hamming_accuracy": hamming_accuracy.compute().detach().float(),
-        "non_target_accuracy": non_target_accuracy.compute().detach().float(),
+        "accuracy": accuracy.compute().detach().clone().float(),
+        "hamming_accuracy": hamming_accuracy.compute().detach().clone().float(),
+        "non_target_accuracy": non_target_accuracy.compute().detach().clone().float(),
     }

### 2023-07-25 source/mlt/language_games/test.py
captioning game

@@ -31,7 +31,7 @@ def captioning_loss(
 ):
     device = receiver_output.device
     accuracy = BinaryAccuracy(device=device)
-    hamming_accuracy = MulticlassAccuracy(device=device)
+    word_by_word_accuracy = MulticlassAccuracy(device=device)
     non_target_accuracy = BinaryAccuracy(device=device)
 
     predicted_captions = receiver_output.max(dim=1).indices
@@ -56,12 +56,15 @@ def captioning_loss(
             torch.tensor(True).unsqueeze(dim=0),
         )
 
-    hamming_accuracy.update(predicted_captions.flatten(), labels.flatten())
+    word_by_word_accuracy.update(predicted_captions.flatten(), labels.flatten())
 
     loss = F.cross_entropy(receiver_output, labels)
 
     return loss, {
         "accuracy": accuracy.compute().detach().clone().float(),
-        "hamming_accuracy": hamming_accuracy.compute().detach().clone().float(),
+        "word-by-word_accuracy": word_by_word_accuracy.compute()
+        .detach()
+        .clone()
+        .float(),
         "non_target_accuracy": non_target_accuracy.compute().detach().clone().float(),
     }

### 2023-07-26 thesis/sections/3-exp-setup.tex
captioning game
description language games

@@
 \cmtDK{vocabulary?}
-
+\cmtDK{batch iterator}
 All language games are set up and run in the EGG framework \citep{Kharitonov2019}, developed in PyTorch.
@@
 
+\cmtDK{always ResNet-3}
+
 First, I will discuss \emph{discrimination games}, because they have the simplest setup.
@@
 In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
+
+\cmtDK{look at emerged language: distribution of vocab/ngrams, how do they relate to the images (clustering?)}
 
@@
 \paragraph{Caption generators:}
+In a next step, it is tested if the agents can learn to extract features of the objects together.
+For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
+Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
+The receiver only sees the image without additional information.
+\cmtDK{human bias vs emergent language}
+
+This setup is based on the single neural model described above.
+The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
+Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
+
+The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
+The image encoders are set up as described in \citet{Johnson2017}.
+A feature extractor extracts basic features of the images.
+Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
+A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
+This size was found to have the best results during the experiments.
+The two resulting encoded images are flattened and concatenated.
+A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
+
+The receiver uses the same architecture to encode the original image as the sender.
+The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
+This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
+During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
+When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
+The loss is calculated with cross entropy.
+
+Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
+The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
+
+During the experiments, the following variables are adjusted, and the results are compared:
+(1) the vocabulary size $|V|$,
+(2) the LSTM hidden size of the sender wrapper $h_s$ and
+(3) the LSTM hidden size of the receiver wrapper $h_r$.
+The same metrics as in the pre-experiments are used to evaluate the results.
 
 \paragraph{Coordinate predictors:}
+In the final experiments, it is tried to eliminate as much human knowledge as possible.
+For that reason, the agents are tasked to only predict the location of the target object.
+This task is approached in two steps.
+In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
+The receiver doesn't come in contact with any human knowledge, not as input nor as output.
+This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
+In the second step, the caption is also removed from the sender.
+Instead, a masked image points the sender towards the target object.
+In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
+With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
+
+The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
+The sender encodes the original image the same as in the previous paragraph.
+In the first setup, the description of the target object is encoded, using an LSTM.
+For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
+These embeddings are the input for the LSTM.
+The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
+The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
+For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
+Choosing a size of 10 for both variables proved to give good results in the experiments.
+In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
+The resulting vector is passed to the sender wrapper LSTM, to generate the message.
+
+For the second approach, the masked image is passed only through a feature extractor.
+Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
+These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
+
+The receiver contains of two parts.
+First, the original image is encoded with an image encode of the same setup.
+This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
+The resulting vector is the passed through the second part of the receiver, the predictor.
+This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
+A \emph{ReLU} function is applied in between.
+This setup is trained to combine the important information from both the image and the message of the sender.
+The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
+
+As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
+In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
+
+During the experiments, the effects of the following variables are compared:
+(1) the vocabulary size $|V|$,
+(2) the LSTM hidden size of the sender wrapper $h_s$,
+(3) the LSTM hidden size of the receiver wrapper $h_r$ and
+(4) the embedding size of the sender $e_s$.

### 2023-07-26 source/mlt/language_games/callbacks.py
fix callback

@@ -51,9 +51,10 @@ class ExcludingInteractionSaver(InteractionSaver):
         logs.sender_input = torch.tensor(0)
         logs.receiver_input = torch.tensor(0)
 
-        to_remove = ["masked_image", "caption", "train_mode", "attribute_tensor"]
-        for key in to_remove:
-            if key in logs.aux_input:
-                logs.aux_input[key] = torch.tensor(0)
+        if logs.aux_input is not None:
+            to_remove = ["masked_image", "caption", "train_mode", "attribute_tensor"]
+            for key in to_remove:
+                if key in logs.aux_input:
+                    logs.aux_input[key] = torch.tensor(0)
 
         torch.save(logs, dump_dir / f"interaction_gpu{rank}")

### 2023-07-26 source/mlt/language_games/run_discriminators.sh
experiments

@@ -8,3 +8,14 @@ python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$d
 python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 100 --receiver_hidden 10 --sender_hidden 10 --receiver_embedding 10 --sender_embedding 10 --max_samples 10000 --model discriminator --n_epochs 25 --save
 python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 100 --receiver_hidden 100 --sender_hidden 100 --receiver_embedding 100 --sender_embedding 100 --max_samples 10000 --model discriminator --n_epochs 25 --save
 python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 100 --receiver_hidden 10 --sender_hidden 10 --receiver_embedding 50 --sender_embedding 50 --max_samples 10000 --model discriminator --n_epochs 25 --save
+
+dataset=clevr-images-unambigous-dale
+
+python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 10 --receiver_hidden 10 --sender_hidden 10 --receiver_embedding 10 --sender_embedding 10 --max_samples 10000 --model discriminator --n_epochs 25 --save
[…]

### 2023-07-26 thesis/sections/4-results.tex
captioning game
result tables

@@
     \centering
+    \begin{tabular}{ccccc|cc|cc|cc|cc}
+        \toprule
+              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-4}} & \multicolumn{2}{c}{\textbf{CLEVR color}} & \multicolumn{2}{c}{\textbf{CLEVR unambiguous}}                                                                     \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}\cmidrule(lr){12-13}
+        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Acc.}                            & \textbf{length}                                & \textbf{Acc.} & \textbf{length} & \textbf{Acc.} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {10}    & {10}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {10}  & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {13}  & {10}    & {10}    & {10}    & {10}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {13}  & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {20}  & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {100} & {10}    & {10}    & {10}    & {10}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {100} & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        {100} & {100}   & {100}   & {100}   & {100}   & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
+    \label{tab:results_discriminator}
+\end{table}
+
+\begin{table}[h]
+    \centering
     \begin{tabular}{c|ccccc|c}
@@
 \subsubsection{Caption generators}
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ccc|ccc|ccc}
+        \toprule
+              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
+        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {2}             & {7,1\%}       & {40\%}                & {2}             \\
+        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {1}             & {7,3\%}       & {38,7\%}              & {2}             \\
+        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {2}             & {6,7\%}       & {38,7\%}              & {2}             \\
+        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {2}             & {7,8\%}       & {40\%}                & {2}             \\
+        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {2}             & {6,5\%}       & {37,8\%}              & {2}             \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
+    \label{tab:results_caption_generator_game}
+\end{table}
 
 \subsubsection{Coordinate predictors}
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ccc|ccc|ccc|ccc}
+        \toprule
+              &         &         & \multicolumn{3}{c}{\textbf{Single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
+        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                      & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {10,1}                              & {98,5\%}                            & {2}                                 & {36,5}         & {19,9\%}      & {2}             & {45,7}         & {14,4\%}      & {2}             \\
+        {13}  & {10}    & {10}    & {6}                                 & {99\%}                              & {1}                                 & {38}           & {20,4\%}      & {2}             & {47,3}         & {10,8\%}      & {2}             \\
+        {20}  & {10}    & {10}    & {9,7}                               & {96,7\%}                            & {2}                                 & {37,3}         & {21,2\%}      & {2}             & {47,3}         & {11,3\%}      & {1}             \\
+        {100} & {10}    & {10}    & {7,7}                               & {98,4\%}                            & {2}                                 & {40,4}         & {21,7\%}      & {2}             & {45,4}         & {10,8\%}      & {2}             \\
+        {100} & {100}   & {100}   & {7,5}                               & {96,9\%}                            & {2}                                 & {40,1}         & {17,8\%}      & {2}             & {44,3}         & {11,8\%}      & {1}             \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
+    \label{tab:results_dale_predictor_game}
+\end{table}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cccc|ccc|ccc|ccc}
+        \toprule
+              &         &         &         & \multicolumn{3}{c}{\textbf{Single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
+        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & \textbf{Dist.}                      & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {34,8}         & {24,3\%}      & {1}             & {}             & {\%}          & {}              \\
+        {10}  & {10}    & {10}    & {512}   & {}                                  & {\%}                                & {}                                  & {36,3}         & {19,9\%}      & {1
+        ,7}   & {}      & {\%}    & {}                                                                                                                                                                                                                              \\
+        {13}  & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {36,3}         & {20,2\%}      & {1}             & {}             & {\%}          & {}              \\
+        {20}  & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {37,8}         & {16,1\%}      & {2}             & {}             & {\%}          & {}              \\
+        {100} & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {37,4}         & {20,1\%}      & {2}             & {}             & {\%}          & {}              \\
+        {100} & {100}   & {100}   & {1024}  & {}                                  & {\%}                                & {}                                  & {36,5}         & {20,7\%}      & {2}             & {}             & {\%}          & {}              \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
+    \label{tab:results_masked_predictor_game}
+\end{table}

### 2023-07-26 source/mlt/language_games/run_masked_predictors.sh

@@ -6,11 +6,3 @@ python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$d
 python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 20 --receiver_hidden 10 --sender_hidden 10 --receiver_embedding 1024 --sender_embedding 1024 --max_samples 10000 --model masked_coordinate_predictor --n_epochs 25 --save
 python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 100 --receiver_hidden 10 --sender_hidden 10 --receiver_embedding 1024 --sender_embedding 1024 --max_samples 10000 --model masked_coordinate_predictor --n_epochs 25 --save
 python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 100 --receiver_hidden 100 --sender_hidden 100 --receiver_embedding 1024 --sender_embedding 1024 --max_samples 10000 --model masked_coordinate_predictor --n_epochs 25 --save
-
-dataset=clevr-images-unambigous-dale
-python source/mlt/language_games/play.py --scene_json_dir /scratch/guskunkdo/"$dataset"/scenes --image_dir /scratch/guskunkdo/"$dataset"/images --feature_file /scratch/guskunkdo/"$dataset"/features/resnet_3_no-avgpool_no-fc.h5 --vocab_size 10 --receiver_hidden 10 --sender_hidden 10 --receiver_embedding 1024 --sender_embedding 1024 --max_samples 10000 --model masked_coordinate_predictor --n_epochs 25 --save
[…]

### 2023-07-26 thesis/sections/4-results.tex
results and discussion

@@
 
+The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
+In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
+The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
+Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
+A similar worse performance can be seen for the 'Dale-5' dataset.
+Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
+The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
+
+When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
+This is only visible for the 'Dale-2' dataset.
+With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
+When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
+However, the increase is relatively small.
+Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
+As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
+
+Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
+Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
+This would need to be confirmed int the future.
+
 \subsubsection{Coordinate predictors}
@@
         $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                      & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {34,8}         & {24,3\%}      & {1}             & {}             & {\%}          & {}              \\
-        {10}  & {10}    & {10}    & {512}   & {}                                  & {\%}                                & {}                                  & {36,3}         & {19,9\%}      & {1
-        ,7}   & {}      & {\%}    & {}                                                                                                                                                                                                                              \\
-        {13}  & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {36,3}         & {20,2\%}      & {1}             & {}             & {\%}          & {}              \\
-        {20}  & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {37,8}         & {16,1\%}      & {2}             & {}             & {\%}          & {}              \\
-        {100} & {10}    & {10}    & {1024}  & {}                                  & {\%}                                & {}                                  & {37,4}         & {20,1\%}      & {2}             & {}             & {\%}          & {}              \\
-        {100} & {100}   & {100}   & {1024}  & {}                                  & {\%}                                & {}                                  & {36,5}         & {20,7\%}      & {2}             & {}             & {\%}          & {}              \\
+        {10}  & {10}    & {10}     & {1024}  & {19,8}                              & {55,1\%}                            & {2}                                 & {34,8}         & {24,3\%}      & {1}             & {44,3}         & {11,8\%}      & {2}             \\
+        {10}  & {10}    & {10}     & {512}   & {9,3}                               & {92\%}                              & {2}                                 & {36,3}         & {19,9\%}      & {1
+        ,7}   & {45,9}  & {12,7\%} & {2}                                                                                                                                                                                                                             \\
+        {13}  & {10}    & {10}     & {1024}  & {7,8}                               & {96,8\%}                            & {2}                                 & {36,3}         & {20,2\%}      & {1}             & {45,4}         & {11,4\%}      & {2}             \\
+        {20}  & {10}    & {10}     & {1024}  & {6,6}                               & {98,3\%}                            & {2}                                 & {37,8}         & {16,1\%}      & {2}             & {45,2}         & {11\%}        & {2}             \\
+        {100} & {10}    & {10}     & {1024}  & {5,2}                               & {98,5\%}                            & {2}                                 & {37,4}         & {20,1\%}      & {2}             & {43,6}         & {16,7\%}      & {2}             \\
+        {100} & {100}   & {100}    & {1024}  & {12,5}                              & {92,1\%}                            & {2}                                 & {36,5}         & {20,7\%}      & {2}             & {44,6}         & {12,7\%}      & {2}             \\
         \bottomrule

### 2023-07-26 thesis/sections/4-results.tex
results and discussion

@@
 \end{table}
+\cmtDK{insert results of new experiments}
 
@@
 Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This would need to be confirmed int the future.
+This would need to be confirmed in future experiments.
+
+An interesting observation can be made about the length of the messages, the sender generates.
+In almost all cases, the average length is 2 symbols.
+Only in one configuration, the sender produces single symbol messages.
+This happens, when the vocabulary size is 13, the same number as possible features the objects can have.
+In this case, the accuracy, as well as the word-by-word accuracy stay the same compared to the configuration with a vocabulary of only 10 symbols.
+
+\cmtDK{look at language more detailed}
 
 \subsubsection{Coordinate predictors}
+In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
+Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
+Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
+In every configuration of the variables, the agents achieve a very high performance.
+The worst average distance across the test dataset is 10 pixels, which still points on an object.
+Also the accuracy, which evaluates how many guesses of the receiver were pointing on an object reflects this fact.
+All configuration achieve an accuracy higher than 96,7\%.
+This aligns also with the results from the single neural models, where the average distance was similarly low.
+In general, this shows that the agents are able to predict coordinates together.
+However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
+When a different sender is used that is receiving the masked image as input, the results indicate that the message has in fact an influence and the receiver learns how to interpret the message.
+The results for this setup are summarized in \ref{tab:results_masked_predictor_game}.
+In three of six experiments, the agents predict much higher average distances.
+In one case, the mean distance lies at 19,8 pixels, while the accuracy is only at 55,1\%.
+Since the receiver is the same in both setups, this observation indicates that a meaningful message is communicated between the agents.
+
 \begin{table}[h]
@@
         \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
+              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
         $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
@@
 
+When the experiments are run, using the 'Dale-2' dataset, the results are much worse.
+For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
+The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
+Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
+Here, the configurations with higher vocabulary sizes fare better, even though the differences are very small and may be due to other factors.
+Interestingly, when comparing the two configurations with a vocabulary size of 100 symbols one can see that the average distance is similar, but the accuracy is significantly higher when the hidden sizes are lower.
+In other words, with lower hidden sizes the agents can make more correct guesses.
+Overall, it can be seen that a higher vocabulary size lets the agents' guesses get in general closer to the target object, but they are less precise on the object itself.
+
+The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
+Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
+The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
+The only metric that stands out is the accuracy of the configuration with a vocabulary size of 10 symbols.
+With 14,4\%, it is 2,6\% points higher than the next best accuracy.
+It is able to predict the center coordinates of the target object the most precise.
+
 \begin{table}[h]
@@
         \toprule
-              &         &          &         & \multicolumn{3}{c}{\textbf{Single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
+              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
         $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
@@
     \label{tab:results_masked_predictor_game}
+
+    Finally looking at the
 \end{table}

### 2023-08-03 paper/paper.tex
paper
Camera-ready version of Semdial paper

@@
   How do artificial agents based on neural networks coordinate on a new language through referential games over 3-d scenes?
-  We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
+  We extend a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
   % complicated
@@
 There is no restriction on the relation between distractors, hence it is possible to have multiple identical distractors in one image.
-Given the ranking of features in the original GRE algorithm, the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{colour} (2) or the \emph{shape}, \emph{colour} and \emph{size} (3).
+Given the ranking of features in the original GRE algorithm, the target object is therefore identifiable either by \emph{shape} (1), \emph{shape} and \emph{colour} (2) or \emph{shape}, \emph{colour} and \emph{size} (3).
 For each image, fixed-size bounding boxes are extracted around the centre-point of each object.
@@
 This is then used as the initial state of the encoder LSTM.
-After, the sequence is created through Gumbel-Softmax relaxation \citep{Jang2016}.
+After, the sequence of labels is generated through Gumbel-Softmax relaxation \citep{Jang2016}.
 %
 The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-After each step of the LSTM, the receiver calculates the dot product between the hidden state and all of its image encodings.
-The receiver then `points' to one of the images by applying the softmax function over the results of the dot products.
+The sequence, received from the sender is the input for its decoder LSTM, where a hidden state with a dimension of $h_r$ is randomly initialised.
+After each step of the LSTM, the receiver calculates a dot product between the hidden state and all of its image encodings.
+The receiver then `points' to one of the images by applying a softmax function over the results of the dot products.
 The loss is calculated using the NLL-loss.
@@
 % The results are evaluated using accuracy if the receiver can identify the target object.
-Table \ref{tab:results} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
+Table \ref{tab:results} shows the accuracy of the models calculated on the basis of the success of communication if the receiver can identify the target object.
 A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
@@
 % The results change, when using the \emph{Date-5} dataset with four distractors.
-Using \emph{Dale-5} with four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
+Using \emph{Dale-5} with four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline at 23\%.
 Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\% which is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
@@
 The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
+Therefore, the models must learn more combinations of symbols in order to attest to this higher level of detail and how to relate them to features in the images.
 % SD 2023-07-10 18:04:24 +0200: It would be interesting to compare then the length of messages between Dale-2 and Dale-5 that becomes used.
@@
 
-In our ongoing work we are investigating deeper the emerged language and the new vocabulary,
-in particular whether
-it uses similar categories as human language and how its words are combined to form complete messages.
-In future work we will also extend the learning to the relations between entities and the features required to capture them.
+In our ongoing work we are extending our analysis of features and agent configurations as well as we are investigating the emerged language and the new vocabulary, whether
+it consists of similar categories as human language and how its words are combined to form complete messages.
+In our future work we will extend the learning to relations between entities which introduce a high level referential underspecification. % and the features required to capture them.
 % Entries for the entire Anthology, followed by custom entries
@@
 \centering
-\frame{\includegraphics[width=\linewidth]{figures/architecture.png}}
+\frame{\includegraphics[width=\linewidth]{figures/schema.png}}
 % \caption{Architectures of the sender and receiver}

### 2023-08-10 thesis/sections/2-background.tex
thesis

@@
 \label{sec:background}
-% subsections are \subsection{title}
-%% subssubsections are \subsubsection{title}
-%% numbering will work automatically
 
 \cmtDK[inline]{7 pages}
+
+\subsection{Language Games}
+% \begin{itemize}
+%     \item why language games?
+%           \begin{itemize}
+%               \item origins of language
+%               \item flexibility
+%               \item efficiency
+%           \end{itemize}
+%     \item setup of language game
+%           \begin{itemize}
+%               \item cooperative vs self-interested
+%               \item asymetry
+%               \item gumbel softmax vs reinforce
+%               \item EGG
+%               \item bottleneck
+%           \end{itemize}
+%     \item how does emerged language look?
+%           \begin{itemize}
+%               \item compositionality
+%               \item generalization
+%               \item efficiency
+%               \item referring expressions
+%               \item discriminative vs description
+%               \item entropy minimization
+%           \end{itemize}
+%     \item why relating objects?
+% \end{itemize}
+
+The center of this thesis evolves around language games between deep neural models.
+Hereby, multiple deep neural agents need to solve a task, by communicating with symbols, which initially are not associated with any meaning.
+By using and interpreting these symbols in there communication, the agents start to give these symbols and their combinations meaning.
+After this process, a new artificial language emerged, with which the agents can communicate with each other.
+
+In section \ref{sec:aims_languages}, I will explain reasons, why research in this field is conducted and how it may help in further research.
+Section \ref{sec:setup-lg} shows in more detail, how the language games are set up and how exactly agents can communicate and the language emerges.
+In section \ref{sec:properties-el}, I will discuss, how the emerged languages can be built up.
+Finally, section \ref{sec:referring} discusses, how the language games will be used in this thesis, to explore how agents can refer to objects in images.
+
+\subsubsection{Aims of emerged languages}
+\label{sec:aims_languages}
+The setup of language games allows for very controlled rules of how agents can behave.
+
+\subsubsection{Setup of language games}
+\label{sec:setup-lg}
+The term of 'language games' was first introduced by \citet{Wittgenstein1953}.
+The author describes language games as uses of language between multiple interlocutors.
+These can be any small parts of conversations, for instance between a teacher and a student teaching a new concept or between two persons, discussing a topic.
+The rules in each of these situations differ and therefore the meanings and semantics of words and sentences may differ from situation to situation.
+An interjection 'Water!' may be a warning, an answer to a question, a request or something else, depending on the context.
+The meanings are bound to the rules of the language games.
+
+This reasoning was taken up, when trying to train artificial entities to produce a language.
+One of the original games is the signaling game, proposed by \citet{Lewis1969}.
+In this game, a sender needs to send a message to a receiver, based on information only available to the sender.
+Based on the message, the receiver proposes an action.
+Both sender and receiver are rewarded in the same way if the proposed action was correct.
+Hence, both agents need to invent a language together, fit to the conditions of the game.
+
+The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge (\citet{Kirby2002}).
+Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
+Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+
+% This reasoning was taken up, when trying to train deep neural models to produce language.
+% These models need to be set into an interactive situation, a language game, that meaning and a language can emerge (\citet{Baroni2020}).
+% Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
+% Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+
+\subsubsection{Properties of the emerged language}
+\label{sec:properties-el}
+
+\subsubsection{Referring to objects and grounding}
+\label{sec:referring}
 
@@
     ---- 3D objects \\
-    ---- simple objects
-}
-\subsection{Language Games}
-\cmtDK[inline]{\\
-    - EGG \\
-    - Resnet \\
+    ---- simple objects \\
     - CLEVR baselines \\
 }
+
+\subsection{Computational methods}

### 2023-08-10 thesis/sections/3-exp-setup.tex
fix callback
thesis

@@
     \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_single.png}
+        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
         \label{fig:clevr-single}
     }
-    \subfigure['CLEVR unambiguous', \emph{large green sphere}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_unambiguous.png}
-        \label{fig:clevr-unambiguous}
-    }
     \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_color.png}
+        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
         \label{fig:clevr-color}
@@
     \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_dale-2.png}
+        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
         \label{fig:clevr-dale-2}
@@
     \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_dale-5.png}
+        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
         \label{fig:clevr-dale-5}
@@
 Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR unambiguous}
-The second dataset that is created is called 'CLEVR unambiguous'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-The target object doesn't share any attributes with the distractors.
-Namely, if the target is for instance green, no other object in the scene is green as well.
-This also applies to the other attributes.
-When referring to this target object, it is therefore enough to only describe one of the attributes.
-
-Consider the scene in Figure \ref{fig:clevr-unambiguous}.
-The target object can be completely described as the \emph{large green sphere}.
-On the other hand, also descriptions like the \emph{green object}, the \emph{large object} or the \emph{sphere} would be unambiguously referring to the target object.
 
@@
 The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'CLEVR unambiguous' and 'CLEVR color' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the number of shared attributes.
-Does it help the model if the target object is unique in every attribute or does it help if there is only one specific attribute that is shared?
-A similar assumption can be made for the 'Dale' datasets.
-How big is the effect of increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
+The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
+Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.8\linewidth]{figures/discriminator.png}
+    \caption{Sender and receiver architectures in the discrimination game}
+    \label{fig:discriminator_architecture}
+\end{figure}
+
+Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
 For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
@@
 (4) the embedding size of the sender $e_s$.
+As before, the metrics to evaluate the results are the same as in the pre-experiments.

### 2023-08-10 thesis/sections/4-results.tex
thesis

@@
 First, a big difference can be seen between the datasets.
-The model doesn't converge on the correct coordinates for the Dale-5 dataset at all, looking at the test data.
-All resulting mean distances are in a similar range around 50 pixels.
+The model only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
+In most cases, it stays above 50 pixels.
 Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and optionally additional average pooling and classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, using less residual layers, the model doesn't converge at all and the mean distances jump up and down between the epochs.
+All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
+Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
 This effect also applies when using less residual blocks.
-Using the VGG, both using VGG-avg and VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-\cmtDK{compare to scratch}
+Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
 
@@
                             & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,26                               & 46,2                                & \textbf{10,81} & 48,33          \\
-        \textbf{VGG-avg}    & \textbf{9,96}                       & 36,9                                & 32,17          & 54,40          \\
+        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
+        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
         \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 34,95                               & \textbf{36,32}                      & 47,63          & \textbf{46,38} \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & 46,77          \\\midrule
-        \textbf{ResNet-3}   & 92,07                               & 91,12                               & 11,52          & 48,52          \\
-        \textbf{ResNet-4}   & \textbf{8,09}                       & \textbf{36,78}                      & \textbf{9,05}  & 46,84          \\
-        \textbf{ResNet-avg} & 29,21                               & 38,94                               & 37,10          & 48,71          \\
-        \textbf{ResNet-cls} & 33,29                               & 38,83                               & 41,57          & \textbf{46,57} \\
+        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
+        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
+        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
+        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
+        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
+        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
         \bottomrule
@@
 \end{table}
-\cmtDK{wrong numbers}
+
 Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a very low mean distance, meaning a high precision in their predictions, as they did in the experiment before.
-There are only exceptions when ResNet-4 is used or respectively VGG-avg for Dale-2 and VGG-0 for Dale-5.
-In other words, that means that the models are again not able to generalize, but in specific cases to learn the patterns in the train data by heart.
+In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
+The only exception is ResNet-3 as a feature extractor.
+In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
 This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially high layers, like the classifier layers and in the case of ResNet also the average pooling layers seem to already encode to task specific knowledge that was learned during the pre-training of these deep neural models.
-This information can't be utilized for the tasks in this research and earlier layers with more abstract information need to be used.
-The experiments in the following sections are set up using these lower layers as ResNet-3, ResNet-4, VGG-0 and VGG-avg.
+Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
+Higher layers, with more specific encoded information need to be used for this research.
+The experiments in the following sections are set up using these higher layers.
 
@@
 
-In a next step, information about the attributes is included using the \emph{GRE-algorithm} from \citet{Dale1995}.
-\cmtDK{change}
+In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
+Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
 
@@
 These results allow two conclusions.
-First, the models are biased to predict
+First, the models are biased to predict coordinates in the center of the image.
+The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
+Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
+The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
 
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning in the direction of the target object.
-This indicates that it is still able to
+Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
+This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
+Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
+Concluding, the model is able to extract, where objects are located in the image, but can't decide, which of these objects is the target object.
 
@@
 
-Table \ref{tab:results} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
+Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
 A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
+Four different vocabulary sizes $|V|$ are tested.
+A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
+If the symbols were similarly used, messages would have lengths between one and three symbols.
+Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
+Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
+Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
+
+The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
+The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
+Hereby, hidden sizes of 10 and 100 are tested.
+On the other hand, three different embedding sizes are tested: 10, 50 and 100.
+The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
 
@@
     \centering
-    \begin{tabular}{ccccc|cc|cc|cc|cc}
+    \begin{tabular}{ccccc|cc|cc|cc}
         \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-4}} & \multicolumn{2}{c}{\textbf{CLEVR color}} & \multicolumn{2}{c}{\textbf{CLEVR unambiguous}}                                                                     \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}\cmidrule(lr){12-13}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Acc.}                            & \textbf{length}                                & \textbf{Acc.} & \textbf{length} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {10}  & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {\%}                                & {}                                  & {\%}                                     & {}                                             & {\%}          & {}              & {\%}          & {}              \\
+              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
+        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {1,99}                              & {24,7\%}                                 & {1}             & {17,3\%}          & {2}             \\
+        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {2}                                 & {21,4\%}                                 & {2}             & {17,8\%}          & {1}             \\
+        {13}  & {10}    & {10}    & {10}    & {10}    & {94,9\%}                            & {2}                                 & {24,8\%}                                 & {2}             & {17,1\%}          & {2}             \\
+        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {2}                                 & {21,9\%}                                 & {2}             & {17,9\%}          & {1}             \\
+        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {1}                                 & {23\%}                                   & {2}             & {15,9\%}          & {2}             \\
+        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {2}                                 & {24\%}                                   & {2}             & {18,1\%}          & {2}             \\
+        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {2}                                 & {24,4\%}                                 & {2}             & {15,8\%}          & {2}             \\
+        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {1}                                 & {25,3\%}                                 & {2}             & {15,6\%}          & {1}             \\
         \bottomrule
@@
     \label{tab:results_discriminator}
-\end{table}
-\cmtDK{insert results of new experiments}
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|ccccc|c}
-        \hline
-        \textbf{Dataset} & $h_{s}$ & $e_{s}$ & $h_{r}$ & $e_{r}$ & $|V|$ & \textbf{Acc.} \\
-        \hline
-        Dale-2           & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
-        Dale-2           & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
-        Dale-5           & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
-        Dale-5           & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
-        Dale-5           & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
-        \hline
-    \end{tabular}
-    \caption{Results: $h$ are different hidden sizes, $e$ embedding sizes and $|V|$ vocabulary sizes.}
-    \label{tab:results}
 \end{table}
@@
 This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
+The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
 
 \cmtDK{calculation of loss?}
+\cmtDK{length}
 
@@
 \cmtDK{look at language}
-\cmtDK{test different sizes}
 \cmtDK{similarity to bounding box classifier}
@@
     \label{tab:results_masked_predictor_game}
+\end{table}
 
 Finally looking at the
-\end{table}

### 2023-08-11 thesis/sections/3-methodology.tex
restructure thesis

--- 2023-07-25 thesis/sections/3-exp-setup.tex
@@
-\section{Experimental Setup}
-\label{sec:exp-setup}
+\section{Methodology}
+\label{sec:methodology}
 
-\cmtDK[inline]{15 pages}
-
-In this section I will describe which experiments were conducted to answer the \cmtDK{?}{questions} from the previous sections as well as the setup of these experiments.
-This will be done in two parts.
-The first part contains the creation of the datasets that will be used in the experiments.
-In the second part I will go deeper into discussing each experiment, including the necessity of the experiment, which datasets the model is trained on, as well as the final architecture of the models.
-
-\subsection{Creation of the dataset}
+\subsection{CLEVR dataset}
 The basis for my datasets is the CLEVR dataset \citep{Johnson2016}.
@@
 
-\cmtDK{ambiguity}
-\cmtDK{referring expressions}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the later learning of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene that should be identified and communicated by the agents.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both group can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR unambiguous', \emph{large green sphere}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_unambiguous.png}
-        \label{fig:clevr-unambiguous}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.42\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR unambiguous}
-The second dataset that is created is called 'CLEVR unambiguous'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-The target object doesn't share any attributes with the distractors.
-Namely, if the target is for instance green, no other object in the scene is green as well.
-This also applies to the other attributes.
-When referring to this target object, it is therefore enough to only describe one of the attributes.
-
-Consider the scene in Figure \ref{fig:clevr-unambiguous}.
-The target object can be completely described as the \emph{large green sphere}.
-On the other hand, also descriptions like the \emph{green object}, the \emph{large object} or the \emph{sphere} would be unambiguously referring to the target object.
-
-\subsubsection*{CLEVR color}
-This task could on the other hand be difficult to learn for neural models.
-Although the target object is unique in the scene with resect to the distractors, the description isn't.
-Considering all unordered combinations of the three attributes and a composite referring expression, there are $\binom{3}{3} + \binom{3}{2} + \binom{3}{1} = 7$ possible descriptions of the target object.
-This can propose a difficulty, when the models try to converge towards one representation.
-For this reason, I created the 'CLEVR color' dataset.
-It has a similar setup as the previous dataset, containing between 6 and 9 distractors.
-But instead of the target object sharing no attribute with the distractors, they share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is always different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-The number of possible descriptions is lowered to three, since the color \emph{brown} needs to be part of it.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-Both of the above described datasets are very restrictive on the disambiguating relations between the target object and the rest of the objects.
-Either no attribute or only one attribute is shared.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all, how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets, following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Description of experiments}
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own \cmtDK{?}{space}.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.
-\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give me a very detailed overview, where the models struggle to learn and in which ways they can be improved.
-Mainly, I separated the tasks into language games with two agents (final experiments) and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone (pre-experiments).
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-\subsubsection{Feature extractors}
+\subsection{Feature extractors}
 To extract features from the images, I make use of different deep neural network architectures that are developed for this task.
@@
 
-\subsubsection{Pre-experiments without language games}
-The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge.
-Here, I structured the experiments into three levels.
-In the first level, the model are trained to learn the position of objects in the image and attend to specific regions of the image.
-In the second level, the models are trained to differentiate objects in the scene from each other.
-The last level adds language to the experiment, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
+\subsection{EGG}
 
-\paragraph{Level 1 - Coordinate predictor:}
-
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, humans are using.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-Since also positional information of the objects should be part of the discrimination, the image will not be separated into several bounding boxes around the objects, but instead the whole image is the basis.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target object.
-
-With this simple setup, the model is in theory able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize. \cmtDK{go deeper, why}
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode the target object.
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The dimensions of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to 10 objects, resulting in a vector with a length of 20.
-The padding objects have zeros for both coordinates, which no other real object can have.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}. \cmtDK{why?}
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this over the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight could help then to design another model that makes use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left, while still identifying the object uniquely.
-For our research, we rely on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any other object, since the target object is the only cylinder in the scene.
-
-For the experiments, each every image is captioned, by a description of the target object using the described algorithm.
-To use it in the model, the caption needs to be padded to an equal length, in this case to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut of if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.\cmtDK{can just rely on masked image when predicting coords, more useful for captioning or language game}
-
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it can't evaluate every prediction individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given information about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\paragraph{Level 2 - Bounding box classifier:}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to a number of 10 bounding boxes, the maximum number of objects in an image with a matrix of zeros.
-These bounding boxes are shuffled.
-
-The model passes each of the bounding boxes through a feature extractor.
-The resulting vectors are concatenated, which basically is a concatenation of the extracted features of each object.
-This long vector is passed through a linear layer with 4096 with a subsequent \emph{ReLU} function.
-Afterwards, a linear layer reduces the dimensions to 10, corresponding to the 10 bounding boxes in the input.
-With a final \emph{softmax} function, the model then points to one of the bounding boxes as the target bounding box.
-The loss is calculated using cross entropy.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'CLEVR unambiguous' and 'CLEVR color' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the number of shared attributes.
-Does it help the model if the target object is unique in every attribute or does it help if there is only one specific attribute that is shared?
-A similar assumption can be made for the 'Dale' datasets.
-How big is the effect of increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\paragraph{Level 3 - Caption generator:}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection{Language games}
-The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{Gumbel-Softmax vs REINFORCE}
-\cmtDK{better in theoretical background?}
-\cmtDK{vocabulary?}
-
-All language games are set up and run in the EGG framework \citep{Kharitonov2019}, developed in PyTorch.
-This framework consists of heavy configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
-The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
-Furthermore, runs of games can be saved to analyze the used messages of the agents and how they change over the duration of the learning.
-
-The framework is hereby setup in three levels.
-Part of the lowest level are the \emph{agents} themselves.
-The agents are neural models that need to be developed from scratch and define how, the agents process their input and combine it with the message (in case of the receiver) and what they output.
-The second level are \emph{wrappers} that take care of generating and parsing the message.
-The sender wrapper uses the output of the sender agent, to produce a message.
-The receiver on the other hand parses the message received by the sender and passes the result as an additional input to the receiver agent.
-The third level, the \emph{game} links all described parts together.
-It provides the agents with the input and passes the message from the sender to the receiver.
-Furthermore, it uses the output of the receiver and calculates the loss, which is then the basis for the adaption of the weights for both wrappers and agents.
-
-For the following language games, the sender will always produce a sequence of symbols as a message, which the receiver will parse.
-Gumbel-Softmax relaxation is applied to produce discrete symbols.
-This is done using two LSTMs, an encoder LSTM in the sender wrapper and a decoder LSTM in the receiver wrapper.
-The output of the sender agent is used as the initial hidden state for the encoder LSTM.
-This LSTM is then producing symbols until it generates an end-of-sequence symbol.
-This sequence is then passed to the receiver wrapper with its decoder LSTM.
-Its hidden state is initialized randomly.
-The received message sequence is processed symbol by symbol.
-After each time, a symbol is processed by the LSTM, the resulting new hidden state is passed to the receiver agent as the parsed message.
-The receiver agent is combining it with its representation of the image input and predicting an output.
-The \emph{game} is afterwards calculating a loss for each of these outputs separately.
-These losses are summed up to a total loss that is used to adapt the weights in both agents as well as in both LSTMs.
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into caption generators that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-\paragraph{Discrimination games:}
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2016}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2016}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\paragraph{Caption generators:}
-
-\paragraph{Coordinate predictors:}
+\subsection{Ethical considerations}

### 2023-08-11 thesis/sections/4-exp-setup.tex
restructure thesis

--- 2023-08-10 thesis/sections/3-exp-setup.tex
@@
 \subsection{Creation of the dataset}
-The basis for my datasets is the CLEVR dataset \citep{Johnson2016}.
-This dataset includes 3D-generated images depicting scenes with different kinds of objects.
-Each of these objects has different combinations attributes, such as \emph{shape}, \emph{color}, \emph{size} and \emph{material}.
-The possible values of these attributes are listed in Table \ref{tab:clevr-attributes}.
-Three to ten objects are randomly placed into the scene and assigned with random attributes.
-To enhance realism and reduce ambiguity, objects do not intersect, have a certain distance from each other, and are at least partially visible.
-Figure \ref{fig:clevr-example} shows an example of a generated image in the CLEVR dataset.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{cccc}
-        \toprule
-        \textbf{ shape } & \textbf{ color } & \textbf{ size } & \textbf{ material } \\
-        cube             & gray             & small           & rubber              \\
-        sphere           & red              & large           & metal               \\
-        cylinder         & blue                                                     \\
-                         & green                                                    \\
-                         & brown                                                    \\
-                         & purple                                                   \\
-                         & cyan                                                     \\
-                         & yellow                                                   \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attributes of objects in the CLEVR dataset}
-    \label{tab:clevr-attributes}
-\end{table}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/CLEVR_example.png}
-    \caption{Example of a generated image in the CLEVR dataset}
-    \label{fig:clevr-example}
-\end{figure}
-
-Furthermore, the dataset contains information about each scene.
-This includes all selected attributes for each object as well as the exact position of the centers of all the objects, both 3D-coordinates in the 3D scene and 2D-coordinates in the final rendered image.
-In addition, simple spatial relations (in front of, behind, left, right) between the objects are calculated and stored.
-These are simply based on the 3D-coordinates of the objects in relation to the position of the camera.
-
 \cmtDK{ambiguity}
@@
 
-\subsection{Description of experiments}
+\subsection{Generating and understanding referring expressions}
 Language games are very complex setups for machine learning models.
@@
 
-\subsubsection{Feature extractors}
-To extract features from the images, I make use of different deep neural network architectures that are developed for this task.
-First, I use the \cmtDK{why ResNet/VGG?} VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
-After the convolutional layers, the data is passed first through an average pooling layer which outputs 512x7x7 dimensions.
-Next follow three linear layers with \emph{ReLU} non-linearities in between.
-After flattening the input, these classification layers output 4069, 4069 and 1000 dimensions respectively.
-
-Secondly, I include the ResNet-101 \citep{He2016} that tries to overcome problems of very deep networks, using \cmtDK{needs to be explained deeper?}{residual blocks}.
-There are four blocks that output 256x56x56, 512x28x28, 1024x14x14 and 2048x1x1 dimensions.
-A following average pooling layer outputs 2048x1x1 dimensions as well.
-The final linear layer reduces the flattened data to 1000 dimensions, corresponding to the ImageNet classes.
-
-Both architectures are available pretrained on an image classification task on the ImageNet dataset.
-Since the task in this research is very different from a classification, multiple different adaptions of these architectures are compared.
-\cmtDK{deeper discussion about different layers of these networks}
-Table \ref{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
-In the later chapters, it will be referred to these adaptions using the name in the table.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rlc}
-        \toprule
-                            & \textbf{description}                                            & \textbf{ output dimensions } \\\midrule
-        \textbf{VGG-0}      & contains only the convolutional layers                          & 512x7x7                      \\
-        \textbf{VGG-avg}    & contains an additional average pooling layer                    & 512x7x7                      \\
-        \textbf{VGG-cls1}   & \makecell[cl]{contains an additional one classification layer,                                 \\ including its non-linearity} & 4069                         \\
-        \textbf{VGG-cls2}   & \makecell[cl]{contains another additional classification layer,                                \\ including its non-linearity} & 4069                         \\
-        \textbf{VGG-cls3}   & the original VGG19 architecture                                 & 1000                         \\\midrule
-        \textbf{ResNet-1}   & contains one residual block                                     & 256x56x56                    \\
-        \textbf{ResNet-2}   & contains two residual blocks                                    & 512x28x28                    \\
-        \textbf{ResNet-3}   & contains three residual blocks                                  & 1024x14x14                   \\
-        \textbf{ResNet-4}   & contains four residual blocks                                   & 2048x1x1                     \\
-        \textbf{ResNet-avg} & contains an additional average pooling layer                    & 2048x1x1                     \\
-        \textbf{ResNet-cls} & the original ResNet-101 architecture                            & 1000                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Different adaptions of VGG19 and ResNet-101 used in this research}
-    \label{tab:feature-extractor-archs}
-\end{table}
-
-Furthermore, I experimented with both the pretrained models as well as with the architectures trained from scratch with a random initialization for the weights.
-This reason for this was to test if the success of an experiment was actually making use of the pretrained knowledge incorporated in the models.
-If that was not the case, the agents were likely not using image features, but instead relying on some other underlying patterns to solve the task.
-Basically, this approach works as an indicator to determine the actual success of the agents aside from measures as the accuracy or precision.
-
-\subsubsection{Pre-experiments without language games}
 The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
@@
 
-\paragraph{Level 1 - Coordinate predictor:}
+\subsubsection{Coordinate predictor}
 
@@
 
-\paragraph{Level 2 - Bounding box classifier:}
+\paragraph*{Results:}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
+    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
+    \label{fig:coordinate-predictor_loss}
+\end{figure}
+
+Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
+The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ significantly with other feature extractors.
+As can be seen, the success between the different datasets differ significantly.
+The more objects are present in an image, the worse the model performs.
+The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
+This prediction even though not perfectly on the center point is always on the object.
+Opposed to that, using the Dale-2 dataset with two objects, the mean distance is from the first epoch between 37 and 38 pixels.
+With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
+
+An interesting observation is the difference of the mean distances between training and testing data.
+The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
+This points to the fact that the model is not generalizing the task and learning abstract patterns that can be applied to unseen data, but instead learning the training data by heart.
+That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
+Applying a higher dropout didn't have an impact on the results.
+
+This behavior is indeed not very surprising.
+First, these results show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
+Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
+Secondly, more than one objects present in an image confuse the model, and it is not able to consistently point to one of them.
+This can have multiple reasons, for instance that models lacks the ability to separate objects in the extracted features.
+But even given, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
+The guess is then more or less random.
+This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
+For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
+This task on the other hand seems very difficult to learn.
+In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
+
+When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
+One factor that now has a much higher impact is the feature extractor that is used.
+Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
+The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
+First, a big difference can be seen between the datasets.
+The model only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
+In most cases, it stays above 50 pixels.
+Using the Dale-2 dataset, the behavior is a little different.
+All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
+Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
+This effect also applies when using less residual blocks.
+Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{rcccc}
+        \toprule
+                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
+                            & train                               & test                                & train          & test           \\\midrule
+        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
+        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
+        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
+        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
+        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
+        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
+        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
+        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
+        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
+    \label{tab:feature-extractor-mean-distances}
+\end{table}
+
+Secondly, the training loss now looks also different.
+In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
+The only exception is ResNet-3 as a feature extractor.
+In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
+This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
+Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
+Higher layers, with more specific encoded information need to be used for this research.
+The experiments in the following sections are set up using these higher layers.
+
+Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
+In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
+This hypothesis doesn't hold.
+All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
+The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
+The model is still not able to understand, which object is the target object.
+For that reason, a better representation of the target object is necessary.
+
+In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
+Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
+
+\begin{figure}[h]
+    \centering
+    \subfigure['Dale-2', train split]{
+        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
+        \label{fig:visualizations_dale-2_train}
+    }
+    \subfigure['Dale-2', test split]{
+        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
+        \label{fig:visualizations_dale-2_test}
+    }
+    \subfigure['Dale-5', train split]{
+        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
+        \label{fig:visualizations_dale-5_train}
+    }
+    \subfigure['Dale-5', test split]{
+        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
+        \label{fig:visualizations_dale-5_test}
+    }
+    \caption{Visualization of the models' predictions in the 'Dale' datasets}
+    \label{fig:visualizations_dale}
+\end{figure}
+
+An interesting pattern appears when doing a qualitative analysis of the models' predictions.
+Here, I visualized the predicted coordinates compared to the ground truth coordinates.
+Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
+The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
+As can be seen, the prediction are very precise.
+Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
+Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are places as green circles.
+The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
+As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
+
+The results look very different for the test split.
+As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
+For all the images, the predictions don't lie on any object.
+In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
+These findings align with the mean distance scores, described in the sections before.
+However, it seems that the model's predictions are all towards the center of the image.
+This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
+Again, the green circles form the shape of rhombus.
+In contrast, the predictions in red almost all cluster in the center of the image.
+They form roughly the shape of a smaller rhombus.
+This behavior can be observed for all datasets and architectures of the model.
+Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
+Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
+Also the combined visualization shows the same clustering of predictions in the center of the scene, but pattern of the smaller rhombus is better visible.
+
+\begin{figure}[h]
+    \centering
+    \subfigure['Dale-2', train split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
+        \label{fig:visualizations_dale-2_complete_train}
+    }
+    \subfigure['Dale-2', test split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
+        \label{fig:visualizations_dale-2_complete_test}
+    }
+    \subfigure['Dale-5', train split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
+        \label{fig:visualizations_dale-5_complete_train}
+    }
+    \subfigure['Dale-5', test split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
+        \label{fig:visualizations_dale-5_complete_test}
+    }
+    \caption{Visualization of the models' predictions in the Dale datasets}
+    \label{fig:visualizations_dale_complete}
+\end{figure}
+
+These results allow two conclusions.
+First, the models are biased to predict coordinates in the center of the image.
+The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
+Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
+The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
+
+Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
+This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
+Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
+Concluding, the model is able to extract, where objects are located in the image, but can't decide, which of these objects is the target object.
+
+\subsubsection{Bounding box classifier}
 \cmtDK{better to call it discriminators?}
@@
 To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to a number of 10 bounding boxes, the maximum number of objects in an image with a matrix of zeros.
+As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
+The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
 These bounding boxes are shuffled.
+Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
 
-The model passes each of the bounding boxes through a feature extractor.
-The resulting vectors are concatenated, which basically is a concatenation of the extracted features of each object.
-This long vector is passed through a linear layer with 4096 with a subsequent \emph{ReLU} function.
-Afterwards, a linear layer reduces the dimensions to 10, corresponding to the 10 bounding boxes in the input.
-With a final \emph{softmax} function, the model then points to one of the bounding boxes as the target bounding box.
+The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
+The same is done with the one-hot vector of the target attributes.
+In the next step, the dot product between each embedded image and the embedded attributes are calculated.
+A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
+The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
 The loss is calculated using cross entropy.
+
+Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
+Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
+Again, the model uses a \emph{softmax} function to point to one of these objects.
+Since no information about the target object is included, the guesses are expected to be random.
 
@@
 
-\paragraph{Level 3 - Caption generator:}
+\paragraph*{Results:}
+As expected, all baselines achieve a random accuracy for each of the datasets.
+The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
+This changes however, when information about the target object is included.
+Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{rcc}
+        \toprule
+                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
+        \textbf{Dale-2}      & 50\%                       & 99\%              \\
+        \textbf{Dale-5}      & 22\%                       & 94\%              \\
+        \textbf{CLEVR color} & 12\%                       & 93\%              \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the bounding box classifier after 10 epochs}
+    \label{tab:results_bounding_box_classifier}
+\end{table}
+
+The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
+For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
+First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
+This worsens slightly, when more objects are introduced.
+More objects lead to more needed attributes to discriminate the objects from each other.
+The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
+The model can generate these high results, even though its architecture is very simple.
+In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
+A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
+
+\subsubsection{Caption generator}
 
@@
 
-\subsubsection{Language games}
+\paragraph*{Results:}
+In the next paragraphs, the results of the caption generator will be discussed.
+Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
+ResNet-3 produced the best results, when it was used as a feature extractor.
+The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{rccc}
+        \toprule
+                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
+        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
+        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
+        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
+        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the caption generator, based on ResNet-3}
+    \label{tab:results_caption-generator}
+\end{table}
+
+The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
+This means that the models produce almost always correct descriptions of an object in the image.
+Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
+Features of both objects are therefore extracted perfectly and related to the vocabulary.
+As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
+With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
+This changes, when also the masked image is presented to the model.
+Then, the accuracy for the target object is with 72\% well above chance.
+Still, the wrong descriptions describe almost every time the distractor object.
+This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
+
+When the model is trained on the 'Dale-5' dataset, the results are much worse.
+Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
+With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
+Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
+The accuracy stays at the same value.
+In contrast, the non-target accuracy is increased by 5\% points.
+The model is more likely to identify a wrong object.
+Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
+This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
+For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
+After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
+This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
+
+The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
+When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
+The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
+
+There are two main possible explanations for the big difference between the two datasets.
+First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
+In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
+The second explanation lies in the used GRE-algorithm.
+When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
+For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
+Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
+With four distractors the algorithm is much more likely to produce longer captions.
+These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
+
+\subsection{Language games}
 The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.
@@
 
-\paragraph{Discrimination games:}
+\subsubsection{Discrimination games}
 In a discrimination game, the agents are presented with two or more images, one of these being the target image.
@@
 
-\paragraph{Caption generators:}
+\paragraph*{Results:}
+Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
+A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
+Four different vocabulary sizes $|V|$ are tested.
+A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
+If the symbols were similarly used, messages would have lengths between one and three symbols.
+Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
+Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
+Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
+
+The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
+The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
+Hereby, hidden sizes of 10 and 100 are tested.
+On the other hand, three different embedding sizes are tested: 10, 50 and 100.
+The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ccccc|cc|cc|cc}
+        \toprule
+              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
+        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {1,99}                              & {24,7\%}                                 & {1}             & {17,3\%}          & {2}             \\
+        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {2}                                 & {21,4\%}                                 & {2}             & {17,8\%}          & {1}             \\
+        {13}  & {10}    & {10}    & {10}    & {10}    & {94,9\%}                            & {2}                                 & {24,8\%}                                 & {2}             & {17,1\%}          & {2}             \\
+        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {2}                                 & {21,9\%}                                 & {2}             & {17,9\%}          & {1}             \\
+        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {1}                                 & {23\%}                                   & {2}             & {15,9\%}          & {2}             \\
+        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {2}                                 & {24\%}                                   & {2}             & {18,1\%}          & {2}             \\
+        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {2}                                 & {24,4\%}                                 & {2}             & {15,8\%}          & {2}             \\
+        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {1}                                 & {25,3\%}                                 & {2}             & {15,6\%}          & {1}             \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
+    \label{tab:results_discriminator}
+\end{table}
+
+For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
+A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
+The receiver identifies almost every sample correctly when all sizes are 10.
+When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
+Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
+
+The results change, when using the \emph{Dale-5} dataset with four distractors.
+With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
+Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
+This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
+The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
+
+\cmtDK{calculation of loss?}
+\cmtDK{length}
+
+Two conclusions can be drawn.
+First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
+This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
+The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
+from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
+
+The second conclusion that can be drawn looks at the differences between the two datasets.
+Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
+Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
+There are 48 possible combinations of attributes.
+Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
+This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
+The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
+Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
+
+\cmtDK{look at language}
+\cmtDK{similarity to bounding box classifier}
+
+\subsubsection{Caption generator games}
 In a next step, it is tested if the agents can learn to extract features of the objects together.
@@
 
-\paragraph{Coordinate predictors:}
+\paragraph*{Results:}
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ccc|ccc|ccc}
+        \toprule
+              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
+        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {2}             & {7,1\%}       & {40\%}                & {2}             \\
+        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {1}             & {7,3\%}       & {38,7\%}              & {2}             \\
+        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {2}             & {6,7\%}       & {38,7\%}              & {2}             \\
+        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {2}             & {7,8\%}       & {40\%}                & {2}             \\
+        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {2}             & {6,5\%}       & {37,8\%}              & {2}             \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
+    \label{tab:results_caption_generator_game}
+\end{table}
+
+The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
+In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
+The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
+Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
+A similar worse performance can be seen for the 'Dale-5' dataset.
+Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
+The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
+
+When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
+This is only visible for the 'Dale-2' dataset.
+With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
+When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
+However, the increase is relatively small.
+Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
+As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
+
+Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
+Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
+This would need to be confirmed in future experiments.
+
+An interesting observation can be made about the length of the messages, the sender generates.
+In almost all cases, the average length is 2 symbols.
+Only in one configuration, the sender produces single symbol messages.
+This happens, when the vocabulary size is 13, the same number as possible features the objects can have.
+In this case, the accuracy, as well as the word-by-word accuracy stay the same compared to the configuration with a vocabulary of only 10 symbols.
+
+\cmtDK{look at language more detailed}
+
+\subsubsection{Coordinate predictor games}
 In the final experiments, it is tried to eliminate as much human knowledge as possible.
@@
 As before, the metrics to evaluate the results are the same as in the pre-experiments.
+
+\paragraph*{Results:}
+In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
+Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
+Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
+In every configuration of the variables, the agents achieve a very high performance.
+The worst average distance across the test dataset is 10 pixels, which still points on an object.
+Also the accuracy, which evaluates how many guesses of the receiver were pointing on an object reflects this fact.
+All configuration achieve an accuracy higher than 96,7\%.
+This aligns also with the results from the single neural models, where the average distance was similarly low.
+In general, this shows that the agents are able to predict coordinates together.
+However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
+When a different sender is used that is receiving the masked image as input, the results indicate that the message has in fact an influence and the receiver learns how to interpret the message.
+The results for this setup are summarized in \ref{tab:results_masked_predictor_game}.
+In three of six experiments, the agents predict much higher average distances.
+In one case, the mean distance lies at 19,8 pixels, while the accuracy is only at 55,1\%.
+Since the receiver is the same in both setups, this observation indicates that a meaningful message is communicated between the agents.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ccc|ccc|ccc|ccc}
+        \toprule
+              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
+        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {2}                                 & {36,5}         & {19,9\%}      & {2}             & {45,7}         & {14,4\%}      & {2}             \\
+        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {1}                                 & {38}           & {20,4\%}      & {2}             & {47,3}         & {10,8\%}      & {2}             \\
+        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {2}                                 & {37,3}         & {21,2\%}      & {2}             & {47,3}         & {11,3\%}      & {1}             \\
+        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {2}                                 & {40,4}         & {21,7\%}      & {2}             & {45,4}         & {10,8\%}      & {2}             \\
+        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {2}                                 & {40,1}         & {17,8\%}      & {2}             & {44,3}         & {11,8\%}      & {1}             \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
+    \label{tab:results_dale_predictor_game}
+\end{table}
+
+When the experiments are run, using the 'Dale-2' dataset, the results are much worse.
+For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
+The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
+Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
+Here, the configurations with higher vocabulary sizes fare better, even though the differences are very small and may be due to other factors.
+Interestingly, when comparing the two configurations with a vocabulary size of 100 symbols one can see that the average distance is similar, but the accuracy is significantly higher when the hidden sizes are lower.
+In other words, with lower hidden sizes the agents can make more correct guesses.
+Overall, it can be seen that a higher vocabulary size lets the agents' guesses get in general closer to the target object, but they are less precise on the object itself.
+
+The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
+Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
+The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
+The only metric that stands out is the accuracy of the configuration with a vocabulary size of 10 symbols.
+With 14,4\%, it is 2,6\% points higher than the next best accuracy.
+It is able to predict the center coordinates of the target object the most precise.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cccc|ccc|ccc|ccc}
+        \toprule
+              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
+        $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
+        {10}  & {10}    & {10}     & {1024}  & {19,8}                                    & {55,1\%}                            & {2}                                 & {34,8}         & {24,3\%}      & {1}             & {44,3}         & {11,8\%}      & {2}             \\
+        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {2}                                 & {36,3}         & {19,9\%}      & {1
+        ,7}   & {45,9}  & {12,7\%} & {2}                                                                                                                                                                                                                                   \\
+        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {2}                                 & {36,3}         & {20,2\%}      & {1}             & {45,4}         & {11,4\%}      & {2}             \\
+        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {2}                                 & {37,8}         & {16,1\%}      & {2}             & {45,2}         & {11\%}        & {2}             \\
+        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {2}                                 & {37,4}         & {20,1\%}      & {2}             & {43,6}         & {16,7\%}      & {2}             \\
+        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {2}                                 & {36,5}         & {20,7\%}      & {2}             & {44,6}         & {12,7\%}      & {2}             \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
+    \label{tab:results_masked_predictor_game}
+\end{table}
+
+Finally looking at the

### 2023-08-11 thesis/sections/4-exp-setup.tex
inlucde architectures

@@
 Here, I try out four different ways to encode the target object.
+
+\begin{figure}[h]
+    \centering
+    \subfigure[Model including one-hot encoded attributes and locations]{
+        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
+        \label{fig:coordinate_predictor_architecture}
+    }
+    \subfigure[Model including GRE description]{
+        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor_dale.png}
+        \label{fig:coordinate_predictor_dale_architecture}
+    }
+    \caption{Simplified architecture of the coordinate predictors}
+\end{figure}
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
+    \caption{Simplified architecture of the bounding box classifier}
+    \label{fig:bounding_box_classifier_architecture}
+\end{figure}
+
 The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
+    \caption{Simplified architecture of the masked caption generator}
+    \label{fig:caption_generator_architecture}
+\end{figure}
+
 I compare two different models against each other.
@@
 As in the section above, a masked version of the image is created and passed to the model.
+This is shown in Figure \ref{fig:caption_generator_architecture}.
 The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
@@
     \centering
-    \includegraphics[width=.8\linewidth]{figures/discriminator.png}
+    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
     \caption{Sender and receiver architectures in the discrimination game}
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
+    \caption{Simplified architecture of the caption generator game}
+    \label{fig:caption_generator_game_architecture}
+\end{figure}
+
 The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
@@
 The resulting vector is passed to the sender wrapper LSTM, to generate the message.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
+    \caption{Simplified architecture of the masked coordinate predictor game}
+    \label{fig:coordinate_predictor_game_architecture}
+\end{figure}
 

### 2023-08-19 thesis/sections/2-background.tex
langauge analyis

@@
 \cmtDK[inline]{7 pages}
+
+\subsection{Referring expressions}
+Even though, recent large language models (LLMs) such as BERT (\citet{Devlin2018}) or GPT-3 (\citet{Brown2020}) produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
+In this way, their only way of learning, how to generate language as well as solve these tasks is to find patterns of how words and sentences are used in these large corpora.
+Many researchers criticize this approach, by arguing that these model can't learn meaning just by learning on text.
+In \citet{Bender2020}, the authors argue that meaning is bound to a communicative intent.
+These are the purposes of, why humans are using language.
+This intent is connected to the real world that exists outside of language, in other words it is grounded in the world.
+This grounding is missing, when models only train on abstract textual representations of the world.
+\citet{Bisk2020} argues that a multimodal approach, including for instance perception as well as social context, is needed to learn meaning in a broader context.
+One added modality to ground language is often vision. \cmtDK{cite}
+Hereby, the model needs to learn how to associate linguistic concepts in text corpora with features, extracted from visual input.
+For instance, a model can learn to associate the noun "dog" with an animal seen in an image or associate the action of "jumping over" with the animal being above an object.
+
+In a first step, models can combine linguistic knowledge with visual input, by referring to objects, seen in an image, with so called referring expressions.
+\cmtDK{write more}
+By doing this, the models ground parts of their language in another perception of the real world and get a step closer to learn the actual meaning.
+The learning of referring expressions is split in two fields.
+In the referring expression generation, models learn how to produce referring expressions, when presented with visual input.
+In the referring expression understanding, sometimes referred to as coreference, models are trained on interpreting referring expressions and link them to visual input.
+
+A major challenge in natural language processing is, how machines can
 
@@
 
-\subsection{Dataset}
+\subsection{Artificial dataset}
 \cmtDK[inline]{\\
@@
 
-\subsection{Computational methods}
+\subsection{Research question}

### 2023-08-19 thesis/sections/4-exp-setup.tex
langauge analyis

@@
-\section{Experimental Setup}
+\section{Experimentals}
 \label{sec:exp-setup}
@@
 \subsubsection{Coordinate predictor}
+\subsubsection*{Setup}
 
@@
     \subfigure[Model including GRE description]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor_dale.png}
+        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
         \label{fig:coordinate_predictor_dale_architecture}
@@
 
-\paragraph*{Results:}
+\subsubsection*{Results}
 
@@
 \subsubsection{Bounding box classifier}
+\subsubsection*{Setup}
 \cmtDK{better to call it discriminators?}
@@
 
-\paragraph*{Results:}
+\subsubsection*{Results}
 As expected, all baselines achieve a random accuracy for each of the datasets.
@@
 \subsubsection{Caption generator}
+\subsubsection*{Setup}
 
@@
 
-\paragraph*{Results:}
+\subsubsection*{Results}
 In the next paragraphs, the results of the caption generator will be discussed.
@@
 \subsection{Language games}
+
 The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.
@@
 \subsubsection{Discrimination games}
+\subsubsection*{Setup}
+
 In a discrimination game, the agents are presented with two or more images, one of these being the target image.
@@
 
-\paragraph*{Results:}
+\subsubsection*{Results}
 Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
@@
         $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {1,99}                              & {24,7\%}                                 & {1}             & {17,3\%}          & {2}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {2}                                 & {21,4\%}                                 & {2}             & {17,8\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {94,9\%}                            & {2}                                 & {24,8\%}                                 & {2}             & {17,1\%}          & {2}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {2}                                 & {21,9\%}                                 & {2}             & {17,9\%}          & {1}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {1}                                 & {23\%}                                   & {2}             & {15,9\%}          & {2}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {2}                                 & {24\%}                                   & {2}             & {18,1\%}          & {2}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {2}                                 & {24,4\%}                                 & {2}             & {15,8\%}          & {2}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {1}                                 & {25,3\%}                                 & {2}             & {15,6\%}          & {1}             \\
+        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
+        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
+        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
+        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
+        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
+        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
+        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
+        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
         \bottomrule
@@
 
+\subsubsection*{Analysis of the emerged languages}
+In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of 10, 13 and 100.
+In this section, these three emerged languages is analyzed in more detail.
+This is done in two parts.
+The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
+In the second part, the emerged languages are compared to English.
+More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
+
+\cmtDK{first part}
+
+To compare the emerged language to English, a probing approach is used.
+Probing can be used to analyze and interpret the hidden representations in neural networks.
+Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
+If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
+If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
+In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \cite{Dale1995}.
+Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
+The encoder LSTM encodes the messages of the emerged language.
+The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
+The resulting vector is used as the initial state of the decoder LSTM.
+The decoder LSTM is then trained to produce the English referring expression.
+While decoding, teacher forcing is applied.
+The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
+Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
+
+In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
+Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
+
+The translation model with this setup can learn two characteristics.
+First, it can learn to find correlations between the emerged language and the English referring expressions.
+Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
+For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
+This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
+In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
+For that reason, two reference key numbers are calculated for each of the attribute importance order.
+The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
+By this, the model is trained to learn the patterns in the English referring expressions.
+The resulting loss is the highest possible loss the model can achieve, independent of the input.
+Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
+If not, the loss will be as high as this baseline.
+On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
+This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
+These two references are used, to normalize the results of the actual emergent languages $L_{emergent}$, using the formula $L_{norm} = \frac{L_{emergent}-L_{English}}{L_{baseline} - L_{English}}$.
+A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
+By doing this, all configurations can be compared directly to each other.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{c|cc|cc|cc|cc}
+        \toprule
+        \textbf{Order}          & \textbf{Baseline} & \textbf{English} & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                           \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+                                &                   &                  & \textbf{org.}                  & \textbf{norm.}                 & \textbf{org.}                   & \textbf{norm.}     & \textbf{org.} & \textbf{norm.}     \\\midrule
+        {shape > color > size}  & {0,659}           & {0,0001}         & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}       & {90,4\%}           \\
+        {shape > size > color}  & {0,589}           & {0,0002}         & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}        & {\textbf{79,78\%}} \\
+        {color > shape > size}  & {0,849}           & {0,0}            & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}       & {93,15\%}          \\
+        {color  > size > shape} & {0,836}           & {0,0}            & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}       & {94\%}             \\
+        {size > shape > color}  & {0,532}           & {0,0052}         & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}       & {\textbf{85,61\%}} \\
+        {size > color > shape}  & {0,599}           & {0,0001}         & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}       & {\textbf{89,87\%}} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
+    \label{tab:probing_discrimnator_language}
+\end{table}
+
+Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
+
 \subsubsection{Caption generator games}
+\subsubsection*{Setup}
+
 In a next step, it is tested if the agents can learn to extract features of the objects together.
@@
 
-\paragraph*{Results:}
+\subsubsection*{Results}
 \begin{table}[h]
@@
         $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {2}             & {7,1\%}       & {40\%}                & {2}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {1}             & {7,3\%}       & {38,7\%}              & {2}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {2}             & {6,7\%}       & {38,7\%}              & {2}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {2}             & {7,8\%}       & {40\%}                & {2}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {2}             & {6,5\%}       & {37,8\%}              & {2}             \\
+        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
+        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
+        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
+        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
+        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
         \bottomrule
@@
 \subsubsection{Coordinate predictor games}
+\subsubsection*{Setup}
 In the final experiments, it is tried to eliminate as much human knowledge as possible.
@@
 
-\paragraph*{Results:}
+\subsubsection*{Results}
 In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
@@
         $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {2}                                 & {36,5}         & {19,9\%}      & {2}             & {45,7}         & {14,4\%}      & {2}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {1}                                 & {38}           & {20,4\%}      & {2}             & {47,3}         & {10,8\%}      & {2}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {2}                                 & {37,3}         & {21,2\%}      & {2}             & {47,3}         & {11,3\%}      & {1}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {2}                                 & {40,4}         & {21,7\%}      & {2}             & {45,4}         & {10,8\%}      & {2}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {2}                                 & {40,1}         & {17,8\%}      & {2}             & {44,3}         & {11,8\%}      & {1}             \\
+        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
+        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
+        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
+        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
+        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
         \bottomrule
@@
         $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {19,8}                                    & {55,1\%}                            & {2}                                 & {34,8}         & {24,3\%}      & {1}             & {44,3}         & {11,8\%}      & {2}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {2}                                 & {36,3}         & {19,9\%}      & {1
-        ,7}   & {45,9}  & {12,7\%} & {2}                                                                                                                                                                                                                                   \\
-        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {2}                                 & {36,3}         & {20,2\%}      & {1}             & {45,4}         & {11,4\%}      & {2}             \\
-        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {2}                                 & {37,8}         & {16,1\%}      & {2}             & {45,2}         & {11\%}        & {2}             \\
-        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {2}                                 & {37,4}         & {20,1\%}      & {2}             & {43,6}         & {16,7\%}      & {2}             \\
-        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {2}                                 & {36,5}         & {20,7\%}      & {2}             & {44,6}         & {12,7\%}      & {2}             \\
+        {10}  & {10}    & {10}     & {1024}  & {19,8}                                    & {55,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
+        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
+        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
+        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
+        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
+        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
+        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
         \bottomrule

### 2023-08-19 source/mlt/language_games/messge_analyzer.py

@@ -1,0 +1,436 @@
+import argparse
+import json
+import math
+import os
+import sys
+from dataclasses import dataclass
+from glob import glob
+
+import numpy as np
+import torch
+from egg.core import Interaction
+from mlt.preexperiments.data_readers import Color, Shape, Size
+from torch import nn, optim
+from torch.utils.data import DataLoader, Dataset
+
+
+@dataclass
+class LanguageTranslationSample:
+    image_id: str
+    emerged_message: tuple
+    english_description: tuple
+
+
+class LanguageTranslationDataset(Dataset):
+    dataset_mapping = {
+        "dale-2": ("CLEVR_UNAMBIGOUS-DALE-TWO", "clevr-images-unambigous-dale-two"),
+        "dale-5": ("CLEVR_UNAMBIGOUS-DALE", "clevr-images-unambigous-dale"),
+        "single": ("CLEVR_RANDOM-SINGLE", "clevr-images-random-single"),
+        "colour": ("CLEVR_UNAMBIGOUS-COLOR", "clevr-images-unambigous-colour"),
+    }
+
+    combinations = [
+        ("shape", "color", "size"),
+        ("shape", "size", "color"),
+        ("color", "shape", "size"),
+        ("color", "size", "shape"),
+        ("size", "shape", "color"),
+        ("size", "color", "shape"),
+    ]
+
+    PAD_TOKEN = "<pad>"
+    SOS_TOKEN = "<sos>"
+    EOS_TOKEN = "<eos>"
+
+    # class variable, because vocab is static
+    english_vocab = {
+        word: index
+        for index, word in enumerate(
+            list(
+                [
+                    PAD_TOKEN,
+                    SOS_TOKEN,
+                    EOS_TOKEN,
+                    *[
+                        word.lower()
+                        for word in [*Size.names(), *Color.names(), *Shape.names()]
+                    ],
+                ]
+            )
+        )
+    }
+
+    def __init__(
+        self, emerged_vocab_size, dataset, interaction_path, scenes_root, combination
+    ) -> None:
+        self.emerged_vocab_size = emerged_vocab_size
+        self.samples: list[LanguageTranslationSample] = []
+
+        interaction: Interaction = torch.load(interaction_path)
+
+        for raw_message, image_id in zip(
+            interaction.message, interaction.aux_input["image_id"]
+        ):
+            image_id = self._get_image_name(image_id, dataset)
+            message = self._remove_eos(raw_message.max(dim=1).indices)
[…]

### 2023-08-20 thesis/sections/3-methodology.tex
review

@@
 \subsection{CLEVR dataset}
-The basis for my datasets is the CLEVR dataset \citep{Johnson2016}.
-This dataset includes 3D-generated images depicting scenes with different kinds of objects.
-Each of these objects has different combinations attributes, such as \emph{shape}, \emph{color}, \emph{size} and \emph{material}.
+The basis for my datasets is the Visual Question Answering (VQA) dataset CLEVR \citep{Johnson2016}.
+This dataset is split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
+Many of the existing VQA datasets come with two problems.
+First, they include many biases, such as biases in the base images and biases in the linguistic properties of the questions and answers.
+A relatively high number of images of dogs in a dataset, might for instance bias a classifier model towards classifying dogs most of the time.
+On the other hand, repeating patterns in the questions and answers might also be exploited by a model, without extracting the needed information from the image.
+For these reasons the CLEVR dataset aims to reduce the biases as much as possible in both images and questions and answers.
+Secondly, datasets may come with only a limited amount of annotations and information about the state in an image.
+The CLEVR dataset uses artificially rendered 3D-scenes.
+By doing so, all information about for instance the location of objects or their relations to each other can be saved and later used in training models or analyzing their results.
+
+For this thesis, only the visual part of the dataset will be used and extended.
+The questions and answers won't be used.
+The visual part contains images of 3D-generated scenes depicting different kinds of simple objects.
+Each of these objects is made up of a different combination of attributes, such as \emph{shape}, \emph{color}, \emph{size} and \emph{material}.
 The possible values of these attributes are listed in Table \ref{tab:clevr-attributes}.
-Three to ten objects are randomly placed into the scene and assigned with random attributes.
-To enhance realism and reduce ambiguity, objects do not intersect, have a certain distance from each other, and are at least partially visible.
+Three to ten objects are placed in random locations into the scene and assigned with random attributes.
+To enhance realism and reduce ambiguity, objects are placed in a way that they do not intersect and have a certain distance from each other
+Furthermore, it is made sure that every object is almost completely visible.
+The position of the light and the camera are slightly jittered for each image to add noise reduce recurring patterns.
 Figure \ref{fig:clevr-example} shows an example of a generated image in the CLEVR dataset.
@@
 
+In this thesis, the code for the generation of the images is extended, and multiple new datasets are created.
+The extension is detailed in section \ref{sec:creation-dataset}.
+
 \subsection{Feature extractors}
-To extract features from the images, I make use of different deep neural network architectures that are developed for this task.
-First, I use the \cmtDK{why ResNet/VGG?} VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
+In computer vision tasks, machines need to analyze images and extract information from them.
+To do this, machines often rely on feature extractors.
+Features are important parts or patterns in an image, which can have different levels of abstractions.
+They can for example be low-level features, as geometric information about lines and edges in an image, or also very abstract information about whole objects.
+Traditional approaches involve extracting key points and descriptors from an image and using them to represent the image (\cite{Harris1988}, \cite{Lowe1999}, \cite{Bay2006}).
+More recently, convolutional neural networks (CNN) became popular due to their ability to learn complex features automatically from raw image data.
+These are also used in this thesis as a first layer to extract important information from the image. Hereby, two different architectures are tested.
+
+First, I use the VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
+Using 16-19 convolutional layers with small convolution filters helps the model to solve localization and classification tasks on the training dataset, but also enables is to generalize onto other datasets.
 After the convolutional layers, the data is passed first through an average pooling layer which outputs 512x7x7 dimensions.
@@
 
-Secondly, I include the ResNet-101 \citep{He2016} that tries to overcome problems of very deep networks, using \cmtDK{needs to be explained deeper?}{residual blocks}.
+Secondly, I include the ResNet-101 \citep{He2016}.
+This architecture tries to overcome the degradation of very deep networks, where the accuracy rapidly drops after it gets saturated.
+This is done using residual blocks.
+A residual block consists of two or three convolutional layers and a residual connection, also known as a shortcut connection.
+The residual connection allows the input to be added directly to the output of the block, allowing the network to learn the residual function with respect to the input.
+This approach enables the network to better preserve information from earlier layers and avoid the problem of information loss that can occur in very deep networks.
 There are four blocks that output 256x56x56, 512x28x28, 1024x14x14 and 2048x1x1 dimensions.
@@
 Both architectures are available pretrained on an image classification task on the ImageNet dataset.
-Since the task in this research is very different from a classification, multiple different adaptions of these architectures are compared.
-\cmtDK{deeper discussion about different layers of these networks}
+In this thesis, the implementations and weights available for PyTorch are used.\footnote{\href{https://pytorch.org/hub/pytorch\_vision\_resnet/}{https://pytorch.org/hub/pytorch\_vision\_resnet/}}\textsuperscript{,}\footnote{\href{https://pytorch.org/hub/pytorch\_vision\_vgg/}{https://pytorch.org/hub/pytorch\_vision\_vgg/}}
+Since the task in this research is very different from a classification, it likely learned representations that are not directly transferrable to other tasks.
+For this reason, multiple different adaptions of these architectures are compared.
 Table \ref{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
@@
 
-\subsection{EGG}
+\subsection{EGG framework}
+The goal of this thesis is to run and compare different setups of language games systematically.
+To do this, all experiments rely on the \emph{\textbf{E}mergence of lan\textbf{G}uage in \textbf{G}ames} (EGG) framework \citep{Kharitonov2019} which is implemented in PyTorch.
+This framework consists of a heavily configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
+The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
+Furthermore, runs of games can be saved to analyze the used messages of the agents and how they vary over the duration of the learning.
+
+The framework is hereby setup in three levels.
+Part of the lowest level are the \emph{agents} themselves.
+The agents are neural models that need to be implemented from scratch and define how the agents process their input and in case of the receiver combine it with the message as well as what is their output.
+The second level are \emph{wrappers} that take care of generating and parsing the message.
+The sender wrapper uses the output of the sender agent, to produce a message.
+The receiver on the other hand parses the message received by the sender and passes the result as an additional input to the receiver agent.
+The third level, the \emph{game} links all described parts together.
+It provides the agents with the input and passes the message from the sender to the receiver.
+Furthermore, it uses the output of the receiver and calculates the loss, which is then the basis for the adaption of the weights for both wrappers and agents.
+
+For the language games which are run in this thesis, the sender will always produce a sequence of symbols as a message, which the receiver will parse.
+Gumbel-Softmax relaxation is applied to produce discrete symbols.
+This is done using two LSTMs, an encoder LSTM in the sender wrapper and a decoder LSTM in the receiver wrapper.
+The output of the sender agent is used as the initial hidden state for the encoder LSTM.
+This LSTM is then producing symbols until it generates an end-of-sequence symbol.
+This sequence is then passed to the receiver wrapper with its decoder LSTM.
+Its hidden state is initialized randomly.
+The received message sequence is processed symbol by symbol.
+After each time, a symbol is processed by the LSTM, the resulting new hidden state is passed to the receiver agent as the parsed message.
+The receiver agent is combining it with its representation of the image input and is predicting an output.
+In other words the receiver agent produces as many outputs as symbols are present in the message.
+The \emph{game} is then calculating a loss for each of these outputs separately.
+These losses are summed up to a total loss that is used to adapt the weights in both agents as well as in both LSTMs.
+
+\cmtDK{Gumbel-Softmax vs REINFORCE}
 

### 2023-08-20 thesis/sections/4-exp-setup.tex
review

@@
-\section{Experimentals}
+\section{Experiments}
 \label{sec:exp-setup}
 
-\cmtDK[inline]{15 pages}
-
-In this section I will describe which experiments were conducted to answer the \cmtDK{?}{questions} from the previous sections as well as the setup of these experiments.
-This will be done in two parts.
-The first part contains the creation of the datasets that will be used in the experiments.
-In the second part I will go deeper into discussing each experiment, including the necessity of the experiment, which datasets the model is trained on, as well as the final architecture of the models.
+In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
+Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
+This is done in three parts.
+Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
+In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
+Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
 
 \subsection{Creation of the dataset}
+\label{sec:creation-dataset}
 \cmtDK{ambiguity}
-\cmtDK{referring expressions}
 
@@
 Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the later learning of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
+To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
 The \emph{material} is always the same for all objects in a generated image.
@@
 First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene that should be identified and communicated by the agents.
+The target object is the main object in the scene and the models are trained to identify and communicate between each other.
 All other objects and their relations are based on this target object.
@@
 These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both group can be controlled.
+Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
+The number of the objects in both groups can be controlled.
 
@@
 10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are partially visible from the camera.
+Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
 
@@
 \subsubsection*{CLEVR single}
-The simplest dataset is called 'CLEVR single'.
+The simplest new dataset is called 'CLEVR single'.
 This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
@@
 \subsubsection*{CLEVR color}
-This task could on the other hand be difficult to learn for neural models.
-Although the target object is unique in the scene with resect to the distractors, the description isn't.
-Considering all unordered combinations of the three attributes and a composite referring expression, there are $\binom{3}{3} + \binom{3}{2} + \binom{3}{1} = 7$ possible descriptions of the target object.
-This can propose a difficulty, when the models try to converge towards one representation.
-For this reason, I created the 'CLEVR color' dataset.
-It has a similar setup as the previous dataset, containing between 6 and 9 distractors.
-But instead of the target object sharing no attribute with the distractors, they share exactly two attributes.
+The second dataset that is created is called 'CLEVR color'.
+The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
+For this reason, there exist only two groups in the scene, the target object and distractors.
+The distractor group can contain in between 6 and 9 objects.
+To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
 Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is always different.
+The distractors have always the same size and shape as the target object, but the color is different.
 The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
@@
 As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-The number of possible descriptions is lowered to three, since the color \emph{brown} needs to be part of it.
+By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
+All attributes, apart from the color are not discriminating the target object from the distractors.
 Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
@@
 \subsubsection*{CLEVR Dale datasets}
-Both of the above described datasets are very restrictive on the disambiguating relations between the target object and the rest of the objects.
-Either no attribute or only one attribute is shared.
+The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
+The number and the type of shared attributes are controlled exactly.
 More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all, how objects or things relate to each other.
+In real situations, there is no restriction at all how objects or things relate to each other.
 Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
@@
 
-Two datasets, following these rules are created.
+Two datasets following these rules are created.
 The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
@@
 These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
+More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
 
@@
 \subsection{Generating and understanding referring expressions}
+\label{sec:preexperiments}
+This chapter serves two purposes.
+First, the generated dataset from the precious section is validated.
+For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
+A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
+
+Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
 Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time to solve the overall problem.
+The models need to solve multiple tasks at the same time in order to solve the overall problem.
 For instance, in a simple setup of a game two agents are involved.
@@
 The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own \cmtDK{?}{space}.
+In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
 In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
@@
 And finally it needs to learn how to identify the target object with this information.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.
-\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give me a very detailed overview, where the models struggle to learn and in which ways they can be improved.
-Mainly, I separated the tasks into language games with two agents (final experiments) and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone (pre-experiments).
+There are many points of possible failure to train the agents.
+
+For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
+This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
+Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
 With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
 
-The final objective of this thesis is to find out, how agents can communicate about relations of objects spatially as well as based on their attributes.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge.
+The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
+Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
 Here, I structured the experiments into three levels.
-In the first level, the model are trained to learn the position of objects in the image and attend to specific regions of the image.
-In the second level, the models are trained to differentiate objects in the scene from each other.
-The last level adds language to the experiment, more specifically the models learn to caption and describe objects in the image.
+In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
+In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
+In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
 These combined experiments should lay the basis for how to build up the agents in the language games.
@@
 The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, humans are using.
+Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
 For this reason, the final task of the receiver will be to 'point' to the target object.
-Since also positional information of the objects should be part of the discrimination, the image will not be separated into several bounding boxes around the objects, but instead the whole image is the basis.
 The models are therefore tasked to predict the center coordinates of the target object.
@@
 Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity between.
+Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
 These reduce the dimensions first to 1024 and finally to 2.
@@
 This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target object.
-
-With this simple setup, the model is in theory able to focus on an object in the image.
+By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
+
+With this simple setup, the model is theoretically able to focus on an object in the image.
 The problem arises as soon as multiple objects are present in the image.
 There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize. \cmtDK{go deeper, why}
+Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
 Therefore, the models need to receive more information.
-Here, I try out four different ways to encode the target object.
+Here, I try out four different ways to encode and refer to the target object.
 
@@
 There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The dimensions of these vectors can either be zero or one, depending on the attributes of the target object.
+The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
 These three encodings of the attributes are then concatenated.
@@
 All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to 10 objects, resulting in a vector with a length of 20.
-The padding objects have zeros for both coordinates, which no other real object can have.
+Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
+The padded locations consist of two zeros for both coordinates.
 For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}. \cmtDK{why?}
+Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
 The image is first passed through the feature extractor.
@@
 If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this over the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight could help then to design another model that makes use of these learnings without human language.
+If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
+This insight would help for succeeding models that make use of these learnings without human language.
 
@@
 The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left, while still identifying the object uniquely.
-For our research, we rely on the following order from most important to least important: shape, color, size
+This order defines, which attributes can be left out, while still identifying the object uniquely.
+This research relies on the following order from most important to least important: shape, color, size
 Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
@@
 This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any other object, since the target object is the only cylinder in the scene.
-
-For the experiments, each every image is captioned, by a description of the target object using the described algorithm.
-To use it in the model, the caption needs to be padded to an equal length, in this case to a length of 3, which is the maximum number of attributes that can be used.
+Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
+
+For the experiments, each image is captioned with a description of the target object using the described algorithm.
+To include it in the model, the captions need to be padded to an equal length.
+In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
 For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
@@
 The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut of if the target object is too close to the border of the image.
+This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
 The square is filled in white, while the rest of the image is filled in black.
@@
 While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.\cmtDK{can just rely on masked image when predicting coords, more useful for captioning or language game}
-
+It therefore can only rely on its own extracted visual features when looking at masked images.
 The model is presented with both the original image and the masked image.
@@
 The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it can't evaluate every prediction individually.
+Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
 If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
@@
 In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given information about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
+Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
 'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
@@
 Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ significantly with other feature extractors.
-As can be seen, the success between the different datasets differ significantly.
+The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
+As can be seen, the success between the different datasets are significant.
 The more objects are present in an image, the worse the model performs.
@@
 This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance is from the first epoch between 37 and 38 pixels.
+Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
 With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
@@
 The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task and learning abstract patterns that can be applied to unseen data, but instead learning the training data by heart.
+This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
 That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
@@
 This behavior is indeed not very surprising.
-First, these results show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
+First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
 Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuse the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that models lacks the ability to separate objects in the extracted features.
-But even given, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
+Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
+This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
+But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
 The guess is then more or less random.
@@
 First, a big difference can be seen between the datasets.
-The model only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
+The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
 In most cases, it stays above 50 pixels.
@@
 The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the prediction are very precise.
+As can be seen, the predictions are very precise.
 Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are places as green circles.
+Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
 The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
@@
 Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but pattern of the smaller rhombus is better visible.
+Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
 
@@
 Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't decide, which of these objects is the target object.
+Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
 
@@
 \subsection{Language games}
+\label{sec:language-games}
 
@@
 All weights of the agents are adapted in the same way.
-\cmtDK{Gumbel-Softmax vs REINFORCE}
-\cmtDK{better in theoretical background?}
 \cmtDK{vocabulary?}
-\cmtDK{batch iterator}
-All language games are set up and run in the EGG framework \citep{Kharitonov2019}, developed in PyTorch.
-This framework consists of heavy configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
-The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
-Furthermore, runs of games can be saved to analyze the used messages of the agents and how they change over the duration of the learning.
-
-The framework is hereby setup in three levels.
-Part of the lowest level are the \emph{agents} themselves.
-The agents are neural models that need to be developed from scratch and define how, the agents process their input and combine it with the message (in case of the receiver) and what they output.
-The second level are \emph{wrappers} that take care of generating and parsing the message.
-The sender wrapper uses the output of the sender agent, to produce a message.
-The receiver on the other hand parses the message received by the sender and passes the result as an additional input to the receiver agent.
-The third level, the \emph{game} links all described parts together.
-It provides the agents with the input and passes the message from the sender to the receiver.
-Furthermore, it uses the output of the receiver and calculates the loss, which is then the basis for the adaption of the weights for both wrappers and agents.
-
-For the following language games, the sender will always produce a sequence of symbols as a message, which the receiver will parse.
-Gumbel-Softmax relaxation is applied to produce discrete symbols.
-This is done using two LSTMs, an encoder LSTM in the sender wrapper and a decoder LSTM in the receiver wrapper.
-The output of the sender agent is used as the initial hidden state for the encoder LSTM.
-This LSTM is then producing symbols until it generates an end-of-sequence symbol.
-This sequence is then passed to the receiver wrapper with its decoder LSTM.
-Its hidden state is initialized randomly.
-The received message sequence is processed symbol by symbol.
-After each time, a symbol is processed by the LSTM, the resulting new hidden state is passed to the receiver agent as the parsed message.
-The receiver agent is combining it with its representation of the image input and predicting an output.
-The \emph{game} is afterwards calculating a loss for each of these outputs separately.
-These losses are summed up to a total loss that is used to adapt the weights in both agents as well as in both LSTMs.
 
@@
 
-\cmtDK{first part}
-
-To compare the emerged language to English, a probing approach is used.
+\begin{figure}[h]
+    \centering
+    \subfigure[$|V| = 10$]{
+        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
+        \label{fig:relative_frequencies_v10}
+    }
+    \subfigure[$|V| = 13$]{
+        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
+        \label{fig:relative_frequencies_v13}
+    }
+    \subfigure[$|V| = 100$, (13 most frequent messages)]{
+        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
+        \label{fig:relative_frequencies_v100}
+    }
+    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
+    \label{fig:relative_frequencies_vocabularies}
+\end{figure}
+
+When looking at the emerged vocabulary, a few properties can be seen.
+First, when a message is transferred, it consists of only one symbol.
+In some rare cases, an empty message is communicated.
+The models therefore don't learn to combine symbols to create new meaning, but rather encode everything in separate symbols.
+First, in all three emerged languages, very few symbols are used
+
+To compare the emerged language to English in a quantitative way, a probing approach is used.
 Probing can be used to analyze and interpret the hidden representations in neural networks.
@@
 This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{emergent}$, using the formula $L_{norm} = \frac{L_{emergent}-L_{English}}{L_{baseline} - L_{English}}$.
+These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
 A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
@@
         \toprule
-        \textbf{Order}          & \textbf{Baseline} & \textbf{English} & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                           \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                   &                  & \textbf{org.}                  & \textbf{norm.}                 & \textbf{org.}                   & \textbf{norm.}     & \textbf{org.} & \textbf{norm.}     \\\midrule
+        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
         {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
@@
 Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
+First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
+Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
+This fact concludes that all emerged languages don't rely on the GRE algorithm by \cite{Dale1995}.
+Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
+
+Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
+The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
+With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
+Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
+The attribute \emph{color} seems to be less important.
+This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
 

### 2023-08-21 thesis/sections/1-introduction.tex
introduction, ethical considerations

\section{Introduction}
\label{sec:introduction}

The study of emergent languages, in other words artificial languages created by agents to communicate and solve a goal are studied deeply in the recent years \citep{Kirby2002,Kirby2008,Steels2009,Lazaridou2017,Baroni2020,Baroni2022}.
The motivation behind this is twofold.
First, the study of the emergence of artificial language can help to shed light into why and how natural languages evolved.
This is done in a controlled environment, where researchers can regulate exactly, which information the agents receive, how and what they are able to communicate and which goals they are trained on.
This approach allows a deep study of the constraints and restrictions that are necessary for language and meanings of symbols to evolve.
Insights in this field might give indications, under which circumstances natural language evolves.

Secondly, emergent languages can have benefits over predefined communication protocols.
Currently, machines communicate, by using protocols based on predefined rules of how messages can be built up.
These are developed for specific scenarios and goals and don't allow flexible usages if the goal or the environment, machines act in are changing.
Furthermore, these protocols are often based on readability for humans and include overhead that is not necessary to communicate a specific meaning, but is rather used for transparency.
A language that emerges, rather than being predefined could mitigate these problems.
Agents can for instance learn a compositional language that allows them to combine already learned symbols to create new meaning, when presented with a changed environment and unseen situations \citep{Kharitonov2020,Lazaridou2018,Gupta2020}.
Additionally, emergent languages can be learned to encode meaning in efficient ways \citep{Chaabouni2019,Zaslavsky2018}.

Different studies conclude that meaning of symbols and sentences in language can only be bound to the real world, both in the literal sense, but also for abstract concepts \citep{Harnad1999,Bender2020,Bisk2020}.
They need to be grounded.
One way of grounding meaning can be achieved in multimodal systems, where models need to combine language with another modality for instance visual input.
Hereby, a task can be to learn how to refer to objects or actions in the visual input, using referring expressions.
Both generating referring expressions, based on the visual input and understanding given referring expressions by linking them to the visual input ground the meaning of the used language in vision.
Following these arguments, also artificial emergent languages need to be grounded that the used symbols contain meaning.

\subsection{Research Questions}
In this thesis, a deeper look is taken into how agents can ground their emerged language in vision.
The focus hereby lies on referring expressions and how agents are able to generate and understand them.
Multiple experiments are designed in a way that agents need to use referring expressions in their messages that are communicated.
These messages are analyzed with respect to the visual input to answer the two following questions:
\begin{enumerate}
    \item Are cooperative agents in language games able to generate and understand artificial referring expressions based on visual input?
    \item If a language emerges, how similar are the referring expressions to referring expressions that are used in English?
\end{enumerate}

\subsection{Contribution}
This thesis aims to add three contributions to the field of research.
First, new artificial visual datasets are created, consisting of images, depicting objects and their attribute and spatial relations.
Many existing datasets that are used to study referring expressions that use real images are based on photos taking by humans.
This adds a lot of inherent bias to the dataset, since these photos often focus on similar objects and actions.
Additionally, they require external knowledge about the world and the functions of objects which is not present in the image.
Furthermore, information about the objects and their relations in the image are not present in a structured way.
The datasets that are created in this thesis aim to reduce this bias and provide all information about the scene and objects in the image.

Secondly, the thesis evaluates these datasets, by training separate models to generate and understand referring expressions, describing objects in the images.
By doing this, it is tested if first the models are able to extract useful visual information that doesn't rely on bias and latent patterns in the dataset and secondly shows the impact of different levels of ambiguity in the datasets on the performance of generating and understanding of referring expressions.

Lastly, the thesis brings the separate tasks of generating and understanding of referring expressions together into on single task.
In language games, one agent needs to extract visual information from an image and generate a referring expression that is sent to the second agent.
The second agent on the other hand needs to understand this referring expressions and combine it with its visual input.
Only if both of these subtasks succeed, a new artificial language can emerge and the overall task can be solved.
The emerging language is then analyzed to understand, how the artificial referring expressions are built up and compare to natural language.

\subsection{Scope}
The focus of this thesis is the study of referring expressions.
Referring expressions can hereby be based on attributes of objects, as well as their spatial relations towards other objects.
In the present study, only attribute relations are studied.
Spatial relations add another level of complexity and may be studied in future work.
Furthermore, the emerged language will be interpreted by comparing it to referring expressions in natural language.
This analysis focuses on the emerged meaning of symbols and if they align with natural language.
A deeper study of its compositionality or complexity is out of scope.

### 2023-08-21 thesis/sections/2-background.tex
review
introduction, ethical considerations

@@
 
-\cmtDK[inline]{7 pages}
-
 \subsection{Referring expressions}
-Even though, recent large language models (LLMs) such as BERT (\citet{Devlin2018}) or GPT-3 (\citet{Brown2020}) produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
+Even though, recent large language models (LLMs) such as BERT \citep{Devlin2019} or GPT-3 \citep{Brown2020} produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
 In this way, their only way of learning, how to generate language as well as solve these tasks is to find patterns of how words and sentences are used in these large corpora.
@@
 
-The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge (\citet{Kirby2002}).
+The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
 Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.

### 2023-08-21 thesis/sections/3-methodology.tex
introduction, ethical considerations

@@
 \subsection{CLEVR dataset}
-The basis for my datasets is the Visual Question Answering (VQA) dataset CLEVR \citep{Johnson2016}.
+The basis for my datasets is the Visual Question Answering (VQA) dataset CLEVR \citep{Johnson2017a}.
 This dataset is split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
@@
 They can for example be low-level features, as geometric information about lines and edges in an image, or also very abstract information about whole objects.
-Traditional approaches involve extracting key points and descriptors from an image and using them to represent the image (\cite{Harris1988}, \cite{Lowe1999}, \cite{Bay2006}).
+Traditional approaches involve extracting key points and descriptors from an image and using them to represent the image \citep{Harris1988,Lowe1999,Bay2006}.
 More recently, convolutional neural networks (CNN) became popular due to their ability to learn complex features automatically from raw image data.
@@
 \subsection{Ethical considerations}
+In the field of natural language processing (NLP) ethical issues often play major roles.
+These can be part of the used datasets, the created models and their training as well as the application of the models.
+For datasets, the role data privacy is increasing with the necessity of larger amounts of data \citep{Klymenko2022}.
+Furthermore, datasets often contain biases, based for instance on the authors of the collected natural language texts.
+Often they also contain biases such as overrepresentations and underrepresentations.
+Even though some of these biases are inherent to the data and not necessarily negative, much research is indicating that undetected and unaddressed biases in datasets might lead to negative consequences \citep{Shah2020,Field2021,Bender2021}.
+Training neural models can also lead to environmental issues, as large models need to process huge amounts of data and require a lot of energy \citep{Bender2021}.
+Finally, the application of trained models can create harm.
+This applies for example to easy accessible large language model (LLMs) that can be used to create information hazard \citep{Weidinger2022}.
+
+The research in this thesis tries to reduce these risks.
+Looking at the datasets that are used in this thesis, all data is created artificially and contains therefore no personal information.
+Even further, the aim of the creation of these datasets is to reduce biases.
+It doesn't include any social information, but on the other hand consists only of abstract scenes.
+The choice of which attributes the objects are made up is inherently biased towards human cognition, but doesn't have a social impact.
+The models and agents are therefore trained, by including as few human biases as possible.
+
+Looking at the environmental issues, the models used in this thesis consist of only few trained layers and the training is therefore short and doesn't require much energy.
+Larger models as the feature extractors are already pretrained and add no additional consumption.
+
+Finally, the purpose of this thesis is to analyze the results and the emerged language and draw conclusions, how emerged languages can be grounded better in the environment.
+For that reason, the final models can't be uses in any real world applications and produce potential harm.

### 2023-08-21 thesis/sections/4-exp-setup.tex
introduction, ethical considerations

@@
-\section{Experiments}
+\section{Experimental setup and results}
 \label{sec:exp-setup}
@@
 
-The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.
+The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
 
@@
 Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into caption generators that are set up as a language game.
+In the next step, I will look into \emph{caption generators} that are set up as a language game.
 Here, the sender describes the scene, while the receiver needs to generate a caption.
@@
 
-\cmtDK{look at emerged language: distribution of vocab/ngrams, how do they relate to the images (clustering?)}
+The languages that emerge are analyzed in two ways.
+An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
+In a second step, the language is compared to several referring expressions in natural language.
+By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
 
@@
 
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2016}.
+The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
 The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
@@
 Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2016}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
+As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
 The sender is assumed to learn which of the input images is the target images by this approach.
@@
 
-\cmtDK{look at language}
 \cmtDK{similarity to bounding box classifier}
@@
 If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \cite{Dale1995}.
+In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
 Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
@@
 Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \cite{Dale1995}.
+This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
 Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.

### 2023-08-21 thesis/sections/3-methodology.tex
language analysis

@@
 This framework consists of a heavily configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
-The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols or an easy switch between different loss functions.
+The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, an easy switch between different loss functions or a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols.
 Furthermore, runs of games can be saved to analyze the used messages of the agents and how they vary over the duration of the learning.
 
-The framework is hereby setup in three levels.
+The traditional approach to solve these problems relies on the REINFORCE algorithm \citep{Williams1992}.
+It uses the concept of Monte Carlo sampling to estimate the gradient of the expected cumulative reward with respect to the policy parameters.
+The basic idea is to collect trajectories by following the policy, compute the cumulative rewards for each trajectory, and then update the policy parameters to increase the probabilities of actions that led to higher rewards.
+REINFORCE is easy to implement and understand, but it suffers from high variance in gradient estimation due to the inherent randomness in the environment and the policy.
+
+Gumbel-Softmax is a relaxation of the discrete categorical distribution to a continuous distribution that can be differentiated \citep{Jang2016}.
+It allows for the application of gradient-based optimization methods to discrete optimization problems.
+In the Gumbel-Softmax approach, randomness is introduced using the Gumbel distribution (a continuous distribution) and then the \emph{softmax} operation creates a distribution over discrete actions that can be differentiated.
+Gumbel-Softmax offers stable gradient estimates and can be more efficient than the traditional REINFORCE algorithm, especially when dealing with large action spaces.
+
+The EGG framework is set up in three levels.
 Part of the lowest level are the \emph{agents} themselves.

### 2023-08-21 thesis/sections/4-exp-setup.tex
language analysis

@@
 \subsubsection{Coordinate predictor}
+\label{sec:coordinate_predictor}
 \subsubsection*{Setup}
@@
 \subsubsection{Caption generator}
+\label{sec:caption_generator}
 \subsubsection*{Setup}
@@
 \subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of 10, 13 and 100.
+In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
+They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
 In this section, these three emerged languages is analyzed in more detail.
@@
     \centering
-    \subfigure[$|V| = 10$]{
+    \subfigure[$Lang_{10}$]{
         \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
@@
     }
-    \subfigure[$|V| = 13$]{
+    \subfigure[$Lang_{13}$]{
         \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
@@
     }
-    \subfigure[$|V| = 100$, (13 most frequent messages)]{
+    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
         \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
@@
 
-When looking at the emerged vocabulary, a few properties can be seen.
-First, when a message is transferred, it consists of only one symbol.
-In some rare cases, an empty message is communicated.
-The models therefore don't learn to combine symbols to create new meaning, but rather encode everything in separate symbols.
-First, in all three emerged languages, very few symbols are used
-
-To compare the emerged language to English in a quantitative way, a probing approach is used.
+When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
+Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
+Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
+By this the languages are easier comparable across different runs and vocabulary sizes.
+Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
+The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
+
+The first property is that when a message is transferred, it consists of only one symbol.
+In some rare cases, also an empty message is communicated.
+The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
+Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
+More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
+Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
+This is especially visible for $Lang_{100}$.
+Symbols that are used for 16,5\% of the training sample are not used at all in test split.
+Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
+
+These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
+However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
+The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
+
+\begin{figure}[h]
+    \centering
+    \subfigure[$Lang_{10}$]{
+        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
+        \label{fig:language_analysis_v10}
+    }
+    \subfigure[$Lang_{13}$]{
+        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
+        \label{fig:language_analysis_v13}
+    }
+    \subfigure[$Lang_{100}$]{
+        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
+        \label{fig:language_analysis_v100}
+    }
+    \caption{Relative share of the described target object's attributes for the top three messages}
+
+    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
+    \textbf{Size:} brown: small, blue: large
+    \label{fig:language_analysis_vocabularies}
+\end{figure}
+
+Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
+Hereby, the relative share each value of all three attributes are displayed.
+For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
+In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
+
+The first thing that can be seen is that different symbols convey different values of attributes.
+In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
+There is not much difference for target objects that are cylinders.
+The distribution for colors is almost constant across symbols (1) and (2).
+Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
+This different distribution may also be caused by the much lower absolute usage of symbol (3).
+The different sizes are encoded by all symbols in the same way.
+
+Looking at $Lang_{13}$ the symbol usage differs.
+The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
+Again, the colors look similar with small deviations for brown, green and yellow objects.
+Most striking however, is the difference for the size.
+Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
+
+Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
+The usage for the color is similar to $Lang_{13}$.
+Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
+
+An interesting observation is that symbols are not used for the same attributes across the languages.
+In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
+The same applies to the values of attributes, especially to the shapes.
+Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
+Which shape are encoded, differs from language to language.
+
+Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
+For instance $Lang_{10}$ only encodes the shape in its symbols.
+This would not be enough to distinguish the target object from the distractor consistently.
+Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
+
+To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
 Probing can be used to analyze and interpret the hidden representations in neural networks.
@@
 Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This would need to be confirmed in future experiments.
-
-An interesting observation can be made about the length of the messages, the sender generates.
-In almost all cases, the average length is 2 symbols.
-Only in one configuration, the sender produces single symbol messages.
-This happens, when the vocabulary size is 13, the same number as possible features the objects can have.
-In this case, the accuracy, as well as the word-by-word accuracy stay the same compared to the configuration with a vocabulary of only 10 symbols.
-
-\cmtDK{look at language more detailed}
+This is confirmed, when looking at emerged languages.
+In all the setups, the same message is communicated for all samples, independently of the input image.
+This is also reflected in the length of the messages.
+For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
+
+These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
+This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
+In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
+This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
 
@@
 In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points on an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing on an object reflects this fact.
+The worst average distance across the test dataset is 10 pixels, which still points onto an object.
+Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
 All configuration achieve an accuracy higher than 96,7\%.
@@
 However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-When a different sender is used that is receiving the masked image as input, the results indicate that the message has in fact an influence and the receiver learns how to interpret the message.
-The results for this setup are summarized in \ref{tab:results_masked_predictor_game}.
-In three of six experiments, the agents predict much higher average distances.
-In one case, the mean distance lies at 19,8 pixels, while the accuracy is only at 55,1\%.
-Since the receiver is the same in both setups, this observation indicates that a meaningful message is communicated between the agents.
+Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
+As in the experiment before, all communicated messages contain the same symbol independent of the input image.
 
@@
 
-When the experiments are run, using the 'Dale-2' dataset, the results are much worse.
+When the experiments are run on the 'Dale-2' dataset, the results are much worse.
 For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
@@
 Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare better, even though the differences are very small and may be due to other factors.
-Interestingly, when comparing the two configurations with a vocabulary size of 100 symbols one can see that the average distance is similar, but the accuracy is significantly higher when the hidden sizes are lower.
-In other words, with lower hidden sizes the agents can make more correct guesses.
-Overall, it can be seen that a higher vocabulary size lets the agents' guesses get in general closer to the target object, but they are less precise on the object itself.
+Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
 
@@
 The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-The only metric that stands out is the accuracy of the configuration with a vocabulary size of 10 symbols.
-With 14,4\%, it is 2,6\% points higher than the next best accuracy.
-It is able to predict the center coordinates of the target object the most precise.
 
@@
         $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {19,8}                                    & {55,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
+        {10}  & {10}    & {10}     & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
         {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
@@
 
-Finally looking at the
+Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
+This is easily explainable with the emerged language.
+For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
+Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
+
+When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
+The model was already not able to solve the task without the increased complexity of a language game.
+This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2023-08-21 thesis/sections/4-exp-setup.tex
revisions

@@
 Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
+In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not necessarily important for classifying tasks.
 To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
+Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects, target object and distractors.
 This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
 To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
+As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels to capture the complete object.
+The bounding boxes of all objects are padded to the maximum number of objects present in an image across the whole dataset with a matrix of zeros.
+For the 'Dale-2', this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset
 These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
+Furthermore, to point the model towards the target object, its attributes are encoded as on-hot vectors.
 
@@
 
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
+The model passes each of the bounding boxes through a linear layer to transfer it into an embedding space of 10 dimensions.
 The same is done with the one-hot vector of the target attributes.
@@
 A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
+The model then points to an object, using the \emph{softmax} function over all concatenated dot products.
 The loss is calculated using cross entropy.
@@
 The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
+The model can generate these high results, even with its relatively simple architecture.
 In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
@@
 
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
+As the bounding box classifier, the caption generator acts as a learning step towards understanding, how the model can learn the attributes of objects.
+The same method using the GRE-algorithm from the section above is used to create the descriptions of the target object for each image.
+
 There are some minor additions concerning the padding of the caption.
 As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
 However, there are three different ways how the padding is applied.
@@
 
+The task for the model is to generate these descriptions, based on the complete scene as input.
+Doing this, the task inherently involves learning human knowledge and natural language structure.
 Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
+Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
 
@@
 The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
+During training teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM.
 The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
 The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
+During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
 Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
 
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
+Using this approach, the model doesn't have any information about the target object.
+Therefore, I extend it in a second step with a masked image.
+A masked version of the image is created ass in the section above and passed to the model.
 This is shown in Figure \ref{fig:caption_generator_architecture}.
@@
 Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
+This should point the model towards which object to describe and discriminate from the distractors.
 
@@
 The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
+This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
 However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
@@
 It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
+Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
 For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
@@
 If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
+The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
 To mitigate this, I also include a \textbf{word-by-word accuracy}.
@@
 
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
 \subsubsection*{Results}
 In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
+Table \ref{tab:results_caption-generator} shows the different scores of the models when trained on 'Dale-2' and 'Dale-5'.
 ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
+The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
 
@@
 The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
+This means that the models produce almost always correct descriptions of one of the objects in the image.
+Moreover, these descriptions are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
+Features of both objects are therefore extracted perfectly and associated with the vocabulary.
 As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
@@
 This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
+Then, the accuracy for the target object rises with 72\% well above chance.
 Still, the wrong descriptions describe almost every time the distractor object.
@@
 For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
+When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
 This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
@@
 The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
+When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
+The final peak stayed exactly the same and the effects were therefore not studied deeper.
 
@@
 With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
+These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
 
@@
 This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
+The receiver needs to parse this message and combine it with the same information that the sender was presented with.
 The receiver then makes a prediction, which is compared to the ground truth.
@@
 All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
+
+The vocabulary that the sender can draw from to produce a message is made up of initially arbitrary symbols.
+The meaning of these symbols is created as soon as the sender uses them in one of the message, and the receiver is able to use it to solve the task successfully.
+Over time, specific meanings are reinforced and a language emerges.
+The number of possible words, namely the size of the vocabulary can be varied.
 
@@
 Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
+In the next step, I will describe \emph{caption generators} that are set up as a language game.
 Here, the sender describes the scene, while the receiver needs to generate a caption.

### 2023-08-30 source/mlt/language_games/models.py
continuous training

@@ -4,7 +4,7 @@ from torch import nn
 
 
 class DummyReferentialSender(nn.Module):
-    def __init__(self, hidden_size, *_args, embedding_dimension=256, **_kwargs) -> None:
+    def __init__(self, hidden_size, *_args, **_kwargs) -> None:
         super().__init__()
         self.hidden_size = hidden_size
 

### 2023-08-30 source/mlt/language_games/play.py

@@ -75,7 +75,7 @@ models = {
     "test_discriminator": ModelDefinition(
         dataset=DaleReferentialGameDataset,
         dataset_args={},
-        split_dataset=True,
+        split_dataset=False,
         image_loader=FeatureImageLoader,
         iterator=DaleReferentialGameBatchIterator,
         sender=DummyReferentialSender,
@@ -87,7 +87,7 @@ models = {
     "discriminator": ModelDefinition(
         dataset=DaleReferentialGameDataset,
         dataset_args={},
-        split_dataset=True,
+        split_dataset=False,
         image_loader=FeatureImageLoader,
         iterator=DaleReferentialGameBatchIterator,
         sender=ReferentialGameSender,
@@ -105,7 +105,7 @@ models = {
             ),
             "image_masker": BasicImageMasker(),
         },
-        split_dataset=True,
+        split_dataset=False,
         image_loader=FeatureImageLoader,
         iterator=CaptionGeneratorGameBatchIterator,
         sender=CaptionGeneratorSender,
@@ -140,7 +140,7 @@ models = {
                 reversed_caption=False,
             )
         },
-        split_dataset=True,
+        split_dataset=False,
         image_loader=FeatureImageLoader,
         iterator=CoordinatePredictorGameBatchIterator,
         sender=DaleAttributeCoordinatePredictorSender,
@@ -159,7 +159,7 @@ models = {
         dataset_args={
             "image_masker": BasicImageMasker(),
         },
-        split_dataset=True,
+        split_dataset=False,
         image_loader=FeatureImageLoader,
         iterator=CoordinatePredictorGameBatchIterator,
         sender=MaskedCoordinatePredictorSender,

### 2023-08-30 source/mlt/language_games/models.py
align all models

@@ -1,5 +1,10 @@
 import torch
-from mlt.feature_extractors import FeatureExtractor
+from mlt.shared_models import (
+    BoundingBoxImageEncoder,
+    CoordinateClassifier,
+    ImageEncoder,
+    MaskedImageEncoder,
+)
 from torch import nn
 
 
@@ -15,8 +20,8 @@ class DummyReferentialSender(nn.Module):
 class ReferentialGameSender(nn.Module):
     def __init__(self, hidden_size, *_args, embedding_dimension=256, **_kwargs) -> None:
         super().__init__()
-        self.image_encoder = nn.Sequential(
-            nn.Flatten(), nn.LazyLinear(embedding_dimension, bias=False)
+        self.image_encoder = BoundingBoxImageEncoder(
+            embedding_dimension=embedding_dimension
         )
 
         self.lin = nn.LazyLinear(hidden_size, bias=False)
@@ -35,8 +40,8 @@ class ReferentialGameSender(nn.Module):
 class ReferentialGameReceiver(nn.Module):
     def __init__(self, *_args, embedding_dimension=256, **_kwargs) -> None:
         super().__init__()
-        self.image_encoder = nn.Sequential(
-            nn.Flatten(), nn.LazyLinear(embedding_dimension, bias=False)
+        self.image_encoder = BoundingBoxImageEncoder(
+            embedding_dimension=embedding_dimension
         )
 
         self.linear_message = nn.LazyLinear(embedding_dimension, bias=False)
@@ -56,10 +61,17 @@ class ReferentialGameReceiver(nn.Module):
 
 
 class CaptionGeneratorSender(nn.Module):
-    def __init__(self, image_encoder, hidden_size, *_args, **_kwargs) -> None:
+    def __init__(
+        self,
+        image_encoder: ImageEncoder,
+        masked_image_encoder: MaskedImageEncoder,
+        hidden_size,
+        *_args,
+        **_kwargs
+    ) -> None:
         super().__init__()
         self.image_encoder = image_encoder
-        self.masked_image_encoder = nn.Sequential(nn.Flatten(), nn.LazyLinear(1024))
+        self.masked_image_encoder = masked_image_encoder
 
         self.linear = nn.LazyLinear(hidden_size)
 
@@ -78,7 +90,12 @@ class CaptionGeneratorSender(nn.Module):
 
 class CaptionGeneratorReceiver(nn.Module):
     def __init__(
[…]

### 2023-08-30 source/mlt/language_games/play.py

@@ -40,7 +40,12 @@ from mlt.preexperiments.data_readers import (
     BasicImageMasker,
     DaleCaptionAttributeEncoder,
 )
-from mlt.preexperiments.models import CaptionDecoder, ImageEncoder
+from mlt.preexperiments.models import CaptionDecoder
+from mlt.shared_models import (
+    ClevrImageEncoder,
+    CoordinateClassifier,
+    MaskedImageEncoder,
+)
 from torch.nn import Module
 from torch.utils.data import Dataset, random_split
 
@@ -110,14 +115,17 @@ models = {
         iterator=CaptionGeneratorGameBatchIterator,
         sender=CaptionGeneratorSender,
         sender_args={
-            "image_encoder": ImageEncoder(
+            "image_encoder": ClevrImageEncoder(
                 encoder_out_dim=1024,
                 feature_extractor=DummyFeatureExtractor(),
+            ),
+            "masked_image_encoder": MaskedImageEncoder(
+                encoder_out_dim=1024,
             ),
         },
         receiver=CaptionGeneratorReceiver,
         receiver_args={
-            "image_encoder": ImageEncoder(
+            "image_encoder": ClevrImageEncoder(
                 encoder_out_dim=1024,
                 feature_extractor=DummyFeatureExtractor(),
             ),
@@ -146,11 +154,18 @@ models = {
         sender=DaleAttributeCoordinatePredictorSender,
         sender_args={
             "vocab_size": len(DaleCaptionAttributeEncoder.vocab),
-            "feature_extractor": DummyFeatureExtractor(),
+            "image_encoder": ClevrImageEncoder(
+                encoder_out_dim=1024,
+                feature_extractor=DummyFeatureExtractor(),
+            ),
         },
         receiver=CoordinatePredictorReceiver,
         receiver_args={
-            "feature_extractor": DummyFeatureExtractor(),
+            "image_encoder": ClevrImageEncoder(
+                encoder_out_dim=1024,
+                feature_extractor=DummyFeatureExtractor(),
+            ),
+            "coordinate_classifier": CoordinateClassifier(),
         },
         loss_function=pixel_loss,
     ),
@@ -164,11 +179,21 @@ models = {
         iterator=CoordinatePredictorGameBatchIterator,
         sender=MaskedCoordinatePredictorSender,
         sender_args={
-            "feature_extractor": DummyFeatureExtractor(),
+            "image_encoder": ClevrImageEncoder(
+                encoder_out_dim=1024,
[…]

### 2023-08-30 source/mlt/preexperiments/models.py

@@ -1,8 +1,6 @@
-from functools import reduce
-
 import torch
-from mlt.feature_extractors import FeatureExtractor
-from torch import Tensor, nn
+from mlt.shared_models import CoordinateClassifier, ImageEncoder, MaskedImageEncoder
+from torch import nn
 
 
 class BoundingBoxClassifier(nn.Module):
@@ -45,12 +43,10 @@ class BoundingBoxAttributeClassifier(nn.Module):
      - bounding boxes of objects
     """
 
-    def __init__(self, embedding_dimension) -> None:
+    def __init__(self, embedding_dimension, image_encoder: ImageEncoder) -> None:
         super().__init__()
 
-        self.image_encoder = nn.Sequential(
-            nn.Flatten(), nn.LazyLinear(embedding_dimension)
-        )
+        self.image_encoder = image_encoder
 
         self.linear_attributes = nn.LazyLinear(embedding_dimension)
 
@@ -79,24 +75,21 @@ class CoordinatePredictor(nn.Module):
      - image
     """
 
-    def __init__(self, feature_extractor: FeatureExtractor) -> None:
+    def __init__(
+        self, image_encoder: ImageEncoder, coordinate_classifier: CoordinateClassifier
+    ) -> None:
         super().__init__()
-        self.feature_extractor = feature_extractor
+        self.image_encoder = image_encoder
 
-        self.classifier = nn.Sequential(
-            nn.Flatten(),
-            nn.LazyLinear(1024),
-            nn.ReLU(),
-            nn.Linear(1024, 2),
-        )
+        self.coordinate_classifier = coordinate_classifier
 
     def forward(self, data):
         image, *_ = data
 
-        extracted_features = self.feature_extractor(image)
-        classified = self.classifier(extracted_features)
+        encoded_image = self.image_encoder(image)
+        coordinates = self.coordinate_classifier(encoded_image)
 
-        return classified
+        return coordinates
 
 
 class AttributeCoordinatePredictor(nn.Module):
@@ -111,30 +104,22 @@ class AttributeCoordinatePredictor(nn.Module):
 
     def __init__(
         self,
-        number_colors,
-        number_shapes,
-        number_sizes,
-        feature_extractor: FeatureExtractor,
+        image_encoder: ImageEncoder,
+        coordinate_classifier: CoordinateClassifier,
     ) -> None:
         super().__init__()
-        self.feature_extractor = feature_extractor
[…]

### 2023-08-30 source/mlt/preexperiments/train.py

@@ -3,19 +3,15 @@ from dataclasses import dataclass
 from typing import Callable
 
 import torch
-from mlt.feature_extractors import DummyFeatureExtractor, ResnetFeatureExtractor
+from mlt.feature_extractors import DummyFeatureExtractor
 from mlt.image_loader import ClevrImageLoader, FeatureImageLoader
 from mlt.preexperiments.data_readers import (
     BasicImageMasker,
     BoundingBoxClassifierDataset,
     CaptionGeneratorDataset,
-    Color,
     CoordinatePredictorDataset,
     DaleCaptionAttributeEncoder,
     OneHotAttributeEncoder,
-    PreprocessScratch,
-    Shape,
-    Size,
 )
 from mlt.preexperiments.models import (
     AttributeCoordinatePredictor,
@@ -26,7 +22,6 @@ from mlt.preexperiments.models import (
     CaptionGenerator,
     CoordinatePredictor,
     DaleAttributeCoordinatePredictor,
-    ImageEncoder,
     MaskedCaptionGenerator,
     MaskedCoordinatePredictor,
 )
@@ -42,6 +37,12 @@ from mlt.preexperiments.test import (
     CaptionGeneratorTester,
     CoordinatePredictorTester,
     Tester,
+)
+from mlt.shared_models import (
+    BoundingBoxImageEncoder,
+    ClevrImageEncoder,
+    CoordinateClassifier,
+    MaskedImageEncoder,
 )
 from torch import nn, optim
 from torch.nn import Module
@@ -75,23 +76,12 @@ models = {
         dataset_args={},
         preprocess=ResNet101_Weights.IMAGENET1K_V2.transforms(),
         model=CoordinatePredictor,
-        model_args={"feature_extractor": DummyFeatureExtractor()},
-        loss_function=pixel_loss,
-        tester=CoordinatePredictorTester,
-        output_processor=PixelOutputProcessor,
-        output_processor_args={
-            "output_fields": ("image_id", "x", "y", "target_x", "target_y")
-        },
-    ),
-    "coordinate_predictor_scratch": ModelDefinition(
-        dataset=CoordinatePredictorDataset,
-        dataset_args={},
-        preprocess=PreprocessScratch(250),
-        model=CoordinatePredictor,
         model_args={
-            "feature_extractor": ResnetFeatureExtractor(
-                pretrained=False, fine_tune=True
-            )
+            "image_encoder": ClevrImageEncoder(
[…]

### 2023-08-30 source/mlt/shared_models.py

@@ -1,0 +1,78 @@
+from abc import ABC
+
+import torch
+from mlt.feature_extractors import FeatureExtractor
+from torch import nn
+
+
+class ImageEncoder(ABC, nn.Module):
+    ...
+
+
+class ClevrImageEncoder(ImageEncoder):
+    def __init__(
+        self,
+        encoder_out_dim,
+        feature_extractor: FeatureExtractor,
+    ) -> None:
+        super().__init__()
+        self.process_image = nn.Sequential(
+            feature_extractor,
+            nn.LazyConv2d(128, kernel_size=1, padding=0),
+            nn.ReLU(),
+            nn.LazyConv2d(128, kernel_size=1, padding=0),
+            nn.ReLU(),
+            nn.MaxPool2d(kernel_size=2, stride=2),
+        )
+        self.mean_reduction = nn.LazyLinear(encoder_out_dim)
+
+    def forward(self, image):
+        processed_image = self.process_image(image)
+        flattened = torch.flatten(processed_image, start_dim=2).permute(0, 2, 1)
+        reduced = self.mean_reduction(flattened.mean(dim=1))
+        return reduced
+
+
+class BoundingBoxImageEncoder(ImageEncoder):
+    def __init__(
+        self,
+        embedding_dimension,
+    ) -> None:
+        super().__init__()
+        self.process_image = nn.Sequential(
+            nn.Flatten(), nn.LazyLinear(embedding_dimension, bias=False)
+        )
+
+    def forward(self, bounding_box):
+        return self.process_image(bounding_box)
+
+
+class MaskedImageEncoder(ImageEncoder):
+    def __init__(
+        self,
+        encoder_out_dim,
+    ) -> None:
+        super().__init__()
+
+        self.process_image = nn.Sequential(nn.Flatten(), nn.LazyLinear(encoder_out_dim))
+
+    def forward(self, masked_image):
+        return self.process_image(masked_image)
+
+
+class CoordinateClassifier(nn.Module):
+    def __init__(self) -> None:
+        super().__init__()
+
+        self.classifier = nn.Sequential(
+            nn.LazyLinear(1024),
+            nn.Dropout(0.2),
+            nn.ReLU(),
+            nn.LazyLinear(1024),
+            nn.Dropout(0.2),
+            nn.ReLU(),
+            nn.LazyLinear(2),
+        )
+
+    def forward(self, data):
+        return self.classifier(data)

### 2023-08-30 thesis/sections/2-background.tex
comments, feedback

@@
 Many researchers criticize this approach, by arguing that these model can't learn meaning just by learning on text.
-In \citet{Bender2020}, the authors argue that meaning is bound to a communicative intent.
-These are the purposes of, why humans are using language.
-This intent is connected to the real world that exists outside of language, in other words it is grounded in the world.
+In \citep{Bender2020}, the authors argue that meaning is bound to a communicative intent.
+These are the purposes of why humans are using language.
+The intent is always embedded in a broader context that exists outside the language itself.
+This context includes the real world, but also the background knowledge of the speaker as well as interlocutors or reason for what the speaker is saying.
+By grounding the intent in this context, it becomes meaningful.
+Only with this step, the intent can be interpreted by the listeners.
+% This intent is connected to the real world that exists outside of language, in other words it is grounded in the world.
+% SD: Not just real world, also to background knowledge and what the person wants to do. 
+% SD: Grounding is more related to interpretability of expressions rather than intent.
+Without grounding they cannot be interpreted.
+The intent shapes WHAT referring expressions are generated in what contexts of the real world.
+Hence, intent is a meta thing connecting world, language, knowledge and agent’s experience.
+% DK: first part included, referring expressions are included below (done)
 This grounding is missing, when models only train on abstract textual representations of the world.
@@
 In a first step, models can combine linguistic knowledge with visual input, by referring to objects, seen in an image, with so called referring expressions.
-\cmtDK{write more}
-By doing this, the models ground parts of their language in another perception of the real world and get a step closer to learn the actual meaning.
-The learning of referring expressions is split in two fields.
+Hereby, the communicative intent as well as the broader context in which it is used determines which referring expressions are applied.
+By doing this, the models learn how to relate referring expressions against a representation of the environment.
+More specifically, the model learns to connect symbols to a non-symbolic representation.
+This is known as the symbol grounding problem \citep{Harnad1999,Roy2002}.
+Hereby, the question is asked how arbitrary symbols can be connected to the representation they actually stand for.
+
+To solve this problem, multiple challenges arise since language describes the world in a very under-specified way.
+For instance, sensory input for humans as well as for machines only show a very limited representation of the world.
+Vision needs to map a 3-d world into 2-d images.
+During this process, information about the world gets lost, as for example depth can't be perceived easily or objects might be obfuscated in certain perspectives.
+Following, language needs to be connected to a representation of the world that is already not complete.
+Secondly, referring expressions in language themselves are often under-specified.
+Given for example a table with five cups on top, an interlocutor is asked to retrieve one of the cups, referring to it by 'the red cup'.
+This referring expression on its own might be ambiguous and refer to multiple red cups on the table.
+The challenge now arises as how the interlocutor associates the correct cup with this referring expressions.
+This gets even more complex, when perspective is part of the referring expression as for example in 'the cup on the left' where \emph{left} is dependent on the speaker's and interlocutors' spatial position in relation to the cup \citep{Dobnik2021}.
+
+% By doing this, the models ground parts of their language in another perception of the real world and get a step closer to learn the actual meaning.
+% SD: They learn how to interpret expressions against some representation.
+Hence, a reference is a relation between a description and some entity, it’s interpretation.
+Without this connection referring expressions are useless as we do not know how to interpret this.
+Textual models might be good for general knowledge such as factual information from news and Wikipedia but as soon as we have NLP applications that interact with other domains, e.g.
+a ticket booking system or a situated robot involved in patient care, we need to connect language to some other representations to make language interpretable.
+% DK: done
+The learning of referring expressions is therefore split in two fields.
 In the referring expression generation, models learn how to produce referring expressions, when presented with visual input.
@@
 After this process, a new artificial language emerged, with which the agents can communicate with each other.
+% SD: Give an example of a language games.
+Symbols are invented at random.
+However, the constraints of interaction control how new symbols are introduced and when existing symbols are re-used and used.
+Agents cannot invent unlimited number of symbols to refer to every event the6 encounter as they have limited memory and therefore they are driven by abstraction.
+On the other hand, the symbols must be flexible enough.
+The other agent must be able to resolve the symbols they hear based on the context.
+Hence, a successful interaction is when a describer is producing such symbols that interpreter can easily interpret within the context, here an image.
+This is also how a reward function is defined and loss is propagated.
+% DK: TODO
 
-In section \ref{sec:aims_languages}, I will explain reasons, why research in this field is conducted and how it may help in further research.
-Section \ref{sec:setup-lg} shows in more detail, how the language games are set up and how exactly agents can communicate and the language emerges.
-In section \ref{sec:properties-el}, I will discuss, how the emerged languages can be built up.
-Finally, section \ref{sec:referring} discusses, how the language games will be used in this thesis, to explore how agents can refer to objects in images.
+Section \ref{sec:aims_languages} explains reasons why research in this field is conducted and how it may help in further research.
+Section \ref{sec:setup-lg} shows in more detail how the language games are set up and how exactly agents can communicate and the language emerges.
+Section \ref{sec:properties-el} discusses, how the emerged languages can be built up.
+Finally, section \ref{sec:referring} discusses how the language games will be used in this thesis, to explore how agents can refer to objects in images.
 
@@
 The setup of language games allows for very controlled rules of how agents can behave.
+% SD: Rephrase, language games provide interactive constraints within which language can emerge.
+But what are these?
+In this thesis we will study some, such as agent’s memory and layer representation and the feature representation, the structure of the world and its ambiguity.
+% DK: TODO
 
@@
 An interjection 'Water!' may be a warning, an answer to a question, a request or something else, depending on the context.
-The meanings are bound to the rules of the language games.
+% SD: Because referring expressions are ambiguous participants rely on language games to resolve the reference in context.
+% SD: Give an example of a game?
+% DK: TODO
+meanings are bound to the rules of the language games.
 
@@
 Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+% SD: Examples of architectures from Baroni. The EGG framework.
+% DK: TODO
 
@@
 \subsection{Artificial dataset}
-\cmtDK[inline]{\\
-    - CLEVR \\
-    ---- 3D objects \\
-    ---- simple objects \\
-    - CLEVR baselines \\
-}
-
-\subsection{Research question}

### 2023-08-30 thesis/sections/3-methodology.tex
comments, feedback

@@
-\section{Methodology}
+\section{Methodology and Frameworks}
 \label{sec:methodology}
 
-\subsection{CLEVR dataset}
-The basis for my datasets is the Visual Question Answering (VQA) dataset CLEVR \citep{Johnson2017a}.
-This dataset is split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
+\subsection{CLEVR framework}
+The basis for my datasets is the Visual Question Answering (VQA) framework CLEVR \citep{Johnson2017a}.
+This framework provides code to generate configurable VQA datasets that are split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
 Many of the existing VQA datasets come with two problems.
@@
 On the other hand, repeating patterns in the questions and answers might also be exploited by a model, without extracting the needed information from the image.
-For these reasons the CLEVR dataset aims to reduce the biases as much as possible in both images and questions and answers.
+For these reasons the CLEVR framework aims to reduce the biases as much as possible in both images and questions and answers as well as precisely control the remaining biases to explore their limits.
+% SD: Rather, being an artificial dataset, it allows us to precisely control the bias and therefore explore its limits.
+% DK: (added; done)
 Secondly, datasets may come with only a limited amount of annotations and information about the state in an image.
 The CLEVR dataset uses artificially rendered 3D-scenes.
-By doing so, all information about for instance the location of objects or their relations to each other can be saved and later used in training models or analyzing their results.
+By doing so, all information about for instance the location of objects or their relations to each other can be later used in training models or analyzing their results.
+Furthermore, it allows making predictions about different effects of varying contexts.
+% SD: We know the ground truth function that generated the scenes and hence we can also make predictions about different effects of contexts.
+% DK: (added; done)
 
-For this thesis, only the visual part of the dataset will be used and extended.
+For this thesis, the CLEVR framework will be extended to have more control over the generation of the images.
+With this extension, several datasets are created.
+The extensions are described in Chapter \ref{sec:creation-dataset}.
+Hereby, only the images and their ground truth properties are interesting for the present study of referring expressions.
+% SD: We will use the code to generate a new dataset of scenes and descriptions with the properties we want to study.
+% SD: You are not just taking their dataset but you are taking the framework and you are applying it on a new task, to generate a new dataset(s) with carefully controlled properties.
+The CLEVR framework is suitable for this because…
+% DK: made clearer, what is done in the thesis, and what described in which chapter (done)
 The questions and answers won't be used.
+In the following section the image generation of the original CLEVR framework is described.
 The visual part contains images of 3D-generated scenes depicting different kinds of simple objects.
@@
 The possible values of these attributes are listed in Table \ref{tab:clevr-attributes}.
+% SD: Reference to GitHub, also perhaps in the following text when you describe your own code, it would be good to include a reference to the code as you go along.
+% DK: this only describes the original framework, my extension is in the next chapter with reference (done)
 Three to ten objects are placed in random locations into the scene and assigned with random attributes.
@@
 The position of the light and the camera are slightly jittered for each image to add noise reduce recurring patterns.
+Since the objects are part of 3-d scenes, they may appear differently for each image, because of different lighting and shadows, distances to the camera and rotations.
+This noise approximates the real world and natural environment, and makes it harder to learn for models compared to relatively noise-free projections on a 2-d plane.
 Figure \ref{fig:clevr-example} shows an example of a generated image in the CLEVR dataset.
@@
     \label{fig:clevr-example}
+    % SD: Give also example of ground truth features that were used to generate this, i.e. objects and attributes.
+    % DK: TODO
 \end{figure}
@@
 In addition, simple spatial relations (in front of, behind, left, right) between the objects are calculated and stored.
+% SD: For this, functions from the existing code are used, right?
+% DK: exactly. Here, it is only described, what is part of the original CLEVR dataset (done)
 These are simply based on the 3D-coordinates of the objects in relation to the position of the camera.
-
-In this thesis, the code for the generation of the images is extended, and multiple new datasets are created.
-The extension is detailed in section \ref{sec:creation-dataset}.
 
@@
 
-First, I use the VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
+First, we use the VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
 Using 16-19 convolutional layers with small convolution filters helps the model to solve localization and classification tasks on the training dataset, but also enables is to generalize onto other datasets.
@@
 
-Secondly, I include the ResNet-101 \citep{He2016}.
+Secondly, we include the ResNet-101 \citep{He2016}.
 This architecture tries to overcome the degradation of very deep networks, where the accuracy rapidly drops after it gets saturated.
@@
 
-Both architectures are available pretrained on an image classification task on the ImageNet dataset.
+Both architectures are available pretrained on an image classification task on the ImageNet dataset \citep{Deng2009}.
 In this thesis, the implementations and weights available for PyTorch are used.\footnote{\href{https://pytorch.org/hub/pytorch\_vision\_resnet/}{https://pytorch.org/hub/pytorch\_vision\_resnet/}}\textsuperscript{,}\footnote{\href{https://pytorch.org/hub/pytorch\_vision\_vgg/}{https://pytorch.org/hub/pytorch\_vision\_vgg/}}
 Since the task in this research is very different from a classification, it likely learned representations that are not directly transferrable to other tasks.
+The pretrained knowledge in both models might not be directly transferrable to new tasks as studied by \citet{Yosinski2014}.
+In this thesis, the task differs for multiple reasons.
+First, the used images for the pretraining and the present study are part of a different domain.
+Even though images of the CLEVR framework are generated to resemble the real world, they still only include abstract geometric objects while ImageNet contains real photos of persons, animals and objects.
+Furthermore, the original task for the pretraining is a classification task, whilst this thesis is interested in generating and understanding referring expressions.
+% SD: What are the differences? Mainly in the domain of objects and scenes. What data are VGG19 and ResNet trained on?
+% DK: (added; done)
 For this reason, multiple different adaptions of these architectures are compared.
@@
 
-Furthermore, I experimented with both the pretrained models as well as with the architectures trained from scratch with a random initialization for the weights.
-This reason for this was to test if the success of an experiment was actually making use of the pretrained knowledge incorporated in the models.
-If that was not the case, the agents were likely not using image features, but instead relying on some other underlying patterns to solve the task.
-Basically, this approach works as an indicator to determine the actual success of the agents aside from measures as the accuracy or precision.
+% Furthermore, we experimented with both the pretrained models as well as with the architectures trained from scratch with a random initialization for the weights.
+% This reason for this was to test if the success of an experiment was actually making use of the pretrained knowledge incorporated in the models.
+% SD: Link to the discussion on what datasets the systems are trained on.
+% If that was not the case, the agents were likely not using image features, but instead relying on some other underlying patterns to solve the task.
+% Basically, this approach works as an indicator to determine the actual success of the agents aside from measures as the accuracy or precision.
+% SD: This is interestingly more complex and deserves much longer discussion.
+Pre-trained knowledge may be unhelpful if it is very different from the new domain.
+Hence, at some point training from scratch may be more successful.
+% SD: What is success? Learning may take longer but then after some time the performance is better than with pre-training. 
+% SD: Since agents are free to invent new language it is not clear whether this will be align with the human language that influenced pre-training the visual models, hence there is both effect of domain and language here.
+% SD: We should make these questions part of a discussion, particularly if we are comparing the effects of pre-training vs training from scratch in the experiments.
+% DK: learning from scratch not part of the thesis anymore, since not the focus and much more complex (done)
 
@@
 To do this, all experiments rely on the \emph{\textbf{E}mergence of lan\textbf{G}uage in \textbf{G}ames} (EGG) framework \citep{Kharitonov2019} which is implemented in PyTorch.
-This framework consists of a heavily configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
-The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, an easy switch between different loss functions or a choice between Gumbel-Softmax relaxation and REINFORCE algorithms to learn neural models containing discrete symbols.
+% SD: What is the framework? We saw that language games can be implemented very differently, with robots, as code, etc. it is a neural model.
+% DK: (added; done)
+This framework allows the implementation of language games in code, where agents are neural models that communicate with each other.
+It consists of a heavily configurable core that controls the generating and parsing of the message, the calculation of the loss and the rules, for how the weights of all neural models are trained.
+% SD: You frequently start discussion by referring to quite specific concepts straight away in a very concise way, e.g.
+Gumbel Softmax, Reinforce, without defining them but then an explanation comes later in the text.
+The problem with this a reader would wonder what they are and would look for explanation.
+It is best to have a very general introduction, e.g.
+we will test two optimisation functions and then introduce them and explain them in one go later.
+% DK: (rephrased; done)
+The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, an easy switch between different loss functions or a choice between two optimization functions (Gumbel-Softmax relaxation and REINFORCE algorithms) to learn neural models containing discrete symbols.
+% SD: More detailed description of all these. Readers will not be familiar with Gumbel-Softmax for example.
+% DK: (rephrased; done)
 Furthermore, runs of games can be saved to analyze the used messages of the agents and how they vary over the duration of the learning.
 
-The traditional approach to solve these problems relies on the REINFORCE algorithm \citep{Williams1992}.
+The traditional approach to solve problems with a discrete channel relies on the REINFORCE algorithm \citep{Williams1992}.
 It uses the concept of Monte Carlo sampling to estimate the gradient of the expected cumulative reward with respect to the policy parameters.
 The basic idea is to collect trajectories by following the policy, compute the cumulative rewards for each trajectory, and then update the policy parameters to increase the probabilities of actions that led to higher rewards.
-REINFORCE is easy to implement and understand, but it suffers from high variance in gradient estimation due to the inherent randomness in the environment and the policy.
+% REINFORCE is easy to implement and understand, but it
+REINFORCE suffers from high variance in gradient estimation due to the inherent randomness in the environment and the policy.
+% SD: :-)
+% DK: (done)
 
@@
 It allows for the application of gradient-based optimization methods to discrete optimization problems.
-In the Gumbel-Softmax approach, randomness is introduced using the Gumbel distribution (a continuous distribution) and then the \emph{softmax} operation creates a distribution over discrete actions that can be differentiated.
+In the Gumbel-Softmax approach, randomness is introduced using the continuous Gumbel distribution.
+Following, the \emph{softmax} operation creates a distribution over discrete actions that can be differentiated.
 Gumbel-Softmax offers stable gradient estimates and can be more efficient than the traditional REINFORCE algorithm, especially when dealing with large action spaces.
+This also applies to language games, where \citet{Havrylov2017} demonstrated that Gumbel-Softmax relaxation is more effective.
+% SD: This is what a technical manual would say - but what does it do really?
+What is the difference between reinforce and Gumbel softmax in practice?
+In our case?
+Reinforce is still superior but here you make it less.
+% DK: I don't think, REINFORCE is superior also in our case, since GS offers more stability in learning (see citing) (TODO, explain GS)
 
@@
 Furthermore, it uses the output of the receiver and calculates the loss, which is then the basis for the adaption of the weights for both wrappers and agents.
+% SD: Include a diagram?
+% DK: TODO
 
@@
 Gumbel-Softmax relaxation is applied to produce discrete symbols.
-This is done using two LSTMs, an encoder LSTM in the sender wrapper and a decoder LSTM in the receiver wrapper.
+This is done in the default method of the EGG framework using two LSTMs, an encoder LSTM in the sender wrapper and a decoder LSTM in the receiver wrapper.
 The output of the sender agent is used as the initial hidden state for the encoder LSTM.
+% SD: This sounds like details of your implementation and should come later.
+Here, we would only need a general description of the EGG framework.
+Also for this text later, it would be good to have a diagram of all these steps and examples what these games actually are.
+% DK: this is actually the default implementation in EGG in therefore described here (TODO, diagram)
 This LSTM is then producing symbols until it generates an end-of-sequence symbol.
@@
 These losses are summed up to a total loss that is used to adapt the weights in both agents as well as in both LSTMs.
-
-\cmtDK{Gumbel-Softmax vs REINFORCE}
 
@@
 Looking at the datasets that are used in this thesis, all data is created artificially and contains therefore no personal information.
-Even further, the aim of the creation of these datasets is to reduce biases.
+Even further, the aim of the creation of these datasets is to reduce and study the remaining biases.
+% SD: to study biases
+% DK: (rephrased; done)
 It doesn't include any social information, but on the other hand consists only of abstract scenes.
@@
 For that reason, the final models can't be uses in any real world applications and produce potential harm.
+On the other hand, this work can on the long run mitigate harm as it provides a study of models, how they would behave on real data.
+This therefore contributes towards interpretability of AI.
+% SD: The work can on the long run mitigate harm as it is provides a study of models, how they would behave on real data and therefore contributes towards interpretability of AI.
+% DK: (added; done)

### 2023-08-30 thesis/sections/4-exp-setup.tex
comments, feedback

@@
 
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
+In this section we will describe which experiments were conducted to answer the research questions described in the previous chapters.
 Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
@@
 Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
+In Chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
+Furthermore, these experiments provide the basis for the language games, which are then described in Chapter \ref{sec:language-games}.
 
@@
 \label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
+This research investigates how agents use referring expressions to discriminate objects seen in images based on their relations.
+For that reason, the original CLEVR framework offers too little control over how a new dataset is created.
+Especially, which objects and in which attributes they share with each other in each image can't be controlled.
+Following, we extended the framework for generating new datasets.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+By this, the objects in the generated images are controlled to have different human-recognizable attributes, namely the \emph{shape}, \emph{size} and \emph{color}.
+These attributes also correspond to referring expressions in natural language such as English.
+% To simplify the generation and the succeeding training of the models, we only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
 The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
+There were three main extensions to the framework:
+% SD: Too informal
+% SD: We create a new dataset based on the CLEVR framework where we control the appearance of scenes in a referring expression task.
+% SD: The scenes are controlled by human-recognisable attributes of object such as object shape, colour and type.
+These attributes also correspond to referring expressions in natural language such as English.
+% SD: We create two contexts of scenes, one with two objects and one with five. 
+% SD: We also control for the number of attributes shared between the target and the distractor.
+% DK: creation of dataset later (here only the general rules), (added; done)
 
@@
 If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
+The number of shared attributes can also be set to a range to control how challenging the referring task is.
+% SD: To control how challenging the referring task is
+% DK: (done)
 
@@
 This allows for a very controlled generation of relations between the objects in the scene.
+% SD: Introduce other images earlier?
+% DK: I think 'Dale-5' is the most interesting image to make the point here.
+The others are introduced later, in the order of the pictures.
+So I don't think I also can change the order the pictures.
+(done)
 Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
@@
 There are no distractor objects.
+% SD: It depends what you mean by distractor objects.
+I’d say they are all distractor objects by the virtue they share attributes.
+The system has to learn understanding of an intersection of these attributes.
+% DK: I refer to the definitions in the paragraphs above, namley target object, target groub and distrctors. (done)
 
@@
 Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
+% SD: Note that since objects are generated as a part of 3-d scenes, they may appear differently on images: size, occlusion, shading, rotation.
+We also expect the model to learn from this noise and which therefore makes the task much harder and a task that approximates natural environment compared to the task where we would use abstract geometric shapes projected on a 2-d plane.
+% DK: added in the general description of CLEVR since it is not my extention (done)
+% TODO: we create the follwoing dataset ....
 
@@
 To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
+% SD: Wouldn’t it make sense to introduce the discussion of how scenes can be generated with attributes somewhere here? Again, I felt we were projecting forward earlier without understanding the task.
+% DK: before, I only wanted to outlie the rules for the creation, i.e.
+what is possible with it.
+Since the code to create the datasets is also a contribution, I wanted to let it stand on its own.
+Here I only describe how these rules are applied for each new dataset.
+(QUESTION)
 Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
@@
 The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
+% SD: But why not test the limits and the harder examples?
+% DK: In the experiments, the models often hda problems to learn anything (with the complex dataset), which is why I reduced the complexity with the 'color' dataset. (done)
 
@@
 Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
+In other words, the choice of attributes is random for the distractors, and they may overlap.
+% SD: In order words, the choice of attributes is random for distractor and there may be an overlap between the same distractor objects.
+This is because we do not control overlap attributes over distractor, I.e.
+random.
+% DK: (added; done)
 
 \subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
+% SD: In the previous dataset the target and distractor are discriminated by ONE attribute exactly.
+% DK: TODO
+The above described dataset is very restrictive in the relation between the objects, where only \emph{one} attribute is used to disambiguate them.
 The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
+In the real world, objects have overlapping attributes and hence objects can only be identified by an intersection of multiple attributes.
 In real situations, there is no restriction at all how objects or things relate to each other.
@@
 This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
+% SD: In real world, objects have over-lapping attributes and hence a single object can only be identified by an intersection of attributes.
+% DK: (rephrased; done)
+
+% SD: However, we do not do this randomly but ge5 inspiration from the Dale and Reiter generation algorithm who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
+This way we approximate the information in the scenes to human cognition and we hope that the system will be able to exploit that (we would really need to study and compare it with another dataset where attributes are add3d randomly to confirm that there is an effect of h7man.
+Ignition).
+% DK: see below (done)
+For this, we created a dataset that allows almost any relation between a target object and the distractors.
+However, the creation is inspired by incremental algorithm for the Generation of Referring Expressions (GRE) described in \citep{Dale1995} who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
 This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
@@
 On the other side, it is not controlled which attributes are shared.
+% SD: The sharing of attributes should be according to the Dale and Reiter hierarchy, shouldn’t it? Two attributes, big blue (cube|sphere)
+% DK: During the creation of the dataset, only the number of shared attributes is shared that we get different numbers of overlaps.
+Only the captions (that are created later and used as input or output for some models/agents) rely directly on the algorithm.
+At this point, I only wanted to bring up the reason for this, which is the usage of the algorithm in human referring expressions and how it compares to machine referring expressions.
+(done)
 These are assigned randomly.
@@
 The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
+% SD: It’s actually the reverse in which descriptions are generated, we start left to right: large purple (are shared), the last attribute must thus be unique.
+Generation proceeds in the opposite order, one would just say sphere.
+We should clarify this.
+Now it is very hard to understand.
+% DK: That's true for the generation of the descriptions.
+This is explained in chapter 4.2.1 where the captions are needed for the training as input.
+As said before, the algorithm was not directly used and is just the reason.
+(done)
 
@@
 There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
+% SD: This should go at the beginning where language games with neural agents are introduced.
+% DK: TODO
+
+For this reason, we decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
 This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
+Mainly, the tasks are separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
+With this division, we can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
+
+% TODO maybe move to front...
 The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
 Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
+Here, we structured the experiments into three levels.
 In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
+% SD: This is a natural language understanding or reference resolution task.
+% DK: TODO maybe put into methods...
 In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
+% SD: Object identification task
+% DK: TODO
 In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
+% SD: Referring expression generation task
+% DK: TODO
 These combined experiments should lay the basis for how to build up the agents in the language games.
+% SD: We are validating the dataset on 3 tasks
+% DK: TODO
 
 \subsubsection{Coordinate predictor}
+% SD: Reference resolution task
+% DK: TODO
 \label{sec:coordinate_predictor}
@@
 The models are therefore tasked to predict the center coordinates of the target object.
+% SD: Reference resolution with identification of location.
+The other task, object identification is also resolving reference but it is easier as you are only picking objects.
+The first task also involves spatial knowledge.
+% DK: TODO
 With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
@@
 In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+% SD: What are the features?
+% SD: Note that these are visual features and not spatial features.
+It is true that visual features also encode some spatial information, how visual features relate to each other, but such information is very different from the spatial information required to predict coordinates in a coordinate frames, and hence we expect the task will be very hard.
+% DK: TODO
 Here the image is first passed through one of the \emph{feature extractors}.
@@
 With this simple setup, the model is theoretically able to focus on an object in the image.
+% SD: Very precisely. This is higher resolution that attention in the visual models that operates on o 7x7 grid, normally.
+% DK: TODO
 The problem arises as soon as multiple objects are present in the image.
@@
 Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
+Here, we try out four different ways to encode and refer to the target object.
 
@@
   \caption{Simplified architecture of the coordinate predictors}
+  % SD: How are these encoded?
+  % DK: TODO
 \end{figure}
 
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
+In the first method, we encode the attributes of the target object as \textbf{one-hot encodings}.
 There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
@@
 
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
+In an extension to this method, we also include the \textbf{center coordinates of all objects} in the image.
+% SD: Encoded locations?
+% DK: TODO
 This should help the model to identify all possible options to chose from, when predicting the target object.
@@
 The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
+For this model, we also made use of a more complex way to encode the image.
+Here, we based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
 The image is first passed through the feature extractor.
@@
 The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
+% SD: Why shuffled? If we order them the way they appear in the image, the model will have more information that they are sequentially related.
+% DK: TODO
 This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
+% SD: Convolutions are only applied on visual features, not object attribute features and locations.
+% DK: TODO
+
+The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citep{Dale1995}.
 This opposes the idea described before, to share as few human knowledge as possible with the model.
 Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
+% SD: The question you are asking is whether it is possible to predict location from the visual appearance of the object.
+This is highly complex task as it requires quite two step reasoning, identification of features with that attribute, e.g.
+blue and then locating that feature in the image.
+% DK: TODO
 If the model is able to match parts of the image with human words it would show that the model learned this attribute.
@@
 Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
+% SD: See my point earlier about the reversed algorithm between the two tasks 
+% DK: TODO
 
@@
 Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
+This architecture is also inspired by the baselines described in \citep{Johnson2017}.
+% SD: \citep and \citet Use Name (year) when you are referring to particular person and (Name, year) when you are referring to a paper. Hence, in this case it would be the latter.
+% DK: done
 
 The forth method, to encode the target object utilizes \textbf{masking} of the image.
+% SD: Oh, before we talked about 3 methods
+% DK: done
 For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
@@
 It therefore can only rely on its own extracted visual features when looking at masked images.
+% SD: The information in the masked image is still not enough to identify the precise location, only the region.
+Hence, it corresponds on attention in the attention models.
+But those are generating labels and not predicting coordinated.
+The task is still challenging.
+% DK: TODO
 The model is presented with both the original image and the masked image.
@@
 If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
+% SD: Yes that’s correct but through several epochs we hope we will refine the distance and standard deviation of the error.
+It should level out.
+It is not a problem.
+What you do with a circle and accuracy is that you make the task easier as your pointer is now not pointing to a point but to a larger area.
+% DK: TODO
 It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
+For that reason, we also introduced an accuracy score.
+For that we defined a fixed size circle with a radius of 20 pixels around the center of each object.
 If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
 These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
+% SD: But here the score will face the same problem with steward deviation. Hence the only difference here is that pointing is less precise.
+% DK: TODO
 This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
@@
 Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
+% SD: Explain. w3 had a paper with John Kelleher on what spatial information is encoded in CNNs
+% DK: TODO
 Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
@@
 'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
+% SD: Point at not just discriminate
+% DK: TODO
 Latter task is assumed to be significantly harder.
@@
 An interesting observation is the difference of the mean distances between training and testing data.
+% SD: How can we have epochs on the testing data? There is no updates so i5 does not make sense.
+% DK: TODO
 The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
@@
 Applying a higher dropout didn't have an impact on the results.
+% SD: The results show that the model is learning something from the training data but this is not the feature that should be learning as the performance on the test data is low.
+Actually, is it low, it is 10, 35 and 45 pixels.
+We should not expect any difference between epochs as there is no training.
+Hence, a flat line is expected.
+% DK: TODO
 
@@
 Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
+% SD: Because it is in the same location?
+% DK: TODO
 Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
@@
 In most cases, it stays above 50 pixels.
+% SD: What is the image size? What is the size of a typical object? 50 sounds quite good. Attention is predicting one of 7x7 blocks. Is 50 more or less than 1/7 of an image? 
+% DK: TODO
 Using the Dale-2 dataset, the behavior is a little different.
@@
   \label{tab:feature-extractor-mean-distances}
+  % SD: Explanation of labels
+  % DK: TODO
 \end{table}
@@
 The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
+% SD: Perhaps arranging the objects sequentially would help the model learn spatial contiguity, see my earlier comment.
+% DK: TODO
 The model is still not able to understand, which object is the target object.
@@
 Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
+% SD: Table with these results?
+% DK: the table would show the same figures as the existing table. Should be still included? (QUESTION)
 
@@
 An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
+Here, we visualized the predicted coordinates compared to the ground truth coordinates.
 Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
@@
 Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
+This shows general patterns of the models predictions over the complete dataset.
 Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
@@
 As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
+% SD: But this image does not show clearly whether the circles for the same object match. You could calculate the average error in distance between the ground truth and predicted coordinates.
+% DK: that is done in the quantitative analysis (mean square error above). The visualization should show general learned patterns of the model (as in this case predictions toard the center)
 
@@
 In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
+% SD: They are closer to the target than distractor and you can see that it is working to a point. But remember this task is very challenging and I wouldn't say the results are so negative.
+% DK: right, this is addressed in the conclusions below
 These findings align with the mean distance scores, described in the sections before.
@@
 They form roughly the shape of a smaller rhombus.
+% SD: There is a bias towards the centre but I would not say that the model has not learned anything.
+I'm surprised that it works so well given the feature representation we have, i.e.
+no geometric features.
+% DK: again, conclusion below
 This behavior can be observed for all datasets and architectures of the model.
@@
 Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
+% SD: The centre bias could be the way the error function is used, that is averages all distances and of course this has a tendency to some middle distance.
+A solution would be to have better geometric features that could take these errors better into account.
+% DK: TODO (future work)
 
@@
   \caption{Visualization of the models' predictions in the Dale datasets}
+  % SD: I Haven't checked the earlier captions, but the captions should be informative in the sense that one does not need to look for the text to understand them.
+Hence, include a brief summary of what each figure contains.
+  % DK: TODO
   \label{fig:visualizations_dale_complete}
@@
 First, the models are biased to predict coordinates in the center of the image.
+% SD: centre - using British English?
+% DK: so far I used always American English (hopefully consistently)
 The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
@@
 Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
+% SD: There is a centre bias and the model can partially locate the target.
+The task is hard.
+The features are not optimal for this task.
+Future would should focus on using geometric features - we should have tried these and I'm sure we would get better results.
+Another extension would be making pointing less precise, i.e.
+focus on a single point, i.e.
+using a circle of 20 pixels or a 7x7 grid as in the attention mechanism.
+This would simplify the task.
+% DK: TODO (future work)
 
@@
 In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not necessarily important for classifying tasks.
-To address this challenge, I restructured the task in this experiment.
+To address this challenge, we restructured the task in this experiment.
+% SD: Just say that this is a task of identifying objects (the previous was about locating them).
+This task does not rely on geometric information but it relies on identifying correct object features, hence explore different features and hence it is useful as abenchmarking task.
+% DK: TODO
 Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects, target object and distractors.
@@
   \label{fig:bounding_box_classifier_architecture}
+  % SD: I haven't commented on this before, but let us be very careful in naming all meta-paramaters of the models in the text or in caption: how many layers, how many dimensions, etc.
+The reader should be able to replicate the code and the experiment from this.
+You can do this in the caption which is a longer explanation or in the text itself.
+  % DK: TODO
 \end{figure}
@@
 The model passes each of the bounding boxes through a linear layer to transfer it into an embedding space of 10 dimensions.
+% SD: This is quite low? Why only 10?
+% DK: to calculate the dot product (inserted; done)
 The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
+By bringing both vectors to the same embedding size, the dot product between each embedded image and the embedded attributes can be calculated in the next step.
 A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model then points to an object, using the \emph{softmax} function over all concatenated dot products.
+The dot products are concatenated and form a vector with as many dimensions as objects present.
+The \emph{softmax} function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
+% The model then points to an object, using the \emph{softmax} function over this resulting vector.
 The loss is calculated using cross entropy.
+% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
+% DK: the output is a probability distriution over all objects (inserted above; done)
 
@@
 Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
+% SD: I don't understand this?
+Have you actually implemented such model?
+But how can you train a random classifier?
+The baseline would normally be 1/5 or 1/5, i.e.
+the number of labels if the number of examples is equal, other a weighted proportion of these.
+By definition of "random" you would expect such classifier to achieve these probabilities.
+% DK: the idea was not to just compare it to the random probabilities, but compare it to a similar setup as the actual model, with the only difference being the missing information of the target object.
+As you say the probability is 1/5 (resp.
+1/2), which is also the result of thi model.
+Better to just compare to random probability?
+(QUESTION)
+
+The bounding box classifier is trained on all datasets that inlcudes more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
+% excluding the 'CLEVR single'.
+% SD: Why not?
+% DK: because there is only one object in the input, and therefore only one possible guess. The model can't point to a distractor (adressed; done)
 The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
@@
 A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
+% SD: The model is able to leatn to discriminate objects based on visual appearance.
+% DK: TODO
 
@@
 As the bounding box classifier, the caption generator acts as a learning step towards understanding, how the model can learn the attributes of objects.
+Additionally, this setup is a natural language generation task, where the model is trained to generate referring expressions, based on visual input.
+% SD: Well, this experiment is a natural language generation task: whether the model is able to generate referring expressions.
+% DK: (addressed; done)
 The same method using the GRE-algorithm from the section above is used to create the descriptions of the target object for each image.
@@
 By this, the positions of the slots are preserved and if not specified just filled with a padding token.
+Each slot has always the same semantic value, e.g. the last slot always contains the shape of an object.
+This can be done since the captions are not free text, but instead the structure and the possible content is given by the dataset.
+This method might help the model to learn the correct referring expressions.
+% SD: Slots have the same semantic value, i.e.
+the last slot is the object type.
+We can do this because we know the structure of the descriptions in this dataset and because we hope this will help the learning.
+Also, we are thinking of generation based on slots.
+% DK: (addressed; done)
 The last variation concerns the order of producing each token.
@@
 This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
+Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the caption.
+% SD: We... I sounds more like you are writing a lab report with casual language but here you want to be more formal.
+% DK: (done)
 Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
@@
 Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
+Can the model solve the task, or do specific attributes used by humans pose challenges to the model?
+% Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
+% SD: In this case it should be using the same attributes as humans as we are training it on human labels.
+% DK: (rephrased; done)
 
@@
 
-I compare two different models against each other.
+We compare two different models against each other.
 The first model just takes the image as the input.
@@
 The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM.
+% SD: Dimensions?
+% DK: 1024? (QUESTION)
+During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
+% SD: Why is this teacher forcing?
+% DK: because the ground truth during generation of the sequence (added subclause; done)
 The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
+The loss is calculated using cross-entropy.
 During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
+Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.
+% SD: There is a greedy decoding algorithm.
+% DK: (added; done)
 
 Using this approach, the model doesn't have any information about the target object.
-Therefore, I extend it in a second step with a masked image.
+Therefore, we extend it in a second step with attention to the target object.
+% SD: You add attention to the target object.
+% DK: (added; done)
 A masked version of the image is created ass in the section above and passed to the model.
@@
 The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
+% SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
+% DK: (QUESTION)
 Both encodings are concatenated and then used as the initial hidden state for the LSTM.
@@
 The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
+% SD: This is too strict.
+if this is a generation task then we can use generation metrics.
+Those compare if the generated string as a whole is kind of similar to the target string, see ROUGE, METEOR, BERTscore.
+% If this is a classification task, and it is, given that we know what type of word we expect in each slot, we could calculate accuracy per class, precision, recall, F-score.
+We could also calculate mean ranks for predicting words per classes, or acucracy at k- if the target word is in the k-top predictions.
+% We could also take it a multi-class classification problem and how to calculate accuracy in this case, would need to check.
+% DK: I agree that classification metrics refelct success better here (therefore I also used accuracy).
+These metrics would give a better overview, but I don't think there is time to implement it (QUESTION)
 This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
@@
 For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
+% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
+% DK: TODO
 The correct caption is \emph{cylinder}.
@@
 The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
+To mitigate this, we also include a \textbf{word-by-word accuracy}.
 This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
+The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would yield 0\%.
+% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
+% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
 This can give a better understanding of the errors the model makes.
 
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
+% SD: Very good If I understand this right this would be a case where the generated description describes another object better than the target and a human would identify that.
+So it is like false negatives in describing.
+Perhaps use another term for this: referring to a distractor?
+(since you are not measuring hwo individually the description fitts a distractor but only if the entire description could perfectly fit a distractor)
+% DK: exactly. I added false negatives. But I would like to keep accuracy in the name, since that is what it is. maybe better distractor accuracy? (QUESTION)
+With the \textbf{non-target accuracy}, we identify if the model described another object, which is not the target image.
 This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
+By this it measures the false negative generated captions.
+For this we generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
 If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
 If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
+Using this measure, we can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
 
@@
 \subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
+In the next paragraphs, the results of the caption generator are discussed.
 Table \ref{tab:results_caption-generator} shows the different scores of the models when trained on 'Dale-2' and 'Dale-5'.
@@
   \caption{Results of the caption generator, based on ResNet-3}
+  % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1.
+It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
+  % DK: might be the case.
+in the last language game i use only a linear layer instead of convolutions, but it doesn't help.
+This however might also be due to other problems.
+Time?
+(QUESTION, future work)
+  % SD: I would present these for individual classes.
+  % DK: I only have accuracy scores at. Time? (QUESTION)
   \label{tab:results_caption-generator}
@@
 
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
+The sum of the two columns for the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
+% SD: Not shown in Table 5
+% DK: sum of two columns (addressed; done)
 This means that the models produce almost always correct descriptions of one of the objects in the image.
@@
 Features of both objects are therefore extracted perfectly and associated with the vocabulary.
+% SD: But this is no-surprise since we have fixed the slots.
+% DK: but this applies to all padding methods. (done)
 As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
@@
 When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
+Now, without masking only in 61\% of the images, the model describes any of the objects in the image, whether it is the target object or a distractor.
+% SD: ?
+% DK: added subclause (done)
 With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
@@
 This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
+For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
 When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
 This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
+% SD: Not in the table?
+% DK: no, because a result of one specific sample (done)
 
@@
 The final peak stayed exactly the same and the effects were therefore not studied deeper.
+% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
+% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
 
@@
 In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
+% SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated.
+I would only pose this as a very vague hypothesis to be tested, as you conclude now it is also the case that images with simialr distractors are more difficult to describe correctly.
+Hence, it is by no way given.
+% DK: TODO
 The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
+When only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
+Otherwise, the shape is enough, and the caption is only one word long.
+The probability for this lies at 66,6\%.
+% SD: Why 66? We have 3 attributes, right, so 33%?
+% DK: but that the other attributes are only needed if the shape is the same. The prob. for the shape being different for the second object is 66% (rephrased; done)
+For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
 Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
@@
 These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
+% SD: I think we need individual accuracies per class and also precision, recall and the f-score.
+If we include the joint accuracy figure when you also need to explain how you avergaed the infividual accuracies.
+% DK: time? (QUESTION)
+% SD: Not sure. Note that <pad> is a token. The model has to learn when to say <pad> just as any word, hence knowing when not to dewscribe is also trickly.
+% DK: true, but <pad> token is probably the most used token in general and therefore a guess of <pad> just by frequency is likely right? (QUESTION)
 
@@
 The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
+% SD: But there is an importnat difference that now we have two neural netweorks, the sender and the receiver.
+% DK: TODO
 
@@
 The number of possible words, namely the size of the vocabulary can be varied.
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
+% SD: But possibly you would describe the language game setup already before on the background or method chapter.
+I would move all text there and make the language games description much more detailed, also giving examples.
+What is happening in terms of interaction?
+How is this modelled with the neural networks?
+It is important to emphasise this difference.
+% DK: TODO
+
+First, we will discuss \emph{discrimination games}, because they have the simplest setup.
+% SD: What is this setup?
+% DK: TODO
 Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will describe \emph{caption generators} that are set up as a language game.
+In the next step, we will describe \emph{caption generators} that are set up as a language game.
 Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
+In the last step, we try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
+% SD: Why are we not testing it on all 3 games?
+% SD: Not sure what you mean here? Why do we loose human bias? 
+% DK: TODO
 
@@
 
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
+The discrimination games in this research have a very similar setup as described in \citep{Lazaridou2017}.
 The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
@@
 Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
+% SD: On the contrary,...
+% DK: TODO
 It is therefore assumed that the sender will communicate these discriminative attributes.
@@
 Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
+As in \citep{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
+% SD: Here it does not receive a boundig box of an object as in Lazaridou. They use whole images, don't they?
+% DK: TODO
 The sender is assumed to learn which of the input images is the target images by this approach.
@@
 This is then used as the initial state of the encoder LSTM in the sender wrapper.
+% SD: Deatils about the layers, their parameters. More explanatory figure caption.
+% DK: TODO
 
@@
 Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
+% SD: Motivation why this is so. The possible number of strings is {#colour} x {#shape} {object type} where all numbers except the last one also contain a padding token.
+% DK: TODO
 
@@
 The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
+% SD: This discussion of dimensions could already be introduced when you discuss simple modles earlier as the same dimensions are used.
+This is a more an evaluation of represnetations rather than interaction and hence it would fit there.
+% DK: TODO
 
@@
   \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
+  % SD: A note about the datasets chosen. Why do we choose these? Initially, we describe more than 3 datasets.
+  % DK: because the 'CLEVR Single' contains only one object. So there is nothing to discriminate (done)
   \label{tab:results_discriminator}
 \end{table}
+% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
+As David pointed out it may be that convergence is unstable and changes happen in between.
+% DK: TODO
 
@@
 Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
+% SD: How do we know? What would happen if the hidden layers are smaller than the vocabulary size? We would also need to test several smaller dimensions.
+% DK: because the task is solved succesfully, TODO
 
@@
 The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-
-\cmtDK{calculation of loss?}
-\cmtDK{length}
+% SD: Could we say that calculation on Dale-5 is successful?
+% DK: TODO
 
@@
 First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
+% SD: Not sure, since we never tested smaller layer sizes.
+If the embedding layers are the same as the vocabulary size then at least a language encoding LSTM is not motivated to geenralise at all as it can keep representations for individual words without abstracting them.
+Leanring word embeddings involves reducing dimensions.
+% DK: TODO
 This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
@@
 from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
+% SD: Hypothesis space to find the target function is just too big and they can never approximate that function.
+% DK: TODO
 
@@
 Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are 48 possible combinations of attributes.
+Since we discriminate objects based on properties that are also distinguished in human cognition (8 colors, 2 sizes, 3 shapes), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
+There are $8*2*3=48$ possible combinations of attributes.
+% SD: Calculation behind this figure?
+% DK: (done)
 Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
@@
 Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
+% SD: We would need to show that the messages in Dale-2 are shorter than in Dale-5 to claim this.
+% DK: TODO
 
@@
 Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
+% SD: I don't understand how test and training set are used in these experiments.
+In language games one only has a training set and success is measured through the success in the game.
+It is not possible to have a test set as learning is continuous.
+% DK: TODO
 Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
 By this the languages are easier comparable across different runs and vocabulary sizes.
+% SD: I don't understand this.
+% DK: TODO
 Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
@@
 The first property is that when a message is transferred, it consists of only one symbol.
+% SD: Is this controlled by the sender.
+The sender must have some policy how it is assigning messages.
+Or is it just random?
+But what makes the sender then to generate a single message vs a longer message.
+If it is not controlled - how could we implement some policies for message length?
+% DK: what do you mean by policy? The sender is only controled by the success of the receiver. If a longer message leads to a better success, it learns to prodcue longer messages. TODO
 In some rare cases, also an empty message is communicated.
@@
 Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
+% SD: A table or a graph with these figures?
+% DK: is part of the existing figure. The sum of the first two columns (QUESTION)
+% SD: Differences between training and testing configuration; what happens during testing, there is no learning involved but the systems see new images but they have to use the same vocabulary?
+% DK: exactly
 
@@
 However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
+% SD: Are they?
+We haven't tested how humans would describe these scene.
+I see what you mean, but they you have to explain that you count as an English expression a description containing sequences of colour, size and type and then following the Dale and Reiter's algorithm.
+% DK: TODO
 The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
+% SD: All this has to do with the sender's policty how to generate messages.
+We should have varied this policy.
+I suspect no policy leads just to one message.
+How is the system motivated to generate more than one symbol?
+% DK: TODO
 
@@
 Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
+% SD: But this is slightly problematic, since we strcutured the world this way to emphasise these attributes visually.
+Hence, a mapping would indicate that the symbols have good semantics if they can discriminate the colours.
+But this could have to do with the compisitionality issue, perhaps the system was biased to use single word expressions and hence this interfered with the grounding.
+% DK: TODO
 
@@
 In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
+% SD: The probing test is to show how well the emerged language can be trasnlated into English descriptions consistting of labels of scene features using Dale and Reiter idea of incrementality.
+However, we would need to know now what was the policy of the sender to generate message.
+Was this similar to Dale and Reiter to enforce longer descriptions?
+% DK: TODO
 Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
@@
 Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
+% SD: Move to the earlier paragraph when you introduce Dale and Reiter.
+% DK: TODO
 
@@
 In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
+For that reason, two reference baselines are calculated for each of the attribute importance order.
+% SD: Baselines?
+% DK: (done)
 The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
@@
 If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
+On the other side, we calculated the lowest loss possible, by using the English referring expressions as both input and target.
 This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
@@
 A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
+% SD: Is this encessary? But when and where is this loss using 0 inouts calculated?
+% DK: TODO, QUESTION
 By doing this, all configurations can be compared directly to each other.
@@
   \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
+  % SD: I have a trouble of understanding and interpreting these results.
+First the loss should be estimated on a test set, after the system has been trained.
+I'm not sure how you estaimate the baseline loss.
+  % DK: TODO, why a test set?
+Isn't the purpose to learn how well the generated captions map to English and not how well they generalize.
+Using only the train set exactly shows if a mapping is possible
   \label{tab:probing_discrimnator_language}
@@
 Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
+% SD: I think at this point a readerwill be confused since we are discussing Dale and Reiter bith for generation of scenes and now for the generation of strings.
+We should have clarified the difference earlier.
+Also, rather than referring to it as Dale and Reiter we should call it "the incremental algorithm" and add a reference at first mention.
+% DK: TODO
 
@@
 The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
+% SD: The sender uses the incremental algorithm?
+% DK: TODO, QUESTION
 Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
@@
   \label{fig:caption_generator_game_architecture}
+  % SD: But here we don't need an LSTM, we just want to identify one of the objects. The LSTM should only encode the inout message.
+  % DK: TODO
 \end{figure}
@@
 The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
+The image encoders are set up as described in \citep{Johnson2017}.
 A feature extractor extracts basic features of the images.
@@
 This size was found to have the best results during the experiments.
+% SD: Reference back to when you reported this?
+% DK: TODO
 The two resulting encoded images are flattened and concatenated.
@@
 This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
+% SD: But receiver is not doing captioning? Is it incrementally identifying the objects as the sender's messages are coming in?
+% DK: TODO
+During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
+% SD: I'm not sure how teacher forcing is implemented, see my earlier comment
+% DK: typo, teacher forcing is actually applied (done)
 When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
@@
 The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
+% SD: So the sender is generating descriptions following the GRE policy?
+% DK: TODO
 
@@
   \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
+  % SD: What is word-by-word?
+  % DK: the same as in the caption generator with a single model (QUESTION)
   \label{tab:results_caption_generator_game}
@@
 A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
+Here, the agents only manage to produce for 7,8\% of the images correct descriptions with a vocabulary size of 100, 13,2\% points less than the single neural model.
 The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
+% SD: With the vocabulary size of 100.
+% DK: (done)
 
 When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
+% SD: Not sure. The difference is still very small, within a couple of %.
+% DK: TODO
 This is only visible for the 'Dale-2' dataset.
@@
 As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
+% SD: Very small decerase for such a large difference in vocabulary size compared to what we have seen earlier.
+% DK: TODO
 
@@
 For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
+% SD: Here it would be much better to have a loss curve to show differences in loss between the number of communication events, the same could also be done for accuracy.
+Then we would have a clearer picture of how vocabulary size affects learning.
+It could be that with 100 vocabulary there is actually a better performance than before the final cut-off point.
+Is this possible?
+How was the final cut-off point determined anyway?
+% DK: TODO
 
 These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
+% SD: Negative result: how can we then interpret that? What could we change? What did we learn from this?
+% DK: TODO
 This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
@@
 \subsubsection*{Setup}
+% SD: But previously we have said that we are not going to study this game. But good that we have!
+% DK: where? (QUESTION)
 In the final experiments, it is tried to eliminate as much human knowledge as possible.
+% SD: Not true: human knowledge is still reflected in the way the environment is structured.
+% DK: TODO
 For that reason, the agents are tasked to only predict the location of the target object.
@@
 In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
+% SD: I'm not following this
+% DK: TODO
 The receiver doesn't come in contact with any human knowledge, not as input nor as output.
@@
 In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
+% SD: Which contains human knowledge!
+% DK: TODO
 With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
@@
 The resulting vector is passed to the sender wrapper LSTM, to generate the message.
+% SD: Just thinking: would there be a more systematic way to compare all the models?
+Perhaps if you write the parameters in a table or a text that is parallel.
+Then you could simply say, modal A + these changes.
+It would be much clearer to the reader how all models are rerlated and it would be easier for the reader to follow.
+% DK: TODO
 
@@
   \label{fig:coordinate_predictor_game_architecture}
+  % SD: Again, why LSTM, are we inceremtally identifying the image while we are processing each input word?
+  % DK: because it's a message sequence of discrete symbols (done)
 \end{figure}
@@
 All configuration achieve an accuracy higher than 96,7\%.
+% SD: Just occured to me: how do we measure accuracy if we are predicting numbers (numeric prediction)?
+% DK: same as in the experiments with one model (done)
 This aligns also with the results from the single neural models, where the average distance was similarly low.
@@
 Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
+% SD: But how could it do so if the visual data is randomly distribution. It must rely on the message as that is the only source of information.
+% DK: it does it in the same way as the single model above (TODO)
 As in the experiment before, all communicated messages contain the same symbol independent of the input image.
@@
 The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
+% SD: Remind the reader what is the size of the image so that they can interprert the distance relatively to that.
+% DK: TODO
 Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
@@
 The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
+% SD: I don't understand this
+% DK: TODO
 
@@
     {10}  & {10}    & {10}    & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
+    {10}  & {10}    & {10}    & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0,7}           & {45,9}         & {12,7\%}      & {1}             \\
     {13}  & {10}    & {10}    & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
@@
 For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
+% SD: A bit gloomy result. Negative result is okay but then we need an explanation why this is so and have some ideas what to change. What we can learn from these results?
+% DK: TODO
 Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
+% SD: This identifies that the problem lies in how the sender encodes the messages; what is its policy to generate longer strings or reuse the word.
+Without a policy the sender will never be motivated to encode longer messages and therefore rely on single-word expressions.
+% SD: The second problem is the size of these networks.
+We should also test smaller embedding sizes than 10.
+Practice shows that the embedding sizes can be even several magnitues smaller than the vocabulary sizes, cf.
+the Bengio paper and word2vec.
+% SD: The reason why the system is not performing well on the pointing task is that it does not have the right features to learn from.
+It ahs visual information WHAT these objects are (what do they look like) but they do not have information WHERE these objects are,.
+John Kelleher and I wrote an opiiuon piece about the lack of spatial knowledge required to model spatial descriptions in CNNs.
+The same problem is probably occuring here.
+Replacing the features or adding geometric features that would communicate geometric relations that allow pointing would improve the task and also we would expect that a vocabulary would emerge such that some words are more biased towards visual features (to identify objects) and some more to geometric features (to locate them), hence WHAT and WHERE.
+% SD: J.
+D.
+Kelleher and S.
+Dobnik.
+What is not where: the challenge of integrating spatial representations into deep learning architectures.
+In S.
+Dobnik and S.
+Lappin, editors, Proceedings of the Conference on Logic and Machine Learning in Natural Language (LaML 2017), Gothenburg, 12 –13 June, volume 1 of CLASP Papers in Computational Linguistics, pages 41–52, Gothenburg, Sweden, November 2017.
+Department of Philosophy, Linguistics and Theory of Science (FLOV), University of Gothenburg, CLASP, Centre for Language and Studies in Probability.https://gup.ub.gu.se/publication/262970?lang=en
+% SD: Overall, the thesis has a lot of potential if we could implement all this
+% DK: TODO
 

### 2023-09-01 thesis/sections/4-datasets.tex
restructuring

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
+\section{Creation of the dataset}
 \label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
+This research investigates how agents use referring expressions to discriminate objects seen in images based on their relations.
+For that reason, the original CLEVR framework offers too little control over how a new dataset is created.
+Especially, which objects and in which attributes they share with each other in each image can't be controlled.
+Following, we extended the framework for generating new datasets.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
+By this, the objects in the generated images are controlled to have different human-recognizable attributes, namely the \emph{shape}, \emph{size} and \emph{color}.
+These attributes also correspond to referring expressions in natural language such as English.
+% To simplify the generation and the succeeding training of the models, we only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
 The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
+There were three main extensions to the framework:
+% SD: Too informal
+% SD: We create a new dataset based on the CLEVR framework where we control the appearance of scenes in a referring expression task.
+% SD: The scenes are controlled by human-recognisable attributes of object such as object shape, colour and type.
+These attributes also correspond to referring expressions in natural language such as English.
+% SD: We create two contexts of scenes, one with two objects and one with five. 
+% SD: We also control for the number of attributes shared between the target and the distractor.
+% DK: creation of dataset later (here only the general rules), (added; done)
 
@@
 Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
+% SD: Note that since objects are generated as a part of 3-d scenes, they may appear differently on images: size, occlusion, shading, rotation.
+We also expect the model to learn from this noise and which therefore makes the task much harder and a task that approximates natural environment compared to the task where we would use abstract geometric shapes projected on a 2-d plane.
+% DK: added in the general description of CLEVR since it is not my extention (done)
+% TODO: we create the follwoing dataset ....
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\subsubsection*{CLEVR single}
+\subsection{CLEVR single}
 The simplest new dataset is called 'CLEVR single'.
@@
 
-\subsubsection*{CLEVR color}
+\subsection{CLEVR color}
 The second dataset that is created is called 'CLEVR color'.
@@
 Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
+In other words, the choice of attributes is random for the distractors, and they may overlap.
+% SD: In order words, the choice of attributes is random for distractor and there may be an overlap between the same distractor objects.
+This is because we do not control overlap attributes over distractor, I.e.
+random.
+% DK: (added; done)
 
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
+\subsection{CLEVR Dale datasets}
+% SD: In the previous dataset the target and distractor are discriminated by ONE attribute exactly.
+% DK: TODO
+The above described dataset is very restrictive in the relation between the objects, where only \emph{one} attribute is used to disambiguate them.
 The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
+In the real world, objects have overlapping attributes and hence objects can only be identified by an intersection of multiple attributes.
 In real situations, there is no restriction at all how objects or things relate to each other.
@@
 This emergence of referring attributes and their combination is studied deeper in this work.
+% SD: In real world, objects have over-lapping attributes and hence a single object can only be identified by an intersection of attributes.
+% DK: (rephrased; done)
 
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
+% SD: However, we do not do this randomly but ge5 inspiration from the Dale and Reiter generation algorithm who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
+This way we approximate the information in the scenes to human cognition and we hope that the system will be able to exploit that (we would really need to study and compare it with another dataset where attributes are add3d randomly to confirm that there is an effect of h7man.
+Ignition).
+% DK: see below (done)
+For this, we created a dataset that allows almost any relation between a target object and the distractors.
+However, the creation is inspired by incremental algorithm for the Generation of Referring Expressions (GRE) described in \citep{Dale1995} who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
 This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
@@
 \cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
-\subsubsection*{Setup}
-
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-\end{figure}
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
-These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-    \caption{Sender and receiver architectures in the discrimination game}
-    \label{fig:discriminator_architecture}
-\end{figure}
-
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
-        \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_discriminator}
-\end{table}
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-
-\cmtDK{calculation of loss?}
-\cmtDK{length}
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are 48 possible combinations of attributes.
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
-
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
-
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
-    \label{fig:caption_generator_game_architecture}
-\end{figure}
-
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
-    \label{fig:coordinate_predictor_game_architecture}
-\end{figure}
-
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
-
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
-        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2023-09-01 thesis/sections/5-grounding.tex
restructuring

--- 2023-08-30 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section we will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In Chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in Chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-This research investigates how agents use referring expressions to discriminate objects seen in images based on their relations.
-For that reason, the original CLEVR framework offers too little control over how a new dataset is created.
-Especially, which objects and in which attributes they share with each other in each image can't be controlled.
-Following, we extended the framework for generating new datasets.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-By this, the objects in the generated images are controlled to have different human-recognizable attributes, namely the \emph{shape}, \emph{size} and \emph{color}.
-These attributes also correspond to referring expressions in natural language such as English.
-% To simplify the generation and the succeeding training of the models, we only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the framework:
-% SD: Too informal
-% SD: We create a new dataset based on the CLEVR framework where we control the appearance of scenes in a referring expression task.
-% SD: The scenes are controlled by human-recognisable attributes of object such as object shape, colour and type.
-These attributes also correspond to referring expressions in natural language such as English.
-% SD: We create two contexts of scenes, one with two objects and one with five. 
-% SD: We also control for the number of attributes shared between the target and the distractor.
-% DK: creation of dataset later (here only the general rules), (added; done)
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to control how challenging the referring task is.
-% SD: To control how challenging the referring task is
-% DK: (done)
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-% SD: Introduce other images earlier?
-% DK: I think 'Dale-5' is the most interesting image to make the point here.
-The others are introduced later, in the order of the pictures.
-So I don't think I also can change the order the pictures.
-(done)
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-% SD: It depends what you mean by distractor objects.
-I’d say they are all distractor objects by the virtue they share attributes.
-The system has to learn understanding of an intersection of these attributes.
-% DK: I refer to the definitions in the paragraphs above, namley target object, target groub and distrctors. (done)
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-% SD: Note that since objects are generated as a part of 3-d scenes, they may appear differently on images: size, occlusion, shading, rotation.
-We also expect the model to learn from this noise and which therefore makes the task much harder and a task that approximates natural environment compared to the task where we would use abstract geometric shapes projected on a 2-d plane.
-% DK: added in the general description of CLEVR since it is not my extention (done)
-% TODO: we create the follwoing dataset ....
-
-\begin{figure}[h]
-  \centering
-  \subfigure['CLEVR single', \emph{large yellow sphere}]{
-    \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-    \label{fig:clevr-single}
-  }
-  \subfigure['CLEVR color', \emph{small brown cylinder}]{
-    \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-    \label{fig:clevr-color}
-  }
-  \subfigure['Dale-2', \emph{small green cylinder}]{
-    \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-    \label{fig:clevr-dale-2}
-  }
-  \subfigure['Dale-5', \emph{large purple cylinder}]{
-    \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-    \label{fig:clevr-dale-5}
-  }
-  \caption{Example images of each dataset, with the target object specified}
-  \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-% SD: Wouldn’t it make sense to introduce the discussion of how scenes can be generated with attributes somewhere here? Again, I felt we were projecting forward earlier without understanding the task.
-% DK: before, I only wanted to outlie the rules for the creation, i.e.
-what is possible with it.
-Since the code to create the datasets is also a contribution, I wanted to let it stand on its own.
-Here I only describe how these rules are applied for each new dataset.
-(QUESTION)
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-% SD: But why not test the limits and the harder examples?
-% DK: In the experiments, the models often hda problems to learn anything (with the complex dataset), which is why I reduced the complexity with the 'color' dataset. (done)
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-In other words, the choice of attributes is random for the distractors, and they may overlap.
-% SD: In order words, the choice of attributes is random for distractor and there may be an overlap between the same distractor objects.
-This is because we do not control overlap attributes over distractor, I.e.
-random.
-% DK: (added; done)
-
-\subsubsection*{CLEVR Dale datasets}
-% SD: In the previous dataset the target and distractor are discriminated by ONE attribute exactly.
-% DK: TODO
-The above described dataset is very restrictive in the relation between the objects, where only \emph{one} attribute is used to disambiguate them.
-The number and the type of shared attributes are controlled exactly.
-In the real world, objects have overlapping attributes and hence objects can only be identified by an intersection of multiple attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-% SD: In real world, objects have over-lapping attributes and hence a single object can only be identified by an intersection of attributes.
-% DK: (rephrased; done)
-
-% SD: However, we do not do this randomly but ge5 inspiration from the Dale and Reiter generation algorithm who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
-This way we approximate the information in the scenes to human cognition and we hope that the system will be able to exploit that (we would really need to study and compare it with another dataset where attributes are add3d randomly to confirm that there is an effect of h7man.
-Ignition).
-% DK: see below (done)
-For this, we created a dataset that allows almost any relation between a target object and the distractors.
-However, the creation is inspired by incremental algorithm for the Generation of Referring Expressions (GRE) described in \citep{Dale1995} who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-% SD: The sharing of attributes should be according to the Dale and Reiter hierarchy, shouldn’t it? Two attributes, big blue (cube|sphere)
-% DK: During the creation of the dataset, only the number of shared attributes is shared that we get different numbers of overlaps.
-Only the captions (that are created later and used as input or output for some models/agents) rely directly on the algorithm.
-At this point, I only wanted to bring up the reason for this, which is the usage of the algorithm in human referring expressions and how it compares to machine referring expressions.
-(done)
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-% SD: It’s actually the reverse in which descriptions are generated, we start left to right: large purple (are shared), the last attribute must thus be unique.
-Generation proceeds in the opposite order, one would just say sphere.
-We should clarify this.
-Now it is very hard to understand.
-% DK: That's true for the generation of the descriptions.
-This is explained in chapter 4.2.1 where the captions are needed for the training as input.
-As said before, the algorithm was not directly used and is just the reason.
-(done)
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
+\section{Generating and understanding referring expressions}
 \label{sec:preexperiments}
@@
 
-\subsubsection{Coordinate predictor}
+\subsection{Reference resolution}
 % SD: Reference resolution task
 % DK: TODO
-\label{sec:coordinate_predictor}
+\label{sec:reference_resolution}
 \subsubsection*{Setup}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\begin{table}[h]
+\begin{table}[ht]
     \centering
@@
 
-\subsubsection{Bounding box classifier}
+\subsection{Object identification}
 \subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
 
@@
 
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
+\subsection{Referring expression generation}
+\label{sec:referring_expression_generation}
 \subsubsection*{Setup}
@@
 % DK: true, but <pad> token is probably the most used token in general and therefore a guess of <pad> just by frequency is likely right? (QUESTION)
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-% SD: But there is an importnat difference that now we have two neural netweorks, the sender and the receiver.
-% DK: TODO
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the same information that the sender was presented with.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-
-The vocabulary that the sender can draw from to produce a message is made up of initially arbitrary symbols.
-The meaning of these symbols is created as soon as the sender uses them in one of the message, and the receiver is able to use it to solve the task successfully.
-Over time, specific meanings are reinforced and a language emerges.
-The number of possible words, namely the size of the vocabulary can be varied.
-% SD: But possibly you would describe the language game setup already before on the background or method chapter.
-I would move all text there and make the language games description much more detailed, also giving examples.
-What is happening in terms of interaction?
-How is this modelled with the neural networks?
-It is important to emphasise this difference.
-% DK: TODO
-
-First, we will discuss \emph{discrimination games}, because they have the simplest setup.
-% SD: What is this setup?
-% DK: TODO
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, we will describe \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, we try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-% SD: Why are we not testing it on all 3 games?
-% SD: Not sure what you mean here? Why do we loose human bias? 
-% DK: TODO
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citep{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-% SD: On the contrary,...
-% DK: TODO
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citep{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-% SD: Here it does not receive a boundig box of an object as in Lazaridou. They use whole images, don't they?
-% DK: TODO
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-  \caption{Sender and receiver architectures in the discrimination game}
-  \label{fig:discriminator_architecture}
-\end{figure}
-
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-% SD: Deatils about the layers, their parameters. More explanatory figure caption.
-% DK: TODO
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-% SD: Motivation why this is so. The possible number of strings is {#colour} x {#shape} {object type} where all numbers except the last one also contain a padding token.
-% DK: TODO
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-% SD: This discussion of dimensions could already be introduced when you discuss simple modles earlier as the same dimensions are used.
-This is a more an evaluation of represnetations rather than interaction and hence it would fit there.
-% DK: TODO
-
-\begin{table}[h]
-  \centering
-  \begin{tabular}{ccccc|cc|cc|cc}
-    \toprule
-          &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-    $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-    {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-    {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-    {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-    {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-    {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-    {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-    {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-    {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-    \bottomrule
-  \end{tabular}
-  \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-  % SD: A note about the datasets chosen. Why do we choose these? Initially, we describe more than 3 datasets.
-  % DK: because the 'CLEVR Single' contains only one object. So there is nothing to discriminate (done)
-  \label{tab:results_discriminator}
-\end{table}
-% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
-As David pointed out it may be that convergence is unstable and changes happen in between.
-% DK: TODO
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-% SD: How do we know? What would happen if the hidden layers are smaller than the vocabulary size? We would also need to test several smaller dimensions.
-% DK: because the task is solved succesfully, TODO
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-% SD: Could we say that calculation on Dale-5 is successful?
-% DK: TODO
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-% SD: Not sure, since we never tested smaller layer sizes.
-If the embedding layers are the same as the vocabulary size then at least a language encoding LSTM is not motivated to geenralise at all as it can keep representations for individual words without abstracting them.
-Leanring word embeddings involves reducing dimensions.
-% DK: TODO
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-% SD: Hypothesis space to find the target function is just too big and they can never approximate that function.
-% DK: TODO
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (8 colors, 2 sizes, 3 shapes), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are $8*2*3=48$ possible combinations of attributes.
-% SD: Calculation behind this figure?
-% DK: (done)
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-% SD: We would need to show that the messages in Dale-2 are shorter than in Dale-5 to claim this.
-% DK: TODO
-
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$Lang_{10}$]{
-    \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-    \label{fig:relative_frequencies_v10}
-  }
-  \subfigure[$Lang_{13}$]{
-    \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-    \label{fig:relative_frequencies_v13}
-  }
-  \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-    \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-    \label{fig:relative_frequencies_v100}
-  }
-  \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-  \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-% SD: I don't understand how test and training set are used in these experiments.
-In language games one only has a training set and success is measured through the success in the game.
-It is not possible to have a test set as learning is continuous.
-% DK: TODO
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-% SD: I don't understand this.
-% DK: TODO
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-% SD: Is this controlled by the sender.
-The sender must have some policy how it is assigning messages.
-Or is it just random?
-But what makes the sender then to generate a single message vs a longer message.
-If it is not controlled - how could we implement some policies for message length?
-% DK: what do you mean by policy? The sender is only controled by the success of the receiver. If a longer message leads to a better success, it learns to prodcue longer messages. TODO
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-% SD: A table or a graph with these figures?
-% DK: is part of the existing figure. The sum of the first two columns (QUESTION)
-% SD: Differences between training and testing configuration; what happens during testing, there is no learning involved but the systems see new images but they have to use the same vocabulary?
-% DK: exactly
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-% SD: Are they?
-We haven't tested how humans would describe these scene.
-I see what you mean, but they you have to explain that you count as an English expression a description containing sequences of colour, size and type and then following the Dale and Reiter's algorithm.
-% DK: TODO
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-% SD: All this has to do with the sender's policty how to generate messages.
-We should have varied this policy.
-I suspect no policy leads just to one message.
-How is the system motivated to generate more than one symbol?
-% DK: TODO
-
-\begin{figure}[h]
-  \centering
-  \subfigure[$Lang_{10}$]{
-    \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-    \label{fig:language_analysis_v10}
-  }
-  \subfigure[$Lang_{13}$]{
-    \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-    \label{fig:language_analysis_v13}
-  }
-  \subfigure[$Lang_{100}$]{
-    \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-    \label{fig:language_analysis_v100}
-  }
-  \caption{Relative share of the described target object's attributes for the top three messages}
-
-  \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-  \textbf{Size:} brown: small, blue: large
-  \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-% SD: But this is slightly problematic, since we strcutured the world this way to emphasise these attributes visually.
-Hence, a mapping would indicate that the symbols have good semantics if they can discriminate the colours.
-But this could have to do with the compisitionality issue, perhaps the system was biased to use single word expressions and hence this interfered with the grounding.
-% DK: TODO
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-% SD: The probing test is to show how well the emerged language can be trasnlated into English descriptions consistting of labels of scene features using Dale and Reiter idea of incrementality.
-However, we would need to know now what was the policy of the sender to generate message.
-Was this similar to Dale and Reiter to enforce longer descriptions?
-% DK: TODO
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-% SD: Move to the earlier paragraph when you introduce Dale and Reiter.
-% DK: TODO
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference baselines are calculated for each of the attribute importance order.
-% SD: Baselines?
-% DK: (done)
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, we calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-% SD: Is this encessary? But when and where is this loss using 0 inouts calculated?
-% DK: TODO, QUESTION
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-  \centering
-  \begin{tabular}{c|cc|cc|cc|cc}
-    \toprule
-    \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                            &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-    {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-    {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-    {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-    {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-    {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-    {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-    \bottomrule
-  \end{tabular}
-  \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-  % SD: I have a trouble of understanding and interpreting these results.
-First the loss should be estimated on a test set, after the system has been trained.
-I'm not sure how you estaimate the baseline loss.
-  % DK: TODO, why a test set?
-Isn't the purpose to learn how well the generated captions map to English and not how well they generalize.
-Using only the train set exactly shows if a mapping is possible
-  \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-% SD: I think at this point a readerwill be confused since we are discussing Dale and Reiter bith for generation of scenes and now for the generation of strings.
-We should have clarified the difference earlier.
-Also, rather than referring to it as Dale and Reiter we should call it "the incremental algorithm" and add a reference at first mention.
-% DK: TODO
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
-
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-% SD: The sender uses the incremental algorithm?
-% DK: TODO, QUESTION
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-  \caption{Simplified architecture of the caption generator game}
-  \label{fig:caption_generator_game_architecture}
-  % SD: But here we don't need an LSTM, we just want to identify one of the objects. The LSTM should only encode the inout message.
-  % DK: TODO
-\end{figure}
-
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citep{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-% SD: Reference back to when you reported this?
-% DK: TODO
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-% SD: But receiver is not doing captioning? Is it incrementally identifying the objects as the sender's messages are coming in?
-% DK: TODO
-During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
-% SD: I'm not sure how teacher forcing is implemented, see my earlier comment
-% DK: typo, teacher forcing is actually applied (done)
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-% SD: So the sender is generating descriptions following the GRE policy?
-% DK: TODO
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[h]
-  \centering
-  \begin{tabular}{ccc|ccc|ccc}
-    \toprule
-          &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-    $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-    {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-    {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-    {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-    {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-    {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-    \bottomrule
-  \end{tabular}
-  \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-  % SD: What is word-by-word?
-  % DK: the same as in the caption generator with a single model (QUESTION)
-  \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions with a vocabulary size of 100, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-% SD: With the vocabulary size of 100.
-% DK: (done)
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-% SD: Not sure. The difference is still very small, within a couple of %.
-% DK: TODO
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-% SD: Very small decerase for such a large difference in vocabulary size compared to what we have seen earlier.
-% DK: TODO
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-% SD: Here it would be much better to have a loss curve to show differences in loss between the number of communication events, the same could also be done for accuracy.
-Then we would have a clearer picture of how vocabulary size affects learning.
-It could be that with 100 vocabulary there is actually a better performance than before the final cut-off point.
-Is this possible?
-How was the final cut-off point determined anyway?
-% DK: TODO
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-% SD: Negative result: how can we then interpret that? What could we change? What did we learn from this?
-% DK: TODO
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-% SD: But previously we have said that we are not going to study this game. But good that we have!
-% DK: where? (QUESTION)
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-% SD: Not true: human knowledge is still reflected in the way the environment is structured.
-% DK: TODO
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-% SD: I'm not following this
-% DK: TODO
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-% SD: Which contains human knowledge!
-% DK: TODO
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-% SD: Just thinking: would there be a more systematic way to compare all the models?
-Perhaps if you write the parameters in a table or a text that is parallel.
-Then you could simply say, modal A + these changes.
-It would be much clearer to the reader how all models are rerlated and it would be easier for the reader to follow.
-% DK: TODO
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-  \caption{Simplified architecture of the masked coordinate predictor game}
-  \label{fig:coordinate_predictor_game_architecture}
-  % SD: Again, why LSTM, are we inceremtally identifying the image while we are processing each input word?
-  % DK: because it's a message sequence of discrete symbols (done)
-\end{figure}
-
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
-
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-% SD: Just occured to me: how do we measure accuracy if we are predicting numbers (numeric prediction)?
-% DK: same as in the experiments with one model (done)
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-% SD: But how could it do so if the visual data is randomly distribution. It must rely on the message as that is the only source of information.
-% DK: it does it in the same way as the single model above (TODO)
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[h]
-  \centering
-  \begin{tabular}{ccc|ccc|ccc|ccc}
-    \toprule
-          &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-    $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-    {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-    {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-    {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-    {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-    {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-    \bottomrule
-  \end{tabular}
-  \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-  \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-% SD: Remind the reader what is the size of the image so that they can interprert the distance relatively to that.
-% DK: TODO
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-% SD: I don't understand this
-% DK: TODO
-
-\begin{table}[h]
-  \centering
-  \begin{tabular}{cccc|ccc|ccc|ccc}
-    \toprule
-          &         &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-    $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-    {10}  & {10}    & {10}    & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-    {10}  & {10}    & {10}    & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0,7}           & {45,9}         & {12,7\%}      & {1}             \\
-    {13}  & {10}    & {10}    & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-    {20}  & {10}    & {10}    & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-    {100} & {10}    & {10}    & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-    {100} & {100}   & {100}   & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-    \bottomrule
-  \end{tabular}
-  \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-  \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-% SD: A bit gloomy result. Negative result is okay but then we need an explanation why this is so and have some ideas what to change. What we can learn from these results?
-% DK: TODO
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-% SD: This identifies that the problem lies in how the sender encodes the messages; what is its policy to generate longer strings or reuse the word.
-Without a policy the sender will never be motivated to encode longer messages and therefore rely on single-word expressions.
-% SD: The second problem is the size of these networks.
-We should also test smaller embedding sizes than 10.
-Practice shows that the embedding sizes can be even several magnitues smaller than the vocabulary sizes, cf.
-the Bengio paper and word2vec.
-% SD: The reason why the system is not performing well on the pointing task is that it does not have the right features to learn from.
-It ahs visual information WHAT these objects are (what do they look like) but they do not have information WHERE these objects are,.
-John Kelleher and I wrote an opiiuon piece about the lack of spatial knowledge required to model spatial descriptions in CNNs.
-The same problem is probably occuring here.
-Replacing the features or adding geometric features that would communicate geometric relations that allow pointing would improve the task and also we would expect that a vocabulary would emerge such that some words are more biased towards visual features (to identify objects) and some more to geometric features (to locate them), hence WHAT and WHERE.
-% SD: J.
-D.
-Kelleher and S.
-Dobnik.
-What is not where: the challenge of integrating spatial representations into deep learning architectures.
-In S.
-Dobnik and S.
-Lappin, editors, Proceedings of the Conference on Logic and Machine Learning in Natural Language (LaML 2017), Gothenburg, 12 –13 June, volume 1 of CLASP Papers in Computational Linguistics, pages 41–52, Gothenburg, Sweden, November 2017.
-Department of Philosophy, Linguistics and Theory of Science (FLOV), University of Gothenburg, CLASP, Centre for Language and Studies in Probability.https://gup.ub.gu.se/publication/262970?lang=en
-% SD: Overall, the thesis has a lot of potential if we could implement all this
-% DK: TODO
-
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2023-09-01 thesis/sections/6-grounding-language-games.tex
restructuring

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
-The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
-\subsubsection*{Setup}
-
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-\end{figure}
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not necessarily important for classifying tasks.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects, target object and distractors.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels to capture the complete object.
-The bounding boxes of all objects are padded to the maximum number of objects present in an image across the whole dataset with a matrix of zeros.
-For the 'Dale-2', this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset
-These bounding boxes are shuffled.
-Furthermore, to point the model towards the target object, its attributes are encoded as on-hot vectors.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to transfer it into an embedding space of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model then points to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even with its relatively simple architecture.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator acts as a learning step towards understanding, how the model can learn the attributes of objects.
-The same method using the GRE-algorithm from the section above is used to create the descriptions of the target object for each image.
-
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-The task for the model is to generate these descriptions, based on the complete scene as input.
-Doing this, the task inherently involves learning human knowledge and natural language structure.
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Using this approach, the model doesn't have any information about the target object.
-Therefore, I extend it in a second step with a masked image.
-A masked version of the image is created ass in the section above and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model towards which object to describe and discriminate from the distractors.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of one of the objects in the image.
-Moreover, these descriptions are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and associated with the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object rises with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
-The final peak stayed exactly the same and the effects were therefore not studied deeper.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
+\section{Language games}
 \label{sec:language-games}
@@
 The number of possible words, namely the size of the vocabulary can be varied.
+% SD: But possibly you would describe the language game setup already before on the background or method chapter.
+I would move all text there and make the language games description much more detailed, also giving examples.
+What is happening in terms of interaction?
+How is this modelled with the neural networks?
+It is important to emphasise this difference.
+% DK: TODO
 
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
+First, we will discuss \emph{discrimination games}, because they have the simplest setup.
+% SD: What is this setup?
+% DK: TODO
 Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will describe \emph{caption generators} that are set up as a language game.
+In the next step, we will describe \emph{caption generators} that are set up as a language game.
 Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
+In the last step, we try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
+% SD: Why are we not testing it on all 3 games?
+% SD: Not sure what you mean here? Why do we loose human bias? 
+% DK: TODO
 
@@
 
-\subsubsection{Discrimination games}
+\subsection{Object identification}
 \subsubsection*{Setup}
@@
 
-\subsubsection*{Results}
+\subsection*{Results}
 Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
@@
     \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
+    % SD: A note about the datasets chosen. Why do we choose these? Initially, we describe more than 3 datasets.
+    % DK: because the 'CLEVR Single' contains only one object. So there is nothing to discriminate (done)
     \label{tab:results_discriminator}
 \end{table}
+% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
+As David pointed out it may be that convergence is unstable and changes happen in between.
+% DK: TODO
 
@@
 
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
-
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
+\subsection{Referring expression generation}
 \subsubsection*{Setup}
@@
     \label{fig:caption_generator_game_architecture}
+    % SD: But here we don't need an LSTM, we just want to identify one of the objects. The LSTM should only encode the inout message.
+    % DK: TODO
 \end{figure}
@@
     \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
+    % SD: What is word-by-word?
+    % DK: the same as in the caption generator with a single model (QUESTION)
     \label{tab:results_caption_generator_game}
@@
 For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
+% SD: Here it would be much better to have a loss curve to show differences in loss between the number of communication events, the same could also be done for accuracy.
+Then we would have a clearer picture of how vocabulary size affects learning.
+It could be that with 100 vocabulary there is actually a better performance than before the final cut-off point.
+Is this possible?
+How was the final cut-off point determined anyway?
+% DK: TODO
 
 These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
+% SD: Negative result: how can we then interpret that? What could we change? What did we learn from this?
+% DK: TODO
+This is especially interesting, compared to single model caption generator in section \ref{sec:referring_expression_generation}.
 In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
@@
 
-\subsubsection{Coordinate predictor games}
+\subsection{Reference resolution}
 \subsubsection*{Setup}
+% SD: But previously we have said that we are not going to study this game. But good that we have!
+% DK: where? (QUESTION)
 In the final experiments, it is tried to eliminate as much human knowledge as possible.
+% SD: Not true: human knowledge is still reflected in the way the environment is structured.
+% DK: TODO
 For that reason, the agents are tasked to only predict the location of the target object.
@@
     \label{fig:coordinate_predictor_game_architecture}
+    % SD: Again, why LSTM, are we inceremtally identifying the image while we are processing each input word?
+    % DK: because it's a message sequence of discrete symbols (done)
 \end{figure}
@@
 
-\subsubsection*{Results}
+\subsection*{Results}
 In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
@@
         {10}  & {10}    & {10}    & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
+        {10}  & {10}    & {10}    & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0,7}           & {45,9}         & {12,7\%}      & {1}             \\
         {13}  & {10}    & {10}    & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
@@
 For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
+% SD: A bit gloomy result. Negative result is okay but then we need an explanation why this is so and have some ideas what to change. What we can learn from these results?
+% DK: TODO
 Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
+% SD: This identifies that the problem lies in how the sender encodes the messages; what is its policy to generate longer strings or reuse the word.
+Without a policy the sender will never be motivated to encode longer messages and therefore rely on single-word expressions.
+% SD: The second problem is the size of these networks.
+We should also test smaller embedding sizes than 10.
+Practice shows that the embedding sizes can be even several magnitues smaller than the vocabulary sizes, cf.
+the Bengio paper and word2vec.
+% SD: The reason why the system is not performing well on the pointing task is that it does not have the right features to learn from.
+It ahs visual information WHAT these objects are (what do they look like) but they do not have information WHERE these objects are,.
+John Kelleher and I wrote an opiiuon piece about the lack of spatial knowledge required to model spatial descriptions in CNNs.
+The same problem is probably occuring here.
+Replacing the features or adding geometric features that would communicate geometric relations that allow pointing would improve the task and also we would expect that a vocabulary would emerge such that some words are more biased towards visual features (to identify objects) and some more to geometric features (to locate them), hence WHAT and WHERE.
+% SD: J.
+D.
+Kelleher and S.
+Dobnik.
+What is not where: the challenge of integrating spatial representations into deep learning architectures.
+In S.
+Dobnik and S.
+Lappin, editors, Proceedings of the Conference on Logic and Machine Learning in Natural Language (LaML 2017), Gothenburg, 12 –13 June, volume 1 of CLASP Papers in Computational Linguistics, pages 41–52, Gothenburg, Sweden, November 2017.
+Department of Philosophy, Linguistics and Theory of Science (FLOV), University of Gothenburg, CLASP, Centre for Language and Studies in Probability.https://gup.ub.gu.se/publication/262970?lang=en
+% SD: Overall, the thesis has a lot of potential if we could implement all this
+% DK: TODO
 
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
+When comparing these results to the neural models in Section \ref{sec:reference_resolution} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
 The model was already not able to solve the task without the increased complexity of a language game.

### 2023-09-01 thesis/sections/7-language-analysis.tex
restructuring

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
-The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
-\subsubsection*{Setup}
-
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-\end{figure}
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
-These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-    \caption{Sender and receiver architectures in the discrimination game}
-    \label{fig:discriminator_architecture}
-\end{figure}
-
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
-        \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_discriminator}
-\end{table}
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-
-\cmtDK{calculation of loss?}
-\cmtDK{length}
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are 48 possible combinations of attributes.
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
+\section{Analysis of the emerged languages}
+\label{sec:analysis_language}
 In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
@@
 However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
+% SD: Are they?
+We haven't tested how humans would describe these scene.
+I see what you mean, but they you have to explain that you count as an English expression a description containing sequences of colour, size and type and then following the Dale and Reiter's algorithm.
+% DK: TODO
 The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
+% SD: All this has to do with the sender's policty how to generate messages.
+We should have varied this policy.
+I suspect no policy leads just to one message.
+How is the system motivated to generate more than one symbol?
+% DK: TODO
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
+% SD: Is this encessary? But when and where is this loss using 0 inouts calculated?
+% DK: TODO, QUESTION
 By doing this, all configurations can be compared directly to each other.
 
-\begin{table}[h]
+\begin{table}[ht]
     \centering
@@
     \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
+    % SD: I have a trouble of understanding and interpreting these results.
+First the loss should be estimated on a test set, after the system has been trained.
+I'm not sure how you estaimate the baseline loss.
+    % DK: TODO, why a test set?
+Isn't the purpose to learn how well the generated captions map to English and not how well they generalize.
+Using only the train set exactly shows if a mapping is possible
     \label{tab:probing_discrimnator_language}
@@
 This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
-
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
-    \label{fig:caption_generator_game_architecture}
-\end{figure}
-
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
-    \label{fig:coordinate_predictor_game_architecture}
-\end{figure}
-
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
-
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
-        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2023-09-01 thesis/sections/1-introduction.tex
comments, feedback
structure and comments

@@
 \label{sec:introduction}
-
+\cmtDK[inline]{lead from grounding to referring expressions to language games}
 The study of emergent languages, in other words artificial languages created by agents to communicate and solve a goal are studied deeply in the recent years \citep{Kirby2002,Kirby2008,Steels2009,Lazaridou2017,Baroni2020,Baroni2022}.
@@
 Following these arguments, also artificial emergent languages need to be grounded that the used symbols contain meaning.
+% SD: I would turn the order of this around, it would make the motivation clearer.
+The introduction should also be much longer with linguistic examples and citations of related work (but I see that you want to keep it for later.)
+% SD: Our goal is to study referring, how we refer to entities and have we interpret the entities referred to. 
+% SD: There is a lot of ambiguity involved as language only maps to the world in an under specified way.
+The first problem is that a word may map to several pixels.
+There is a loss of information/abstraction.
+The second problem is that referring expressions are under-specified, chair can be any of the 5 chairs.
+This is to compress information in communication, we say less than we mean.
+Illustrate both points with examples.
+% SD: Instead, humans rely on communicative protocols to disambiguate referring expressions.
+The Dale and Reiter algorithm and the literature on GRE.
+Communicative protocols are established through language games, some parts seen to be universal, some parts are negotiated on the fly.
+% SD: Artificial language games - describe in more detail what they are and how they are implemented - allow us to study language with8n this communicative setting. 
+% SD: In this thesis we look at referring within the context of communicative games to explore both theoretical and practical (computational) limits of grounded referring expressions in interactive setting.
+% SD: The novelty of this thesis is that we study referring to entities through language games involving sequences of descriptions.
+% SD: Here we need to cite: (1) literature on grounding, connecting language and vision, Harnad, Roy, us, etc.
+(2) referring expression generation, Dale and Reiter, (4) reference and co- reference (Poesio, us), (5) language games (Wittgenstein) and referring as a collaborative process (Clark, David Lewis), (6) language games (work by Kazakov and Kirby), within robotics (Steels), language games within neural models (Baroni).
+% SD: Mathias’ work is also relevant https://era.ed.ac.uk/handle/1842/38727
+% DK: TODO
 
@@
 \begin{enumerate}
-    \item Are cooperative agents in language games able to generate and understand artificial referring expressions based on visual input?
-    \item If a language emerges, how similar are the referring expressions to referring expressions that are used in English?
+      % \item Are cooperative agents in language games able to generate and understand artificial referring expressions based on visual input?
+      \item What are the limits of the agent architectures and input representations on learning successfully grounding referring expressions through language games?
+            % SD: What are the limits of the agent architecture and and input representation on learning successful grounding referring expressions through language games?
+            % DK: (done)
+            % \item If a language emerges, how similar are the referring expressions to referring expressions that are used in English?
+      \item To what degree do emergent referring expressions align with referring expressions in a natural language such as English and what constraints can be imposed on the environment and the agents themselves that languages align?
+            % SD: To what degree does the emergent referring expressions align with referring expressions in a natural language such as English what constraints can be imposed on the environment and the agents themselves that languages align?
+            % DK: (done)
 \end{enumerate}
@@
 The datasets that are created in this thesis aim to reduce this bias and provide all information about the scene and objects in the image.
+% SD: We also know the ground truth about the scenes as we know the function that generated them.
+% DK: TODO
 
@@
 By doing this, it is tested if first the models are able to extract useful visual information that doesn't rely on bias and latent patterns in the dataset and secondly shows the impact of different levels of ambiguity in the datasets on the performance of generating and understanding of referring expressions.
+% SD: We can control the properties of these datasets and we can introduce as much bias and ambiguity as we see firm in each experiment to compare with the natural datasets.
+% DK: TODO
 
 Lastly, the thesis brings the separate tasks of generating and understanding of referring expressions together into on single task.
+% SD: Which is how this is done in practice, cf. referring as a collaborative process, the paper by Clark.
+% DK: TODO
 In language games, one agent needs to extract visual information from an image and generate a referring expression that is sent to the second agent.
@@
 In the present study, only attribute relations are studied.
+% SD: What are those?
+% DK: TODO
 Spatial relations add another level of complexity and may be studied in future work.
+% SD: What kind of complexity?
+% DK: TODO
 Furthermore, the emerged language will be interpreted by comparing it to referring expressions in natural language.
@@
 A deeper study of its compositionality or complexity is out of scope.
+% SD: Leave for the last chapter, limitations and future work.
+% DK: TODO

### 2023-09-01 thesis/sections/2-background.tex
structure and comments

@@
 
+\subsection{Grounding}
+\cmtDK[inline]{importance of grounding and what it is}
+
 \subsection{Referring expressions}
+\cmtDK[inline]{what are referring expressions and how are they studied (generation, understanding, ...)}
+\cmtDK[inline]{incremental GRE algorithm from Dale}
 Even though, recent large language models (LLMs) such as BERT \citep{Devlin2019} or GPT-3 \citep{Brown2020} produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
@@
 \subsection{Language Games}
+\cmtDK[inline]{language games combine generation and understanding}
+\cmtDK[inline]{describe as 'dialogue turns'}
+\cmtDK[inline]{different approaches (robotics, neural models, ...)}
+\cmtDK[inline]{different study interests (linguistically, non-linguistically)}
+\cmtDK[inline]{continuous learning}
 % \begin{itemize}
@@
 \subsection{Artificial dataset}
+\cmtDK[inline]{'world shapes language'}
+\cmtDK[inline]{different types of biases -> reduce and control}

### 2023-09-01 thesis/sections/3-methodology.tex
restructuring
bibliography
structure and comments

@@
 \subsection{Feature extractors}
+\cmtDK[inline]{preprocessing}
 In computer vision tasks, machines need to analyze images and extract information from them.
@@
 
+\subsection{Image processing}
+\cmtDK[inline]{how are feature extractors used in my models}
+\cmtDK[inline]{refer to Johnson2017a}
+\cmtDK[inline]{masking and how are they processed (convolutions to bring it to the same shape)}
+
 \subsection{EGG framework}
@@
 Furthermore, runs of games can be saved to analyze the used messages of the agents and how they vary over the duration of the learning.
-
-The traditional approach to solve problems with a discrete channel relies on the REINFORCE algorithm \citep{Williams1992}.
-It uses the concept of Monte Carlo sampling to estimate the gradient of the expected cumulative reward with respect to the policy parameters.
-The basic idea is to collect trajectories by following the policy, compute the cumulative rewards for each trajectory, and then update the policy parameters to increase the probabilities of actions that led to higher rewards.
-% REINFORCE is easy to implement and understand, but it
-REINFORCE suffers from high variance in gradient estimation due to the inherent randomness in the environment and the policy.
-% SD: :-)
-% DK: (done)
-
-Gumbel-Softmax is a relaxation of the discrete categorical distribution to a continuous distribution that can be differentiated \citep{Jang2016}.
-It allows for the application of gradient-based optimization methods to discrete optimization problems.
-In the Gumbel-Softmax approach, randomness is introduced using the continuous Gumbel distribution.
-Following, the \emph{softmax} operation creates a distribution over discrete actions that can be differentiated.
-Gumbel-Softmax offers stable gradient estimates and can be more efficient than the traditional REINFORCE algorithm, especially when dealing with large action spaces.
-This also applies to language games, where \citet{Havrylov2017} demonstrated that Gumbel-Softmax relaxation is more effective.
-% SD: This is what a technical manual would say - but what does it do really?
-What is the difference between reinforce and Gumbel softmax in practice?
-In our case?
-Reinforce is still superior but here you make it less.
-% DK: I don't think, REINFORCE is superior also in our case, since GS offers more stability in learning (see citing) (TODO, explain GS)
 
@@
 
+\subsection{Optimization in language games}
+The traditional approach to solve problems with a discrete channel relies on the REINFORCE algorithm \citep{Williams1992}.
+It uses the concept of Monte Carlo sampling to estimate the gradient of the expected cumulative reward with respect to the policy parameters.
+The basic idea is to collect trajectories by following the policy, compute the cumulative rewards for each trajectory, and then update the policy parameters to increase the probabilities of actions that led to higher rewards.
+% REINFORCE is easy to implement and understand, but it
+REINFORCE suffers from high variance in gradient estimation due to the inherent randomness in the environment and the policy.
+% SD: :-)
+% DK: (done)
+
+Gumbel-Softmax is a relaxation of the discrete categorical distribution to a continuous distribution that can be differentiated \citep{Jang2017}.
+It allows for the application of gradient-based optimization methods to discrete optimization problems.
+In the Gumbel-Softmax approach, randomness is introduced using the continuous Gumbel distribution.
+Following, the \emph{softmax} operation creates a distribution over discrete actions that can be differentiated.
+Gumbel-Softmax offers stable gradient estimates and can be more efficient than the traditional REINFORCE algorithm, especially when dealing with large action spaces.
+This also applies to language games, where \citet{Havrylov2017} demonstrated that Gumbel-Softmax relaxation is more effective.
+% SD: This is what a technical manual would say - but what does it do really?
+What is the difference between reinforce and Gumbel softmax in practice?
+In our case?
+Reinforce is still superior but here you make it less.
+% DK: I don't think, REINFORCE is superior also in our case, since GS offers more stability in learning (see citing) (TODO, explain GS)
+
+\subsection{Probing?}
+\cmtDK[inline]{reason for usage and techniques fits more into theoretical background}
+
 \subsection{Ethical considerations}

### 2023-09-01 thesis/sections/5-grounding.tex
structure and comments

@@
-\section{Generating and understanding referring expressions}
+\section{Grounding referring expressions}
 \label{sec:preexperiments}
+
+\cmtDK[inline]{general setup: epochs, optimizer, batch size, ...}
 This chapter serves two purposes.
@@
 
+\subsection{Object identification}
+\subsubsection*{Setup}
+\cmtDK[inline]{include how to process bounding boxes}
+\cmtDK[inline]{dot product + discussion about embedding size}
+One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
+Feature extractors like VGG or ResNet are trained on extracting visual information.
+In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not necessarily important for classifying tasks.
+To address this challenge, we restructured the task in this experiment.
+% SD: Just say that this is a task of identifying objects (the previous was about locating them).
+This task does not rely on geometric information but it relies on identifying correct object features, hence explore different features and hence it is useful as abenchmarking task.
+% DK: TODO
+Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects, target object and distractors.
+This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
+To do this, fixed-size bounding boxes are extracted around each of the object in the image.
+As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels to capture the complete object.
+The bounding boxes of all objects are padded to the maximum number of objects present in an image across the whole dataset with a matrix of zeros.
+For the 'Dale-2', this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset
+These bounding boxes are shuffled.
+Furthermore, to point the model towards the target object, its attributes are encoded as on-hot vectors.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
+    \caption{Simplified architecture of the bounding box classifier}
+    \label{fig:bounding_box_classifier_architecture}
+    % SD: I haven't commented on this before, but let us be very careful in naming all meta-paramaters of the models in the text or in caption: how many layers, how many dimensions, etc.
+The reader should be able to replicate the code and the experiment from this.
+You can do this in the caption which is a longer explanation or in the text itself.
+    % DK: TODO
+\end{figure}
+
+The model passes each of the bounding boxes through a linear layer to transfer it into an embedding space of 10 dimensions.
+% SD: This is quite low? Why only 10?
+% DK: to calculate the dot product (inserted; done)
+The same is done with the one-hot vector of the target attributes.
+By bringing both vectors to the same embedding size, the dot product between each embedded image and the embedded attributes can be calculated in the next step.
+A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
+The dot products are concatenated and form a vector with as many dimensions as objects present.
+The \emph{softmax} function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
+% The model then points to an object, using the \emph{softmax} function over this resulting vector.
+The loss is calculated using cross entropy.
+% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
+% DK: the output is a probability distriution over all objects (inserted above; done)
+
+Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
+Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
+Again, the model uses a \emph{softmax} function to point to one of these objects.
+Since no information about the target object is included, the guesses are expected to be random.
+% SD: I don't understand this?
+Have you actually implemented such model?
+But how can you train a random classifier?
+The baseline would normally be 1/5 or 1/5, i.e.
+the number of labels if the number of examples is equal, other a weighted proportion of these.
+By definition of "random" you would expect such classifier to achieve these probabilities.
+% DK: the idea was not to just compare it to the random probabilities, but compare it to a similar setup as the actual model, with the only difference being the missing information of the target object.
+As you say the probability is 1/5 (resp.
+1/2), which is also the result of thi model.
+Better to just compare to random probability?
+(QUESTION)
+
+The bounding box classifier is trained on all datasets that inlcudes more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
+% excluding the 'CLEVR single'.
+% SD: Why not?
+% DK: because there is only one object in the input, and therefore only one possible guess. The model can't point to a distractor (adressed; done)
+The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
+Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
+
+\subsubsection*{Results}
+As expected, all baselines achieve a random accuracy for each of the datasets.
+The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
+This changes however, when information about the target object is included.
+Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rcc}
+        \toprule
+                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
+        \textbf{Dale-2}      & 50\%                       & 99\%              \\
+        \textbf{Dale-5}      & 22\%                       & 94\%              \\
+        \textbf{CLEVR color} & 12\%                       & 93\%              \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the bounding box classifier after 10 epochs}
+    \label{tab:results_bounding_box_classifier}
+\end{table}
+
+The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
+For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
+First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
+This worsens slightly, when more objects are introduced.
+More objects lead to more needed attributes to discriminate the objects from each other.
+The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
+The model can generate these high results, even with its relatively simple architecture.
+In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
+A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
+% SD: The model is able to leatn to discriminate objects based on visual appearance.
+% DK: TODO
+
+\subsection{Referring expression generation}
+\label{sec:referring_expression_generation}
+\subsubsection*{Setup}
+\cmtDK[inline]{new metrics}
+
+As the bounding box classifier, the caption generator acts as a learning step towards understanding, how the model can learn the attributes of objects.
+Additionally, this setup is a natural language generation task, where the model is trained to generate referring expressions, based on visual input.
+% SD: Well, this experiment is a natural language generation task: whether the model is able to generate referring expressions.
+% DK: (addressed; done)
+The same method using the GRE-algorithm from the section above is used to create the descriptions of the target object for each image.
+
+There are some minor additions concerning the padding of the caption.
+As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
+However, there are three different ways how the padding is applied.
+First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
+A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
+More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
+The second last token if existing describes the color, while the third last token if existing describes the size.
+As soon as this sequence is padded at the end, these slots suddenly disappear.
+A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
+Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
+For this reason, the second method of padding the caption is prepending the caption with padding tokens.
+By this, the positions of the slots are preserved and if not specified just filled with a padding token.
+Each slot has always the same semantic value, e.g. the last slot always contains the shape of an object.
+This can be done since the captions are not free text, but instead the structure and the possible content is given by the dataset.
+This method might help the model to learn the correct referring expressions.
+% SD: Slots have the same semantic value, i.e.
+the last slot is the object type.
+We can do this because we know the structure of the descriptions in this dataset and because we hope this will help the learning.
+Also, we are thinking of generation based on slots.
+% DK: (addressed; done)
+The last variation concerns the order of producing each token.
+When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
+This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
+Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the caption.
+% SD: We... I sounds more like you are writing a lab report with casual language but here you want to be more formal.
+% DK: (done)
+Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
+Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
+
+The task for the model is to generate these descriptions, based on the complete scene as input.
+Doing this, the task inherently involves learning human knowledge and natural language structure.
+Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
+Can the model solve the task, or do specific attributes used by humans pose challenges to the model?
+% Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
+% SD: In this case it should be using the same attributes as humans as we are training it on human labels.
+% DK: (rephrased; done)
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
+    \caption{Simplified architecture of the masked caption generator}
+    \label{fig:caption_generator_architecture}
+\end{figure}
+
+We compare two different models against each other.
+The first model just takes the image as the input.
+During training, also the ground truth caption is passed to the model.
+The image is processed as in the section before.
+First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
+After each convolutional layer the \emph{ReLU} function is applied.
+The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
+The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
+% SD: Dimensions?
+% DK: 1024? (QUESTION)
+During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
+% SD: Why is this teacher forcing?
+% DK: because the ground truth during generation of the sequence (added subclause; done)
+The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
+The loss is calculated using cross-entropy.
+During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
+Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.
+% SD: There is a greedy decoding algorithm.
+% DK: (added; done)
+
+Using this approach, the model doesn't have any information about the target object.
+Therefore, we extend it in a second step with attention to the target object.
+% SD: You add attention to the target object.
+% DK: (added; done)
+A masked version of the image is created ass in the section above and passed to the model.
+This is shown in Figure \ref{fig:caption_generator_architecture}.
+The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
+% SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
+% DK: (QUESTION)
+Both encodings are concatenated and then used as the initial hidden state for the LSTM.
+This should point the model towards which object to describe and discriminate from the distractors.
+
+To test the success of the model, three measures are calculated.
+The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
+% SD: This is too strict.
+if this is a generation task then we can use generation metrics.
+Those compare if the generated string as a whole is kind of similar to the target string, see ROUGE, METEOR, BERTscore.
+% If this is a classification task, and it is, given that we know what type of word we expect in each slot, we could calculate accuracy per class, precision, recall, F-score.
+We could also calculate mean ranks for predicting words per classes, or acucracy at k- if the target word is in the k-top predictions.
+% We could also take it a multi-class classification problem and how to calculate accuracy in this case, would need to check.
+% DK: I agree that classification metrics refelct success better here (therefore I also used accuracy).
+These metrics would give a better overview, but I don't think there is time to implement it (QUESTION)
+This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
+However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
+
+It could be the case that the model predicted the correct shape, but wrong color.
+Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
+For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
+% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
+% DK: TODO
+The correct caption is \emph{cylinder}.
+If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
+The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
+To mitigate this, we also include a \textbf{word-by-word accuracy}.
+This measure calculates the accuracy not on sentence level, but on word level.
+The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would yield 0\%.
+% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
+% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
+This can give a better understanding of the errors the model makes.
+
+% SD: Very good If I understand this right this would be a case where the generated description describes another object better than the target and a human would identify that.
+So it is like false negatives in describing.
+Perhaps use another term for this: referring to a distractor?
+(since you are not measuring hwo individually the description fitts a distractor but only if the entire description could perfectly fit a distractor)
+% DK: exactly. I added false negatives. But I would like to keep accuracy in the name, since that is what it is. maybe better distractor accuracy? (QUESTION)
+With the \textbf{non-target accuracy}, we identify if the model described another object, which is not the target image.
+This is basically an inverted accuracy score; the lower the score, the better the model fares.
+By this it measures the false negative generated captions.
+For this we generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
+If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
+If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
+Using this measure, we can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
+
+The caption generator models are trained on both 'Dale' datasets.
+Again, each of these datasets increases the complexity of the description.
+While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
+Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
+
+\subsubsection*{Results}
+In the next paragraphs, the results of the caption generator are discussed.
+Table \ref{tab:results_caption-generator} shows the different scores of the models when trained on 'Dale-2' and 'Dale-5'.
+ResNet-3 produced the best results, when it was used as a feature extractor.
+The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rccc}
+        \toprule
+                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
+        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
+        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
+        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
+        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of the caption generator, based on ResNet-3}
+    % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1.
+It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
+    % DK: might be the case.
+in the last language game i use only a linear layer instead of convolutions, but it doesn't help.
+This however might also be due to other problems.
+Time?
+(QUESTION, future work)
+    % SD: I would present these for individual classes.
+    % DK: I only have accuracy scores at. Time? (QUESTION)
+    \label{tab:results_caption-generator}
+\end{table}
+
+The sum of the two columns for the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
+% SD: Not shown in Table 5
+% DK: sum of two columns (addressed; done)
+This means that the models produce almost always correct descriptions of one of the objects in the image.
+Moreover, these descriptions are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
+Features of both objects are therefore extracted perfectly and associated with the vocabulary.
+% SD: But this is no-surprise since we have fixed the slots.
+% DK: but this applies to all padding methods. (done)
+As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
+With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
+This changes, when also the masked image is presented to the model.
+Then, the accuracy for the target object rises with 72\% well above chance.
+Still, the wrong descriptions describe almost every time the distractor object.
+This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
+
+When the model is trained on the 'Dale-5' dataset, the results are much worse.
+Now, without masking only in 61\% of the images, the model describes any of the objects in the image, whether it is the target object or a distractor.
+% SD: ?
+% DK: added subclause (done)
+With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
+Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
+The accuracy stays at the same value.
+In contrast, the non-target accuracy is increased by 5\% points.
+The model is more likely to identify a wrong object.
+Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
+This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
+For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
+When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
+This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
+% SD: Not in the table?
+% DK: no, because a result of one specific sample (done)
+
+The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
+When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
+The final peak stayed exactly the same and the effects were therefore not studied deeper.
+% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
+% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
+
+There are two main possible explanations for the big difference between the two datasets.
+First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
+In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
+% SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated.
+I would only pose this as a very vague hypothesis to be tested, as you conclude now it is also the case that images with simialr distractors are more difficult to describe correctly.
+Hence, it is by no way given.
+% DK: TODO
+The second explanation lies in the used GRE-algorithm.
+When only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
+Otherwise, the shape is enough, and the caption is only one word long.
+The probability for this lies at 66,6\%.
+% SD: Why 66? We have 3 attributes, right, so 33%?
+% DK: but that the other attributes are only needed if the shape is the same. The prob. for the shape being different for the second object is 66% (rephrased; done)
+For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
+Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
+With four distractors the algorithm is much more likely to produce longer captions.
+These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
+% SD: I think we need individual accuracies per class and also precision, recall and the f-score.
+If we include the joint accuracy figure when you also need to explain how you avergaed the infividual accuracies.
+% DK: time? (QUESTION)
+% SD: Not sure. Note that <pad> is a token. The model has to learn when to say <pad> just as any word, hence knowing when not to dewscribe is also trickly.
+% DK: true, but <pad> token is probably the most used token in general and therefore a guess of <pad> just by frequency is likely right? (QUESTION)
 \subsection{Reference resolution}
@@
 \subsubsection*{Setup}
+\cmtDK[inline]{attribute encodings, coordinate predictor}
 
@@
 % DK: TODO (future work)
-
-\subsection{Object identification}
-\subsubsection*{Setup}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not necessarily important for classifying tasks.
-To address this challenge, we restructured the task in this experiment.
-% SD: Just say that this is a task of identifying objects (the previous was about locating them).
-This task does not rely on geometric information but it relies on identifying correct object features, hence explore different features and hence it is useful as abenchmarking task.
-% DK: TODO
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects, target object and distractors.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels to capture the complete object.
-The bounding boxes of all objects are padded to the maximum number of objects present in an image across the whole dataset with a matrix of zeros.
-For the 'Dale-2', this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset
-These bounding boxes are shuffled.
-Furthermore, to point the model towards the target object, its attributes are encoded as on-hot vectors.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-    % SD: I haven't commented on this before, but let us be very careful in naming all meta-paramaters of the models in the text or in caption: how many layers, how many dimensions, etc.
-The reader should be able to replicate the code and the experiment from this.
-You can do this in the caption which is a longer explanation or in the text itself.
-    % DK: TODO
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to transfer it into an embedding space of 10 dimensions.
-% SD: This is quite low? Why only 10?
-% DK: to calculate the dot product (inserted; done)
-The same is done with the one-hot vector of the target attributes.
-By bringing both vectors to the same embedding size, the dot product between each embedded image and the embedded attributes can be calculated in the next step.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The dot products are concatenated and form a vector with as many dimensions as objects present.
-The \emph{softmax} function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
-% The model then points to an object, using the \emph{softmax} function over this resulting vector.
-The loss is calculated using cross entropy.
-% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
-% DK: the output is a probability distriution over all objects (inserted above; done)
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-% SD: I don't understand this?
-Have you actually implemented such model?
-But how can you train a random classifier?
-The baseline would normally be 1/5 or 1/5, i.e.
-the number of labels if the number of examples is equal, other a weighted proportion of these.
-By definition of "random" you would expect such classifier to achieve these probabilities.
-% DK: the idea was not to just compare it to the random probabilities, but compare it to a similar setup as the actual model, with the only difference being the missing information of the target object.
-As you say the probability is 1/5 (resp.
-1/2), which is also the result of thi model.
-Better to just compare to random probability?
-(QUESTION)
-
-The bounding box classifier is trained on all datasets that inlcudes more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
-% excluding the 'CLEVR single'.
-% SD: Why not?
-% DK: because there is only one object in the input, and therefore only one possible guess. The model can't point to a distractor (adressed; done)
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even with its relatively simple architecture.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-% SD: The model is able to leatn to discriminate objects based on visual appearance.
-% DK: TODO
-
-\subsection{Referring expression generation}
-\label{sec:referring_expression_generation}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator acts as a learning step towards understanding, how the model can learn the attributes of objects.
-Additionally, this setup is a natural language generation task, where the model is trained to generate referring expressions, based on visual input.
-% SD: Well, this experiment is a natural language generation task: whether the model is able to generate referring expressions.
-% DK: (addressed; done)
-The same method using the GRE-algorithm from the section above is used to create the descriptions of the target object for each image.
-
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-Each slot has always the same semantic value, e.g. the last slot always contains the shape of an object.
-This can be done since the captions are not free text, but instead the structure and the possible content is given by the dataset.
-This method might help the model to learn the correct referring expressions.
-% SD: Slots have the same semantic value, i.e.
-the last slot is the object type.
-We can do this because we know the structure of the descriptions in this dataset and because we hope this will help the learning.
-Also, we are thinking of generation based on slots.
-% DK: (addressed; done)
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the caption.
-% SD: We... I sounds more like you are writing a lab report with casual language but here you want to be more formal.
-% DK: (done)
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-The task for the model is to generate these descriptions, based on the complete scene as input.
-Doing this, the task inherently involves learning human knowledge and natural language structure.
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Can the model solve the task, or do specific attributes used by humans pose challenges to the model?
-% Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
-% SD: In this case it should be using the same attributes as humans as we are training it on human labels.
-% DK: (rephrased; done)
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-We compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-% SD: Dimensions?
-% DK: 1024? (QUESTION)
-During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
-% SD: Why is this teacher forcing?
-% DK: because the ground truth during generation of the sequence (added subclause; done)
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross-entropy.
-During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.
-% SD: There is a greedy decoding algorithm.
-% DK: (added; done)
-
-Using this approach, the model doesn't have any information about the target object.
-Therefore, we extend it in a second step with attention to the target object.
-% SD: You add attention to the target object.
-% DK: (added; done)
-A masked version of the image is created ass in the section above and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-% SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
-% DK: (QUESTION)
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model towards which object to describe and discriminate from the distractors.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-% SD: This is too strict.
-if this is a generation task then we can use generation metrics.
-Those compare if the generated string as a whole is kind of similar to the target string, see ROUGE, METEOR, BERTscore.
-% If this is a classification task, and it is, given that we know what type of word we expect in each slot, we could calculate accuracy per class, precision, recall, F-score.
-We could also calculate mean ranks for predicting words per classes, or acucracy at k- if the target word is in the k-top predictions.
-% We could also take it a multi-class classification problem and how to calculate accuracy in this case, would need to check.
-% DK: I agree that classification metrics refelct success better here (therefore I also used accuracy).
-These metrics would give a better overview, but I don't think there is time to implement it (QUESTION)
-This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
-% DK: TODO
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
-To mitigate this, we also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would yield 0\%.
-% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
-% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
-This can give a better understanding of the errors the model makes.
-
-% SD: Very good If I understand this right this would be a case where the generated description describes another object better than the target and a human would identify that.
-So it is like false negatives in describing.
-Perhaps use another term for this: referring to a distractor?
-(since you are not measuring hwo individually the description fitts a distractor but only if the entire description could perfectly fit a distractor)
-% DK: exactly. I added false negatives. But I would like to keep accuracy in the name, since that is what it is. maybe better distractor accuracy? (QUESTION)
-With the \textbf{non-target accuracy}, we identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-By this it measures the false negative generated captions.
-For this we generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, we can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator are discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1.
-It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
-    % DK: might be the case.
-in the last language game i use only a linear layer instead of convolutions, but it doesn't help.
-This however might also be due to other problems.
-Time?
-(QUESTION, future work)
-    % SD: I would present these for individual classes.
-    % DK: I only have accuracy scores at. Time? (QUESTION)
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the two columns for the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-% SD: Not shown in Table 5
-% DK: sum of two columns (addressed; done)
-This means that the models produce almost always correct descriptions of one of the objects in the image.
-Moreover, these descriptions are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and associated with the vocabulary.
-% SD: But this is no-surprise since we have fixed the slots.
-% DK: but this applies to all padding methods. (done)
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object rises with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image, whether it is the target object or a distractor.
-% SD: ?
-% DK: added subclause (done)
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-% SD: Not in the table?
-% DK: no, because a result of one specific sample (done)
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
-The final peak stayed exactly the same and the effects were therefore not studied deeper.
-% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
-% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-% SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated.
-I would only pose this as a very vague hypothesis to be tested, as you conclude now it is also the case that images with simialr distractors are more difficult to describe correctly.
-Hence, it is by no way given.
-% DK: TODO
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
-Otherwise, the shape is enough, and the caption is only one word long.
-The probability for this lies at 66,6\%.
-% SD: Why 66? We have 3 attributes, right, so 33%?
-% DK: but that the other attributes are only needed if the shape is the same. The prob. for the shape being different for the second object is 66% (rephrased; done)
-For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
-% SD: I think we need individual accuracies per class and also precision, recall and the f-score.
-If we include the joint accuracy figure when you also need to explain how you avergaed the infividual accuracies.
-% DK: time? (QUESTION)
-% SD: Not sure. Note that <pad> is a token. The model has to learn when to say <pad> just as any word, hence knowing when not to dewscribe is also trickly.
-% DK: true, but <pad> token is probably the most used token in general and therefore a guess of <pad> just by frequency is likely right? (QUESTION)

### 2023-09-02 thesis/sections/3-methodology.tex
comments

@@
 These are also used in this thesis as a first layer to extract important information from the image. Hereby, two different architectures are tested.
+\cmtDK[inline]{explain advantage of CNN}
 
@@
                             & \textbf{description}                                            & \textbf{ output dimensions } \\\midrule
-        \textbf{VGG-0}      & contains only the convolutional layers                          & 512x7x7                      \\
-        \textbf{VGG-avg}    & contains an additional average pooling layer                    & 512x7x7                      \\
+        \textbf{VGG-0}      & contains only the convolutional layers                          & 512 \times\ 7 \times\ 7      \\
+        \textbf{VGG-avg}    & contains an additional average pooling layer                    & 512 \times\ 7 \times\ 7      \\
         \textbf{VGG-cls1}   & \makecell[cl]{contains an additional one classification layer,                                 \\ including its non-linearity} & 4069                         \\
@@
         \textbf{VGG-cls3}   & the original VGG19 architecture                                 & 1000                         \\\midrule
-        \textbf{ResNet-1}   & contains one residual block                                     & 256x56x56                    \\
-        \textbf{ResNet-2}   & contains two residual blocks                                    & 512x28x28                    \\
-        \textbf{ResNet-3}   & contains three residual blocks                                  & 1024x14x14                   \\
-        \textbf{ResNet-4}   & contains four residual blocks                                   & 2048x1x1                     \\
-        \textbf{ResNet-avg} & contains an additional average pooling layer                    & 2048x1x1                     \\
+        \textbf{ResNet-1}   & contains one residual block                                     & 256 \times\ 56 \times\ 56    \\
+        \textbf{ResNet-2}   & contains two residual blocks                                    & 512 \times\ 28 \times\ 28    \\
+        \textbf{ResNet-3}   & contains three residual blocks                                  & 1024 \times\ 14 \times\ 14   \\
+        \textbf{ResNet-4}   & contains four residual blocks                                   & 2048 \times\ 1 \times\ 1     \\
+        \textbf{ResNet-avg} & contains an additional average pooling layer                    & 2048 \times\ 1 \times\ 1     \\
         \textbf{ResNet-cls} & the original ResNet-101 architecture                            & 1000                         \\
@@
 \subsection{Image processing}
-\cmtDK[inline]{how are feature extractors used in my models}
-\cmtDK[inline]{refer to Johnson2017a}
 \cmtDK[inline]{masking and how are they processed (convolutions to bring it to the same shape)}
+
+The above described feature extractors are not enough to be solely used to encode the images used in this thesis.
+Even though final layers that contain very task and domain specific knowledge can be removed, the previous layers still don't contain any information about the new domain and task.
+For this reason they need to be extened with further layers that are not frozen and can learn and store information about the new domain by connecting the pretrained general knowledge with the specific task.
+In this thesis, we use one architecture to encode images for all the conducted experiment.
+The additional layers on top of the feature extractors however are always trained from scratch for each of the experiments and tasks.
+By doing this, the models are made up of multiple smaller modules, as for example the image processing module or other modules that encode different inputs.
+These modules can be added and combined step by step, whilst the architecture of each module always stays the same.
+Following, the experiments stay comparable, since differences are only added or removed modules.
+
+\citet{Johnson2017} describe an architecture that was used for training baseline models on the CLEVR dataset.
+This architecture will be used for all experiments.
+Hereby, the image is first passed through a frozen feature extractor.
+Two convolutional networks with subsequent \emph{ReLU} non-linearities condense the important information from the output of the feature extractor.
+The convolutional layers reduce the channels to 128 channels, using a kernel size of 3 and a stride and padding of 1.
+Afterwards, a 2-dimensional max pooling is applied with a kernel size and stride of 2.
+Finally, the resulting matrix is flattened and passed through a linear layer to reduce it to an encoding size.
+This vector represents the encoded image with its extracted features.
+Table \ref{tab:image_encoder} shows the layers and their output dimensions when using ResNet-3 as the feature extractor and an encoding size of 2048 dimensions.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        \textbf{layer}                      & \textbf{output dimensions} \\\midrule
+        Input image                         & 3 \times\ 224 \times\ 224  \\
+        ResNet-3                            & 1024 \times\ 14 \times\ 14 \\
+        Conv(3 \times\ 3, \rightarrow\ 128) & 128 \times\ 14 \times\ 14  \\
+        ReLU                                & 128 \times\ 14 \times\ 14  \\
+        Conv(3 \times\ 3, \rightarrow\ 128) & 128 \times\ 14 \times\ 14  \\
+        ReLU                                & 128 \times\ 14 \times\ 14  \\
+        MaxPool(2 \times\ 2, stride 2)      & 128 \times\ 7 \times\ 7    \\
+        Flatten                             & 6272                       \\
+        Linear Layer(\rightarrow\ 2048)     & 2048                       \\
+        \bottomrule
+    \end{tabular}
+    \caption{Image encoder with ResNet-3 and an encoding size of 2048}
+    \label{tab:image_encoder}
+\end{table}
 

### 2023-09-02 source/mlt/shared_models.py

@@ -1,7 +1,7 @@
 from abc import ABC
 
 import torch
-from mlt.feature_extractors import FeatureExtractor
+from mlt.feature_extractors import DummyFeatureExtractor, FeatureExtractor
 from torch import nn
 
 
@@ -18,18 +18,21 @@ class ClevrImageEncoder(ImageEncoder):
         super().__init__()
         self.process_image = nn.Sequential(
             feature_extractor,
-            nn.LazyConv2d(128, kernel_size=1, padding=0),
+            nn.LazyConv2d(128, kernel_size=3, padding=1),
             nn.ReLU(),
-            nn.LazyConv2d(128, kernel_size=1, padding=0),
+            nn.LazyConv2d(128, kernel_size=3, padding=1),
             nn.ReLU(),
             nn.MaxPool2d(kernel_size=2, stride=2),
         )
-        self.mean_reduction = nn.LazyLinear(encoder_out_dim)
+        self.reduction = nn.Sequential(nn.Flatten(), nn.LazyLinear(encoder_out_dim))
 
     def forward(self, image):
         processed_image = self.process_image(image)
-        flattened = torch.flatten(processed_image, start_dim=2).permute(0, 2, 1)
-        reduced = self.mean_reduction(flattened.mean(dim=1))
+
+        # flattened = torch.flatten(processed_image, start_dim=2).permute(0, 2, 1)
+        # reduced = self.mean_reduction(flattened.mean(dim=1))
+        reduced = self.reduction(processed_image)
+
         return reduced
 
 

### 2023-10-26 thesis/sections/3-methodology.tex
bugfix, thesis

@@
 \subsection{Feature extractors}
+\label{sec:feature-extractors}
 \cmtDK[inline]{preprocessing}
@@
 Even though final layers that contain very task and domain specific knowledge can be removed, the previous layers still don't contain any information about the new domain and task.
-For this reason they need to be extened with further layers that are not frozen and can learn and store information about the new domain by connecting the pretrained general knowledge with the specific task.
+For this reason they need to be extended with further layers that are not frozen and can learn and store information about the new domain by connecting the pretrained general knowledge with the specific task.
 In this thesis, we use one architecture to encode images for all the conducted experiment.

### 2023-10-26 thesis/sections/5-grounding.tex
bugfix, thesis

@@
 \subsubsection*{Setup}
-\cmtDK[inline]{include how to process bounding boxes}
-\cmtDK[inline]{dot product + discussion about embedding size}
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not necessarily important for classifying tasks.
-To address this challenge, we restructured the task in this experiment.
-% SD: Just say that this is a task of identifying objects (the previous was about locating them).
-This task does not rely on geometric information but it relies on identifying correct object features, hence explore different features and hence it is useful as abenchmarking task.
-% DK: TODO
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects, target object and distractors.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels to capture the complete object.
-The bounding boxes of all objects are padded to the maximum number of objects present in an image across the whole dataset with a matrix of zeros.
-For the 'Dale-2', this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset
-These bounding boxes are shuffled.
-Furthermore, to point the model towards the target object, its attributes are encoded as on-hot vectors.
+
+In a first task a neural model is trained to discriminate multiple objects from each other.
+Hereby, the model is shown bounding boxes of all objects in the scene as well as a description of the target object.
+The model needs to combine all information and then point to the correct bounding box.
+This task forces the model to extract visual features from the objects in the shown inputs and connect them to a representation of human given attributes.
+Since the model is shown only bounding boxes around each object rather than the whole scene, geometric information about the objects' locations doesn't play a role for the success in this task.
+
+In a first step, bounding boxes are extracted from each scene.
+Each bounding box is a square with a side length of 96 pixels around the center coordinates of each object.
+By doing this large objects in the front of the scene fill the bounding box completely, while smaller objects in the back of the scene contain some space around them.
+They also might contain noise in the form of parts of adjacent objects.
+Preliminary experiments indicated however that a more complex method for extracting the bounding boxes, depending on spatial position in the scene to reduce the noise didn't improve the results.
+Therefore, bounding boxes in this thesis rely on the former simpler approach.
+Since, each bounding box will be passed through on of the feature extractors, described in section \ref{sec:feature-extractors} they need to be normalized first.
+As described in \citep{He2016,Simonyan2015}, the bounding boxes are resized to 256 pixels for the shorter side, then cropped around the center to a square of 224 \times\ 224 pixels.
+Finally, the RGB channels are normalized, by subtracting the means (0,485; 0,456; 0,406) and dividing the result by the standard deviation (0,229; 0,224; 0,225) for each channel respectively.
+The array of bounding boxes is padded with matrix of zeros to the maximum possible number of objects present in a scene across the dataset.
+For the 'Dale-2' dataset, this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset.
+For each sample in the dataset the bounding boxes are shuffled.
+
+The attributes of the target object are encoded as one-hot encodings.
+There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the \emph{color} and a two-dimensional vector encoding the two different \emph{sizes}.
+The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
+These three encodings of the attributes are then concatenated to a vector with 13 dimensions.
 
@@
     \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
+    \caption{Simplified architecture of the object identification task}
     \label{fig:bounding_box_classifier_architecture}
@@
 
-The model passes each of the bounding boxes through a linear layer to transfer it into an embedding space of 10 dimensions.
+The model is split into two parts.
+First, each normalized bounding box of the sample is passed through \emph{ResNet-avg}.
+The preceding layers \emph{ResNet-1} to \emph{ResNet-4} couldn't be compared due to memory restrictions on the GPU.
+A deeper analysis of the effects of different feature extractors will be done in the following chapters.
+The resulting vector is flattened and projected to the embedding dimension $e$ via a linear layer.
 % SD: This is quite low? Why only 10?
-% DK: to calculate the dot product (inserted; done)
-The same is done with the one-hot vector of the target attributes.
-By bringing both vectors to the same embedding size, the dot product between each embedded image and the embedded attributes can be calculated in the next step.
+% DK: different embedding sizes are tested (done)
+Correspondingly, the one-hot vector of the attributes is projected to the same embedding dimension with another linear layer.
+In the second part, each bounding box is combined with the representation of the attributes.
+By bringing both vectors to the same embedding size, the dot product between each embedded bounding box and the embedded attributes can be calculated.
 A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
+The model can so learn to connect the object representation with the attributes.
 The dot products are concatenated and form a vector with as many dimensions as objects present.
@@
 % The model then points to an object, using the \emph{softmax} function over this resulting vector.
+
+The experiments are conducted with the following hyperparameters: a learning rate of 0,0002, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
+8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
 The loss is calculated using cross entropy.
+The following embedding dimension $e$ are compared: 2, 10, 50, 100, 500, 1000.
 % SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
@@
 
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-% SD: I don't understand this?
-Have you actually implemented such model?
-But how can you train a random classifier?
-The baseline would normally be 1/5 or 1/5, i.e.
-the number of labels if the number of examples is equal, other a weighted proportion of these.
-By definition of "random" you would expect such classifier to achieve these probabilities.
-% DK: the idea was not to just compare it to the random probabilities, but compare it to a similar setup as the actual model, with the only difference being the missing information of the target object.
-As you say the probability is 1/5 (resp.
-1/2), which is also the result of thi model.
-Better to just compare to random probability?
-(QUESTION)
-
-The bounding box classifier is trained on all datasets that inlcudes more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
+The object identification task is trained on all datasets that include more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
+Random baselines for each dataset would yield accuracies of 10\%, 50\% and 20\% respectively.
 % excluding the 'CLEVR single'.
@@
 
-As the bounding box classifier, the caption generator acts as a learning step towards understanding, how the model can learn the attributes of objects.
-Additionally, this setup is a natural language generation task, where the model is trained to generate referring expressions, based on visual input.
-% SD: Well, this experiment is a natural language generation task: whether the model is able to generate referring expressions.
-% DK: (addressed; done)
-The same method using the GRE-algorithm from the section above is used to create the descriptions of the target object for each image.
+Opposed to the previous experiment, where the focus lied on extracting visual features, the model is now tasked to generate referring expressions.
+This is done in two setups.
+The first setup uses bounding boxes as in the previous section as input, where the model needs to describe the target object of one of the bounding boxes.
+In the second setup, the model is presented with the complete image and therefore also needs to accommodate to geometric information of the scene.
 
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
+The bounding boxes are extracted and normalized in the same way as for the object identification task.
+Since for the second setup, the images will also be passed through the feature extractors, they are normalized as well in like manner.
+The referring expressions for the target object are generated using the incremental GRE-algorithm, described in section \ref{GRE}.
+By this, the model needs to describe the target object with respect to the distractor objects.
+There are some minor additions concerning the padding of the referring expression.
+As before, the referring expression is padded to a number of three tokens, corresponding to the maximum of three attributes.
 However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
+First, the referring expression are, as usual in captioning tasks padded at the end with a specified padding token.
+A problem could arise when the referring expression is not viewed as a natural language sentence, but as slots filled with tokens.
+More specifically, following the GRE-algorithm, the last token in the referring expression is always the shape.
 The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
+As soon as this sequence is padded at the end, these slots disappear.
+A referring expression that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
+Since this task is not focussing on producing natural language with a correct grammar, but focuses instead on extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
+For this reason, the second method of padding the referring expression is prepending the referring expression with padding tokens.
 By this, the positions of the slots are preserved and if not specified just filled with a padding token.
 Each slot has always the same semantic value, e.g. the last slot always contains the shape of an object.
-This can be done since the captions are not free text, but instead the structure and the possible content is given by the dataset.
+This can be done since the referring expressions are not free text, but instead the structure and the possible content is given by the dataset.
 This method might help the model to learn the correct referring expressions.
@@
 The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
+When the referring expressions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
 This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the caption.
+Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the referring expression.
 % SD: We... I sounds more like you are writing a lab report with casual language but here you want to be more formal.
@@
 Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
+Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the referring expression are still preserved.
 
-The task for the model is to generate these descriptions, based on the complete scene as input.
-Doing this, the task inherently involves learning human knowledge and natural language structure.
+This task inherently involves learning human knowledge and natural language structure.
 Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
@@
 % DK: (rephrased; done)
+
+% TODO include figure
+
+In the first setup, the model receives the bounding boxes as input.
+The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
+As in the previous experiment, each bounding box is passed through \emph{ResNet-avg} and projected afterwards to an image embedding dimension $e_i$ with a linear layer.
+All thus encoded bounding boxes are concatenated and again compressed to the decoder output dimension $d$ using another linear layer.
+This representation of all objects serves as the initial hidden state of an LSTM, which generates the referring expression.
+Tokens used in the LSTM are embedded with embedding dimension $e_t$.
+During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
+The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
+The loss is calculated using cross-entropy.
+During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
+Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.
 

### 2023-10-26 source/mlt/preexperiments/data_readers.py

@@ -593,7 +593,7 @@ class BoundingBoxCaptioningDataset(Dataset, Persistable):
                 "caption", data=torch.stack([sample.caption for sample in self.samples])
             )
             f.create_dataset(
-                "non_target_caption",
+                "non_target_captions",
                 data=torch.stack(
                     [sample.non_target_caption for sample in self.samples]
                 ),
@@ -615,7 +615,7 @@ class BoundingBoxCaptioningDataset(Dataset, Persistable):
                                 "bounding_boxes",
                                 "image_id",
                                 "caption",
-                                "non_target_caption",
+                                "non_target_captions",
                             ],
                             sample,
                         )
@@ -625,7 +625,7 @@ class BoundingBoxCaptioningDataset(Dataset, Persistable):
                     [torch.from_numpy(b) for b in f["bounding_boxes"]],
                     [str(i, "utf-8") for i in f["image_id"]],
                     [torch.from_numpy(c) for c in f["caption"]],
-                    [torch.from_numpy(n) for n in f["non_target_caption"]],
+                    [torch.from_numpy(n) for n in f["non_target_captions"]],
                 )
             ]
             captioner = pickle.loads(f.attrs["captioner"].tobytes())

### 2023-10-26 source/mlt/preexperiments/data_readers.py
bugfix

@@ -595,7 +595,7 @@ class BoundingBoxCaptioningDataset(Dataset, Persistable):
             f.create_dataset(
                 "non_target_captions",
                 data=torch.stack(
-                    [sample.non_target_caption for sample in self.samples]
+                    [sample.non_target_captions for sample in self.samples]
                 ),
             )
             f.create_dataset(

### 2023-10-26 source/mlt/preexperiments/run_experiments.py
bugfix

@@ -2,24 +2,24 @@ import itertools
 import subprocess
 
 options = [
-    # "--dataset_base_dir=/scratch/guskunkdo",
-    "--dataset_base_dir=/home/dominik/Development/",
+    "--dataset_base_dir=/scratch/guskunkdo",
+    # "--dataset_base_dir=/home/dominik/Development/",
     "--epochs=30",
     "--lr=0.0002",
     "--feature_file=bounding_box_resnet_4_avgpool_no-fc.h5",
     "--max_samples=10000",
     "--device=cuda",
-    "--model=bounding_box_classifier_attributes",
-    # "--out=/scratch/guskunkdo/out/",
+    "--model=bounding_box_caption_generator",
+    "--out=/scratch/guskunkdo/out/",
 ]
 
 variables = {
-    "--dataset": ["dale-2", "dale-5", "colour"],
-    "--image_embedding_dimension": [2, 10, 50, 100, 500, 1000],
+    "--dataset": ["colour"],
+    "--image_embedding_dimension": [100, 500, 1000],
     # "--coordinate_classifier_dimension": [1024, 2048],
-    # "--decoder_out_dim": [500, 1000, 1500, 2000],
+    "--decoder_out_dim": [100, 500, 1000],
     # "--encoder_out_dim": [500, 1000, 1500, 2000],
-    # "--embedding_dim": [15, 30],
+    "--embedding_dim": [10, 15, 30],
 }
 
 for index, combination in enumerate(itertools.product(*variables.values())):

### 2023-10-26 thesis/sections/5-grounding.tex
thesis

@@
 \subsection{Object identification}
+\label{sec:object-identification}
 \subsubsection*{Setup}
@@
 
-The experiments are conducted with the following hyperparameters: a learning rate of 0,0002, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
 The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
@@
 The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross-entropy.
 During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
@@
 
-We compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-% SD: Dimensions?
-% DK: 1024? (QUESTION)
-During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
-% SD: Why is this teacher forcing?
-% DK: because the ground truth during generation of the sequence (added subclause; done)
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross-entropy.
-During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.
-% SD: There is a greedy decoding algorithm.
-% DK: (added; done)
+In the second setup two different models are compared against each other.
+The first model acts as the baseline and only receives the complete image as the input.
+The image is passed through \emph{ResNet-3} and subsequently processed by additional convolution layers, described in section \ref{sec:image-processing}.
+The resulting vector is then reduced to decoder output dimension $d$ by a linear layer and as for the bounding boxes acts as the initial hidden state for the LSTM.
+The LSTM is trained in the same way as above.
 
-Using this approach, the model doesn't have any information about the target object.
-Therefore, we extend it in a second step with attention to the target object.
+Using this approach, the model doesn't have any information about the target object and is therefore expected to produce referring expressions for a randomly selected object on the scene.
+Therefore, it is extended in a second step with attention to the target object.
 % SD: You add attention to the target object.
 % DK: (added; done)
-A masked version of the image is created ass in the section above and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
+A masked version of the image is created and passed to the model.
+For this, a fixed size squared area of 96 \times\ 96 pixels around the center of the target object is filled in white, while the rest of the image is filled in black.
+Both original and masked image are processed together with convolutional layers as described in section \ref{sec:image-processing}.
 % SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
 % DK: (QUESTION)
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
+The result is then again projected to the decoder output dimension $d$ by a linear layer and then used as the initial hidden state for the LSTM.
 This should point the model towards which object to describe and discriminate from the distractors.
 
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
+For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $e_i$ (only for bounding boxes): 100, 500, 1000
+    \item $d$: 10, 15, 30
+    \item $e_t$: 100, 500, 1000
+\end{itemize}
+
+As already discussed before, this task can be interpreted as a classification task rather than a natural language generation task.
+The main reason for this is that the model is tasked to assign specific attributes to the target object instead of producing free text with a large vocabulary.
+Following, we are interested in the classification mistakes the model makes.
+For this, the model's success is validated on accuracy, recall and precision scores.
+These are calculated in the following ways.
+The first measure is the \textbf{overall accuracy} if the model predicted every word in the caption correctly.
 % SD: This is too strict.
@@
 The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
-To mitigate this, we also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would yield 0\%.
+To mitigate this, the accuracies for each class are included, as well as the macro average.
 % SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
@@
 This can give a better understanding of the errors the model makes.
+The same is done for \emph{precision} and \emph{recall}.
 
@@
 If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, we can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
+Using this measure, we can get an overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
 
@@
 
+With the reference resolution task, the model is trained to understand referring expressions, by pointing towards the target object in the scene.
 This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
@@
 The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
+Including human referring expressions into the task would bias also the emerged language towards attributes and words, used in natural language.
 For this reason, the final task of the receiver will be to 'point' to the target object.
@@
 % DK: TODO
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
+With this approach, the models receive few natural language information, but are still able to rely on all information present in the image to discriminate the objects.
 
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+To achieve this goal, multiple setups of models are trained.
+In the simplest setup, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
 % SD: What are the features?
@@
 % DK: TODO
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
+In a first step, the image is encoded.
+This is done by first passing it through one of the \emph{feature extractors} and then further processing it with convolutional layers as described in section \ref{sec:image-processing}.
+In the next step the encoded image is used to predict coordinates using the \emph{coordinate predictor}.
+Hereby, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
 These reduce the dimensions first to 1024 and finally to 2.
@@
 Therefore, the models need to receive more information.
-Here, we try out four different ways to encode and refer to the target object.
+Here, four different ways to encode and refer to the target object are tested.
 
@@
 
-In the first method, we encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, we also include the \textbf{center coordinates of all objects} in the image.
+In the first method, the target object is encoded as \textbf{one-hot encodings}, as described in section \ref{sec:object-identification}.
+In an extension to this method, also the \textbf{center coordinates of all objects} in the image are included.
 % SD: Encoded locations?
@@
 The padded locations consist of two zeros for both coordinates.
-For this model, we also made use of a more complex way to encode the image.
-Here, we based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
 % SD: Why shuffled? If we order them the way they appear in the image, the model will have more information that they are sequentially related.
 % DK: TODO
-This is then passed again through a final linear layer to predict x- and y-coordinates.
 % SD: Convolutions are only applied on visual features, not object attribute features and locations.
@@
 
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citep{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
+The third method encodes the attributes of the target object with human language using the \textbf{incremental GRE-algorithm}.
+This opposes the idea described before, to share as few natural language information as possible with the model.
 Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
@@
 
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-% SD: See my point earlier about the reversed algorithm between the two tasks 
-% DK: TODO
+% Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
+% In other words, the object is described unambiguously using the lowest number of words.
+% The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
+% This order defines, which attributes can be left out, while still identifying the object uniquely.
+% This research relies on the following order from most important to least important: shape, color, size
+% Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
+% Using all three attributes, this description identifies the object perfectly and uniquely.
+% Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
+% This can be taken even one step further by removing also the \emph{color}.
+% Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
+% % SD: See my point earlier about the reversed algorithm between the two tasks 
+% % DK: TODO
 
@@
 
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
+In the model, the referring expression is encoded, using an LSTM.
 Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citep{Johnson2017}.
+Both the processed image and the final hidden state of the LSTM are flattened and concatenated, which is then passed through the coordinate predictor.
 % SD: \citep and \citet Use Name (year) when you are referring to particular person and (Name, year) when you are referring to a paper. Hence, in this case it would be the latter.
@@
 % DK: done
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
+The image is masked in the same way as in section \ref{sec:referring_expression_generation}.
 While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
+It therefore can only rely on its own extracted visual features and the inherent human bias in the image when looking at masked images.
 % SD: The information in the masked image is still not enough to identify the precise location, only the region.
@@
 % DK: TODO
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
+Both original and masked image are processed as described in \ref{sec:image-processing} and afterwards passed through the coordinate predictor.
+
+For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lcccc}
+        \toprule
+                               & $e_i$              & $c$               & $LSTM_e$  & $LSTM_o$               \\
+                               & $[100, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[500,1000,1500,2000]$ \\\midrule
+        Coordinate predictor   & \times             & \times            & -         & -                      \\
+        + one-hot              & \times             & \times            & -         & -                      \\
+        + one-hot + locations  & \times             & \times            & -         & -                      \\
+        + referring expression & \times             & \times            & \times    & \times                 \\
+        + masking              & \times             & \times            & -         & -                      \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
+    \label{tab:feature-extractor-mean-distances}
+    % SD: Explanation of labels
+    % DK: TODO
+\end{table}
 
@@
 
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
+The reference resolution models are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
 The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.

### 2023-10-27 thesis/sections/5-grounding.tex
describe experimental setups

@@
 
-\cmtDK[inline]{general setup: epochs, optimizer, batch size, ...}
 This chapter serves two purposes.
@@
 
-The model is split into two parts.
+The model, the \emph{object identifier} is split into two parts.
 First, each normalized bounding box of the sample is passed through \emph{ResNet-avg}.
@@
 
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
 The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
@@
 
-In the first setup, the model receives the bounding boxes as input.
+In the first setup, the \emph{bounding box RE generator}, the model receives the bounding boxes as input.
 The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
@@
 In the second setup two different models are compared against each other.
-The first model acts as the baseline and only receives the complete image as the input.
+The first model, the \emph{basic RE generator} acts as the baseline and only receives the complete image as the input.
 The image is passed through \emph{ResNet-3} and subsequently processed by additional convolution layers, described in section \ref{sec:image-processing}.
@@
 Using this approach, the model doesn't have any information about the target object and is therefore expected to produce referring expressions for a randomly selected object on the scene.
-Therefore, it is extended in a second step with attention to the target object.
+Therefore, it is extended in a second step with attention to the target object, the \emph{masked RE generator}.
 % SD: You add attention to the target object.
@@
 
-For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
 8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
@@
     \item $e_i$ (only for bounding boxes): 100, 500, 1000
-    \item $d$: 10, 15, 30
-    \item $e_t$: 100, 500, 1000
+    \item $d$: 100, 500, 1000
+    \item $e_t$: 10, 15, 30
 \end{itemize}
@@
 To achieve this goal, multiple setups of models are trained.
-In the simplest setup, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+In the simplest setup, the \emph{reference resolver}, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
 % SD: What are the features?
@@
 This is done by first passing it through one of the \emph{feature extractors} and then further processing it with convolutional layers as described in section \ref{sec:image-processing}.
+Finally, it is projected to the image embedding size $e_i$.
 In the next step the encoded image is used to predict coordinates using the \emph{coordinate predictor}.
 Hereby, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
+These reduce the dimensions first to the coordinate predictor dimension $c$ and finally to 2.
 
@@
 The padded locations consist of two zeros for both coordinates.
+For both methods, the encodings of the target object are concatenated with the encoded image and then passed to the coordinate predictor.
 % SD: Why shuffled? If we order them the way they appear in the image, the model will have more information that they are sequentially related.
@@
 Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
+Tokens are embedded with embedding dimensions $LSTM_e$ and the output of the LSTM has the size if $LSTM_o$.
+The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
 Both the processed image and the final hidden state of the LSTM are flattened and concatenated, which is then passed through the coordinate predictor.
@@
 
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
+The fourth method, to encode the target object utilizes \textbf{masking} of the image.
 % SD: Oh, before we talked about 3 methods
@@
 
-For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, Adam \citep{Kingma2015} as optimizer.
+For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
 8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
 The loss is calculated using cross entropy.
-The following values for the variables are compared:
+Table \ref{tab:variables-reference-resolution} shows the variables that are changed during the experiments for each of the models.
 
@@
                                & $[100, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[500,1000,1500,2000]$ \\\midrule
-        Coordinate predictor   & \times             & \times            & -         & -                      \\
+        reference resolver     & \times             & \times            & -         & -                      \\
         + one-hot              & \times             & \times            & -         & -                      \\
@@
     \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-    % SD: Explanation of labels
-    % DK: TODO
+    \caption{Variables for each model where $e_i$ is the image embedding dimension, $c$ the dimensions of the coordinate predictor, $LSTM_e$ the embedding dimensions for tokens in the LSTM and $LSTM_o$ the output dimension of the LSTM}
+    \label{tab:variables-reference-resolution}
 \end{table}

### 2023-10-27 thesis/sections/6-grounding-language-games.tex
structure and comments
describe experimental setups

@@
-\section{Language games}
+\section{Grounding referring expressions in language games}
 \label{sec:language-games}
+\cmtDK[inline]{'epochs', training/testing (success), optimizer, ...}
+\cmtDK[inline]{max message length}
 
@@
 
+\cmtDK[inline]{discuss training/epochs/batches, vocab size, message length, gs temperature}
+
 \subsection{Object identification}
@@
 
+The object identification game is set up as a discrimination game.
 In a discrimination game, the agents are presented with two or more images, one of these being the target image.
@@
 
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citep{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-% SD: Here it does not receive a boundig box of an object as in Lazaridou. They use whole images, don't they?
-% DK: TODO
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
+The data is prepared in the same way as for the single model object identification task in section \ref{sec:object-identification}.
+Bounding boxes of all objects in the scene are extracted and preprocessed to be used for the feature extractors.
+Accordingly, also the models from the single model tasks serve as basis for the object identification task.
+The sender uses the architecture of the model described in section \ref{sec:referring_expression_generation} apart from the final LSTM.
+Instead, it will use the LSTM with the hidden size $h_s$ included in the \emph{EGG} framework to produce a message.
+Based on the results from the single model object identifiers, the image embedding size $e_s$ is fixed to 100.
+The bounding box of the target object is passed first, the remaining are shuffled.
 
-\begin{figure}[h]
+The receiver instead uses the architecture of the single model object identifier, described in section \ref{sec:object-identification}.
+Similarly to the sender, the LSTM of the \emph{EGG} framework with the hidden size $h_e$ is used to encode the sender's message instead of the one-hot encoded attributes in the single model.
+The hidden state of the LSTM is projected to the embedding size $e_r=500$ with a linear layer, based on the results in the previous experiments.
+As before the dot product between the embedded message and the encoded image is calculated to let the model point towards one of the bounding boxes.
+Opposed to the sender, the receiver is shown all bounding boxes in shuffled order.
+Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
+
+\begin{figure}[ht]
     \centering
@@
 
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-% SD: Deatils about the layers, their parameters. More explanatory figure caption.
-% DK: TODO
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $h_s$: 100, 500, 1000
+    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $|V|$: 1, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 5, 6
+\end{itemize}
+The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
 
@@
 
-\cmtDK{similarity to bounding box classifier}
+\cmtDK[inline]{similarity to bounding box classifier}
 
 \subsection{Referring expression generation}
+\label{sec:referring-expression-generation-game}
 \subsubsection*{Setup}
@@
 The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
+\cmtDK[inline]{human bias vs emergent language}
 
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-% SD: The sender uses the incremental algorithm?
-% DK: TODO, QUESTION
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
+For this experiment, the data is prepared in the same way as described in section \ref{sec:referring_expression_generation}.
+Since the position of the padding as well as the order of the words didn't have an effect on the results, the usual approach in natural generation task is used: padding tokens are appended and the referring expression is not reversed.
+Respectively, both agents are based on the models in these experiments.
+The sender uses the \emph{masked RE generator} with its LSTM replaced with the LSTM included in the \emph{EGG} framework, using the hidden size $h_s$.
+\cmtDK[inline]{fixed parameters}
+The receiver is built up similarly.
+The image is encoded in the same way, but afterwards it is concatenated with the decoded message, received be the sender with the hidden size $h_r$.
+The concatenated is reduced to $d=$ dimensions, based on the results of the previous experiments, and used as the initial state of the captioning LSTM.
+\cmtDK[inline]{fixed parameters}
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citep{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-% SD: Reference back to when you reported this?
-% DK: TODO
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-% SD: But receiver is not doing captioning? Is it incrementally identifying the objects as the sender's messages are coming in?
-% DK: TODO
 During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
@@
 When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
 
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $h_s$: 100, 500, 1000
+    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $|V|$: 1, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 5, 6
+\end{itemize}
+\cmtDK[inline]{real values}
 Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
@@
 % DK: TODO
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
+The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results.
 
 \subsubsection*{Results}
-\begin{table}[h]
+\begin{table}[ht]
     \centering
@@
 \subsubsection*{Setup}
-% SD: But previously we have said that we are not going to study this game. But good that we have!
-% DK: where? (QUESTION)
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-% SD: Not true: human knowledge is still reflected in the way the environment is structured.
-% DK: TODO
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-% SD: I'm not following this
-% DK: TODO
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-% SD: Which contains human knowledge!
-% DK: TODO
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
+In the final experiments, the agents are tasked to produced and understand referring expressions only based on the implicit human knowledge present in the scenes as for example the attributes of the objects.
+Opposed to the previous experiments, explicit human language information as human referring expressions or one-hot encoded attributes are avoided.
+By this, the agents can't just reuse these human referring expressions, but need to learn to generate them themselves during the language games just based on the visual input.
+This is done similarly to the single model reference resolution task, described in section \ref{sec:reference_resolution}.
+The agents are tasked to predict the center coordinates of the target object, after being presented with the complete scene.
+However, the agents, more specifically the sender needs to be pointed towards the target object.
+In the final setup, the \emph{masked reference resolution game}, this is done providing the sender with an additional masked visual input.
+However, this can prove to be a challenging task for the agents.
+For this reason, in an approach to reduce the complexity, for the \emph{RE reference resolution game} the sender is shown a human referring expression, generated with the incremental GRE algorithm instead of the additional masked image.
+This might lead to a heavy bias of the emerged language towards the human referring expressions and will be compared to the unbiased emerged language.
 
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
+The masked sender is the same as the \emph{RE generation sender}, described in section \ref{sec:referring-expression-generation-game}.
+The RE sender works similarly, but instead encodes the human referring expression with an LSTM.
+The final hidden state is concatenated with the encoded image.
+The resulting vector is reduced to the hidden size $h_s$ and used as the initial hidden state of the LSTM included in the \emph{EGG} framework to generate the message.
+\cmtDK[inline]{fixed variables}
+
+The receiver is the same in both setups and is based on the \emph{reference resolver} + referring expressions, described in section \ref{sec:reference_resolution}.
+Again, instead of the included LSTM, the receiver uses the \emph{EGG} LSTM with the hidden size $h_r$ to encode the message.
+The dimensions of the coordinate predictor are fixed to $c=$, based on the previous results.
+\cmtDK[inline]{fixed variables}
 % SD: Just thinking: would there be a more systematic way to compare all the models?
@@
 % DK: TODO
+The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $h_s$: 100, 500, 1000
+    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $|V|$: 1, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 5, 6
+\end{itemize}
+\cmtDK[inline]{real values}
 
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
+As in section \ref{sec:reference_resolution}, the agents are trained first on the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
 In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
+The same metrics are used to evaluate the results.
 

### 2023-10-27 source/mlt/evaluate_results.py

@@ -24,6 +24,9 @@ if __name__ == "__main__":
             for line in f:
                 if line.startswith("{"):
                     epoch = json.loads(line)
+                    if epoch.get("mode", "test") == "train":
+                        continue
+
                     flattened = {}
                     for k, v in epoch.items():
                         if isinstance(v, dict):
@@ -31,9 +34,12 @@ if __name__ == "__main__":
                                 flattened[f"{k}_{sub_k}"] = sub_v
                         else:
                             flattened[k] = v
+
                     results.append(flattened)
-            metrics.update(results[0].keys())
-            all_results[name] = results
+
+            if len(results) != 0:
+                metrics.update(results[0].keys())
+                all_results[name] = results
 
     sorted_metrics = list(sorted(metrics))
     print(list(enumerate(sorted_metrics)))

### 2023-10-29 thesis/sections/5-grounding.tex
results

@@
     \centering
-    \begin{tabular}{rcc}
+    \begin{tabular}{c|c|c|c}
         \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
+               & \textbf{Dale-2}   & \textbf{Dale-5}   & \textbf{CLEVR Color} \\\cmidrule(lr){2-2}\cmidrule(lr){3-3}\cmidrule(lr){4-4}
+        $e$    & \textbf{Accuracy} & \textbf{Accuracy} & \textbf{Accuracy}    \\\midrule
+        {2}    & {92\%}            & {72\%}            & {40\%}               \\
+        {10}   & \textbf{100\%}    & {94\%}            & {92\%}               \\
+        {50}   & {99\%}            & \textbf{95\%}     & \textbf{94\%}        \\
+        {100}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
+        {500}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
+        {1000} & \textbf{100\%}    & {94\%}            & \textbf{94\%}        \\
         \bottomrule
     \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
+    \caption{Accuracy scores of the object identifier: $e$ are different embedding sizes}
     \label{tab:results_bounding_box_classifier}
@@
 The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
-As in the previous experiment, each bounding box is passed through \emph{ResNet-avg} and projected afterwards to an image embedding dimension $e_i$ with a linear layer.
-All thus encoded bounding boxes are concatenated and again compressed to the decoder output dimension $d$ using another linear layer.
+As in the previous experiment, each bounding box is passed through \emph{ResNet-avg} and projected afterwards to an image embedding dimension $e$ with a linear layer.
+All thus encoded bounding boxes are concatenated and again compressed to the decoder output dimension $LSTM_o$ using another linear layer.
 This representation of all objects serves as the initial hidden state of an LSTM, which generates the referring expression.
-Tokens used in the LSTM are embedded with embedding dimension $e_t$.
+Tokens used in the LSTM are embedded with embedding dimension $LSTM_e$.
 During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
@@
 The image is passed through \emph{ResNet-3} and subsequently processed by additional convolution layers, described in section \ref{sec:image-processing}.
-The resulting vector is then reduced to decoder output dimension $d$ by a linear layer and as for the bounding boxes acts as the initial hidden state for the LSTM.
+The resulting vector is then reduced to decoder output dimension $LSTM_o$ by a linear layer and as for the bounding boxes acts as the initial hidden state for the LSTM.
 The LSTM is trained in the same way as above.
@@
 \begin{itemize}
-    \item $e_i$ (only for bounding boxes): 100, 500, 1000
-    \item $d$: 100, 500, 1000
-    \item $e_t$: 10, 15, 30
+    \item $e$ (only for bounding boxes): 100, 500, 1000
+    \item $LSMT_o$: 100, 500, 1000
+    \item $LSTM_e$: 10, 15, 30
 \end{itemize}
@@
 This can give a better understanding of the errors the model makes.
-The same is done for \emph{precision} and \emph{recall}.
+The same is done for \textbf{precision} and \textbf{recall}.
 
@@
     \centering
-    \begin{tabular}{rccc}
+    \begin{tabular}{ccc|ccc|ccc|ccc}
         \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
+               &          &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                                 \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
+        $e$    & $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}     & \textbf{NT}  & \textbf{Acc.} & \textbf{F1}   & \textbf{NT}  \\\midrule
+        {100}  & {100}    & {10}     & {97\%}                              & {97,5\%}                            & {1\%}                                    & {65\%}        & {88\%}          & {2\%}        & {92\%}        & {81\%}        & {3\%}        \\
+        {100}  & {100}    & {15}     & {97\%}                              & {98\%}                              & {0\%}                                    & {62\%}        & {86,47\%}       & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
+        {100}  & {100}    & {30}     & {98\%}                              & {98\%}                              & {1\%}                                    & {65\%}        & {87,99\%}       & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
+        {100}  & {500}    & {10}     & {98\%}                              & {98,5\%}                            & {0\%}                                    & {67\%}        & {88\%}          & {2\%}        & {93\%}        & {82\%}        & {4\%}        \\
+        {100}  & {500}    & {30}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
+        {100}  & {1000}   & {10}     & {98\%}                              & {98\%}                              & {0\%}                                    & {68\%}        & {88\%}          & {2\%}        & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
+        {500}  & {500}    & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & {67\%}        & {88,99\%}       & {2\%}        & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
+        {500}  & {1000}   & {30}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & {69\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
+        {1000} & {100}    & {10}     & {97\%}                              & {97,99\%}                           & {0\%}                                    & {59\%}        & {85,5\%}        & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
+        {1000} & {100}    & {30}     & {98\%}                              & {98,5\%}                            & {0\%}                                    & {61\%}        & {87\%}          & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
+        {1000} & {1000}   & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & {93\%}        & {82\%}        & {4\%}        \\
+        {1000} & {1000}   & {15}     & {99\%}                              & {98,5\%}                            & {0\%}                                    & {69\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
         \bottomrule
     \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \label{tab:results_bounding-box-caption-generator}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|ccc|ccc|ccc}
+        \toprule
+                 &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                         \\\cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1} & \textbf{NT} & \textbf{Acc.} & \textbf{F1} & \textbf{NT} \\\midrule
+        {100}    & {10}     & {\%}                                & {\%}                                & {\%}                                     & {65\%}        & {88\%}      & {2\%}       & {92\%}        & {81\%}      & {3\%}       \\
+        {100}    & {15}     & {78\%}                              & {71,99\%}                           & {11\%}                                   & {62\%}        & {86,47\%}   & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
+        {100}    & {30}     & {68\%}                              & {42,79\%}                           & {4\%}                                    & {65\%}        & {87,99\%}   & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
+        {500}    & {10}     & {83\%}                              & {79,95\%}                           & {6\%}                                    & {67\%}        & {88\%}      & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
+        {500}    & {15}     & {83\%}                              & {79,5\%}                            & {7\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
+        {500}    & {30}     & {84\%}                              & {83,43\%}                           & {5\%}                                    & {68\%}        & {88\%}      & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
+        {1000}   & {10}     & {84\%}                              & {82,42\%}                           & {5\%}                                    & {67\%}        & {88,99\%}   & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
+        {1000}   & {15}     & {85\%}                              & {82,42\%}                           & {5\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
+        {1000}   & {30}     & {85\%}                              & {82,81\%}                           & {4\%}                                    & {59\%}        & {85,5\%}    & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
+        {1500}   & {10}     & {84\%}                              & {81,42\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
+        {1500}   & {15}     & \textbf{85\%}                       & \textbf{84,43\%}                    & \textbf{4\%}                             & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
+        {1500}   & {30}     & {84\%}                              & {80,95\%}                           & {4\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
+        {2000}   & {10}     & {83\%}                              & {82,89\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
+        {2000}   & {15}     & {84\%}                              & {81,89\%}                           & {4\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
+        {2000}   & {30}     & {85\%}                              & {82,95\%}                           & {4\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
+        {3000}   & {10}     & {82\%}                              & {79,89\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
+        {3000}   & {15}     & {85\%}                              & {81,89\%}                           & {4\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
+        {3000}   & {30}     & {83\%}                              & {80,89\%}                           & {3\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
+        \bottomrule
+    \end{tabular}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \label{tab:results_masked-caption-generator}
     % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1.
@@
     % DK: I only have accuracy scores at. Time? (QUESTION)
-    \label{tab:results_caption-generator}
 \end{table}

### 2023-10-29 source/mlt/print_results.py

@@ -1,0 +1,60 @@
+import argparse
+import glob
+import json
+import os
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser()
+
+    parser.add_argument(
+        "--folder",
+        type=str,
+        help="path to the results folders (can contain wildcards)",
+    )
+
+    args = parser.parse_args()
+    print(args)
+
+    metrics = set()
+    all_results = {}
+    for folder in glob.glob(args.folder):
+        name = folder.split("/")[-1]
+        with open(os.path.join(folder, "log.txt"), "r", encoding="utf-8") as f:
+            results = []
+            for line in f:
+                if line.startswith("{"):
+                    epoch = json.loads(line)
+                    if epoch.get("mode", "test") == "train":
+                        continue
+
+                    flattened = {}
+                    for k, v in epoch.items():
+                        if isinstance(v, dict):
+                            for sub_k, sub_v in v.items():
+                                flattened[f"{k}_{sub_k}"] = sub_v
+                        else:
+                            flattened[k] = v
+
+                    results.append(flattened)
+
+            if len(results) != 0:
+                metrics.update(results[0].keys())
+                all_results[name] = results
+
+    sorted_metrics = list(sorted(metrics))
+    print(list(enumerate(sorted_metrics)))
+
+    choice = input("choose important metrics (separate by comma):")
+    chosen_metrics = [sorted_metrics[int(c)] for c in choice.split(",")]
+
+    last_runs = {run_name: runs[-1] for run_name, runs in all_results.items()}
+
+    sorted_runs = sorted(
+        last_runs.items(),
+        key=lambda x: [x[1][metric] for metric in chosen_metrics],
+        reverse=True,
+    )
+
+    print(chosen_metrics)
+    for run_name, run_metrics in sorted_runs:
+        print(run_name, [run_metrics[metric] for metric in chosen_metrics])

### 2023-10-31 source/mlt/print_results.py
evaluation

@@ -2,6 +2,32 @@ import argparse
 import glob
 import json
 import os
+import pprint
+import re
+from dataclasses import dataclass
+
+
+@dataclass
+class Experiment:
+    folder_name: str
+    variable_string: str
+    dataset: str
+    variables: dict
+    runs: list[dict]
+    last_run: dict
+
+
+def sort_by_variable(runs):
+    return dict(
+        sorted(
+            runs.items(),
+            key=lambda x: x[1][1],
+            reverse=False,
+        )
+    )
+
+
+sorts = {"metric": sort_by_metric, "variable": sort_by_variable}
 
 if __name__ == "__main__":
     parser = argparse.ArgumentParser()
@@ -12,15 +38,52 @@ if __name__ == "__main__":
         help="path to the results folders (can contain wildcards)",
     )
 
+    parser.add_argument(
+        "--variables",
+        nargs="+",
+        help="variables in the same order as in the folder name",
+    )
+
+    parser.add_argument(
+        "--variable_order",
+        nargs="+",
+        type=int,
+        help="variables in the same order as in the folder name",
+    )
+
+    parser.add_argument(
+        "--sort",
+        choices=sorts.keys(),
+        help="sorting of the results",
+    )
+
     args = parser.parse_args()
+    if len(args.variable_order) == 0:
+        args.variable_order = range(len(args.variables))
+    if len(args.variable_order) != len(args.variables):
+        raise ValueError(
+            f"Number of variables and orders don't match: {args.variable_order}, {args.variables}"
+        )
     print(args)
 
-    metrics = set()
+    experiments: list[Experiment] = []
     all_results = {}
     for folder in glob.glob(args.folder):
-        name = folder.split("/")[-1]
+        folder_name = folder.split("/")[-1]
+
+        pattern = r"_([^_]+)((_\d+)*)$"
+        match = re.search(pattern, folder_name)
+        if match:
+            dataset = match.group(1)
+            variables = [int(var) for var in match.group(2).split("_") if var != ""]
+            if len(variables) != len(args.variables):
+                raise ValueError(f"Number of Variables don't match: {folder_name}")
+            ordered_variables = [variables[o] for o in args.variable_order]
+
+        variable_string = "_".join(map(str, ordered_variables))
+
         with open(os.path.join(folder, "log.txt"), "r", encoding="utf-8") as f:
-            results = []
+            runs = []
             for line in f:
                 if line.startswith("{"):
[…]

### 2023-11-01 thesis/sections/5-grounding.tex
evaluation

@@
 
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|ccc|ccc|ccc}
+        \toprule
+                                        &                           & \multicolumn{3}{c}{\textbf{dale-5}} & \multicolumn{3}{c}{\textbf{dale-2}} & \multicolumn{3}{c}{\textbf{colour}}                                                                                                                                                             \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        coordinate-classifier-dimension & image-embedding-dimension & \textbf{accuracy}                   & \textbf{word-by-word-f1}            & \textbf{non-target-accuracy}        & \textbf{accuracy} & \textbf{word-by-word-f1} & \textbf{non-target-accuracy} & \textbf{accuracy} & \textbf{word-by-word-f1} & \textbf{non-target-accuracy} \\\midrule
+        {10}                            & {1500}                    & {33\%}                              & {67,47\%}                           & {14\%}                              & {84\%}            & {81,42\%}                & {4\%}                        & {12\%}            & {33,43\%}                & {34\%}                       \\
+        {10}                            & {2000}                    & {36\%}                              & {69\%}                              & {13\%}                              & {83\%}            & {82,89\%}                & {4\%}                        & {13\%}            & {36\%}                   & {34\%}                       \\
+        {10}                            & {3000}                    & {34\%}                              & {68\%}                              & {12\%}                              & {82\%}            & {79,89\%}                & {4\%}                        & {12\%}            & {32,31\%}                & {34\%}                       \\
+        {15}                            & {1500}                    & {35\%}                              & {68,99\%}                           & {12\%}                              & {85\%}            & {84,43\%}                & {4\%}                        & {17\%}            & {38,49\%}                & {31\%}                       \\
+        {15}                            & {2000}                    & {34\%}                              & {67\%}                              & {14\%}                              & {84\%}            & {81,89\%}                & {4\%}                        & {14\%}            & {33,31\%}                & {33\%}                       \\
+        {15}                            & {3000}                    & {32\%}                              & {65,5\%}                            & {15\%}                              & {85\%}            & {81,89\%}                & {4\%}                        & {13\%}            & {32,73\%}                & {35\%}                       \\
+        {30}                            & {1500}                    & {36\%}                              & {68,99\%}                           & {15\%}                              & {84\%}            & {80,95\%}                & {4\%}                        & {14\%}            & {36,97\%}                & {33\%}                       \\
+        {30}                            & {2000}                    & {32\%}                              & {65,5\%}                            & {16\%}                              & {85\%}            & {82,95\%}                & {4\%}                        & {18\%}            & {39\%}                   & {32\%}                       \\
+        {30}                            & {3000}                    & {3\%}                               & {63,5\%}                            & {16\%}                              & {83\%}            & {80,89\%}                & {3\%}                        & {12\%}            & {32,73\%}                & {34\%}                       \\
+        {100}                           & {10}                      & {21\%}                              & {57\%}                              & {23\%}                              & {16\%}            & {37\%}                   & {31\%}                                                                                                     \\
+        {100}                           & {15}                      & {17\%}                              & {52,98\%}                           & {25\%}                              & {78\%}            & {71,99\%}                & {11\%}                       & {26\%}            & {43,49\%}                & {3\%}                        \\
+        {100}                           & {30}                      & {19\%}                              & {54,46\%}                           & {24\%}                              & {68\%}            & {42,79\%}                & {4\%}                        & {29\%}            & {45,49\%}                & {26\%}                       \\
+        {500}                           & {10}                      & {27\%}                              & {61,98\%}                           & {16\%}                              & {83\%}            & {79,95\%}                & {6\%}                        & {17\%}            & {38,49\%}                & {32\%}                       \\
+        {500}                           & {15}                      & {28\%}                              & {63,5\%}                            & {16\%}                              & {83\%}            & {79,5\%}                 & {7\%}                        & {21\%}            & {4\%}                    & {31\%}                       \\
+        {500}                           & {30}                      & {3\%}                               & {63,98\%}                           & {17\%}                              & {84\%}            & {83,43\%}                & {5\%}                        & {21\%}            & {4\%}                    & {31\%}                       \\
+        {1000}                          & {10}                      & {31\%}                              & {65,5\%}                            & {13\%}                              & {84\%}            & {82,42\%}                & {5\%}                        & {12\%}            & {32,12\%}                & {34\%}                       \\
+        {1000}                          & {15}                      & {31\%}                              & {64,5\%}                            & {16\%}                              & {85\%}            & {82,42\%}                & {5\%}                        & {16\%}            & {37\%}                   & {32\%}                       \\
+        {1000}                          & {30}                      & {31\%}                              & {63,98\%}                           & {17\%}                              & {85\%}            & {82,81\%}                & {4\%}                        & {13\%}            & {36\%}                   & {34\%}                       \\
+        \bottomrule
+    \end{tabular}
+    \caption{test}
+    \label{test}
+\end{table}
 The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.

### 2023-11-01 source/mlt/print_results.py

@@ -2,7 +2,6 @@ import argparse
 import glob
 import json
 import os
-import pprint
 import re
 from dataclasses import dataclass
 
@@ -10,24 +9,31 @@ from dataclasses import dataclass
 @dataclass
 class Experiment:
     folder_name: str
-    variable_string: str
     dataset: str
-    variables: dict
+    variables: tuple
     runs: list[dict]
-    last_run: dict
+
+    @property
+    def last_run(self):
+        return self.runs[-1]
 
 
-def sort_by_variable(runs):
-    return dict(
-        sorted(
-            runs.items(),
-            key=lambda x: x[1][1],
-            reverse=False,
+def calculate_f1(run):
+    if "word_by_word_precision" in run.keys() and "word_by_word_recall" in run.keys():
+        run["word_by_word_f1"] = (
+            2
+            * (run["word_by_word_precision"] * run["word_by_word_recall"])
+            / (run["word_by_word_precision"] + run["word_by_word_recall"])
         )
-    )
+
+    return run
 
 
-sorts = {"metric": sort_by_metric, "variable": sort_by_variable}
+def calculate_metrics(run):
+    run = calculate_f1(run)
+
+    return run
+
 
 if __name__ == "__main__":
     parser = argparse.ArgumentParser()
@@ -37,24 +43,22 @@ if __name__ == "__main__":
         type=str,
         help="path to the results folders (can contain wildcards)",
     )
-
     parser.add_argument(
         "--variables",
         nargs="+",
         help="variables in the same order as in the folder name",
     )
-
     parser.add_argument(
         "--variable_order",
         nargs="+",
         type=int,
-        help="variables in the same order as in the folder name",
+        help="order of the chosen variables",
     )
-
     parser.add_argument(
-        "--sort",
-        choices=sorts.keys(),
-        help="sorting of the results",
+        "--datasets",
+        nargs="+",
+        type=str,
+        help="datasets in order in which they should appear in the table",
     )
 
     args = parser.parse_args()
@@ -78,9 +82,7 @@ if __name__ == "__main__":
             variables = [int(var) for var in match.group(2).split("_") if var != ""]
             if len(variables) != len(args.variables):
                 raise ValueError(f"Number of Variables don't match: {folder_name}")
-            ordered_variables = [variables[o] for o in args.variable_order]
-
-        variable_string = "_".join(map(str, ordered_variables))
+            ordered_variables = tuple(variables[o] for o in args.variable_order)
 
[…]

### 2023-11-02 thesis/sections/5-grounding.tex
results

--- 2023-10-29 thesis/sections/5-grounding.tex
@@
         {1000} & {1000}   & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & {93\%}        & {82\%}        & {4\%}        \\
-        {1000} & {1000}   & {15}     & {99\%}                              & {98,5\%}                            & {0\%}                                    & {69\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
+        {1000} & {1000}   & {15}     & {99\%}                              & {98,5\%}                            & {0\%}                                    & {68\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
         \bottomrule
@@
         \toprule
-                 &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                         \\\cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+                 &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                         \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
         $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1} & \textbf{NT} & \textbf{Acc.} & \textbf{F1} & \textbf{NT} \\\midrule
-        {100}    & {10}     & {\%}                                & {\%}                                & {\%}                                     & {65\%}        & {88\%}      & {2\%}       & {92\%}        & {81\%}      & {3\%}       \\
-        {100}    & {15}     & {78\%}                              & {71,99\%}                           & {11\%}                                   & {62\%}        & {86,47\%}   & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
-        {100}    & {30}     & {68\%}                              & {42,79\%}                           & {4\%}                                    & {65\%}        & {87,99\%}   & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
-        {500}    & {10}     & {83\%}                              & {79,95\%}                           & {6\%}                                    & {67\%}        & {88\%}      & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
-        {500}    & {15}     & {83\%}                              & {79,5\%}                            & {7\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
-        {500}    & {30}     & {84\%}                              & {83,43\%}                           & {5\%}                                    & {68\%}        & {88\%}      & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
-        {1000}   & {10}     & {84\%}                              & {82,42\%}                           & {5\%}                                    & {67\%}        & {88,99\%}   & {2\%}       & {93\%}        & {82\%}      & {3\%}       \\
-        {1000}   & {15}     & {85\%}                              & {82,42\%}                           & {5\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
-        {1000}   & {30}     & {85\%}                              & {82,81\%}                           & {4\%}                                    & {59\%}        & {85,5\%}    & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
-        {1500}   & {10}     & {84\%}                              & {81,42\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
-        {1500}   & {15}     & \textbf{85\%}                       & \textbf{84,43\%}                    & \textbf{4\%}                             & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
-        {1500}   & {30}     & {84\%}                              & {80,95\%}                           & {4\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
-        {2000}   & {10}     & {83\%}                              & {82,89\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
-        {2000}   & {15}     & {84\%}                              & {81,89\%}                           & {4\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
-        {2000}   & {30}     & {85\%}                              & {82,95\%}                           & {4\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
-        {3000}   & {10}     & {82\%}                              & {79,89\%}                           & {4\%}                                    & {61\%}        & {87\%}      & {2\%}       & {92\%}        & {81\%}      & {4\%}       \\
-        {3000}   & {15}     & {85\%}                              & {81,89\%}                           & {4\%}                                    & {69\%}        & {89,5\%}    & {2\%}       & {93\%}        & {82\%}      & {4\%}       \\
-        {3000}   & {30}     & {83\%}                              & {80,89\%}                           & {3\%}                                    & {69\%}        & {89\%}      & {2\%}       & {93\%}        & {81\%}      & {4\%}       \\
+        {100}    & {10}     & {81\%}                              & {75,95\%}                           & {9\%}                                    & {21\%}        & {57\%}      & {23\%}      & {16\%}        & {37\%}      & {31\%}      \\
+        {100}    & {15}     & {78\%}                              & {71,99\%}                           & {11\%}                                   & {17\%}        & {52,98\%}   & {25\%}      & {26\%}        & {43,49\%}   & {30\%}      \\
+        {100}    & {30}     & {83\%}                              & {78,99\%}                           & {7\%}                                    & {19\%}        & {54,46\%}   & {24\%}      & {29\%}        & {45,49\%}   & {26\%}      \\
+        {500}    & {10}     & {83\%}                              & {79,95\%}                           & {6\%}                                    & {27\%}        & {61,98\%}   & {16\%}      & {17\%}        & {38,49\%}   & {32\%}      \\
+        {500}    & {15}     & {83\%}                              & {79,5\%}                            & {7\%}                                    & {28\%}        & {63,5\%}    & {16\%}      & {21\%}        & {40\%}      & {31\%}      \\
+        {500}    & {30}     & {84\%}                              & {83,43\%}                           & {5\%}                                    & {30\%}        & {63,98\%}   & {17\%}      & {21\%}        & {40\%}      & {31\%}      \\
+        {1000}   & {10}     & {84\%}                              & {82,42\%}                           & {5\%}                                    & {31\%}        & {65,5\%}    & {13\%}      & {12\%}        & {32,12\%}   & {34\%}      \\
+        {1000}   & {15}     & {85\%}                              & {82,42\%}                           & {5\%}                                    & {31\%}        & {64,5\%}    & {16\%}      & {16\%}        & {37\%}      & {32\%}      \\
+        {1000}   & {30}     & {85\%}                              & {82,81\%}                           & {4\%}                                    & {31\%}        & {63,98\%}   & {17\%}      & {13\%}        & {36\%}      & {34\%}      \\
+        {1500}   & {10}     & {84\%}                              & {81,42\%}                           & {4\%}                                    & {33\%}        & {67,47\%}   & {14\%}      & {12\%}        & {33,43\%}   & {34\%}      \\
+        {1500}   & {15}     & {85\%}                              & {84,43\%}                           & {4\%}                                    & {35\%}        & {68,99\%}   & {12\%}      & {17\%}        & {38,49\%}   & {31\%}      \\
+        {1500}   & {30}     & {84\%}                              & {80,95\%}                           & {4\%}                                    & {36\%}        & {68,99\%}   & {15\%}      & {14\%}        & {36,97\%}   & {33\%}      \\
+        {2000}   & {10}     & {83\%}                              & {82,89\%}                           & {4\%}                                    & {36\%}        & {69\%}      & {13\%}      & {13\%}        & {36\%}      & {34\%}      \\
+        {2000}   & {15}     & {84\%}                              & {81,89\%}                           & {4\%}                                    & {34\%}        & {67\%}      & {14\%}      & {14\%}        & {33,31\%}   & {33\%}      \\
+        {2000}   & {30}     & {85\%}                              & {82,95\%}                           & {4\%}                                    & {32\%}        & {65,5\%}    & {16\%}      & {18\%}        & {39\%}      & {32\%}      \\
+        {3000}   & {10}     & {82\%}                              & {79,89\%}                           & {4\%}                                    & {34\%}        & {68\%}      & {12\%}      & {12\%}        & {32,31\%}   & {34\%}      \\
+        {3000}   & {15}     & {85\%}                              & {81,89\%}                           & {4\%}                                    & {32\%}        & {65,5\%}    & {15\%}      & {13\%}        & {32,73\%}   & {35\%}      \\
+        {3000}   & {30}     & {83\%}                              & {80,89\%}                           & {3\%}                                    & {30\%}        & {63,5\%}    & {16\%}      & {12\%}        & {32,73\%}   & {34\%}      \\
         \bottomrule
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the masked caption generator: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
     \label{tab:results_masked-caption-generator}

### 2023-11-02 source/mlt/print_results.py

@@ -26,13 +26,75 @@ def calculate_f1(run):
             / (run["word_by_word_precision"] + run["word_by_word_recall"])
         )
 
+    for attribute in ["shape", "size", "color"]:
+        if (
+            f"precision_{attribute}" in run.keys()
+            and f"recall_{attribute}" in run.keys()
+        ):
+            run[f"f1_{attribute}"] = (
+                (
+                    2
+                    * (run[f"precision_{attribute}"] * run[f"recall_{attribute}"])
+                    / (run[f"precision_{attribute}"] + run[f"recall_{attribute}"])
+                )
+                if run[f"precision_{attribute}"] + run[f"recall_{attribute}"] != 0
+                else float("nan")
+            )
+
+    return run
+
+
+def calculate_metrics_by_attribute(run):
+    attributes = {
+        "shape": ["sphere", "cube", "cylinder"],
+        "size": [
+            "large",
+            "small",
+        ],
+        "color": [
+            "gray",
+            "red",
+            "blue",
+            "green",
+            "brown",
+            "purple",
+            "cyan",
+            "yellow",
+        ],
+    }
+
+    for metric in ["precision", "recall", "accuracy"]:
+        for attribute, values in attributes.items():
+            for value in values:
+                if f"{metric}_by_word_{value}" not in run.keys():
+                    return run
+
+            run[f"{metric}_{attribute}"] = sum(
+                run[f"{metric}_by_word_{value}"] for value in values
+            ) / len(values)
+
     return run
 
 
 def calculate_metrics(run):
+    run = calculate_metrics_by_attribute(run)
     run = calculate_f1(run)
-
     return run
+
+
+def to_percent(number):
+    rounded = str(round(number * 100, 2)).replace(".", ",").rstrip("0").rstrip(",")
+    return rf"{rounded}\%"
+
+
+def to_number(number):
+    return rf"{str(round(number, 2)).rstrip('0.').replace('.', ',')}"
+
+
+metric_types = {
+    "%": to_percent,
+    "num": to_number,
+}
 
 
 if __name__ == "__main__":
@@ -55,6 +117,12 @@ if __name__ == "__main__":
         help="order of the chosen variables",
     )
     parser.add_argument(
+        "--metric_types",
+        nargs="+",
+        choices=metric_types.keys(),
+        help="type of the variable in order of --variables",
+    )
+    parser.add_argument(
         "--datasets",
         nargs="+",
         type=str,
@@ -62,8 +130,9 @@ if __name__ == "__main__":
     )
 
[…]

### 2023-11-04 thesis/sections/5-grounding.tex
output processor

@@
 The loss is calculated using cross entropy.
-The following values for the variables are compared:
-\begin{itemize}
-    \item $e$ (only for bounding boxes): 100, 500, 1000
-    \item $LSMT_o$: 100, 500, 1000
-    \item $LSTM_e$: 10, 15, 30
-\end{itemize}
+Table \ref{tab:variables-reference-expression-generation} shows which variables are compared for each model:
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lccc}
+        \toprule
+                                  & $e$                & $LSTM_o$           & $LSTM_e$       \\
+                                  & $[100, 500, 1000]$ & $[100, 500, 1000]$ & $[10, 15, 30]$ \\\midrule
+        bounding box RE generator & \times             & \times             & \times         \\
+        basic RE generator        & -                  & \times             & \times         \\
+        masked RE generator       & -                  & \times             & \times         \\
+        \bottomrule
+    \end{tabular}
+    \caption{Variables for each model where $e$ is the embedding dimension, $LSTM_e$ the embedding dimensions for tokens in the LSTM and $LSTM_o$ the output dimension of the LSTM}
+    \label{tab:variables-reference-expression-generation}
+\end{table}
 

### 2023-11-04 source/mlt/feature_extractors.py

@@ -266,7 +266,6 @@ if __name__ == "__main__":
     if args.bounding_boxes:
         scenes = sorted(os.listdir(args.scene_dir))
 
-        # TODO optimize performance
         max_number_objects = 0
         for scene_file in scenes:
             with open(
@@ -315,6 +314,7 @@ if __name__ == "__main__":
                     number_images = int(len(batch) / max_number_objects)
 
                     with torch.no_grad():
+                        # pylint: disable-next=not-callable
                         features = feature_extractor(torch.stack(batch)).cpu()
 
                     features_by_image = features.view(
@@ -352,6 +352,7 @@ if __name__ == "__main__":
 
                 if len(batch) == args.batch_size or image_index == len(images) - 1:
                     with torch.no_grad():
+                        # pylint: disable-next=not-callable
                         features = feature_extractor(torch.stack(batch)).cpu()
                     i_end = i_start + len(batch)
                     feature_dataset[i_start:i_end] = features

### 2023-11-04 source/mlt/language_games/play.py

@@ -6,7 +6,6 @@ from dataclasses import dataclass
 from time import gmtime, strftime
 from typing import Callable
 
-import torch
 from egg import core
 from mlt.feature_extractors import DummyFeatureExtractor, ResnetFeatureExtractor
 from mlt.image_loader import FeatureImageLoader, ImageLoader

### 2023-11-04 source/mlt/preexperiments/train.py

@@ -43,11 +43,7 @@ from mlt.preexperiments.test import (
     CoordinatePredictorTester,
     Tester,
 )
-from mlt.shared_models import (
-    BoundingBoxImageEncoder,
-    ClevrImageEncoder,
-    CoordinateClassifier,
-)
+from mlt.shared_models import ClevrImageEncoder, CoordinateClassifier
 from torch import nn, optim
 from torch.nn import Module
 from torch.utils.data import DataLoader, Dataset, random_split

### 2023-11-04 source/mlt/print_results.py

@@ -3,86 +3,14 @@ import glob
 import json
 import os
 import re
+
+# necessary for input()
+# pylint: disable-next=unused-import
+import readline
 from abc import abstractmethod
 from dataclasses import dataclass
 
 from tabulate import tabulate
-
-
-class Experiments:
-    def __init__(self, root_folder, datasets, variables, variable_order) -> None:
-        self.datasets = datasets
-        self.variables = variables
-        self.variable_order = variable_order
-
-        self.experiments: list[Experiment] = []
-        for folder in glob.glob(root_folder):
-            folder_name = folder.split("/")[-1]
-
-            dataset, experiment_variables = self._extract_information_from_folder(
-                folder_name
-            )
-            if dataset not in datasets:
-                continue
-
-            if len(experiment_variables) != len(variables):
-                raise ValueError(f"Number of Variables don't match: {folder_name}")
-            ordered_variables = tuple(experiment_variables[o] for o in variable_order)
-
-            run = Run(os.path.join(folder, "log.txt"))
-            if len(run.epochs) != 0:
-                self.experiments.append(
-                    Experiment(
-                        folder_name=folder_name,
-                        dataset=dataset,
-                        variables=ordered_variables,
-                        run=run,
-                    )
-                )
-
-    def _extract_information_from_folder(self, folder_name):
-        pattern = r"_([^_]+)((_\d+)*)$"
-        match = re.search(pattern, folder_name)
-        if match:
-            dataset = match.group(1)
-            experiment_variables = [
-                int(var) for var in match.group(2).split("_") if var != ""
-            ]
-
-        return dataset, experiment_variables
-
-    def _get_excluded_folders(self):
-        for index, folder in enumerate(
-            [experiment.folder_name for experiment in self.experiments]
-        ):
-            print(f"{index}: {folder}")
-        return input("exclude folders (separated by comma, ranges with hyphen):")
-
-    def exclude_folders(self):
-        excluded_folders = self._get_excluded_folders()
-
-        if len(excluded_folders) != 0:
-            excluded_indices = []
-            for index in excluded_folders.split(","):
-                if "-" in index:
-                    range_indices = index.split("-")
-                    excluded_indices.extend(
-                        range(int(range_indices[0]), int(range_indices[1]) + 1)
-                    )
[…]

### 2023-11-07 thesis/sections/5-grounding.tex
resulst

@@
 \subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
+Table \ref{tab:results:bounding_box_classifier} lists the accuracies of the models' predictions after 30 epochs for all three different datasets and different embedding sizes $e$.
 
@@
     \end{tabular}
-    \caption{Accuracy scores of the object identifier: $e$ are different embedding sizes}
-    \label{tab:results_bounding_box_classifier}
+    \caption{Accuracy scores of the object identifier after 30 epochs: $e$ are different embedding sizes}
+    \label{tab:results:bounding_box_classifier}
 \end{table}
 
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
+The first trend that is visible is that fewer distractor increase the accuracy of the model.
+The model achieves almost perfect accuracy in all configurations for the \emph{Dale 2} dataset with only two objects.
+With five objects in the \emph{Dale-5} dataset, the accuracy drops to 95\% and with a maximum possible number of 10 objects in the \emph{CLEVR color} dataset the model only achieves 94\% in the best configuration.
+This is not very surprising, since the model has a higher chance of predicting a wrong object, given the random baselines of 50\%, 20\% and 10\%.
+Furthermore, when more objects are shown to the model, more attributes are needed to discriminate the target object from the distractor, since it is more likely that the distractors share attributes with the target object.
+The task is getting therefore more challenging.
+
+A second conclusion is that the model needs a certain embedding space, to represent the features of each image.
+Across all dataset, an embedding space of only 2 dimensions result in a much lower performance compared to the best configuration.
+Even though the model is still able to achieve an accuracy of 30\% to 50\% points over the random baseline, this shows that a bigger embedding size is beneficial to extract and represent the features of the objects.
+Between 50 and 1000 dimensions, the model performs the best and varies only by 1\% point, which may be due to different random initialization of the weights in the model.
+
+In conclusion, the model is able to learn to discriminate the objects based on the visual appearance.
 The model can generate these high results, even with its relatively simple architecture.
@@
 % SD: The model is able to leatn to discriminate objects based on visual appearance.
-% DK: TODO
+% DK: (done)
 
@@
 \subsubsection*{Setup}
-\cmtDK[inline]{new metrics}
 
@@
 \subsubsection*{Results}
-In the next paragraphs, the results of the caption generator are discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
+Table \ref{tab:results:bb-re-generator} shows the \textbf{overall accuracy}, \textbf{F1 scores} for each word and the \textbf{non-target accuracies} of the \emph{bounding box RE generator} when trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
+As can be seen, the overall accuracies, in other words perfect matches of the generated referring expression depend very much on the dataset.
+With the \emph{Dale-2} dataset, the model can achieve perfect matches in 99\% of the cases in its best configuration.
+Also the \emph{CLEVR color} dataset allows the model to predict the correct referring expression in 93\% of the samples.
+Opposed to that the model can only generate perfect referring expressions in 69\% of the samples of the \emph{Dale-5} dataset.
+Looking at the \textbf{non-target accuracy}, one of the problems can be identified for the \emph{CLEVR color} dataset.
+Summing the \textbf{non-target accuracy} and the \textbf{overall accuracy}, one gets the accuracy that any of the shown object was described independently if it was the target object.
+For the \emph{CLEVR color} dataset, this score lies at 96\% to 97\% in the best configurations, which means that the model described a shown object for almost all the samples.
+In 3\% to 4\% of the cases, the model only picked the wrong object to describe.
+This looks different for the \emph{Dale-5} dataset.
+Here, the model describes only in 71\% of the cases one of the shown objects, 69\% the target object and 2\% a distractor.
+The model therefore struggles more with the correct generation of words than with choosing, which object to describe.
+
+When looking at the different configurations, in especially different values for $e$, $LSTM_o$ and $LSTM_e$, it can be seen that the results are not very different for the \emph{Dale-2} and \emph{CLEVR color} datasets.
+However, bigger effects can be identified for the \emph{Dale-5} dataset.
+Here, a low output dimension of the LSTM $LSTM_o$ tends to give lower scores.
+This especially enhanced, when also the embedding dimensions of the tokens $LSTM_e$ is low.
+Again, the image embedding size $e$ doesn't seem to have a big effect, when over 50 dimensions.
+Indeed, the model achieves an accuracy of 69\% with all three tested embedding sizes.
 
@@
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
-    \label{tab:results_bounding-box-caption-generator}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator after 30 epochs: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \label{tab:results:bb-re-generator}
 \end{table}
 
+Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results.
+They list accuracy, precision and recall metrics for each token for the best configuration of the model with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+The tokens are grouped by attribute and also show the metrics averaged over each of the attributes.
+The metrics of the \emph{<pad>} token indicate if the model produced the correct length of the referring expression, in other words if it was able to determine which attributes are necessary to discriminate the target object from the distractors.
+For both \emph{Dale-2} and the \emph{CLEVR color}, the scores are prefect.
+Former follows the almost perfect \emph{overall accuracy} discussed before, while latter is also not surprising, because all referring expressions for the \emph{CLEVR color} dataset consist of two attributes, shape and color.
+The \emph{<pad>} token is therefore easy to learn.
+For the \emph{Dale-5} dataset, the model struggles more, to predict the correct length of the referring expression.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{r|c|cc|c|ccc|c|c}
+        \toprule
+        \textbf{Dataset}                      & {Metric} & {small} & {large} & \textbf{size}   & {cube}  & {cylinder} & {sphere} & \textbf{shape}   & {<pad>} \\\midrule
+        \multirow{3}{*}{\textbf{Dale-2}}      & {Acc.}   & {98\%}  & {94\%}  & \textbf{96\%}   & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
+                                              & {Prec.}  & {99\%}  & {98\%}  & \textbf{98,5\%} & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
+                                              & {Rec.}   & {98\%}  & {94\%}  & \textbf{96\%}   & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\\midrule
+        \multirow{3}{*}{\textbf{Dale-5}}      & {Acc.}   & {62\%}  & {66\%}  & \textbf{64\%}   & {99\%}  & {98\%}     & {98\%}   & \textbf{98,33\%} & {85\%}  \\
+                                              & {Prec.}  & {70\%}  & {69\%}  & \textbf{69,5\%} & {98\%}  & {98\%}     & {98\%}   & \textbf{98\%}    & {82\%}  \\
+                                              & {Rec.}   & {62\%}  & {66\%}  & \textbf{64\%}   & {99\%}  & {98\%}     & {98\%}   & \textbf{98,33\%} & {85\%}  \\\midrule
+        \multirow{3}{*}{\textbf{CLEVR color}} & {Acc.}   & {-}     & {-}     & \textbf{-}      & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
+                                              & {Prec.}  & {-}     & {-}     & \textbf{-}      & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
+                                              & {Rec.}   & {-}     & {-}     & \textbf{-}      & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Accuracy, Precision and Recall for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+Columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
+    \label{tab:results:bb-re-generator_size-shape}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{r|c|cccccccc|c}
+        \toprule
+        \textbf{Dataset}                      & {Metric} & {blue} & {brown} & {cyan} & {gray} & {green} & {purple} & {red}   & {yellow} & \textbf{color}   \\\midrule
+        \multirow{3}{*}{\textbf{Dale-2}}      & {Acc.}   & {98\%} & {100\%} & {99\%} & {97\%} & {97\%}  & {99\%}   & {100\%} & {99\%}   & \textbf{98,62\%} \\
+                                              & {Prec.}  & {95\%} & {99\%}  & {98\%} & {99\%} & {99\%}  & {99\%}   & {97\%}  & {100\%}  & \textbf{98,25\%} \\
+                                              & {Rec.}   & {98\%} & {100\%} & {99\%} & {97\%} & {97\%}  & {99\%}   & {100\%} & {99\%}   & \textbf{98,62\%} \\\midrule
+        \multirow{3}{*}{\textbf{Dale-5}}      & {Acc.}   & {92\%} & {90\%}  & {95\%} & {95\%} & {96\%}  & {92\%}   & {89\%}  & {95\%}   & \textbf{93\%}    \\
+                                              & {Prec.}  & {92\%} & {94\%}  & {89\%} & {89\%} & {93\%}  & {91\%}   & {97\%}  & {94\%}   & \textbf{92,38\%} \\
+                                              & {Rec.}   & {92\%} & {90\%}  & {95\%} & {95\%} & {96\%}  & {92\%}   & {89\%}  & {95\%}   & \textbf{93\%}    \\\midrule
+        \multirow{3}{*}{\textbf{CLEVR color}} & {Acc.}   & {93\%} & {92\%}  & {96\%} & {90\%} & {94\%}  & {91\%}   & {94\%}  & {92\%}   & \textbf{92,75\%} \\
+                                              & {Prec.}  & {93\%} & {92\%}  & {94\%} & {94\%} & {92\%}  & {91\%}   & {90\%}  & {95\%}   & \textbf{92,63\%} \\
+                                              & {Rec.}   & {93\%} & {92\%}  & {96\%} & {90\%} & {94\%}  & {91\%}   & {94\%}  & {92\%}   & \textbf{92,75\%} \\
+
+        \bottomrule
+    \end{tabular}
+    \caption{Accuracy, Precision and Recall for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The Column \textbf{color} shows the average across all colors.}
+    \label{tab:results:bb-re-generator_color}
+\end{table}
+
+ResNet-3 produced the best results, when it was used as a feature extractor.
+The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
 \begin{table}[ht]
@@
     \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the masked caption generator: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
-    \label{tab:results_masked-caption-generator}
+    \label{tab:results:masked-re-generator}
     % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1.

### 2023-11-07 thesis/sections/5-grounding.tex
results

@@
 \subsubsection*{Results}
-Table \ref{tab:results:bb-re-generator} shows the \textbf{overall accuracy}, \textbf{F1 scores} for each word and the \textbf{non-target accuracies} of the \emph{bounding box RE generator} when trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
+Table \ref{tab:results:bb-re-generator} shows the \emph{overall accuracy}, \emph{F1 scores} for each word and the \emph{non-target accuracies} of the \textbf{bounding box RE generator} when trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
 As can be seen, the overall accuracies, in other words perfect matches of the generated referring expression depend very much on the dataset.
@@
 Opposed to that the model can only generate perfect referring expressions in 69\% of the samples of the \emph{Dale-5} dataset.
-Looking at the \textbf{non-target accuracy}, one of the problems can be identified for the \emph{CLEVR color} dataset.
-Summing the \textbf{non-target accuracy} and the \textbf{overall accuracy}, one gets the accuracy that any of the shown object was described independently if it was the target object.
+Looking at the \emph{non-target accuracy}, one of the problems can be identified for the \emph{CLEVR color} dataset.
+Summing the \emph{non-target accuracy} and the \emph{overall accuracy}, one gets the accuracy that any of the shown object was described independently if it was the target object.
 For the \emph{CLEVR color} dataset, this score lies at 96\% to 97\% in the best configurations, which means that the model described a shown object for almost all the samples.
@@
         $e$    & $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} \\\midrule
-        {100}  & {100}    & {10}     & {97\%}                              & {97,5\%}                            & {1\%}                                    & {65\%}        & {88\%}          & {2\%}        & {92\%}        & {81\%}        & {3\%}        \\
-        {100}  & {100}    & {15}     & {97\%}                              & {98\%}                              & {0\%}                                    & {62\%}        & {86,47\%}       & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
-        {100}  & {100}    & {30}     & {98\%}                              & {98\%}                              & {1\%}                                    & {65\%}        & {87,99\%}       & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
-        {100}  & {500}    & {10}     & {98\%}                              & {98,5\%}                            & {0\%}                                    & {67\%}        & {88\%}          & {2\%}        & {93\%}        & {82\%}        & {4\%}        \\
-        {100}  & {500}    & {30}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
-        {100}  & {1000}   & {10}     & {98\%}                              & {98\%}                              & {0\%}                                    & {68\%}        & {88\%}          & {2\%}        & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
-        {500}  & {500}    & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & {67\%}        & {88,99\%}       & {2\%}        & \textbf{93\%} & \textbf{82\%} & \textbf{3\%} \\
-        {500}  & {1000}   & {30}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & {69\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
-        {1000} & {100}    & {10}     & {97\%}                              & {97,99\%}                           & {0\%}                                    & {59\%}        & {85,5\%}        & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
-        {1000} & {100}    & {30}     & {98\%}                              & {98,5\%}                            & {0\%}                                    & {61\%}        & {87\%}          & {2\%}        & {92\%}        & {81\%}        & {4\%}        \\
-        {1000} & {1000}   & {10}     & \textbf{99\%}                       & \textbf{99\%}                       & \textbf{0\%}                             & \textbf{69\%} & \textbf{89,5\%} & \textbf{2\%} & {93\%}        & {82\%}        & {4\%}        \\
-        {1000} & {1000}   & {15}     & {99\%}                              & {98,5\%}                            & {0\%}                                    & {68\%}        & {89\%}          & {2\%}        & {93\%}        & {81\%}        & {4\%}        \\
+        {100}  & {100}    & {10}     & {97}                                & {97,67}                             & {1}                                      & {65}          & {88,17}        & {2}         & {92}          & {94,95}        & {3}         \\
+        {100}  & {100}    & {15}     & {97}                                & {97,77}                             & {0}                                      & {62}          & {86,51}        & {2}         & {92}          & {94,45}        & {4}         \\
+        {100}  & {100}    & {30}     & {98}                                & {97,81}                             & {1}                                      & {65}          & {87,97}        & {2}         & {92}          & {94,47}        & {4}         \\
+        {100}  & {500}    & {10}     & {98}                                & {98,16}                             & {0}                                      & {67}          & {88,13}        & {2}         & {93}          & {95,18}        & {4}         \\
+        {100}  & {500}    & {30}     & \textbf{99}                         & \textbf{98,57}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,53} & \textbf{2}  & \textbf{93}   & \textbf{95,17} & \textbf{3}  \\
+        {100}  & {1000}   & {10}     & {98}                                & {98,28}                             & {0}                                      & {68}          & {88,23}        & {2}         & \textbf{93}   & \textbf{95,41} & \textbf{3}  \\
+        {500}  & {1000}   & {30}     & \textbf{99}                         & \textbf{98,85}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,15} & \textbf{2}  & {93}          & {95,05}        & {4}         \\
+        {500}  & {500}    & {10}     & \textbf{99}                         & \textbf{98,71}                      & \textbf{0}                               & {67}          & {88,99}        & {2}         & \textbf{93}   & \textbf{95,4}  & \textbf{3}  \\
+        {1000} & {100}    & {10}     & {97}                                & {97,72}                             & {0}                                      & {59}          & {85,43}        & {2}         & {93}          & {95,04}        & {4}         \\
+        {1000} & {100}    & {30}     & {98}                                & {98,16}                             & {0}                                      & {61}          & {86,77}        & {2}         & {92}          & {94,68}        & {4}         \\
+        {1000} & {1000}   & {10}     & \textbf{99}                         & \textbf{98,95}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,21} & \textbf{2}  & {93}          & {95,15}        & {4}         \\
+        {1000} & {1000}   & {15}     & {99}                                & {98,67}                             & {0}                                      & {68}          & {88,98}        & {2}         & {93}          & {95,01}        & {4}         \\
         \bottomrule
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the bounding box caption generator after 30 epochs: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the bounding box caption generator after 30 epochs: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
     \label{tab:results:bb-re-generator}
@@
 
-Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results.
-They list accuracy, precision and recall metrics for each token for the best configuration of the model with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results and especially what mistakes the model is making for both the \emph{Dale-5} and \emph{CLEVR color} datasets.
+They list \emph{precision} and \emph{recall} metrics for each token for the best configuration of the model with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
 The tokens are grouped by attribute and also show the metrics averaged over each of the attributes.
+Since the overall results for the \emph{Dale-2} dataset are already close to perfect, the focus for this analysis lies on the remaining two datasets.
 The metrics of the \emph{<pad>} token indicate if the model produced the correct length of the referring expression, in other words if it was able to determine which attributes are necessary to discriminate the target object from the distractors.
-For both \emph{Dale-2} and the \emph{CLEVR color}, the scores are prefect.
-Former follows the almost perfect \emph{overall accuracy} discussed before, while latter is also not surprising, because all referring expressions for the \emph{CLEVR color} dataset consist of two attributes, shape and color.
+For the \emph{CLEVR color} dataset, the scores are perfect.
+This is not surprising, because all referring expressions for the \emph{CLEVR color} dataset consist of exactly two attributes, shape and color, and the first generated token will always be the only \emph{<pad>} token in the referring expression (corresponding to the unspecified size).
 The \emph{<pad>} token is therefore easy to learn.
-For the \emph{Dale-5} dataset, the model struggles more, to predict the correct length of the referring expression.
+For the \emph{Dale-5} dataset, the model struggles more to predict the correct length of the referring expression.
+\cmtDK[inline]{why?}
+
+When looking at the tokens for the shape, it can be seen that the model is able to identify it very well across all datasets.
+The model predicts the correct shape for all samples using the \emph{CLEVR color} dataset, while both \emph{precision} and \emph{recall} lie around 98,3\% when using the \emph{Dale-5} dataset.
+Even though the score is almost perfect, the slight difference might stem from the fact that all distractors have the same shape in the first case, while distractors can be different in the second case.
+Consequently, the model is only exposed to one shape at a time for each sample, which might simplify its identification.
+
+For the color attribute, the metrics drop significantly for both \emph{Dale-5} and \emph{CLEVR color} to an average of around 93\%.
+Hereby, no meaningful difference can be seen across the datasets, but there are differences between the colors.
+Some colors are predicted with \emph{precision} and \emph{recall} around 95\% to 96\%, while others are only around 90\%.
+However, these differences are not reproducible across multiple runs and configurations.
+The best and worst predicted colors vary and no conclusions can be drawn which colors are easier to predict for the model.
+
+Finally, the size is the most difficult attribute to predict for the model.
+Apart from the \emph{CLEVR color} dataset, where a size never needs to be predicted and also is never predicted, the metrics for the prediction of size tokens are the lowest across all tokens.
+They are the only mistakes, the model makes, when exposed to the \emph{Dale-2} dataset and the average \emph{precision} lies around 23\% below the average of predictions of the color for the \emph{Dale-5} dataset, while the average \emph{recall} lies around 28,82\% below.
+The reason why the \emph{precision} is higher than the \emph{recall} is the \emph{<pad>} token, which is predicted very often instead of a token specifying the size.
+In fact, the opposite relationship is visible for the \emph{precision} and \emph{recall} for said token.
+The much higher absolute number of \emph{<pad>} tokens leads to a smaller relative difference of \%-points shown in the table.
+Again, no conclusion can be drawn if larger or smaller objects are easier to predict, since the results vary across runs and configurations.
 
@@
     \centering
-    \begin{tabular}{r|c|cc|c|ccc|c|c}
+    \begin{tabular}{rr|cc|c|ccc|c|c}
         \toprule
-        \textbf{Dataset}                      & {Metric} & {small} & {large} & \textbf{size}   & {cube}  & {cylinder} & {sphere} & \textbf{shape}   & {<pad>} \\\midrule
-        \multirow{3}{*}{\textbf{Dale-2}}      & {Acc.}   & {98\%}  & {94\%}  & \textbf{96\%}   & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
-                                              & {Prec.}  & {99\%}  & {98\%}  & \textbf{98,5\%} & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
-                                              & {Rec.}   & {98\%}  & {94\%}  & \textbf{96\%}   & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\\midrule
-        \multirow{3}{*}{\textbf{Dale-5}}      & {Acc.}   & {62\%}  & {66\%}  & \textbf{64\%}   & {99\%}  & {98\%}     & {98\%}   & \textbf{98,33\%} & {85\%}  \\
-                                              & {Prec.}  & {70\%}  & {69\%}  & \textbf{69,5\%} & {98\%}  & {98\%}     & {98\%}   & \textbf{98\%}    & {82\%}  \\
-                                              & {Rec.}   & {62\%}  & {66\%}  & \textbf{64\%}   & {99\%}  & {98\%}     & {98\%}   & \textbf{98,33\%} & {85\%}  \\\midrule
-        \multirow{3}{*}{\textbf{CLEVR color}} & {Acc.}   & {-}     & {-}     & \textbf{-}      & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
-                                              & {Prec.}  & {-}     & {-}     & \textbf{-}      & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
-                                              & {Rec.}   & {-}     & {-}     & \textbf{-}      & {100\%} & {100\%}    & {100\%}  & \textbf{100\%}   & {100\%} \\
+                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {99,17} & {98,29} & \textbf{98,73} & {99,86} & {99,71}    & {99,67}  & \textbf{99,75} & {99,64} \\
+                                         & {Recall}    & {97,54} & {94,26} & \textbf{95,9}  & {100}   & {99,56}    & {99,67}  & \textbf{99,74} & {99,77} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {69,65} & {69,21} & \textbf{69,43} & {98,19} & {98,32}    & {98,39}  & \textbf{98,3}  & {82,22} \\
+                                         & {Recall}    & {62,11} & {66,15} & \textbf{64,13} & {98,79} & {97,87}    & {98,25}  & \textbf{98,3}  & {84,59} \\\midrule
+        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
+                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
         \bottomrule
     \end{tabular}
-    \caption{Accuracy, Precision and Recall for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
-Columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
+    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
     \label{tab:results:bb-re-generator_size-shape}
@@
     \centering
-    \begin{tabular}{r|c|cccccccc|c}
+    \begin{tabular}{rr|cccccccc|c}
         \toprule
-        \textbf{Dataset}                      & {Metric} & {blue} & {brown} & {cyan} & {gray} & {green} & {purple} & {red}   & {yellow} & \textbf{color}   \\\midrule
-        \multirow{3}{*}{\textbf{Dale-2}}      & {Acc.}   & {98\%} & {100\%} & {99\%} & {97\%} & {97\%}  & {99\%}   & {100\%} & {99\%}   & \textbf{98,62\%} \\
-                                              & {Prec.}  & {95\%} & {99\%}  & {98\%} & {99\%} & {99\%}  & {99\%}   & {97\%}  & {100\%}  & \textbf{98,25\%} \\
-                                              & {Rec.}   & {98\%} & {100\%} & {99\%} & {97\%} & {97\%}  & {99\%}   & {100\%} & {99\%}   & \textbf{98,62\%} \\\midrule
-        \multirow{3}{*}{\textbf{Dale-5}}      & {Acc.}   & {92\%} & {90\%}  & {95\%} & {95\%} & {96\%}  & {92\%}   & {89\%}  & {95\%}   & \textbf{93\%}    \\
-                                              & {Prec.}  & {92\%} & {94\%}  & {89\%} & {89\%} & {93\%}  & {91\%}   & {97\%}  & {94\%}   & \textbf{92,38\%} \\
-                                              & {Rec.}   & {92\%} & {90\%}  & {95\%} & {95\%} & {96\%}  & {92\%}   & {89\%}  & {95\%}   & \textbf{93\%}    \\\midrule
-        \multirow{3}{*}{\textbf{CLEVR color}} & {Acc.}   & {93\%} & {92\%}  & {96\%} & {90\%} & {94\%}  & {91\%}   & {94\%}  & {92\%}   & \textbf{92,75\%} \\
-                                              & {Prec.}  & {93\%} & {92\%}  & {94\%} & {94\%} & {92\%}  & {91\%}   & {90\%}  & {95\%}   & \textbf{92,63\%} \\
-                                              & {Rec.}   & {93\%} & {92\%}  & {96\%} & {90\%} & {94\%}  & {91\%}   & {94\%}  & {92\%}   & \textbf{92,75\%} \\
+                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {94,51} & {98,77} & {97,59} & {98,68} & {98,89} & {98,8}   & {97,47} & {100}    & \textbf{98,09} \\
+                                         & {Recall}    & {97,73} & {100}   & {98,78} & {97,4}  & {96,74} & {98,8}   & {100}   & {98,8}   & \textbf{98,53} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {92,12} & {93,82} & {89,13} & {89,12} & {92,63} & {91,12}  & {97,24} & {94,36}  & \textbf{92,44} \\
+                                         & {Recall}    & {92,12} & {89,78} & {94,91} & {94,51} & {95,71} & {92,42}  & {89,34} & {94,85}  & \textbf{92,95} \\\midrule
+        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                           \\\textbf{color}} & {Precision}           & {93,46} & {92,37} & {94,47} & {93,86} & {92,04} & {91,13}  & {90,07} & {94,7}   & \textbf{92,76} \\
+                                         & {Recall}    & {92,75} & {92}    & {95,98} & {89,92} & {94,12} & {91,13}  & {94,23} & {91,91}  & \textbf{92,76} \\
 
@@
     \end{tabular}
-    \caption{Accuracy, Precision and Recall for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The Column \textbf{color} shows the average across all colors.}
+    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
     \label{tab:results:bb-re-generator_color}
@@
 
+The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
+When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
+The final peak stayed exactly the same and the effects were therefore not studied deeper.
+% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
+% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
+
+In conclusion, it can be said that the model is able to extract discriminative features from the shown bounding boxes and produce referring expressions.
+However, the result highly depends on the amount of distractors and the resulting need of more discriminative features to describe the target object.
+While the shape is easily identified, the model has bigger problems to identify the color and especially the size of the target object.
+
 The sum of the two columns for the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
@@
 % DK: no, because a result of one specific sample (done)
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
-The final peak stayed exactly the same and the effects were therefore not studied deeper.
-% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
-% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
 

### 2023-11-07 source/mlt/preexperiments/test.py

@@ -42,8 +42,8 @@ class CoordinatePredictorTester(Tester):
         return (
             json.dumps(
                 {
-                    "accuracy": f"{accuracy.compute():.2f}",
-                    "mean test loss": f"{mean.compute():.2f}",
+                    "accuracy": f"{accuracy.compute():.4f}",
+                    "mean test loss": f"{mean.compute():.4f}",
                 }
             ),
             test_outputs,
@@ -68,7 +68,7 @@ class BoundingBoxClassifierTester(Tester):
         return (
             json.dumps(
                 {
-                    "accuracy": f"{accuracy.compute():.2f}",
+                    "accuracy": f"{accuracy.compute():.4f}",
                 }
             ),
             test_outputs,
@@ -149,19 +149,19 @@ class CaptionGeneratorTester(Tester):
         logging.getLogger().setLevel(logging.WARNING)
 
         accuracy_by_word = {
-            word: round(accuracy.item(), 2)
+            word: round(accuracy.item(), 4)
             for word, accuracy in zip(
                 DaleCaptionAttributeEncoder.vocab, computed_class_accuracy
             )
         }
         precision_by_word = {
-            word: round(precision.item(), 2)
+            word: round(precision.item(), 4)
             for word, precision in zip(
                 DaleCaptionAttributeEncoder.vocab, computed_class_precisions
             )
         }
         recall_by_word = {
-            word: round(recall.item(), 2)
+            word: round(recall.item(), 4)
             for word, recall in zip(
                 DaleCaptionAttributeEncoder.vocab, computed_class_recall
             )
@@ -176,14 +176,14 @@ class CaptionGeneratorTester(Tester):
         return (
             json.dumps(
                 {
-                    "accuracy": f"{accuracy.compute():.2f}",
-                    "word_by_word_accuracy": f"{computed_accuracy:.2f}",
+                    "accuracy": f"{accuracy.compute():.4f}",
+                    "word_by_word_accuracy": f"{computed_accuracy:.4f}",
                     "accuracy_by_word": accuracy_by_word,
-                    "word_by_word_precision": f"{torch.mean(computed_class_precisions[included_indices]):.2f}",
+                    "word_by_word_precision": f"{torch.mean(computed_class_precisions[included_indices]):.4f}",
                     "precision_by_word": precision_by_word,
-                    "word_by_word_recall": f"{torch.mean(computed_class_recall[included_indices]):.2f}",
[…]

### 2023-11-07 thesis/sections/5-grounding.tex
results

@@
                                          & {Recall}    & {92,75} & {92}    & {95,98} & {89,92} & {94,12} & {91,13}  & {94,23} & {91,91}  & \textbf{92,76} \\
-
         \bottomrule
@@
 
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very little in the following epochs.
+The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
+When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
+The final peak stayed exactly the same and the effects were therefore not studied deeper.
+% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
+% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
+
+In conclusion, it can be said that the model is able to extract discriminative features from the shown bounding boxes and produce referring expressions.
+However, the result highly depends on the amount of distractors and the resulting need of more discriminative features to describe the target object.
+While the shape is easily identified, the model has bigger problems to identify the color and especially the size of the target object.
+
+In the following paragraphs, the results of the \textbf{basic RE generator} as well as the \textbf{masked RE generator} are evaluated.
+Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 30 epochs.
+Compared to the \emph{bounding box RE generator}, the task differs substantially.
+Instead of bounding boxes, the model is shown the image of the whole scene.
+Additionally to extracting features of the target object and distractors separately, the model is now tasked to extract these at once for all objects.
+Furthermore, it needs to learn which of the objects is the target object by combining the image with the masked image.
+
 \begin{table}[ht]
@@
         $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} \\\midrule
-        {100}    & {10}     & {81\%}                              & {75,95\%}                           & {9\%}                                    & {21\%}        & {57\%}      & {23\%}      & {16\%}        & {37\%}      & {31\%}      \\
-        {100}    & {15}     & {78\%}                              & {71,99\%}                           & {11\%}                                   & {17\%}        & {52,98\%}   & {25\%}      & {26\%}        & {43,49\%}   & {30\%}      \\
-        {100}    & {30}     & {83\%}                              & {78,99\%}                           & {7\%}                                    & {19\%}        & {54,46\%}   & {24\%}      & {29\%}        & {45,49\%}   & {26\%}      \\
-        {500}    & {10}     & {83\%}                              & {79,95\%}                           & {6\%}                                    & {27\%}        & {61,98\%}   & {16\%}      & {17\%}        & {38,49\%}   & {32\%}      \\
-        {500}    & {15}     & {83\%}                              & {79,5\%}                            & {7\%}                                    & {28\%}        & {63,5\%}    & {16\%}      & {21\%}        & {40\%}      & {31\%}      \\
-        {500}    & {30}     & {84\%}                              & {83,43\%}                           & {5\%}                                    & {30\%}        & {63,98\%}   & {17\%}      & {21\%}        & {40\%}      & {31\%}      \\
-        {1000}   & {10}     & {84\%}                              & {82,42\%}                           & {5\%}                                    & {31\%}        & {65,5\%}    & {13\%}      & {12\%}        & {32,12\%}   & {34\%}      \\
-        {1000}   & {15}     & {85\%}                              & {82,42\%}                           & {5\%}                                    & {31\%}        & {64,5\%}    & {16\%}      & {16\%}        & {37\%}      & {32\%}      \\
-        {1000}   & {30}     & {85\%}                              & {82,81\%}                           & {4\%}                                    & {31\%}        & {63,98\%}   & {17\%}      & {13\%}        & {36\%}      & {34\%}      \\
-        {1500}   & {10}     & {84\%}                              & {81,42\%}                           & {4\%}                                    & {33\%}        & {67,47\%}   & {14\%}      & {12\%}        & {33,43\%}   & {34\%}      \\
-        {1500}   & {15}     & {85\%}                              & {84,43\%}                           & {4\%}                                    & {35\%}        & {68,99\%}   & {12\%}      & {17\%}        & {38,49\%}   & {31\%}      \\
-        {1500}   & {30}     & {84\%}                              & {80,95\%}                           & {4\%}                                    & {36\%}        & {68,99\%}   & {15\%}      & {14\%}        & {36,97\%}   & {33\%}      \\
-        {2000}   & {10}     & {83\%}                              & {82,89\%}                           & {4\%}                                    & {36\%}        & {69\%}      & {13\%}      & {13\%}        & {36\%}      & {34\%}      \\
-        {2000}   & {15}     & {84\%}                              & {81,89\%}                           & {4\%}                                    & {34\%}        & {67\%}      & {14\%}      & {14\%}        & {33,31\%}   & {33\%}      \\
-        {2000}   & {30}     & {85\%}                              & {82,95\%}                           & {4\%}                                    & {32\%}        & {65,5\%}    & {16\%}      & {18\%}        & {39\%}      & {32\%}      \\
-        {3000}   & {10}     & {82\%}                              & {79,89\%}                           & {4\%}                                    & {34\%}        & {68\%}      & {12\%}      & {12\%}        & {32,31\%}   & {34\%}      \\
-        {3000}   & {15}     & {85\%}                              & {81,89\%}                           & {4\%}                                    & {32\%}        & {65,5\%}    & {15\%}      & {13\%}        & {32,73\%}   & {35\%}      \\
-        {3000}   & {30}     & {83\%}                              & {80,89\%}                           & {3\%}                                    & {30\%}        & {63,5\%}    & {16\%}      & {12\%}        & {32,73\%}   & {34\%}      \\
+        {100}     & {10}      & {81}                                & {76,16}                             & {9}                                      & {21}          & {57,07}        & {23}        & {16}          & {43,69}        & {31}        \\
+        {100}     & {15}      & {78}                                & {72,06}                             & {11}                                     & {17}          & {53,29}        & {25}        & {26}          & {50,87}        & {30}        \\
+        {100}     & {30}      & {83}                                & {78,76}                             & {7}                                      & {19}          & {54,91}        & {24}        & \textbf{29}   & \textbf{53,37} & \textbf{26} \\
+        {500}     & {10}      & {83}                                & {79,78}                             & {6}                                      & {27}          & {62,05}        & {16}        & {17}          & {44,93}        & {32}        \\
+        {500}     & {15}      & {83}                                & {79,75}                             & {7}                                      & {28}          & {63,78}        & {16}        & {21}          & {47,08}        & {31}        \\
+        {500}     & {30}      & {84}                                & {83,03}                             & {5}                                      & {30}          & {64,36}        & {17}        & {21}          & {40}           & {31}        \\
+        {1000}    & {10}      & {84}                                & {82,2}                              & {5}                                      & {31}          & {65,58}        & {13}        & {12}          & {37,67}        & {34}        \\
+        {1000}    & {15}      & {85}                                & {82,62}                             & {5}                                      & {31}          & {64,28}        & {16}        & {16}          & {43,27}        & {32}        \\
+        {1000}    & {30}      & {85}                                & {83,13}                             & {4}                                      & {31}          & {64,06}        & {17}        & {13}          & {41,79}        & {34}        \\
+        {1500}    & {10}      & {84}                                & {81,27}                             & {4}                                      & {33}          & {67,5}         & {14}        & {12}          & {39,04}        & {34}        \\
+        {1500}    & {15}      & \textbf{85}                         & \textbf{84,11}                      & \textbf{4}                               & {35}          & {69,05}        & {12}        & {17}          & {44,98}        & {31}        \\
+        {1500}    & {30}      & {84}                                & {80,93}                             & {4}                                      & \textbf{36}   & \textbf{68,97} & \textbf{15} & {14}          & {42,99}        & {33}        \\
+        {2000}    & {10}      & {83}                                & {82,69}                             & {4}                                      & {36}          & {69,11}        & {13}        & {13}          & {41,85}        & {34}        \\
+        {2000}    & {15}      & {84}                                & {81,95}                             & {4}                                      & {34}          & {66,94}        & {14}        & {14}          & {38,79}        & {33}        \\
+        {2000}    & {30}      & {85}                                & {82,8}                              & {4}                                      & {32}          & {65,52}        & {16}        & {18}          & {45,41}        & {32}        \\
+        {3000}    & {10}      & {82}                                & {79,83}                             & {4}                                      & {34}          & {67,61}        & {12}        & {12}          & {38}           & {34}        \\
+        {3000}    & {15}      & {85}                                & {81,8}                              & {4}                                      & {32}          & {65,64}        & {15}        & {13}          & {38,52}        & {35}        \\
+        {3000}    & {30}      & {83}                                & {80,6}                              & {3}                                      & {30}          & {63,5}         & {16}        & {12}          & {38,45}        & {34}        \\
         \bottomrule
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) of the masked caption generator: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked caption generator: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
     \label{tab:results:masked-re-generator}
@@
 
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
-The final peak stayed exactly the same and the effects were therefore not studied deeper.
-% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
-% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
+Consequently, the results are much less accurate.
+This applies to all datasets, even though it is most apparent for the \emph{CLEVR color} dataset.
+While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the \emph{Dale-2} dataset, the model only reaches 36\% for the \emph{Dale-5} dataset and 29\% for the \emph{CLEVR color} dataset.
+However, these scores are all well above the random baselines of 50\%, 20\% and 10\% respectively.
+Moreover, the referring expression generated for the \emph{Dale-2} dataset are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
+Features of both target object and distractor are therefore extracted and associated with the vocabulary.
+Striking in these results is also the \emph{non-target accuracy}, which is much higher than for the \emph{bounding box RE generator}.
+For the \emph{Dale-2} dataset, it arrives at 11\% in one configuration, for the \emph{Dale-5} dataset, the model describes for at least 12\% of the samples a distractor, for one configuration even in over 25\% of the cases.
+The model is most uncertain, which object to describe when exposed to the \emph{CLEVR color} dataset, where in almost all configurations, it describes a distractor every third sample.
+This shows that the model struggles to learn, which of the objects in the scene is the target object, more specifically, to combine the whole image with the masked image.
+\cmtDK[inline]{compare to basic RE generator}
+Yet, the model is able to extract the attributes from the shown objects.
+This can be seen, when summing up the \emph{overall accuracy} and the \emph{non-target accuracy}.
+This number represents in how many cases any of the shown objects were described.
+For the \emph{Dale-2} dataset, this number lies at 90\% for the best configuration, at 51\% for the \emph{Dale-5} dataset and at 56\% for the \emph{CLEVR color} dataset.
+While being still lower than the same metric for the \emph{bounding box RE generator}, the model fares well in comparison for the \emph{Dale} datasets given the highly increased complexity of the task.
 
-In conclusion, it can be said that the model is able to extract discriminative features from the shown bounding boxes and produce referring expressions.
-However, the result highly depends on the amount of distractors and the resulting need of more discriminative features to describe the target object.
-While the shape is easily identified, the model has bigger problems to identify the color and especially the size of the target object.
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rr|cc|c|ccc|c|c}
+        \toprule
+                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {68,63} & {76}    & \textbf{72,31} & {98,41} & {98,18}    & {98,92}  & \textbf{98,5}  & {94,77} \\
+                                         & {Recall}    & {30,17} & {34,55} & \textbf{32,36} & {98,7}  & {98,33}    & {98,47}  & \textbf{98,5}  & {98,64} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {49,47} & {57,3}  & \textbf{53,38} & {87,09} & {83,26}    & {84,78}  & \textbf{85,04} & {64,99} \\
+                                         & {Recall}    & {24,67} & {26,7}  & \textbf{25,69} & {83,66} & {85,28}    & {86,03}  & \textbf{84,99} & {77,87} \\\midrule
+        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100} & {100} & {99,7} & \textbf{99,9} & {100}   \\
+                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {99,85} & {99,85}    & {100}    & \textbf{99,9}  & {100}   \\
+        \bottomrule
+    \end{tabular}
+    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
+    \label{tab:results:bb-re-generator_size-shape}
+\end{table}
 
-The sum of the two columns for the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-% SD: Not shown in Table 5
-% DK: sum of two columns (addressed; done)
-This means that the models produce almost always correct descriptions of one of the objects in the image.
-Moreover, these descriptions are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and associated with the vocabulary.
-% SD: But this is no-surprise since we have fixed the slots.
-% DK: but this applies to all padding methods. (done)
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object rises with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rr|cccccccc|c}
+        \toprule
+                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {88,89} & {85,9}  & {81,82} & {85,53} & {79,76} & {84,52}  & {87,76} & {87,36}  & \textbf{85,19} \\
+                                         & {Recall}    & {89,89} & {83,75} & {87,8}  & {79,27} & {93,06} & {76,34}  & {90,53} & {81,72}  & \textbf{85,3}  \\\midrule
+        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {69,51} & {72,73} & {73,48} & {66,27} & {63,23} & {64,32}  & {67,98} & {70,76}  & \textbf{68,53} \\
+                                         & {Recall}    & {72,43} & {71,29} & {73,48} & {60,77} & {74,6}  & {74,87}  & {71,13} & {73,57}  & \textbf{71,52} \\\midrule
+        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                           \\\textbf{color}} & {Precision}           & {37,63} & {39,23} & {22,09} & {20,23} & {33,33} & {22,97} & {24,9} & {43,96} & \textbf{30,54} \\
+                                         & {Recall}    & {29,44} & {42,68} & {27,48} & {22,22} & {25,44} & {26,53}  & {26,23} & {37,14}  & \textbf{29,64} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
+    \label{tab:results:bb-re-generator_color}
+\end{table}
 
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image, whether it is the target object or a distractor.
-% SD: ?
-% DK: added subclause (done)
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
+% With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
+% Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
+% The accuracy stays at the same value.
+% In contrast, the non-target accuracy is increased by 5\% points.
+% The model is more likely to identify a wrong object.
+% Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
+% This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
+% For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
+% When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
+% This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
 % SD: Not in the table?
@@
 
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
+There are two main explanations for the big difference between the datasets.
+First, as already seen in the previous experiments, a bigger number of distractors confuses the model more where to focus on.
+In the \emph{Dale-2} dataset, there are only two possibilities, while the four distractors in the \emph{Dale-5} dataset and up to nine distractors in the \emph{CLEVR color} give the model a bigger choice.
 % SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated.

### 2023-11-08 thesis/sections/5-grounding.tex
results

@@
 In the following paragraphs, the results of the \textbf{basic RE generator} as well as the \textbf{masked RE generator} are evaluated.
-Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 30 epochs.
+Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 40 epochs.
 Compared to the \emph{bounding box RE generator}, the task differs substantially.
@@
         {1500}   & {15}     & \textbf{85}                         & \textbf{84,11}                      & \textbf{4}                               & {35}          & {69,05}        & {12}        & {17}          & {44,98}        & {31}        \\
-        {1500}    & {30}      & {84}                                & {80,93}                             & {4}                                      & \textbf{36}   & \textbf{68,97} & \textbf{15} & {14}          & {42,99}        & {33}        \\
-        {2000}    & {10}      & {83}                                & {82,69}                             & {4}                                      & {36}          & {69,11}        & {13}        & {13}          & {41,85}        & {34}        \\
+        {1500}   & {30}     & {84}                                & {80,93}                             & {4}                                      & {36}          & {68,97}        & {15}        & {14}          & {42,99}        & {33}        \\
+        {2000}   & {10}     & {83}                                & {82,69}                             & {4}                                      & \textbf{36}   & \textbf{69,11} & \textbf{13} & {13}          & {41,85}        & {34}        \\
         {2000}   & {15}     & {84}                                & {81,95}                             & {4}                                      & {34}          & {66,94}        & {14}        & {14}          & {38,79}        & {33}        \\
@@
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked caption generator: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked caption generator after 40 epochs: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
     \label{tab:results:masked-re-generator}
@@
 
+Looking at the effects of the LSTM output size $LSTM_o$ and the LSTM embedding size $LSTM_e$, a difference between the \emph{Dale} datasets and the \emph{CLEVR color} dataset can be identified.
+A higher $LSTM_o$ tends to increase the performance of the model, when presented with the \emph{Dale} datasets until a size of 1500 to 2000 dimensions.
+When it is further increased the model starts to perform worse.
+Furthermore, with a low $LSTM_o$, the model is more likely to identify and describe a distractor, while a higher $LSTM_o$ helps the model to focus on the target object.
+The $LSTM_e$ however doesn't seem to have a big consistent effect on the results.
+With an $LSTM_o$ of around 1500 to 2000, the model seems to fare slightly better with higher $LSTM_e$ of 15 or 30 dimensions.
+For the \emph{CLEVR color} dataset, the results look different.
+Opposed to the former datasets, a lower $LSTM_o$ helps the model to describe the correct target object.
+On the other side, the \emph{non-target accuracy} stays relatively constant independently of the variables, in other words
+
 \begin{table}[ht]
@@
     \end{tabular}
-    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+    \caption{Precision and Recall in \% of <pad>, size and shape tokens in the best configuration for each dataset (marked in bold in Table \ref{tab:results:masked-re-generator}).
 The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
-    \label{tab:results:bb-re-generator_size-shape}
+    \label{tab:results:masked-re-generator_size-shape}
 \end{table}
@@
     \end{tabular}
-    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
-    \label{tab:results:bb-re-generator_color}
+    \caption{Precision and Recall in \% of color tokens in the best configuration for each dataset (marked in bold in Table \ref{tab:results:masked-re-generator}).
+The column \textbf{color} shows the average across all colors.}
+    \label{tab:results:masked-re-generator_color}
 \end{table}

### 2023-11-08 thesis/sections/5-grounding.tex
results

--- 2023-11-07 thesis/sections/5-grounding.tex
@@
 By this it measures the false negative generated captions.
-For this we generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
+For this, referring generations are generated for all the non-target objects and distractors in the images using the GRE-algorithm.
+Importantly to notice is that the referring expressions may not uniquely identify a distractor, since multiple distractors with the same shape, color and size are allowed.
 If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.

### 2023-11-08 source/mlt/preexperiments/models.py
coordinate predictor

@@ -296,6 +296,62 @@ class DaleAttributeCoordinatePredictor(nn.Module):
         return coordinates
 
 
+class MaskedDaleAttributeCoordinatePredictor(nn.Module):
+    """
+    Output:
+     - x and y coordinates of target object
+
+    Input:
+     - image
+     - attributes (shape, size, color)
+     - center coordinates of all objects
+    """
+
+    def __init__(
+        self,
+        vocab_size,
+        image_embedding_dimension,
+        embedding_dim,
+        encoder_out_dim,
+        image_encoder: ImageEncoder,
+        masked_image_encoder: MaskedImageEncoder,
+        coordinate_classifier: CoordinateClassifier,
+        *_args,
+        **_kwargs
+    ) -> None:
+        super().__init__()
+        self.image_encoder = image_encoder
+        self.masked_image_encoder = masked_image_encoder
+        self.reduction = nn.Sequential(
+            nn.Flatten(), nn.LazyLinear(image_embedding_dimension)
+        )
+
+        self.embedding = nn.Embedding(vocab_size, embedding_dim)
+        self.lstm = nn.LSTM(embedding_dim, encoder_out_dim, batch_first=True)
+
+        self.coordinate_classifier = coordinate_classifier
+
+    def forward(self, data):
+        image, attribute_tensor, _, masked_image, *_ = data
+
+        encoded_image = self.image_encoder(image)
+        encoded_masked_image = self.masked_image_encoder(masked_image)
+        concatenated = torch.cat(
+            (encoded_image, encoded_masked_image),
+            dim=1,
+        )
+
+        reduced = self.reduction(concatenated)
+
+        embedded = self.embedding(attribute_tensor)
+        _, (hidden_state, _) = self.lstm(embedded)
+
+        concatenated = torch.cat((reduced, hidden_state.squeeze()), dim=1)
+        coordinates = self.coordinate_classifier(concatenated)
+
+        return coordinates
+
+
 class MaskedCoordinatePredictor(nn.Module):
     """
     Output:

### 2023-11-10 thesis/sections/1-introduction.tex
introduction

@@
 \label{sec:introduction}
-\cmtDK[inline]{lead from grounding to referring expressions to language games}
-The study of emergent languages, in other words artificial languages created by agents to communicate and solve a goal are studied deeply in the recent years \citep{Kirby2002,Kirby2008,Steels2009,Lazaridou2017,Baroni2020,Baroni2022}.
-The motivation behind this is twofold.
-First, the study of the emergence of artificial language can help to shed light into why and how natural languages evolved.
-This is done in a controlled environment, where researchers can regulate exactly, which information the agents receive, how and what they are able to communicate and which goals they are trained on.
-This approach allows a deep study of the constraints and restrictions that are necessary for language and meanings of symbols to evolve.
-Insights in this field might give indications, under which circumstances natural language evolves.
+Language is a complex system that enables humans to communicate and convey meaning.
+However, the mere existence of words and grammatical structures is not enough to guarantee effective communication.
+Without grounding, language remains detached from the physical world, making it difficult for individuals to comprehend and interpret linguistic expressions accurately.
+Grounding bridges the gap between abstract linguistic representations and concrete experiences, providing a shared context and reference point for language users \citep{Roy2002,Bisk2020,Bender2020}.
+By linking language to sensory perceptions, physical experiences, and the situational context, grounding enhances comprehension, facilitates communication.
+However, one challenge lies in how arbitrary symbols in language can be connected to real-world objects and concepts and acquire meaning, described as the symbol grounding problem by \citet{Harnad1999}.
+Artificial agents can be taught grounding in different ways.
+One approach is to show the agent an image of the scene as visual input together with a textual description of what is to be grounded in this scene.
+This can be for example a description of the whole scene, certain visible objects or also actions that are happening \citep{Ahrens2022,Ilinykh2022,Lu2016,Mitchell2013}.
+By doing this, the agents can learn to identify the same descriptions and symbols for similar visual input.
+In other approaches, the agents can learn to ground language in visual input through interaction and dialogue, either with humans or other agents.
+Thus, agents might develop certain strategies to ground language efficiently \citep{Dobnik2018,Dobnik2021,Zarriess2019}.
 
-Secondly, emergent languages can have benefits over predefined communication protocols.
-Currently, machines communicate, by using protocols based on predefined rules of how messages can be built up.
-These are developed for specific scenarios and goals and don't allow flexible usages if the goal or the environment, machines act in are changing.
-Furthermore, these protocols are often based on readability for humans and include overhead that is not necessary to communicate a specific meaning, but is rather used for transparency.
-A language that emerges, rather than being predefined could mitigate these problems.
+A central part in many of the ways to learn to ground are referring expressions.
+Referring expressions are used to point out a particular subset of entities from a set of similar entities or the surrounding context.
+By that, they can take many forms as noun phrases (e.g.
+'the large table', 'the second chair from the right'), proper nouns (e.g.
+'Francesca', 'the Big Ben') or complete descriptions (e.g.
+'the men that are wearing glasses').
+When referring to objects, they often include inherent attributes of these objects as their color or shape.
+Furthermore, spatial terms play a vital role to refer to and disambiguate objects \citep{Regier1996,  Dobnik2013, Dobnik2021,Ghanimifard2017, Ramisa2015}.
+
+One problem that arises is the question of ambiguity in language and referring expressions.
+Natural language is often used in an underspecified way.
+Interlocutors rely on pragmatic and semantic context to communicate.
+Given for example the following situation:
+Two people are in a room with two chairs.
+If one person asks for 'a chair', the second person would give him the chair that is closer to him.
+In other situations the second person would give him his preferred chair even if it's further away.
+The referring expression is the same in both contexts, but refers to two different entities.
+The difference lies in the pragmatic context that both interlocutors are aware of.
+In case this context is not shared, referring expressions might be misunderstood, because of the underspecification.
+
+Furthermore, the real world is characterized by its continuous nature.
+On the other side, natural language is a set of discrete symbols.
+In the process of mapping this continuum to the discrete symbols, information is lost, since the whole context of the real world can't be captured by the natural language.
+By this, symbols get ambiguous and refer to multiple concepts in the world.
+One example are the shades of color \citep{Zaslavsky2018}.
+Discrete English terms like 'red', 'green' or 'blue' might refer to an infinite number of colors in the real world.
+
+To deal with this, humans rely on communicative protocols to disambiguate referring expressions.
+\citet{Dale1995}, for instance, describe an incremental algorithm how referring expressions are generated.
+The goal of the algorithm is to construct referring expressions incrementally, taking into account the salience of different properties and abstracting away from differences in lexical choice.
+It works by gradually building up a referring expression based on the properties of the object being referred to.
+Hereby, it considers the most salient or distinguishing properties first and then adds additional properties as needed.
+This helps to create concise, effective and disambiguous referring expressions.
+\cmtDK[inline]{co-reference}
+
+Meaning of symbols and referring expression emerge through dialogue between interlocutors \citep{Wittgenstein1953,Clark1986}.
+Moreover, \citet{Wittgenstein1953} introduces the concept of \emph{language games}: small parts of conversation between interlocutors.
+In these language games, the emerged meaning of the referring expressions depends on their context.
+This reasoning was taken up in training artificial agents to learn to refer to entities.
+Hereby, agents need to communicate arbitrary symbols to each other in multiple turns to solve a given task.
+By solving the task, they assign meaning to these symbols and an artificial language emerges.
+This setup allows to control the architecture of the agents, the information the agents receive, how and what they are able to communicate and which goals they are trained on.
+By this, their behavior and the emergence of referring expressions can be studied precisely and consistently.
+In early research, agents and their communication were studied in the context of robotics and rule-based systems \citep{Steels2009,Roy2002,Kirby2002,Kirby2008}, while in current research, agents are based on deep neural networks \citep{Lazaridou2017,Baroni2020,Baroni2022,Kottur2017}
+Furthermore, a focus lies on the nature of the emerged language.
 Agents can for instance learn a compositional language that allows them to combine already learned symbols to create new meaning, when presented with a changed environment and unseen situations \citep{Kharitonov2020,Lazaridou2018,Gupta2020}.
@@
 
-Different studies conclude that meaning of symbols and sentences in language can only be bound to the real world, both in the literal sense, but also for abstract concepts \citep{Harnad1999,Bender2020,Bisk2020}.
-They need to be grounded.
-One way of grounding meaning can be achieved in multimodal systems, where models need to combine language with another modality for instance visual input.
-Hereby, a task can be to learn how to refer to objects or actions in the visual input, using referring expressions.
-Both generating referring expressions, based on the visual input and understanding given referring expressions by linking them to the visual input ground the meaning of the used language in vision.
-Following these arguments, also artificial emergent languages need to be grounded that the used symbols contain meaning.
 % SD: I would turn the order of this around, it would make the motivation clearer.
@@
 \subsection{Research Questions}
-In this thesis, a deeper look is taken into how agents can ground their emerged language in vision.
+In this thesis, a deeper look is taken into how agents can ground their emerged language in visual input.
 The focus hereby lies on referring expressions and how agents are able to generate and understand them.
@@
       % \item Are cooperative agents in language games able to generate and understand artificial referring expressions based on visual input?
-      \item What are the limits of the agent architectures and input representations on learning successfully grounding referring expressions through language games?
+      \item What are the limits of the agents' architectures and input representations on learning successfully grounding referring expressions through language games?
             % SD: What are the limits of the agent architecture and and input representation on learning successful grounding referring expressions through language games?
@@
 The datasets that are created in this thesis aim to reduce this bias and provide all information about the scene and objects in the image.
+In fact, the artificial creation of the dataset allows controlling precisely bias in each scene.
 % SD: We also know the ground truth about the scenes as we know the function that generated them.
-% DK: TODO
+% DK: (done)
 
 Secondly, the thesis evaluates these datasets, by training separate models to generate and understand referring expressions, describing objects in the images.
-By doing this, it is tested if first the models are able to extract useful visual information that doesn't rely on bias and latent patterns in the dataset and secondly shows the impact of different levels of ambiguity in the datasets on the performance of generating and understanding of referring expressions.
+By doing this, it is tested if first the models are able to extract useful visual information that doesn't rely on bias and latent patterns in the textual information in the dataset and secondly shows the impact of different levels of ambiguity in the datasets on the performance of generating and understanding of referring expressions.
 % SD: We can control the properties of these datasets and we can introduce as much bias and ambiguity as we see firm in each experiment to compare with the natural datasets.
-% DK: TODO
+% DK: (done)
 
-Lastly, the thesis brings the separate tasks of generating and understanding of referring expressions together into on single task.
+Lastly, the thesis brings the separate tasks of generating and understanding of referring expressions together into one single task.
+This process resembles the learning of referring expressions in natural language \citep{Clark1986}.
 % SD: Which is how this is done in practice, cf. referring as a collaborative process, the paper by Clark.
-% DK: TODO
+% DK: (done)
 In language games, one agent needs to extract visual information from an image and generate a referring expression that is sent to the second agent.
@@
 The focus of this thesis is the study of referring expressions.
-Referring expressions can hereby be based on attributes of objects, as well as their spatial relations towards other objects.
-In the present study, only attribute relations are studied.
+Referring expressions can hereby be based on inherent attributes of objects, as well as their spatial relations towards other objects.
+In the present study, only the relations of the inherent attributes are studied.
 % SD: What are those?
-% DK: TODO
+% DK: (done)
 Spatial relations add another level of complexity and may be studied in future work.
 % SD: What kind of complexity?
-% DK: TODO
+% DK: (done)
 Furthermore, the emerged language will be interpreted by comparing it to referring expressions in natural language.
@@
 % SD: Leave for the last chapter, limitations and future work.
-% DK: TODO
+% DK: (done)

### 2023-11-11 thesis/sections/1-introduction.tex
bibliography

@@
 One approach is to show the agent an image of the scene as visual input together with a textual description of what is to be grounded in this scene.
-This can be for example a description of the whole scene, certain visible objects or also actions that are happening \citep{Ahrens2022,Ilinykh2022,Lu2016,Mitchell2013}.
+This can be for example a description of the whole scene, certain visible objects or also actions that are happening \citep{Ahrens2022,Ilinykh2022,Lu2017,Mitchell2013}.
 By doing this, the agents can learn to identify the same descriptions and symbols for similar visual input.
 In other approaches, the agents can learn to ground language in visual input through interaction and dialogue, either with humans or other agents.
-Thus, agents might develop certain strategies to ground language efficiently \citep{Dobnik2018,Dobnik2021,Zarriess2019}.
+Thus, agents might develop certain strategies to ground language efficiently \citep{Dobnik2018a,Dobnik2021,Zarriess2019}.
 
@@
 Moreover, \citet{Wittgenstein1953} introduces the concept of \emph{language games}: small parts of conversation between interlocutors.
-In these language games, the emerged meaning of the referring expressions depends on their context.
+In these language games, their context defines which meaning emerges for the referring expressions.
 This reasoning was taken up in training artificial agents to learn to refer to entities.

### 2023-11-11 thesis/sections/2-background.tex
describe experimental setups
bibliography

@@
 \cmtDK[inline]{importance of grounding and what it is}
-
-\subsection{Referring expressions}
-\cmtDK[inline]{what are referring expressions and how are they studied (generation, understanding, ...)}
-\cmtDK[inline]{incremental GRE algorithm from Dale}
 Even though, recent large language models (LLMs) such as BERT \citep{Devlin2019} or GPT-3 \citep{Brown2020} produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
@@
 \citet{Bisk2020} argues that a multimodal approach, including for instance perception as well as social context, is needed to learn meaning in a broader context.
-One added modality to ground language is often vision. \cmtDK{cite}
+One added modality to ground language is often vision.
 Hereby, the model needs to learn how to associate linguistic concepts in text corpora with features, extracted from visual input.
 For instance, a model can learn to associate the noun "dog" with an animal seen in an image or associate the action of "jumping over" with the animal being above an object.
+
+\subsection{Referring expressions}
+\cmtDK[inline]{what are referring expressions and how are they studied (generation, understanding, ...)}
+\cmtDK[inline]{incremental GRE algorithm from Dale}
 

### 2023-11-11 source/mlt/preexperiments/run_experiments.py

@@ -2,7 +2,7 @@ import itertools
 import subprocess
 
 options = [
-    # "--dataset_base_dir=/scratch/guskunkdo",
+    #"--dataset_base_dir=/scratch/guskunkdo",
     "--dataset_base_dir=/home/dominik/Development/",
     # "--lr=0.0002",
     "--epochs=40",
@@ -15,10 +15,10 @@ options = [
 ]
 
 variables = {
-    "--dataset": ["dale-2"],
-    # "--dataset": ["dale-2", "dale-5", "colour"],
-    "--image_embedding_dimension": [300, 500],
-    # "--image_embedding_dimension": [50, 100, 300, 500],
+    #"--dataset": ["dale-2"],
+    "--dataset": ["dale-5", "colour"],
+    #"--image_embedding_dimension": [300, 500],
+    "--image_embedding_dimension": [50, 100, 300, 500],
     "--coordinate_classifier_dimension": [1024, 2048],
     # "--decoder_out_dim": [100, 500, 1000],
     "--encoder_out_dim": [1000, 1500],

### 2023-11-11 thesis/sections/1-introduction.tex
backgorund

@@
 By linking language to sensory perceptions, physical experiences, and the situational context, grounding enhances comprehension, facilitates communication.
-However, one challenge lies in how arbitrary symbols in language can be connected to real-world objects and concepts and acquire meaning, described as the symbol grounding problem by \citet{Harnad1999}.
+However, one challenge lies in how arbitrary symbols in language can be connected to real-world objects and concepts and acquire meaning, described as the symbol grounding problem by \citet{Harnad1990}.
 Artificial agents can be taught grounding in different ways.

### 2023-11-11 thesis/sections/2-background.tex
backgorund

@@
 \subsection{Grounding}
-\cmtDK[inline]{importance of grounding and what it is}
-Even though, recent large language models (LLMs) such as BERT \citep{Devlin2019} or GPT-3 \citep{Brown2020} produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
+Symbols of a symbolic system such as words in a natural language are typically processed and manipulated in a purely formal manner, without direct reference to their real-world concepts.
+The question arises, how these symbols can acquire meaning and be connected to the real world without any external interpreter, in other words how the symbols can be grounded in the real world.
+This is known as the symbol grounding problem \citep{Harnad1990}.
+Hereby, \citet{Harnad1990} describes that a high-order \emph{symbolic representation} needs to be grounded bottom-up in two non-symbolic representations.
+\emph{Iconic representations} refer to analogue representations of sensory input like for instance vision in the retina.
+These representations are mapped to \emph{categorical representations} which are filtered clusters of the iconic representation.
+These categories are assigned so-called elementary symbols, in which symbolic representation are grounded.
+
+In computational models this problem is tried to be solved in multiple ways.
+Recent large language models (LLMs) such as BERT \citep{Devlin2019} or GPT-3 \citep{Brown2020} are often trained on big amounts of text corpora produce impressive results on many tasks in Natural Language Processing.
+Meaning of symbols emerges through relating it to its surrounding context of other symbols it appears in.
 In this way, their only way of learning, how to generate language as well as solve these tasks is to find patterns of how words and sentences are used in these large corpora.
-Many researchers criticize this approach, by arguing that these model can't learn meaning just by learning on text.
+This approach is however criticized by many researchers, since the meaning of symbols is still not connected to the real world and only relies on other symbols in an infinite recursion.
+
 In \citep{Bender2020}, the authors argue that meaning is bound to a communicative intent.
@@
 % DK: first part included, referring expressions are included below (done)
-This grounding is missing, when models only train on abstract textual representations of the world.
+This grounding is missing when models only train on abstract textual representations of the world \citep{Regier1996,Landau1998}.
+
 \citet{Bisk2020} argues that a multimodal approach, including for instance perception as well as social context, is needed to learn meaning in a broader context.
-One added modality to ground language is often vision.
+One added modality to ground language is often visual input.
 Hereby, the model needs to learn how to associate linguistic concepts in text corpora with features, extracted from visual input.
 For instance, a model can learn to associate the noun "dog" with an animal seen in an image or associate the action of "jumping over" with the animal being above an object.
+\cite{Roy2002} for example studies how an artificial agent can ground spatial relations with a visual scene of geometric shapes.
+Hereby, the agent builds up hierarchical knowledge that is used to produce correct description in a rule-based manner.
+In more recent research, complete natural language utterances as well as the visual representations are mapped into an embedding space.
+By aligning both representations, natural language can be connected to a second modality.
+Other research explores how to ground smaller parts of the utterances, as phrases and words and combine them compositional \citep{Larsson2018,Kollar2010}.
+This might allow artificial machines to learn grounded concepts more general.
+
+Furthermore, artificial agents can learn to ground language through dialogue, either with human tutors or with other artificial agents.
+\citet{Skocaj2011} for example teach a robot different objects and concepts.
+By asking questions and participating in the dialogue with a human tutor, the robot builds up and verifies beliefs, and connects them to language and its visual input.
+\citet{Lauria2001} follow a similar approach where a robot is taught to follow instructions to generate a map of its physical surroundings.
+By incorporating dialogue and deep-level reasoning questions the learning gains are enhanced.
+How artificial agents can ground language in dialogue between each other will be studied in this thesis.
+Section \ref{sec:background:language-games} gives an insight in the state of the research.
 
@@
 More specifically, the model learns to connect symbols to a non-symbolic representation.
-This is known as the symbol grounding problem \citep{Harnad1999,Roy2002}.
+This is known as the symbol grounding problem \citep{Harnad1990,Roy2002}.
 Hereby, the question is asked how arbitrary symbols can be connected to the representation they actually stand for.
@@
 \subsection{Language Games}
+\label{sec:background:language-games}
 \cmtDK[inline]{language games combine generation and understanding}

### 2023-11-12 thesis/sections/2-background.tex
background

@@
 \subsection{Referring expressions}
-\cmtDK[inline]{what are referring expressions and how are they studied (generation, understanding, ...)}
-\cmtDK[inline]{incremental GRE algorithm from Dale}
+\label{sec:background:re}
+A specific case of grounding linguistic information in visual input are referring expressions.
+While a general linguistic expression can be grounded in multiple regions of a visual scene, referring expressions are expected to refer to a unique region \citep{Sanchez2022}.
+Given for example a visual scene with multiple pens lying on a table, the word 'pen' can be grounded in all visible objects.
+A referring expression like 'the red pen on the left' in contrast would only refer to one specific object.
+Referring expressions are however not limited to refer to one single entity, but might also refer to a group of objects as for instance with 'the green pens'.
+Still they refer to a specific subset of entities and concepts.
+
+Hereby, \citet{Krahmer2012} classify referring expressions into four categories:
+\emph{One-place predicates} refer to one single entity in the visual scene without.
+By this, the entity is identifiable without using another entity as basis.
+Often, the referring expressions includes intrinsic attributes of an object or absolute locations in the scene as in 'the red pen on the left'.
+\emph{Relational} referring expressions refer to entities via other entities in the scene.
+Accordingly, a relating between the target entity and distractors is described.
+This includes for example spatial relations like 'the pen behind the cup' or relations based on the entities' properties like 'the largest pen'.
+Referring expressions can also identify \emph{sets} of entities.
+Instead of uniquely referring to on single pen, the referring expression 'the green pens' refers to all objects that are pens and that are green.
+Finally, \emph{gradable} referring expressions involve gradable properties as opposed to absolute values.
+For example, given the referring expression 'the tall building', 'tall' is a gradable property because it can be measured on a scale.
+The success of the referring expression depends on how well it captures the intended level of height compared to other buildings.
+By this, vagueness is introduced and referring expressions depend on context, either present in the visual scene or pragmatic context.
+
+One of the main challenges of referring expressions lies in ambiguity.
+Ambiguity is introduced in two forms.
+First, there is an inherent difference in natural language and the real world it is grounded it, also described in the symbol grounding problem.
+The nature of the world is continuous, while natural language being a formal symbol system uses discrete symbols.
+In especially, natural language utilizes a very limited set symbols to represent the world.
+When trying to map continuous concepts to this limited set of discrete symbols, information nesecarilly gets lost.
+Apparent examples are the color systems in human languages.
+While color grades exist on an infinite spectrum in the real world, languages combine certain color grades into categories \citep{Zaslavsky2018}.
+The word 'red' for example refers to many different shades of red.
+Even more, the shades of color it actually refers to depends also on the pragmatic context and the people uttering it \citep{Monroe2017}.
+The same also applies to spatial references.
+References like 'near to' or 'left to' might refer to objects that are touching each other, but in other contexts also are kilometers away from each other.
+
+The other form of ambiguity is introduced through under-specification of the speakers (e.g. \citet{Dobnik2021}).
+When interlocutors interact in a dialogue, they share some common knowledge.
+The first interlocutor might presuppose that the second interlocutor knows a certain fact.
+Based on this, they might utter a referring expression that together with this shared fact uniquely identifies an entity.
+However, the referring expression on its own is under-specified and might refer to multiple entities.
+Given for example the following situation:
+Two people are in a room with their beverages standing on the table and Person A requests his beverage from Person B.
+A referring expression like 'the beverage close to me' would uniquely identify on of the beverages.
+Despite that, Person A might more likely use 'the beverage' to refer to theirs.
+On its own, the referring expression is under-specified and can refer to both of the present beverages.
+However, both persons share the fact that the targeted beverage belongs to person A, which disambiguates the referring expression in this context.
+In another context, the shared fact would be that Person A wants to try the beverage of Person B.
+Again, this fact would disambiguate the referring expression, but now refer to the second beverage.
+The referring expression is the same in both contexts, but refers to two different entities.
+This creates a challenge for computational models, that need to understand and generate referring expressions if they are only presented with the text, but not with the shared context.
+
+The research of referring expressions can be split into two fields: referring expression generation (REG) and referring expression comprehension (REC).
+In \textbf{referring expression generation}, computational models are trained to generate a referring expression that uniquely identifies an entity given some perceptual, often visual input.
+The research goes back into the 70s, where \citet{Winograd1972} developed an algorithm to generate referring expressions step by step, taking into account the context and the information available at each stage of the generation process.
+A similar approach is described in \citep{Dale1995}.
+Their incremental algorithm is based on the salience of the properties of the present target object and distractor objects.
+The most salient properties are considered first, and added to the referring expression until the referring expression is unambiguous.
+In other words, the object is described unambiguously using the lowest number of words.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.5\linewidth]{figures/CLEVR_dale-5.png}
+    \caption{The target object in this scene is the \emph{small turquoise sphere}. Using the incremental GRE-algorithm, the target object can be uniquely referred to as \emph{turquoise sphere}}
+    \label{fig:incremental-algorithm}
+\end{figure}
+
+Given for example the image in Figure \ref{fig:incremental-algorithm}.
+Objects in this scene have three attributes that we consider in the following order of salience: a shape, a color and a size.
+This order defines, which attributes can be left out, while still identifying the object uniquely.
+The target object in the given scene is the \emph{small turquoise sphere}.
+In the first step, the attribute with the highest salience, the shape, is selected from the target object.
+This produces the referring expression \emph{sphere}.
+In the next step, it is checked if there are also distractors, described by this referring expression.
+In our case, both cubes and the cylinder are already disambiguated, but the purple sphere can still be described by the produced referring expression.
+Therefore, the algorithm also adds the next lower salient attribute, the color to the referring expression, which produces the \emph{turquoise sphere}.
+With this, also the last remaining distractor is as well disambiguated, and the referring expression describes the target object uniquely.
+The size, the least salient property is therefore not necessary in the referring expression.
+This algorithm will be utilized in this thesis and will be referred to as \emph{the incremental GRE-algorithm}.
+
+Additional to generating an efficient disambiguating referring expressions, a challenge also lies in processing the visual input.
+This especially applies to spatial and geometric positions and relations of objects in a scene, where models struggle to extract geometric information from visual features \citep{Kelleher2017}.
 

### 2023-11-13 thesis/sections/2-background.tex
bibliography

@@
 
-Given for example the image in Figure \ref{fig:incremental-algorithm}.
+Given for example the scene in Figure \ref{fig:incremental-algorithm}.
 Objects in this scene have three attributes that we consider in the following order of salience: a shape, a color and a size.
@@
 This especially applies to spatial and geometric positions and relations of objects in a scene, where models struggle to extract geometric information from visual features \citep{Kelleher2017}.
+More complex scenes also involve multiple perspectives, where the models are tasked to produce referring expressions that depend on the spatial perspective of the speaker and the listener \citep{Ahrens2022,Lee2022}.
+In \citep{Dobnik2021}, visual dialogue is analyzed where speakers have different views on cups that are placed on a table.
+When speakers refer to a certain cup, they need to take the perspective of the listener to disambiguate the referring expression; for instance in the referring expression 'the cup on the left', \emph{left} is dependent on the speaker's and listeners' spatial position in relation to the cup.
 
-In a first step, models can combine linguistic knowledge with visual input, by referring to objects, seen in an image, with so called referring expressions.
-Hereby, the communicative intent as well as the broader context in which it is used determines which referring expressions are applied.
-By doing this, the models learn how to relate referring expressions against a representation of the environment.
-More specifically, the model learns to connect symbols to a non-symbolic representation.
-This is known as the symbol grounding problem \citep{Harnad1990,Roy2002}.
-Hereby, the question is asked how arbitrary symbols can be connected to the representation they actually stand for.
-
-To solve this problem, multiple challenges arise since language describes the world in a very under-specified way.
-For instance, sensory input for humans as well as for machines only show a very limited representation of the world.
-Vision needs to map a 3-d world into 2-d images.
-During this process, information about the world gets lost, as for example depth can't be perceived easily or objects might be obfuscated in certain perspectives.
-Following, language needs to be connected to a representation of the world that is already not complete.
-Secondly, referring expressions in language themselves are often under-specified.
-Given for example a table with five cups on top, an interlocutor is asked to retrieve one of the cups, referring to it by 'the red cup'.
-This referring expression on its own might be ambiguous and refer to multiple red cups on the table.
-The challenge now arises as how the interlocutor associates the correct cup with this referring expressions.
-This gets even more complex, when perspective is part of the referring expression as for example in 'the cup on the left' where \emph{left} is dependent on the speaker's and interlocutors' spatial position in relation to the cup \citep{Dobnik2021}.
+Current research focuses hereby on generating relational referring expressions, which add more complexity to the task \citep{Ghanimifard2017,Ghanimifard2019,Ramisa2015,Liu2022a}.
+Models need to first extract visual features of multiple objects from the scene.
+Of these extracted objects, a meaningful landmark object needs to be found which helps to refer to the target object.
+Finally, they need to find proper relations between them to describe the target object.
+Relations can be mostly based on the geometric and spatial relation of two objects (e.g.
+above/below), while the function of objects play a stronger role in other relations (e.g.
+over/under) \citep{Coventry2005,Dobnik2013}.
 
@@
 % DK: done
-The learning of referring expressions is therefore split in two fields.
-In the referring expression generation, models learn how to produce referring expressions, when presented with visual input.
-In the referring expression understanding, sometimes referred to as coreference, models are trained on interpreting referring expressions and link them to visual input.
 
-A major challenge in natural language processing is, how machines can
+Referring generation tasks can be hard to evaluate, since often there are many possible referring expressions that are possible and uniquely identify the target object \citep{Mao2016}.
+Instead, the task can be turned around: A model needs to identify an object when presented with a referring expression.
+This is called \textbf{referring expressions comprehension}.
+Many challenges remain similar to the REG task, since the model needs to extract disambiguating visual features from the scenes as well.
+However, the goal is typically to select the best matching region of a given choice in a visual scene \citep{Qiao2020}.
+\cmtDK[inline]{describe methods in paper}
+
+Closely related to the REC task is visual question and answering (VQA).
+In these tasks, a model is given a question about a visual scene.
+The goal is to find an answer in the scene and generate an answer.
+Central to this is to extract referring expressions from the question and ground them in the visual scene.
+In \citep{Johnson2017}, a model is for example shown an image similar to Figure \ref{fig:incremental-algorithm} and asked the question 'Is there a blue cube with the same size as the purple cylinder?'.
 

### 2023-11-14 source/mlt/image_loader.py
memory efficiency

@@ -39,16 +39,19 @@ class FeatureImageLoader(ImageLoader):
         with h5py.File(feature_file, "r") as f:
             feature_data_set = f["features"]
             self.image_size = feature_data_set.attrs["image_size"]
-            self.features = list(feature_data_set)
 
     def get_image(self, image_id):
         image_index = int(image_id[-6:])
+
+        with h5py.File(self.feature_file, "r") as f:
+            feature_data_set = f["features"]
+            features = feature_data_set[image_index]
 
         image = Image.open(os.path.join(self.image_dir, image_id + ".png")).convert(
             "RGB"
         )
 
-        return image, torch.from_numpy(self.features[image_index]), self.image_size
+        return image, torch.from_numpy(features), self.image_size
 
     def __repr__(self) -> str:
         return f"{self.__class__.__name__}({self.feature_file=}, {self.image_dir=})"

### 2023-11-14 source/mlt/language_games/data_readers.py

@@ -16,7 +16,7 @@ from mlt.preexperiments.data_readers import (
     CoordinatePredictorDataset,
     CoordinatePredictorSample,
 )
-from mlt.util import Persistable
+from mlt.util import Persistable, load_tensor
 from torch.utils.data import DataLoader, Dataset, Subset
 
 
@@ -343,7 +343,14 @@ class DaleReferentialGameBatchIterator(GameBatchIterator):
 
 class CaptionGeneratorGameDataset(CaptionGeneratorDataset):
     def __getitem__(self, index):
-        return self.samples[index]
+        with h5py.File(self.file, "r") as f:
+            return CaptionGeneratorSample(
+                image_id=str(f["image_id"][index], "utf-8"),
+                image=load_tensor(f["image"][index]),
+                caption=load_tensor(f["caption"][index]),
+                masked_image=load_tensor(f["masked_image"][index]),
+                non_target_captions=load_tensor(f["non_target_captions"][index]),
+            )
 
 
 class CaptionGeneratorGameBatchIterator(GameBatchIterator):

### 2023-11-14 source/mlt/preexperiments/data_readers.py

@@ -826,12 +826,23 @@ class CaptionGeneratorDataset(Dataset, Persistable):
     """
 
     def __init__(
-        self, captioner: Captioner, samples: list[CaptionGeneratorSample]
+        self,
+        captioner: Captioner,
+        samples: list[CaptionGeneratorSample],
+        save_to,
     ) -> None:
         super().__init__()
 
         self.captioner = captioner
         self.samples = samples
+        self.file = save_to
+
+        with h5py.File(save_to, "r") as f:
+            self.num_samples = len(f["image"])
+
+        if len(samples) != 0:
+            self.save(save_to)
+            self.samples = []
 
     @classmethod
     def load(
@@ -840,6 +851,7 @@ class CaptionGeneratorDataset(Dataset, Persistable):
         image_loader: ImageLoader,
         max_number_samples,
         captioner: Captioner,
+        save_to: str,
         *_args,
         image_masker: ImageMasker = None,
         preprocess=ResNet101_Weights.DEFAULT.transforms(),
@@ -906,23 +918,23 @@ class CaptionGeneratorDataset(Dataset, Persistable):
         print()
         print("loaded data.")
 
-        return cls(captioner, samples)
+        return cls(captioner, samples, save_to)
 
     def __getitem__(self, index):
-        sample = self.samples[index]
-        return (
-            (
-                sample.image,
-                sample.caption,
-                sample.non_target_captions[:, 1:],
-                sample.masked_image,
-            ),
-            sample.caption[1:],
-            sample.image_id,
-        )
+        with h5py.File(self.file, "r") as f:
+            return (
+                (
+                    load_tensor(f["image"][index]),
+                    load_tensor(f["caption"][index]),
+                    load_tensor(f["non_target_captions"][index])[:, 1:],
+                    load_tensor(f["masked_image"][index]),
+                ),
+                load_tensor(f["caption"][index])[1:],
+                str(f["image_id"][index], "utf-8"),
+            )
 
     def __len__(self) -> int:
-        return len(self.samples)
+        return self.num_samples
 
     def save(self, file_path):
         with h5py.File(file_path, "w") as f:
@@ -951,30 +963,30 @@ class CaptionGeneratorDataset(Dataset, Persistable):
     @classmethod
     def load_file(cls, file_path):
         with h5py.File(file_path, "r") as f:
-            samples = [
-                CaptionGeneratorSample(
-                    **dict(
-                        zip(
-                            [
-                                "image",
-                                "image_id",
-                                "caption",
[…]

### 2023-11-14 source/mlt/preexperiments/train.py

@@ -515,19 +515,20 @@ if __name__ == "__main__":
         dataset = model_name.dataset.load_file(dataset_file)
         print(f"Dataset {dataset_identifier} loaded.   ")
     else:
-        dataset = model_name.dataset.load(**dataset_args)
+        dataset = model_name.dataset.load(**dataset_args, save_to=dataset_file)
 
-        print(f"Saving dataset {dataset_identifier}...", end="\r")
-        if not os.path.exists(dataset_dir):
-            os.makedirs(dataset_dir)
-        dataset.save(dataset_file)
-        print(f"Dataset {dataset_identifier} saved.   ")
+        # print(f"Saving dataset {dataset_identifier}...", end="\r")
+        # if not os.path.exists(dataset_dir):
+        #     os.makedirs(dataset_dir)
+        # dataset.save(dataset_file)
+        # print(f"Dataset {dataset_identifier} saved.   ")
 
     train_dataset_length = int(0.8 * len(dataset))
     test_dataset_length = len(dataset) - train_dataset_length
     train_dataset, test_dataset = random_split(
         dataset, (train_dataset_length, test_dataset_length)
     )
+
     train_loader = DataLoader(
         train_dataset, batch_size=args.batch_size, shuffle=True, num_workers=1
     )

### 2023-11-14 source/mlt/preexperiments/data_readers.py
cleanup

@@ -828,21 +828,15 @@ class CaptionGeneratorDataset(Dataset, Persistable):
     def __init__(
         self,
         captioner: Captioner,
-        samples: list[CaptionGeneratorSample],
         save_to,
     ) -> None:
         super().__init__()
 
         self.captioner = captioner
-        self.samples = samples
         self.file = save_to
 
         with h5py.File(save_to, "r") as f:
             self.num_samples = len(f["image"])
-
-        if len(samples) != 0:
-            self.save(save_to)
-            self.samples = []
 
     @classmethod
     def load(
@@ -918,7 +912,8 @@ class CaptionGeneratorDataset(Dataset, Persistable):
         print()
         print("loaded data.")
 
-        return cls(captioner, samples, save_to)
+        cls.save(save_to, samples, captioner)
+        return cls(captioner, save_to)
 
     def __getitem__(self, index):
         with h5py.File(self.file, "r") as f:
@@ -936,57 +931,31 @@ class CaptionGeneratorDataset(Dataset, Persistable):
     def __len__(self) -> int:
         return self.num_samples
 
-    def save(self, file_path):
+    @classmethod
+    def save(cls, file_path, samples, captioner):
         with h5py.File(file_path, "w") as f:
             f.create_dataset(
                 "image",
-                data=torch.stack([sample.image for sample in self.samples]),
+                data=torch.stack([sample.image for sample in samples]),
             )
             f.create_dataset(
-                "caption", data=torch.stack([sample.caption for sample in self.samples])
+                "caption", data=torch.stack([sample.caption for sample in samples])
             )
             f.create_dataset(
                 "non_target_captions",
-                data=torch.stack(
-                    [sample.non_target_captions for sample in self.samples]
-                ),
+                data=torch.stack([sample.non_target_captions for sample in samples]),
             )
             f.create_dataset(
                 "masked_image",
-                data=torch.stack([sample.masked_image for sample in self.samples]),
+                data=torch.stack([sample.masked_image for sample in samples]),
             )
-            f.create_dataset(
-                "image_id", data=[sample.image_id for sample in self.samples]
-            )
-            f.attrs["captioner"] = np.void(pickle.dumps(self.captioner))
+            f.create_dataset("image_id", data=[sample.image_id for sample in samples])
+            f.attrs["captioner"] = np.void(pickle.dumps(captioner))
 
     @classmethod
     def load_file(cls, file_path):
         with h5py.File(file_path, "r") as f:
[…]

### 2023-11-14 source/mlt/preexperiments/train.py

@@ -515,13 +515,10 @@ if __name__ == "__main__":
         dataset = model_name.dataset.load_file(dataset_file)
         print(f"Dataset {dataset_identifier} loaded.   ")
     else:
+        if not os.path.exists(dataset_dir):
+            os.makedirs(dataset_dir)
         dataset = model_name.dataset.load(**dataset_args, save_to=dataset_file)
-
-        # print(f"Saving dataset {dataset_identifier}...", end="\r")
-        # if not os.path.exists(dataset_dir):
-        #     os.makedirs(dataset_dir)
-        # dataset.save(dataset_file)
-        # print(f"Dataset {dataset_identifier} saved.   ")
+        print(f"Dataset {dataset_identifier} saved.   ")
 
     train_dataset_length = int(0.8 * len(dataset))
     test_dataset_length = len(dataset) - train_dataset_length

### 2023-11-14 source/mlt/util.py

@@ -6,7 +6,7 @@ import torch
 
 class Persistable(ABC):
     @abstractmethod
-    def save(self, file_path):
+    def save(self, file_path, *args):
         ...
 
     @classmethod

### 2023-11-14 source/mlt/language_games/callbacks.py
cleanup

@@ -78,3 +78,14 @@ class PrintMessages(Callback):
 
     def on_early_stopping(self, _train_loss, _train_logs, epoch, _test_loss, test_logs):
         self.print_events(test_logs)
+
+
+class ProgressLogger(Callback):
+    def __init__(self, batches_per_epoch) -> None:
+        super().__init__()
+        self.batches_per_epoch = batches_per_epoch
+
+    def on_batch_end(
+        self, logs: Interaction, loss: float, batch_id: int, is_training: bool = True
+    ):
+        print(f"Processing batch {batch_id} / {self.batches_per_epoch}...", end="\r")

### 2023-11-14 source/mlt/language_games/play.py

@@ -13,6 +13,7 @@ from mlt.language_games.callbacks import (
     ExcludingInteractionSaver,
     LogSaver,
     PrintMessages,
+    ProgressLogger,
 )
 from mlt.language_games.data_readers import (
     CaptionGeneratorGameBatchIterator,
@@ -428,12 +429,10 @@ def main(params):
         dataset = model.dataset.load_file(dataset_file)
         print(f"Dataset {dataset_identifier} loaded.   ")
     else:
-        dataset = model.dataset.load(**dataset_args)
-
-        print(f"Saving dataset {dataset_identifier}...", end="\r")
         if not os.path.exists(dataset_dir):
             os.makedirs(dataset_dir)
-        dataset.save(dataset_file)
+
+        dataset = model.dataset.load(**dataset_args)
         print(f"Dataset {dataset_identifier} saved.   ")
 
     if model.split_dataset:
@@ -526,10 +525,13 @@ def main(params):
                 )
             )
         else:
-            callbacks.append(
-                core.ConsoleLogger(
-                    print_train_loss=True,
-                ),
+            callbacks.extend(
+                [
+                    core.ConsoleLogger(
+                        print_train_loss=True,
+                    ),
+                    ProgressLogger(batches_per_epoch=opts.batches_per_epoch),
+                ]
             )
 
     if opts.save:

### 2023-11-15 thesis/sections/2-background.tex
memory efficiency
background

@@
 However, the goal is typically to select the best matching region of a given choice in a visual scene \citep{Qiao2020}.
-\cmtDK[inline]{describe methods in paper}
+They describe seven different approaches how the challenge is approached in the current research.
+In \emph{joint embedding} approaches, both visual input and the textual referring expressions are embedded into the same embedding space.
+By doing this, the model is able to find similarities and combine them.
+In some cases, an attention mechanism is applied on top.
+This allows the model to focus on important parts of each representation for the specific task.
+\emph{Modular models} are split up in multiple submodels that split up the complete task and solve them step by step.
+This helps to reduce the complexity of a task, as well as making more transparent, where models fail and succeed.
+\emph{Graph-based models} try to make use of the structure in a visual scene when multiple objects are present and the model is tasked to disambiguate them.
+In a first step, the visual features of the image are transferred into a graph that captures objects' properties and their relations.
+This graph is then combined with the referring expression to extract, which of the nodes is represented by it.
+To give a bigger focus on syntactic and semantic structure in a referring expression (which might get underrepresented by sentence-level representations of e.g.
+an LSTM), some models utilize \emph{external parsers}.
+These make use classical methods as for example dependency parsers, to parse the referring expressions.
+\emph{Weakly supervised} models are used when there is a lack of large datasets that annotate the relation between a referring expression and the corresponding regions in the image.
+Instead, they can operate on datasets that only include the visual scene with a matching referring expression.
+This is done by selecting a matching region in the visual scene and then reproducing the input referring expression from that region.
+During testing, the model only extracts the region from a given referring expression.
+While most of the described approaches grounding the referring expression in multiple steps, by first identifying possible regions around objects in the visual scene and then combining each proposal with the representation of the referring expression, \emph{one-stage approaches} combine them into one step.
+The image is processed as a whole and combined with the embedded referring expression.
+The model then predicts a bounding box around the referred entity.
+Finally, \emph{Visual-Language pre-training} tries to jointly learn from visual and textual representations, instead of separate visual and textual processors in the other approaches.
+Models like ViLBERT \citep{Lu2019} learn joint representations of language and visual input.
+These can then be fine-tuned on the specific referring expression comprehension task.
 
-Closely related to the REC task is visual question and answering (VQA).
+Closely related to the REC task is visual question and answering (VQA) \citep{Ahrens2022,Antol2015,Ilinykh2022,Xu2016}.
 In these tasks, a model is given a question about a visual scene.
@@
 In \citep{Johnson2017}, a model is for example shown an image similar to Figure \ref{fig:incremental-algorithm} and asked the question 'Is there a blue cube with the same size as the purple cylinder?'.
+To answer this, the model needs to understand two referring expressions and try to ground them in the image.
+In some cases the model can't ground a referring expression, when no object in the scene matches the referring expression.
+In the paper, it is solved by using a modular design of the model, where in a first step the question is translated into a sequence of instructions that defines, how the model is supposed to extract information from the image.
+In the first part, the referring expression is parsed, while in the second part the model grounds it in the visual input.
 
@@
 After this process, a new artificial language emerged, with which the agents can communicate with each other.
+
 % SD: Give an example of a language games.

### 2023-11-17 thesis/sections/2-background.tex
background

@@
 
+In this thesis, referring expression generation and referring expression comprehension are studied at the same time.
+This is done using language games between two artificial agents.
+The term of 'language games' was first introduced by \citet{Wittgenstein1953}.
+Following his argumentation, meaning of words and language is not absolute, but is situated in human dialogue.
+Hereby, two interlocutors converse in language games, that define how the utterances are meant and interpreted.
+The words' meaning is not just based on the context or surrounding words, but more on the situation, including pragmatic and social context.
+The rules in each of these situations differ and therefore the meanings and semantics of words and sentences may differ from situation to situation.
+This is summarized as a language game.
+These can be any small parts of conversations, for instance between a teacher and a student teaching a new concept or between two persons, discussing a specific topic.
+An interjection 'Water!' may be a warning, an answer to a question, a request or something else, depending on the context.
+% SD: Because referring expressions are ambiguous participants rely on language games to resolve the reference in context.
+% SD: Give an example of a game?
+% DK: TODO
+This applies as well to the ambiguous referring expressions, as in the example about the beverages described in section \ref{sec:background:re}.
+Which referring expression Person A uses to ask for the beverage, depends on the assumed shared knowledge.
+With other interlocutors or situations, other referring expressions are uttered.
+
+This reasoning was taken up, when trying to train artificial entities to produce a language.
+One of the original games is the signaling game, proposed by \citet{Lewis1969}.
+In this game, a sender needs to send a message to a receiver, based on information only available to the sender.
+The sender interprets the message and proposes an action to take.
+Both sender and receiver are rewarded in the same way if the proposed action was correct.
+Hence, both agents need to invent a language together, fit to the conditions of the game.
+
+The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
+Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
+Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+% SD: Examples of architectures from Baroni. The EGG framework.
+% DK: TODO
+
+In a language game, multiple agents need to solve a task by communicating with each other.
+The means of communication are very restricted.
+Typically, the agents are given a set of arbitrary symbols, a vocabulary, the agents can use to convey information.
+These symbols initially don't have any meaning, but by using and interpreting symbols repeatedly to solve the task the agents learn to use certain symbols, to encode specific information.
+By this, a new artificial language emerges between these agents.
+The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve.
+
 The center of this thesis evolves around language games between deep neural models.

### 2023-11-21 thesis/sections/2-background.tex
background

@@
 This applies as well to the ambiguous referring expressions, as in the example about the beverages described in section \ref{sec:background:re}.
-Which referring expression Person A uses to ask for the beverage, depends on the assumed shared knowledge.
+Which referring expression, Person A uses to ask for the beverage, depends on the assumed shared knowledge.
 With other interlocutors or situations, other referring expressions are uttered.
@@
 One of the original games is the signaling game, proposed by \citet{Lewis1969}.
-In this game, a sender needs to send a message to a receiver, based on information only available to the sender.
-The sender interprets the message and proposes an action to take.
+These games are made up of a sender and a receiver.
+The sender has access to the state of the world, while the receiver does not.
+The sender can choose from a fixed set of signals to communicate information to the receiver.
+The receiver, upon receiving the signal, must take an action based on that information.
+The goal of both the sender and the receiver is for the receiver to take the correct action in every state, as both players have a common interest in achieving the same outcomes.
 Both sender and receiver are rewarded in the same way if the proposed action was correct.
@@
 
-The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
+Central in language games is that the means of communication are very restricted.
+Typically, the agents are given a set of arbitrary symbols, a vocabulary, the agents can use to convey information.
+These symbols initially don't have any meaning, but by using and interpreting symbols repeatedly to solve the task the agents learn to use certain symbols, to encode specific information.
+By this, a new artificial language emerges between these agents.
+The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve.
+
+Language games and the emergence of language are studied in different environments.
+Agents can be robots that are situated in the real world.
+In \citep{Steels2009} the authors explore spatial language, in especially language that focuses on spatial positions movements.
+Two robots are moving freely in a space until they see each other as well as target object.
+The object is target is moved, which is registered by both robots.
+The robots now need to communicate the movement to each other using language games.
+Hereby, the experiments show that perspective and the capability of aligning the perspective with the other robot is central for the success of communication and an emergence of a language, when agents are embodied and have a specific view on the surrounding world.
+
+In more recent work, deep neural models are utilized for the agents.
+
+The agents can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
 Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
@@
 % DK: TODO
-
-In a language game, multiple agents need to solve a task by communicating with each other.
-The means of communication are very restricted.
-Typically, the agents are given a set of arbitrary symbols, a vocabulary, the agents can use to convey information.
-These symbols initially don't have any meaning, but by using and interpreting symbols repeatedly to solve the task the agents learn to use certain symbols, to encode specific information.
-By this, a new artificial language emerges between these agents.
-The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve.
 

### 2023-11-21 thesis/sections/2-background.tex
background

--- 2023-11-17 thesis/sections/2-background.tex
@@
 
-The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
-Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
-Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
-% SD: Examples of architectures from Baroni. The EGG framework.
-% DK: TODO
+The study of the emergence of an artificial language serves multiple goals.
+First, the study of the emergence of artificial language can help to shed light into why and how natural languages evolved \citep{Bartlett2005,Kirby2002,Kirby2008,Noukhovitch2021}.
+This approach allows a deep study of the constraints and restrictions that are necessary that symbols start to be connected with meaning and get grounded in the real world.
+Insights in this field might give indications, under which circumstances natural language evolves.
+On the other hand, research analyzes how the emerged language looks and how the agents represent information in messages and symbols \citep{Baroni2022,Lazaridou2017,Chaabouni2022,Kottur2017}.
+Some studies observe different properties of the emerged languages, such as compositionality, complexity or efficiency, and study how the agents can be biased towards producing a language that holds certain properties.
+Eventually, this can produce languages that agents can use in practical applications instead of relying on predefined unflexible protocols.
 
@@
 By this, a new artificial language emerges between these agents.
-The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve.
+The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve \citep{Baroni2022}.
+
+The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
+Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
+Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+% SD: Examples of architectures from Baroni. The EGG framework.
+% DK: TODO
 

### 2023-11-21 thesis/sections/2-background.tex

--- 2023-11-21 thesis/sections/2-background.tex
@@
 
+The study of the emergence of an artificial language serves multiple goals.
+First, the study of the emergence of artificial language can help to shed light into why and how natural languages evolved \citep{Bartlett2005,Kirby2002,Kirby2008,Noukhovitch2021}.
+This approach allows a deep study of the constraints and restrictions that are necessary that symbols start to be connected with meaning and get grounded in the real world.
+Insights in this field might give indications, under which circumstances natural language evolves.
+On the other hand, research analyzes how the emerged language looks and how the agents represent information in messages and symbols \citep{Baroni2022,Lazaridou2017,Chaabouni2022,Kottur2017}.
+Some studies observe different properties of the emerged languages, such as compositionality, complexity or efficiency, and study how the agents can be biased towards producing a language that holds certain properties.
+Eventually, this can produce languages that agents can use in practical applications instead of relying on predefined unflexible protocols.
+
 Central in language games is that the means of communication are very restricted.
@@
 By this, a new artificial language emerges between these agents.
-The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve.
+The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve \citep{Baroni2022}.
 
@@
 The agents can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
+Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
+Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+% SD: Examples of architectures from Baroni. The EGG framework.
+% DK: TODO
+
+The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
 Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.

### 2023-11-24 thesis/sections/2-background.tex
background

--- 2023-11-21 thesis/sections/2-background.tex
@@
 
-Current research focuses hereby on generating relational referring expressions, which add more complexity to the task \citep{Ghanimifard2017,Ghanimifard2019,Ramisa2015,Liu2022a}.
+Current research focuses hereby on generating relational referring expressions, which add more complexity to the task \citep{Ghanimifard2017,Ghanimifard2019,Ramisa2015,Liu2023}.
 Models need to first extract visual features of multiple objects from the scene.
@@
 \label{sec:background:language-games}
-\cmtDK[inline]{language games combine generation and understanding}
-\cmtDK[inline]{describe as 'dialogue turns'}
-\cmtDK[inline]{different approaches (robotics, neural models, ...)}
-\cmtDK[inline]{different study interests (linguistically, non-linguistically)}
-\cmtDK[inline]{continuous learning}
 % \begin{itemize}
@@
 In this thesis, referring expression generation and referring expression comprehension are studied at the same time.
-This is done using language games between two artificial agents.
-The term of 'language games' was first introduced by \citet{Wittgenstein1953}.
+This is done by using language games between two artificial agents.
+The term 'language games' was first introduced by \citet{Wittgenstein1953}.
 Following his argumentation, meaning of words and language is not absolute, but is situated in human dialogue.
@@
 This is summarized as a language game.
-These can be any small parts of conversations, for instance between a teacher and a student teaching a new concept or between two persons, discussing a specific topic.
+These can be any small parts of conversations, for instance between a teacher explaining a new concept to a student, or between two people, discussing a specific topic.
 An interjection 'Water!' may be a warning, an answer to a question, a request or something else, depending on the context.
@@
 Which referring expression Person A uses to ask for the beverage, depends on the assumed shared knowledge.
-With other interlocutors or situations, other referring expressions are uttered.
+With other interlocutors or situations, different referring expressions are uttered.
 
-This reasoning was taken up, when trying to train artificial entities to produce a language.
+This reasoning was taken up when trying to train artificial entities to produce a language.
 One of the original games is the signaling game, proposed by \citet{Lewis1969}.
-In this game, a sender needs to send a message to a receiver, based on information only available to the sender.
-The sender interprets the message and proposes an action to take.
+These games are composed of a sender and a receiver.
+The sender has access to the state of the world, while the receiver does not.
+The sender can choose from a fixed set of signals to communicate information to the receiver.
+The receiver, upon receiving the signal, must take an action based on that information.
+The goal of both sender and receiver is for the receiver to take the correct action in every state, as both players have a common interest in achieving the same outcomes.
 Both sender and receiver are rewarded in the same way if the proposed action was correct.
@@
 First, the study of the emergence of artificial language can help to shed light into why and how natural languages evolved \citep{Bartlett2005,Kirby2002,Kirby2008,Noukhovitch2021}.
-This approach allows a deep study of the constraints and restrictions that are necessary that symbols start to be connected with meaning and get grounded in the real world.
-Insights in this field might give indications, under which circumstances natural language evolves.
+This approach allows a deep study of the constraints and restrictions that are necessary so that symbols start to be connected with meaning and get grounded in the real world.
+Insights in this field might give indications for in which circumstances natural language evolves.
 On the other hand, research analyzes how the emerged language looks and how the agents represent information in messages and symbols \citep{Baroni2022,Lazaridou2017,Chaabouni2022,Kottur2017}.
 Some studies observe different properties of the emerged languages, such as compositionality, complexity or efficiency, and study how the agents can be biased towards producing a language that holds certain properties.
-Eventually, this can produce languages that agents can use in practical applications instead of relying on predefined unflexible protocols.
+Eventually, this can produce languages that agents can use in practical applications instead of relying on the predefined unflexible protocols that are currently used.
 
-In a language game, multiple agents need to solve a task by communicating with each other.
-The means of communication are very restricted.
-Typically, the agents are given a set of arbitrary symbols, a vocabulary, the agents can use to convey information.
-These symbols initially don't have any meaning, but by using and interpreting symbols repeatedly to solve the task the agents learn to use certain symbols, to encode specific information.
+In language games it is central that the means of communication are very restricted.
+Typically, the agents are given a set of arbitrary symbols, a vocabulary, that the agents can use to convey information.
+These symbols initially don't have any meaning, but by using and interpreting them repeatedly to solve the task the agents learn to use certain of those symbols, to encode specific information.
 By this, a new artificial language emerges between these agents.
-The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task, they need to solve \citep{Baroni2022}.
+The emergence of a language however is highly dependent on the constraints posed on (1) the agents' architectures, (2) the way of communication, (3) the perceptual input for the agents and (4) the task they need to solve \citep{Baroni2022}.
 
-The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
-Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
-Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+In \citep{Steels2009} two robots are trained by using language games.
+The authors explore how spatial language, more specifically language that focuses on spatial positions and movements, emerges when the agents are situated in the real world and can perceive each other.
+Two robots equipped with a camera are moving freely in a space until they see each other as well as the target object of the game.
+The target object is moved by the researchers, and that movement is registered by both robots.
+The robots now need to communicate the movement to each other using language games.
+Hereby, the experiments show that when agents are embodied and have a specific view on the surrounding world the capability of aligning their own perspective with the other's is central for the success of communication and the emergence of a language.
+
+In more recent work, deep neural models are utilized for the agents.
+Instead of being embodied, agents only receive visual input, without any perception that the other agent is present.
+In \citep{Lazaridou2017}, two agents play a discrimination game.
+The same set of images is shown to both agents.
+While they are unordered for the receiver, the sender always sees the target image as its first image.
+The task is now for the sender to communicate the target image in a way that the receiver can discriminate it from the distractors and then point towards it.
+Hereby, two of the constraints mentioned above are studied at once.
+First, the way of communication is constrained in the way that the sender can only communicate one symbol per message, which results in the agents clustering images into categories that are described with a symbol.
+Secondly, two different sender architectures are compared: the 'agnostic' sender that encodes each image separately and then produces the message, and the 'informed' sender that uses convolutions to combine both images for better discrimination.
+While both architectures are able to solve the task with an emerged language, the emerged language of the informed sender converges faster and makes use of more symbols than the agnostic sender.
+
+While the previously described examples always study prosocial agents (agents that have a common goal), \citet{Cao2018} examine self-interested agents.
+Hereby, multiple agents need to distribute a set of items between them.
+Each item has a utility score for each agent that is only visible to themselves.
+The goal for each agent is to gain the highest utility score by choosing the correct items.
+To do that, the agents first negotiate with arbitrary symbols in a 'linguistic' channel.
+At any time, any agent can choose to end the negotiation by proposing how to distribute the items, using predefined symbols in the 'proposal channel'.
+These can be accepted or rejected by the other agents.
+Their experiments show that in this self-interested setup, the agents utilize the proposal channel well to solve the task, but no language emerges in the linguistic channel.
+However, in a similar setup with prosocial agents, they start to utilize the linguistic channel and create an artificial language.
 % SD: Examples of architectures from Baroni. The EGG framework.
 % DK: TODO
-
-The center of this thesis evolves around language games between deep neural models.
-Hereby, multiple deep neural agents need to solve a task, by communicating with symbols, which initially are not associated with any meaning.
-By using and interpreting these symbols in there communication, the agents start to give these symbols and their combinations meaning.
-After this process, a new artificial language emerged, with which the agents can communicate with each other.
 
@@
 % DK: TODO
+\subsubsection{Setup of language games}
+\label{sec:background:lg:setup}
+In this thesis, the language games are studied with prosocial non-embodied agents that are deep neural networks.
+They can exchange multiple symbols to communicate after being presented with images.
+This is done using the \emph{EGG} framework that allows a high configuration of the agents' architectures, of the ways of communication, and of the technical constraints \citep{Kharitonov2019}.
 
-Section \ref{sec:aims_languages} explains reasons why research in this field is conducted and how it may help in further research.
-Section \ref{sec:setup-lg} shows in more detail how the language games are set up and how exactly agents can communicate and the language emerges.
-Section \ref{sec:properties-el} discusses, how the emerged languages can be built up.
-Finally, section \ref{sec:referring} discusses how the language games will be used in this thesis, to explore how agents can refer to objects in images.
+Using the framework, how the messages of the sender are produced can be controlled.
+For instance the sender can send single symbols or use different kinds of RNNs to produce a sequence as a message.
+The same applies to the receiver for parsing the message.
 
-\subsubsection{Aims of emerged languages}
-\label{sec:aims_languages}
-The setup of language games allows for very controlled rules of how agents can behave.
-% SD: Rephrase, language games provide interactive constraints within which language can emerge.
-But what are these?
-In this thesis we will study some, such as agent’s memory and layer representation and the feature representation, the structure of the world and its ambiguity.
-% DK: TODO
+Furthermore, a central problem in language games is how to train the sender.
+Usually every layer in the receiver model is differentiable and therefore standard backpropagation methods can be used to update the weights in the receiver.
+However, the sender's messages are discrete symbols and the loss coming from the receiver can therefore not be simply passed on to the sender.
+A classic solution to this problem is to use REINFORCE algorithms \citep{Williams1992} that can handle discrete symbols.
+More recently, Gumbel-Softmax relaxation \citep{Jang2017} is used to turn the discrete categorical distribution into a continuous probability distribution.
+Doing this, receiver and sender act as a whole deep neural model that can be completely backpropagated with standard methods.
+The methods are discussed deeper in section \ref{sec:methodology:optimization}.
+When setting up the language game with the EGG, both REINFORCE and Gumbel-Softmax relaxation can be used.
 
-\subsubsection{Setup of language games}
-\label{sec:setup-lg}
-The term of 'language games' was first introduced by \citet{Wittgenstein1953}.
-The author describes language games as uses of language between multiple interlocutors.
-These can be any small parts of conversations, for instance between a teacher and a student teaching a new concept or between two persons, discussing a topic.
-The rules in each of these situations differ and therefore the meanings and semantics of words and sentences may differ from situation to situation.
-An interjection 'Water!' may be a warning, an answer to a question, a request or something else, depending on the context.
-% SD: Because referring expressions are ambiguous participants rely on language games to resolve the reference in context.
-% SD: Give an example of a game?
-% DK: TODO
-meanings are bound to the rules of the language games.
+In classical machine learning problems, the models are trained using a train dataset for multiple epochs.
+To verify the performance, the model is verified on unseen test samples, which allows to test the ability to generalize its learned knowledge.
+Opposed to that, the main goal of language games is not to solve a specific task, but instead the success of communication with a useful emerged language.
+Consequently, the agents are trained continuously, and the testing doesn't validate the model on unseen samples, but instead is used to extract the exchanged messages, namely the emerged language.
+In other words, during training, the agents are shown randomly drawn samples from the dataset for each game.
+This is done repeatedly for $k$ 'games' so that the agents improve and ground their knowledge in the used symbols.
+During the testing, the agents are frozen and again randomly drawn samples are shown.
+The messages that the agents are exchanging now represent the language that emerged after $k$ language games which can be analyzed.
 
-This reasoning was taken up, when trying to train artificial entities to produce a language.
-One of the original games is the signaling game, proposed by \citet{Lewis1969}.
-In this game, a sender needs to send a message to a receiver, based on information only available to the sender.
-Based on the message, the receiver proposes an action.
-Both sender and receiver are rewarded in the same way if the proposed action was correct.
-Hence, both agents need to invent a language together, fit to the conditions of the game.
-
-The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
-Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
-Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
-% SD: Examples of architectures from Baroni. The EGG framework.
-% DK: TODO
-
-% This reasoning was taken up, when trying to train deep neural models to produce language.
-% These models need to be set into an interactive situation, a language game, that meaning and a language can emerge (\citet{Baroni2020}).
-% Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
-% Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.
+Section \ref{sec:methodology:egg} explains the framework in more detail.
 
 \subsubsection{Properties of the emerged language}
-\label{sec:properties-el}
+\label{sec:background:lg:properties}
+When agents develop an artificial language, they can learn to use it in different ways.
+Several studies research the properties of emerged languages, in particular if they share properties with human languages.
+One important property is the capability of forming new meanings by combining multiple symbols in a message, so-called compositionality \citep{Chaabouni2020, Gupta2020,Kharitonov2020, Lazaridou2018}.
+\citet{Lazaridou2018} construct a referential game where agents are presented with symbolic input instead of visual input in form of images.
+More specifically, each of the objects, target object and distractors, are represented as one-hot vectors where each dimension corresponds to an attribute of the object.
+This setup should bias the agents to rely directly on these given attributes in the messages to communicate the target object, instead of first having to extract every possible attribute, for instance  from visual features.
+Hence, this helps them to develop a compositional language.
+A compositional language should help the agents to communicate objects with unseen combination of attributes during training.
+Therefore, to analyze the degree of compositionality in the emerged language, the authors present the agents with new combinations of objects and measure the accuracy.
+Even if the accuracy drops lower, the fact that it is still above the random baseline shows that the agents are able to generalize to a certain degree.
+Furthermore, they utilize new combinations of symbols in their messages.
+However, \citet{Kharitonov2020} show that compositionality is not necessarily tied to the ability to generalize.
+In their experiments, non-compositional languages that emerge are able to generalize better than their compositional counterparts.
 
-\subsubsection{Referring to objects and grounding}
-\label{sec:referring}
+One essential property of human languages is \emph{Zip's Law of Abbreviation}, in other words that more frequent words tend to be shorter.
+In \citep{Chaabouni2019}, the researchers analyze if this law also applies to emerged languages.
+They find that without any pressure to produce shorter strings the agents produce a language that follows the inverse rule: the most frequent inputs are associated with longer messages.
+Only when long messages are penalized, a language which follows similar rules as human languages emerges.
+
+Finally, a strong focus in the study of emerged languages lies in interpretability \citep{Dessi2021,Lazaridou2017,Chaabouni2021}.
+While the agents ground the symbols in their input and certain symbols correspond to certain categories, they do not necessarily align with human categories and words.
+However, agents can be biased to use similar categories, as shown in \citep{Lazaridou2017}.
+This is done by infusing human labels during training so that the agents ground the visual input not only in their artificial language, but also in human language.
+More specifically, their experimental setup is a referential game, where the sender needs to describe the target image in a set of distractors to the receiver.
+Alternating with describing the target image to the receiver, the sender is trained to classify its input with human labels.
+The results show that this setup doesn't have any negative effect on the communicative success, but instead it only biases the agents to align their vocabulary with human descriptions.
+This makes the emerged language much easier to interpret for humans.
 
 \subsection{Artificial dataset}
-\cmtDK[inline]{'world shapes language'}
-\cmtDK[inline]{different types of biases -> reduce and control}
+Language is heavily influenced by the perception and the environment \citep{Hofstadter2013,Dobnik2017,Ji2022}.
+This is for instance very apparent in color naming systems across different languages \citep{Zaslavsky2018}.
+Which and how many words are used to refer to a specific color is dependent on the perceptual and social context around the speakers.
+Different studies show that while all languages tend to share similar basic color categories, some languages' systems are based on different underlying principles.
+Color names in the language of the Berinmo people from Papua New Guinea for example  reference natural objects and hence color categories are different from non-referential English color names \citep{Davidoff2001,Steels2005}.
+Further, this fact can be applied to more abstract constructs like metaphors, which often reference perceived events in the world \citep{Lakoff1980}.
+Perceiving that cats always land on the ground on their feet is for example the origin of the metaphor 'as nimble as a cat'.
+
+This influence on the language by the perception of the world is not only limited to human language learning and evolution, but it as well applies to machines that are trained to produce language.
+Many models in different disciplines of Natural Language Processing are trained on datasets such as the MS COCO dataset \citep{Lin2014} or the Flickr30k dataset \citep{Young2014}, which are sourced from uploaded photos in social networks.
+While these datasets often contain a large amount of samples, they can contain structural biases.
+The photos usually feature and center on specific objects, people or animals.
+This is not necessarily a faithful representation of human perception of the world, in which these entities are for instance partially hidden, shown from the back or appear at the edges of the field of view.
+The datasets might therefore give a euphemized view on the world.
+Furthermore, the datasets can contain underlying patterns in the semantic information of the photos \citep{Hendricks2018,Johnson2017a,Hudson2019}.
+For instance, when dogs are photographed, they usually appear running on a meadow.
+An image captioning model might learn to predict the caption 'a dog running on a meadow' for all images containing a dog, independently if the dog is actually running on a meadow or sitting in the living room, just because of the higher frequency in the dataset.
+In other words, the model might utilize these underlying structural biases rather than relying on the extracted visual information from the image.
+
+Different approaches are taken to mitigate this problem.
+While some research focuses on improving the models (e.g. \citet{Hendricks2018,Lu2017}), others create artificial datasets.
+Using artificial datasets, one can control exactly which structural patterns occur.
+Furthermore, all ground truths as for example spatial positions of objects are known and can be used for training models.
+Figure \ref{fig:artificial-datasets} shows several examples of different artificial datasets.
+Real world photos often contain a lot of distracting visual information that poses challenges for models to identify the relevant information.
+In \citep{Ghanimifard2017} and \citep{Ji2022}, the authors for example create abstract visual scenes that only contain specific concepts, like for instance spatial positions.
+With no distracting concepts present, researchers can focus only on the effects of the concept that is relevant for the respective study.
+Other datasets consist for example of comic-like scenes \citep{Zitnick2013,Zhang2016}.
+These scenes, while still being abstract also contain functional objects.
+This adds more complexity for predicting spatial relations.
+More recently, datasets are created using 3D environments \citep{Johnson2017a,Lee2022,Ahrens2022}.
+Scenes created in this way resemble photos taken in the real world and models can more easily transfer the learned knowledge to real world photos.
+Extending this reasoning, several datasets include embodied perspectives.
+Images are for example taken from the perspective of a person in the scene \citep{Dobnik2015} or from the perspective of an embodied robot navigating a 3D environment \citep{Hill2021}.
+
+\begin{figure}[ht]
+    \centering
+    \subfigure[Abstract spatial templates \citep{Ghanimifard2017}]{
+        \includegraphics[width=0.3\linewidth]{figures/dataset_Ghanimifard.png}
+    }
+    \subfigure[Abstract tangram shapes \citep{Ji2022}]{
+        \includegraphics[width=0.3\linewidth]{figures/dataset_Ji.png}
+    }
+    \subfigure[comic-like scenes \citep{Zhang2016}]{
+        \includegraphics[width=0.3\linewidth]{figures/dataset_Zhang.png}
+    }
+    \subfigure[3D-generated scenes \citep{Lee2022}]{
+        \includegraphics[width=0.3\linewidth]{figures/dataset_Lee.png}
+    }
+    \subfigure[3D-generated scenes including perspective \citep{Dobnik2015}]{
+        \includegraphics[width=0.3\linewidth]{figures/dataset_Dobnik.png}
+    }
+    \subfigure[embodied robot in a 3D environment \citep{Hill2021}]{
+        \includegraphics[width=0.3\linewidth]{figures/dataset_Hill.png}
+    }
+    \caption{Example images from different artificial datasets}
+    \label{fig:artificial-datasets}
+\end{figure}
+
+This thesis utilizes an artificial dataset for two reasons.
+First, all ground truth information about the scenes can be extracted during their generation.
+This includes spatial positions of entities placed in the scene, their properties and relations to other entities.
+Having all information present, several different tasks for the models can be created using the same generated images.
+More importantly, the usage of an artificial dataset allows for perfect control over the bias present in the images, in particular which entities are present, and which properties and attributes they have.
+As detailed before, perceptual input shapes the language that describes and interacts with it.
+When agents in a language game describe and refer to entities present in a scene, the emerged language is expected to utilize these biases by making use of the controlled attributes in the scenes.

### 2023-11-24 thesis/sections/3-methodology.tex
thesis
background

@@
 \subsection{Image processing}
+\label{sec:image-processing}
 \cmtDK[inline]{masking and how are they processed (convolutions to bring it to the same shape)}
@@
 \subsection{EGG framework}
+\label{sec:methodology:egg}
 The goal of this thesis is to run and compare different setups of language games systematically.
@@
 \subsection{Optimization in language games}
+\label{sec:methodology:optimization}
 The traditional approach to solve problems with a discrete channel relies on the REINFORCE algorithm \citep{Williams1992}.
@@
 Finally, the purpose of this thesis is to analyze the results and the emerged language and draw conclusions, how emerged languages can be grounded better in the environment.
-For that reason, the final models can't be uses in any real world applications and produce potential harm.
+For that reason, the final models can't be used in any real world applications and produce potential harm.
 On the other hand, this work can on the long run mitigate harm as it provides a study of models, how they would behave on real data.

### 2023-11-26 thesis/sections/1-introduction.tex
background
methodology, diagrams

@@
 Natural language is often used in an underspecified way.
-Interlocutors rely on pragmatic and semantic context to communicate.
-Given for example the following situation:
-Two people are in a room with two chairs.
-If one person asks for 'a chair', the second person would give him the chair that is closer to him.
-In other situations the second person would give him his preferred chair even if it's further away.
-The referring expression is the same in both contexts, but refers to two different entities.
-The difference lies in the pragmatic context that both interlocutors are aware of.
-In case this context is not shared, referring expressions might be misunderstood, because of the underspecification.
+Interlocutors rely on pragmatic and semantic context to communicate, particular interlocutors share a common ground which is used to disambiguate referring expressions.
+In case a common ground is not shared, referring expressions might be misunderstood, because of their underspecification.
 
@@
 Hereby, it considers the most salient or distinguishing properties first and then adds additional properties as needed.
-This helps to create concise, effective and disambiguous referring expressions.
-\cmtDK[inline]{co-reference}
+This helps to create concise, effective and unambiguous referring expressions.
+
+On the other side, humans as well as machines need to be able to comprehend referring expressions.
+Often this task is defined as identifying a unique target object from a set of distractors in a visual scene \citep{Mao2016}.
+Hereby, many different architectures for models are studied that try to solve this task \citep{Qiao2020}.
+In some approaches the models for example embed both the referring expression and bounding boxes around objects in the visual scene in the same embedding space to align them.
+In other approaches, models don't rely on bounding boxes, but extract features from the whole visual scene and combine them with the referring expression in one step.
+An extension of referring expression comprehension are visual question and answering tasks (VQA), in which a model is presented with a question and a visual scene that should be used to answer the question (e.g.
+\citep{Antol2015}).
+In these tasks, the model often needs to extract multiple referring expressions at the same time and ground them in the visual scene.
 
@@
 First, new artificial visual datasets are created, consisting of images, depicting objects and their attribute and spatial relations.
-Many existing datasets that are used to study referring expressions that use real images are based on photos taking by humans.
-This adds a lot of inherent bias to the dataset, since these photos often focus on similar objects and actions.
+Many existing datasets that are used to study referring expressions use real photos taking by humans.
+This adds a lot of inherent bias to the dataset, since these photos for instance often focus on similar objects and actions.
 Additionally, they require external knowledge about the world and the functions of objects which is not present in the image.
@@
 The datasets that are created in this thesis aim to reduce this bias and provide all information about the scene and objects in the image.
-In fact, the artificial creation of the dataset allows controlling precisely bias in each scene.
+In fact, the artificial creation of the dataset allows controlling precisely for the bias in each scene which can be utilized by agents in a language game.
 % SD: We also know the ground truth about the scenes as we know the function that generated them.

### 2023-11-26 thesis/sections/3-methodology.tex
methodology, diagrams

@@
 The basis for my datasets is the Visual Question Answering (VQA) framework CLEVR \citep{Johnson2017a}.
-This framework provides code to generate configurable VQA datasets that are split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
+This framework provides code to generate configurable VQA datasets that are split in two parts: a collection of images, and a set of questions and answers that refer to and describe each image.
 Many of the existing VQA datasets come with two problems.
@@
 By doing so, all information about for instance the location of objects or their relations to each other can be later used in training models or analyzing their results.
-Furthermore, it allows making predictions about different effects of varying contexts.
+Furthermore, it allows making predictions about the effects of varying contexts and information present in the scene.
 % SD: We know the ground truth function that generated the scenes and hence we can also make predictions about different effects of contexts.
@@
 Three to ten objects are placed in random locations into the scene and assigned with random attributes.
-To enhance realism and reduce ambiguity, objects are placed in a way that they do not intersect and have a certain distance from each other
+To enhance realism and reduce ambiguity, objects are placed in a way so that they do not intersect and have a certain distance from each other.
 Furthermore, it is made sure that every object is almost completely visible.
-The position of the light and the camera are slightly jittered for each image to add noise reduce recurring patterns.
-Since the objects are part of 3-d scenes, they may appear differently for each image, because of different lighting and shadows, distances to the camera and rotations.
-This noise approximates the real world and natural environment, and makes it harder to learn for models compared to relatively noise-free projections on a 2-d plane.
+The positions of the light and the camera are slightly jittered for each image to add noise and reduce recurring patterns.
+Since the objects are part of 3D scenes, they may look different in each image, because of different lighting and shadows, distances to the camera and rotations.
+This noise approximates the real world and natural environment, and makes it harder for models to learn compared to relatively noise-free projections on a 2D plane.
 Figure \ref{fig:clevr-example} shows an example of a generated image in the CLEVR dataset.
@@
 \label{sec:feature-extractors}
-\cmtDK[inline]{preprocessing}
 In computer vision tasks, machines need to analyze images and extract information from them.
@@
 These are also used in this thesis as a first layer to extract important information from the image. Hereby, two different architectures are tested.
-\cmtDK[inline]{explain advantage of CNN}
 
 First, we use the VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
-Using 16-19 convolutional layers with small convolution filters helps the model to solve localization and classification tasks on the training dataset, but also enables is to generalize onto other datasets.
-After the convolutional layers, the data is passed first through an average pooling layer which outputs 512x7x7 dimensions.
+Using 16 to 19 convolutional layers with small convolution filters helps the model to solve localization and classification tasks on the training dataset, but also enables it to generalize onto other datasets.
+After the convolutional layers, the data is passed first through an average pooling layer which outputs 512\times7\times7 dimensions.
 Next follow three linear layers with \emph{ReLU} non-linearities in between.
@@
 This approach enables the network to better preserve information from earlier layers and avoid the problem of information loss that can occur in very deep networks.
-There are four blocks that output 256x56x56, 512x28x28, 1024x14x14 and 2048x1x1 dimensions.
-A following average pooling layer outputs 2048x1x1 dimensions as well.
+There are four blocks that output 256\times56\times56, 512\times28\times28, 1024\times14\times14 and 2048\times1\times1 dimensions.
+A following average pooling layer outputs 2048\times1\times1 dimensions as well.
 The final linear layer reduces the flattened data to 1000 dimensions, corresponding to the ImageNet classes.
@@
 In this thesis, the implementations and weights available for PyTorch are used.\footnote{\href{https://pytorch.org/hub/pytorch\_vision\_resnet/}{https://pytorch.org/hub/pytorch\_vision\_resnet/}}\textsuperscript{,}\footnote{\href{https://pytorch.org/hub/pytorch\_vision\_vgg/}{https://pytorch.org/hub/pytorch\_vision\_vgg/}}
+In both cases, images that should be passed through the feature extractors need to have the same structure and shape as the images that were used in the pretraining.
+Hence, all images passed to them in the experiments in this thesis are preprocessed in the following way.
+As described in \citep{He2016,Simonyan2015}, the images are resized to 256 pixels for the shorter side, then cropped around the center to a square of 224\times224 pixels.
+Finally, the RGB channels are normalized, by subtracting the means (0,485; 0,456; 0,406) and dividing the result by the standard deviation (0,229; 0,224; 0,225) for each channel respectively.
+
 Since the task in this research is very different from a classification, it likely learned representations that are not directly transferrable to other tasks.
@@
 \label{sec:image-processing}
-\cmtDK[inline]{masking and how are they processed (convolutions to bring it to the same shape)}
+In the experiments different agent architectures are compared.
+To make these architectures comparable, the agents' models are composed of submodules that solve a specific subtask.
+Submodules can for example be image processing modules or text parsing modules.
+Hereby, different ways of solving the subtask are manifested in multiple exchangeable submodules.
+For instance, one image processing module uses CNNs while another image processing module uses a traditional approach.
+These can simply be exchanged in an agent's model without needing to adapt the whole agent.
+Consequently, the experiments stay comparable, since the differences are only added or removed submodules.
 
-The above described feature extractors are not enough to be solely used to encode the images used in this thesis.
+The feature extractors described in section \ref{sec:feature-extractors} are not enough to be solely used to encode the images used in this thesis.
 Even though final layers that contain very task and domain specific knowledge can be removed, the previous layers still don't contain any information about the new domain and task.
 For this reason they need to be extended with further layers that are not frozen and can learn and store information about the new domain by connecting the pretrained general knowledge with the specific task.
-In this thesis, we use one architecture to encode images for all the conducted experiment.
-The additional layers on top of the feature extractors however are always trained from scratch for each of the experiments and tasks.
-By doing this, the models are made up of multiple smaller modules, as for example the image processing module or other modules that encode different inputs.
-These modules can be added and combined step by step, whilst the architecture of each module always stays the same.
-Following, the experiments stay comparable, since differences are only added or removed modules.
+These additional layers on top of the feature extractors that make up the image processing submodule, however, are always trained from scratch for each of the experiments and tasks.
+In this thesis, only this submodule is used to encode images for all the conducted experiments.
 
 \citet{Johnson2017} describe an architecture that was used for training baseline models on the CLEVR dataset.
-This architecture will be used for all experiments.
+This architecture will be used as the submodule.
 Hereby, the image is first passed through a frozen feature extractor.
@@
 This vector represents the encoded image with its extracted features.
-Table \ref{tab:image_encoder} shows the layers and their output dimensions when using ResNet-3 as the feature extractor and an encoding size of 2048 dimensions.
+Table \ref{tab:image_encoder} shows the layers and their output dimensions when using ResNet-3 as the feature extractor and an image embedding size $e_i$.
 
@@
     \centering
-    \begin{tabular}{cc}
+    \begin{tabular}{ccc}
         \toprule
-        \textbf{layer}                      & \textbf{output dimensions} \\\midrule
-        Input image                         & 3 \times\ 224 \times\ 224  \\
-        ResNet-3                            & 1024 \times\ 14 \times\ 14 \\
-        Conv(3 \times\ 3, \rightarrow\ 128) & 128 \times\ 14 \times\ 14  \\
-        ReLU                                & 128 \times\ 14 \times\ 14  \\
-        Conv(3 \times\ 3, \rightarrow\ 128) & 128 \times\ 14 \times\ 14  \\
-        ReLU                                & 128 \times\ 14 \times\ 14  \\
-        MaxPool(2 \times\ 2, stride 2)      & 128 \times\ 7 \times\ 7    \\
-        Flatten                             & 6272                       \\
-        Linear Layer(\rightarrow\ 2048)     & 2048                       \\
+          & \textbf{layer}                      & \textbf{output dimensions} \\\midrule
+          & Input image                         & 3 \times\ 224 \times\ 224  \\
+        1 & ResNet-3                            & 1024 \times\ 14 \times\ 14 \\
+        2 & Conv(3 \times\ 3, \rightarrow\ 128) & 128 \times\ 14 \times\ 14  \\
+        3 & ReLU                                & 128 \times\ 14 \times\ 14  \\
+        4 & Conv(3 \times\ 3, \rightarrow\ 128) & 128 \times\ 14 \times\ 14  \\
+        5 & ReLU                                & 128 \times\ 14 \times\ 14  \\
+        6 & MaxPool(2 \times\ 2, stride 2)      & 128 \times\ 7 \times\ 7    \\
+        7 & Flatten                             & 6272                       \\
+        8 & Linear Layer(\rightarrow\ $e_i$)    & $e_i$                      \\
         \bottomrule
     \end{tabular}
-    \caption{Image encoder with ResNet-3 and an encoding size of 2048}
+    \caption{Image encoder with ResNet-3 and an image embedding size $e_i$}
     \label{tab:image_encoder}
 \end{table}
+
+In some experiments, the focus of the model should be pointed to certain regions in the image.
+This is done by providing the model additionally to the original image of the scene with a masked version of the scene.
+In particular, the masked version only contains black and white regions.
+All pixels in the relevant region are white, while all remaining pixels outside this region are black.
+This is expected to help the model to learn which are the important parts in the original image.
+However, a challenge lies in combining both original image and masked image.
+When extracting features from an image using the previously described submodule, the output of the final linear layer doesn't directly correspond to pixels and regions in the image, but only contains the important extracted features sequentially.
+In other words it looses the 2D geometrical structure of the image.
+The model will have problems to align these extracted features with the masked image, still having the 2D geometrical structure.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.48\linewidth]{figures/masked_image_combination.png}
+    \caption{Combination of the original image and its masked version}
+    \label{fig:masked-combination}
+\end{figure}
+
+For that reason, another submodule for masked images is used.
+In this submodule, the original image and the masked image are combined at an earlier stage, when both have the same structure (see Figure \ref{fig:masked-combination}).
+In particular, both images are processed in exactly the same way, by passing it through a feature extractor and then through the above described extending layers until layer 6, the max pool layer.
+Even though  the masked image doesn't contain similar features to the images the feature extractors are pretrained on, this method still provided better results in the experiments.
+After this, the images are concatenated along the channels, which results in a matrix with the dimensions 256\times7\times7 (still including the geometrical structure with 7\times7 grids).
+Finally, this matrix is flattened and reduced to the image embedding size $e_i$.
+Doing this, the model can directly align a region in the original image with a region in the masked image.
 
@@
 % DK: (rephrased; done)
-The configuration includes for example a choice between single symbol and sequence messages with varying RNNs, an easy switch between different loss functions or a choice between two optimization functions (Gumbel-Softmax relaxation and REINFORCE algorithms) to learn neural models containing discrete symbols.
+The configuration includes for example:
+\begin{itemize}[noitemsep,topsep=0pt]
+    \item a choice between single symbol and sequence messages with varying RNNs,
+    \item an easy switch between different loss functions,
+    \item or a choice between two optimization functions (Gumbel-Softmax relaxation and REINFORCE algorithms) to learn neural models containing discrete symbols.
+\end{itemize}
 % SD: More detailed description of all these. Readers will not be familiar with Gumbel-Softmax for example.
@@
 
-The EGG framework is set up in three levels.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.8\linewidth]{figures/egg_framework.png}
+    \caption{The EGG framework is made up of three levels: the game (gray), the wrappers (blue) and the agents (orange)}
+    \label{fig:egg-framework}
+\end{figure}
+
+The EGG framework is set up in three levels, shown in .
 Part of the lowest level are the \emph{agents} themselves.
-The agents are neural models that need to be implemented from scratch and define how the agents process their input and in case of the receiver combine it with the message as well as what is their output.
-The second level are \emph{wrappers} that take care of generating and parsing the message.
+The agents are neural models that need to be implemented from scratch and define how the agents process their input and in case of the receiver combine it with the message.
+The second level consists of \emph{wrappers} that take care of generating and parsing the message.
 The sender wrapper uses the output of the sender agent, to produce a message.
-The receiver on the other hand parses the message received by the sender and passes the result as an additional input to the receiver agent.
-The third level, the \emph{game} links all described parts together.
+The receiver wrapper on the other hand parses the message received by the sender and passes the result as an additional input to the receiver agent.
+The third level, the \emph{game}, links all described parts together.
 It provides the agents with the input and passes the message from the sender to the receiver.
@@
 % SD: Include a diagram?
-% DK: TODO
+% DK: done
 
@@
 \label{sec:methodology:optimization}
-The traditional approach to solve problems with a discrete channel relies on the REINFORCE algorithm \citep{Williams1992}.
-It uses the concept of Monte Carlo sampling to estimate the gradient of the expected cumulative reward with respect to the policy parameters.
-The basic idea is to collect trajectories by following the policy, compute the cumulative rewards for each trajectory, and then update the policy parameters to increase the probabilities of actions that led to higher rewards.
-% REINFORCE is easy to implement and understand, but it
-REINFORCE suffers from high variance in gradient estimation due to the inherent randomness in the environment and the policy.
-% SD: :-)
-% DK: (done)
+Training agents in a language game is not a straight forward task.
+In this thesis, the agents share a common goal, in particular optimizing the predictions of the receiver.
+The loss function of the whole game can therefore be defined as the loss of the receiver's predictions given its input and the sender's message.
+By this, the receiver can be trained relatively easily by backpropagating the game's loss through all layers in the model of the receiver.
+A problem however arises when passing on the loss to the sender.
+The generation of the message involves sampling from a categorical distribution and is therefore not differentiable.
+Consequently, the gradients of the loss function can't be estimated with standard backpropagation methods for the sender.
+To solve this problem two dominant methods are currently used: REINFORCE algorithms \citep{Williams1992} and Gumbel-Softmax relaxation \citep{Jang2017}.
 
-Gumbel-Softmax is a relaxation of the discrete categorical distribution to a continuous distribution that can be differentiated \citep{Jang2017}.
-It allows for the application of gradient-based optimization methods to discrete optimization problems.
-In the Gumbel-Softmax approach, randomness is introduced using the continuous Gumbel distribution.
-Following, the \emph{softmax} operation creates a distribution over discrete actions that can be differentiated.
-Gumbel-Softmax offers stable gradient estimates and can be more efficient than the traditional REINFORCE algorithm, especially when dealing with large action spaces.
+The traditional approach to solve problems with a discrete channel formulates the problem as a reinforcement learning task and relies on the REINFORCE algorithms.
+In reinforcement learning, the neural models are trained to optimize their policy given a reward.
+Given for example the referential game, described in \citep{Lazaridou2017}, where the sender is presented with two images and needs to communicate the target image to the receiver.
+The receiver is shown the same images and is tasked to point to the correct target image given the sender's message.
+The sender model and receiver model are following a policy with parameters $\theta$ respectively.
+The policy of the sender can be defined as $s(\theta_S(i_L,i_R,t)) \in V$ with $\theta_s$ being the parameters of the senders neural model, $i_L$ and $i_R$ the presented images, $t$ the target image and $V$ the vocabulary.
+The receiver's policy is dependent on the message by the sender and therefore directly dependent on the sender's policy and can be defined like this: $r(i_L,i_R,s(\theta_S(i_L,i_R,t))) \in \{L,R\}$.
+The reward is given if the receiver's policy successfully produced the target image:
+\begin{equation}
+    R =
+    \begin{cases}
+        1, & \text{iff } r(i_L,i_R,s(\theta_S(i_L,i_R,t))) = t \\
+        0, & \text{otherwise}
+    \end{cases}
+\end{equation}
+
+The task for the agents to learn is to maximize the reward function $R$ across all possible rewards (dependent on receiver's policy and target image) $\mathbb{E}_{\tilde{r}}[R(\tilde{r})]$.
+The problem hereby is that not all possible rewards are known.
+Without training, it can't be determined if a certain configuration of the agents, the policy, yields the correct target image.
+The rewards are in fact the outcome of the training.
+To solve this, Monte Carlo sampling is used to approximate the expected value $\mathbb{E}$.
+Monte Carlo sampling works on the fact that the mean of all drawn samples from an unknown probability distribution approximates its expected value.
+The more samples are drawn, the better the approximation.
+Using the approximation, the function becomes differentiable and can be used to calculate the gradients to update the policies.
+In this case, the approximation is based on the already determined rewards, in other words on the already played games.
+Early games are more or less trial and error, because the approximation is very bad.
+Only when many games are played, the approximation becomes better and the models can be updated properly.
+This however also means that the training can take very long until the models start to converge, and it suffers from high variance.
+
+For this reason, more recently Gumbel-Softmax relaxation is used to represent the discrete message.
+Basically, this method is a way to turn a discrete categorical distribution into a continuous distribution that can be differentiated.
+By doing this, sender and receiver are completely differentiable and act as whole deep neural model where the loss can be backpropagated with standard methods.
+In particular the problem lies in sampling from the categorical distribution $C$  over $|V|$ symbols, which can't be differentiated.
+Therefore, the method applies a reparameterization trick: random samples are instead drawn from the Gumbel distribution $G$ (Equation \ref{eq:gumbel-sample}) and summed with the logarithmic probabilities of each symbol (Equation \ref{eq:reparameterization}).
+\begin{align}
+    g   & = -log(-log(u))  & \text{ with }u\sim Uniform(0,1) \label{eq:gumbel-sample} \\
+    s_k & = log(c_k) + g_k & \text{ for }k = 1,...,|V| \label{eq:reparameterization}
+\end{align}
+
+To determine the symbol to use, usually the $argmax$ function would be applied on the resulting vector $s$.
+However, this function is again undifferentiable and is therefore approximated using the $softmax$ function:
+\begin{equation}
+    y_k = \frac{exp(s_k / \tau)}{\sum_{i=1}^{|V|}{exp(s_i / \tau)}}
+\end{equation}
+The temperature $\tau$ of the $softmax$ function determines how well it approximates the $argmax$ function.
+For $\tau \to 0$, the samples from the Gumbel-Softmax distribution are identical with the categorical distribution while for $\tau \to \infty$ the Gumbel-Softmax distribution becomes uniform.
+
+The Gumbel-Softmax method offers stable gradient estimates and can be more efficient than the traditional REINFORCE algorithm, especially when dealing with large action spaces.
 This also applies to language games, where \citet{Havrylov2017} demonstrated that Gumbel-Softmax relaxation is more effective.
+For that reason, the experiments in this thesis utilize Gumbel-Softmax relaxation.
 % SD: This is what a technical manual would say - but what does it do really?
@@
 % DK: I don't think, REINFORCE is superior also in our case, since GS offers more stability in learning (see citing) (TODO, explain GS)
-
-\subsection{Probing?}
-\cmtDK[inline]{reason for usage and techniques fits more into theoretical background}
 

### 2023-11-26 thesis/sections/5-grounding.tex
background
methodology, diagrams

@@
 Therefore, bounding boxes in this thesis rely on the former simpler approach.
-Since, each bounding box will be passed through on of the feature extractors, described in section \ref{sec:feature-extractors} they need to be normalized first.
-As described in \citep{He2016,Simonyan2015}, the bounding boxes are resized to 256 pixels for the shorter side, then cropped around the center to a square of 224 \times\ 224 pixels.
-Finally, the RGB channels are normalized, by subtracting the means (0,485; 0,456; 0,406) and dividing the result by the standard deviation (0,229; 0,224; 0,225) for each channel respectively.
+Since, each bounding box will be passed through on of the feature extractors, they are preprocessed as described in section \ref{sec:feature-extractors}.
 The array of bounding boxes is padded with matrix of zeros to the maximum possible number of objects present in a scene across the dataset.
@@
 Since for the second setup, the images will also be passed through the feature extractors, they are normalized as well in like manner.
-The referring expressions for the target object are generated using the incremental GRE-algorithm, described in section \ref{GRE}.
+The referring expressions for the target object are generated using the incremental GRE-algorithm, described in section \ref{sec:background:re}.
 By this, the model needs to describe the target object with respect to the distractor objects.
@@
 
-% TODO include figure
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/arch_bb_caption_generator.png}
+    \caption{Simplified architecture of the bounding box RE generator}
+    \label{fig:bb_re_generator_architecture}
+\end{figure}
 
-In the first setup, the \emph{bounding box RE generator}, the model receives the bounding boxes as input.
+In the first setup, the \emph{bounding box RE generator}, the model receives the bounding boxes as input (see Figure \ref{fig:bb_re_generator_architecture}).
 The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
@@
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
 In the second setup two different models are compared against each other.
@@
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
+    \caption{Simplified architecture of the masked RE generator}
+    \label{fig:masked_re_generator_architecture}
+\end{figure}
+
 Using this approach, the model doesn't have any information about the target object and is therefore expected to produce referring expressions for a randomly selected object on the scene.
-Therefore, it is extended in a second step with attention to the target object, the \emph{masked RE generator}.
+Therefore, it is extended in a second step with attention to the target object, the \emph{masked RE generator} (see Figure \ref{fig:masked_re_generator_architecture}).
 % SD: You add attention to the target object.
@@
 % SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
-% DK: (QUESTION)
+% DK: (done)
 The result is then again projected to the decoder output dimension $d$ by a linear layer and then used as the initial hidden state for the LSTM.

### 2023-11-27 thesis/sections/3-methodology.tex
setup grounding

@@
 To make these architectures comparable, the agents' models are composed of submodules that solve a specific subtask.
-Submodules can for example be image processing modules or text parsing modules.
+Submodules can for example be image encoder modules or text parsing modules.
 Hereby, different ways of solving the subtask are manifested in multiple exchangeable submodules.
-For instance, one image processing module uses CNNs while another image processing module uses a traditional approach.
+For instance, one image encoder module uses CNNs while another image encoder module uses a traditional approach.
 These can simply be exchanged in an agent's model without needing to adapt the whole agent.
@@
 For this reason they need to be extended with further layers that are not frozen and can learn and store information about the new domain by connecting the pretrained general knowledge with the specific task.
-These additional layers on top of the feature extractors that make up the image processing submodule, however, are always trained from scratch for each of the experiments and tasks.
+These additional layers on top of the feature extractors that make up the \emph{image encoder} submodule, however, are always trained from scratch for each of the experiments and tasks.
 In this thesis, only this submodule is used to encode images for all the conducted experiments.
@@
 
-For that reason, another submodule for masked images is used.
+For that reason, another submodule for masked images is used, the \emph{masked image encoder}.
 In this submodule, the original image and the masked image are combined at an earlier stage, when both have the same structure (see Figure \ref{fig:masked-combination}).

### 2023-11-27 thesis/sections/5-grounding.tex
setup grounding

--- 2023-11-08 thesis/sections/5-grounding.tex
@@
 % SD: We are validating the dataset on 3 tasks
-% DK: TODO
-
-\subsection{Object identification}
-\label{sec:object-identification}
-\subsubsection*{Setup}
-
-In a first task a neural model is trained to discriminate multiple objects from each other.
-Hereby, the model is shown bounding boxes of all objects in the scene as well as a description of the target object.
-The model needs to combine all information and then point to the correct bounding box.
-This task forces the model to extract visual features from the objects in the shown inputs and connect them to a representation of human given attributes.
-Since the model is shown only bounding boxes around each object rather than the whole scene, geometric information about the objects' locations doesn't play a role for the success in this task.
-
-In a first step, bounding boxes are extracted from each scene.
-Each bounding box is a square with a side length of 96 pixels around the center coordinates of each object.
-By doing this large objects in the front of the scene fill the bounding box completely, while smaller objects in the back of the scene contain some space around them.
-They also might contain noise in the form of parts of adjacent objects.
-Preliminary experiments indicated however that a more complex method for extracting the bounding boxes, depending on spatial position in the scene to reduce the noise didn't improve the results.
-Therefore, bounding boxes in this thesis rely on the former simpler approach.
-Since, each bounding box will be passed through on of the feature extractors, described in section \ref{sec:feature-extractors} they need to be normalized first.
-As described in \citep{He2016,Simonyan2015}, the bounding boxes are resized to 256 pixels for the shorter side, then cropped around the center to a square of 224 \times\ 224 pixels.
-Finally, the RGB channels are normalized, by subtracting the means (0,485; 0,456; 0,406) and dividing the result by the standard deviation (0,229; 0,224; 0,225) for each channel respectively.
-The array of bounding boxes is padded with matrix of zeros to the maximum possible number of objects present in a scene across the dataset.
-For the 'Dale-2' dataset, this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset.
-For each sample in the dataset the bounding boxes are shuffled.
-
-The attributes of the target object are encoded as one-hot encodings.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the \emph{color} and a two-dimensional vector encoding the two different \emph{sizes}.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated to a vector with 13 dimensions.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the object identification task}
-    \label{fig:bounding_box_classifier_architecture}
-    % SD: I haven't commented on this before, but let us be very careful in naming all meta-paramaters of the models in the text or in caption: how many layers, how many dimensions, etc.
-The reader should be able to replicate the code and the experiment from this.
-You can do this in the caption which is a longer explanation or in the text itself.
-    % DK: TODO
-\end{figure}
-
-The model, the \emph{object identifier} is split into two parts.
-First, each normalized bounding box of the sample is passed through \emph{ResNet-avg}.
-The preceding layers \emph{ResNet-1} to \emph{ResNet-4} couldn't be compared due to memory restrictions on the GPU.
-A deeper analysis of the effects of different feature extractors will be done in the following chapters.
-The resulting vector is flattened and projected to the embedding dimension $e$ via a linear layer.
-% SD: This is quite low? Why only 10?
-% DK: different embedding sizes are tested (done)
-Correspondingly, the one-hot vector of the attributes is projected to the same embedding dimension with another linear layer.
-In the second part, each bounding box is combined with the representation of the attributes.
-By bringing both vectors to the same embedding size, the dot product between each embedded bounding box and the embedded attributes can be calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model can so learn to connect the object representation with the attributes.
-The dot products are concatenated and form a vector with as many dimensions as objects present.
-The \emph{softmax} function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
-% The model then points to an object, using the \emph{softmax} function over this resulting vector.
-
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
-The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
-8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
-The loss is calculated using cross entropy.
-The following embedding dimension $e$ are compared: 2, 10, 50, 100, 500, 1000.
-% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
-% DK: the output is a probability distriution over all objects (inserted above; done)
-
-The object identification task is trained on all datasets that include more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
-Random baselines for each dataset would yield accuracies of 10\%, 50\% and 20\% respectively.
-% excluding the 'CLEVR single'.
-% SD: Why not?
-% DK: because there is only one object in the input, and therefore only one possible guess. The model can't point to a distractor (adressed; done)
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-Table \ref{tab:results:bounding_box_classifier} lists the accuracies of the models' predictions after 30 epochs for all three different datasets and different embedding sizes $e$.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{c|c|c|c}
-        \toprule
-               & \textbf{Dale-2}   & \textbf{Dale-5}   & \textbf{CLEVR Color} \\\cmidrule(lr){2-2}\cmidrule(lr){3-3}\cmidrule(lr){4-4}
-        $e$    & \textbf{Accuracy} & \textbf{Accuracy} & \textbf{Accuracy}    \\\midrule
-        {2}    & {92\%}            & {72\%}            & {40\%}               \\
-        {10}   & \textbf{100\%}    & {94\%}            & {92\%}               \\
-        {50}   & {99\%}            & \textbf{95\%}     & \textbf{94\%}        \\
-        {100}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
-        {500}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
-        {1000} & \textbf{100\%}    & {94\%}            & \textbf{94\%}        \\
-        \bottomrule
-    \end{tabular}
-    \caption{Accuracy scores of the object identifier after 30 epochs: $e$ are different embedding sizes}
-    \label{tab:results:bounding_box_classifier}
-\end{table}
-
-The first trend that is visible is that fewer distractor increase the accuracy of the model.
-The model achieves almost perfect accuracy in all configurations for the \emph{Dale 2} dataset with only two objects.
-With five objects in the \emph{Dale-5} dataset, the accuracy drops to 95\% and with a maximum possible number of 10 objects in the \emph{CLEVR color} dataset the model only achieves 94\% in the best configuration.
-This is not very surprising, since the model has a higher chance of predicting a wrong object, given the random baselines of 50\%, 20\% and 10\%.
-Furthermore, when more objects are shown to the model, more attributes are needed to discriminate the target object from the distractor, since it is more likely that the distractors share attributes with the target object.
-The task is getting therefore more challenging.
-
-A second conclusion is that the model needs a certain embedding space, to represent the features of each image.
-Across all dataset, an embedding space of only 2 dimensions result in a much lower performance compared to the best configuration.
-Even though the model is still able to achieve an accuracy of 30\% to 50\% points over the random baseline, this shows that a bigger embedding size is beneficial to extract and represent the features of the objects.
-Between 50 and 1000 dimensions, the model performs the best and varies only by 1\% point, which may be due to different random initialization of the weights in the model.
-
-In conclusion, the model is able to learn to discriminate the objects based on the visual appearance.
-The model can generate these high results, even with its relatively simple architecture.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-% SD: The model is able to leatn to discriminate objects based on visual appearance.
-% DK: (done)
-
-\subsection{Referring expression generation}
-\label{sec:referring_expression_generation}
-\subsubsection*{Setup}
-
-Opposed to the previous experiment, where the focus lied on extracting visual features, the model is now tasked to generate referring expressions.
-This is done in two setups.
-The first setup uses bounding boxes as in the previous section as input, where the model needs to describe the target object of one of the bounding boxes.
-In the second setup, the model is presented with the complete image and therefore also needs to accommodate to geometric information of the scene.
-
-The bounding boxes are extracted and normalized in the same way as for the object identification task.
-Since for the second setup, the images will also be passed through the feature extractors, they are normalized as well in like manner.
-The referring expressions for the target object are generated using the incremental GRE-algorithm, described in section \ref{GRE}.
-By this, the model needs to describe the target object with respect to the distractor objects.
-There are some minor additions concerning the padding of the referring expression.
-As before, the referring expression is padded to a number of three tokens, corresponding to the maximum of three attributes.
-However, there are three different ways how the padding is applied.
-First, the referring expression are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the referring expression is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the referring expression is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots disappear.
-A referring expression that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since this task is not focussing on producing natural language with a correct grammar, but focuses instead on extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the referring expression is prepending the referring expression with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-Each slot has always the same semantic value, e.g. the last slot always contains the shape of an object.
-This can be done since the referring expressions are not free text, but instead the structure and the possible content is given by the dataset.
-This method might help the model to learn the correct referring expressions.
-% SD: Slots have the same semantic value, i.e.
-the last slot is the object type.
-We can do this because we know the structure of the descriptions in this dataset and because we hope this will help the learning.
-Also, we are thinking of generation based on slots.
-% DK: (addressed; done)
-The last variation concerns the order of producing each token.
-When the referring expressions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the referring expression.
-% SD: We... I sounds more like you are writing a lab report with casual language but here you want to be more formal.
-% DK: (done)
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the referring expression are still preserved.
-
-This task inherently involves learning human knowledge and natural language structure.
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Can the model solve the task, or do specific attributes used by humans pose challenges to the model?
-% Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
-% SD: In this case it should be using the same attributes as humans as we are training it on human labels.
-% DK: (rephrased; done)
-
-% TODO include figure
-
-In the first setup, the \emph{bounding box RE generator}, the model receives the bounding boxes as input.
-The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
-As in the previous experiment, each bounding box is passed through \emph{ResNet-avg} and projected afterwards to an image embedding dimension $e$ with a linear layer.
-All thus encoded bounding boxes are concatenated and again compressed to the decoder output dimension $LSTM_o$ using another linear layer.
-This representation of all objects serves as the initial hidden state of an LSTM, which generates the referring expression.
-Tokens used in the LSTM are embedded with embedding dimension $LSTM_e$.
-During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-In the second setup two different models are compared against each other.
-The first model, the \emph{basic RE generator} acts as the baseline and only receives the complete image as the input.
-The image is passed through \emph{ResNet-3} and subsequently processed by additional convolution layers, described in section \ref{sec:image-processing}.
-The resulting vector is then reduced to decoder output dimension $LSTM_o$ by a linear layer and as for the bounding boxes acts as the initial hidden state for the LSTM.
-The LSTM is trained in the same way as above.
-
-Using this approach, the model doesn't have any information about the target object and is therefore expected to produce referring expressions for a randomly selected object on the scene.
-Therefore, it is extended in a second step with attention to the target object, the \emph{masked RE generator}.
-% SD: You add attention to the target object.
-% DK: (added; done)
-A masked version of the image is created and passed to the model.
-For this, a fixed size squared area of 96 \times\ 96 pixels around the center of the target object is filled in white, while the rest of the image is filled in black.
-Both original and masked image are processed together with convolutional layers as described in section \ref{sec:image-processing}.
-% SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
-% DK: (QUESTION)
-The result is then again projected to the decoder output dimension $d$ by a linear layer and then used as the initial hidden state for the LSTM.
-This should point the model towards which object to describe and discriminate from the distractors.
-
-For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
-8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
-The loss is calculated using cross entropy.
-Table \ref{tab:variables-reference-expression-generation} shows which variables are compared for each model:
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{lccc}
-        \toprule
-                                  & $e$                & $LSTM_o$           & $LSTM_e$       \\
-                                  & $[100, 500, 1000]$ & $[100, 500, 1000]$ & $[10, 15, 30]$ \\\midrule
-        bounding box RE generator & \times             & \times             & \times         \\
-        basic RE generator        & -                  & \times             & \times         \\
-        masked RE generator       & -                  & \times             & \times         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Variables for each model where $e$ is the embedding dimension, $LSTM_e$ the embedding dimensions for tokens in the LSTM and $LSTM_o$ the output dimension of the LSTM}
-    \label{tab:variables-reference-expression-generation}
-\end{table}
-
-As already discussed before, this task can be interpreted as a classification task rather than a natural language generation task.
-The main reason for this is that the model is tasked to assign specific attributes to the target object instead of producing free text with a large vocabulary.
-Following, we are interested in the classification mistakes the model makes.
-For this, the model's success is validated on accuracy, recall and precision scores.
-These are calculated in the following ways.
-The first measure is the \textbf{overall accuracy} if the model predicted every word in the caption correctly.
-% SD: This is too strict.
-if this is a generation task then we can use generation metrics.
-Those compare if the generated string as a whole is kind of similar to the target string, see ROUGE, METEOR, BERTscore.
-% If this is a classification task, and it is, given that we know what type of word we expect in each slot, we could calculate accuracy per class, precision, recall, F-score.
-We could also calculate mean ranks for predicting words per classes, or acucracy at k- if the target word is in the k-top predictions.
-% We could also take it a multi-class classification problem and how to calculate accuracy in this case, would need to check.
-% DK: I agree that classification metrics refelct success better here (therefore I also used accuracy).
-These metrics would give a better overview, but I don't think there is time to implement it (QUESTION)
-This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
-% DK: TODO
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
-To mitigate this, the accuracies for each class are included, as well as the macro average.
-% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
-% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
-This can give a better understanding of the errors the model makes.
-The same is done for \textbf{precision} and \textbf{recall}.
-
-% SD: Very good If I understand this right this would be a case where the generated description describes another object better than the target and a human would identify that.
-So it is like false negatives in describing.
-Perhaps use another term for this: referring to a distractor?
-(since you are not measuring hwo individually the description fitts a distractor but only if the entire description could perfectly fit a distractor)
-% DK: exactly. I added false negatives. But I would like to keep accuracy in the name, since that is what it is. maybe better distractor accuracy? (QUESTION)
-With the \textbf{non-target accuracy}, we identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-By this it measures the false negative generated captions.
-For this, referring generations are generated for all the non-target objects and distractors in the images using the GRE-algorithm.
-Importantly to notice is that the referring expressions may not uniquely identify a distractor, since multiple distractors with the same shape, color and size are allowed.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, we can get an overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\subsubsection*{Results}
-Table \ref{tab:results:bb-re-generator} shows the \emph{overall accuracy}, \emph{F1 scores} for each word and the \emph{non-target accuracies} of the \textbf{bounding box RE generator} when trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
-As can be seen, the overall accuracies, in other words perfect matches of the generated referring expression depend very much on the dataset.
-With the \emph{Dale-2} dataset, the model can achieve perfect matches in 99\% of the cases in its best configuration.
-Also the \emph{CLEVR color} dataset allows the model to predict the correct referring expression in 93\% of the samples.
-Opposed to that the model can only generate perfect referring expressions in 69\% of the samples of the \emph{Dale-5} dataset.
-Looking at the \emph{non-target accuracy}, one of the problems can be identified for the \emph{CLEVR color} dataset.
-Summing the \emph{non-target accuracy} and the \emph{overall accuracy}, one gets the accuracy that any of the shown object was described independently if it was the target object.
-For the \emph{CLEVR color} dataset, this score lies at 96\% to 97\% in the best configurations, which means that the model described a shown object for almost all the samples.
-In 3\% to 4\% of the cases, the model only picked the wrong object to describe.
-This looks different for the \emph{Dale-5} dataset.
-Here, the model describes only in 71\% of the cases one of the shown objects, 69\% the target object and 2\% a distractor.
-The model therefore struggles more with the correct generation of words than with choosing, which object to describe.
-
-When looking at the different configurations, in especially different values for $e$, $LSTM_o$ and $LSTM_e$, it can be seen that the results are not very different for the \emph{Dale-2} and \emph{CLEVR color} datasets.
-However, bigger effects can be identified for the \emph{Dale-5} dataset.
-Here, a low output dimension of the LSTM $LSTM_o$ tends to give lower scores.
-This especially enhanced, when also the embedding dimensions of the tokens $LSTM_e$ is low.
-Again, the image embedding size $e$ doesn't seem to have a big effect, when over 50 dimensions.
-Indeed, the model achieves an accuracy of 69\% with all three tested embedding sizes.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-               &          &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                               \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $e$    & $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} \\\midrule
-        {100}  & {100}    & {10}     & {97}                                & {97,67}                             & {1}                                      & {65}          & {88,17}        & {2}         & {92}          & {94,95}        & {3}         \\
-        {100}  & {100}    & {15}     & {97}                                & {97,77}                             & {0}                                      & {62}          & {86,51}        & {2}         & {92}          & {94,45}        & {4}         \\
-        {100}  & {100}    & {30}     & {98}                                & {97,81}                             & {1}                                      & {65}          & {87,97}        & {2}         & {92}          & {94,47}        & {4}         \\
-        {100}  & {500}    & {10}     & {98}                                & {98,16}                             & {0}                                      & {67}          & {88,13}        & {2}         & {93}          & {95,18}        & {4}         \\
-        {100}  & {500}    & {30}     & \textbf{99}                         & \textbf{98,57}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,53} & \textbf{2}  & \textbf{93}   & \textbf{95,17} & \textbf{3}  \\
-        {100}  & {1000}   & {10}     & {98}                                & {98,28}                             & {0}                                      & {68}          & {88,23}        & {2}         & \textbf{93}   & \textbf{95,41} & \textbf{3}  \\
-        {500}  & {1000}   & {30}     & \textbf{99}                         & \textbf{98,85}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,15} & \textbf{2}  & {93}          & {95,05}        & {4}         \\
-        {500}  & {500}    & {10}     & \textbf{99}                         & \textbf{98,71}                      & \textbf{0}                               & {67}          & {88,99}        & {2}         & \textbf{93}   & \textbf{95,4}  & \textbf{3}  \\
-        {1000} & {100}    & {10}     & {97}                                & {97,72}                             & {0}                                      & {59}          & {85,43}        & {2}         & {93}          & {95,04}        & {4}         \\
-        {1000} & {100}    & {30}     & {98}                                & {98,16}                             & {0}                                      & {61}          & {86,77}        & {2}         & {92}          & {94,68}        & {4}         \\
-        {1000} & {1000}   & {10}     & \textbf{99}                         & \textbf{98,95}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,21} & \textbf{2}  & {93}          & {95,15}        & {4}         \\
-        {1000} & {1000}   & {15}     & {99}                                & {98,67}                             & {0}                                      & {68}          & {88,98}        & {2}         & {93}          & {95,01}        & {4}         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the bounding box caption generator after 30 epochs: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
-    \label{tab:results:bb-re-generator}
-\end{table}
-
-Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results and especially what mistakes the model is making for both the \emph{Dale-5} and \emph{CLEVR color} datasets.
-They list \emph{precision} and \emph{recall} metrics for each token for the best configuration of the model with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
-The tokens are grouped by attribute and also show the metrics averaged over each of the attributes.
-Since the overall results for the \emph{Dale-2} dataset are already close to perfect, the focus for this analysis lies on the remaining two datasets.
-The metrics of the \emph{<pad>} token indicate if the model produced the correct length of the referring expression, in other words if it was able to determine which attributes are necessary to discriminate the target object from the distractors.
-For the \emph{CLEVR color} dataset, the scores are perfect.
-This is not surprising, because all referring expressions for the \emph{CLEVR color} dataset consist of exactly two attributes, shape and color, and the first generated token will always be the only \emph{<pad>} token in the referring expression (corresponding to the unspecified size).
-The \emph{<pad>} token is therefore easy to learn.
-For the \emph{Dale-5} dataset, the model struggles more to predict the correct length of the referring expression.
-\cmtDK[inline]{why?}
-
-When looking at the tokens for the shape, it can be seen that the model is able to identify it very well across all datasets.
-The model predicts the correct shape for all samples using the \emph{CLEVR color} dataset, while both \emph{precision} and \emph{recall} lie around 98,3\% when using the \emph{Dale-5} dataset.
-Even though the score is almost perfect, the slight difference might stem from the fact that all distractors have the same shape in the first case, while distractors can be different in the second case.
-Consequently, the model is only exposed to one shape at a time for each sample, which might simplify its identification.
-
-For the color attribute, the metrics drop significantly for both \emph{Dale-5} and \emph{CLEVR color} to an average of around 93\%.
-Hereby, no meaningful difference can be seen across the datasets, but there are differences between the colors.
-Some colors are predicted with \emph{precision} and \emph{recall} around 95\% to 96\%, while others are only around 90\%.
-However, these differences are not reproducible across multiple runs and configurations.
-The best and worst predicted colors vary and no conclusions can be drawn which colors are easier to predict for the model.
-
-Finally, the size is the most difficult attribute to predict for the model.
-Apart from the \emph{CLEVR color} dataset, where a size never needs to be predicted and also is never predicted, the metrics for the prediction of size tokens are the lowest across all tokens.
-They are the only mistakes, the model makes, when exposed to the \emph{Dale-2} dataset and the average \emph{precision} lies around 23\% below the average of predictions of the color for the \emph{Dale-5} dataset, while the average \emph{recall} lies around 28,82\% below.
-The reason why the \emph{precision} is higher than the \emph{recall} is the \emph{<pad>} token, which is predicted very often instead of a token specifying the size.
-In fact, the opposite relationship is visible for the \emph{precision} and \emph{recall} for said token.
-The much higher absolute number of \emph{<pad>} tokens leads to a smaller relative difference of \%-points shown in the table.
-Again, no conclusion can be drawn if larger or smaller objects are easier to predict, since the results vary across runs and configurations.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rr|cc|c|ccc|c|c}
-        \toprule
-                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {99,17} & {98,29} & \textbf{98,73} & {99,86} & {99,71}    & {99,67}  & \textbf{99,75} & {99,64} \\
-                                         & {Recall}    & {97,54} & {94,26} & \textbf{95,9}  & {100}   & {99,56}    & {99,67}  & \textbf{99,74} & {99,77} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {69,65} & {69,21} & \textbf{69,43} & {98,19} & {98,32}    & {98,39}  & \textbf{98,3}  & {82,22} \\
-                                         & {Recall}    & {62,11} & {66,15} & \textbf{64,13} & {98,79} & {97,87}    & {98,25}  & \textbf{98,3}  & {84,59} \\\midrule
-        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
-                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
-        \bottomrule
-    \end{tabular}
-    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
-The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
-    \label{tab:results:bb-re-generator_size-shape}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rr|cccccccc|c}
-        \toprule
-                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {94,51} & {98,77} & {97,59} & {98,68} & {98,89} & {98,8}   & {97,47} & {100}    & \textbf{98,09} \\
-                                         & {Recall}    & {97,73} & {100}   & {98,78} & {97,4}  & {96,74} & {98,8}   & {100}   & {98,8}   & \textbf{98,53} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {92,12} & {93,82} & {89,13} & {89,12} & {92,63} & {91,12}  & {97,24} & {94,36}  & \textbf{92,44} \\
-                                         & {Recall}    & {92,12} & {89,78} & {94,91} & {94,51} & {95,71} & {92,42}  & {89,34} & {94,85}  & \textbf{92,95} \\\midrule
-        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                           \\\textbf{color}} & {Precision}           & {93,46} & {92,37} & {94,47} & {93,86} & {92,04} & {91,13}  & {90,07} & {94,7}   & \textbf{92,76} \\
-                                         & {Recall}    & {92,75} & {92}    & {95,98} & {89,92} & {94,12} & {91,13}  & {94,23} & {91,91}  & \textbf{92,76} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
-    \label{tab:results:bb-re-generator_color}
-\end{table}
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
-The final peak stayed exactly the same and the effects were therefore not studied deeper.
-% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
-% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)
-
-In conclusion, it can be said that the model is able to extract discriminative features from the shown bounding boxes and produce referring expressions.
-However, the result highly depends on the amount of distractors and the resulting need of more discriminative features to describe the target object.
-While the shape is easily identified, the model has bigger problems to identify the color and especially the size of the target object.
-
-In the following paragraphs, the results of the \textbf{basic RE generator} as well as the \textbf{masked RE generator} are evaluated.
-Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 30 epochs.
-Compared to the \emph{bounding box RE generator}, the task differs substantially.
-Instead of bounding boxes, the model is shown the image of the whole scene.
-Additionally to extracting features of the target object and distractors separately, the model is now tasked to extract these at once for all objects.
-Furthermore, it needs to learn which of the objects is the target object by combining the image with the masked image.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|ccc|ccc|ccc}
-        \toprule
-                 &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                               \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
-        $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} \\\midrule
-        {100}    & {10}     & {81}                                & {76,16}                             & {9}                                      & {21}          & {57,07}        & {23}        & {16}          & {43,69}        & {31}        \\
-        {100}    & {15}     & {78}                                & {72,06}                             & {11}                                     & {17}          & {53,29}        & {25}        & {26}          & {50,87}        & {30}        \\
-        {100}    & {30}     & {83}                                & {78,76}                             & {7}                                      & {19}          & {54,91}        & {24}        & \textbf{29}   & \textbf{53,37} & \textbf{26} \\
-        {500}    & {10}     & {83}                                & {79,78}                             & {6}                                      & {27}          & {62,05}        & {16}        & {17}          & {44,93}        & {32}        \\
-        {500}    & {15}     & {83}                                & {79,75}                             & {7}                                      & {28}          & {63,78}        & {16}        & {21}          & {47,08}        & {31}        \\
-        {500}    & {30}     & {84}                                & {83,03}                             & {5}                                      & {30}          & {64,36}        & {17}        & {21}          & {40}           & {31}        \\
-        {1000}   & {10}     & {84}                                & {82,2}                              & {5}                                      & {31}          & {65,58}        & {13}        & {12}          & {37,67}        & {34}        \\
-        {1000}   & {15}     & {85}                                & {82,62}                             & {5}                                      & {31}          & {64,28}        & {16}        & {16}          & {43,27}        & {32}        \\
-        {1000}   & {30}     & {85}                                & {83,13}                             & {4}                                      & {31}          & {64,06}        & {17}        & {13}          & {41,79}        & {34}        \\
-        {1500}   & {10}     & {84}                                & {81,27}                             & {4}                                      & {33}          & {67,5}         & {14}        & {12}          & {39,04}        & {34}        \\
-        {1500}   & {15}     & \textbf{85}                         & \textbf{84,11}                      & \textbf{4}                               & {35}          & {69,05}        & {12}        & {17}          & {44,98}        & {31}        \\
-        {1500}   & {30}     & {84}                                & {80,93}                             & {4}                                      & \textbf{36}   & \textbf{68,97} & \textbf{15} & {14}          & {42,99}        & {33}        \\
-        {2000}   & {10}     & {83}                                & {82,69}                             & {4}                                      & {36}          & {69,11}        & {13}        & {13}          & {41,85}        & {34}        \\
-        {2000}   & {15}     & {84}                                & {81,95}                             & {4}                                      & {34}          & {66,94}        & {14}        & {14}          & {38,79}        & {33}        \\
-        {2000}   & {30}     & {85}                                & {82,8}                              & {4}                                      & {32}          & {65,52}        & {16}        & {18}          & {45,41}        & {32}        \\
-        {3000}   & {10}     & {82}                                & {79,83}                             & {4}                                      & {34}          & {67,61}        & {12}        & {12}          & {38}           & {34}        \\
-        {3000}   & {15}     & {85}                                & {81,8}                              & {4}                                      & {32}          & {65,64}        & {15}        & {13}          & {38,52}        & {35}        \\
-        {3000}   & {30}     & {83}                                & {80,6}                              & {3}                                      & {30}          & {63,5}         & {16}        & {12}          & {38,45}        & {34}        \\
-        \bottomrule
-    \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked caption generator: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
-    \label{tab:results:masked-re-generator}
-    % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1.
-It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
-    % DK: might be the case.
-in the last language game i use only a linear layer instead of convolutions, but it doesn't help.
-This however might also be due to other problems.
-Time?
-(QUESTION, future work)
-    % SD: I would present these for individual classes.
-    % DK: I only have accuracy scores at. Time? (QUESTION)
-\end{table}
-
-Consequently, the results are much less accurate.
-This applies to all datasets, even though it is most apparent for the \emph{CLEVR color} dataset.
-While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the \emph{Dale-2} dataset, the model only reaches 36\% for the \emph{Dale-5} dataset and 29\% for the \emph{CLEVR color} dataset.
-However, these scores are all well above the random baselines of 50\%, 20\% and 10\% respectively.
-Moreover, the referring expression generated for the \emph{Dale-2} dataset are also efficient in the sense that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both target object and distractor are therefore extracted and associated with the vocabulary.
-Striking in these results is also the \emph{non-target accuracy}, which is much higher than for the \emph{bounding box RE generator}.
-For the \emph{Dale-2} dataset, it arrives at 11\% in one configuration, for the \emph{Dale-5} dataset, the model describes for at least 12\% of the samples a distractor, for one configuration even in over 25\% of the cases.
-The model is most uncertain, which object to describe when exposed to the \emph{CLEVR color} dataset, where in almost all configurations, it describes a distractor every third sample.
-This shows that the model struggles to learn, which of the objects in the scene is the target object, more specifically, to combine the whole image with the masked image.
-\cmtDK[inline]{compare to basic RE generator}
-Yet, the model is able to extract the attributes from the shown objects.
-This can be seen, when summing up the \emph{overall accuracy} and the \emph{non-target accuracy}.
-This number represents in how many cases any of the shown objects were described.
-For the \emph{Dale-2} dataset, this number lies at 90\% for the best configuration, at 51\% for the \emph{Dale-5} dataset and at 56\% for the \emph{CLEVR color} dataset.
-While being still lower than the same metric for the \emph{bounding box RE generator}, the model fares well in comparison for the \emph{Dale} datasets given the highly increased complexity of the task.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rr|cc|c|ccc|c|c}
-        \toprule
-                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {68,63} & {76}    & \textbf{72,31} & {98,41} & {98,18}    & {98,92}  & \textbf{98,5}  & {94,77} \\
-                                         & {Recall}    & {30,17} & {34,55} & \textbf{32,36} & {98,7}  & {98,33}    & {98,47}  & \textbf{98,5}  & {98,64} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {49,47} & {57,3}  & \textbf{53,38} & {87,09} & {83,26}    & {84,78}  & \textbf{85,04} & {64,99} \\
-                                         & {Recall}    & {24,67} & {26,7}  & \textbf{25,69} & {83,66} & {85,28}    & {86,03}  & \textbf{84,99} & {77,87} \\\midrule
-        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100} & {100} & {99,7} & \textbf{99,9} & {100}   \\
-                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {99,85} & {99,85}    & {100}    & \textbf{99,9}  & {100}   \\
-        \bottomrule
-    \end{tabular}
-    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
-The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
-    \label{tab:results:bb-re-generator_size-shape}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rr|cccccccc|c}
-        \toprule
-                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color} \\\midrule
-        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {88,89} & {85,9}  & {81,82} & {85,53} & {79,76} & {84,52}  & {87,76} & {87,36}  & \textbf{85,19} \\
-                                         & {Recall}    & {89,89} & {83,75} & {87,8}  & {79,27} & {93,06} & {76,34}  & {90,53} & {81,72}  & \textbf{85,3}  \\\midrule
-        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {69,51} & {72,73} & {73,48} & {66,27} & {63,23} & {64,32}  & {67,98} & {70,76}  & \textbf{68,53} \\
-                                         & {Recall}    & {72,43} & {71,29} & {73,48} & {60,77} & {74,6}  & {74,87}  & {71,13} & {73,57}  & \textbf{71,52} \\\midrule
-        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                           \\\textbf{color}} & {Precision}           & {37,63} & {39,23} & {22,09} & {20,23} & {33,33} & {22,97} & {24,9} & {43,96} & \textbf{30,54} \\
-                                         & {Recall}    & {29,44} & {42,68} & {27,48} & {22,22} & {25,44} & {26,53}  & {26,23} & {37,14}  & \textbf{29,64} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
-    \label{tab:results:bb-re-generator_color}
-\end{table}
-
-% With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-% Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-% The accuracy stays at the same value.
-% In contrast, the non-target accuracy is increased by 5\% points.
-% The model is more likely to identify a wrong object.
-% Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-% This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-% For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-% When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
-% This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-% SD: Not in the table?
-% DK: no, because a result of one specific sample (done)
-
-There are two main explanations for the big difference between the datasets.
-First, as already seen in the previous experiments, a bigger number of distractors confuses the model more where to focus on.
-In the \emph{Dale-2} dataset, there are only two possibilities, while the four distractors in the \emph{Dale-5} dataset and up to nine distractors in the \emph{CLEVR color} give the model a bigger choice.
-% SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated.
-I would only pose this as a very vague hypothesis to be tested, as you conclude now it is also the case that images with simialr distractors are more difficult to describe correctly.
-Hence, it is by no way given.
-% DK: TODO
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
-Otherwise, the shape is enough, and the caption is only one word long.
-The probability for this lies at 66,6\%.
-% SD: Why 66? We have 3 attributes, right, so 33%?
-% DK: but that the other attributes are only needed if the shape is the same. The prob. for the shape being different for the second object is 66% (rephrased; done)
-For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
-% SD: I think we need individual accuracies per class and also precision, recall and the f-score.
-If we include the joint accuracy figure when you also need to explain how you avergaed the infividual accuracies.
-% DK: time? (QUESTION)
-% SD: Not sure. Note that <pad> is a token. The model has to learn when to say <pad> just as any word, hence knowing when not to dewscribe is also trickly.
-% DK: true, but <pad> token is probably the most used token in general and therefore a guess of <pad> just by frequency is likely right? (QUESTION)
-\subsection{Reference resolution}
-% SD: Reference resolution task
-% DK: TODO
-\label{sec:reference_resolution}
-\subsubsection*{Setup}
-\cmtDK[inline]{attribute encodings, coordinate predictor}
-
-With the reference resolution task, the model is trained to understand referring expressions, by pointing towards the target object in the scene.
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human referring expressions into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-% SD: Reference resolution with identification of location.
-The other task, object identification is also resolving reference but it is easier as you are only picking objects.
-The first task also involves spatial knowledge.
-% DK: TODO
-With this approach, the models receive few natural language information, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are trained.
-In the simplest setup, the \emph{reference resolver}, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-% SD: What are the features?
-% SD: Note that these are visual features and not spatial features.
-It is true that visual features also encode some spatial information, how visual features relate to each other, but such information is very different from the spatial information required to predict coordinates in a coordinate frames, and hence we expect the task will be very hard.
-% DK: TODO
-In a first step, the image is encoded.
-This is done by first passing it through one of the \emph{feature extractors} and then further processing it with convolutional layers as described in section \ref{sec:image-processing}.
-Finally, it is projected to the image embedding size $e_i$.
-In the next step the encoded image is used to predict coordinates using the \emph{coordinate predictor}.
-Hereby, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to the coordinate predictor dimension $c$ and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-% SD: Very precisely. This is higher resolution that attention in the visual models that operates on o 7x7 grid, normally.
-% DK: TODO
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, four different ways to encode and refer to the target object are tested.
-
-\begin{figure}[ht]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-    % SD: How are these encoded?
-    % DK: TODO
-\end{figure}
-
-In the first method, the target object is encoded as \textbf{one-hot encodings}, as described in section \ref{sec:object-identification}.
-In an extension to this method, also the \textbf{center coordinates of all objects} in the image are included.
-% SD: Encoded locations?
-% DK: TODO
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For both methods, the encodings of the target object are concatenated with the encoded image and then passed to the coordinate predictor.
-% SD: Why shuffled? If we order them the way they appear in the image, the model will have more information that they are sequentially related.
-% DK: TODO
-% SD: Convolutions are only applied on visual features, not object attribute features and locations.
-% DK: TODO
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental GRE-algorithm}.
-This opposes the idea described before, to share as few natural language information as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-% SD: The question you are asking is whether it is possible to predict location from the visual appearance of the object.
-This is highly complex task as it requires quite two step reasoning, identification of features with that attribute, e.g.
-blue and then locating that feature in the image.
-% DK: TODO
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-% Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-% In other words, the object is described unambiguously using the lowest number of words.
-% The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-% This order defines, which attributes can be left out, while still identifying the object uniquely.
-% This research relies on the following order from most important to least important: shape, color, size
-% Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-% Using all three attributes, this description identifies the object perfectly and uniquely.
-% Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-% This can be taken even one step further by removing also the \emph{color}.
-% Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-% % SD: See my point earlier about the reversed algorithm between the two tasks 
-% % DK: TODO
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-In the model, the referring expression is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-Tokens are embedded with embedding dimensions $LSTM_e$ and the output of the LSTM has the size if $LSTM_o$.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated, which is then passed through the coordinate predictor.
-% SD: \citep and \citet Use Name (year) when you are referring to particular person and (Name, year) when you are referring to a paper. Hence, in this case it would be the latter.
 % DK: done
-
-The fourth method, to encode the target object utilizes \textbf{masking} of the image.
-% SD: Oh, before we talked about 3 methods
-% DK: done
-The image is masked in the same way as in section \ref{sec:referring_expression_generation}.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features and the inherent human bias in the image when looking at masked images.
-% SD: The information in the masked image is still not enough to identify the precise location, only the region.
-Hence, it corresponds on attention in the attention models.
-But those are generating labels and not predicting coordinated.
-The task is still challenging.
-% DK: TODO
-Both original and masked image are processed as described in \ref{sec:image-processing} and afterwards passed through the coordinate predictor.
-
-For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
-8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
-The loss is calculated using cross entropy.
-Table \ref{tab:variables-reference-resolution} shows the variables that are changed during the experiments for each of the models.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{lcccc}
-        \toprule
-                               & $e_i$              & $c$               & $LSTM_e$  & $LSTM_o$               \\
-                               & $[100, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[500,1000,1500,2000]$ \\\midrule
-        reference resolver     & \times             & \times            & -         & -                      \\
-        + one-hot              & \times             & \times            & -         & -                      \\
-        + one-hot + locations  & \times             & \times            & -         & -                      \\
-        + referring expression & \times             & \times            & \times    & \times                 \\
-        + masking              & \times             & \times            & -         & -                      \\
-        \bottomrule
-    \end{tabular}
-    \caption{Variables for each model where $e_i$ is the image embedding dimension, $c$ the dimensions of the coordinate predictor, $LSTM_e$ the embedding dimensions for tokens in the LSTM and $LSTM_o$ the output dimension of the LSTM}
-    \label{tab:variables-reference-resolution}
-\end{table}
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-% SD: Yes that’s correct but through several epochs we hope we will refine the distance and standard deviation of the error.
-It should level out.
-It is not a problem.
-What you do with a circle and accuracy is that you make the task easier as your pointer is now not pointing to a point but to a larger area.
-% DK: TODO
-It doesn't reflect this change.
-For that reason, we also introduced an accuracy score.
-For that we defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-% SD: But here the score will face the same problem with steward deviation. Hence the only difference here is that pointing is less precise.
-% DK: TODO
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The reference resolution models are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-% SD: Explain. w3 had a paper with John Kelleher on what spatial information is encoded in CNNs
-% DK: TODO
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-% SD: Point at not just discriminate
-% DK: TODO
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-% SD: How can we have epochs on the testing data? There is no updates so i5 does not make sense.
-% DK: TODO
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-% SD: The results show that the model is learning something from the training data but this is not the feature that should be learning as the performance on the test data is low.
-Actually, is it low, it is 10, 35 and 45 pixels.
-We should not expect any difference between epochs as there is no training.
-Hence, a flat line is expected.
-% DK: TODO
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-% SD: Because it is in the same location?
-% DK: TODO
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-% SD: What is the image size? What is the size of a typical object? 50 sounds quite good. Attention is predicting one of 7x7 blocks. Is 50 more or less than 1/7 of an image? 
-% DK: TODO
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-    % SD: Explanation of labels
-    % DK: TODO
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-% SD: Perhaps arranging the objects sequentially would help the model learn spatial contiguity, see my earlier comment.
-% DK: TODO
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-% SD: Table with these results?
-% DK: the table would show the same figures as the existing table. Should be still included? (QUESTION)
-
-\begin{figure}[ht]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, we visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-This shows general patterns of the models predictions over the complete dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-% SD: But this image does not show clearly whether the circles for the same object match. You could calculate the average error in distance between the ground truth and predicted coordinates.
-% DK: that is done in the quantitative analysis (mean square error above). The visualization should show general learned patterns of the model (as in this case predictions toard the center)
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-% SD: They are closer to the target than distractor and you can see that it is working to a point. But remember this task is very challenging and I wouldn't say the results are so negative.
-% DK: right, this is addressed in the conclusions below
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-% SD: There is a bias towards the centre but I would not say that the model has not learned anything.
-I'm surprised that it works so well given the feature representation we have, i.e.
-no geometric features.
-% DK: again, conclusion below
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-% SD: The centre bias could be the way the error function is used, that is averages all distances and of course this has a tendency to some middle distance.
-A solution would be to have better geometric features that could take these errors better into account.
-% DK: TODO (future work)
-
-\begin{figure}[ht]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    % SD: I Haven't checked the earlier captions, but the captions should be informative in the sense that one does not need to look for the text to understand them.
-Hence, include a brief summary of what each figure contains.
-    % DK: TODO
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-% SD: centre - using British English?
-% DK: so far I used always American English (hopefully consistently)
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-% SD: There is a centre bias and the model can partially locate the target.
-The task is hard.
-The features are not optimal for this task.
-Future would should focus on using geometric features - we should have tried these and I'm sure we would get better results.
-Another extension would be making pointing less precise, i.e.
-focus on a single point, i.e.
-using a circle of 20 pixels or a 7x7 grid as in the attention mechanism.
-This would simplify the task.
-% DK: TODO (future work)

### 2023-11-27 thesis/sections/5.1-object-identification.tex
setup grounding

\subsection{Object identification}
\label{sec:object-identification}
\subsubsection*{Setup}

In a first task a neural model is trained to discriminate multiple objects from each other.
Hereby, the model is shown bounding boxes of all objects in the scene as well as a description of the target object.
The model needs to combine all information and then point to the correct bounding box.
This task forces the model to extract visual features from the objects in the shown inputs and connect them to a representation of human given attributes.
Since the model is shown only bounding boxes around each object rather than the whole scene, geometric information about the objects' locations doesn't play a role for the success in this task.

In a first step, bounding boxes are extracted from each scene.
Each bounding box is a square with a side length of 96 pixels around the center coordinates of each object.
By doing this large objects in the front of the scene fill the bounding box completely, while smaller objects in the back of the scene contain some space around them.
They also might contain noise in the form of parts of adjacent objects.
Preliminary experiments indicated however that a more complex method for extracting the bounding boxes, depending on spatial position in the scene to reduce the noise didn't improve the results.
Therefore, bounding boxes in this thesis rely on the former simpler approach.
Since, each bounding box will be passed through on of the feature extractors, they are preprocessed as described in Section \ref{sec:feature-extractors}.
The array of bounding boxes is padded with matrix of zeros to the maximum possible number of objects present in a scene across the dataset.
For the 'Dale-2' dataset, this corresponds to a maximum of two bounding boxes, 5 bounding boxes for the 'Dale-5' dataset and 10 for the 'CLEVR color' dataset.
For each sample in the dataset the bounding boxes are shuffled.

The attributes of the target object are encoded as one-hot encodings.
There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the \emph{color} and a two-dimensional vector encoding the two different \emph{sizes}.
The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
These three encodings of the attributes are then concatenated to a vector with 13 dimensions.

\begin{figure}[ht]
    \centering
    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
    \caption{Simplified architecture of the object identification task}
    \label{fig:bounding_box_classifier_architecture}
    % SD: I haven't commented on this before, but let us be very careful in naming all meta-paramaters of the models in the text or in caption: how many layers, how many dimensions, etc. The reader should be able to replicate the code and the experiment from this. You can do this in the caption which is a longer explanation or in the text itself.
    % DK: TODO
\end{figure}

The model, the \textbf{object identifier} is split into three submodules, the \emph{bounding box encoder}, the \emph{attribute encoder} and the \emph{identifier}.
Each normalized bounding box of the sample is encoded with the \emph{bounding box encoder}.
Hereby, they are passed through \emph{ResNet-avg}.
The preceding layers \emph{ResNet-1} to \emph{ResNet-4} couldn't be compared due to memory restrictions on the GPU.
A deeper analysis of the effects of different feature extractors will be done in the following chapters.
The resulting vector is flattened and projected to the embedding dimension $e$ via a linear layer.
% SD: This is quite low? Why only 10?
% DK: different embedding sizes are tested (done)
Correspondingly, the one-hot vector of the attributes is projected to the same embedding dimension with another linear layer, using the \emph{attribute encoder}.
In the \emph{identifier} submodule, each bounding box is combined with the representation of the attributes.
By bringing both vectors to the same embedding size, the dot product between each embedded bounding box and the embedded attributes can be calculated.
A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
The model can so learn to connect the object representation with the attributes.
The dot products are concatenated and form a vector with as many dimensions as objects present.
The $softmax$ function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
% The model then points to an object, using the \emph{softmax} function over this resulting vector.

The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
The following embedding dimension $e$ are compared: 2, 10, 50, 100, 500, 1000.
% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
% DK: the output is a probability distriution over all objects (inserted above; done)

The object identification task is trained on all datasets that include more than one object in each sample, namely the 'CLEVR color', 'Dale-2' and 'Dale-5' dataset.
Random baselines for each dataset would yield accuracies of 10\%, 50\% and 20\% respectively.
% excluding the 'CLEVR single'.
% SD: Why not?
% DK: because there is only one object in the input, and therefore only one possible guess. The model can't point to a distractor (adressed; done)
The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.

\subsubsection*{Results}
Table \ref{tab:results:bounding_box_classifier} lists the accuracies of the models' predictions after 30 epochs for all three different datasets and different embedding sizes $e$.

\begin{table}[ht]
    \centering
    \begin{tabular}{c|c|c|c}
        \toprule
               & \textbf{Dale-2}   & \textbf{Dale-5}   & \textbf{CLEVR Color} \\\cmidrule(lr){2-2}\cmidrule(lr){3-3}\cmidrule(lr){4-4}
        $e$    & \textbf{Accuracy} & \textbf{Accuracy} & \textbf{Accuracy}    \\\midrule
        {2}    & {92\%}            & {72\%}            & {40\%}               \\
        {10}   & \textbf{100\%}    & {94\%}            & {92\%}               \\
        {50}   & {99\%}            & \textbf{95\%}     & \textbf{94\%}        \\
        {100}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
        {500}  & \textbf{100\%}    & \textbf{95\%}     & {93\%}               \\
        {1000} & \textbf{100\%}    & {94\%}            & \textbf{94\%}        \\
        \bottomrule
    \end{tabular}
    \caption{Accuracy scores of the object identifier after 30 epochs: $e$ are different embedding sizes}
    \label{tab:results:bounding_box_classifier}
\end{table}

The first trend that is visible is that fewer distractor increase the accuracy of the model.
The model achieves almost perfect accuracy in all configurations for the \emph{Dale 2} dataset with only two objects.
With five objects in the \emph{Dale-5} dataset, the accuracy drops to 95\% and with a maximum possible number of 10 objects in the \emph{CLEVR color} dataset the model only achieves 94\% in the best configuration.
This is not very surprising, since the model has a higher chance of predicting a wrong object, given the random baselines of 50\%, 20\% and 10\%.
Furthermore, when more objects are shown to the model, more attributes are needed to discriminate the target object from the distractor, since it is more likely that the distractors share attributes with the target object.
The task is getting therefore more challenging.

A second conclusion is that the model needs a certain embedding space, to represent the features of each image.
Across all dataset, an embedding space of only 2 dimensions result in a much lower performance compared to the best configuration.
Even though the model is still able to achieve an accuracy of 30\% to 50\% points over the random baseline, this shows that a bigger embedding size is beneficial to extract and represent the features of the objects.
Between 50 and 1000 dimensions, the model performs the best and varies only by 1\% point, which may be due to different random initialization of the weights in the model.

In conclusion, the model is able to learn to discriminate the objects based on the visual appearance.
The model can generate these high results, even with its relatively simple architecture.
In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
% SD: The model is able to leatn to discriminate objects based on visual appearance.
% DK: (done)

### 2023-11-27 thesis/sections/5.2-re-generation.tex
setup grounding

\subsection{Referring expression generation}
\label{sec:referring_expression_generation}
\subsubsection*{Setup}

Opposed to the previous experiment, where the focus lied on extracting visual features, the model is now tasked to generate referring expressions.
This is done in two setups.
The first setup uses bounding boxes as in the previous section as input, where the model needs to describe the target object of one of the bounding boxes.
In the second setup, the model is presented with the complete image and therefore also needs to accommodate to geometric information of the scene.

The bounding boxes are extracted and normalized in the same way as for the object identification task.
Since for the second setup, the images will also be passed through the feature extractors, they are normalized as well in like manner.
The referring expressions for the target object are generated using the incremental GRE-algorithm, described in Section \ref{sec:background:re}.
By this, the model needs to describe the target object with respect to the distractor objects.
There are some minor additions concerning the padding of the referring expression.
As before, the referring expression is padded to a number of three tokens, corresponding to the maximum of three attributes.
However, there are three different ways how the padding is applied.
First, the referring expression are, as usual in captioning tasks padded at the end with a specified padding token.
A problem could arise when the referring expression is not viewed as a natural language sentence, but as slots filled with tokens.
More specifically, following the GRE-algorithm, the last token in the referring expression is always the shape.
The second last token if existing describes the color, while the third last token if existing describes the size.
As soon as this sequence is padded at the end, these slots disappear.
A referring expression that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
Since this task is not focussing on producing natural language with a correct grammar, but focuses instead on extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
For this reason, the second method of padding the referring expression is prepending the referring expression with padding tokens.
By this, the positions of the slots are preserved and if not specified just filled with a padding token.
Each slot has always the same semantic value, e.g. the last slot always contains the shape of an object.
This can be done since the referring expressions are not free text, but instead the structure and the possible content is given by the dataset.
This method might help the model to learn the correct referring expressions.
% SD: Slots have the same semantic value, i.e. the last slot is the object type. We can do this because we know the structure of the descriptions in this dataset and because we hope this will help the learning. Also, we are thinking of generation based on slots.
% DK: (addressed; done)
The last variation concerns the order of producing each token.
When the referring expressions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
Even though a sequence of just three tokens may not be long enough for this factor to be a problem, we experimented to reverse the referring expression.
% SD: We... I sounds more like you are writing a lab report with casual language but here you want to be more formal.
% DK: (done)
Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the referring expression are still preserved.

This task inherently involves learning human knowledge and natural language structure.
Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
Can the model solve the task, or do specific attributes used by humans pose challenges to the model?
% Does it rely on the same attributes humans are using, or does it find other important differences, or is it not able to solve the task at all?
% SD: In this case it should be using the same attributes as humans as we are training it on human labels.
% DK: (rephrased; done)

\begin{figure}[ht]
    \centering
    \includegraphics[width=.5\linewidth]{figures/arch_bb_caption_generator.png}
    \caption{Simplified architecture of the bounding box RE generator}
    \label{fig:bb_re_generator_architecture}
\end{figure}

In the first setup, the \textbf{bounding box RE generator}, the model receives the bounding boxes as input (see Figure \ref{fig:bb_re_generator_architecture}).
The first bounding box is always the target object, while the remaining bounding boxes are shuffled.
As in the previous experiment, each bounding box is embedded, using the same \emph{bounding box encoder} submodule: first it is passed through \emph{ResNet-avg} and afterwards projected to an image embedding dimension $e$ with a linear layer.
The \emph{RE generator} submodule uses these embedding to produce a referring expression.
Hereby, all encoded bounding boxes are concatenated and again compressed to the decoder output dimension $LSTM_o$ using another linear layer.
This representation of all objects serves as the initial hidden state of an LSTM, which generates the referring expression.
Tokens used in the LSTM are embedded with embedding dimension $LSTM_e$.
During training, teacher forcing is applied by using embeddings of the ground truth tokens as the input sequence for the LSTM, instead of the output of the LSTM.
The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
During testing, the LSTM is always forced to generate three tokens, with an embedded start-of-sequence token as first input to the LSTM.
Each token in the sequence is determined greedily, by selecting the highest logit in the output of each step in the LSTM.

In the second setup two different models are compared against each other.
The first model, the \textbf{basic RE generator} acts as the baseline and only receives the complete image as the input.
The image is encoded, using the \emph{image encoder} submodule, described in Section \ref{sec:image-processing}, in particular it is passed through \emph{ResNet-3} and subsequently processed by several additional convolution layers.
The same \emph{RE generator} submodule as for the bounding boxes is used to produce the referring expression.
Hereby, since there is only one input scene, its encoding is directly reduced to the decoder output dimension $LSTM_o$.
The LSTM is trained in the same way as above.

\begin{figure}[ht]
    \centering
    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
    \caption{Simplified architecture of the masked RE generator}
    \label{fig:masked_re_generator_architecture}
    \cmtDK[inline]{add linear layer for LSTM}
\end{figure}

Using this approach, the model doesn't have any information about the target object and is therefore expected to produce referring expressions for a randomly selected object on the scene.
Therefore, it is extended in a second step with attention to the target object, the \textbf{masked RE generator} (see Figure \ref{fig:masked_re_generator_architecture}).
% SD: You add attention to the target object.
% DK: (added; done)
A masked version of the image is created and passed to the model.
For this, a fixed size squared area of 96 \times\ 96 pixels around the center of the target object is filled in white, while the rest of the image is filled in black.
Both original and masked image are processed together with the \emph{masked image encoder} submodule described in Section \ref{sec:image-processing}.
% SD: This is normally implemented as a filter of 1 and 0 corresponding to pixels. As such it then corresponds to attention, cf. paper with Mehdi.
% DK: (done)
Accordingly, the same \emph{RE generator} submodule is utilized here.
This should point the model towards which object to describe and discriminate from the distractors.

For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
Table \ref{tab:variables-reference-expression-generation} shows which variables are compared for each model:

\begin{table}[ht]
    \centering
    \begin{tabular}{lccc}
        \toprule
                                  & $e$                & $LSTM_o$           & $LSTM_e$       \\
                                  & $[100, 500, 1000]$ & $[100, 500, 1000]$ & $[10, 15, 30]$ \\\midrule
        bounding box RE generator & \times             & \times             & \times         \\
        basic RE generator        & -                  & \times             & \times         \\
        masked RE generator       & -                  & \times             & \times         \\
        \bottomrule
    \end{tabular}
    \caption{Variables for each model where $e$ is the embedding dimension, $LSTM_e$ the embedding dimensions for tokens in the LSTM and $LSTM_o$ the output dimension of the LSTM}
    \label{tab:variables-reference-expression-generation}
\end{table}

As already discussed before, this task can be interpreted as a classification task rather than a natural language generation task.
The main reason for this is that the model is tasked to assign specific attributes to the target object instead of producing free text with a large vocabulary.
Following, we are interested in the classification mistakes the model makes.
For this, the model's success is validated on accuracy, recall and precision scores.
These are calculated in the following ways.
The first measure is the \textbf{overall accuracy} if the model predicted every word in the referring expression correctly.
% SD: This is too strict. if this is a generation task then we can use generation metrics. Those compare if the generated string as a whole is kind of similar to the target string, see ROUGE, METEOR, BERTscore.
% If this is a classification task, and it is, given that we know what type of word we expect in each slot, we could calculate accuracy per class, precision, recall, F-score. We could also calculate mean ranks for predicting words per classes, or acucracy at k- if the target word is in the k-top predictions.
% We could also take it a multi-class classification problem and how to calculate accuracy in this case, would need to check.
% DK: I agree that classification metrics refelct success better here (therefore I also used accuracy). These metrics would give a better overview, but I don't think there is time to implement it (QUESTION)
This gives a hint, how the model fares in general and if it is able to predict any of the attributes.
However, a 'false' prediction doesn't give much insight into why the model predicted a wrong referring expression.

It could be the case that the model predicted the correct shape, but wrong color.
Even worse, the model could have predicted more attributes than necessary to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
% SD: Too far away, might be worth repeating here, you say still stay Fiigure 2d, repeated asFigure 6 here
% DK: TODO
The correct referring expression is \emph{cylinder}.
If the model would predict \emph{purple cylinder}, the accuracy determines it as false as generating the referring expression \emph{large purple cylinder} as well \emph{small green cube}.
The first two descriptions identify the target object perfectly, but the model only didn't learn to exclude unnecessary attributes.
To mitigate this, the accuracies for each class are included, as well as the macro average.
% SD: Is this figure for a particular class, i.e. object type, or do you then do something to individual accuracy, i.e. average them?
% DK: the metric is averaged over all samples (not by class). Might be definitly worth to do it class by class. Time? (QUESTION)
This can give a better understanding of the errors the model makes.
The same is done for \textbf{precision} and \textbf{recall}.

% SD: Very good If I understand this right this would be a case where the generated description describes another object better than the target and a human would identify that. So it is like false negatives in describing. Perhaps use another term for this: referring to a distractor? (since you are not measuring hwo individually the description fitts a distractor but only if the entire description could perfectly fit a distractor)
% DK: exactly. I added false negatives. But I would like to keep accuracy in the name, since that is what it is. maybe better distractor accuracy? (QUESTION)
With the \textbf{non-target accuracy}, we identify if the model described another object, which is not the target image.
This is basically an inverted accuracy score; the lower the score, the better the model fares.
By this, it measures the false negative generated referring expressions.
For this, referring generations are generated for all the non-target objects and distractors in the images using the GRE-algorithm.
Importantly to notice is that the referring expressions may not uniquely identify a distractor, since multiple distractors with the same shape, color and size are allowed.
If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
Using this measure, we can get an overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.

The RE generator models are trained on both 'Dale' datasets.
Again, each of these datasets increases the complexity of the description.
While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
Additionally, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
Furthermore, the models are trained on the 'CLEVR color' dataset.
In this dataset, the focus lies mainly on only one discriminating attribute, the color.
This might help the model in its prediction and will be tested in the experiments.

\subsubsection*{Results}
Table \ref{tab:results:bb-re-generator} shows the \emph{overall accuracy}, \emph{F1 scores} for each word and the \emph{non-target accuracies} of the \textbf{bounding box RE generator} when trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
As can be seen, the overall accuracies, in other words perfect matches of the generated referring expression depend very much on the dataset.
With the \emph{Dale-2} dataset, the model can achieve perfect matches in 99\% of the cases in its best configuration.
Also the \emph{CLEVR color} dataset allows the model to predict the correct referring expression in 93\% of the samples.
Opposed to that the model can only generate perfect referring expressions in 69\% of the samples of the \emph{Dale-5} dataset.
Looking at the \emph{non-target accuracy}, one of the problems can be identified for the \emph{CLEVR color} dataset.
Summing the \emph{non-target accuracy} and the \emph{overall accuracy}, one gets the accuracy that any of the shown object was described independently if it was the target object.
For the \emph{CLEVR color} dataset, this score lies at 96\% to 97\% in the best configurations, which means that the model described a shown object for almost all the samples.
In 3\% to 4\% of the cases, the model only picked the wrong object to describe.
This looks different for the \emph{Dale-5} dataset.
Here, the model describes only in 71\% of the cases one of the shown objects, 69\% the target object and 2\% a distractor.
The model therefore struggles more with the correct generation of words than with choosing, which object to describe.

When looking at the different configurations, in especially different values for $e$, $LSTM_o$ and $LSTM_e$, it can be seen that the results are not very different for the \emph{Dale-2} and \emph{CLEVR color} datasets.
However, bigger effects can be identified for the \emph{Dale-5} dataset.
Here, a low output dimension of the LSTM $LSTM_o$ tends to give lower scores.
This especially enhanced, when also the embedding dimensions of the tokens $LSTM_e$ is low.
Again, the image embedding size $e$ doesn't seem to have a big effect, when over 50 dimensions.
Indeed, the model achieves an accuracy of 69\% with all three tested embedding sizes.

\begin{table}[ht]
    \centering
    \begin{tabular}{ccc|ccc|ccc|ccc}
        \toprule
               &          &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                               \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
        $e$    & $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} \\\midrule
        {100}  & {100}    & {10}     & {97}                                & {97,67}                             & {1}                                      & {65}          & {88,17}        & {2}         & {92}          & {94,95}        & {3}         \\
        {100}  & {100}    & {15}     & {97}                                & {97,77}                             & {0}                                      & {62}          & {86,51}        & {2}         & {92}          & {94,45}        & {4}         \\
        {100}  & {100}    & {30}     & {98}                                & {97,81}                             & {1}                                      & {65}          & {87,97}        & {2}         & {92}          & {94,47}        & {4}         \\
        {100}  & {500}    & {10}     & {98}                                & {98,16}                             & {0}                                      & {67}          & {88,13}        & {2}         & {93}          & {95,18}        & {4}         \\
        {100}  & {500}    & {30}     & \textbf{99}                         & \textbf{98,57}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,53} & \textbf{2}  & \textbf{93}   & \textbf{95,17} & \textbf{3}  \\
        {100}  & {1000}   & {10}     & {98}                                & {98,28}                             & {0}                                      & {68}          & {88,23}        & {2}         & \textbf{93}   & \textbf{95,41} & \textbf{3}  \\
        {500}  & {1000}   & {30}     & \textbf{99}                         & \textbf{98,85}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,15} & \textbf{2}  & {93}          & {95,05}        & {4}         \\
        {500}  & {500}    & {10}     & \textbf{99}                         & \textbf{98,71}                      & \textbf{0}                               & {67}          & {88,99}        & {2}         & \textbf{93}   & \textbf{95,4}  & \textbf{3}  \\
        {1000} & {100}    & {10}     & {97}                                & {97,72}                             & {0}                                      & {59}          & {85,43}        & {2}         & {93}          & {95,04}        & {4}         \\
        {1000} & {100}    & {30}     & {98}                                & {98,16}                             & {0}                                      & {61}          & {86,77}        & {2}         & {92}          & {94,68}        & {4}         \\
        {1000} & {1000}   & {10}     & \textbf{99}                         & \textbf{98,95}                      & \textbf{0}                               & \textbf{69}   & \textbf{89,21} & \textbf{2}  & {93}          & {95,15}        & {4}         \\
        {1000} & {1000}   & {15}     & {99}                                & {98,67}                             & {0}                                      & {68}          & {88,98}        & {2}         & {93}          & {95,01}        & {4}         \\
        \bottomrule
    \end{tabular}
    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the bounding box RE generator after 30 epochs: $e$ are different embedding sizes, $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
    \label{tab:results:bb-re-generator}
\end{table}

Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results and especially what mistakes the model is making for both the \emph{Dale-5} and \emph{CLEVR color} datasets.
They list \emph{precision} and \emph{recall} metrics for each token for the best configuration of the model with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
The tokens are grouped by attribute and also show the metrics averaged over each of the attributes.
Since the overall results for the \emph{Dale-2} dataset are already close to perfect, the focus for this analysis lies on the remaining two datasets.
The metrics of the \emph{<pad>} token indicate if the model produced the correct length of the referring expression, in other words if it was able to determine which attributes are necessary to discriminate the target object from the distractors.
For the \emph{CLEVR color} dataset, the scores are perfect.
This is not surprising, because all referring expressions for the \emph{CLEVR color} dataset consist of exactly two attributes, shape and color, and the first generated token will always be the only \emph{<pad>} token in the referring expression (corresponding to the unspecified size).
The \emph{<pad>} token is therefore easy to learn.
For the \emph{Dale-5} dataset, the model struggles more to predict the correct length of the referring expression.
\cmtDK[inline]{why?}

When looking at the tokens for the shape, it can be seen that the model is able to identify it very well across all datasets.
The model predicts the correct shape for all samples using the \emph{CLEVR color} dataset, while both \emph{precision} and \emph{recall} lie around 98,3\% when using the \emph{Dale-5} dataset.
Even though the score is almost perfect, the slight difference might stem from the fact that all distractors have the same shape in the first case, while distractors can be different in the second case.
Consequently, the model is only exposed to one shape at a time for each sample, which might simplify its identification.

For the color attribute, the metrics drop significantly for both \emph{Dale-5} and \emph{CLEVR color} to an average of around 93\%.
Hereby, no meaningful difference can be seen across the datasets, but there are differences between the colors.
Some colors are predicted with \emph{precision} and \emph{recall} around 95\% to 96\%, while others are only around 90\%.
However, these differences are not reproducible across multiple runs and configurations.
The best and worst predicted colors vary and no conclusions can be drawn which colors are easier to predict for the model.

Finally, the size is the most difficult attribute to predict for the model.
Apart from the \emph{CLEVR color} dataset, where a size never needs to be predicted and also is never predicted, the metrics for the prediction of size tokens are the lowest across all tokens.
They are the only mistakes, the model makes, when exposed to the \emph{Dale-2} dataset and the average \emph{precision} lies around 23\% below the average of predictions of the color for the \emph{Dale-5} dataset, while the average \emph{recall} lies around 28,82\% below.
The reason why the \emph{precision} is higher than the \emph{recall} is the \emph{<pad>} token, which is predicted very often instead of a token specifying the size.
In fact, the opposite relationship is visible for the \emph{precision} and \emph{recall} for said token.
The much higher absolute number of \emph{<pad>} tokens leads to a smaller relative difference of \%-points shown in the table.
Again, no conclusion can be drawn if larger or smaller objects are easier to predict, since the results vary across runs and configurations.

\begin{table}[ht]
    \centering
    \begin{tabular}{rr|cc|c|ccc|c|c}
        \toprule
                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {99,17} & {98,29} & \textbf{98,73} & {99,86} & {99,71}    & {99,67}  & \textbf{99,75} & {99,64} \\
                                         & {Recall}    & {97,54} & {94,26} & \textbf{95,9}  & {100}   & {99,56}    & {99,67}  & \textbf{99,74} & {99,77} \\\midrule
        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {69,65} & {69,21} & \textbf{69,43} & {98,19} & {98,32}    & {98,39}  & \textbf{98,3}  & {82,22} \\
                                         & {Recall}    & {62,11} & {66,15} & \textbf{64,13} & {98,79} & {97,87}    & {98,25}  & \textbf{98,3}  & {84,59} \\\midrule
        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
        \bottomrule
    \end{tabular}
    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
    \label{tab:results:bb-re-generator_size-shape}
\end{table}

\begin{table}[ht]
    \centering
    \begin{tabular}{rr|cccccccc|c}
        \toprule
                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color} \\\midrule
        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {94,51} & {98,77} & {97,59} & {98,68} & {98,89} & {98,8}   & {97,47} & {100}    & \textbf{98,09} \\
                                         & {Recall}    & {97,73} & {100}   & {98,78} & {97,4}  & {96,74} & {98,8}   & {100}   & {98,8}   & \textbf{98,53} \\\midrule
        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {92,12} & {93,82} & {89,13} & {89,12} & {92,63} & {91,12}  & {97,24} & {94,36}  & \textbf{92,44} \\
                                         & {Recall}    & {92,12} & {89,78} & {94,91} & {94,51} & {95,71} & {92,42}  & {89,34} & {94,85}  & \textbf{92,95} \\\midrule
        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                           \\\textbf{color}} & {Precision}           & {93,46} & {92,37} & {94,47} & {93,86} & {92,04} & {91,13}  & {90,07} & {94,7}   & \textbf{92,76} \\
                                         & {Recall}    & {92,75} & {92}    & {95,98} & {89,92} & {94,12} & {91,13}  & {94,23} & {91,91}  & \textbf{92,76} \\
        \bottomrule
    \end{tabular}
    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
    \label{tab:results:bb-re-generator_color}
\end{table}

The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
When the order was reversed and the padding appended, the model was converging slightly faster and reached the limit around two to three epochs earlier.
The final peak stayed exactly the same and the effects were therefore not studied deeper.
% SD: Earlier you say that you will only try one method but now you have tried both. Change earlier text; also do you have full performance figures for this setup to include?
% DK: where? I can't find it. (QUESTION). Since they are the same (and are therefore not central in this study), i didn't inlcude them. (QUESTION)

In conclusion, it can be said that the model is able to extract discriminative features from the shown bounding boxes and produce referring expressions.
However, the result highly depends on the amount of distractors and the resulting need of more discriminative features to describe the target object.
While the shape is easily identified, the model has bigger problems to identify the color and especially the size of the target object.

In the following paragraphs, the results of the \textbf{basic RE generator} as well as the \textbf{masked RE generator} are evaluated.
Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 40 epochs.
Compared to the \emph{bounding box RE generator}, the task differs substantially.
Instead of bounding boxes, the model is shown the image of the whole scene.
Additionally to extracting features of the target object and distractors separately, the model is now tasked to extract these at once for all objects.
Furthermore, it needs to learn which of the objects is the target object by combining the image with the masked image.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|ccc|ccc|ccc}
        \toprule
                 &          & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                               \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
        $LSTM_o$ & $LSTM_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} & \textbf{Acc.} & \textbf{F1}    & \textbf{NT} \\\midrule
        {100}    & {10}     & {81}                                & {76,16}                             & {9}                                      & {21}          & {57,07}        & {23}        & {16}          & {43,69}        & {31}        \\
        {100}    & {15}     & {78}                                & {72,06}                             & {11}                                     & {17}          & {53,29}        & {25}        & {26}          & {50,87}        & {30}        \\
        {100}    & {30}     & {83}                                & {78,76}                             & {7}                                      & {19}          & {54,91}        & {24}        & \textbf{29}   & \textbf{53,37} & \textbf{26} \\
        {500}    & {10}     & {83}                                & {79,78}                             & {6}                                      & {27}          & {62,05}        & {16}        & {17}          & {44,93}        & {32}        \\
        {500}    & {15}     & {83}                                & {79,75}                             & {7}                                      & {28}          & {63,78}        & {16}        & {21}          & {47,08}        & {31}        \\
        {500}    & {30}     & {84}                                & {83,03}                             & {5}                                      & {30}          & {64,36}        & {17}        & {21}          & {40}           & {31}        \\
        {1000}   & {10}     & {84}                                & {82,2}                              & {5}                                      & {31}          & {65,58}        & {13}        & {12}          & {37,67}        & {34}        \\
        {1000}   & {15}     & {85}                                & {82,62}                             & {5}                                      & {31}          & {64,28}        & {16}        & {16}          & {43,27}        & {32}        \\
        {1000}   & {30}     & {85}                                & {83,13}                             & {4}                                      & {31}          & {64,06}        & {17}        & {13}          & {41,79}        & {34}        \\
        {1500}   & {10}     & {84}                                & {81,27}                             & {4}                                      & {33}          & {67,5}         & {14}        & {12}          & {39,04}        & {34}        \\
        {1500}   & {15}     & \textbf{85}                         & \textbf{84,11}                      & \textbf{4}                               & {35}          & {69,05}        & {12}        & {17}          & {44,98}        & {31}        \\
        {1500}   & {30}     & {84}                                & {80,93}                             & {4}                                      & {36}          & {68,97}        & {15}        & {14}          & {42,99}        & {33}        \\
        {2000}   & {10}     & {83}                                & {82,69}                             & {4}                                      & \textbf{36}   & \textbf{69,11} & \textbf{13} & {13}          & {41,85}        & {34}        \\
        {2000}   & {15}     & {84}                                & {81,95}                             & {4}                                      & {34}          & {66,94}        & {14}        & {14}          & {38,79}        & {33}        \\
        {2000}   & {30}     & {85}                                & {82,8}                              & {4}                                      & {32}          & {65,52}        & {16}        & {18}          & {45,41}        & {32}        \\
        {3000}   & {10}     & {82}                                & {79,83}                             & {4}                                      & {34}          & {67,61}        & {12}        & {12}          & {38}           & {34}        \\
        {3000}   & {15}     & {85}                                & {81,8}                              & {4}                                      & {32}          & {65,64}        & {15}        & {13}          & {38,52}        & {35}        \\
        {3000}   & {30}     & {83}                                & {80,6}                              & {3}                                      & {30}          & {63,5}         & {16}        & {12}          & {38,45}        & {34}        \\
        \bottomrule
    \end{tabular}
    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked RE generator after 40 epochs: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
    \label{tab:results:masked-re-generator}
    % SD: Masking does not work but perhaps this is because you applied convolutional pre-trained filters rather than 0-1. It is harder to detect meaningful patterns since the targets will be smaller and more integrated with other objects.
    % DK: might be the case. in the last language game i use only a linear layer instead of convolutions, but it doesn't help. This however might also be due to other problems. Time? (QUESTION, future work)
    % SD: I would present these for individual classes.
    % DK: I only have accuracy scores at. Time? (QUESTION)
\end{table}

Consequently, the results are much less accurate.
This applies to all datasets, even though it is most apparent for the \emph{CLEVR color} dataset.
While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the \emph{Dale-2} dataset, the model only reaches 36\% for the \emph{Dale-5} dataset and 29\% for the \emph{CLEVR color} dataset.
However, these scores are all well above the random baselines of 50\%, 20\% and 10\% respectively.
Moreover, the referring expression generated for the \emph{Dale-2} dataset are also efficient in the sense that they follow incremental GRE-algorithm and only use necessary discriminative attributes.
Features of both target object and distractor are therefore extracted and associated with the vocabulary.
Striking in these results is also the \emph{non-target accuracy}, which is much higher than for the \emph{bounding box RE generator}.
For the \emph{Dale-2} dataset, it arrives at 11\% in one configuration, for the \emph{Dale-5} dataset, the model describes for at least 12\% of the samples a distractor, for one configuration even in over 25\% of the cases.
The model is most uncertain which object to describe when exposed to the \emph{CLEVR color} dataset where in almost all configurations, it describes a distractor every third sample.
This shows that the model struggles to learn which of the objects in the scene is the target object, more specifically, to combine the whole image with the masked image.
\cmtDK[inline]{compare to basic RE generator}
Yet, the model is able to extract the attributes from the shown objects.
This can be seen, when summing up the \emph{overall accuracy} and the \emph{non-target accuracy}.
This number represents in how many cases any of the shown objects were described.
For the \emph{Dale-2} dataset, this number lies at 90\% for the best configuration, at 51\% for the \emph{Dale-5} dataset and at 56\% for the \emph{CLEVR color} dataset.
While being still lower than the same metric for the \emph{bounding box RE generator}, the model fares well in comparison for the \emph{Dale} datasets given the highly increased complexity of the task.

Looking at the effects of the LSTM output size $LSTM_o$ and the LSTM embedding size $LSTM_e$, a difference between the \emph{Dale} datasets and the \emph{CLEVR color} dataset can be identified.
A higher $LSTM_o$ tends to increase the performance of the model, when presented with the \emph{Dale} datasets until a size of 1500 to 2000 dimensions.
When it is further increased the model starts to perform worse.
Furthermore, with a low $LSTM_o$, the model is more likely to identify and describe a distractor, while a higher $LSTM_o$ helps the model to focus on the target object.
The $LSTM_e$ however doesn't seem to have a big consistent effect on the results. With an $LSTM_o$ of around 1500 to 2000, the model seems to fare slightly better with higher $LSTM_e$ of 15 or 30 dimensions.
For the \emph{CLEVR color} dataset, the results look different.
Opposed to the former datasets, a lower $LSTM_o$ helps the model to describe the correct target object.
On the other side, the \emph{non-target accuracy} stays relatively constant independently of the variables, in other words

\cmtDK[inline]{discuss attributes}
\begin{table}[ht]
    \centering
    \begin{tabular}{rr|cc|c|ccc|c|c}
        \toprule
                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {68,63} & {76}    & \textbf{72,31} & {98,41} & {98,18}    & {98,92}  & \textbf{98,5}  & {94,77} \\
                                         & {Recall}    & {30,17} & {34,55} & \textbf{32,36} & {98,7}  & {98,33}    & {98,47}  & \textbf{98,5}  & {98,64} \\\midrule
        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {49,47} & {57,3}  & \textbf{53,38} & {87,09} & {83,26}    & {84,78}  & \textbf{85,04} & {64,99} \\
                                         & {Recall}    & {24,67} & {26,7}  & \textbf{25,69} & {83,66} & {85,28}    & {86,03}  & \textbf{84,99} & {77,87} \\\midrule
        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100} & {100} & {99,7} & \textbf{99,9} & {100}   \\
                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {99,85} & {99,85}    & {100}    & \textbf{99,9}  & {100}   \\
        \bottomrule
    \end{tabular}
    \caption{Precision and Recall in \% of <pad>, size and shape tokens in the best configuration for each dataset (marked in bold in Table \ref{tab:results:masked-re-generator}). The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
    \label{tab:results:masked-re-generator_size-shape}
\end{table}

\begin{table}[ht]
    \centering
    \begin{tabular}{rr|cccccccc|c}
        \toprule
                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color} \\\midrule
        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {88,89} & {85,9}  & {81,82} & {85,53} & {79,76} & {84,52}  & {87,76} & {87,36}  & \textbf{85,19} \\
                                         & {Recall}    & {89,89} & {83,75} & {87,8}  & {79,27} & {93,06} & {76,34}  & {90,53} & {81,72}  & \textbf{85,3}  \\\midrule
        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {69,51} & {72,73} & {73,48} & {66,27} & {63,23} & {64,32}  & {67,98} & {70,76}  & \textbf{68,53} \\
                                         & {Recall}    & {72,43} & {71,29} & {73,48} & {60,77} & {74,6}  & {74,87}  & {71,13} & {73,57}  & \textbf{71,52} \\\midrule
        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                           \\\textbf{color}} & {Precision}           & {37,63} & {39,23} & {22,09} & {20,23} & {33,33} & {22,97} & {24,9} & {43,96} & \textbf{30,54} \\
                                         & {Recall}    & {29,44} & {42,68} & {27,48} & {22,22} & {25,44} & {26,53}  & {26,23} & {37,14}  & \textbf{29,64} \\
        \bottomrule
    \end{tabular}
    \caption{Precision and Recall in \% of color tokens in the best configuration for each dataset (marked in bold in Table \ref{tab:results:masked-re-generator}). The column \textbf{color} shows the average across all colors.}
    \label{tab:results:masked-re-generator_color}
\end{table}

% With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
% Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
% The accuracy stays at the same value.
% In contrast, the non-target accuracy is increased by 5\% points.
% The model is more likely to identify a wrong object.
% Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
% This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
% For instance a sample with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
% When also the masked image is passed to the model, it generates the description \emph{small blue cube}.
% This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
% SD: Not in the table?
% DK: no, because a result of one specific sample (done)

There are two main explanations for the big difference between the datasets.
First, as already seen in the previous experiments, a bigger number of distractors confuses the model more where to focus on.
In the \emph{Dale-2} dataset, there are only two possibilities, while the four distractors in the \emph{Dale-5} dataset and up to nine distractors in the \emph{CLEVR color} give the model a bigger choice.
% SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated. I would only pose this as a very vague hypothesis to be tested, as you conclude now it is also the case that images with simialr distractors are more difficult to describe correctly. Hence, it is by no way given.
% DK: TODO
The second explanation lies in the used GRE-algorithm.
When only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
Otherwise, the shape is enough, and the RE is only one word long.
The probability for this lies at 66,6\%.
% SD: Why 66? We have 3 attributes, right, so 33%?
% DK: but that the other attributes are only needed if the shape is the same. The prob. for the shape being different for the second object is 66% (rephrased; done)
For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
With four distractors the algorithm is much more likely to produce longer referring expressions.
These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
% SD: I think we need individual accuracies per class and also precision, recall and the f-score. If we include the joint accuracy figure when you also need to explain how you avergaed the infividual accuracies.
% DK: time? (QUESTION)
% SD: Not sure. Note that <pad> is a token. The model has to learn when to say <pad> just as any word, hence knowing when not to dewscribe is also trickly.
% DK: true, but <pad> token is probably the most used token in general and therefore a guess of <pad> just by frequency is likely right? (QUESTION)

### 2023-11-27 thesis/sections/5.3-reference-resolution.tex
setup grounding

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
-The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
+\subsection{Reference resolution}
+% SD: Reference resolution task
+% DK: (done)
+\label{sec:reference_resolution}
 \subsubsection*{Setup}
-
+With the reference resolution task, the model is trained to understand referring expressions, by pointing towards the target object in the scene.
+Compared to the object identification task in Section \ref{sec:object-identification}, this task is more complex since it also includes identifying locations instead of choosing from given objects.
 This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
@@
 The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
+Including human referring expressions into the task would bias also the emerged language towards attributes and words, used in natural language.
 For this reason, the final task of the receiver will be to 'point' to the target object.
 The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
+% SD: Reference resolution with identification of location.
+The other task, object identification is also resolving reference but it is easier as you are only picking objects.
+The first task also involves spatial knowledge.
+% DK: (done)
+With this approach, the models receive few natural language information, but are still able to rely on all information present in the image to discriminate the objects.
 
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
+To achieve this goal, multiple setups of models are trained.
+In the simplest setup, the \textbf{reference resolver}, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+% SD: What are the features?
+% SD: Note that these are visual features and not spatial features.
+It is true that visual features also encode some spatial information, how visual features relate to each other, but such information is very different from the spatial information required to predict coordinates in a coordinate frames, and hence we expect the task will be very hard.
+% DK: TODO
+Again, the image is encoded, using the \emph{image encoder} submodule, described in Section \ref{sec:image-processing}.
+The \emph{coordinate predictor} submodule takes the encoder image to predict x- and and y-coordinates of the target image,
+Hereby, the encoded image is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
+These reduce the dimensions first to the coordinate predictor dimension $c$ and finally to 2.
 
@@
 
-With this simple setup, the model is theoretically able to focus on an object in the image.
+With this simple setup, the model is theoretically able to very precisely focus on an object in the image.
+% SD: Very precisely. This is higher resolution that attention in the visual models that operates on o 7x7 grid, normally.
+% DK: (done)
 The problem arises as soon as multiple objects are present in the image.
@@
 Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
+Here, four different ways to encode and refer to the target object are tested.
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
     }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
+    \subfigure[Model including incremental referring expression]{
+        \includegraphics[width=0.34\linewidth]{figures/arch_coordinate_predictor_dale.png}
         \label{fig:coordinate_predictor_dale_architecture}
@@
     \caption{Simplified architecture of the coordinate predictors}
+    % SD: How are these encoded?
+    % DK: (done)
 \end{figure}
 
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
+In the first method, the target objects' attributes are encoded as \textbf{one-hot encodings}, as described in Section \ref{sec:object-identification}.
+In an extension to this method, also the \textbf{center coordinates of all objects} in the image are included.
+% SD: Encoded locations?
+% DK: (done)
+The center coordinates of all objects are simply extracted.
+Since there is no direct way of ordering objects, distributed in 2 dimensions into a list, these are shuffled, to just provide the model with all possible targets to choose from.
 Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
 The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
+In both methods, the encodings of the target object are concatenated with the encoded image and then passed to the coordinate predictor.
+% SD: Why shuffled? If we order them the way they appear in the image, the model will have more information that they are sequentially related.
+% DK: (done)
+% SD: Convolutions are only applied on visual features, not object attribute features and locations.
+% DK: (done)
 
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
+The third method encodes the attributes of the target object with human language using the \textbf{incremental GRE-algorithm}.
+This opposes the idea described before, to share as few natural language information as possible with the model.
 Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
+% SD: The question you are asking is whether it is possible to predict location from the visual appearance of the object.
+This is highly complex task as it requires quite two step reasoning, identification of features with that attribute, e.g.
+blue and then locating that feature in the image.
+% DK: TODO
 If the model is able to match parts of the image with human words it would show that the model learned this attribute.
@@
 
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
+% Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
+% In other words, the object is described unambiguously using the lowest number of words.
+% The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
+% This order defines, which attributes can be left out, while still identifying the object uniquely.
+% This research relies on the following order from most important to least important: shape, color, size
+% Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
+% Using all three attributes, this description identifies the object perfectly and uniquely.
+% Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
+% This can be taken even one step further by removing also the \emph{color}.
+% Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
+% % SD: See my point earlier about the reversed algorithm between the two tasks 
+% % DK: TODO
 
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
+For the experiments, each image is captioned with a referring expression of the target object using the described algorithm.
+To include it in the model, the referring expressions need to be padded to an equal length.
 In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
+For this, as standard practice in captioning tasks, the referring expressions are padded at the end with a specified padding token.
 
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
+In the model, the referring expression is encoded, using the \emph{GRE parser} submodule.
+Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete referring expression.
+Tokens are embedded with embedding dimensions $LSTM_e$ and the output of the LSTM has the size if $LSTM_o$.
+The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
+Both the processed image and the final hidden state of the LSTM are flattened and concatenated, which is then passed through the coordinate predictor.
+% SD: \citep and \citet Use Name (year) when you are referring to particular person and (Name, year) when you are referring to a paper. Hence, in this case it would be the latter.
+% DK: done
 
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
+The fourth method, to encode the target object utilizes \textbf{masking} of the image.
+% SD: Oh, before we talked about 3 methods
+% DK: done
+The image is masked in the same way as in Section \ref{sec:referring_expression_generation}.
+While even the one-hot encodings contain human language knowledge by translating human referring descriptions into a vector, the masking method will only point the model towards the target object without giving more information.
+It therefore can only rely on its own extracted visual features and the inherent human bias in the image when looking at masked images.
+% SD: The information in the masked image is still not enough to identify the precise location, only the region.
+Hence, it corresponds on attention in the attention models.
+But those are generating labels and not predicting coordinated.
+The task is still challenging.
+% DK: TODO
+Both original and masked image are processed as described in \ref{sec:image-processing} and afterwards passed through the coordinate predictor.
+
+For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
+8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
+The loss is calculated using cross entropy.
+Table \ref{tab:variables-reference-resolution} shows the variables that are changed during the experiments for each of the models.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lcccc}
+        \toprule
+                              & $e_i$              & $c$               & $LSTM_e$  & $LSTM_o$               \\
+                              & $[100, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[500,1000,1500,2000]$ \\\midrule
+        reference resolver    & \times             & \times            & -         & -                      \\
+        + one-hot             & \times             & \times            & -         & -                      \\
+        + one-hot + locations & \times             & \times            & -         & -                      \\
+        + incremental RE      & \times             & \times            & \times    & \times                 \\
+        + masking             & \times             & \times            & -         & -                      \\
+        \bottomrule
+    \end{tabular}
+    \caption{Variables for each model where $e_i$ is the image embedding dimension, $c$ the dimensions of the coordinate predictor, $LSTM_e$ the embedding dimensions for tokens in the LSTM and $LSTM_o$ the output dimension of the LSTM}
+    \label{tab:variables-reference-resolution}
+\end{table}
 
@@
 
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
+The reference resolution models are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
 The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
+Since the model relies on the extracted features of either VGG or ResNet, locational information of the objects in the image might not be present anymore.
+This is due the fact that CNNs compress information of the image summarizing small regions (convolutions) several times.
+While this extracts the key features of objects in an image and might capture relational position between these objects, absolute locations might get lost as shown by \citet{Kelleher2017}.
+% SD: Explain. w3 had a paper with John Kelleher on what spatial information is encoded in CNNs
+% DK: (done)
 Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
 In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
+Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the incremental GRE-algorithm or the encoded locations.
+'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors and then point towards it.
+% SD: Point at not just discriminate
+% DK: (done)
 Latter task is assumed to be significantly harder.
@@
     \label{tab:feature-extractor-mean-distances}
+    % SD: Explanation of labels
+    % DK: TODO
 \end{table}
@@
 
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
+In a next step, information about the attributes is included using the \textbf{incremental GRE-algorithm}.
 Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
+% SD: Table with these results?
+% DK: the table would show the same figures as the existing table. Should be still included? (QUESTION)
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
+% SD: The centre bias could be the way the error function is used, that is averages all distances and of course this has a tendency to some middle distance.
+A solution would be to have better geometric features that could take these errors better into account.
+% DK: TODO (future work)
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
     \caption{Visualization of the models' predictions in the Dale datasets}
+    % SD: I Haven't checked the earlier captions, but the captions should be informative in the sense that one does not need to look for the text to understand them.
+Hence, include a brief summary of what each figure contains.
+    % DK: TODO
     \label{fig:visualizations_dale_complete}
@@
 Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
-These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-    \caption{Sender and receiver architectures in the discrimination game}
-    \label{fig:discriminator_architecture}
-\end{figure}
-
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
-        \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_discriminator}
-\end{table}
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-
-\cmtDK{calculation of loss?}
-\cmtDK{length}
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are 48 possible combinations of attributes.
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
-
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
-
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
-    \label{fig:caption_generator_game_architecture}
-\end{figure}
-
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
-    \label{fig:coordinate_predictor_game_architecture}
-\end{figure}
-
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
-
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
-        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.
+% SD: There is a centre bias and the model can partially locate the target.
+The task is hard.
+The features are not optimal for this task.
+Future would should focus on using geometric features - we should have tried these and I'm sure we would get better results.
+Another extension would be making pointing less precise, i.e.
+focus on a single point, i.e.
+using a circle of 20 pixels or a 7x7 grid as in the attention mechanism.
+This would simplify the task.
+% DK: TODO (future work)

### 2023-11-27 thesis/sections/6-grounding-language-games.tex
setup grounding

@@
 \cmtDK[inline]{discuss training/epochs/batches, vocab size, message length, gs temperature}
-
-\subsection{Object identification}
-\subsubsection*{Setup}
-
-The object identification game is set up as a discrimination game.
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citep{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-% SD: On the contrary,...
-% DK: TODO
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-The data is prepared in the same way as for the single model object identification task in section \ref{sec:object-identification}.
-Bounding boxes of all objects in the scene are extracted and preprocessed to be used for the feature extractors.
-Accordingly, also the models from the single model tasks serve as basis for the object identification task.
-The sender uses the architecture of the model described in section \ref{sec:referring_expression_generation} apart from the final LSTM.
-Instead, it will use the LSTM with the hidden size $h_s$ included in the \emph{EGG} framework to produce a message.
-Based on the results from the single model object identifiers, the image embedding size $e_s$ is fixed to 100.
-The bounding box of the target object is passed first, the remaining are shuffled.
-
-The receiver instead uses the architecture of the single model object identifier, described in section \ref{sec:object-identification}.
-Similarly to the sender, the LSTM of the \emph{EGG} framework with the hidden size $h_e$ is used to encode the sender's message instead of the one-hot encoded attributes in the single model.
-The hidden state of the LSTM is projected to the embedding size $e_r=500$ with a linear layer, based on the results in the previous experiments.
-As before the dot product between the embedded message and the encoded image is calculated to let the model point towards one of the bounding boxes.
-Opposed to the sender, the receiver is shown all bounding boxes in shuffled order.
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-    \caption{Sender and receiver architectures in the discrimination game}
-    \label{fig:discriminator_architecture}
-\end{figure}
-
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
-The loss is calculated using cross entropy.
-The following values for the variables are compared:
-\begin{itemize}
-    \item $h_s$: 100, 500, 1000
-    \item $h_r$: 10, 30, 50, 100, 500, 1000
-    \item $|V|$: 1, 10, 16, 50, 100
-    \item $n$: 1, 2, 3, 4, 5, 6
-\end{itemize}
-The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
-
-\subsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-% SD: Motivation why this is so. The possible number of strings is {#colour} x {#shape} {object type} where all numbers except the last one also contain a padding token.
-% DK: TODO
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-% SD: This discussion of dimensions could already be introduced when you discuss simple modles earlier as the same dimensions are used.
-This is a more an evaluation of represnetations rather than interaction and hence it would fit there.
-% DK: TODO
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
-        \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    % SD: A note about the datasets chosen. Why do we choose these? Initially, we describe more than 3 datasets.
-    % DK: because the 'CLEVR Single' contains only one object. So there is nothing to discriminate (done)
-    \label{tab:results_discriminator}
-\end{table}
-% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
-As David pointed out it may be that convergence is unstable and changes happen in between.
-% DK: TODO
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-% SD: How do we know? What would happen if the hidden layers are smaller than the vocabulary size? We would also need to test several smaller dimensions.
-% DK: because the task is solved succesfully, TODO
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-% SD: Could we say that calculation on Dale-5 is successful?
-% DK: TODO
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-% SD: Not sure, since we never tested smaller layer sizes.
-If the embedding layers are the same as the vocabulary size then at least a language encoding LSTM is not motivated to geenralise at all as it can keep representations for individual words without abstracting them.
-Leanring word embeddings involves reducing dimensions.
-% DK: TODO
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-% SD: Hypothesis space to find the target function is just too big and they can never approximate that function.
-% DK: TODO
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (8 colors, 2 sizes, 3 shapes), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are $8*2*3=48$ possible combinations of attributes.
-% SD: Calculation behind this figure?
-% DK: (done)
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-% SD: We would need to show that the messages in Dale-2 are shorter than in Dale-5 to claim this.
-% DK: TODO
-
-\cmtDK[inline]{similarity to bounding box classifier}
-
-\subsection{Referring expression generation}
-\label{sec:referring-expression-generation-game}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK[inline]{human bias vs emergent language}
-
-For this experiment, the data is prepared in the same way as described in section \ref{sec:referring_expression_generation}.
-Since the position of the padding as well as the order of the words didn't have an effect on the results, the usual approach in natural generation task is used: padding tokens are appended and the referring expression is not reversed.
-Respectively, both agents are based on the models in these experiments.
-The sender uses the \emph{masked RE generator} with its LSTM replaced with the LSTM included in the \emph{EGG} framework, using the hidden size $h_s$.
-\cmtDK[inline]{fixed parameters}
-The receiver is built up similarly.
-The image is encoded in the same way, but afterwards it is concatenated with the decoded message, received be the sender with the hidden size $h_r$.
-The concatenated is reduced to $d=$ dimensions, based on the results of the previous experiments, and used as the initial state of the captioning LSTM.
-\cmtDK[inline]{fixed parameters}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
-    \label{fig:caption_generator_game_architecture}
-    % SD: But here we don't need an LSTM, we just want to identify one of the objects. The LSTM should only encode the inout message.
-    % DK: TODO
-\end{figure}
-
-During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
-% SD: I'm not sure how teacher forcing is implemented, see my earlier comment
-% DK: typo, teacher forcing is actually applied (done)
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
-The loss is calculated using cross entropy.
-The following values for the variables are compared:
-\begin{itemize}
-    \item $h_s$: 100, 500, 1000
-    \item $h_r$: 10, 30, 50, 100, 500, 1000
-    \item $|V|$: 1, 10, 16, 50, 100
-    \item $n$: 1, 2, 3, 4, 5, 6
-\end{itemize}
-\cmtDK[inline]{real values}
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-% SD: So the sender is generating descriptions following the GRE policy?
-% DK: TODO
-The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    % SD: What is word-by-word?
-    % DK: the same as in the caption generator with a single model (QUESTION)
-    \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions with a vocabulary size of 100, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-% SD: With the vocabulary size of 100.
-% DK: (done)
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-% SD: Not sure. The difference is still very small, within a couple of %.
-% DK: TODO
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-% SD: Very small decerase for such a large difference in vocabulary size compared to what we have seen earlier.
-% DK: TODO
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-% SD: Here it would be much better to have a loss curve to show differences in loss between the number of communication events, the same could also be done for accuracy.
-Then we would have a clearer picture of how vocabulary size affects learning.
-It could be that with 100 vocabulary there is actually a better performance than before the final cut-off point.
-Is this possible?
-How was the final cut-off point determined anyway?
-% DK: TODO
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-% SD: Negative result: how can we then interpret that? What could we change? What did we learn from this?
-% DK: TODO
-This is especially interesting, compared to single model caption generator in section \ref{sec:referring_expression_generation}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsection{Reference resolution}
-\subsubsection*{Setup}
-In the final experiments, the agents are tasked to produced and understand referring expressions only based on the implicit human knowledge present in the scenes as for example the attributes of the objects.
-Opposed to the previous experiments, explicit human language information as human referring expressions or one-hot encoded attributes are avoided.
-By this, the agents can't just reuse these human referring expressions, but need to learn to generate them themselves during the language games just based on the visual input.
-This is done similarly to the single model reference resolution task, described in section \ref{sec:reference_resolution}.
-The agents are tasked to predict the center coordinates of the target object, after being presented with the complete scene.
-However, the agents, more specifically the sender needs to be pointed towards the target object.
-In the final setup, the \emph{masked reference resolution game}, this is done providing the sender with an additional masked visual input.
-However, this can prove to be a challenging task for the agents.
-For this reason, in an approach to reduce the complexity, for the \emph{RE reference resolution game} the sender is shown a human referring expression, generated with the incremental GRE algorithm instead of the additional masked image.
-This might lead to a heavy bias of the emerged language towards the human referring expressions and will be compared to the unbiased emerged language.
-
-The masked sender is the same as the \emph{RE generation sender}, described in section \ref{sec:referring-expression-generation-game}.
-The RE sender works similarly, but instead encodes the human referring expression with an LSTM.
-The final hidden state is concatenated with the encoded image.
-The resulting vector is reduced to the hidden size $h_s$ and used as the initial hidden state of the LSTM included in the \emph{EGG} framework to generate the message.
-\cmtDK[inline]{fixed variables}
-
-The receiver is the same in both setups and is based on the \emph{reference resolver} + referring expressions, described in section \ref{sec:reference_resolution}.
-Again, instead of the included LSTM, the receiver uses the \emph{EGG} LSTM with the hidden size $h_r$ to encode the message.
-The dimensions of the coordinate predictor are fixed to $c=$, based on the previous results.
-\cmtDK[inline]{fixed variables}
-% SD: Just thinking: would there be a more systematic way to compare all the models?
-Perhaps if you write the parameters in a table or a text that is parallel.
-Then you could simply say, modal A + these changes.
-It would be much clearer to the reader how all models are rerlated and it would be easier for the reader to follow.
-% DK: TODO
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
-    \label{fig:coordinate_predictor_game_architecture}
-    % SD: Again, why LSTM, are we inceremtally identifying the image while we are processing each input word?
-    % DK: because it's a message sequence of discrete symbols (done)
-\end{figure}
-
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
-The loss is calculated using cross entropy.
-The following values for the variables are compared:
-\begin{itemize}
-    \item $h_s$: 100, 500, 1000
-    \item $h_r$: 10, 30, 50, 100, 500, 1000
-    \item $|V|$: 1, 10, 16, 50, 100
-    \item $n$: 1, 2, 3, 4, 5, 6
-\end{itemize}
-\cmtDK[inline]{real values}
-
-As in section \ref{sec:reference_resolution}, the agents are trained first on the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-The same metrics are used to evaluate the results.
-
-\subsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-% SD: Just occured to me: how do we measure accuracy if we are predicting numbers (numeric prediction)?
-% DK: same as in the experiments with one model (done)
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-% SD: But how could it do so if the visual data is randomly distribution. It must rely on the message as that is the only source of information.
-% DK: it does it in the same way as the single model above (TODO)
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-% SD: Remind the reader what is the size of the image so that they can interprert the distance relatively to that.
-% DK: TODO
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-% SD: I don't understand this
-% DK: TODO
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}    & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0,7}           & {45,9}         & {12,7\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}    & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}   & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-% SD: A bit gloomy result. Negative result is okay but then we need an explanation why this is so and have some ideas what to change. What we can learn from these results?
-% DK: TODO
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-% SD: This identifies that the problem lies in how the sender encodes the messages; what is its policy to generate longer strings or reuse the word.
-Without a policy the sender will never be motivated to encode longer messages and therefore rely on single-word expressions.
-% SD: The second problem is the size of these networks.
-We should also test smaller embedding sizes than 10.
-Practice shows that the embedding sizes can be even several magnitues smaller than the vocabulary sizes, cf.
-the Bengio paper and word2vec.
-% SD: The reason why the system is not performing well on the pointing task is that it does not have the right features to learn from.
-It ahs visual information WHAT these objects are (what do they look like) but they do not have information WHERE these objects are,.
-John Kelleher and I wrote an opiiuon piece about the lack of spatial knowledge required to model spatial descriptions in CNNs.
-The same problem is probably occuring here.
-Replacing the features or adding geometric features that would communicate geometric relations that allow pointing would improve the task and also we would expect that a vocabulary would emerge such that some words are more biased towards visual features (to identify objects) and some more to geometric features (to locate them), hence WHAT and WHERE.
-% SD: J.
-D.
-Kelleher and S.
-Dobnik.
-What is not where: the challenge of integrating spatial representations into deep learning architectures.
-In S.
-Dobnik and S.
-Lappin, editors, Proceedings of the Conference on Logic and Machine Learning in Natural Language (LaML 2017), Gothenburg, 12 –13 June, volume 1 of CLASP Papers in Computational Linguistics, pages 41–52, Gothenburg, Sweden, November 2017.
-Department of Philosophy, Linguistics and Theory of Science (FLOV), University of Gothenburg, CLASP, Centre for Language and Studies in Probability.https://gup.ub.gu.se/publication/262970?lang=en
-% SD: Overall, the thesis has a lot of potential if we could implement all this
-% DK: TODO
-
-When comparing these results to the neural models in Section \ref{sec:reference_resolution} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2023-11-27 thesis/sections/6.1-object-identification.tex
setup grounding

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
-The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
+\subsection{Object identification}
 \subsubsection*{Setup}
 
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-\end{figure}
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
-These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
+The object identification game is set up as a discrimination game.
 In a discrimination game, the agents are presented with two or more images, one of these being the target image.
@@
 
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
+The data is prepared in the same way as for the single model object identification task in section \ref{sec:object-identification}.
+Bounding boxes of all objects in the scene are extracted and preprocessed to be used for the feature extractors.
+Accordingly, also the models from the single model tasks serve as basis for the object identification task.
+The sender uses the architecture of the model described in section \ref{sec:referring_expression_generation} apart from the final LSTM.
+Instead, it will use the LSTM with the hidden size $h_s$ included in the \emph{EGG} framework to produce a message.
+Based on the results from the single model object identifiers, the image embedding size $e_s$ is fixed to 100.
+The bounding box of the target object is passed first, the remaining are shuffled.
 
-\begin{figure}[h]
+The receiver instead uses the architecture of the single model object identifier, described in section \ref{sec:object-identification}.
+Similarly to the sender, the LSTM of the \emph{EGG} framework with the hidden size $h_e$ is used to encode the sender's message instead of the one-hot encoded attributes in the single model.
+The hidden state of the LSTM is projected to the embedding size $e_r=500$ with a linear layer, based on the results in the previous experiments.
+As before the dot product between the embedded message and the encoded image is calculated to let the model point towards one of the bounding boxes.
+Opposed to the sender, the receiver is shown all bounding boxes in shuffled order.
+Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
+
+\begin{figure}[ht]
     \centering
@@
 
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $h_s$: 100, 500, 1000
+    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $|V|$: 1, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 5, 6
+\end{itemize}
+The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
 
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
+\subsection*{Results}
 Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
@@
 The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
+% SD: This discussion of dimensions could already be introduced when you discuss simple modles earlier as the same dimensions are used.
+This is a more an evaluation of represnetations rather than interaction and hence it would fit there.
+% DK: TODO
 
-\begin{table}[h]
+\begin{table}[ht]
     \centering
@@
 Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
+% SD: We would need to show that the messages in Dale-2 are shorter than in Dale-5 to claim this.
+% DK: TODO
 
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
-
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
-
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
-    \label{fig:caption_generator_game_architecture}
-\end{figure}
-
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
-    \label{fig:coordinate_predictor_game_architecture}
-\end{figure}
-
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
-
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
-        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.
+\cmtDK[inline]{similarity to bounding box classifier}

### 2023-11-27 thesis/sections/6.2-re-generation.tex
setup grounding

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
-
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
-
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-\cmtDK{ambiguity}
-
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
-The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
-\subsubsection*{Setup}
-
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-\end{figure}
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
-These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-    \caption{Sender and receiver architectures in the discrimination game}
-    \label{fig:discriminator_architecture}
-\end{figure}
-
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
-        \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_discriminator}
-\end{table}
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-
-\cmtDK{calculation of loss?}
-\cmtDK{length}
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are 48 possible combinations of attributes.
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
-
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
+\subsection{Referring expression generation}
+\label{sec:referring-expression-generation-game}
 \subsubsection*{Setup}
@@
 The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
+\cmtDK[inline]{human bias vs emergent language}
 
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
+For this experiment, the data is prepared in the same way as described in section \ref{sec:referring_expression_generation}.
+Since the position of the padding as well as the order of the words didn't have an effect on the results, the usual approach in natural generation task is used: padding tokens are appended and the referring expression is not reversed.
+Respectively, both agents are based on the models in these experiments.
+The sender uses the \emph{masked RE generator} with its LSTM replaced with the LSTM included in the \emph{EGG} framework, using the hidden size $h_s$.
+\cmtDK[inline]{fixed parameters}
+The receiver is built up similarly.
+The image is encoded in the same way, but afterwards it is concatenated with the decoded message, received be the sender with the hidden size $h_r$.
+The concatenated is reduced to $d=$ dimensions, based on the results of the previous experiments, and used as the initial state of the captioning LSTM.
+\cmtDK[inline]{fixed parameters}
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
     \label{fig:caption_generator_game_architecture}
+    % SD: But here we don't need an LSTM, we just want to identify one of the objects. The LSTM should only encode the inout message.
+    % DK: TODO
 \end{figure}
 
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
+During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
+% SD: I'm not sure how teacher forcing is implemented, see my earlier comment
+% DK: typo, teacher forcing is actually applied (done)
+When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
 
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $h_s$: 100, 500, 1000
+    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $|V|$: 1, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 5, 6
+\end{itemize}
+\cmtDK[inline]{real values}
 Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
 The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
+% SD: So the sender is generating descriptions following the GRE policy?
+% DK: TODO
+The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results.
 
 \subsubsection*{Results}
-\begin{table}[h]
+\begin{table}[ht]
     \centering
@@
 For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
+% SD: Here it would be much better to have a loss curve to show differences in loss between the number of communication events, the same could also be done for accuracy.
+Then we would have a clearer picture of how vocabulary size affects learning.
+It could be that with 100 vocabulary there is actually a better performance than before the final cut-off point.
+Is this possible?
+How was the final cut-off point determined anyway?
+% DK: TODO
 
 These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
+% SD: Negative result: how can we then interpret that? What could we change? What did we learn from this?
+% DK: TODO
+This is especially interesting, compared to single model caption generator in section \ref{sec:referring_expression_generation}.
 In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
 This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
-    \label{fig:coordinate_predictor_game_architecture}
-\end{figure}
-
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
-
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
-
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
-
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
-
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &          &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$  & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}     & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}     & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0
-        ,7}   & {45,9}  & {12,7\%} & {1}                                                                                                                                                                                                                                   \\
-        {13}  & {10}    & {10}     & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}     & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}     & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}    & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
-
-When comparing these results to the neural models in Section \ref{sec:coordinate_predictor} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2023-11-27 thesis/sections/6.3-reference-resolution.tex
setup grounding

--- 2023-08-21 thesis/sections/4-exp-setup.tex
@@
-\section{Experimental setup and results}
-\label{sec:exp-setup}
+\subsection{Reference resolution}
+\subsubsection*{Setup}
+In the final experiments, the agents are tasked to produced and understand referring expressions only based on the implicit human knowledge present in the scenes as for example the attributes of the objects.
+Opposed to the previous experiments, explicit human language information as human referring expressions or one-hot encoded attributes are avoided.
+By this, the agents can't just reuse these human referring expressions, but need to learn to generate them themselves during the language games just based on the visual input.
+This is done similarly to the single model reference resolution task, described in section \ref{sec:reference_resolution}.
+The agents are tasked to predict the center coordinates of the target object, after being presented with the complete scene.
+However, the agents, more specifically the sender needs to be pointed towards the target object.
+In the final setup, the \emph{masked reference resolution game}, this is done providing the sender with an additional masked visual input.
+However, this can prove to be a challenging task for the agents.
+For this reason, in an approach to reduce the complexity, for the \emph{RE reference resolution game} the sender is shown a human referring expression, generated with the incremental GRE algorithm instead of the additional masked image.
+This might lead to a heavy bias of the emerged language towards the human referring expressions and will be compared to the unbiased emerged language.
 
-In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
-Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
-This is done in three parts.
-Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
-In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
-Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.
+The masked sender is the same as the \emph{RE generation sender}, described in section \ref{sec:referring-expression-generation-game}.
+The RE sender works similarly, but instead encodes the human referring expression with an LSTM.
+The final hidden state is concatenated with the encoded image.
+The resulting vector is reduced to the hidden size $h_s$ and used as the initial hidden state of the LSTM included in the \emph{EGG} framework to generate the message.
+\cmtDK[inline]{fixed variables}
 
-\subsection{Creation of the dataset}
-\label{sec:creation-dataset}
-\cmtDK{ambiguity}
+The receiver is the same in both setups and is based on the \emph{reference resolver} + referring expressions, described in section \ref{sec:reference_resolution}.
+Again, instead of the included LSTM, the receiver uses the \emph{EGG} LSTM with the hidden size $h_r$ to encode the message.
+The dimensions of the coordinate predictor are fixed to $c=$, based on the previous results.
+\cmtDK[inline]{fixed variables}
+% SD: Just thinking: would there be a more systematic way to compare all the models?
+Perhaps if you write the parameters in a table or a text that is parallel.
+Then you could simply say, modal A + these changes.
+It would be much clearer to the reader how all models are rerlated and it would be easier for the reader to follow.
+% DK: TODO
+The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
 
-This research investigates how agents communicate about the relations of objects seen in images.
-For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
-Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
-To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
-The \emph{material} is always the same for all objects in a generated image.
-There were three main extensions to the code:
-
-First, objects in the scene were separated into three categories: one \emph{target object}, objects in a \emph{target group} and \emph{distractor} objects.
-The target object is the main object in the scene and the models are trained to identify and communicate between each other.
-All other objects and their relations are based on this target object.
-The target group contains similar objects to the target object.
-These are objects that the agents need to discriminate the target object from.
-Finally, the distractors are objects that add noise to the scene and should make it more complex. They are expected to teach the agents more precise descriptions of the target object.
-The number of the objects in both groups can be controlled.
-
-In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
-The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
-For example the target object is a \emph{small red cube}.
-If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
-The number of shared attributes can also be set to a range to make the discrimination task more challenging.
-
-Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
-For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
-This allows for a very controlled generation of relations between the objects in the scene.
-Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
-Here, the target object is the large purple cylinder.
-The target group contains four objects that share zero to a maximum of two attributes.
-It is not controlled, which attributes are shared (they are selected randomly).
-The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
-There are no distractor objects.
-
-For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
-The size of the generated images is 480x320 pixels.
-10.000 images are created for each of the datasets.
-Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['CLEVR single', \emph{large yellow sphere}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_single.png}
-        \label{fig:clevr-single}
-    }
-    \subfigure['CLEVR color', \emph{small brown cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_color.png}
-        \label{fig:clevr-color}
-    }
-    \subfigure['Dale-2', \emph{small green cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-2.png}
-        \label{fig:clevr-dale-2}
-    }
-    \subfigure['Dale-5', \emph{large purple cylinder}]{
-        \includegraphics[width=0.43\linewidth]{figures/CLEVR_dale-5.png}
-        \label{fig:clevr-dale-5}
-    }
-    \caption{Example images of each dataset, with the target object specified}
-    \label{fig:clevr-examples}
-\end{figure}
-
-\subsubsection*{CLEVR single}
-The simplest new dataset is called 'CLEVR single'.
-This is a very simple dataset and has the purpose to simplify the problem the model needs to learn as much as possible.
-Each scene in the dataset contains only one single object, the target object.
-There are neither objects in the target nor in the distractor group.
-All attributes are assigned randomly to the target object.
-The differences across the whole dataset are the locations and rotations of the objects.
-With this dataset, neural models can focus on only the features, as well as the locations of this single object.
-There are no objects that distract the model from extracting features from the target object.
-This helps to understand if the models are actually able to assign features or learn locations of these features in an image.
-Figure \ref{fig:clevr-single} shows an example with the only object being the \emph{large yellow sphere}.
-
-\subsubsection*{CLEVR color}
-The second dataset that is created is called 'CLEVR color'.
-The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
-For this reason, there exist only two groups in the scene, the target object and distractors.
-The distractor group can contain in between 6 and 9 objects.
-To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
-Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
-The distractors have always the same size and shape as the target object, but the color is different.
-The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
-
-As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
-By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
-All attributes, apart from the color are not discriminating the target object from the distractors.
-Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
-
-\subsubsection*{CLEVR Dale datasets}
-The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
-The number and the type of shared attributes are controlled exactly.
-More realistically would be if objects can share a variable number of attributes.
-In real situations, there is no restriction at all how objects or things relate to each other.
-Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
-This emergence of referring attributes and their combination is studied deeper in this work.
-
-For this, I created a dataset that allows almost any relation between a target object and the distractors.
-The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
-This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
-Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
-This ensures the uniqueness of the target object.
-On the other side, it is not controlled which attributes are shared.
-These are assigned randomly.
-There is again no control over the relations between distractors, which means that distractors can appear multiple times.
-
-Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
-Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
-The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
-
-These two datasets allow a more realistic look in how models can acquire knowledge about attributes of objects.
-More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}
-
-\subsection{Generating and understanding referring expressions}
-\label{sec:preexperiments}
-This chapter serves two purposes.
-First, the generated dataset from the precious section is validated.
-For this, models are trained to both generate referring expressions of the target object and understand existing referring expressions.
-A success of these experiments indicates that the target objects in the datasets are possible to refer to and the datasets can be used in more complex setups in language games.
-
-Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
-Language games are very complex setups for machine learning models.
-The models need to solve multiple tasks at the same time in order to solve the overall problem.
-For instance, in a simple setup of a game two agents are involved.
-The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
-The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
-In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
-In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
-The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
-And finally it needs to learn how to identify the target object with this information.
-There are many points of possible failure to train the agents.
-
-For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
-This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
-Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
-With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.
-
-The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
-Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
-Here, I structured the experiments into three levels.
-In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
-In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
-In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
-These combined experiments should lay the basis for how to build up the agents in the language games.
-
-\subsubsection{Coordinate predictor}
-\label{sec:coordinate_predictor}
-\subsubsection*{Setup}
-
-This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
-As described before, the sender should communicate an object in the image and the receiver needs to identify it.
-The challenge lies in how the receiver refers to the identified object.
-There are multiple possibilities, how it can be done.
-One of them could be to describe the target object with human language, using the attributes.
-The main goal however is to let the language of the agents emerge as natural as possible.
-Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
-For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
-With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.
-
-To achieve this goal, multiple setups of models are tested.
-In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
-Here the image is first passed through one of the \emph{feature extractors}.
-Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
-These reduce the dimensions first to 1024 and finally to 2.
-
-To determine the loss, the euclidean distance between the resulting predicted point on the image and the ground truth point are calculated.
-This distance is learned to be minimized.
-By doing that, the model learns to focus and attend on a specific part in the image, in a perfect model the center of the target objects.
-
-With this simple setup, the model is theoretically able to focus on an object in the image.
-The problem arises as soon as multiple objects are present in the image.
-There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
-Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
-Therefore, the models need to receive more information.
-Here, I try out four different ways to encode and refer to the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[Model including one-hot encoded attributes and locations]{
-        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
-        \label{fig:coordinate_predictor_architecture}
-    }
-    \subfigure[Model including GRE description]{
-        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
-        \label{fig:coordinate_predictor_dale_architecture}
-    }
-    \caption{Simplified architecture of the coordinate predictors}
-\end{figure}
-
-In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
-There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
-The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
-These three encodings of the attributes are then concatenated.
-The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
-The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.
-
-In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
-This should help the model to identify all possible options to chose from, when predicting the target object.
-All the center coordinates are simply extracted and shuffled.
-Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
-The padded locations consist of two zeros for both coordinates.
-For this model, I also made use of a more complex way to encode the image.
-Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
-The image is first passed through the feature extractor.
-Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
-After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
-The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
-This is then passed again through a final linear layer to predict x- and y-coordinates.
-
-The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
-This opposes the idea described before, to share as few human knowledge as possible with the model.
-Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
-If the model is able to match parts of the image with human words it would show that the model learned this attribute.
-If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
-This insight would help for succeeding models that make use of these learnings without human language.
-
-Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
-In other words, the object is described unambiguously using the lowest number of words.
-The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
-This order defines, which attributes can be left out, while still identifying the object uniquely.
-This research relies on the following order from most important to least important: shape, color, size
-Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
-Using all three attributes, this description identifies the object perfectly and uniquely.
-Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
-This can be taken even one step further by removing also the \emph{color}.
-Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
-
-For the experiments, each image is captioned with a description of the target object using the described algorithm.
-To include it in the model, the captions need to be padded to an equal length.
-In this case they are padded to a length of 3, which is the maximum number of attributes that can be used.
-For this, as standard practice in captioning tasks, the captions are padded at the end with a specified padding token.
-
-The model is made up of three parts:
-The first part extracts features from the image.
-Here, the setup is similar to the previous model.
-The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
-A \emph{ReLU} function is applied, after both convolutional layers.
-As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
-In the second part, the caption is encoded, using an LSTM.
-Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
-The third part is again the predictor of the coordinates of the target object.
-Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
-The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
-Between each linear layer, the \emph{ReLU} function is applied.
-This architecture is also inspired by the baselines described in \citet{Johnson2017}.
-
-The forth method, to encode the target object utilizes \textbf{masking} of the image.
-For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
-The side of the square is always two fifth of the image width, in this case 192 pixels.
-This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
-The square is filled in white, while the rest of the image is filled in black.
-This approach has the advantage of providing as few as possible human bias to the model.
-While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
-It therefore can only rely on its own extracted visual features when looking at masked images.
-The model is presented with both the original image and the masked image.
-Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
-The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.
-
-The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
-This distance needs to be minimized.
-The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
-Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
-If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
-It doesn't reflect this change.
-For that reason, I also introduced an accuracy score.
-For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
-If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
-These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
-This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
-A high accuracy may indicate that the model could identify this specific object better.
-
-The coordinate predictors are trained on the 'CLEVR single' as well as on both 'Dale' datasets.
-The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information about the image could have gone lost.
-Training on this dataset should make sure that the model can converge towards the correct pixels, utilizing these features.
-In a next step, the 'Dale' datasets provide the actual problem of discriminating objects from each other and afterwards pointing to the correct one.
-Here, the models should make use of the additional given referring expressions about the scene, as one-hot encodings of the attributes, descriptions using the GRE-algorithm or the encoded locations.
-'Dale-2' and 'Dale-5' provide two different difficulties for the model, where it needs to discriminate a target object from one or four distractors.
-Latter task is assumed to be significantly harder.
-
-\subsubsection*{Results}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
-    \label{fig:coordinate-predictor_loss}
-\end{figure}
-
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{\emph{GRE-algorithm}} from \citet{Dale1995}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
-        \label{fig:visualizations_dale-2_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
-        \label{fig:visualizations_dale-2_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
-        \label{fig:visualizations_dale-5_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
-        \label{fig:visualizations_dale-5_test}
-    }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
-\end{figure}
-
-An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, I visualized the predicted coordinates compared to the ground truth coordinates.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
-The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
-As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
-Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
-As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
-
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
-These findings align with the mean distance scores, described in the sections before.
-However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
-Again, the green circles form the shape of rhombus.
-In contrast, the predictions in red almost all cluster in the center of the image.
-They form roughly the shape of a smaller rhombus.
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
-
-\begin{figure}[h]
-    \centering
-    \subfigure['Dale-2', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_train_complete.png}
-        \label{fig:visualizations_dale-2_complete_train}
-    }
-    \subfigure['Dale-2', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-2_test_complete.png}
-        \label{fig:visualizations_dale-2_complete_test}
-    }
-    \subfigure['Dale-5', train split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_train_complete.png}
-        \label{fig:visualizations_dale-5_complete_train}
-    }
-    \subfigure['Dale-5', test split]{
-        \includegraphics[width=0.42\linewidth]{figures/visualization_dale-5_test_complete.png}
-        \label{fig:visualizations_dale-5_complete_test}
-    }
-    \caption{Visualization of the models' predictions in the Dale datasets}
-    \label{fig:visualizations_dale_complete}
-\end{figure}
-
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
-The reason for this is likely that the model can produce a relatively low loss, without relying on many extracted features of the objects.
-Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
-
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
-Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
-
-\subsubsection{Bounding box classifier}
-\subsubsection*{Setup}
-\cmtDK{better to call it discriminators?}
-
-One problem that arises, when only predicting the coordinates is that the models may not be able to predict locations of objects from just visual features.
-Feature extractors like VGG or ResNet are trained on extracting visual information.
-In this process, absolute locations of these features in the images may get lost in the resulting vectors, since they are not important for this task.
-To address this challenge, I restructured the task in this experiment.
-Instead of predicting two coordinates of a target object, the model now needs to classify between all available objects.
-This will take all positional information out and just focus on the attributes and discriminating factors between the objects.
-To do this, fixed-size bounding boxes are extracted around each of the object in the image.
-As for the masking in the experiment before, the bounding boxes are a square of 192x192 pixels, to capture the complete object. \cmtDK{preprocess bounding box}
-The bounding boxes of all objects are padded to the maximum number of object present in an image across the whole dataset with a matrix of zeros.
-These bounding boxes are shuffled.
-Furthermore, the attributes of the target object are encoded as on-hot vectors to point the model towards it.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_bounding_box_classifier.png}
-    \caption{Simplified architecture of the bounding box classifier}
-    \label{fig:bounding_box_classifier_architecture}
-\end{figure}
-
-The model passes each of the bounding boxes through a linear layer to compress it to an embedding vector of 10 dimensions.
-The same is done with the one-hot vector of the target attributes.
-In the next step, the dot product between each embedded image and the embedded attributes are calculated.
-A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
-The model points then to an object, using the \emph{softmax} function over all concatenated dot products.
-The loss is calculated using cross entropy.
-
-Furthermore, the model is compared to a baseline that doesn't include any information about the target object.
-Here, the model passes all flattened bounding box representations through a linear layer resulting in a vector with as many objects as are present in the image.
-Again, the model uses a \emph{softmax} function to point to one of these objects.
-Since no information about the target object is included, the guesses are expected to be random.
-
-The bounding box classifier is trained on all datasets excluding the 'CLEVR single'.
-The 'Dale' datasets are directly comparable to each other for the similar setup of their creation.
-Especially interesting is the effect of the increasing the number of distractors and the growing number of attributes that are needed to discriminate the objects.
-
-\subsubsection*{Results}
-As expected, all baselines achieve a random accuracy for each of the datasets.
-The accuracy lies at 50\% for the 'Dale-2' dataset, 22\% for the 'Dale-5' dataset and 12\% for the 'CLEVR color' dataset.
-This changes however, when information about the target object is included.
-Table \ref{tab:results_bounding_box_classifier} lists the accuracies of the models' predictions after 10 epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rcc}
-        \toprule
-                             & \textbf{Accuracy baseline} & \textbf{Accuracy} \\
-        \textbf{Dale-2}      & 50\%                       & 99\%              \\
-        \textbf{Dale-5}      & 22\%                       & 94\%              \\
-        \textbf{CLEVR color} & 12\%                       & 93\%              \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the bounding box classifier after 10 epochs}
-    \label{tab:results_bounding_box_classifier}
-\end{table}
-
-The model achieves almost perfect accuracy on the 'Dale-2' dataset with 99\%.
-For both other datasets, the accuracy is slightly lower with 94\% and 93\% respectively, but still close to perfect, especially compared to the random baseline.
-First, this shows that the model is able to extract attribute features from the image and manages to combine them with the given information about the target object.
-This worsens slightly, when more objects are introduced.
-More objects lead to more needed attributes to discriminate the objects from each other.
-The model is still able to extract and utilize these additional attributes and combine them with the given attribute encoding of the target object.
-The model can generate these high results, even though its architecture is very simple.
-In this architecture, the model doesn't compare the objects directly to each other, but each object is only associated with the attribute encodings.
-A more complex architecture, in which the model is additionally tasked to discriminate the objects directly from each other might even improve the results.
-
-\subsubsection{Caption generator}
-\label{sec:caption_generator}
-\subsubsection*{Setup}
-
-As the bounding box classifier, the caption generator only acts as a learning step towards understanding, how the model can learn the attributes of objects.
-It is used the same way to create the captions for each image as in the section above with the GRE-algorithm.
-This method focuses even more on human knowledge and structure of language than the experiment above, since it is even the final task of the model.
-There are some minor additions concerning the padding of the caption.
-As before, the caption is padded to a number of three tokens, corresponding to the maximum of three attributes.
-
-However, there are three different ways how the padding is applied.
-First, the captions are, as usual in captioning tasks padded at the end with a specified padding token.
-A problem could arise when the caption is not viewed as a natural language sentence, but as slots filled with tokens.
-More specifically, following the GRE-algorithm, the last token in the caption is always the shape.
-The second last token if existing describes the color, while the third last token if existing describes the size.
-As soon as this sequence is padded at the end, these slots suddenly disappear.
-A caption that only describes the shape, such as \emph{cube} will be padded to \emph{cube <pad> <pad>}, where the third last slot is filled up with the shape instead of the size.
-Since in this task we are not focussing on producing natural language with a correct grammar, but focus instead of extracting attributes, having a slot structure could help the model to express the extracted attributes correctly.
-For this reason, the second method of padding the caption is prepending the caption with padding tokens.
-By this, the positions of the slots are preserved and if not specified just filled with a padding token.
-The last variation concerns the order of producing each token.
-When the captions are prepended, the model would need first produce two padding tokens, before it finally can produce a much more meaningful token for the shape.
-This could be difficult to learn for a model, as the longer a sequence of tokens is, the more information about the beginning of the sequence gets lost.
-Even though a sequence of just three tokens may not be long enough for this factor to be a problem, I experimented to reverse the caption.
-Instead of producing for instance \emph{<pad> green sphere} as correct in English, the model would now need to produce \emph{sphere green <pad>}.
-Notice that the padding token is again at the end of the generation, but the order of slots as well as the amount of information in the caption are still preserved.
-
-Nonetheless, this helps to understand more detailed if and how the model discriminates objects.
-Does it rely on the same attributes, humans are using, or does it find other important differences, or is it able to solve the task at all?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.5\linewidth]{figures/arch_caption_generator.png}
-    \caption{Simplified architecture of the masked caption generator}
-    \label{fig:caption_generator_architecture}
-\end{figure}
-
-I compare two different models against each other.
-The first model just takes the image as the input.
-During training, also the ground truth caption is passed to the model.
-The image is processed as in the section before.
-First, it is passed through a feature extractor with two following 2d convolutional layers, reducing the channels to 128.
-After each convolutional layer the \emph{ReLU} function is applied.
-The result is pooled using max pooling and then reduced to 1024 dimension using a linear layer.
-The such encoded image serves as the initial hidden state of an LSTM, which generates the caption.
-During training the input sequence for the LSTM at each time step are embeddings of the tokens in the ground truth caption.
-No teacher forcing is applied.
-The output of the LSTM is passed through a linear layer at each step to determine logits over the symbols of the vocabulary.
-The loss is calculated using cross entropy.
-When testing, the LSTM always generates three tokens, with a start-of-sequence token as first input to the LSTM.
-Each token in the sequence is determined, by selecting the highest logit in the output of each step in the LSTM.
-
-Since with this approach, the model doesn't have any information about the target object, I extend it with a masked image.
-As in the section above, a masked version of the image is created and passed to the model.
-This is shown in Figure \ref{fig:caption_generator_architecture}.
-The model works exactly the same, except for the additional separate image encoder that encodes the masked image in the same way as the original image.
-Both encodings are concatenated and then used as the initial hidden state for the LSTM.
-This should point the model to which object to describe and discriminate from the other objects.
-
-To test the success of the model, three measures are calculated.
-The first measure is the \textbf{accuracy} if the model predicted every word in the caption correctly.
-This gives a hint, how the model fares in general and also able to predict actually any of the attributes at all.
-However, a 'false' prediction doesn't give much insight into why the model predicted a wrong caption.
-
-It could be the case that the model predicted the correct shape, but wrong color.
-Even worse, the model could have predicted more attributes than necessarily to uniquely identify the target object and didn't follow the rules of the GRE-algorithm.
-For instance, consider the scene in Figure \ref{fig:clevr-dale-5}.
-The correct caption is \emph{cylinder}.
-If the model would predict \emph{purple cylinder}, the accuracy determines it as false as captioning \emph{large purple cylinder} as well \emph{small green cube}.
-The first two descriptions identify the target object perfectly, but the model only didn't learn yet to leave unnecessary attributes out.
-To mitigate this, I also include a \textbf{word-by-word accuracy}.
-This measure calculates the accuracy not on sentence level, but on word level.
-The first predicted caption of the model would yield a word-by-word accuracy of 66\% (including padding tokens), the second 33\%, while the third prediction would
-yield 0\%.
-This can give a better understanding of the errors the model makes.
-
-With the \textbf{non-target accuracy}, I identify if the model described another object, which is not the target image.
-This is basically an inverted accuracy score; the lower the score, the better the model fares.
-For this I generated captions for all the non-target objects and distractors in the images using the GRE-algorithm.
-If the generated description of the model describes an object that is not the target object, it gets assigned 100\%.
-If not, independently of describing the target object, no object, or one of the objects insufficiently, it gets assigned 0\%.
-Using this measure, I can get a quick overview if the model's problem lies in extracting and relating attributes or in understanding which of the presented objects is the target object.
-
-The caption generator models are trained on both 'Dale' datasets.
-Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
-Furthermore, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-
-\cmtDK{align 'description', 'caption' and 'referring expression' over the whole section}
-
-\subsubsection*{Results}
-In the next paragraphs, the results of the caption generator will be discussed.
-Table \ref{tab:results_caption-generator} shows the different scores of the models, when trained on 'Dale-2' and 'Dale-5'.
-ResNet-3 produced the best results, when it was used as a feature extractor.
-The results are shown after 50 epochs, but the scores already start to converge after 20 epochs and changed only very in the following epochs.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{rccc}
-        \toprule
-                               & \textbf{Accuracy} & \textbf{Word-by-word accuracy} & \textbf{Non-target accuracy} \\
-        \textbf{Dale-2}        & 49\%              & 82\%                           & 48\%                         \\
-        \textbf{Dale-2 masked} & 72\%              & 90\%                           & 25\%                         \\
-        \textbf{Dale-5}        & 21\%              & 45\%                           & 40\%                         \\
-        \textbf{Dale-5 masked} & 21\%              & 54\%                           & 45\%                         \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator, based on ResNet-3}
-    \label{tab:results_caption-generator}
-\end{table}
-
-The sum of the \emph{accuracy} and the \emph{non-target-accuracy} adds to 97\%, when trained on the 'Dale-2' dataset.
-This means that the models produce almost always correct descriptions of an object in the image.
-Moreover, these descriptions are also efficient in the way that they follow the GRE-algorithm of \citet{Dale1995} and only use necessary discriminative attributes.
-Features of both objects are therefore extracted perfectly and related to the vocabulary.
-As expected, the model has difficulties to decide, which of both objects is the target object, since no information is passed to the model.
-With 48\% of the images describing the target object and 49\% of the images describing the distractor, the model uses a random guess.
-This changes, when also the masked image is presented to the model.
-Then, the accuracy for the target object is with 72\% well above chance.
-Still, the wrong descriptions describe almost every time the distractor object.
-This indicates that the problem doesn't lie in extracting features of objects, but only in deciding which object in the image to describe.
-
-When the model is trained on the 'Dale-5' dataset, the results are much worse.
-Now, without masking only in 61\% of the images, the model describes any of the objects in the image.
-With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
-Interestingly, passing the masked image to the model doesn't help it to identify the correct object.
-The accuracy stays at the same value.
-In contrast, the non-target accuracy is increased by 5\% points.
-The model is more likely to identify a wrong object.
-Furthermore, the word-by-word accuracy increases from 45\% to 54\%.
-This increase is likely due to the fact that the description of the target object often shares some attributes with distractor objects.
-For instance an image with a target object being a \emph{small red cube} was identified as a \emph{<pad> <pad> cube} when the model didn't receive the masked image.
-After also the masked image was passed to the model, it generated the description \emph{small blue cube}.
-This was the correct description of a distractor and because of the overlap of the attributes, the word-by-word accuracy increased from 33\% to 66\%.
-
-The approach, how the padding is produced and in which order the attributes are concatenated didn't have an effect on the described metrics.
-When the order was reversed and the padding appended, it seemed like the model was converging slightly faster and reaching the limit around tow to three epochs earlier.
-The final peak stayed exactly the same and the effect was therefore not studied deeper if there is an actual significant difference.
-
-There are two main possible explanations for the big difference between the two datasets.
-First, as already seen in the experiments before, the bigger number of distractors confuses the model more, where to focus on.
-In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset give the model a bigger choice.
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, the probability that only the shape is enough to discriminate the objects is at 66,6\%, namely the caption is only one word long.
-For shape and color, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
-With four distractors the algorithm is much more likely to produce longer captions.
-These are harder to generate for the model, since it needs to take more attributes into account to discriminate the target object from the distractors.
-
-\subsection{Language games}
-\label{sec:language-games}
-
-The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
-
-The language games in this research have an asymmetric setup.
-One agent, the sender is shown some information and needs to generate a message.
-This message is received by the second agent, the receiver.
-The receiver needs to parse this message and combine it with the information they are given as well.
-The receiver then makes a prediction, which is compared to the ground truth.
-The game is set up prosocial, which means that both agents receive the same loss based on the receiver's prediction.
-All weights of the agents are adapted in the same way.
-\cmtDK{vocabulary?}
-
-\cmtDK{always ResNet-3}
-
-First, I will discuss \emph{discrimination games}, because they have the simplest setup.
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, I will look into \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-
-\subsubsection{Discrimination games}
-\subsubsection*{Setup}
-
-In a discrimination game, the agents are presented with two or more images, one of these being the target image.
-The sender needs to communicate this target image to the receiver by discriminating it from the other distractor images.
-The receiver then needs to decide based on the message, which of the images is the target image.
-
-The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
-The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
-One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
-It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
-
-Both 'Dale' datasets are used to train the agents in this mode.
-Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
-As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
-The sender is assumed to learn which of the input images is the target images by this approach.
-The receiver receives the images in completely shuffled order.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/arch_discriminator.png}
-    \caption{Sender and receiver architectures in the discrimination game}
-    \label{fig:discriminator_architecture}
-\end{figure}
-
-Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
-For the sender, the images are passed through a feature extractor and a following linear layer that reduces the dimensions to an embedding size $e_s$.
-All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
-This is then used as the initial state of the encoder LSTM in the sender wrapper.
-
-The receiver also encodes all images using a feature extractor with a following linear layer, reducing it to $e_r$.
-The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
-As mentioned above, the receiver combines this image representation with the hidden state of each symbol separately.
-This is done by passing the hidden state through a linear layer to scale its dimensions to the same $e_r$.
-This allows the calculation of the dot product between it and the image representation.
-If the message describes an object well, the resulting dot product should be higher.
-The receiver then `points' to one of the images by applying the \emph{softmax} function over the results of the dot products.
-The loss is calculated using the NLL-loss.
-
-During the experiments, five variables are adjusted to compare their effects:
-(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender wrapper $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver wrapper $h_r$ and (5) the size of the vocabulary $|V|$.
-
-\subsubsection*{Results}
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
-        \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_discriminator}
-\end{table}
-
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-
-\cmtDK{calculation of loss?}
-\cmtDK{length}
-
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (color, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are 48 possible combinations of attributes.
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-
-\cmtDK{similarity to bounding box classifier}
-
-\subsubsection*{Analysis of the emerged languages}
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
-
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
-
-The first property is that when a message is transferred, it consists of only one symbol.
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-
-\begin{figure}[h]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
-
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
-
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
-
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
-
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
-
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
-
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
-
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference key numbers are calculated for each of the attribute importance order.
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, I calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\subsubsection{Caption generator games}
-\subsubsection*{Setup}
-
-In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK{human bias vs emergent language}
-
-This setup is based on the single neural model described above.
-The target caption for each image is created using the GRE-algorithm of \citet{Dale1995}.
-Since the results of the experiments show that the position of the padding doesn't have an effect on the final converging scores, the padding tokens are only prepended to the caption.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
-    \label{fig:caption_generator_game_architecture}
-\end{figure}
-
-The sender is built up of two image encoders, one encodes the original target image and one encodes the masked image.
-The image encoders are set up as described in \citet{Johnson2017}.
-A feature extractor extracts basic features of the images.
-Two following convolutional layers with a \emph{ReLU} function are trained to condense the most important information from the resulting matrices.
-A max pooling layer and a linear layer reduce the dimensions to 1024 dimensions.
-This size was found to have the best results during the experiments.
-The two resulting encoded images are flattened and concatenated.
-A final linear layer reduces this long vector to the hidden size of the message LSTM $h_s$, which is used as the initial hidden state of this LSTM.
-
-The receiver uses the same architecture to encode the original image as the sender.
-The resulting flattened vector is concatenated with the decoded message of hidden size $h_r$.
-This is then passed through a linear layer, again reducing it to 1024 dimensions and is then used as the initial state of the captioning LSTM.
-During training, the ground truth caption is used as the input to the LSTM without teacher forcing.
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
-The loss is calculated with cross entropy.
-
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
-
-During the experiments, the following variables are adjusted, and the results are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$ and
-(3) the LSTM hidden size of the receiver wrapper $h_r$.
-The same metrics as in the pre-experiments are used to evaluate the results.
-
-\subsubsection*{Results}
-\begin{table}[h]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_caption_generator_game}
-\end{table}
-
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-This is especially interesting, compared to single model caption generator in section \ref{sec:caption_generator}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
-
-\subsubsection{Coordinate predictor games}
-\subsubsection*{Setup}
-In the final experiments, it is tried to eliminate as much human knowledge as possible.
-For that reason, the agents are tasked to only predict the location of the target object.
-This task is approached in two steps.
-In the first step the sender is still shown human knowledge in form of a description of the target object, since the previous experiments showed that the models were able to relate them to their own extracted features.
-The receiver doesn't come in contact with any human knowledge, not as input nor as output.
-This approach acts as a sanity check if the agents are able to converge together on the correct target object coordinates.
-In the second step, the caption is also removed from the sender.
-Instead, a masked image points the sender towards the target object.
-In this setup, no human knowledge is explicitly present, that can bias the emerged language to form similarly to human language, except for the implicit information in the image itself.
-With this, it can be analyzed, how the language between the agents emerges and which features or patterns are represented with symbols.
-
-The agents are both set up in the same way as the single neural model, predicting the target object's coordinates.
-The sender encodes the original image the same as in the previous paragraph.
-In the first setup, the description of the target object is encoded, using an LSTM.
-For this, an embedding with $emb_{descr}$ dimensions is learned to represent each word.
-These embeddings are the input for the LSTM.
-The final hidden state of $LSTM_{descr}$ dimensions is used as the representation of the whole description.
-The vocabulary that is used for the descriptions is based on 14 symbols, including the padding token.
-For the LSTM to learn a representation of each token as well as of the complete description, both $emb_{descr}$ and $LSTM_{descr}$ need to be smaller than the size of this vocabulary.
-Choosing a size of 10 for both variables proved to give good results in the experiments.
-In a next step, the image encoding and the final hidden state of the description are concatenated and passed through a linear layer to reduce the dimension to the hidden size $h_s$.
-The resulting vector is passed to the sender wrapper LSTM, to generate the message.
-
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
     \label{fig:coordinate_predictor_game_architecture}
+    % SD: Again, why LSTM, are we inceremtally identifying the image while we are processing each input word?
+    % DK: because it's a message sequence of discrete symbols (done)
 \end{figure}
 
-For the second approach, the masked image is passed only through a feature extractor.
-Both the resulting encoding of the masked image and the encoding of the original image are each passed through separate linear layer to adjust their dimensions to the same embedding size $e_s$.
-These embeddings are subsequently concatenated and reduced to the hidden size $h_s$ with a final linear layer.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
+The loss is calculated using cross entropy.
+The following values for the variables are compared:
+\begin{itemize}
+    \item $h_s$: 100, 500, 1000
+    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $|V|$: 1, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 5, 6
+\end{itemize}
+\cmtDK[inline]{real values}
 
-The receiver contains of two parts.
-First, the original image is encoded with an image encode of the same setup.
-This encoded image is flattened and concatenated with the final hidden state of the wrapper LSTM encoding the message received by the receiver.
-The resulting vector is the passed through the second part of the receiver, the predictor.
-This predictor contains of three linear layers, reducing the dimensions to 1024, 1024 and 2 respectively.
-A \emph{ReLU} function is applied in between.
-This setup is trained to combine the important information from both the image and the message of the sender.
-The euclidean distance between the resulting prediction of the center point and the true center of the target object is calculated and the weights of both agents are adapted accordingly.
+As in section \ref{sec:reference_resolution}, the agents are trained first on the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
+In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
+The same metrics are used to evaluate the results.
 
-As in the pre-experiments, the agents are trained first the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-
-During the experiments, the effects of the following variables are compared:
-(1) the vocabulary size $|V|$,
-(2) the LSTM hidden size of the sender wrapper $h_s$,
-(3) the LSTM hidden size of the receiver wrapper $h_r$ and
-(4) the embedding size of the sender $e_s$.
-As before, the metrics to evaluate the results are the same as in the pre-experiments.
-
-\subsubsection*{Results}
+\subsection*{Results}
 In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
@@
 Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
+% SD: But how could it do so if the visual data is randomly distribution. It must rely on the message as that is the only source of information.
+% DK: it does it in the same way as the single model above (TODO)
 As in the experiment before, all communicated messages contain the same symbol independent of the input image.
 
-\begin{table}[h]
+\begin{table}[ht]
     \centering
@@
 The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
+% SD: I don't understand this
+% DK: TODO
 
-\begin{table}[h]
+\begin{table}[ht]
     \centering

### 2023-11-28 thesis/sections/3-methodology.tex
results

@@
 More recently, convolutional neural networks (CNN) became popular due to their ability to learn complex features automatically from raw image data.
-These are also used in this thesis as a first layer to extract important information from the image. Hereby, two different architectures are tested.
+These are also used in this thesis as a first layer to extract important information from the image. Hereby, there are currently two different architectures.
 
-First, we use the VGG19 \citep{Simonyan2015} which is an architecture based on many convolutional layers.
+First, the VGG19 \citep{Simonyan2015} is used which is an architecture based on many convolutional layers.
 Using 16 to 19 convolutional layers with small convolution filters helps the model to solve localization and classification tasks on the training dataset, but also enables it to generalize onto other datasets.
@@
 
-Secondly, we include the ResNet-101 \citep{He2016}.
+The second architecture is ResNet-101 \citep{He2016}.
 This architecture tries to overcome the degradation of very deep networks, where the accuracy rapidly drops after it gets saturated.
@@
 % DK: (added; done)
-For this reason, multiple different adaptions of these architectures are compared.
-Table \ref{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101 that will be used in this research.
-In the later chapters, it will be referred to these adaptions using the name in the table.
+For this reason, multiple different adaptions of these architectures are usually used.
+Table \ref{tab:feature-extractor-archs} lists the different adaptions for both VGG19 and ResNet-101, where later layers are omitted.
+Preliminary results showed that early layers like \emph{ResNet-1} to \emph{ResNet-3} or \emph{VGG-0} and \emph{VGG-avg} provided better results.
+This corresponds to the hypothesis that later layers include too much task specific knowledge from pretraining.
+There was no visible difference between \emph{VGG} and \emph{ResNet}, so \emph{ResNet} is used in all experiments due to its more efficient architecture.
+However, \emph{ResNet-1} and \emph{ResNet-2} contain a high number of dimensions, that slows the training and raises the amount of needed memory on the GPU.
+Some experiments couldn't be performed using these layers, so all experiments are run, using \emph{ResNet-3}.
 

### 2023-11-28 thesis/sections/5.3-reference-resolution.tex
results

@@
     }
-    \caption{Simplified architecture of the coordinate predictors}
+    \caption{Simplified architectures of the reference resolvers}
     % SD: How are these encoded?
@@
 
-For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
+For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $1\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
 8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
@@
                               & $e_i$                       & $c$               & $LSTM_e$  & $LSTM_o$      \\
-                              & $[100, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[500,1000,1500,2000]$ \\\midrule
+                              & $[50, 100, 300, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[1000,1500]$ \\\midrule
         reference resolver    & \times                      & \times            & -         & -             \\
@@
 The 'CLEVR single' dataset should test the model if it can actually learn locations of an object.
-Since the model relies on the extracted features of either VGG or ResNet, locational information of the objects in the image might not be present anymore.
+Since the model relies on the extracted features of ResNet, locational information of the objects in the image might not be present anymore.
 This is due the fact that CNNs compress information of the image summarizing small regions (convolutions) several times.
@@
 \subsubsection*{Results}
+Tables \ref{tab:results:reference-resolver} to \ref{tab:results:reference-resolver-masked} show the results of the five adaptions of the reference resolver.
+Across the first three models (basic reference resolver, +one-hot, +one-hot+locations), the same patterns can be seen.
+
+The \textbf{basic reference resolver} acts as a baseline for the other models.
+Since this model doesn't have any information about the target object, it shouldn't be able to point towards the correct target object.
+However, this model should learn underlying structures in the images that it can use to make better predictions.
+Indeed, the model converges on all datasets to a certain limit.
+Random guesses of the coordinates correspond to a mean distance of 95 pixels, given the image size of 224\times224 pixels, which the model can beat for all datasets.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|cc|cc|cc|cc}
+        \toprule
+               &        & \multicolumn{2}{c}{\textbf{CLEVR single}} & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                                 \\  \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}\cmidrule(lr){9-10}
+        $e_i$  & $c$    & \textbf{loss}                             & \textbf{Acc.}                       & \textbf{loss}                       & \textbf{Acc.}                            & \textbf{loss} & \textbf{Acc.} & \textbf{loss} & \textbf{Acc.} \\\midrule
+        {100}  & {512}  & {2,45}                                    & {96\%}                              & {38,16}                             & {35\%}                                   & {46,44}       & {18\%}        & {51,8}        & {12\%}        \\
+        {100}  & {1024} & {2,18}                                    & {94\%}                              & {37,23}                             & {38\%}                                   & {47,38}       & {18\%}        & {52,04}       & {13\%}        \\
+        {100}  & {2048} & {3,44}                                    & {99\%}                              & {36,8}                              & {40\%}                                   & {46,77}       & {18\%}        & {51,29}       & {13\%}        \\
+        {500}  & {512}  & {2,71}                                    & {96\%}                              & {37,1}                              & {37\%}                                   & {47,28}       & {18\%}        & {51,01}       & {14\%}        \\
+        {500}  & {1024} & {2,61}                                    & {97\%}                              & {36,2}                              & {39\%}                                   & {48,6}        & {17\%}        & {52,73}       & {12\%}        \\
+        {500}  & {2048} & {5,18}                                    & {99\%}                              & {36,09}                             & {40\%}                                   & {44,81}       & {18\%}        & {51,8}        & {13\%}        \\
+        {1000} & {512}  & {3,11}                                    & {98\%}                              & {37,58}                             & {36\%}                                   & {46,8}        & {19\%}        & {49,95}       & {14\%}        \\
+        {1000} & {1024} & {7,91}                                    & {100\%}                             & {37,78}                             & {39\%}                                   & {47,11}       & {18\%}        & {54,18}       & {11\%}        \\
+        {1000} & {2048} & {4,64}                                    & {100\%}                             & {37,01}                             & {42\%}                                   & {45,72}       & {21\%}        & {49,98}       & {13\%}        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean loss and accuracy of the basic reference resolver for different configurations of the model after 40 epochs: $e_i$ is the image embedding dimension and $c$ the coordinate predictor dimension}
+    \label{tab:results:reference-resolver}
+\end{table}
+
+\begin{table}[b]
+    \centering
+    \begin{tabular}{cc|cc|cc|cc|cc}
+        \toprule
+               &        & \multicolumn{2}{c}{\textbf{CLEVR single}} & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                                 \\  \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}\cmidrule(lr){9-10}
+        $e_i$  & $c$    & \textbf{loss}                             & \textbf{Acc.}                       & \textbf{loss}                       & \textbf{Acc.}                            & \textbf{loss} & \textbf{Acc.} & \textbf{loss} & \textbf{Acc.} \\\midrule
+        {100}  & {512}  & {2,95}                                    & {98\%}                              & {37,27}                             & {34\%}                                   & {47,05}       & {18\%}        & {51,25}       & {12\%}        \\
+        {100}  & {1024} & {3,64}                                    & {98\%}                              & {37,19}                             & {37\%}                                   & {46,39}       & {19\%}        & {50,25}       & {14\%}        \\
+        {100}  & {2048} & {6,15}                                    & {99\%}                              & {37,04}                             & {39\%}                                   & {45,77}       & {18\%}        & {50,55}       & {14\%}        \\
+        {500}  & {512}  & {3,29}                                    & {98\%}                              & {38,58}                             & {34\%}                                   & {47,25}       & {18\%}        & {50,56}       & {14\%}        \\
+        {500}  & {1024} & {2,15}                                    & {93\%}                              & {37,18}                             & {40\%}                                   & {47,57}       & {18\%}        & {51,75}       & {14\%}        \\
+        {500}  & {2048} & {2,86}                                    & {98\%}                              & {35,71}                             & {41\%}                                   & {46,62}       & {18\%}        & {52,67}       & {11\%}        \\
+        {1000} & {512}  & {3,6}                                     & {97\%}                              & {36,37}                             & {36\%}                                   & {46,37}       & {18\%}        & {51,6}        & {12\%}        \\
+        {1000} & {1024} & {2,02}                                    & {94\%}                              & {36,86}                             & {38\%}                                   & {48,35}       & {17\%}        & {51,41}       & {15\%}        \\
+        {1000} & {2048} & {4,49}                                    & {99\%}                              & {36,5}                              & {41\%}                                   & {46,04}       & {18\%}        & {51,22}       & {13\%}        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean loss and accuracy of the \emph{reference resolver + one-hot} for different configurations of the model after 40 epochs: $e_i$ is the image embedding dimension and $c$ the coordinate predictor dimension}
+    \label{tab:results:reference-resolver-one-hot}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|cc|cc|cc|cc}
+        \toprule
+               &        & \multicolumn{2}{c}{\textbf{CLEVR single}} & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                                 \\  \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}\cmidrule(lr){9-10}
+        $e_i$  & $c$    & \textbf{loss}                             & \textbf{Acc.}                       & \textbf{loss}                       & \textbf{Acc.}                            & \textbf{loss} & \textbf{Acc.} & \textbf{loss} & \textbf{Acc.} \\\midrule
+        {100}  & {512}  & {3,02}                                    & {97\%}                              & {36,88}                             & {37\%}                                   & {47,04}       & {16\%}        & {51,36}       & {12\%}        \\
+        {100}  & {1024} & {2,48}                                    & {95\%}                              & {38,24}                             & {33\%}                                   & {46,91}       & {15\%}        & {52,67}       & {12\%}        \\
+        {100}  & {2048} & {3,8}                                     & {100\%}                             & {37,89}                             & {32\%}                                   & {45,33}       & {13\%}        & {49,12}       & {11\%}        \\
+        {500}  & {512}  & {3,6}                                     & {98\%}                              & {35,56}                             & {42\%}                                   & {47,41}       & {16\%}        & {51,86}       & {12\%}        \\
+        {500}  & {1024} & {2,51}                                    & {98\%}                              & {36,56}                             & {39\%}                                   & {45,88}       & {17\%}        & {51,73}       & {12\%}        \\
+        {500}  & {2048} & {2,37}                                    & {97\%}                              & {37,33}                             & {35\%}                                   & {48,04}       & {16\%}        & {53,68}       & {11\%}        \\
+        {1000} & {512}  & {1,97}                                    & {93\%}                              & {36,07}                             & {42\%}                                   & {47,34}       & {19\%}        & {51,74}       & {12\%}        \\
+        {1000} & {1024} & {2,99}                                    & {94\%}                              & {36,23}                             & {41\%}                                   & {45,64}       & {19\%}        & {53,27}       & {12\%}        \\
+        {1000} & {2048} & {3}                                       & {98\%}                              & {37,62}                             & {33\%}                                   & {46,54}       & {17\%}        & {52,26}       & {14\%}        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean loss and accuracy of the \emph{reference resolver + one-hot + locations} for different configurations of the model after 40 epochs: $e_i$ is the image embedding dimension and $c$ the coordinate predictor dimension}
+    \label{tab:results:reference-resolver-one-hot-location}
+\end{table}
+
+Table \ref{tab:results:reference-resolver} shows the results.
+When trained on the 'CLEVR single' dataset, the model can minimize the mean loss consistently across all variations of the variables to a few pixels.
+In effect, this means, that the model is able to almost always predict the correct location of the center of the target object.
+Since no distractor is present, the model also doesn't need any description of the target object and only needs to discriminate the object from the background.
+This shows that the model is in general able to derive geometrical information from abstract feature, extracted by a feature extractor.
+Geometrical information therefore doesn't get completely lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
+% SD: Because it is in the same location?
+% DK: (done)
+
+However, the model has more difficulties when distractors are present in the image.
+For the 'Dale-2' dataset, the mean loss converges towards 35 to 39 pixels
+The model was able to extract some structural information from the image, but the predictions are not consistently on the target object.
+Looking at the accuracy, the best variations of the models are only able to point in 41\% to 42\% of the samples towards the correct object, which lies just below a random guess of 50\%.
+Using the 'Dale-5' and the 'CLEVR color' datasets, the results are similar.
+While the mean loss stays consistently around 46 to 48 pixels and 50 to 53 pixels respectively, they are still far below the random baseline.
+Also the accuracies, lie a bit below the random guesses of 20\% and 15\% respectively.
+However, it is visible that more present distractors in the scene challenge the model more.
+% SD: What is the image size? What is the size of a typical object? 50 sounds quite good. Attention is predicting one of 7x7 blocks. Is 50 more or less than 1/7 of an image? 
+% DK: (done)
+
+Looking at the image embedding size $e_i$ and the coordinate predictor dimension $c$, it seems, they don't have any significant effect on the results.
+The results staying the same across all variations indicates that the increase compared to the random baseline is just due to learning an underlying structural pattern of how scenes are set up.
+This hypothesis will be discussed further below with examples from the predictions.
+
+Interestingly, providing the model with the target object encoded as a \textbf{one-hot vector} and the \textbf{locations} of all objects doesn't improve the results.
+% SD: Perhaps arranging the objects sequentially would help the model learn spatial contiguity, see my earlier comment.
+% DK: TODO
+Tables \ref{tab:results:reference-resolver-one-hot} and \ref{tab:results:reference-resolver-one-hot-location} show the same results as for the \emph{basic reference resolver}.
+It seems that the model is not able to combine the image representation with the representation of that target object or the locations of all objects and another representation is needed.
+
+\begin{table}[hb]
+    \centering
+    \begin{tabular}{cccc|cc|cc|cc}
+        \toprule
+              &        &          &          & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                 \\  \cmidrule(lr){5-6}\cmidrule(lr){7-8}\cmidrule(lr){9-10}
+        $e_i$ & $c$    & $LSTM_o$ & $LSTM_e$ & \textbf{loss}                       & \textbf{Acc.}                       & \textbf{loss}                            & \textbf{Acc.} & \textbf{loss} & \textbf{Acc.} \\\midrule
+        {50}  & {1024} & {1000}   & {15}     & {22,64}                             & {63,55\%}                           & {48,23}                                  & {14,95\%}     & {52,62}       & {11\%}        \\
+        {50}  & {2048} & {1500}   & {15}     & {21,35}                             & {65,9\%}                            & {46,95}                                  & {16,75\%}     & {52,2}        & {11,35\%}     \\
+        {100} & {1024} & {1000}   & {15}     & \textbf{16,08}                      & \textbf{75,5\%}                     & {47,62}                                  & {16,5\%}      & {50,65}       & {12,8\%}      \\
+        {100} & {1024} & {1000}   & {30}     & {20,69}                             & {69,5\%}                            & {46,9}                                   & {16,45\%}     & {51,23}       & {12,2\%}      \\
+        {100} & {1024} & {1500}   & {15}     & {21,24}                             & {66,3\%}                            & {48,78}                                  & {15,55\%}     & {52,21}       & {12,25\%}     \\
+        {100} & {2048} & {1000}   & {15}     & {19,46}                             & {70,1\%}                            & {46,79}                                  & {16,9\%}      & {50,24}       & {13,3\%}      \\
+        {100} & {2048} & {1500}   & {15}     & {21,19}                             & {66,65\%}                           & {46,79}                                  & {16,65\%}     & {50,55}       & {11,3\%}      \\
+        {300} & {1024} & {1000}   & {15}     & {20,63}                             & {68\%}                              & {46,91}                                  & {17,4\%}      & {50,39}       & {11,75\%}     \\
+        {300} & {1024} & {1500}   & {15}     & {21,38}                             & {66,15\%}                           & {46,36}                                  & {15,55\%}     & {51,29}       & {12,25\%}     \\
+        {300} & {2048} & {1000}   & {15}     & {21,12}                             & {66,95\%}                           & {46,98}                                  & {16,65\%}     & {51,4}        & {12,85\%}     \\
+        {300} & {2048} & {1500}   & {15}     & \textbf{14,01}                      & \textbf{79,45\%}                    & {47,41}                                  & {16,9\%}      & {51,32}       & {12,6\%}      \\
+        {300} & {2048} & {1500}   & {30}     & {22,09}                             & {64,75\%}                           & {46,77}                                  & {17,65\%}     & {49,75}       & {13,4\%}      \\
+        {500} & {1024} & {1000}   & {15}     & {35,14}                             & {41,25\%}                           & {46,41}                                  & {16,6\%}      & {50,63}       & {12,8\%}      \\
+        {500} & {2048} & {1500}   & {15}     & {36,52}                             & {41,95\%}                           & {44,9}                                   & {18,65\%}     & {50,02}       & {13,4\%}      \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean loss and accuracy of the \emph{reference resolver + incremental RE} for different configurations of the model after 40 epochs: $e_i$ is the image embedding dimension, $c$ the coordinate predictor dimension, $LSTM_o$ the output dimension of the LSTM and $LSTM_e$ the embedding dimensions for tokens in the LSTM}
+    \label{tab:results:reference-resolver-gre}
+\end{table}
+
+Encoding the attributes of the target object with the \textbf{incremental GRE-algorithm} finally improves the model's performance.
+Table \ref{tab:results:reference-resolver-gre} shows the results of selected configurations.
+Using the 'Dale-2' dataset, the model is able to predict the center coordinates of the target object for most of the images.
+The results, hereby, are dependent on the different variables.
+The best results are achieved with $e_i=300$, $c=2048$, $LSTM_o=1500$ and $LSTM_e=15$ and this configuration will be used in the succeeding analysis and the remaining language games experiments.
+With this configuration the mean distance of the model's predictions is around 14 pixels, and they lie in around 79,45\% of the cases on the target object.
+This is far below the random guess of the center coordinates as well as better than a random choice between the objects.
+Comparing it to the other configurations, it becomes apparent that an image embedding size $e_i$ between 100 and 300 dimensions provides the model with enough information from the image.
+Contrary to this, a relatively high decoder size of the LSTM $LSTM_o$ that decodes the referring expression is necessary.
+However, it seems to be dependent on $e_i$.
+A larger $e_i$ requires a larger $LSTM_o$.
+Preliminary experiments have shown that lower sizes didn't provide any good results.
+Furthermore, lower embedding dimensions of the tokens $LSTM_e$ tend to give better results.
+An $LSTM_e$ around the vocabulary size of 14 gave the best results while for instance more dimensions with 30 always worsened the model.
+
+Using the 'Dale-5' and 'CLEVR color' datasets, providing the incremental referring expression to the model didn't help.
+The results are similar to the baseline with no information about the target object.
+
+Finally, Table \ref{tab:results:reference-resolver-masked} shows the results, when the model is additionally provided with the \textbf{masked image}.
+Hereby, the model is able to predict correct locations using all datasets.
+All configurations achieve mean distances of a few pixels and accuracies of almost 100\%.
+This shows that the model can make use of the masked image.
+However, this is also not surprising since the model doesn't need to use the original image, but can completely rely on the mask.
+Since the mask only includes the target object but no distractors, it corresponds to the 'CLEVR single' dataset which as already seen in the results before is an easy task to learn for the model.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|cc|cc|cc}
+        \toprule
+               &        & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                 \\  \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
+        $e_i$  & $c$    & \textbf{loss}                       & \textbf{Acc.}                       & \textbf{loss}                            & \textbf{Acc.} & \textbf{loss} & \textbf{Acc.} \\\midrule
+        {100}  & {512}  & {8,01}                              & {98\%}                              & {6,77}                                   & {100\%}       & {6,05}        & {95\%}        \\
+        {100}  & {1024} & {8,77}                              & {99\%}                              & {3,61}                                   & {98\%}        & {5,29}        & {93\%}        \\
+        {100}  & {2048} & {8,71}                              & {99\%}                              & {4,28}                                   & {96\%}        & {3,24}        & {93\%}        \\
+        {500}  & {512}  & {14,27}                             & {99\%}                              & {9,97}                                   & {99\%}        & {5,9}         & {94\%}        \\
+        {500}  & {1024} & {8,6}                               & {99\%}                              & {2,07}                                   & {94\%}        & {13,03}       & {95\%}        \\
+        {500}  & {2048} & {2,67}                              & {95\%}                              & {10,57}                                  & {96\%}        & {4,24}        & {98\%}        \\
+        {1000} & {512}  & {6,59}                              & {95\%}                              & {9,2}                                    & {99\%}        & {7,76}        & {100\%}       \\
+        {1000} & {1024} & {5,85}                              & {99\%}                              & {12,24}                                  & {98\%}        & {8,08}        & {99\%}        \\
+        {1000} & {2048} & {6,03}                              & {99\%}                              & {9,12}                                   & {97\%}        & {14,29}       & {99\%}        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Mean loss and accuracy of the \emph{reference resolver + masking} for different configurations of the model after 40 epochs: $e_i$ is the image embedding dimension and $c$ the coordinate predictor dimension}
+    \label{tab:results:reference-resolver-masked}
+\end{table}
 
@@
     \includegraphics[width=0.8\linewidth]{figures/coordinate-predictor_loss.png}
-    \caption{Mean distance between predicted coordinates and ground truth in pixels on different datasets}
+    \caption{Mean loss of training and testing data with all datasets, using attributes encoded with the incremental GRE algorithm.
+The following variables are used for all models: $e_i=300$, $c=2048$, $LSTM_o=1500$, $LSTM_e=15$}
     \label{fig:coordinate-predictor_loss}
@@
 
-Figure \ref{fig:coordinate-predictor_loss} shows the results of the coordinate predictor that doesn't include any information about the target object.
-The used feature extractor for these results is \emph{Resnet-3}, but the results don't differ meaningfully from results with other feature extractors.
-As can be seen, the success between the different datasets are significant.
-The more objects are present in an image, the worse the model performs.
-The model converges for the CLEVR single dataset after around 20 epochs to a mean distance of around 2 pixels.
-This prediction even though not perfectly on the center point is always on the object.
-Opposed to that, using the Dale-2 dataset with two objects, the mean distance lies between 37 and 38 pixels already after the first epoch and doesn't drop with increasing number of epochs.
-With five objects in the Dale-5 dataset, the model only predicts a mean distance of around 45 pixels in the beginning, which worsens with a rising number of epochs to 48 pixels.
-
-An interesting observation is the difference of the mean distances between training and testing data.
-% SD: How can we have epochs on the testing data? There is no updates so i5 does not make sense.
-% DK: TODO
-The training distance is constantly approaching zero, while the testing loss is staying constant or even getting higher.
-This points to the fact that the model is not generalizing the task by learning abstract patterns that can be applied to unseen data, but is instead memorizing the training data.
-That is especially visible for the Dale-5 dataset, where learning the patterns of the training data looses even the ability to interpret some patterns in the testing data.
-Applying a higher dropout didn't have an impact on the results.
+Figure \ref{fig:coordinate-predictor_loss} gives more insight in why the models don't converge to a lower mean loss.
+It shows how the mean loss changes during the training after each epoch.
+Hereby, the figure shows the graph for the best performing configuration of the model that is presented with the incremental referring expression, using $e_i=300$, $c=2048$, $LSTM_o=1500$ and $LSTM_e=15$.
+The different colors correspond to the different datasets.
+Two lines are shown for each dataset.
+The solid line represents the loss when trained on the training data.
+The dotted line shows the mean distance to the target object of the model's predictions when presented with the unseen test data.
+In particular, after each epoch, the model is frozen and tested on the test split.
+After this, the weights are unfrozen again and the training is continued.
+As can be seen, the training loss of the model always converges towards a low value.
+Unsurprisingly, the datasets with fewer distractors are faster to learn than ones with more.
+For both successful datasets, 'CLEVR single' and 'Dale-2', also the training loss converges towards the same value.
+However, the testing curves for the remaining two datasets don't match their corresponding training curve.
+Indeed, both curves stay at the same value until around epoch 9.
+Then the loss of the training starts to drop while the loss in the testing starts to rise.
+That indicates that until epoch 9, the model learns to general structural patterns in the images that apply to the complete dataset, both training and testing split.
+After this, the model begins to memorize the training samples.
+This knowledge is not transferrable to the test split and indeed even harms the test performance slightly.
+That behavior can be observed for all models and configurations that don't outperform the baseline.
+Preliminary experiments showed that applying dropout to help the model to generalize didn't increase the performance which shows that the architecture is not able to extract the necessary features from the images and the referring expression and combine them.
 % SD: The results show that the model is learning something from the training data but this is not the feature that should be learning as the performance on the test data is low.
@@
 Hence, a flat line is expected.
-% DK: TODO
-
-This behavior is indeed not very surprising.
-First, results with the 'CLEVR single' dataset show that the model is able to derive geometrical information from abstract feature, extracted by a feature extractor.
-Geometrical information therefore doesn't get lost during this abstraction, but the model is able to point to a specific object, as long as only one object is part of the image.
-% SD: Because it is in the same location?
-% DK: TODO
-Secondly, more than one objects present in an image confuses the model, and it is not able to consistently point to one of them.
-This can have multiple reasons, for instance that the models lacks the ability to separate objects in the extracted features.
-But even in case, the model is able to do that and could determine the location of each object in the scene given the feature, it would not be able to tell, which of these is the actual target object.
-The guess is then more or less random.
-This especially applies to the Dale-2 dataset, where an identification of the target object just based on one distractor is impossible; both of the present objects are unique.
-For the Dale-5 dataset, the model could in theory learn that the target object is always the one object which is unique in respect to its and the distractors' attributes.
-This task on the other hand seems very difficult to learn.
-In conclusion, the models are able to predict geometrical coordinates, but need more information about the target object to identify it.
-
-When the target object \textbf{attributes are encoded as one-hot vectors} and added to the input, the results don't improve.
-One factor that now has a much higher impact is the feature extractor that is used.
-Table \ref{tab:feature-extractor-mean-distances} compares the mean distances the models predict for the different feature extractors.
-The results are shown for the models trained on the Dale-2 and Dale-5 datasets for training and test data.
-First, a big difference can be seen between the datasets.
-The models only converge to a minimum mean distance of around 46 pixels to the correct coordinates for the Dale-5 dataset, looking at the test data.
-In most cases, it stays above 50 pixels.
-% SD: What is the image size? What is the size of a typical object? 50 sounds quite good. Attention is predicting one of 7x7 blocks. Is 50 more or less than 1/7 of an image? 
-% DK: TODO
-Using the Dale-2 dataset, the behavior is a little different.
-All ResNet extractors with four residual blocks and an additional average pooling and optionally a classifier layers reach similar scores as the experiments before without any one-hot encodings.
-Interestingly, without the classifier layer, the model doesn't converge at all and the mean distances jump up and down between the epochs.
-This effect also applies when using less residual blocks.
-Using the VGG, only VGG-cls2 achieve a similar performance, while the others predict coordinates between 43 and 46 pixels away.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{rcccc}
-        \toprule
-                            & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}}                                   \\\cmidrule(lr){2-3}\cmidrule(lr){4-5}
-                            & train                               & test                                & train          & test           \\\midrule
-        \textbf{VGG-0}      & 30,27                               & 46,20                               & \textbf{32,17} & 54,40          \\
-        \textbf{VGG-avg}    & \textbf{29,99}                      & 45,08                               & 32,32          & 52,67          \\
-        \textbf{VGG-cls1}   & 37,99                               & 43,28                               & 46,57          & 50,75          \\
-        \textbf{VGG-cls2}   & 38,87                               & \textbf{39,02}                      & 47,48          & 49,91          \\
-        \textbf{VGG-cls3}   & 39,99                               & 44,32                               & 47,26          & \textbf{46,77} \\\midrule
-        \textbf{ResNet-3}   & 78,26                               & 65,23                               & 92,07          & 91,12          \\
-        \textbf{ResNet-4}   & 44,14                               & 55,24                               & \textbf{36,48} & 58,28          \\
-        \textbf{ResNet-avg} & \textbf{33,06}                      & \textbf{39,18}                      & 47,64          & 46,38          \\
-        \textbf{ResNet-cls} & 37,57                               & 38,10                               & 44,72          & \textbf{45,92} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Mean test losses for different feature extractors with one-hot attribute encodings after 20 epochs}
-    \label{tab:feature-extractor-mean-distances}
-    % SD: Explanation of labels
-    % DK: TODO
-\end{table}
-
-Secondly, the training loss now looks also different.
-In almost no cases, the models converge to a lower mean distance than with the test data, meaning a higher precision in their predictions, as they did in the experiment before.
-The only exception is ResNet-3 as a feature extractor.
-In other words, the models are again not able to generalize, but in specific cases memorize the patterns in the train data.
-This hints to the fact that only specific layers of the feature extractors contain information that is generally usable to identify and discriminate objects.
-Especially the lower layers with fewer residual blocks in the case of ResNet and no classifier layers for the VGG seem to not encode knowledge that can be utilized for this task.
-Higher layers, with more specific encoded information need to be used for this research.
-The experiments in the following sections are set up using these higher layers.
-
-Adding \textbf{information about the center coordinates} of all objects should have helped the models to get a list of possible predictions.
-In theory, the model could learn to choose between these coordinates by relating them to the extracted features of the image.
-This hypothesis doesn't hold.
-All results for both datasets Dale-2 and Dale-5 are the exact same as without included information about the locations.
-The problem therefore doesn't seem to lie in predicting coordinates in general, but predicting the coordinates of the target object.
-% SD: Perhaps arranging the objects sequentially would help the model learn spatial contiguity, see my earlier comment.
-% DK: TODO
-The model is still not able to understand, which object is the target object.
-For that reason, a better representation of the target object is necessary.
-
-In a next step, information about the attributes is included using the \textbf{incremental GRE-algorithm}.
-Again, the mean distance of the predictions as well as the accuracy doesn't improve compared to the previous experiments.
-% SD: Table with these results?
-% DK: the table would show the same figures as the existing table. Should be still included? (QUESTION)
+% DK: (done)
 
@@
     \subfigure['Dale-2', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_train.png}
+        \includegraphics[width=.839\linewidth]{figures/visualization_dale-2_train.png}
         \label{fig:visualizations_dale-2_train}
@@
     \subfigure['Dale-2', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-2_test.png}
+        \includegraphics[width=.839\linewidth]{figures/visualization_dale-2_test.png}
         \label{fig:visualizations_dale-2_test}
@@
     \subfigure['Dale-5', train split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_train.png}
+        \includegraphics[width=.839\linewidth]{figures/visualization_dale-5_train.png}
         \label{fig:visualizations_dale-5_train}
@@
     \subfigure['Dale-5', test split]{
-        \includegraphics[width=.92\linewidth]{figures/visualization_dale-5_test.png}
+        \includegraphics[width=.839\linewidth]{figures/visualization_dale-5_test.png}
         \label{fig:visualizations_dale-5_test}
     }
-    \caption{Visualization of the models' predictions in the 'Dale' datasets}
-    \label{fig:visualizations_dale}
+    \subfigure['CLEVR color', train split]{
+        \includegraphics[width=.839\linewidth]{figures/visualization_colour_train.png}
+        \label{fig:visualizations_colour_train}
+    }
+    \subfigure['CLEVR color', test split]{
+        \includegraphics[width=.839\linewidth]{figures/visualization_colour_test.png}
+        \label{fig:visualizations_colour_test}
+    }
+    \caption{Examples of the models' predictions on the 'Dale' and 'CLEVR color' datasets.
+The green circle represents the target object, the red object the model's prediction.
+Notice that the images are shown in the original size while the model processes them normalized.
+Hence, the distances don't correspond to the axes.}
+    \label{fig:visualizations}
 \end{figure}
@@
 An interesting pattern appears when doing a qualitative analysis of the models' predictions.
-Here, we visualized the predicted coordinates compared to the ground truth coordinates.
+Here, the predicted coordinates compared to the ground truth coordinates are visualized (Figure \ref{fig:visualizations}), using the best performing model with $e_i=300$, $c=2048$, $LSTM_o=1500$ and $LSTM_e=15$.
 Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
@@
 As can be seen, the predictions are very precise.
-Figure \ref{fig:visualizations_dale-2_complete_train} combines the predictions and ground truths across all images in the train dataset.
+Figure \ref{fig:visualizations_complete} combines the predictions and ground truths across all images in the train dataset.
 This shows general patterns of the models predictions over the complete dataset.
 Here, all predicted coordinates are placed as red circles into the image, while all ground truth coordinates are placed as green circles.
-The resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene.
+For the predictions on the 'Dale-2' dataset, the resulting shape is a rhombus, which reflects that all objects are placed usually central into the scene (see Figure \ref{fig:visualizations_dale-2_complete_train}).
 As expected the green and red rhombus align mostly in the same area for the train split of the 'Dale-2' dataset.
@@
 % DK: that is done in the quantitative analysis (mean square error above). The visualization should show general learned patterns of the model (as in this case predictions toard the center)
+Figures \ref{fig:visualizations_dale-5_train} and \ref{fig:visualizations_dale-5_complete_train} show the same patterns for the 'Dale-5' dataset, while Figures \ref{fig:visualizations_colour_train} and \ref{fig:visualizations_colour_complete_train} show it for the 'CLEVR color' dataset.
 
-The results look very different for the test split.
-As can be seen in Figure \ref{fig:visualizations_dale-2_test}, the three randomly selected predictions don't align with the ground truth coordinates.
-For all the images, the predictions don't lie on any object.
-In the left image as well as in the central image, the predictions are closer to the target object than towards the distractor, but are still quite imprecise.
+The results look very different when using the predictions for the test split.
+As discussed in the paragraphs above, the model performs well on the 'Dale-2' dataset which is also reflected in the qualitative analysis.
+The visualization of the test split resembles the train split very much.
+However, the other two datasets show differences.
+As can be seen in Figures \ref{fig:visualizations_dale-5_test} and \ref{fig:visualizations_colour_test}, the three randomly selected predictions don't align with the ground truth coordinates.
+While for some images they lie on a distractor, on some images, the prediction is on the background.
+However, they tend to be in an area towards the target object, but are still quite imprecise.
 % SD: They are closer to the target than distractor and you can see that it is working to a point. But remember this task is very challenging and I wouldn't say the results are so negative.
 % DK: right, this is addressed in the conclusions below
-These findings align with the mean distance scores, described in the sections before.
+These findings align with the mean distance scores, described in the paragraphs before.
 However, it seems that the model's predictions are all towards the center of the image.
-This can be seen clearer in Figure \ref{fig:visualizations_dale-2_complete_test}.
+This can be seen clearer in Figures \ref{fig:visualizations_dale-5_complete_test} and \ref{fig:visualizations_colour_complete_test}.
 Again, the green circles form the shape of rhombus.
@@
 % DK: again, conclusion below
-This behavior can be observed for all datasets and architectures of the model.
-Figures \ref{fig:visualizations_dale-5_train}, \ref{fig:visualizations_dale-5_test}, \ref{fig:visualizations_dale-5_complete_train} and \ref{fig:visualizations_dale-5_complete_test} show the results for the 'Dale-5' dataset.
-Here, the model more likely predicts the center coordinates of a distractor object as seen in the right image, which is also reflected in the lower score of the mean distance.
-Also the combined visualization shows the same clustering of predictions in the center of the scene, but the pattern of the smaller rhombus is more visible.
 % SD: The centre bias could be the way the error function is used, that is averages all distances and of course this has a tendency to some middle distance.
@@
     }
-    \caption{Visualization of the models' predictions in the Dale datasets}
+    \subfigure['CLEVR color', train split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_colour_train_complete.png}
+        \label{fig:visualizations_colour_complete_train}
+    }
+    \subfigure['CLEVR color', test split]{
+        \includegraphics[width=0.42\linewidth]{figures/visualization_colour_test_complete.png}
+        \label{fig:visualizations_colour_complete_test}
+    }
+    \caption{The images show all the model's predictions (red circles) and the ground truths (green circles) combined for the 'Dale' and 'CLEVR color' datasets}
     % SD: I Haven't checked the earlier captions, but the captions should be informative in the sense that one does not need to look for the text to understand them.
 Hence, include a brief summary of what each figure contains.
-    % DK: TODO
-    \label{fig:visualizations_dale_complete}
+    % DK: (done)
+    \label{fig:visualizations_complete}
 \end{figure}
 
-These results allow two conclusions.
-First, the models are biased to predict coordinates in the center of the image.
+These results allow three conclusions.
+First, the underlying structure in all images is the placement of objects in the center of the scene.
+The models, learn to use the structure and are biased to predict coordinates towards the center.
 % SD: centre - using British English?
@@
 Since all objects are always located in the center of the image and never in its corners, a prediction of any coordinate in the center is on average closer to the target object than any random prediction or predictions of coordinates at the borders of the image.
-The model therefore learns only, where any object is likely located and can minimize the mean distance to a certain extent with this strategy.
+The model therefore learns only, where \emph{any} object is likely located and can minimize the mean distance to a certain extent with this strategy.
 
-Second, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
-This can be seen for the 'Dale-5' dataset, especially in left and central image in Figure \ref{fig:visualizations_dale-5_test}.
+Secondly, even though the models are biased towards the center of the image, the predictions are still often leaning towards the location where many objects lie.
+This can be seen for the 'CLEVR color' dataset, especially in left and central image in Figure \ref{fig:visualizations_colour_test}.
 Again, by this strategy, the model can minimize the mean distance, since the probability is high that the target object lies in this cluster of objects.
-Concluding, the model is able to extract, where objects are located in the image, but can't make use of the referring expressions, to decide which of these objects is the target object.
+
+Thirdly, when few distractors are present in the scene, the model is able to combine the encoding of the referring expression with the encoding of the image.
+In particular, it is able to learn which of the objects in the image is described by the referring expression.
+As soon as more objects are present, the task gets to difficult and the model only relies on the general patterns in the dataset.
+Concluding, the model is able to extract, where objects are located in the image, but can't consistently make use of the referring expressions, to decide which of these objects is the target object.
 % SD: There is a centre bias and the model can partially locate the target.

### 2023-11-28 source/mlt/preexperiments/models.py

@@ -1,3 +1,5 @@
+import random
+
 import torch
 from mlt.shared_models import (
     BoundingBoxImageEncoder,
@@ -395,6 +397,28 @@ class MaskedCoordinatePredictor(nn.Module):
         coordinates = self.coordinate_classifier(reduced)
 
         return coordinates
+
+
+class RandomCoordinatePredictor(nn.Module):
+    """
+    Output:
+     - x and y coordinates of target object
+
+    Input:
+    """
+
+    def __init__(self, *_args, **_kwargs) -> None:
+        super().__init__()
+
+        self.dummy = nn.Linear(1, 1)
+
+    def forward(self, data):
+        image, *_ = data
+        # 224 = image width and height
+        return (
+            torch.rand((image.shape[0], 2), device=image.device, requires_grad=True)
+            * 224
+        )
 
 
 class CaptionDecoder(nn.Module):

### 2023-11-28 source/mlt/preexperiments/train.py

@@ -30,6 +30,7 @@ from mlt.preexperiments.models import (
     MaskedCaptionGenerator,
     MaskedCoordinatePredictor,
     MaskedDaleAttributeCoordinatePredictor,
+    RandomCoordinatePredictor,
 )
 from mlt.preexperiments.save import (
     BoundingBoxOutputProcessor,
@@ -267,6 +268,19 @@ models = {
             "image_embedding_dimension": 4096,
             "coordinate_classifier": CoordinateClassifier,
         },
+        loss_function=pixel_loss,
+        tester=CoordinatePredictorTester,
+        output_processor=PixelOutputProcessor,
+        output_processor_args={
+            "output_fields": ("image_id", "x", "y", "target_x", "target_y")
+        },
+    ),
+    "random_coordinate_predictor": ModelDefinition(
+        dataset=CoordinatePredictorDataset,
+        dataset_args={},
+        preprocess=ResNet101_Weights.IMAGENET1K_V2.transforms(),
+        model=RandomCoordinatePredictor,
+        model_args={},
         loss_function=pixel_loss,
         tester=CoordinatePredictorTester,
         output_processor=PixelOutputProcessor,

### 2023-11-28 source/mlt/print_results.py

@@ -223,13 +223,13 @@ class Experiments:
 
 class OutputProcessor:
     @classmethod
-    def to_percent(cls, number):
-        rounded = str(round(number * 100, 2)).replace(".", ",").rstrip("0").rstrip(",")
+    def to_percent(cls, number, r=2):
+        rounded = str(round(number * 100, r)).replace(".", ",").rstrip("0").rstrip(",")
         return rf"{rounded}%"
 
     @classmethod
-    def to_number(cls, number):
-        return rf"{str(round(number, 2)).rstrip('0.').replace('.', ',')}"
+    def to_number(cls, number, r=2):
+        return rf"{str(round(number, r)).replace('.', ',').rstrip('0').rstrip(',')}"
 
     def choose_metrics(self, experiments):
         all_metrics = set(
@@ -358,7 +358,7 @@ class LatexOutputProcessor(OutputProcessor):
             ]
 
             variable_string = "        " + " & ".join(
-                [f"{{{variable}}}" for variable in variation]
+                [f"{{{self.to_number(variable, 10)}}}" for variable in variation]
             )
 
             result_string = " & ".join(

### 2023-11-28 thesis/sections/5.2-re-generation.tex
results
results re generation

@@
     \label{fig:masked_re_generator_architecture}
-    \cmtDK[inline]{add linear layer for LSTM}
 \end{figure}
@@
 Again, each of these datasets increases the complexity of the description.
-While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
+While the referring expression for the \emph{Dale-2} datasets are generally shorter, expressions of the \emph{Dale-5} datasets need to be more specific and use more attributes.
 Additionally, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-Furthermore, the models are trained on the 'CLEVR color' dataset.
+Furthermore, the models are trained on the \emph{CLEVR color} dataset.
 In this dataset, the focus lies mainly on only one discriminating attribute, the color.
@@
 For the \emph{Dale-5} dataset, the model struggles more to predict the correct length of the referring expression.
-\cmtDK[inline]{why?}
 
@@
 In the following paragraphs, the results of the \textbf{basic RE generator} as well as the \textbf{masked RE generator} are evaluated.
-Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 40 epochs.
 Compared to the \emph{bounding box RE generator}, the task differs substantially.
 Instead of bounding boxes, the model is shown the image of the whole scene.
-Additionally to extracting features of the target object and distractors separately, the model is now tasked to extract these at once for all objects.
+In addition to extracting features of the target object and distractors separately, the model is now tasked to extract these at once for all objects.
 Furthermore, it needs to learn which of the objects is the target object by combining the image with the masked image.
+The \emph{basic RE generator} acts hereby as a baseline.
+Since there is no information about the target object present, the model is expected to extract the features of all objects, but not to be able to focus on the target object.
+Masking should help the model, to also solve the second task.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|ccc|ccc|ccc}
+        \toprule
+                  &           & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR colour}}                                                                                         \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        $LSTM\_o$ & $LSTM\_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                               & \textbf{Acc.} & \textbf{F1} & \textbf{NT} & \textbf{Acc.} & \textbf{F1} & \textbf{NT} \\\midrule
+        {100}     & {30}      & {38}                                & {41,16}                             & {40}                                      & {14}          & {48,91}     & {34}        & {15}          & {36,98}     & {34}        \\
+        {1500}    & {15}      & {45}                                & {62,66}                             & {45}                                      & {17}          & {51,89}     & {38}        & {13}          & {35,85}     & {33}        \\
+        {2000}    & {10}      & {45}                                & {60,72}                             & {44}                                      & {16}          & {50,77}     & {37}        & {13}          & {34,96}     & {36}        \\
+        \bottomrule
+    \end{tabular}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the basic RE generator after 30 epochs for selected configurations: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \label{tab:results:basic-re-generator}
+\end{table}
+
+Table \ref{tab:results:basic-re-generator} shows some selected results for the \emph{basic RE generator} after 30 epochs.
+The sum of the \emph{overall accuracy} and \emph{non-target accuracy} represents in how many cases any of the shown objects were described, independently if it is a distractor or the target object.
+As can be seen, the sum is always relatively high independently of the dataset.
+This shows verifies that the model is able to extract information about the objects in the image.
+However, it can't decide which object to describe.
+Most clearly, this is shown with the \emph{Dale-2} dataset where the sum lies around 90\%.
+In almost all samples, the  model describes an object perfectly, but in 45\% of the cases it is the target object, in 45\% the distractor.
+Which object, the model chooses is pure chance.
+The same pattern is visible for the other datasets.
+Since there are more distractors in the scenes, the model is more likely to describe on of those, when deciding randomly, and the \emph{non-target accuracy} is higher than the \emph{overall accuracy}.
+However, another fact plays a role here.
+The incremental GRE-algorithm can only generate unique referring expressions, when the referred to object is unique in the scene.
+This applies to all target objects, but there is no such restriction on the distractors.
+One scene can include for instance multiple \emph{large red spheres} as distractors.
+The algorithm then generates two identical referring expressions which can affect the \emph{non-target accuracy}.
+Indeed, while the average number of distractors per scene in the \emph{CLEVR color} dataset is 7,5, the average number of unique distractors lies at 2,7.
+As seen later, the model learns to correctly never predict a \emph{size} and always predict the correct \emph{shape}.
+Given this fact, a random guess of a color would predict in $\frac{1}{8} * 2,7 = 33,9\%$ of the cases a unique distractor while in $\frac{1}{8}=12,5\%$ of the cases it would predict the target object.
+This corresponds the \emph{non-target accuracies} and \emph{overall accuracies} the \emph{CLEVR color} dataset.
+The high \emph{non-target accuracy} is therefore misleading and doesn't represent that the model identified other objects in the scene.
 
@@
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked RE generator after 40 epochs: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
+    \caption{Overall accuracies (Acc.), F1-Score (F1) and non-target accuracies (NT) in \% of the masked RE generator after 30 epochs: $LSTM_o$ are different LSTM output sizes and $LSTM_e$ are different embedding sizes for the tokens in the LSTM.}
     \label{tab:results:masked-re-generator}
@@
 
-Consequently, the results are much less accurate.
+Table \ref{tab:results:masked-re-generator} shows the results for the \emph{masked RE generator} after 30 epochs.
+A first look at the sum of the \emph{overall accuracy} and the \emph{non-target accuracy} shows the model is able to extract the attributes from the shown objects.
+For the \emph{Dale-2} dataset, this number lies at 90\% for the best configuration, at 51\% for the \emph{Dale-5} dataset and at 55\% for the \emph{CLEVR color} dataset.
+Those sums correspond more or less to the sums of the model without masking shown in Table \ref{tab:results:basic-re-generator}.
+In other words, the masking doesn't hinder the model to extract and describe attributes in the image.
+However, the balance in between the two accuracy scores shifts towards the \emph{overall accuracy}, meaning that the model is now able to identify the correct target object more often.
+While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the \emph{Dale-2} dataset, the model only reaches 36\% for the \emph{Dale-5} dataset and 29\% for the \emph{CLEVR color} dataset.
+The scores for all datasets are all well above the \emph{basic RE generator} baselines of 45\%, 16\% and 15\% respectively, while in all cases the \emph{non-target accuracies} drop lower.
+In other words, the masked image helps the model to focus on the correct object.
+
+Comparing it to the less complex \emph{bounding box RE generator}, the results are much less accurate.
 This applies to all datasets, even though it is most apparent for the \emph{CLEVR color} dataset.
-While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the \emph{Dale-2} dataset, the model only reaches 36\% for the \emph{Dale-5} dataset and 29\% for the \emph{CLEVR color} dataset.
-However, these scores are all well above the random baselines of 50\%, 20\% and 10\% respectively.
-Moreover, the referring expression generated for the \emph{Dale-2} dataset are also efficient in the sense that they follow incremental GRE-algorithm and only use necessary discriminative attributes.
+Given the highly increased complexity of the task, the model fares well in comparison on the \emph{Dale} datasets.
+
+Moreover, the referring expression generated for the \emph{Dale-2} dataset are also efficient in the sense that they follow the incremental GRE-algorithm and only use necessary discriminative attributes.
 Features of both target object and distractor are therefore extracted and associated with the vocabulary.
@@
 For the \emph{Dale-2} dataset, it arrives at 11\% in one configuration, for the \emph{Dale-5} dataset, the model describes for at least 12\% of the samples a distractor, for one configuration even in over 25\% of the cases.
-The model is most uncertain which object to describe when exposed to the \emph{CLEVR color} dataset where in almost all configurations, it describes a distractor every third sample.
-This shows that the model struggles to learn which of the objects in the scene is the target object, more specifically, to combine the whole image with the masked image.
-\cmtDK[inline]{compare to basic RE generator}
-Yet, the model is able to extract the attributes from the shown objects.
-This can be seen, when summing up the \emph{overall accuracy} and the \emph{non-target accuracy}.
-This number represents in how many cases any of the shown objects were described.
-For the \emph{Dale-2} dataset, this number lies at 90\% for the best configuration, at 51\% for the \emph{Dale-5} dataset and at 56\% for the \emph{CLEVR color} dataset.
-While being still lower than the same metric for the \emph{bounding box RE generator}, the model fares well in comparison for the \emph{Dale} datasets given the highly increased complexity of the task.
+The \emph{non-target accuracy} of the \emph{CLEVR color} is misleading, and is therefore not considered.
+This shows that the model still struggles to learn which of the objects in the scene is the target object, even though masking is applied.
 
@@
 Opposed to the former datasets, a lower $LSTM_o$ helps the model to describe the correct target object.
-On the other side, the \emph{non-target accuracy} stays relatively constant independently of the variables, in other words
+On the other side, the \emph{non-target accuracy} stays relatively constant independently of the variables at the same level as for the \emph{basic RE generator}.
+This indicates that the model uses random guesses of the color when it is not able to extract the color of the target object.
+To confirm this, the model's performance on different tokens and attributes are analyzed.
 
-\cmtDK[inline]{discuss attributes}
 \begin{table}[ht]
@@
 
+Tables \ref{tab:results:masked-re-generator_size-shape} and \ref{tab:results:masked-re-generator_color} shows the precision and recall for each token.
+The metrics on the \emph{<pad>} token show that the model always predicts the correct length of the referring expression when presented with the \emph{CLEVR color} dataset.
+For the \emph{Dale-2} dataset, the model is able to do this in almost all the cases, while the model struggles the most with the \emph{Dale-5} dataset.
+While this was similar to the \emph{bounding box RE generator}, the effect here is amplified.
+Furthermore, the easiest attribute to predict is the shape.
+Again, the metrics for the \emph{Dale-2} and \emph{CLEVR color} dataset are almost the same, but the model is much less able to predict the correct shape for images from the \emph{Dale-5} dataset with only around 85\%.
+Hereby, the precision is the highest for predicting \emph{cubes} and lowest for predicting \emph{cylinders}, but this varies across runs.
+
+The prediction of the color drops very much in comparison.
+While the metrics all lied above 90\% for the \emph{bounding box RE generator}, now none of datasets passes this limit.
+With the \emph{Dale-2} dataset, the precision and recall lie around 85\%, with the \emph{Dale-5} dataset around 70\% and around 30\% with the \emph{CLEVR color} dataset.
+The drop for the 'Dale' datasets is expected, because of the higher complexity of the task, but the model's performance on the \emph{CLEVR color} dataset is much worse.
+Only every third prediction is correct.
+Hereby, a difference across the colors is visible.
+Some colors reach both precision and recall of around 40\%, while others only of only around 20\%.
+Again, which color is captured better differs from run to run.
+Interesting is also the margin between the precision and recall for the colors 'blue', 'cyan', 'green', 'purple' and 'yellow'.
+A low precision and high recall indicates that the model predicts the specific color more often than needed while the opposite indicates that it predicts the color less often than needed.
+In particular, 'blue', 'green' and 'yellow' are predicted too often and 'cyan' and 'purple' are predicted too few.
+
+Finally, the size is again the most difficult attribute to predict.
+As the \emph{bounding box RE generator}, the \emph{masked RE generator} learns correctly to never predict a size for the \emph{CLEVR color} dataset.
+Interestingly is the fact that the recall is low which shows that the model likely predicts a \emph{<pad>} token instead.
+This might be due to the fact that the \emph{<pad>} token is very frequent in the size slot across all referring expressions and the model amplified this fact.
+Between both sizes, 'large' objects are easier to predict for the model than 'small' objects.
+The difference lies for both \emph{Dale} datasets at around 8\%.
+However, the precision is still far above a random guess, when used on the \emph{Dale-2} dataset.
+Using the \emph{Dale-5} dataset, the precision only lies at around 50\% to 57\%.
+
 % With 21\% of these descriptions being a description of the target object, the model again uses a random guess.
@@
 
-There are two main explanations for the big difference between the datasets.
+In conclusion, when the model doesn't need to extract the objects from the scene it performs very well.
+In this case, it is able to achieve almost perfect results on the \emph{Dale-2} dataset and very good results on the remaining two datasets.
+However, using complete scenes impairs the performance highly.
+The \emph{overall accuracies} decrease and the \emph{non-target accuracies} increase; the model has more difficulties to focus on the correct object.
+The most difficulties the model has with the \emph{CLEVR color} dataset where only the color is a discriminative attribute.
+Looking at the attributes, the last slot (shape) is always the easiest to predict, while the first slot (size for the \emph{Dale} dataset and color for the \emph{CLEVR color} dataset) is the most difficult.
+This effect is again amplified when the model needs to extract the objects itself instead of being presented with bounding boxes.
+
+There are two main explanations for these results.
 First, as already seen in the previous experiments, a bigger number of distractors confuses the model more where to focus on.
 In the \emph{Dale-2} dataset, there are only two possibilities, while the four distractors in the \emph{Dale-5} dataset and up to nine distractors in the \emph{CLEVR color} give the model a bigger choice.
+Moreover, having only one attribute to discriminate the target object, poses the model with more difficulties.
 % SD: But before you argued, rather pesimistically, that having distractors with the same attributes actaully contributes to greater performance as it is moire likely that the correct word is generated.
@@
 % DK: TODO
-The second explanation lies in the used GRE-algorithm.
-When only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
+The second explanation lies in the incremental GRE-algorithm.
+For instance using the \emph{Dale-2} dataset, when only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
 Otherwise, the shape is enough, and the RE is only one word long.
@@
 For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors are 19,8\%, 47\% and 33,2\% respectively.
+Opposed to that the probabilities with four distractors as in the \emph{Dale-5} dataset are 19,8\%, 47\% and 33,2\% respectively.
 With four distractors the algorithm is much more likely to produce longer referring expressions.
 These are harder to learn and generate for the model since it needs to take more extracted attributes into account to discriminate the target object from the distractors.
+This is verified by the lower precisions and recalls of the size and color tokens.
+Furthermore, the low recall of the size tokens shows that the model produces too short referring expressions too often.
 % SD: I think we need individual accuracies per class and also precision, recall and the f-score.

### 2023-11-28 thesis/sections/5.3-reference-resolution.tex
results re generation

@@
     }
-    \caption{The images show all the model's predictions (red circles) and the ground truths (green circles) combined for the 'Dale' and 'CLEVR color' datasets}
+    \caption{The images show all the model's predictions (red dots) and the ground truths (green dots) combined for the 'Dale' and 'CLEVR color' datasets}
     % SD: I Haven't checked the earlier captions, but the captions should be informative in the sense that one does not need to look for the text to understand them.

### 2023-11-28 thesis/sections/6.1-object-identification.tex
results re generation

@@
 
-\subsection*{Results}
+\subsubsection*{Results}
 Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.

### 2023-11-28 source/mlt/preexperiments/models.py

@@ -1,5 +1,3 @@
-import random
-
 import torch
 from mlt.shared_models import (
     BoundingBoxImageEncoder,

### 2023-11-28 source/mlt/preexperiments/run_experiments.py

@@ -2,28 +2,27 @@ import itertools
 import subprocess
 
 options = [
-    #"--dataset_base_dir=/scratch/guskunkdo",
+    # "--dataset_base_dir=/scratch/guskunkdo",
     "--dataset_base_dir=/home/dominik/Development/",
-    # "--lr=0.0002",
-    "--epochs=40",
+    "--lr=0.0002",
+    "--epochs=30",
     "--feature_file=resnet_3_no-avgpool_no-fc.h5",
     "--max_samples=10000",
     "--device=cuda",
-    "--model=dale_attribute_coordinate_predictor",
+    "--model=caption_generator",
     "--out=out/",
     # "--out=/scratch/guskunkdo/out/",
 ]
 
 variables = {
-    #"--dataset": ["dale-2"],
-    "--dataset": ["dale-5", "colour"],
-    #"--image_embedding_dimension": [300, 500],
-    "--image_embedding_dimension": [50, 100, 300, 500],
-    "--coordinate_classifier_dimension": [1024, 2048],
-    # "--decoder_out_dim": [100, 500, 1000],
-    "--encoder_out_dim": [1000, 1500],
-    "--embedding_dim": [15, 30],
-    "--lr": [0.0001, 0.00005, 0.00003],
+    # "--dataset": ["dale-2"],
+    "--dataset": ["dale-2", "dale-5", "colour"],
+    # "--image_embedding_dimension": [50, 100, 300, 500],
+    # "--coordinate_classifier_dimension": [1024, 2048],
+    "--decoder_out_dim": [2000],
+    # "--encoder_out_dim": [1000, 1500],
+    "--embedding_dim": [10],
+    # "--lr": [0.0001, 0.00005, 0.00003],
 }
 
 for index, combination in enumerate(itertools.product(*variables.values())):

### 2023-11-28 source/mlt/preexperiments/train.py

@@ -455,7 +455,7 @@ if __name__ == "__main__":
     parser.add_argument("--decoder_out_dim", type=int, default=10)
     parser.add_argument("--encoder_out_dim", type=int, default=10)
     parser.add_argument("--embedding_dim", type=int, default=10)
-    parser.add_argument("--image_embedding_dimension", type=int, default=10)
+    parser.add_argument("--image_embedding_dimension", type=int, default=None)
     parser.add_argument("--coordinate_classifier_dimension", type=int, default=10)
 
     # -- TRAINING --

### 2023-11-30 thesis/sections/6-grounding-language-games.tex
language games setup

@@
 \label{sec:language-games}
-\cmtDK[inline]{'epochs', training/testing (success), optimizer, ...}
-\cmtDK[inline]{max message length}
-
 The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
@@
 Over time, specific meanings are reinforced and a language emerges.
-The number of possible words, namely the size of the vocabulary can be varied.
+Hereby, the production of the messages is constrained in two ways.
+First, the size of the vocabulary is varied.
+A smaller size forces the agents to compress their information more which is a prerequisite for the emergence of a language.
+With more different words, they can express more information, but it is harder to learn.
+Different sizes are compared to analyze its effect on the emerged language.
+Secondly, the maximum possible length of a message is constrained.
+Again, a shorter message requires more compression, while the longer message allows for more information exchange.
 % SD: But possibly you would describe the language game setup already before on the background or method chapter.
@@
 It is important to emphasise this difference.
-% DK: TODO
+% DK: (in both beginning and here, done)
 
-First, we will discuss \emph{discrimination games}, because they have the simplest setup.
-% SD: What is this setup?
-% DK: TODO
-Furthermore, other language games that research this topic use a very similar setup.
-In the next step, we will describe \emph{caption generators} that are set up as a language game.
-Here, the sender describes the scene, while the receiver needs to generate a caption.
-In the last step, we try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.
-% SD: Why are we not testing it on all 3 games?
-% SD: Not sure what you mean here? Why do we loose human bias? 
-% DK: TODO
+As the experiments before, three different tasks are set up.
+For the \emph{object identification} task, the agents play discrimination games similar to the setup in \citep{Lazaridou2017}.
+For the \emph{RE generation} task, the receiver needs to generate a human language referring expression based on the emerged language referring expression by the sender.
+Since, it is possible that the agents simply learn to align these two languages, the agents are tasked to point towards the center coordinates of the target object in the final task, the \emph{reference resolution} task.
 
-The languages that emerge are analyzed in two ways.
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
+As discussed in Section \ref{sec:background:lg:setup}, language games are evaluated on the success of communication and the emergence of a language.
+The testing phase is not necessarily used to test the generalization capabilities of the agents, but more to extract a sample of the emerged language.
+It is therefore not as important that the samples in the testing phase are unseen samples.
+Hence, for both training and testing a certain number of samples are randomly drawn from the complete dataset.
+In these experiments, the dataset consists of 10.000 scenes in total.
+To train the agents, 12.800 games are played (using 12.800 randomly drawn samples from the dataset) with a batch size of 32 samples.
+To retrieve the samples of the emerged language, after the agents are trained, they are frozen and play 2.048 games with randomly drawn samples from the dataset.
+This yields 2.048 interactions that will be used to analyze the emerged language.
 
-\cmtDK[inline]{discuss training/epochs/batches, vocab size, message length, gs temperature}
+All agents are trained using the \emph{Adam} optimizer \citep{Kingma2015}.
+LSTMs are used to produce and parse the message between the agents.
+Gumbel-Softmax relaxation is applied to train the agents with those discrete messages.
+Hereby, a temperature of $\tau=1$ is used which showed the best results in preliminary experiments.

### 2023-11-30 thesis/sections/7-language-analysis.tex
results re generation
language games setup

@@
 \label{sec:analysis_language}
+The languages that emerge are analyzed in two ways.
+An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
+In a second step, the language is compared to several referring expressions in natural language.
+By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
+% from prev chapter
+
 In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
@@
 \end{figure}
+
+\cmtDK[inline]{differnces in different languages correspond to differences in color prediction of single model caption generation}
 

### 2023-12-07 thesis/sections/5.3-reference-resolution.tex
experiments

@@
 For this reason, the final task of the receiver will be to 'point' to the target object.
-The models are therefore tasked to predict the center coordinates of the target object.
+In this section, this is done in two ways.
+First, the models are tasked to predict the center coordinates of the target object (\emph{coordinate reference resolver}).
 % SD: Reference resolution with identification of location.
@@
 % DK: (done)
-With this approach, the models receive few natural language information, but are still able to rely on all information present in the image to discriminate the objects.
+This requires a very high precision of the model in extracting spatial information.
+However, the exact location is not necessarily important to highlight, which object the model predicts to be the target object, and might challenge the model too much.
+Therefore, in a second approach, the model is tasked to predict a region around the target object (\emph{attention reference resolver}).
+This reduces the necessary amount of spatial information and might be easier to learn for the models.
+With both approaches, the models receive few natural language information, but are still able to rely on all information present in the image to discriminate the objects.
 
-To achieve this goal, multiple setups of models are trained.
-In the simplest setup, the \textbf{reference resolver}, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
+In the simplest setup, the \textbf{coordinate reference resolver}, the model receives only the image as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
 % SD: What are the features?
@@
 Again, the image is encoded, using the \emph{image encoder} submodule, described in Section \ref{sec:image-processing}.
-The \emph{coordinate predictor} submodule takes the encoder image to predict x- and and y-coordinates of the target image,
+The \emph{coordinate predictor} submodule takes the encoder image to predict x- and y-coordinates of the target image,
 Hereby, the encoded image is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
@@
     }
-    \caption{Simplified architectures of the reference resolvers}
+    \caption{Simplified architectures of the \emph{coordinate reference resolvers}}
     % SD: How are these encoded?
@@
                                       & $e_i$                 & $c$               & $LSTM_e$  & $LSTM_o$      \\
-                              & $[50, 100, 300, 500, 1000]$ & $[512,1024,2048]$ & $[15,30]$ & $[1000,1500]$ \\\midrule
-        reference resolver    & \times                      & \times            & -         & -             \\
+                                      & $[50, 100, 300, 500]$ & $[512,1024,2048]$ & $[15,30]$ & $[1000,1500]$ \\\midrule
+        coordinate reference resolver & \times                & \times            & -         & -             \\
         + one-hot                     & \times                & \times            & -         & -             \\
@@
 A high accuracy may indicate that the model could identify this specific object better.
+
+In the second approach, the \textbf{attention reference resolver} predicts a region around the center of the target object.
+For this, the image is divided into 14\times14 regions.
+The target area is located around the center of the target object, consisting of 3\times3 regions.
+Hereby, the center of the target object is located in the central region.
+The model is then tasked to predict the matrix $A = (a_{ij})$, where:
+\begin{align*}
+    a_{ij} =
+    \begin{cases}
+        1, & \text{if region $i,j$ in target area} \\
+        0, & \text{otherwise}
+    \end{cases}
+\end{align*}
+
+To do this, the model combines a representation of the image and a representation of the target object.
+More specifically, the image is encoded using the \emph{image encoder} submodule
 
@@
         {500} & {2048} & {5,18}                                    & {99\%}                              & {36,09}                             & {40\%}                                   & {44,81}       & {18\%}        & {51,8}        & {13\%}        \\
-        {1000} & {512}  & {3,11}                                    & {98\%}                              & {37,58}                             & {36\%}                                   & {46,8}        & {19\%}        & {49,95}       & {14\%}        \\
-        {1000} & {1024} & {7,91}                                    & {100\%}                             & {37,78}                             & {39\%}                                   & {47,11}       & {18\%}        & {54,18}       & {11\%}        \\
-        {1000} & {2048} & {4,64}                                    & {100\%}                             & {37,01}                             & {42\%}                                   & {45,72}       & {21\%}        & {49,98}       & {13\%}        \\
         \bottomrule
@@
         {500} & {2048} & {2,86}                                    & {98\%}                              & {35,71}                             & {41\%}                                   & {46,62}       & {18\%}        & {52,67}       & {11\%}        \\
-        {1000} & {512}  & {3,6}                                     & {97\%}                              & {36,37}                             & {36\%}                                   & {46,37}       & {18\%}        & {51,6}        & {12\%}        \\
-        {1000} & {1024} & {2,02}                                    & {94\%}                              & {36,86}                             & {38\%}                                   & {48,35}       & {17\%}        & {51,41}       & {15\%}        \\
-        {1000} & {2048} & {4,49}                                    & {99\%}                              & {36,5}                              & {41\%}                                   & {46,04}       & {18\%}        & {51,22}       & {13\%}        \\
         \bottomrule
@@
         {500} & {2048} & {2,37}                                    & {97\%}                              & {37,33}                             & {35\%}                                   & {48,04}       & {16\%}        & {53,68}       & {11\%}        \\
-        {1000} & {512}  & {1,97}                                    & {93\%}                              & {36,07}                             & {42\%}                                   & {47,34}       & {19\%}        & {51,74}       & {12\%}        \\
-        {1000} & {1024} & {2,99}                                    & {94\%}                              & {36,23}                             & {41\%}                                   & {45,64}       & {19\%}        & {53,27}       & {12\%}        \\
-        {1000} & {2048} & {3}                                       & {98\%}                              & {37,62}                             & {33\%}                                   & {46,54}       & {17\%}        & {52,26}       & {14\%}        \\
         \bottomrule
@@
         {500} & {2048} & {2,67}                              & {95\%}                              & {10,57}                                  & {96\%}        & {4,24}        & {98\%}        \\
-        {1000} & {512}  & {6,59}                              & {95\%}                              & {9,2}                                    & {99\%}        & {7,76}        & {100\%}       \\
-        {1000} & {1024} & {5,85}                              & {99\%}                              & {12,24}                                  & {98\%}        & {8,08}        & {99\%}        \\
-        {1000} & {2048} & {6,03}                              & {99\%}                              & {9,12}                                   & {97\%}        & {14,29}       & {99\%}        \\
         \bottomrule

### 2023-12-07 source/mlt/language_games/run_experiments.py

@@ -3,30 +3,30 @@ import itertools
 import subprocess
 
 options = [
-    "--n_epochs=9",
-    "--validation_freq=9",
-    "--batch_size=32",
-    "--batches_per_epoch=40",
-    "--lr=0.0002",
-    "--feature_file=resnet_3_no-avgpool_no-fc.h5",
-    "--max_samples=10000",
-    "--model=caption_generator",
-    "--mode=gs",
+    "--save",
     "--sender_cell=lstm",
     "--receiver_cell=lstm",
-    "--save",
-    # -- model specifics --
-    "--sender_image_embedding=500",
-    # "--sender_encoder_dim=",
-    "--receiver_image_embedding=100",
-    "--receiver_decoder_out_dim=1500",
-    "--receiver_embedding=500",
-    # "--coordinate_classifier_dimension=",
-    "--temperature=1",
+    "--image_feature_file=resnet_3_no-avgpool_no-fc.h5",
+    "--bounding_box_feature_file=bounding_box_resnet_4_avgpool_no-fc.h5",
 ]
 
 variables = {
+    "--n_epochs": [100],
+    "--batch_size": [32],
+    "--batches_per_epoch": [40],
+    "--validation_freq": [10],
+    "--lr": [0.0002],
+    "--max_samples": [10000],
+    "--temperature": [1],
+    "--model": ["caption_generator"],
     "--dataset": ["dale-2", "dale-5", "colour"],
+    # -- model specifics --
+    "--sender_image_embedding": [500],
+    # "--sender_encoder_dim": [100],
+    "--receiver_image_embedding": [100],
+    "--receiver_decoder_out": [1500],
+    "--receiver_embedding": [500],
+    # "--receiver_coordinate_classifier_dimension": [100],
     "--sender_hidden": [100, 500],
     "--receiver_hidden": [10, 50, 500],
     "--vocab_size": [1, 10, 16, 50, 100],
@@ -60,9 +60,6 @@ if __name__ == "__main__":
         ]
 
     for index, combination in enumerate(itertools.product(*variables.values())):
-        save_appendix = "_".join(
-            str(i) for i in combination if i not in variables.get("--dataset", [])
-        )
         subprocess.run(
             [
                 "python",
@@ -73,7 +70,6 @@ if __name__ == "__main__":
                     f"{option}={value}"
                     for option, value in zip(variables.keys(), combination)
                 ],
-                f"--save_appendix={save_appendix}",
                 *additional,
             ],
             check=False,

### 2023-12-07 thesis/sections/5.3-reference-resolution.tex
experiments

@@
 To do this, the model combines a representation of the image and a representation of the target object.
-More specifically, the image is encoded using the \emph{image encoder} submodule
+More specifically, the image is encoded using the \emph{image encoder} submodule, omitting the last \emph{max pool} layer.
+By this, the resulting matrix has 128\times14\times14 dimension, corresponding to the 14\time14 regions.
+For the target object, the incremental RE is passed through an LSTM as described before.
+Both encodings are passed through two linear layers separately to project them to the projection dimension $p$, followed by the $tanh$ function.
 

### 2023-12-07 source/mlt/language_games/run_experiments.py

@@ -18,17 +18,18 @@ variables = {
     "--lr": [0.0002],
     "--max_samples": [10000],
     "--temperature": [1],
-    "--model": ["caption_generator"],
+    "--model": ["bounding_box_attention_predictor"],
     "--dataset": ["dale-2", "dale-5", "colour"],
     # -- model specifics --
-    "--sender_image_embedding": [500],
+    # "--sender_image_embedding": [500],
     # "--sender_encoder_dim": [100],
-    "--receiver_image_embedding": [100],
-    "--receiver_decoder_out": [1500],
-    "--receiver_embedding": [500],
+    # "--receiver_image_embedding": [100],
+    # "--receiver_decoder_out": [1500],
+    "--receiver_projection": [500],
+    "--sender_embedding": [100],
     # "--receiver_coordinate_classifier_dimension": [100],
-    "--sender_hidden": [100, 500],
-    "--receiver_hidden": [10, 50, 500],
+    "--sender_hidden": [10, 100, 500],
+    "--receiver_hidden": [10, 100, 500],
     "--vocab_size": [1, 10, 16, 50, 100],
     "--max_len": [1, 2, 3, 4, 6],
 }

### 2023-12-12 thesis/sections/5.1-object-identification.tex
results
experiments

@@
 The preceding layers \emph{ResNet-1} to \emph{ResNet-4} couldn't be compared due to memory restrictions on the GPU.
-A deeper analysis of the effects of different feature extractors will be done in the following chapters.
 The resulting vector is flattened and projected to the embedding dimension $e$ via a linear layer.
@@
 
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
+The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$ and 30 epochs.
 The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
-8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
 The loss is calculated using cross entropy.

### 2023-12-12 thesis/sections/5.2-re-generation.tex
experiments

@@
 
-For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
-8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
+For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$ and 30 epochs.
 The loss is calculated using cross entropy.

### 2023-12-12 thesis/sections/5.3-reference-resolution.tex
experiments

@@
 
-For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $1\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
-8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
+For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $1\times10^{-4}$ and 30 epochs.
 The loss is calculated using cross entropy.
@@
 Both encodings are passed through two linear layers separately to project them to the projection dimension $p$, followed by the $tanh$ function.
+Following, the dot product is calculated between the projected image and the projected attributes which results in a 14\times14 matrix.
+This emphasizes the correlation between the attributes and each region.
+A high dot product for a region indicates a high correlation between the attributes and the specific region, while a low dot product indicates the opposite.
+Training a model like this should therefore highlight the regions in the image that are described by the referring expression, namely the region around the target object.
+To calculate the loss, the $softmax$ function is applied over the prediction and compared to the ground truth matrix $A$ using the \emph{binary cross entropy}.
+Hereby, the models converged faster and were trained for 20 epochs with a learning rate of $2\times10^{-4}$.
+An $LSTM_e=15$ and $LSTM_o=1500$ showed the best results.
+Projection sizes $p=\{50,100,300,500,1000\}$ are compared.
+
+The \emph{attention reference resolver} is evaluated on the summed predicted probability for the regions in the target area.
+In particular, the predicted matrix, consisting of probabilities for each region is multiplied with the ground truth matrix $A$, consisting of only ones and zeros.
+The result is summed and returns the probability mass for the target area.
+If the model predicts the target area perfectly, the probabilities in the target area sum to 1.
+If the model for instance focuses on the wrong object, the probability mass in the target area is lower.
 

### 2023-12-12 thesis/sections/6-grounding-language-games.tex
experiments

@@
 Since, it is possible that the agents simply learn to align these two languages, the agents are tasked to point towards the center coordinates of the target object in the final task, the \emph{reference resolution} task.
+For all three setups, the sender uses the same setup and only the receiver is switched.
+By this, differences in the produced referring expression in the message by the sender are only dependent on the receiver's ability to solve the different tasks.
+In particular, the impact of the task on the emerged language can be analyzed in more detail.
 
 As discussed in Section \ref{sec:background:lg:setup}, language games are evaluated on the success of communication and the emergence of a language.
-The testing phase is not necessarily used to test the generalization capabilities of the agents, but more to extract a sample of the emerged language.
+The testing phase is not necessarily used to test the generalization capabilities of the agents, but rather to extract a sample of the emerged language.
 It is therefore not as important that the samples in the testing phase are unseen samples.
@@
 In these experiments, the dataset consists of 10.000 scenes in total.
-To train the agents, 12.800 games are played (using 12.800 randomly drawn samples from the dataset) with a batch size of 32 samples.
+To train the agents, 128.000 games are played (using 128.000 randomly drawn samples from the dataset) with a batch size of 32 samples.
 To retrieve the samples of the emerged language, after the agents are trained, they are frozen and play 2.048 games with randomly drawn samples from the dataset.

### 2023-12-12 thesis/sections/6.1-object-identification.tex
experiments

@@
 One central difference is the production of the message.
-The main goal of their language game was the identification of the concept that and image was related to.
-Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
-Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
+The main goal of their language games was the identification of the concept that the shown image was related to.
+Therefore, the sender communicated only single-symbol messages to the receiver which should describe the concept of the target image.
+Opposed to that, in this research, the agents are tasked to discriminate objects from each other based on their attributes.
 % SD: On the contrary,...
-% DK: TODO
+% DK: (done)
 It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message.
+For that reason, the sender is allowed to generate sequences as a message where for instance each symbol in the sequence might correspond to on attribute.
 
@@
 
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
-The loss is calculated using cross entropy.
+The experiments are conducted with a learning rate of $2\times10^{-4}$, the loss is calculated using cross entropy.
 The following values for the variables are compared:

### 2023-12-12 source/mlt/language_games/data_readers.py

@@ -571,6 +571,53 @@ class CoordinatePredictorGameBatchIterator(GameBatchIterator):
         )
 
 
+class BoundingBoxCoordinatePredictorGameBatchIterator(GameBatchIterator):
+    def __init__(self, loader, batch_size, n_batches, train_mode, seed) -> None:
+        self.loader = loader
+        self.batch_size = batch_size
+        self.n_batches = n_batches
+        self.batches_generated = 0
+        self.train_mode = train_mode
+        self.random_seed = random.Random(seed)
+
+    def __next__(self):
+        if self.batches_generated > self.n_batches:
+            raise StopIteration()
+
+        batch_data = self.get_batch()
+        self.batches_generated += 1
+        return batch_data
+
+    def get_batch(self):
+        sampled_indices = self.random_seed.sample(
+            range(len(self.loader.dataset)), self.batch_size
+        )
+        samples: list[CoordinatePredictorSample] = [
+            self.loader.dataset[i] for i in sampled_indices
+        ]
+
+        sender_inputs = []
+        targets = []
+        receiver_inputs = []
+        image_ids = []
+
+        for sample in samples:
+            sender_inputs.append(sample.bounding_boxes)
+            targets.append(sample.target_pixels)
+
+            receiver_inputs.append(sample.image)
+            image_ids.append(int(sample.image_id[-6:]))
+
+        return (
+            torch.stack(sender_inputs),
+            torch.stack(targets),
+            torch.stack(receiver_inputs),
+            {
+                "image_id": torch.tensor(image_ids),
+            },
+        )
+
+
 class AttentionPredictorGameBatchIterator(GameBatchIterator):
     def __init__(self, loader, batch_size, n_batches, train_mode, seed) -> None:
         self.loader = loader

### 2023-12-12 source/mlt/language_games/play.py

@@ -19,6 +19,7 @@ from mlt.language_games.data_readers import (
     AttentionPredictorGameBatchIterator,
     BoundingBoxAttentionPredictorGameBatchIterator,
     BoundingBoxCaptionGeneratorGameBatchIterator,
+    BoundingBoxCoordinatePredictorGameBatchIterator,
     CaptionGeneratorGameBatchIterator,
     CaptionGeneratorGameDataset,
     CoordinatePredictorGameBatchIterator,
@@ -340,6 +341,24 @@ models = {
         },
         loss_function=attention_loss,
     ),
+    "bounding_box_coordinate_predictor": ModelDefinition(
+        dataset=CoordinatePredictorGameDataset,
+        dataset_args={},
+        split_dataset=False,
+        image_loader=FeatureImageLoader,
+        bounding_box_loader=FeatureImageLoader,
+        iterator=BoundingBoxCoordinatePredictorGameBatchIterator,
+        sender=ReferentialGameSender,
+        sender_args={},
+        receiver=CoordinatePredictorReceiver,
+        receiver_args={
+            "image_encoder": ClevrImageEncoder(
+                feature_extractor=DummyFeatureExtractor(), max_pool=True
+            ),
+            "coordinate_classifier": CoordinateClassifier,
+        },
+        loss_function=pixel_loss,
+    ),
     "bounding_box_attention_predictor": ModelDefinition(
         dataset=CoordinatePredictorGameDataset,
         dataset_args={
@@ -469,11 +488,6 @@ def get_params(params):
         help="Size of the LSTM encoder of Sender when attributes are encoded with descriptions (default: 10)",
     )
     parser.add_argument(
-        "--sender_coordinate_classifier",
-        type=int,
-        help="Dimensions for the coordinate predictor for Receiver (default: 10)",
-    )
-    parser.add_argument(
         "--sender_projection",
         type=int,
         help="Projection dimension to combin message and image (default: 10)",
@@ -512,6 +526,11 @@ def get_params(params):
     )
     parser.add_argument(
         "--receiver_projection",
+        type=int,
+        help="Projection dimension to combin message and image (default: 10)",
+    )
+    parser.add_argument(
+        "--receiver_coordinate_classifier_dimension",
         type=int,
         help="Projection dimension to combin message and image (default: 10)",
     )
@@ -679,7 +698,9 @@ def main(params):
         )
     if "coordinate_classifier" in receiver_args.keys():
[…]

### 2023-12-12 source/mlt/language_games/run_experiments.py

@@ -18,19 +18,20 @@ variables = {
     "--lr": [0.0002],
     "--max_samples": [10000],
     "--temperature": [1],
-    "--model": ["bounding_box_attention_predictor"],
+    "--model": ["discriminator"],
     "--dataset": ["dale-2", "dale-5", "colour"],
     # -- model specifics --
-    # "--sender_image_embedding": [500],
-    # "--sender_encoder_dim": [100],
-    # "--receiver_image_embedding": [100],
-    # "--receiver_decoder_out": [1500],
-    "--receiver_projection": [500],
     "--sender_embedding": [100],
-    # "--receiver_coordinate_classifier_dimension": [100],
-    "--sender_hidden": [10, 100, 500],
-    "--receiver_hidden": [10, 100, 500],
-    "--vocab_size": [1, 10, 16, 50, 100],
+    "--sender_hidden": [100, 500, 1000],
+    "--sender_image_embedding": [500],
+    "--receiver_embedding": [100],
+    "--receiver_hidden": [100, 500, 1000],
+    # "--receiver_projection": [100],
+    "--receiver_image_embedding": [500],
+    # "--receiver_decoder_out": [1500],
+    # "--receiver_decoder_embedding": [15],
+    # "--receiver_coordinate_classifier_dimension": [2048],
+    "--vocab_size": [2, 10, 16, 50, 100],
     "--max_len": [1, 2, 3, 4, 6],
 }
 

### 2023-12-12 source/mlt/print_results.py

@@ -1,4 +1,5 @@
 import argparse
+import ast
 import glob
 import json
 import os
@@ -18,6 +19,11 @@ class Run:
         self.epochs = []
         with open(file, "r", encoding="utf-8") as f:
             for line in f:
+                if line.startswith("appendix: "):
+                    self.variable_names = ast.literal_eval(
+                        line.removeprefix("appendix: ")
+                    )
+
                 if line.startswith("{"):
                     epoch = json.loads(line)
                     if epoch.get("mode", "test") == "train":
@@ -137,19 +143,30 @@ class Run:
         return set(self.last_epoch.keys())
 
 
+variable_dict = {
+    "max_len": "$n$",
+    "receiver_embedding": "$e_r$",
+    "receiver_hidden": "$h_r$",
+    "receiver_projection": "$p$",
+    "sender_embedding": "$e_s$",
+    "sender_hidden": "$h_s$",
+    "sender_image_embedding": "$e_i$",
+    "vocab_size": "$|V|$",
+}
+
+
 @dataclass
 class Experiment:
     folder_name: str
     dataset: str
-    variables: tuple
+    variables: dict
     run: Run
 
 
 class Experiments:
-    def __init__(self, root_folder, datasets, variables, variable_order) -> None:
+    def __init__(self, root_folder, datasets) -> None:
         self.datasets = datasets
-        self.variables = variables
-        self.variable_order = variable_order
+        self.chosen_variables = tuple()
 
         self.experiments: list[Experiment] = []
         for folder in glob.glob(root_folder):
@@ -161,17 +178,19 @@ class Experiments:
             if dataset not in datasets:
                 continue
 
-            if len(experiment_variables) != len(variables):
-                raise ValueError(f"Number of Variables don't match: {folder_name}")
-            ordered_variables = tuple(experiment_variables[o] for o in variable_order)
-
             run = Run(os.path.join(folder, "log.txt"))
+
             if len(run.epochs) != 0:
                 self.experiments.append(
                     Experiment(
                         folder_name=folder_name,
                         dataset=dataset,
-                        variables=ordered_variables,
+                        variables={
+                            variable_dict[variable_name]: variable
+                            for variable_name, variable in zip(
+                                run.variable_names, experiment_variables
+                            )
+                        },
                         run=run,
                     )
                 )
@@ -186,13 +205,6 @@ class Experiments:
             ]
 
         return dataset, experiment_variables
-
-    def _get_excluded_folders(self):
-        for index, folder in enumerate(
[…]

### 2023-12-12 source/mlt/print_results.py
print

@@ -148,9 +148,10 @@ variable_dict = {
     "receiver_embedding": "$e_r$",
     "receiver_hidden": "$h_r$",
     "receiver_projection": "$p$",
+    "receiver_image_embedding": "$e_{ri}$",
     "sender_embedding": "$e_s$",
     "sender_hidden": "$h_s$",
-    "sender_image_embedding": "$e_i$",
+    "sender_image_embedding": "$e_{si}$",
     "vocab_size": "$|V|$",
 }
 

### 2024-02-27 thesis/sections/6-grounding-language-games.tex
thesis

@@
 The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.
+However, every task is solved by two neural networks that communicate through a discrete bottleneck.
 % SD: But there is an importnat difference that now we have two neural netweorks, the sender and the receiver.
@@
 
+Across all experiments, the same restriction on the vocabulary and the produced messages are compared.
+First, five different vocabulary sizes $|V|$ are tested: 2, 10, 16, 50 and 100 symbols.
+Notice that these vocabularies include the \emph{<eos>} token which signals the end of the message.
+A vocabulary of two symbols therefore consists effectively of only one token that can be learned by the agents.
+The center of attention lies on the vocabularies with 16 and 50 symbols.
+The vocabulary with 16 symbols corresponds the 13 attributes the objects can have and align with human language referring expressions.
+However, a smaller vocabulary size increases the pressure to converge on meaning, which is my also 10 and 2 symbols are tested.
+On the other side, the agents might instead of a compositional language learn to assign one symbol with every combination of attributes.
+This results in $2 * 3 * 8 = 42$ unique combinations, which corresponds to the vocabulary with 50 symbols.
+Furthermore, a vocabulary of 100 symbols is used to test the effect of a much bigger bottleneck.
+
+Secondly, also different maximum message lengths are compared: 1, 2, 3, 4 and 6 symbols.
+Notice that this applies only to the maximum possible length; the agents can choose to use shorter messages.
+Again, the corresponding length to human referring expressions in English is 3 for the three attributes.
+However, the agents can be pressured to condense the information more precisely with shorter messages or given less restrictions with longer messages.
+Vocabulary size and message length play hereby an opposing role in the pressure on condensing information.
+% If the symbols were similarly used, messages would have lengths between one and three symbols.
+% Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
+% Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
+% Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
+% SD: Motivation why this is so. The possible number of strings is {#colour} x {#shape} {object type} where all numbers except the last one also contain a padding token.
+% DK: TODO
+
 All agents are trained using the \emph{Adam} optimizer \citep{Kingma2015}.
@@
 Hereby, a temperature of $\tau=1$ is used which showed the best results in preliminary experiments.
+All setups are additionally compared to a baseline.
+In the baseline the sender is "blinded", meaning that the sender is not shown any input but produces random messages.
+Effectively, the receiver needs to solve the task on their own, by exploiting underlying structures in the shown data and the target data.
+If a non-baseline model doesn't achieve a better performance than the baseline, the sender didn't communicate any meaningful messages and no language emerged.

### 2024-02-27 thesis/sections/6.1-object-identification.tex
thesis

@@
 \subsection{Object identification}
+\label{sec:language-games:identification}
+
 \subsubsection*{Setup}
@@
 It is therefore assumed that the sender will communicate these discriminative attributes.
-For that reason, the sender is allowed to generate sequences as a message where for instance each symbol in the sequence might correspond to on attribute.
+For that reason, the sender is allowed to generate sequences as a message where for instance each symbol in the sequence might correspond to one attribute.
 
@@
 Bounding boxes of all objects in the scene are extracted and preprocessed to be used for the feature extractors.
+For datasets with a variable number of objects, the list bounding of bounding boxes is padded with a matrix of zeros to the maximum possible number of objects present in a scene across the
+dataset.
+Opposed to the experiments with one neural model, no additional information about the target object's attributes is provided.
+Instead, the sender agent is presented with the target object always as the first object, while the distractors are shuffled.
+The receiver is presented with completely shuffled bounding boxes.
+
 Accordingly, also the models from the single model tasks serve as basis for the object identification task.
 The sender uses the architecture of the model described in section \ref{sec:referring_expression_generation} apart from the final LSTM.
-Instead, it will use the LSTM with the hidden size $h_s$ included in the \emph{EGG} framework to produce a message.
+Instead, it will use the LSTM with a hidden size $h_s$ included in the \emph{EGG} framework to produce a message.
 Based on the results from the single model object identifiers, the image embedding size $e_s$ is fixed to 100.
-The bounding box of the target object is passed first, the remaining are shuffled.
 
@@
 The hidden state of the LSTM is projected to the embedding size $e_r=500$ with a linear layer, based on the results in the previous experiments.
-As before the dot product between the embedded message and the encoded image is calculated to let the model point towards one of the bounding boxes.
-Opposed to the sender, the receiver is shown all bounding boxes in shuffled order.
+As before the dot product between the embedded message and the encoded image is calculated to let the model point towards the bounding box that is described best by the message.
 Figure \ref{fig:discriminator_architecture} shows, how the sender and the receiver of the discriminator are built up.
@@
     \item $h_r$: 10, 30, 50, 100, 500, 1000
-    \item $|V|$: 1, 10, 16, 50, 100
-    \item $n$: 1, 2, 3, 4, 5, 6
+    \item $|V|$: 2, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 6
 \end{itemize}
+
 The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
+A random guess corresponds to 50\% in the \emph{Dale-2} dataset, 20\% in the \emph{Dale-5} dataset and 13\% in the \emph{CLEVR color} dataset.
 
 \subsubsection*{Results}
+When looking at the results of the experiments, it can be seen that the agents are successfully discriminating the objects across all datasets.
+Table TODO gives an overview over the accuracy scores of selected configurations.
+For each dataset there are many configurations that outperform the baseline models by a high margin.
+However, there are still clear differences between the accuracy scores of each dataset.
+
+When the agents are tasked to discriminate between the two objects of the \emph{Dale-2} dataset, it succeeds in almost all configurations with an accuracy of over 94\%.
+In many configurations the accuracy lies above 99\%, while only 5 configurations have an accuracy of only 61\% to 64\%.
+An indicator for
+
 Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.
-Four different vocabulary sizes $|V|$ are tested.
-A size of 13 symbols corresponds to the 13 attributes the objects can have and align with human language.
-If the symbols were similarly used, messages would have lengths between one and three symbols.
-Opposed to that a slightly smaller vocabulary with 10 symbols is used to create a smaller bottleneck with a higher pressure to condense the information.
-Similarly, a bigger vocabulary consisting of 20 symbols tests how a bigger bottleneck changes the results.
-Lastly, a big vocabulary of 100 symbols should give the model all options to encode the information, including one symbol per attribute or one symbol describing a combination of attributes.
-% SD: Motivation why this is so. The possible number of strings is {#colour} x {#shape} {object type} where all numbers except the last one also contain a padding token.
-% DK: TODO
 

### 2024-02-27 thesis/sections/6.2-re-generation.tex
thesis

@@
 In a next step, it is tested if the agents can learn to extract features of the objects together.
-For this, the receiver is tasked to describe the target object in natural language, while the sender needs to communicate, which object is the target.
-Again, the setup is asymmetrical: the sender receives the image and information, which of the objects in the image is the target object in form of a masked image.
-The receiver only sees the image without additional information.
-\cmtDK[inline]{human bias vs emergent language}
-
-For this experiment, the data is prepared in the same way as described in section \ref{sec:referring_expression_generation}.
-Since the position of the padding as well as the order of the words didn't have an effect on the results, the usual approach in natural generation task is used: padding tokens are appended and the referring expression is not reversed.
-Respectively, both agents are based on the models in these experiments.
-The sender uses the \emph{masked RE generator} with its LSTM replaced with the LSTM included in the \emph{EGG} framework, using the hidden size $h_s$.
-\cmtDK[inline]{fixed parameters}
-The receiver is built up similarly.
-The image is encoded in the same way, but afterwards it is concatenated with the decoded message, received be the sender with the hidden size $h_r$.
-The concatenated is reduced to $d=$ dimensions, based on the results of the previous experiments, and used as the initial state of the captioning LSTM.
-\cmtDK[inline]{fixed parameters}
+For this the same sender as in the previous section is used.
+Based on the results, $h_s$ is fixed to $500$ and $e_s$ to $100$
+However, the task and by that the receiver's architecture is adapted in two different ways.
+In the first setup, the \textbf{referring expression generator}, the receiver is tasked to describe the target object with natural language following the \emph{incremental GRE-algorithm} based on the sender's message (see Figure \ref{fig:caption_generator_game_architecture}).
+This follows the approach of the approach of the \emph{RE generators} described in section \ref{sec:referring_expression_generation}.
+The whole scene is encoded using the \emph{image encoder} submodule, projecting it to the image encoding dimension $e_{ri}=100$.
+The sender's message is decoded with the hidden size $h_r=500$ and $e_r=100$, and concatenated with the encoded image.
+This is then reduced to $LSTM_o=1500$ dimensions based on the results of the previous experiments, and used as the initial state of the captioning LSTM.
+Tokens are embedded with $LSTM_e=15$ dimensions.
+Since the position of the padding as well as the order of the words in the target didn't have an effect on the results, the usual approach in natural generation task is used: padding tokens are appended and the referring expression is not reversed.
+During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
+% SD: I'm not sure how teacher forcing is implemented, see my earlier comment
+% DK: typo, teacher forcing is actually applied (done)
+When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
+The loss is calculated using cross entropy.
 
@@
 
-During training, the ground truth caption is used as the input to the LSTM using teacher forcing.
-% SD: I'm not sure how teacher forcing is implemented, see my earlier comment
-% DK: typo, teacher forcing is actually applied (done)
-When presented with test data, the LSTM always produces three tokens, by using its own predicted words as the input for the next step.
+However, this setup could lead to two problems:
+First, the task is quite complex for both agents to learn, since two language generation steps are involved.
+This might stop the agents from converging towards a useful language.
+Secondly, in case a language emerges, it could be aligned with the natural language of the target referring expressions.
+In other words, the sender might just learn to effectively repeat the target referring expressions without extracting the necessary features themselves.
+For that reason, in the second setup the receiver is tasked to predict only the attributes of the target object in the shape of a one-hot vector.
+By this, no natural language including its salience order is involved and the complexity of the complete game is reduced.
+The receiver uses the same way to encode and combine the complete scene and the sender's message with $e_{ri}=100$, $h_r=500$ and $e_r=100$.
+This is then passed to a linear layer that produces a vector with 13 dimension, corresponding to the 2 sizes, 3 shapes, and 8 colors.
+The loss is then calculated using binary cross entropy.
 
-The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
-The loss is calculated using cross entropy.
-The following values for the variables are compared:
+\cmtDK[inline]{figure}
+
+The experiments for both setups are conducted with a learning rate of $2\times10^{-4}$.
+As in the previous section, the following values for the variables are compared:
 \begin{itemize}
-    \item $h_s$: 100, 500, 1000
-    \item $h_r$: 10, 30, 50, 100, 500, 1000
-    \item $|V|$: 1, 10, 16, 50, 100
-    \item $n$: 1, 2, 3, 4, 5, 6
+    \item $|V|$: 2, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4, 6
 \end{itemize}
-\cmtDK[inline]{real values}
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2' and 'Dale-5' dataset.
-The 'Dale-5' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
+
+Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} dataset.
+The \emph{Dale-5} and the \emph{CLEVR color} should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
 % SD: So the sender is generating descriptions following the GRE policy?
 % DK: TODO
-The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results.
+The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results for the first setup.
+When predicting one-hot vectors, an overall \textbf{accuracy} is reported that reflects if all attributes were predicted correctly.
+Additionally, the accuracy for each attribute is calculated separately to show which attributes give bigger challenges for the models.
 

### 2024-02-27 thesis/sections/6.3-reference-resolution.tex
thesis

@@
 \subsubsection*{Setup}
-In the final experiments, the agents are tasked to produced and understand referring expressions only based on the implicit human knowledge present in the scenes as for example the attributes of the objects.
+In the final experiments, the agents are tasked to produced and understand referring expressions only based on the implicit human knowledge present in the scenes, as for example the attributes of the objects.
 Opposed to the previous experiments, explicit human language information as human referring expressions or one-hot encoded attributes are avoided.
@@
 This is done similarly to the single model reference resolution task, described in section \ref{sec:reference_resolution}.
-The agents are tasked to predict the center coordinates of the target object, after being presented with the complete scene.
-However, the agents, more specifically the sender needs to be pointed towards the target object.
-In the final setup, the \emph{masked reference resolution game}, this is done providing the sender with an additional masked visual input.
-However, this can prove to be a challenging task for the agents.
-For this reason, in an approach to reduce the complexity, for the \emph{RE reference resolution game} the sender is shown a human referring expression, generated with the incremental GRE algorithm instead of the additional masked image.
-This might lead to a heavy bias of the emerged language towards the human referring expressions and will be compared to the unbiased emerged language.
+As before, the sender has the same architecture as in the previous language games, encoding the bounding boxes of all objects.
 
-The masked sender is the same as the \emph{RE generation sender}, described in section \ref{sec:referring-expression-generation-game}.
-The RE sender works similarly, but instead encodes the human referring expression with an LSTM.
-The final hidden state is concatenated with the encoded image.
-The resulting vector is reduced to the hidden size $h_s$ and used as the initial hidden state of the LSTM included in the \emph{EGG} framework to generate the message.
-\cmtDK[inline]{fixed variables}
-
+The receiver is tasked to solve two tasks.
+First, in the \textbf{coordinate reference resolver}, the agents are tasked to predict the center coordinates of the target object (see Figure \ref{fig:coordinate_predictor_game_architecture}).
 The receiver is the same in both setups and is based on the \emph{reference resolver} + referring expressions, described in section \ref{sec:reference_resolution}.
-Again, instead of the included LSTM, the receiver uses the \emph{EGG} LSTM with the hidden size $h_r$ to encode the message.
-The dimensions of the coordinate predictor are fixed to $c=$, based on the previous results.
-\cmtDK[inline]{fixed variables}
+Again, instead of the included LSTM, the receiver uses the \emph{EGG} LSTM with the hidden size $h_r=500$ and $e_r=100$ to encode the message.
+The dimensions of the coordinate predictor are fixed to $c=2048$, based on the previous results.
 % SD: Just thinking: would there be a more systematic way to compare all the models?
@@
 
+Secondly, in the \textbf{attention reference resolver}, the agents learn to point towards regions in the image, which is based on the \emph{attention reference resolver} described in section \ref{sec:reference_resolution}.
+As before, the LSTM is replaced with the \emph{EGG} LSTM to parse the sender's message with the hidden size $h_r=500$ and $e_r=100$.
+The dot product is calculated between each encoded region of the image and the encoded message, and the $softmax$ function is applied subsequently.
+The loss is calculated using \emph{binary cross entropy} between the predictions and ground truth regions.
+
+\cmtDK[inline]{figure}
+
 The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
-The loss is calculated using cross entropy.
 The following values for the variables are compared:
 \begin{itemize}
-    \item $h_s$: 100, 500, 1000
-    \item $h_r$: 10, 30, 50, 100, 500, 1000
-    \item $|V|$: 1, 10, 16, 50, 100
-    \item $n$: 1, 2, 3, 4, 5, 6
+    \item $|V|$: 2, 10, 16, 50, 100
+    \item $n$: 1, 2, 3, 4,  6
 \end{itemize}
-\cmtDK[inline]{real values}
 
-As in section \ref{sec:reference_resolution}, the agents are trained first on the 'CLEVR single' dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2' and 'Dale-5' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
-The same metrics are used to evaluate the results.
+As in section \ref{sec:reference_resolution}, the agents are trained first on the \emph{CLEVR single} dataset to understand if they are capable of predicting locations in an image together.
+In a next step, the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
+The same metrics are used to evaluate the results as described in \ref{sec:reference_resolution}.
 

### 2024-02-27 source/mlt/print_results.py

@@ -151,6 +151,7 @@ variable_dict = {
     "receiver_image_embedding": "$e_{ri}$",
     "receiver_decoder_embedding": "$LSTM_e$",
     "receiver_decoder_out": "$LSTM_o$",
+    "receiver_coordinate_classifier_dimension": "$c$",
     "sender_embedding": "$e_s$",
     "sender_hidden": "$h_s$",
     "sender_image_embedding": "$e_{si}$",
@@ -312,8 +313,7 @@ class OutputProcessor:
         }
 
     @abstractmethod
-    def print(self):
-        ...
+    def print(self): ...
 
 
 class SortedOutputProcessor(OutputProcessor):

### 2024-03-10 thesis/sections/6.1-object-identification.tex
results

@@
     \item $h_s$: 100, 500, 1000
-    \item $h_r$: 10, 30, 50, 100, 500, 1000
+    \item $h_r$: 100, 500, 1000
     \item $|V|$: 2, 10, 16, 50, 100
@@
 When looking at the results of the experiments, it can be seen that the agents are successfully discriminating the objects across all datasets.
-Table TODO gives an overview over the accuracy scores of selected configurations.
+Tables \ref{tab:results_discriminator_dale-2} to \ref{tab:results_discriminator_color} give an overview over the accuracy scores of selected configurations.
 For each dataset there are many configurations that outperform the baseline models by a high margin.
@@
 
-When the agents are tasked to discriminate between the two objects of the \emph{Dale-2} dataset, it succeeds in almost all configurations with an accuracy of over 94\%.
-In many configurations the accuracy lies above 99\%, while only 5 configurations have an accuracy of only 61\% to 64\%.
-An indicator for
-
-Table \ref{tab:results_discriminator} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
-
-The hidden sizes $h_s$ and $h_r$ as well as the embedding sizes $e_s$ and $e_r$ are chosen in alignment with the vocabulary size.
-The hidden sizes are always smaller or equal to the vocabulary size since the information about each word needs to be compressed in a smaller dimension to learn meaning.
-Hereby, hidden sizes of 10 and 100 are tested.
-On the other hand, three different embedding sizes are tested: 10, 50 and 100.
-The reason for this is to experiment, what is the optimal middle ground between compressing features of an image encoded in high dimension vectors and upscaling encoded messages in low dimension vectors.
-% SD: This discussion of dimensions could already be introduced when you discuss simple modles earlier as the same dimensions are used.
-This is a more an evaluation of represnetations rather than interaction and hence it would fit there.
+\cmtDK[inline]{baseline}
+\cmtDK[inline]{learning curve}
+% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
+As David pointed out it may be that convergence is unstable and changes happen in between.
 % DK: TODO
@@
     \centering
-    \begin{tabular}{ccccc|cc|cc|cc}
+    \begin{tabular}{cccc|c}
         \toprule
-              &         &         &         &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                         \\\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & $e_{r}$ & \textbf{Accuracy}                   & \textbf{length}                     & \textbf{Accuracy}                        & \textbf{length} & \textbf{Accuracy} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10}    & {10}    & {96,4\%}                            & {0,99}                              & {24,7\%}                                 & {0}             & {17,3\%}          & {1}             \\
-        {10}  & {50}    & {50}    & {50}    & {50}    & {50\%}                              & {1}                                 & {21,4\%}                                 & {1}             & {17,8\%}          & {0}             \\
-        {13}  & {10}    & {10}    & {10}    & {10}    & {96,16\%}                           & {1}                                 & {24,8\%}                                 & {1}             & {17,1\%}          & {1}             \\
-        {13}  & {10}    & {10}    & {50}    & {50}    & {49,6\%}                            & {1}                                 & {21,9\%}                                 & {1}             & {17,9\%}          & {0}             \\
-        {20}  & {10}    & {10}    & {50}    & {50}    & {50,9\%}                            & {0}                                 & {23\%}                                   & {1}             & {15,9\%}          & {1}             \\
-        {100} & {10}    & {10}    & {10}    & {10}    & {97,3\%}                            & {1}                                 & {24\%}                                   & {1}             & {18,1\%}          & {1}             \\
-        {100} & {10}    & {10}    & {50}    & {50}    & {49,9\%}                            & {1}                                 & {24,4\%}                                 & {1}             & {15,8\%}          & {1}             \\
-        {100} & {100}   & {100}   & {100}   & {100}   & {49\%}                              & {0}                                 & {25,3\%}                                 & {1}             & {15,6\%}          & {0}             \\
+               &        &     &       & \multicolumn{1}{c}{\textbf{Dale-2}} \\  \cmidrule(lr){5-5}
+        $h_r$  & $h_s$  & $n$ & $|V|$ & \textbf{Accuracy}                   \\\midrule
+        {100}  & {100}  & {2} & {10}  & {99,96\%}                           \\
+        {100}  & {1000} & {2} & {10}  & {99,96\%}                           \\
+        {1000} & {100}  & {3} & {10}  & {99,87\%}                           \\
+        {500}  & {100}  & {3} & {100} & {99,83\%}                           \\
+        {100}  & {100}  & {6} & {50}  & {99,83\%}                           \\
+        {100}  & {500}  & {3} & {16}  & {99,7\%}                            \\
+        {1000} & {100}  & {6} & {16}  & {99,61\%}                           \\
+        {100}  & {100}  & {6} & {100} & {99,57\%}                           \\
+        {100}  & {500}  & {1} & {100} & {99,52\%}                           \\
+        {100}  & {100}  & {1} & {50}  & {99,35\%}                           \\
+        {500}  & {100}  & {1} & {2}   & {99,26\%}                           \\
+        {100}  & {500}  & {1} & {50}  & {99,26\%}                           \\
+        {500}  & {500}  & {1} & {100} & {99,26\%}                           \\
+        {100}  & {500}  & {3} & {100} & {99,83\%}                           \\
+        {500}  & {1000} & {1} & {16}  & {98\%}                              \\
+        {100}  & {500}  & {1} & {10}  & {96,61\%}                           \\
+        {100}  & {1000} & {1} & {100} & {99,44\%}                           \\
+        {500}  & {500}  & {6} & {2}   & {95,92\%}                           \\
+        {100}  & {1000} & {4} & {2}   & {94,88\%}                           \\
+        {100}  & {1000} & {6} & {10}  & {63,15\%}                           \\
+        {1000} & {500}  & {2} & {2}   & {62,98\%}                           \\\midrule
+        {-}    & {-}    & {-} & {-}   & {62,72\% (baseline)}                \\\midrule
+        {100}  & {1000} & {3} & {2}   & {61,2\%}                            \\
         \bottomrule
     \end{tabular}
-    \caption{Results of the discriminators: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    % SD: A note about the datasets chosen. Why do we choose these? Initially, we describe more than 3 datasets.
-    % DK: because the 'CLEVR Single' contains only one object. So there is nothing to discriminate (done)
-    \label{tab:results_discriminator}
+    \caption{Selected results of the object identification games on the Dale-2 dataset with $e_s=100$ and $e_r=500$}
+    \label{tab:results_discriminator_dale-2}
 \end{table}
-% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
-As David pointed out it may be that convergence is unstable and changes happen in between.
-% DK: TODO
 
-For the \emph{Dale-2}, a clear correlation between the hidden sizes, embedding sizes and the size of the vocabulary can be identified.
-A hidden/embedding size as high as the vocabulary size is beneficial for identifying the correct object.
-The receiver identifies almost every sample correctly when all sizes are 10.
-When the hidden and embedding sizes are increased, the guesses by the receiver are random with 50\% accuracy.
-Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message when the model is trained on the \emph{Dale-2} dataset.
-% SD: How do we know? What would happen if the hidden layers are smaller than the vocabulary size? We would also need to test several smaller dimensions.
-% DK: because the task is solved succesfully, TODO
+When the agents are tasked to discriminate between the two objects of the \emph{Dale-2} dataset, it succeeds in almost all configurations with an accuracy of over 94\%.
+Selected results are listed in Table \ref{tab:results_discriminator_dale-2}.
+In many configurations the accuracy lies above 99\%, while only 4 configurations have an accuracy of only 61\% to 64\%.
+An antecedent for a low accuracy seems to be a low vocabulary size $|V|$.
+All experiments with a vocabulary size of only two symbols (including the <eos> symbol) tend to give worse results.
+Three of the four results that don't surpass the baseline use a vocabulary size of 2.
+On the other side, only few configurations with $|V| = 2$ are among the overall best results.
+When the vocabulary size is increased, the results become better.
+Especially configurations with $|V| = 10$, $|V| = 16$, $|V| = 50$ and $|V| = 100$ consistently achieve high accuracies.
+While there are some few exceptions, the results show that the agents tend to perform better, when they can exchange messages with more than one symbol.
+Hereby, even games with an $n = 6$ perform as well as games with $n = 2$ as seen for example for $|V| = 10$.
+However, when only one symbol is allowed in the exchange, the accuracy still surpasses the baseline by a big margin and configurations with a high vocabulary of $|V| = 50$ and $|V| = 100$ tend to achieve higher accuracies.
+Finally looking at the hidden sizes, no direct effect can be identified and the differences in the results seem to be explained by vocabulary size and message length.
 
-The results change, when using the \emph{Dale-5} dataset with four distractors.
-With four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
-Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% points to 43\%.
-This is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
-The same applies to the 'CLEVR color' dataset, where all models achieve a very low accuracy of around 15 to 17\%, corresponding to random guesses.
-% SD: Could we say that calculation on Dale-5 is successful?
-% DK: TODO
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cccc|c}
+        \toprule
+               &        &     &       & \multicolumn{1}{c}{\textbf{Dale-5}} \\  \cmidrule(lr){5-5}
+        $h_r$  & $h_s$  & $n$ & $|V|$ & \textbf{Accuracy}                   \\\midrule
+        {1000} & {500}  & {3} & {50}  & {91,75\%}                           \\
+        {100}  & {100}  & {6} & {10}  & {91,02\%}                           \\
+        {100}  & {100}  & {3} & {10}  & {90,76\%}                           \\
+        {100}  & {100}  & {4} & {16}  & {89,76\%}                           \\
+        {100}  & {100}  & {2} & {10}  & {88,85\%}                           \\
+        {500}  & {1000} & {1} & {100} & {87,11\%}                           \\
+        {1000} & {1000} & {6} & {50}  & {85,07\%}                           \\
+        {500}  & {1000} & {1} & {10}  & {84,46\%}                           \\
+        {500}  & {500}  & {1} & {50}  & {83,29\%}                           \\
+        {100}  & {100}  & {1} & {10}  & {80,86\%}                           \\
+        {500}  & {100}  & {1} & {100} & {76,74\%}                           \\
+        {500}  & {500}  & {4} & {2}   & {77,78\%}                           \\
+        {1000} & {1000} & {3} & {2}   & {75,78\%}                           \\
+        {100}  & {100}  & {3} & {2}   & {74\%}                              \\
+        {500}  & {500}  & {1} & {2}   & {63,41\%}                           \\
+        {500}  & {1000} & {2} & {100} & {34,98\%}                           \\
+        {500}  & {500}  & {3} & {16}  & {34,77\%}                           \\
+        {500}  & {500}  & {6} & {2}   & {34,38\%}                           \\
+        {1000} & {1000} & {4} & {2}   & {35,33\%}                           \\
+        {1000} & {1000} & {6} & {100} & {35,24\%}                           \\
+        \bottomrule
+    \end{tabular}
+    \caption{Selected results of the object identification games on the Dale-5 dataset with $e_s=100$ and $e_r=500$}
+    \label{tab:results_discriminator_dale-5}
+\end{table}
 
-Two conclusions can be drawn.
-First, the hidden as well as the embedding sizes need to be close to the vocabulary size.
-% SD: Not sure, since we never tested smaller layer sizes.
-If the embedding layers are the same as the vocabulary size then at least a language encoding LSTM is not motivated to geenralise at all as it can keep representations for individual words without abstracting them.
-Leanring word embeddings involves reducing dimensions.
-% DK: TODO
-This even applies for very low vocabulary sizes, which means that the image encodings need to be compressed to the same low dimensions.
-The reason for this is very likely that neural models have difficulties to upscale from lower dimensions (e.g.
-from low $h_r$ to high $e_r$) as opposed to learn how to extract the important information from a vector with many dimensions.
-% SD: Hypothesis space to find the target function is just too big and they can never approximate that function.
-% DK: TODO
+When the agents are presented with the \emph{Dale-5} dataset, the results are less clear (see Table \ref{tab:results_discriminator_dale-5}).
+Again, two groups can be identified: One group of configurations doesn't surpass the baseline of 35\% accuracy, while the other group achieves accuracies in between 62\% and 94\%.
+For this dataset, the group of successful communication has a wider range of results compared to the \emph{Dale-2} dataset.
+Only 16 configurations surpass 90\% accuracy, while the majority lies between 80\% and 90\%.
+As before, for the hidden sizes, no pattern is visible.
+However, there is now a clearer influence of the message length $n$ and the vocabulary size $|V|$.
+First, 15 of the 16 configurations above 90\% accuracy are allowed messages with more than 3 symbols.
+8 of these configurations even have a message length $n = 6$.
+Although there are a few configurations with $n = 1$ that perform well, most of them have an accuracy of below 85\%.
+Overall, a longer message tends to help the agents to communicate the correct object.
+Concerning the vocabulary size, a different pattern can be seen.
+A vocabulary size of $|V| = 10$, $|V| = 16$ and $|V| = 50$ gives the best results.
+All the configurations with results above 90\% use these vocabulary sizes.
+While a large vocabulary size of $|V| = 100$ can still give good results, most of the time, the accuracy stays around 80\% to 84\%.
+When given a vocabulary size of only two symbols, the agents still perform far above the baseline.
+However, compared to the other sizes, the accuracy scores are at the lower end, mainly between 62\% and 80\%.
+As expected, when the message length is allowed to be longer, these configurations perform the best.
 
-The second conclusion that can be drawn looks at the differences between the two datasets.
-Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
-Since we discriminate objects based on properties that are also distinguished in human cognition (8 colors, 2 sizes, 3 shapes), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
-There are $8*2*3=48$ possible combinations of attributes.
-% SD: Calculation behind this figure?
-% DK: (done)
-Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
-This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
-The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
-Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
-% SD: We would need to show that the messages in Dale-2 are shorter than in Dale-5 to claim this.
-% DK: TODO
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cccc|c}
+        \toprule
+               &        &     &       & \multicolumn{1}{c}{\textbf{CLEVR color}} \\  \cmidrule(lr){5-5}
+        $h\_r$ & $h\_s$ & $n$ & $|V|$ & \textbf{Accuracy}                        \\\midrule
+        {100}  & {1000} & {3} & {10}  & {73,65\%}                                \\
+        {100}  & {1000} & {3} & {16}  & {73,57\%}                                \\
+        {500}  & {500}  & {4} & {10}  & {73,57\%}                                \\
+        {100}  & {500}  & {3} & {16}  & {70,75\%}                                \\
+        {100}  & {1000} & {4} & {50}  & {69,31\%}                                \\
+        {1000} & {500}  & {1} & {16}  & {62,85\%}                                \\
+        {100}  & {500}  & {1} & {10}  & {57,12\%}                                \\
+        {1000} & {500}  & {6} & {16}  & {54,69\%}                                \\
+        {100}  & {100}  & {4} & {2}   & {52,95\%}                                \\
+        {1000} & {500}  & {1} & {100} & {52,82\%}                                \\
+        {100}  & {1000} & {1} & {100} & {52,34\%}                                \\
+        {100}  & {500}  & {3} & {2}   & {51,48\%}                                \\
+        {500}  & {1000} & {3} & {16}  & {24,61\%}                                \\
+        {100}  & {1000} & {3} & {50}  & {24,48\%}                                \\
+        {1000} & {1000} & {4} & {100} & {24\%}                                   \\
+        {500}  & {1000} & {1} & {2}   & {23,96\%}                                \\
+        {1000} & {100}  & {6} & {10}  & {23,87\%}                                \\
+        {500}  & {1000} & {6} & {50}  & {23,87\%}                                \\
+        {500}  & {500}  & {2} & {2}   & {22,35\%}                                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Selected results of the object identification games on the CLEVR color dataset with $e_s=100$ and $e_r=500$}
+    \label{tab:results_discriminator_color}
+\end{table}
 
-\cmtDK[inline]{similarity to bounding box classifier}
+Finally, the results for the \emph{CLEVR color} dataset show a similar structure (see Table \ref{tab:results_discriminator_color}).
+Two main groups can be identified: Configurations that perform as the baseline with an accuracy of around 23\% and configurations that successfully communicate with accuracies of 40\% to 73\%.
+Compared to the \emph{Dale-5} dataset, the highest accuracy is much lower.
+Furthermore, almost 70\% of the configurations are unsuccessful, which is much bigger compared to the \emph{Dale-5} dataset.
+Still, similar conclusions can be drawn.
+Different hidden sizes don't seem to have an influence.
+For this dataset, medium message lengths of $n=2$, $n=3$ and $n=4$ provide the best results for the agents.
+However, the agents can still communicate successfully with $n=1$ and $n=6$, but the accuracy reaches maximum 64\% and mainly stays around 50\%.
+As before, the agents can utilize a vocabulary size of $|V|=10$ and $|V|=16$ the best.
+Some configurations with $|V|=2$ and $|V|=100$ achieve accuracies of up to 50\%, but most lead to an unsuccessful communication.
+Interesting hereby is that a vocabulary size of $|V|=2$ requires $n$ to be bigger for better results, while the opposite is true for $|V|=100$.
+
+Overall these results show that across all datasets successful communication between the agents emerge.
+The sender is able to encode a description of the target object into the message which the receiver is able to decode.
+Unsurprisingly, this works the best for simplest dataset \emph{Dale-2} with only one distractor.
+With an increasing number of distractors, fewer games are successful, and successful games perform worse.
+Interestingly, with all vocabulary sizes a language can emerge between the agents.
+While the vocabulary size $|V|=2$ usually requires longer messages to succeed, there are some games that also achieve high accuracies with short messages.
+This fact suggests that the agents don't necessarily rely on the attributes shape, color or size, since they would require more symbols to be encoded.
+Instead, the agents likely utilize other patterns that are not used in human referring expressions.
+This will be further analyzed in section \ref{sec:analysis_language}.

### 2024-03-17 source/mlt/language_games/run_experiments.py
results

@@ -18,20 +18,20 @@ variables = {
     "--lr": [0.0002],
     "--max_samples": [10000],
     "--temperature": [1],
-    "--model": ["discriminator"],
-    "--dataset": ["dale-2", "dale-5", "colour"],
+    "--model": ["bounding_box_caption_generator"],
+    "--dataset": ["dale-5", "colour"],
     # -- model specifics --
     "--sender_embedding": [100],
-    "--sender_hidden": [100, 500, 1000],
+    "--sender_hidden": [500],
     "--sender_image_embedding": [500],
     "--receiver_embedding": [100],
-    "--receiver_hidden": [100, 500, 1000],
+    "--receiver_hidden": [500],
     # "--receiver_projection": [100],
-    "--receiver_image_embedding": [500],
-    # "--receiver_decoder_out": [1500],
-    # "--receiver_decoder_embedding": [15],
+    "--receiver_image_embedding": [100],
+    "--receiver_decoder_out": [1500],
+    "--receiver_decoder_embedding": [15],
     # "--receiver_coordinate_classifier_dimension": [2048],
-    "--vocab_size": [2, 10, 16, 50, 100],
+    "--vocab_size": [2],
     "--max_len": [1, 2, 3, 4, 6],
 }
 

### 2024-03-17 thesis/sections/5.3-reference-resolution.tex
results

@@
 \subsubsection*{Results}
+\cmtDK[inline]{attention predictor}
+
 Tables \ref{tab:results:reference-resolver} to \ref{tab:results:reference-resolver-masked} show the results of the five adaptions of the reference resolver.

### 2024-03-17 thesis/sections/6.2-re-generation.tex
results

@@
     \centering
+    \begin{tabular}{cc|ccc|ccc|ccc}
+        \toprule
+                                      &        & \multicolumn{3}{c}{\textbf{Dale 2}} & \multicolumn{3}{c}{\textbf{Dale 5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                                    \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        $n$                           & $|V|$  & \textbf{Acc}                        & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc}     & \textbf{F1}      & \textbf{NT}     & \textbf{Acc} & \textbf{F1} & \textbf{NT} \\\midrule
+        \multicolumn{2}{c|}{baseline} & {41,8} & {42,36}                             & {37,11}                             & {62,66}                                  & {81,16}          & {18,36}          & {13,52}         & {33,08}      & {34,53}                   \\\midrule
+        {1}                           & {2}    & {70,49}                             & {54,73}                             & {7,29}                                   & {70,16}          & {85,06}          & {13,06}         & {11,68}      & {32,25}     & {34,03}     \\
+        {1}                           & {10}   & {78,26}                             & {63,33}                             & {3,56}                                   & {68,97}          & {85,01}          & {10,85}         & {12,72}      & {32,77}     & {33,85}     \\
+        {1}                           & {16}   & {74,18}                             & {52,21}                             & {5,38}                                   & {71,57}          & {86,87}          & {10,72}         & {13,19}      & {33,64}     & {33,33}     \\
+        {1}                           & {50}   & {70,83}                             & {43,56}                             & {4,21}                                   & {65,23}          & {83,11}          & {12,54}         & {11,81}      & {33,14}     & {34,72}     \\
+        {1}                           & {100}  & {73,18}                             & {52,14}                             & {3,86}                                   & {60,94}          & {80,85}          & {20,36}         & {13,67}      & {33,71}     & {33,64}     \\
+        {2}                           & {2}    & {72,31}                             & {50,42}                             & {5,52}                                   & {66,2}           & {83,39}          & {14,42}         & {13,32}      & {33,62}     & {32,73}     \\
+        {2}                           & {10}   & {75,22}                             & {58,4}                              & {3,99}                                   & {73,22}          & {87,68}          & {9,03}          & {12,37}      & {33,3}      & {34,11}     \\
+        {2}                           & {16}   & {\textbf{81,64}}                    & {\textbf{73,66}}                    & {\textbf{2,86}}                          & {73,44}          & {88,21}          & {8,85}          & {12,98}      & {33,94}     & {34,24}     \\
+        {2}                           & {50}   & {74,39}                             & {48,53}                             & {3,17}                                   & {\textbf{75,35}} & {\textbf{88,77}} & {\textbf{8,07}} & {14,19}      & {33,04}     & {33,46}     \\
+        {2}                           & {100}  & {74,09}                             & {54,88}                             & {3,65}                                   & {54,56}          & {76,95}          & {22,57}         & {13,28}      & {32,8}      & {33,38}     \\
+        {3}                           & {2}    & {71,03}                             & {50,76}                             & {4,62}                                   & {65,18}          & {82,68}          & {14,23}         & {12,76}      & {33,58}     & {33,64}     \\
+        {3}                           & {10}   & {76,91}                             & {59,42}                             & {2,6}                                    & {71,66}          & {87,14}          & {9,9}           & {12,54}      & {32,75}     & {33,85}     \\
+        {3}                           & {16}   & {79,95}                             & {67,55}                             & {3,17}                                   & {50,13}          & {73,24}          & {24,22}         & {12,93}      & {33,79}     & {34,46}     \\
+        {3}                           & {50}   & {62,93}                             & {48,03}                             & {16,58}                                  & {\textbf{76,09}} & {\textbf{88,5}}  & {\textbf{9,2}}  & {13,67}      & {33,73}     & {34,29}     \\
+        {3}                           & {100}  & {75,74}                             & {57,51}                             & {2,43}                                   & {69,7}           & {86,14}          & {9,38}          & {13,02}      & {33,64}     & {34,2}      \\
+        {4}                           & {2}    & {41,23}                             & {42,53}                             & {35,33}                                  & {63,93}          & {82,34}          & {15,3}          & {12,2}       & {33,25}     & {33,64}     \\
+        {4}                           & {10}   & {78,99}                             & {66,4}                              & {2,08}                                   & {73,22}          & {88,01}          & {10,85}         & {12,5}       & {33,55}     & {33,16}     \\
+        {4}                           & {16}   & {\textbf{80,08}}                    & {\textbf{70,48}}                    & {\textbf{3,47}}                          & {69,4}           & {86,09}          & {10,63}         & {11,72}      & {33,1}      & {34,03}     \\
+        {4}                           & {50}   & {75,48}                             & {54,44}                             & {3,34}                                   & {71,35}          & {86,61}          & {10,68}         & {11,81}      & {33,18}     & {34,33}     \\
+        {4}                           & {100}  & {74,78}                             & {57,09}                             & {3,86}                                   & {59,2}           & {80,14}          & {20,57}         & {12,67}      & {32,3}      & {34,55}     \\
+        {6}                           & {2}    & {72,87}                             & {51,44}                             & {2,78}                                   & {60,82}          & {80,42}          & {17,07}         & {13,45}      & {33,73}     & {33,64}     \\
+        {6}                           & {10}   & {73,61}                             & {48,73}                             & {2,82}                                   & {32,64}          & {58,42}          & {33,12}         & {11,89}      & {33,15}     & {35,03}     \\
+        {6}                           & {16}   & {73,52}                             & {49,67}                             & {2,43}                                   & {49,39}          & {73,19}          & {23,78}         & {13,06}      & {32,31}     & {34,64}     \\
+        {6}                           & {50}   & {44,31}                             & {45,07}                             & {33,29}                                  & {\textbf{78,82}} & {\textbf{90,22}} & {\textbf{7,29}} & {12,72}      & {33,29}     & {32,99}     \\
+        {6}                           & {100}  & {\textbf{80,95}}                    & {\textbf{69,88}}                    & {\textbf{3,34}}                          & {62,2}           & {80,62}          & {19,88}         & {14,24}      & {32,36}     & {32,94}     \\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO: caption}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rr|cc|c|ccc|c|c}
+        \toprule
+                                         &             & {small} & {large} & \textbf{size}  & {cube}  & {cylinder} & {sphere} & \textbf{shape} & {<pad>} \\\midrule
+        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {45,15} & {37,96} & \textbf{41,56} & {99,25} & {99,65}    & {99,21}  & \textbf{99,37} & {94,58} \\
+                                         & {Recall}    & {16,69} & {6,23}  & \textbf{11,46} & {99,59} & {98,55}    & {99,84}  & \textbf{99,33} & {99,3}  \\\midrule
+        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {80,9}  & {89,63} & \textbf{85,27} & {97,67} & {97,7}     & {95,21}  & \textbf{96,86} & {88,1}  \\
+                                         & {Recall}    & {76,99} & {69,14} & \textbf{73,07} & {96,58} & {97,22}    & {96,72}  & \textbf{96,84} & {96,1}  \\\midrule
+        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                          \\\textbf{color}} & {Precision}  & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
+                                         & {Recall}    & {-}     & {-}     & \textbf{-}     & {100}   & {100}      & {100}    & \textbf{100}   & {100}   \\
+        \bottomrule
+    \end{tabular}
+    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
+    \label{tab:results:bb-re-generator_size-shape}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{rr|cccccccc|c}
+        \toprule
+                                         &             & {blue}  & {brown} & {cyan}  & {gray}  & {green} & {purple} & {red}   & {yellow} & \textbf{color}   \\\midrule
+        \multirow{2}{*}{\textbf{Dale-2}} & {Precision} & {66}    & {79,35} & {71,14} & {86,11} & {70,48} & {70,74}  & {67,95} & {76,26}  & \textbf{73,5}    \\
+                                         & {Recall}    & {88,96} & {63,43} & {91,9}  & {46,02} & {69,29} & {65,99}  & {75,35} & {79,54}  & \textbf{72,56}   \\\midrule
+        \multirow{2}{*}{\textbf{Dale-5}} & {Precision} & {92,61} & {93,3}  & {94,73} & {91,12} & {93,66} & {88,71}  & {88,04} & {89,79}  & \textbf{91,5}    \\
+                                         & {Recall}    & {89,09} & {86,31} & {91,08} & {89,08} & {83,23} & {92,16}  & {93,05} & {88,77}  & {\textbf{89,1}}  \\\midrule
+        \multirowcell{2}[0pt][r]{\textbf{CLEVR}                                                                                                             \\\textbf{color}} & {Precision}           & {0} & {0} & {0} & {0} & {12,91} & {13,9}  & {0} & {0}   & \textbf{3,35} \\
+                                         & {Recall}    & {0}     & {0}     & {0}     & {0}     & {36,25} & {69,85}  & {0}     & {0}      & {\textbf{13,26}} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
+    \label{tab:results:bb-re-generator_color}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|ccc|ccc|ccc}
+        \toprule
+                                      &         & \multicolumn{3}{c}{\textbf{Dale 2}} & \multicolumn{3}{c}{\textbf{Dale 5}} & \multicolumn{3}{c}{\textbf{CLEVR colour}}                                                                                                                                                                                                           \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        $n$                           & $|V|$   & \textbf{A\textsubscript{color}}     & \textbf{A\textsubscript{shape}}     & \textbf{A\textsubscript{size}}            & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{size}} & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{size}} \\\midrule
+        \multicolumn{2}{c|}{baseline} & {92,03} & {88,91}                             & {50,55}                             & {89,38}                                   & {80,78}                         & {49,92}                         & {90,78}                        & {100}                           & {76,72}                                                          \\\midrule
+        {1}                           & {2}     & {93,68}                             & {89,69}                             & {52,04}                                   & {90,12}                         & {88,09}                         & {49,74}                        & {87,37}                         & {100}                           & {76,56}                        \\
+        {1}                           & {10}    & {94,21}                             & {96,31}                             & {52,78}                                   & {90,15}                         & {90,6}                          & {49,74}                        & \textbf{91,32}                  & {100}                           & {76,52}                        \\
+        {1}                           & {16}    & {92,8}                              & {94,97}                             & {52,78}                                   & {93,45}                         & {91,97}                         & {49,74}                        & {91,03}                         & {100}                           & {76,52}                        \\
+        {1}                           & {50}    & {91,58}                             & {90,1}                              & {52,78}                                   & {93,39}                         & {94,66}                         & {49,74}                        & {89,41}                         & {100}                           & {76,52}                        \\
+        {1}                           & {100}   & {92,66}                             & {90,41}                             & {52,78}                                   & {90,23}                         & {83,03}                         & {49,74}                        & {86,72}                         & {100}                           & {76,52}                        \\
+        {2}                           & {2}     & {92,45}                             & {92,45}                             & {52,04}                                   & \textbf{94,38}                  & {88,72}                         & {49,74}                        & {88,89}                         & {100}                           & {76,56}                        \\
+        {2}                           & {10}    & {\textbf{94,44}}                    & {96,09}                             & {52,78}                                   & {90,71}                         & {91,54}                         & {49,74}                        & {89,54}                         & {100}                           & {76,52}                        \\
+        {2}                           & {16}    & {92,66}                             & {90,58}                             & {52,78}                                   & {94,14}                         & \textbf{97,18}                  & {49,74}                        & {86,98}                         & {100}                           & {76,52}                        \\
+        {2}                           & {50}    & {92,45}                             & {89,28}                             & {52,78}                                   & {93,01}                         & {92,01}                         & {49,74}                        & {86,5}                          & {100}                           & {76,52}                        \\
+        {2}                           & {100}   & {92,97}                             & {\textbf{98,91}}                    & {52,78}                                   & {92,49}                         & {92,66}                         & {49,74}                        & {86,68}                         & {100}                           & {76,52}                        \\
+        {3}                           & {2}     & {93,32}                             & {90,97}                             & {52,04}                                   & {92,53}                         & {87,88}                         & {49,74}                        & {90,58}                         & {100}                           & {76,56}                        \\
+        {3}                           & {10}    & \textbf{{94,79}}                    & {97,01}                             & {52,78}                                   & {92,8}                          & {94,4}                          & {49,74}                        & \textbf{91,17}                  & {100}                           & {76,52}                        \\
+        {3}                           & {16}    & {93,1}                              & {95,36}                             & {52,78}                                   & {91,88}                         & {86,72}                         & {49,74}                        & {88,06}                         & {100}                           & {76,52}                        \\
+        {3}                           & {50}    & {90,84}                             & {90,93}                             & {52,78}                                   & \textbf{94,18}                  & {94,92}                         & {49,74}                        & {88,59}                         & {100}                           & {76,52}                        \\
+        {3}                           & {100}   & {90,76}                             & {89,37}                             & {52,78}                                   & {92,8}                          & {94,1}                          & {49,74}                        & \textbf{91,84}                  & {100}                           & {76,52}                        \\
+        {4}                           & {2}     & {92,11}                             & {94,86}                             & {52,04}                                   & {93,86}                         & {88,56}                         & {49,74}                        & {87,11}                         & {100}                           & {76,56}                        \\
+        {4}                           & {10}    & {92,75}                             & {90,58}                             & {52,78}                                   & {93,71}                         & {95,05}                         & {49,74}                        & {89,89}                         & {100}                           & {76,52}                        \\
+        {4}                           & {16}    & {94,14}                             & {\textbf{98,57}}                    & {52,78}                                   & {\textbf{95,23}}                & \textbf{96,14}                  & {49,74}                        & {88,06}                         & {100}                           & {76,52}                        \\
+        {4}                           & {50}    & {93,32}                             & {95,66}                             & {52,78}                                   & {90,45}                         & {91,01}                         & {49,74}                        & {88,85}                         & {100}                           & {76,52}                        \\
+        {4}                           & {100}   & {93,58}                             & {89,41}                             & {52,78}                                   & {93,19}                         & {94,66}                         & {49,74}                        & {86,98}                         & {100}                           & {76,52}                        \\
+        {6}                           & {2}     & {92,94}                             & {93,39}                             & {52,04}                                   & {92,16}                         & {82,18}                         & {49,74}                        & {90,05}                         & {100}                           & {76,56}                        \\
+        {6}                           & {10}    & {\textbf{94,23}}                    & {\textbf{98,78}}                    & {52,78}                                   & {92,75}                         & \textbf{96,79}                  & {49,74}                        & {87,67}                         & {100}                           & {76,52}                        \\
+        {6}                           & {16}    & {91,32}                             & {96,96}                             & {52,78}                                   & {93,75}                         & {95,57}                         & {49,74}                        & {89,37}                         & {100}                           & {76,52}                        \\
+        {6}                           & {50}    & {91,45}                             & {88,28}                             & {52,78}                                   & {86,24}                         & {91,71}                         & {49,74}                        & {88,32}                         & {100}                           & {76,52}                        \\
+        {6}                           & {100}   & {92,45}                             & {94,92}                             & {52,78}                                   & {93,62}                         & {95,3}                          & {49,74}                        & {89,19}                         & {100}                           & {76,52}                        \\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO: caption}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
     \begin{tabular}{ccc|ccc|ccc}

### 2024-03-24 thesis/sections/6.2-re-generation.tex
results

@@
         \toprule
-                                      &        & \multicolumn{3}{c}{\textbf{Dale 2}} & \multicolumn{3}{c}{\textbf{Dale 5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                                    \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
-        $n$                           & $|V|$  & \textbf{Acc}                        & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc}     & \textbf{F1}      & \textbf{NT}     & \textbf{Acc} & \textbf{F1} & \textbf{NT} \\\midrule
+                                      &        & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                                     \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        $n$                           & $|V|$  & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.}    & \textbf{F1}      & \textbf{NT}     & \textbf{Acc.} & \textbf{F1} & \textbf{NT} \\\midrule
         \multicolumn{2}{c|}{baseline} & {41,8} & {42,36}                             & {37,11}                             & {62,66}                                  & {81,16}          & {18,36}          & {13,52}         & {33,08}       & {34,53}                   \\\midrule
@@
     \end{tabular}
-    \caption{TODO: caption}
-    \label{TODO: label}
+    \caption{Overall accuracies (Acc.), F1-score (F1) and non-target accuracies (NT) in \% of the RE generator games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
+    \label{tab:results:re-generator-game}
 \end{table}
+
+Table \ref{tab:results:re-generator-game} shows the \emph{overall accuracies}, \emph{F1-scores} and \emph{non-target accuracies} across all datasets for different message lengths $n$ and vocabulary sizes $|V|$.
+The baselines are shown in the first row.
+Hereby, the sender is sending random messages, so that the receiver needs to solve the task on their own.
+In all configurations that don't pass this baseline, no meaningful messages are exchanged, and no language emerged.
+If configurations perform better on the other hand means that the agents learned to encode some meaning in their messages.
+
+When first having a look at the results of the \emph{Dale-2} dataset, it can be seen that the baseline already identifies correct referring expressions in 41,8\% of the samples.
+However, almost all other configurations perform much better.
+This shows that the agents consistently learn languages to solve the tasks.
+Still, differences in how well the agents perform can be identified.
+While a small vocabulary with $|V| = 2$ mostly beats the baseline, the emerged language only lets the agents generate the correct referring expression in a maximum of around 72\% of the samples.
+The similar result is visible for a bigger vocabulary size of $|V|=50$ which has the highest accuracy at around 74\%.
+A vocabulary size of $|V|=100$ performs consistently slightly better, with accuracies of around 73\% to 75\%.
+Interestingly, when the agents are allowed to exchange messages with a message length of $n=6$, the accuracy peaks at almost 81\%.
+Additionally, also the \emph{F1-score} almost reaches 70\% whereas it usually stayed around 50\% to 60\% in the previously discussed configuration.
+The \emph{F1-score} evaluates each word in the referring expression separately instead of the binary \emph{overall accuracy} for the complete referring expression.
+A small increase of the accuracy accompanied by a large increase of the F1-score indicates that the agents learned to encode an additional attribute, with the larger vocabulary and longer messages.
+Finally, vocabulary sizes of $|V|=10$ and $|V|=16$ consistently perform the best with accuracies of 74\% to 81\%.
+However, a difference in the \emph{F1-score} is visible: configurations with $|V|=16$ reach 67,55\% to 73,66\% with message lengths $n=[2,4]$, where $|V|=10$ only reaches an F1-score of 66,4\% with $n=4$ and otherwise performs worse.
+This indicates as before that the agents were able to encode more attributes with the larger vocabulary.
+
+The results look different when the agents are trained to generate referring expressions for the \emph{Dale-5} dataset.
+First, the baseline, in other words the receiver on their own, performs much better.
+It already produces the whole correct referring expressions in 62,66\% of the samples, and achieves and \emph{F1-score} of 81,36\%.
+However, the \emph{non-target accuracy} is quite high with producing correct referring expressions for a distractor in 18,36\% of the cases.
+This allows two conclusions: On the one side, the baseline can pretty successfully extract a bias from the images.
+Since there are multiple distractors in the image, the probability that they share attributes that are relevant for discriminating the target is higher.
+The model therefore can learn to generate \emph{shapes} and \emph{colors} that appear more often correctly, which leads to a higher \emph{F1-score}.
+On the other side, the larger number of distractors makes it possible for the model to single out the target from the image alone, as it needs to be uniquely identifiable.
+If there are for instance two large red cubes, two blue spheres, but only one green sphere, the target needs to be the green sphere, and no additional textual information is needed.
+The model can exploit this bias in the dataset, which leads to a higher accuracy.
+However, distractors might also be unique in the scene, so the model won't be able to consistently single out the correct unique object, which leads to a higher \emph{non-target accuracy}.
+Messages by a sender would need to solve these problems.
+
+Interestingly, many configurations with a sender perform as the baseline or even worse.
+Only 16 out of 30 configurations beat the baseline, some only by a few percent points.
+The best configuration however, has an accuracy of 78,82\%, an \emph{F1-score} of 90,22\% and a \emph{non-target accuracy} of 7,29\% which is 16\% points, 9\% points and 11\% points better than the baseline respectively.
+The models tend to perform better with a message length of $n>1$ and $n<6$, and consistently the best across all vocabulary sizes with $n=2$.
+However, there are good results for all $n$, with the best configuration utilizing $n=6$.
+Looking at the vocabulary sizes, configurations with $|V|=2$ beat the baseline only slightly, except with $n=1$.
+When the agents use a vocabulary with $|V|=10$, they beat the baseline consistently and push the \emph{non-target accuracy} close to and below 10\%.
+However, when longer messages are allowed with $n=6$, the accuracy drops to only 32,64\%.
+The reason might be that the agents learned to exchange messages about patterns that are only present in the training samples and that were not transferrable to unseen samples.
+The learned language was therefore not general enough.
+A similar result is visible for $|V|=16$, but the model additionally performs badly with $n=3$.
+Using a vocabulary with $|V|=50$, the models can generate the best referring expressions.
+Almost all configurations have an accuracy of above 70\%, three of them even above 75\%.
+Having more symbols to describe the objects apparently helps the agents to learn a useful language.
+However, when the bottleneck is too big, the agents are not able to encode useful messages as seen with $|V|=100$.
+The agents are only able to beat the baseline with $n=3$; all other configurations perform equal or worse.
+
+Finally, the results of the experiments with the \emph{CLEVR color} dataset show that the agents are not able to learn a language to describe the target objects.
+The baseline corresponds to a random guess of the color while the other attributes are identified correctly (since they are identical for target and distractor in the complete scene).
+No configuration beats this baseline which shows that no meaningful language emerges independently of vocabulary size $|V|$ and message length $n$.
 
@@
     \end{tabular}
-    \caption{Precision and Recall in \% for <pad>, size and shape tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
+    \caption{Precision and Recall in \% for <pad>, size and shape tokens for the configurations with the highest accuracy in Table \ref{tab:results:re-generator-game}.
 The columns \textbf{shape} and \textbf{size} show the average across all tokens of the respective attribute.}
-    \label{tab:results:bb-re-generator_size-shape}
+    \label{tab:results:re-generator-game_size-shape}
 \end{table}
@@
     \end{tabular}
-    \caption{Precision and Recall in \% for color tokens with $e=100$, $LSTM_o=500$ and $LSTM_e=30$. The column \textbf{color} shows the average across all colors.}
-    \label{tab:results:bb-re-generator_color}
+    \caption{Precision and Recall in \% for color tokens for the configurations with the highest accuracy in Table \ref{tab:results:re-generator-game}.
+The column \textbf{color} shows the average across all colors.}
+    \label{tab:results:re-generator-game_color}
 \end{table}
+
+Tables \ref{tab:results:re-generator-game_size-shape} and \ref{tab:results:re-generator-game_color} show the precision and recall divided by attributes and tokens.
+The results have a similar general pattern to the single model experiments in Section \ref{sec:referring_expression_generation}.
+For the \emph{Dale} dataset, the \emph{shape} is identified well while the agents struggle more with the \emph{color} and the most with the \emph{size}.
+The agents identify the \emph{shape} and \emph{size} perfectly on the \emph{CLEVR color} dataset, but only predict few \emph{colors}.
+
+Looking at the \emph{shape}, the agents perform almost perfectly with the \emph{Dale-2} dataset.
+The precision and recall are slightly lower with the \emph{Dale-5} dataset, and the agents predict the \emph{sphere} too often, while predicting the \emph{cube} too few.
+The difference however lies only around 1\%.
+
+Interestingly, the agents can predict the \emph{color} the best, when presented with the \emph{Dale-5} dataset.
+The precision and recall lie around 90\%, while for the \emph{Dale-2} dataset they lie only around 73\%.
+However, there is a bigger difference across the colors.
+Some colors are predicted more often (e.g.
+purple and red for \emph{Dale-5}, and yellow red, cyan and blue for \emph{Dale-2}) than others (e.g.
+brown and green for \emph{Dale-5}, and purple, gray and brown for \emph{Dale-2}).
+As can be seen, the colors are not consistent and "difficult" or "easy" colors can't be identified.
+Additionally, these colors also differ across several runs of the same configurations and datasets.
+The differences between the two \emph{Dale} dataset, might be explainable by the number of scenes in which the color is part of the referring expression.
+As seen in Table \ref{tab:results:re-generator-game}, the \emph{overall accuracy} is higher for the \emph{Dale-2} dataset, even though the prediction of the \emph{color} attribute is worse.
+The reason for this is that many referring expressions in the \emph{Dale-2} dataset only rely on the \emph{shape} and no \emph{color} is needed.
+The agents therefore can perform well even though predictions of the \emph{color} are wrong.
+On the other side, the \emph{color} is a central part of almost all referring expressions in the \emph{Dale-5} dataset and the agents are pushed to learn this attribute more intensely.
+
+Since the single model in section \ref{sec:referring_expression_generation} already struggled with predicting the \emph{color} on the \emph{CLEVR color} dataset, it is no surprise that it performs badly in the more complex setup of language games.
+Hereby, the agents are not able to communicate anything and the receiver only generates two colors (green and purple) for all shown samples.
+This results in a random guess of one of the 8 \emph{colors}, explaining the \emph{overall accuracy} of $\frac{1}{8}=12,5\%$.
+
+The final attribute is the \emph{size}.
+While the agents easily learn that the \emph{size} is never part of the referring expression for the \emph{CLEVR color} dataset, they struggle with the \emph{Dale} datasets.
+As in section \ref{sec:referring_expression_generation}, the agents again struggle to predict the correct length of the referring expression.
+Especially, the agents predict too short referring expressions as the precision of the \emph{<pad>} token in lower than its recall.
+This is mostly due to the \emph{size} attribute, since a large difference of the precision and recall is visible.
+As for the \emph{color}, the agents perform better on the \emph{Dale-5} dataset.
+This reason might be the same.
+The \emph{size} attribute is more often needed to discriminate the target object and appears therefore in more samples and is more central in the referring expressions.
+Similarly to the \emph{color} attribute, there are differences between the single tokens.
+However, these are not constant across configurations, datasets and runs and no "difficult" token can be identified.
 
@@
         \toprule
-                                      &         & \multicolumn{3}{c}{\textbf{Dale 2}} & \multicolumn{3}{c}{\textbf{Dale 5}} & \multicolumn{3}{c}{\textbf{CLEVR colour}}                                                                                                                                                                                                           \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+                                      &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR colour}}                                                                                                                                                                                                           \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
         $n$                           & $|V|$   & \textbf{A\textsubscript{color}}     & \textbf{A\textsubscript{shape}}     & \textbf{A\textsubscript{size}}            & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{size}} & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{size}} \\\midrule
@@
     \end{tabular}
-    \caption{TODO: caption}
-    \label{TODO: label}
+    \caption{Accuracies for each attribute in \% of the attribute predictor games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
+    \label{tab:results:attribute-predictor-game}
 \end{table}
+
+Finally, Table \ref{tab:results:attribute-predictor-game} show the results for the games in which the receiver predicts a one-hot vector of the attributes.
+Again, the baselines are games, the receiver plays alone.
+Hereby, it can be seen that the receiver can already predict the \emph{color} and \emph{shape} very well: 92\% and 88,91\% respectively on the \emph{Dale-2} dataset, 89,38\% and 80,78\% respectively on the \emph{Dale-5} dataset, and 90,78\% and 100\% respectively on the \emph{CLEVR color} dataset.
+The \emph{size} however corresponds to a random guess on the \emph{Dale} dataset and only achieves 76,72\% on the \emph{CLEVR color} dataset.
 

### 2024-03-25 thesis/sections/5.2-re-generation.tex
results

@@
         \toprule
-                  &           & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR colour}}                                                                                         \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+                  &           & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                         \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
         $LSTM\_o$ & $LSTM\_e$ & \textbf{Acc.}                       & \textbf{F1}                         & \textbf{NT}                              & \textbf{Acc.} & \textbf{F1} & \textbf{NT} & \textbf{Acc.} & \textbf{F1} & \textbf{NT} \\\midrule

### 2024-03-25 thesis/sections/6.2-re-generation.tex
results

@@
         \multicolumn{2}{c|}{baseline} & {41,8} & {42,36}                             & {37,11}                             & {62,66}                                  & {81,16}                   & {18,36}                   & {13,52}                   & {33,08}                   & {34,53}                                               \\\midrule
-        {1}                           & {2}    & {70,49}                             & {54,73}                             & {7,29}                                   & {70,16}          & {85,06}          & {13,06}         & {11,68}       & {32,25}     & {34,03}     \\
-        {1}                           & {10}   & {78,26}                             & {63,33}                             & {3,56}                                   & {68,97}          & {85,01}          & {10,85}         & {12,72}       & {32,77}     & {33,85}     \\
-        {1}                           & {16}   & {74,18}                             & {52,21}                             & {5,38}                                   & {71,57}          & {86,87}          & {10,72}         & {13,19}       & {33,64}     & {33,33}     \\
-        {1}                           & {50}   & {70,83}                             & {43,56}                             & {4,21}                                   & {65,23}          & {83,11}          & {12,54}         & {11,81}       & {33,14}     & {34,72}     \\
-        {1}                           & {100}  & {73,18}                             & {52,14}                             & {3,86}                                   & {60,94}          & {80,85}          & {20,36}         & {13,67}       & {33,71}     & {33,64}     \\
-        {2}                           & {2}    & {72,31}                             & {50,42}                             & {5,52}                                   & {66,2}           & {83,39}          & {14,42}         & {13,32}       & {33,62}     & {32,73}     \\
-        {2}                           & {10}   & {75,22}                             & {58,4}                              & {3,99}                                   & {73,22}          & {87,68}          & {9,03}          & {12,37}       & {33,3}      & {34,11}     \\
-        {2}                           & {16}   & {\textbf{81,64}}                    & {\textbf{73,66}}                    & {\textbf{2,86}}                          & {73,44}          & {88,21}          & {8,85}          & {12,98}       & {33,94}     & {34,24}     \\
-        {2}                           & {50}   & {74,39}                             & {48,53}                             & {3,17}                                   & {\textbf{75,35}} & {\textbf{88,77}} & {\textbf{8,07}} & {14,19}       & {33,04}     & {33,46}     \\
-        {2}                           & {100}  & {74,09}                             & {54,88}                             & {3,65}                                   & {54,56}          & {76,95}          & {22,57}         & {13,28}       & {32,8}      & {33,38}     \\
-        {3}                           & {2}    & {71,03}                             & {50,76}                             & {4,62}                                   & {65,18}          & {82,68}          & {14,23}         & {12,76}       & {33,58}     & {33,64}     \\
-        {3}                           & {10}   & {76,91}                             & {59,42}                             & {2,6}                                    & {71,66}          & {87,14}          & {9,9}           & {12,54}       & {32,75}     & {33,85}     \\
-        {3}                           & {16}   & {79,95}                             & {67,55}                             & {3,17}                                   & {50,13}          & {73,24}          & {24,22}         & {12,93}       & {33,79}     & {34,46}     \\
-        {3}                           & {50}   & {62,93}                             & {48,03}                             & {16,58}                                  & {\textbf{76,09}} & {\textbf{88,5}}  & {\textbf{9,2}}  & {13,67}       & {33,73}     & {34,29}     \\
-        {3}                           & {100}  & {75,74}                             & {57,51}                             & {2,43}                                   & {69,7}           & {86,14}          & {9,38}          & {13,02}       & {33,64}     & {34,2}      \\
-        {4}                           & {2}    & {41,23}                             & {42,53}                             & {35,33}                                  & {63,93}          & {82,34}          & {15,3}          & {12,2}        & {33,25}     & {33,64}     \\
-        {4}                           & {10}   & {78,99}                             & {66,4}                              & {2,08}                                   & {73,22}          & {88,01}          & {10,85}         & {12,5}        & {33,55}     & {33,16}     \\
-        {4}                           & {16}   & {\textbf{80,08}}                    & {\textbf{70,48}}                    & {\textbf{3,47}}                          & {69,4}           & {86,09}          & {10,63}         & {11,72}       & {33,1}      & {34,03}     \\
-        {4}                           & {50}   & {75,48}                             & {54,44}                             & {3,34}                                   & {71,35}          & {86,61}          & {10,68}         & {11,81}       & {33,18}     & {34,33}     \\
-        {4}                           & {100}  & {74,78}                             & {57,09}                             & {3,86}                                   & {59,2}           & {80,14}          & {20,57}         & {12,67}       & {32,3}      & {34,55}     \\
-        {6}                           & {2}    & {72,87}                             & {51,44}                             & {2,78}                                   & {60,82}          & {80,42}          & {17,07}         & {13,45}       & {33,73}     & {33,64}     \\
-        {6}                           & {10}   & {73,61}                             & {48,73}                             & {2,82}                                   & {32,64}          & {58,42}          & {33,12}         & {11,89}       & {33,15}     & {35,03}     \\
-        {6}                           & {16}   & {73,52}                             & {49,67}                             & {2,43}                                   & {49,39}          & {73,19}          & {23,78}         & {13,06}       & {32,31}     & {34,64}     \\
-        {6}                           & {50}   & {44,31}                             & {45,07}                             & {33,29}                                  & {\textbf{78,82}} & {\textbf{90,22}} & {\textbf{7,29}} & {12,72}       & {33,29}     & {32,99}     \\
-        {6}                           & {100}  & {\textbf{80,95}}                    & {\textbf{69,88}}                    & {\textbf{3,34}}                          & {62,2}           & {80,62}          & {19,88}         & {14,24}       & {32,36}     & {32,94}     \\
+        {1}                           & {2}    & {70,49}                             & {54,73}                             & {7,29}                                   & {70,16}                   & {85,06}                   & {13,06}                   & \textcolor{purple}{11,68} & \textcolor{purple}{32,25} & \textcolor{purple}{34,03} \\
+        {1}                           & {10}   & {78,26}                             & {63,33}                             & {3,56}                                   & {68,97}                   & {85,01}                   & {10,85}                   & \textcolor{purple}{12,72} & \textcolor{purple}{32,77} & \textcolor{purple}{33,85} \\
+        {1}                           & {16}   & {74,18}                             & {52,21}                             & {5,38}                                   & {71,57}                   & {86,87}                   & {10,72}                   & \textcolor{purple}{13,19} & \textcolor{purple}{33,64} & \textcolor{purple}{33,33} \\
+        {1}                           & {50}   & {70,83}                             & \textcolor{purple}{43,56}           & {4,21}                                   & {65,23}                   & {83,11}                   & {12,54}                   & \textcolor{purple}{11,81} & \textcolor{purple}{33,14} & \textcolor{purple}{34,72} \\
+        {1}                           & {100}  & {73,18}                             & {52,14}                             & {3,86}                                   & \textcolor{purple}{60,94} & \textcolor{purple}{80,85} & \textcolor{purple}{20,36} & \textcolor{purple}{13,67} & \textcolor{purple}{33,71} & \textcolor{purple}{33,64} \\
+        {2}                           & {2}    & {72,31}                             & {50,42}                             & {5,52}                                   & {66,2}                    & {83,39}                   & {14,42}                   & \textcolor{purple}{13,32} & \textcolor{purple}{33,62} & \textcolor{purple}{32,73} \\
+        {2}                           & {10}   & {75,22}                             & {58,4}                              & {3,99}                                   & {73,22}                   & {87,68}                   & {9,03}                    & \textcolor{purple}{12,37} & \textcolor{purple}{33,3}  & \textcolor{purple}{34,11} \\
+        {2}                           & {16}   & {\textbf{81,64}}                    & {\textbf{73,66}}                    & {\textbf{2,86}}                          & {73,44}                   & {88,21}                   & {8,85}                    & \textcolor{purple}{12,98} & \textcolor{purple}{33,94} & \textcolor{purple}{34,24} \\
+        {2}                           & {50}   & {74,39}                             & {48,53}                             & {3,17}                                   & {\textbf{75,35}}          & \textbf{88,77}            & \textbf{8,07}             & \textcolor{purple}{14,19} & \textcolor{purple}{33,04} & \textcolor{purple}{33,46} \\
+        {2}                           & {100}  & {74,09}                             & {54,88}                             & {3,65}                                   & \textcolor{purple}{54,56} & \textcolor{purple}{76,95} & \textcolor{purple}{22,57} & \textcolor{purple}{13,28} & \textcolor{purple}{32,8}  & \textcolor{purple}{33,38} \\
+        {3}                           & {2}    & {71,03}                             & {50,76}                             & {4,62}                                   & {65,18}                   & \textcolor{purple}{82,68} & {14,23}                   & \textcolor{purple}{12,76} & \textcolor{purple}{33,58} & \textcolor{purple}{33,64} \\
+        {3}                           & {10}   & {76,91}                             & {59,42}                             & {2,6}                                    & {71,66}                   & {87,14}                   & {9,9}                     & \textcolor{purple}{12,54} & \textcolor{purple}{32,75} & \textcolor{purple}{33,85} \\
+        {3}                           & {16}   & {79,95}                             & {67,55}                             & {3,17}                                   & \textcolor{purple}{50,13} & \textcolor{purple}{73,24} & {24,22}                   & \textcolor{purple}{12,93} & \textcolor{purple}{33,79} & \textcolor{purple}{34,46} \\
+        {3}                           & {50}   & {62,93}                             & {48,03}                             & {16,58}                                  & {\textbf{76,09}}          & {\textbf{88,5}}           & \textbf{9,2}              & \textcolor{purple}{13,67} & \textcolor{purple}{33,73} & \textcolor{purple}{34,29} \\
+        {3}                           & {100}  & {75,74}                             & {57,51}                             & {2,43}                                   & {69,7}                    & {86,14}                   & {9,38}                    & \textcolor{purple}{13,02} & \textcolor{purple}{33,64} & \textcolor{purple}{34,2}  \\
+        {4}                           & {2}    & \textcolor{purple}{41,23}           & \textcolor{purple}{42,53}           & \textcolor{purple}{35,33}                & {63,93}                   & \textcolor{purple}{82,34} & {15,3}                    & \textcolor{purple}{12,2}  & \textcolor{purple}{33,25} & \textcolor{purple}{33,64} \\
+        {4}                           & {10}   & {78,99}                             & {66,4}                              & {2,08}                                   & {73,22}                   & {88,01}                   & {10,85}                   & \textcolor{purple}{12,5}  & \textcolor{purple}{33,55} & \textcolor{purple}{33,16} \\
+        {4}                           & {16}   & {\textbf{80,08}}                    & {\textbf{70,48}}                    & {\textbf{3,47}}                          & {69,4}                    & {86,09}                   & {10,63}                   & \textcolor{purple}{11,72} & \textcolor{purple}{33,1}  & \textcolor{purple}{34,03} \\
+        {4}                           & {50}   & {75,48}                             & {54,44}                             & {3,34}                                   & {71,35}                   & {86,61}                   & {10,68}                   & \textcolor{purple}{11,81} & \textcolor{purple}{33,18} & \textcolor{purple}{34,33} \\
+        {4}                           & {100}  & {74,78}                             & {57,09}                             & {3,86}                                   & \textcolor{purple}{59,2}  & \textcolor{purple}{80,14} & \textcolor{purple}{20,57} & \textcolor{purple}{12,67} & \textcolor{purple}{32,3}  & \textcolor{purple}{34,55} \\
+        {6}                           & {2}    & {72,87}                             & {51,44}                             & {2,78}                                   & {60,82}                   & \textcolor{purple}{80,42} & {17,07}                   & \textcolor{purple}{13,45} & \textcolor{purple}{33,73} & \textcolor{purple}{33,64} \\
+        {6}                           & {10}   & {73,61}                             & {48,73}                             & {2,82}                                   & \textcolor{purple}{32,64} & \textcolor{purple}{58,42} & \textcolor{purple}{33,12} & \textcolor{purple}{11,89} & \textcolor{purple}{33,15} & \textcolor{purple}{35,03} \\
+        {6}                           & {16}   & {73,52}                             & {49,67}                             & {2,43}                                   & \textcolor{purple}{49,39} & \textcolor{purple}{73,19} & \textcolor{purple}{23,78} & \textcolor{purple}{13,06} & \textcolor{purple}{32,31} & \textcolor{purple}{34,64} \\
+        {6}                           & {50}   & \textcolor{purple}{44,31}           & {45,07}                             & \textcolor{purple}{33,29}                & {\textbf{78,82}}          & {\textbf{90,22}}          & {\textbf{7,29}}           & \textcolor{purple}{12,72} & \textcolor{purple}{33,29} & \textcolor{purple}{32,99} \\
+        {6}                           & {100}  & {\textbf{80,95}}                    & {\textbf{69,88}}                    & {\textbf{3,34}}                          & {62,2}                    & \textcolor{purple}{80,62} & \textcolor{purple}{19,88} & \textcolor{purple}{14,24} & \textcolor{purple}{32,36} & \textcolor{purple}{32,94} \\
         \bottomrule
@@
         \toprule
-                                      &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR colour}}                                                                                                                                                                                                           \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
-        $n$                           & $|V|$   & \textbf{A\textsubscript{color}}     & \textbf{A\textsubscript{shape}}     & \textbf{A\textsubscript{size}}            & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{size}} & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{size}} \\\midrule
-        \multicolumn{2}{c|}{baseline} & {92,03} & {88,91}                             & {50,55}                             & {89,38}                                   & {80,78}                         & {49,92}                         & {90,78}                        & {100}                           & {76,72}                                                          \\\midrule
-        {1}                           & {2}     & {93,68}                             & {89,69}                             & {52,04}                                   & {90,12}                         & {88,09}                         & {49,74}                        & {87,37}                         & {100}                           & {76,56}                        \\
-        {1}                           & {10}    & {94,21}                             & {96,31}                             & {52,78}                                   & {90,15}                         & {90,6}                          & {49,74}                        & \textbf{91,32}                  & {100}                           & {76,52}                        \\
-        {1}                           & {16}    & {92,8}                              & {94,97}                             & {52,78}                                   & {93,45}                         & {91,97}                         & {49,74}                        & {91,03}                         & {100}                           & {76,52}                        \\
-        {1}                           & {50}    & {91,58}                             & {90,1}                              & {52,78}                                   & {93,39}                         & {94,66}                         & {49,74}                        & {89,41}                         & {100}                           & {76,52}                        \\
-        {1}                           & {100}   & {92,66}                             & {90,41}                             & {52,78}                                   & {90,23}                         & {83,03}                         & {49,74}                        & {86,72}                         & {100}                           & {76,52}                        \\
-        {2}                           & {2}     & {92,45}                             & {92,45}                             & {52,04}                                   & \textbf{94,38}                  & {88,72}                         & {49,74}                        & {88,89}                         & {100}                           & {76,56}                        \\
-        {2}                           & {10}    & {\textbf{94,44}}                    & {96,09}                             & {52,78}                                   & {90,71}                         & {91,54}                         & {49,74}                        & {89,54}                         & {100}                           & {76,52}                        \\
-        {2}                           & {16}    & {92,66}                             & {90,58}                             & {52,78}                                   & {94,14}                         & \textbf{97,18}                  & {49,74}                        & {86,98}                         & {100}                           & {76,52}                        \\
-        {2}                           & {50}    & {92,45}                             & {89,28}                             & {52,78}                                   & {93,01}                         & {92,01}                         & {49,74}                        & {86,5}                          & {100}                           & {76,52}                        \\
-        {2}                           & {100}   & {92,97}                             & {\textbf{98,91}}                    & {52,78}                                   & {92,49}                         & {92,66}                         & {49,74}                        & {86,68}                         & {100}                           & {76,52}                        \\
-        {3}                           & {2}     & {93,32}                             & {90,97}                             & {52,04}                                   & {92,53}                         & {87,88}                         & {49,74}                        & {90,58}                         & {100}                           & {76,56}                        \\
-        {3}                           & {10}    & \textbf{{94,79}}                    & {97,01}                             & {52,78}                                   & {92,8}                          & {94,4}                          & {49,74}                        & \textbf{91,17}                  & {100}                           & {76,52}                        \\
-        {3}                           & {16}    & {93,1}                              & {95,36}                             & {52,78}                                   & {91,88}                         & {86,72}                         & {49,74}                        & {88,06}                         & {100}                           & {76,52}                        \\
-        {3}                           & {50}    & {90,84}                             & {90,93}                             & {52,78}                                   & \textbf{94,18}                  & {94,92}                         & {49,74}                        & {88,59}                         & {100}                           & {76,52}                        \\
-        {3}                           & {100}   & {90,76}                             & {89,37}                             & {52,78}                                   & {92,8}                          & {94,1}                          & {49,74}                        & \textbf{91,84}                  & {100}                           & {76,52}                        \\
-        {4}                           & {2}     & {92,11}                             & {94,86}                             & {52,04}                                   & {93,86}                         & {88,56}                         & {49,74}                        & {87,11}                         & {100}                           & {76,56}                        \\
-        {4}                           & {10}    & {92,75}                             & {90,58}                             & {52,78}                                   & {93,71}                         & {95,05}                         & {49,74}                        & {89,89}                         & {100}                           & {76,52}                        \\
-        {4}                           & {16}    & {94,14}                             & {\textbf{98,57}}                    & {52,78}                                   & {\textbf{95,23}}                & \textbf{96,14}                  & {49,74}                        & {88,06}                         & {100}                           & {76,52}                        \\
-        {4}                           & {50}    & {93,32}                             & {95,66}                             & {52,78}                                   & {90,45}                         & {91,01}                         & {49,74}                        & {88,85}                         & {100}                           & {76,52}                        \\
-        {4}                           & {100}   & {93,58}                             & {89,41}                             & {52,78}                                   & {93,19}                         & {94,66}                         & {49,74}                        & {86,98}                         & {100}                           & {76,52}                        \\
-        {6}                           & {2}     & {92,94}                             & {93,39}                             & {52,04}                                   & {92,16}                         & {82,18}                         & {49,74}                        & {90,05}                         & {100}                           & {76,56}                        \\
-        {6}                           & {10}    & {\textbf{94,23}}                    & {\textbf{98,78}}                    & {52,78}                                   & {92,75}                         & \textbf{96,79}                  & {49,74}                        & {87,67}                         & {100}                           & {76,52}                        \\
-        {6}                           & {16}    & {91,32}                             & {96,96}                             & {52,78}                                   & {93,75}                         & {95,57}                         & {49,74}                        & {89,37}                         & {100}                           & {76,52}                        \\
-        {6}                           & {50}    & {91,45}                             & {88,28}                             & {52,78}                                   & {86,24}                         & {91,71}                         & {49,74}                        & {88,32}                         & {100}                           & {76,52}                        \\
-        {6}                           & {100}   & {92,45}                             & {94,92}                             & {52,78}                                   & {93,62}                         & {95,3}                          & {49,74}                        & {89,19}                         & {100}                           & {76,52}                        \\
+                                      &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}} & \multicolumn{3}{c}{\textbf{CLEVR color}}                                                                                                                                                                                                           \\  \cmidrule(lr){3-5}\cmidrule(lr){6-8}\cmidrule(lr){9-11}
+        $n$                           & $|V|$   & \textbf{A\textsubscript{shape}}     & \textbf{A\textsubscript{color}}     & \textbf{A\textsubscript{size}}           & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{size}} & \textbf{A\textsubscript{shape}} & \textbf{A\textsubscript{color}} & \textbf{A\textsubscript{size}} \\\midrule
+        \multicolumn{2}{c|}{baseline} & {88,91} & {92,03}                             & {50,55}                             & {80,78}                                  & {89,38}                         & {49,92}                         & {100}                          & {90,78}                         & {76,72}                                                          \\\midrule
+        {1}                           & {2}     & \textcolor{purple}{89,69}           & {93,68}                             & \textcolor{purple}{52,04}                & {88,09}                         & \textcolor{purple}{90,12}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{87,37}       & \textcolor{purple}{76,56}      \\
+        {1}                           & {10}    & {96,31}                             & {94,21}                             & \textcolor{purple}{52,78}                & {90,6}                          & \textcolor{purple}{90,15}       & \textcolor{purple}{49,74}      & {100}                           & \textbf{91,32}                  & \textcolor{purple}{76,52}      \\
+        {1}                           & {16}    & {94,97}                             & \textcolor{purple}{92,8}            & \textcolor{purple}{52,78}                & {91,97}                         & {93,45}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{91,03}       & \textcolor{purple}{76,52}      \\
+        {1}                           & {50}    & \textcolor{purple}{90,1}            & \textcolor{purple}{91,58}           & \textcolor{purple}{52,78}                & {94,66}                         & {93,39}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,41}       & \textcolor{purple}{76,52}      \\
+        {1}                           & {100}   & {90,41}                             & \textcolor{purple}{92,66}           & \textcolor{purple}{52,78}                & \textcolor{purple}{83,03}       & \textcolor{purple}{90,23}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,72}       & \textcolor{purple}{76,52}      \\
+        {2}                           & {2}     & {92,45}                             & \textcolor{purple}{92,45}           & \textcolor{purple}{52,04}                & {88,72}                         & \textbf{94,38}                  & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,89}       & \textcolor{purple}{76,56}      \\
+        {2}                           & {10}    & {96,09}                             & \textbf{94,44}                      & \textcolor{purple}{52,78}                & {91,54}                         & \textcolor{purple}{90,71}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,54}       & \textcolor{purple}{76,52}      \\
+        {2}                           & {16}    & {90,58}                             & \textcolor{purple}{92,66}           & \textcolor{purple}{52,78}                & \textbf{97,18}                  & {94,14}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,98}       & \textcolor{purple}{76,52}      \\
+        {2}                           & {50}    & {89,28}                             & \textcolor{purple}{92,45}           & \textcolor{purple}{52,78}                & {92,01}                         & {93,01}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,5}        & \textcolor{purple}{76,52}      \\
+        {2}                           & {100}   & \textbf{98,91}                      & \textcolor{purple}{92,97}           & \textcolor{purple}{52,78}                & {92,66}                         & {92,49}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,68}       & \textcolor{purple}{76,52}      \\
+        {3}                           & {2}     & {90,97}                             & {93,32}                             & \textcolor{purple}{52,04}                & {87,88}                         & {92,53}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{90,58}       & \textcolor{purple}{76,56}      \\
+        {3}                           & {10}    & {97,01}                             & \textbf{{94,79}}                    & \textcolor{purple}{52,78}                & {94,4}                          & {92,8}                          & \textcolor{purple}{49,74}      & {100}                           & \textbf{91,17}                  & \textcolor{purple}{76,52}      \\
+        {3}                           & {16}    & {95,36}                             & {93,1}                              & \textcolor{purple}{52,78}                & {86,72}                         & {91,88}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,06}       & \textcolor{purple}{76,52}      \\
+        {3}                           & {50}    & {90,93}                             & \textcolor{purple}{90,84}           & \textcolor{purple}{52,78}                & {94,92}                         & \textbf{94,18}                  & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,59}       & \textcolor{purple}{76,52}      \\
+        {3}                           & {100}   & \textcolor{purple}{89,37}           & \textcolor{purple}{90,76}           & \textcolor{purple}{52,78}                & {94,1}                          & {92,8}                          & \textcolor{purple}{49,74}      & {100}                           & \textbf{91,84}                  & \textcolor{purple}{76,52}      \\
+        {4}                           & {2}     & {94,86}                             & \textcolor{purple}{92,11}           & \textcolor{purple}{52,04}                & {88,56}                         & {93,86}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{87,11}       & \textcolor{purple}{76,56}      \\
+        {4}                           & {10}    & {90,58}                             & \textcolor{purple}{92,75}           & \textcolor{purple}{52,78}                & {95,05}                         & {93,71}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,89}       & \textcolor{purple}{76,52}      \\
+        {4}                           & {16}    & \textbf{98,57}                      & {94,14}                             & \textcolor{purple}{52,78}                & \textbf{96,14}                  & \textbf{95,23}                  & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,06}       & \textcolor{purple}{76,52}      \\
+        {4}                           & {50}    & {95,66}                             & {93,32}                             & \textcolor{purple}{52,78}                & {91,01}                         & \textcolor{purple}{90,45}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,85}       & \textcolor{purple}{76,52}      \\
+        {4}                           & {100}   & \textcolor{purple}{89,41}           & {93,58}                             & \textcolor{purple}{52,78}                & {94,66}                         & {93,19}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,98}       & \textcolor{purple}{76,52}      \\
+        {6}                           & {2}     & {93,39}                             & {92,94}                             & \textcolor{purple}{52,04}                & \textcolor{purple}{82,18}       & {92,16}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{90,05}       & \textcolor{purple}{76,56}      \\
+        {6}                           & {10}    & \textbf{98,78}                      & \textbf{94,23}                      & \textcolor{purple}{52,78}                & \textbf{96,79}                  & {92,75}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{87,67}       & \textcolor{purple}{76,52}      \\
+        {6}                           & {16}    & {96,96}                             & \textcolor{purple}{91,32}           & \textcolor{purple}{52,78}                & {95,57}                         & {93,75}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,37}       & \textcolor{purple}{76,52}      \\
+        {6}                           & {50}    & \textcolor{purple}{88,28}           & \textcolor{purple}{91,45}           & \textcolor{purple}{52,78}                & {91,71}                         & \textcolor{purple}{86,24}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,32}       & \textcolor{purple}{76,52}      \\
+        {6}                           & {100}   & {94,92}                             & \textcolor{purple}{92,45}           & \textcolor{purple}{52,78}                & {95,3}                          & {93,62}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,19}       & \textcolor{purple}{76,52}      \\
         \bottomrule
@@
 
-Finally, Table \ref{tab:results:attribute-predictor-game} show the results for the games in which the receiver predicts a one-hot vector of the attributes.
+Finally, Table \ref{tab:results:attribute-predictor-game} shows the results for the games in which the receiver predicts a one-hot vector of the attributes.
 Again, the baselines are games, the receiver plays alone.
-Hereby, it can be seen that the receiver can already predict the \emph{color} and \emph{shape} very well: 92\% and 88,91\% respectively on the \emph{Dale-2} dataset, 89,38\% and 80,78\% respectively on the \emph{Dale-5} dataset, and 90,78\% and 100\% respectively on the \emph{CLEVR color} dataset.
+Hereby, it can be seen that the receiver can already predict the \emph{shape} and \emph{color}  very well: 88,91\% and 92\% respectively on the \emph{Dale-2} dataset, 80,78\% and 89,38\% respectively on the \emph{Dale-5} dataset, and 100\% and 90,78\% respectively on the \emph{CLEVR color} dataset.
 The \emph{size} however corresponds to a random guess on the \emph{Dale} dataset and only achieves 76,72\% on the \emph{CLEVR color} dataset.
+This shows that the receiver can already identify attributes of the target object on their own and discriminate it from the distractors.
+However, the size still seems to be an attribute which is difficult to extract from the visual scene.
+The increase in performance compared to the generation of referring expression can be explained by two reasons:
+First, the complexity of the receiver is greatly reduced.
+When only predicting one-hot vectors, much fewer layers and weights are involved that need to be trained and hence the models can learn the given task.
+Secondly, the complexity of the task is reduced as well.
+The generation of referring expressions not only includes the learning of the tokens (and their embeddings), but also their order in the referring expression and their salience for the incremental algorithm.
+Using one-hot vectors, both order and salience are irrelevant and only the attributes themselves are needed to be learned.
+The goal of these experiments is now to see if the information of the sender improves the predictions of the receiver.
 
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                             \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Acc.}                       & \textbf{word-by-word}               & \textbf{length} & \textbf{Acc.} & \textbf{word-by-word} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {22,9\%}                            & {62,8\%}                            & {1}             & {7,1\%}       & {40\%}                & {1}             \\
-        {13}  & {10}    & {10}    & {22,8\%}                            & {62,9\%}                            & {0}             & {7,3\%}       & {38,7\%}              & {1}             \\
-        {20}  & {10}    & {10}    & {24,6\%}                            & {64\%}                              & {1}             & {6,7\%}       & {38,7\%}              & {1}             \\
-        {100} & {10}    & {10}    & {24,4\%}                            & {62\%}                              & {1}             & {7,8\%}       & {40\%}                & {1}             \\
-        {100} & {100}   & {100}   & {21\%}                              & {62\%}                              & {1}             & {6,5\%}       & {37,8\%}              & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the caption generator: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    % SD: What is word-by-word?
-    % DK: the same as in the caption generator with a single model (QUESTION)
-    \label{tab:results_caption_generator_game}
-\end{table}
+As can be seen in the table, the improvement is dependent on the attributes.
+When the agents are trained on the \emph{Dale-2} dataset, almost all configurations improve the prediction of the shape.
+While 11 out of 30 configurations only improve by a maximum of 2\% points, most can predict the \emph{shape} much better, with an accuracy of up to 98,91\%.
+A pattern based on the vocabulary size $|V|$ and message length $n$ is however difficult to identify.
+A smaller vocabulary size of $|V|=10$ generally fares well, but also a large vocabulary size of $|V|=100$ can help to predict the shape both with short messages ($n=2$) and long messages ($n=6$).
+Opposed to the \emph{shape}, the prediction of the \emph{color} can only be improved by a small margin of a maximum of 2,7\% points.
+Most of the configurations produce the same or slightly worse results than the baseline.
+Again, a vocabulary of $|V|=10$ seems to help the most.
+Interestingly, when the agents can only use one token to encode their message ($|V|=2$), an improvement of predictions of the \emph{shape} comes to the expense of the accuracy of the \emph{color} and vice versa.
+This indicates that the sender doesn't have enough space to communicate both attributes to the receiver.
+Looking at the \emph{size}, a constant improvement of 2\% points across all configurations can be seen.
+The final accuracy however still lies slightly above a random guess.
 
-The results of the caption generator game are summarized in Table \ref{tab:results_caption_generator_game}.
-In general, it can be seen that the agents have much bigger problems, to solve the task together than a single neural network.
-The highest accuracy for descriptions, the agents manage to predict correctly is at 24,6\% for images of the 'Dale-2' dataset.
-Compared to the (masked) accuracy of the single model with 72\%, the agents predict 47,4\% points less correct descriptions.
-A similar worse performance can be seen for the 'Dale-5' dataset.
-Here, the agents only manage to produce for 7,8\% of the images correct descriptions with a vocabulary size of 100, 13,2\% points less than the single neural model.
-The same effect can be seen for the word-by-word accuracy, which is much lower than the metric for the single neural model for both datasets.
-% SD: With the vocabulary size of 100.
-% DK: (done)
+The results are similar when the agents need to describe the \emph{Dale-5} dataset.
+The baseline accuracies of the \emph{shape} and \emph{color} are lower, but can be improved to similar peaks as for the \emph{Dale-2} dataset.
+The highest accuracy for the \emph{shape} lies at 97,18\%, while it lies at 95,23\% for the \emph{color}.
+This indicates that the larger number of distractors makes it difficult to focus on the correct object and instead describes distractors when the receiver is on their own.
+But when the sender can communicate the correct object, the receiver describes the attributes as well as with fewer distractors.
+As before, the predictions of the \emph{size} correspond to a random guess.
 
-When looking, how the different variables affect the performance, it can be seen that a bigger vocabulary size tends to help the agents.
-% SD: Not sure. The difference is still very small, within a couple of %.
-% DK: TODO
-This is only visible for the 'Dale-2' dataset.
-With constant hidden sizes of 10, the agents score around 22,9\% with only 10 and 13 available symbols.
-When this is increased to 20 and respectively 100 symbols, the agents can increase their accuracy to around 24,5\%.
-However, the increase is relatively small.
-Interestingly, this effect only occurs, when the hidden sizes are small with only 10 dimensions.
-As soon as they are increased to 100 dimensions with a vocabulary size of 100 symbols, the accuracy drops to 21\%.
-% SD: Very small decerase for such a large difference in vocabulary size compared to what we have seen earlier.
-% DK: TODO
+Using the \emph{CLEVR color} dataset, almost no improvements can be seen.
+The predictions of the \emph{shape} stay at 100\% accuracy, while the predictions of the \emph{color} get slightly worse for most configurations and if they improve they only improve by maximum 1\% point.
+The predictions of the \emph{size} stay the same.
+The fact that the sender has a smaller influence on the \emph{CLEVR color} dataset than on the \emph{Dale} dataset is surprising since the sender would need to communicate only the \emph{color} instead of additionally the \emph{shape} (and \emph{size}).
+A reason why this happens might be the larger number of distractors.
+This number might hinder the receiver to converge and understand the task, by connecting the one-hot vector to the colors of the target object.
+While for the \emph{Dale} datasets the receiver has a choice between only 2 to 5 objects, they need to decide between up to 10 objects on the \emph{CLEVR color} dataset.
+That might confuse the model where to focus on and subsequently to connect it to the on-hot vector.
+Future experiments with a similar dataset including fewer distractors would need to verify this hypothesis.
 
-Looking at the 'Dale-5' dataset, the increase is still there, when the vocabulary is increased to 100 symbols.
-Nonetheless, the difference is with 0,5\% points even smaller and the reason may be due to other influences, such as the random initialization of the weights of the agents.
-This is confirmed, when looking at emerged languages.
-In all the setups, the same message is communicated for all samples, independently of the input image.
-This is also reflected in the length of the messages.
-For the setup with a vocabulary size of $|V| = 13$, no message is transferred, and the accuracy stays the same as in the other setups.
-% SD: Here it would be much better to have a loss curve to show differences in loss between the number of communication events, the same could also be done for accuracy.
-Then we would have a clearer picture of how vocabulary size affects learning.
-It could be that with 100 vocabulary there is actually a better performance than before the final cut-off point.
-Is this possible?
-How was the final cut-off point determined anyway?
-% DK: TODO
-
-These results show that the agents are not at all able to encode meaning about the images and target objects in their messages.
-% SD: Negative result: how can we then interpret that? What could we change? What did we learn from this?
-% DK: TODO
-This is especially interesting, compared to single model caption generator in section \ref{sec:referring_expression_generation}.
-In these experiments, the model was able to converge towards correct captions and therefore able to extract the necessary information.
-This shows that a main challenge for the agents lies in grounding symbols in these extracted features.
+Concluding, these experiments show that the models are able to extract and communicate information about the \emph{shape} and \emph{color} from the images when few objects are present.
+When the agents are forced to generate referring expressions which are more complex in their structure, they produce perfect results less frequently.
+However, the communication is successfully very often as it helps the receiver in identifying the correct attributes and referring expression.

### 2024-03-25 thesis/sections/6.3-reference-resolution.tex
results

@@
 \subsection*{Results}
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|cc|cc|cc}
+        \toprule
+                                      &         & \multicolumn{2}{c}{\textbf{Dale-2}} & \multicolumn{2}{c}{\textbf{Dale-5}} & \multicolumn{2}{c}{\textbf{CLEVR color}}                                                                                         \\  \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
+        $n$                           & $|V|$   & \textbf{loss}                       & \textbf{Acc.}                       & \textbf{loss}                            & \textbf{Acc.}               & \textbf{loss}             & \textbf{Acc.}               \\\midrule
+        \multicolumn{2}{c|}{baseline} & {34,49} & {34,3\%}                            & {38,13}                             & {30,31\%}                                & {42,72}                     & {18,28\%}                                               \\\midrule
+        {1}                           & {2}     & \textcolor{purple}{34,91}           & \textcolor{purple}{34,81\%}         & \textcolor{purple}{40,61}                & {25,48\%}                   & {39,13}                   & {25\%}                      \\
+        {1}                           & {10}    & \textcolor{purple}{35,31}           & {40,15\%}                           & \textcolor{purple}{38,73}                & {28,12\%}                   & \textcolor{purple}{44,19} & {20,53\%}                   \\
+        {1}                           & {16}    & \textcolor{purple}{35,4}            & {42,93\%}                           & \textcolor{purple}{39,58}                & {26,17\%}                   & {39,64}                   & {24,18\%}                   \\
+        {1}                           & {50}    & {32,64}                             & {42,01\%}                           & \textcolor{purple}{38,99}                & {28,52\%}                   & \textcolor{purple}{42,61} & {20,79\%}                   \\
+        {1}                           & {100}   & \textcolor{purple}{36,73}           & \textcolor{purple}{31,21\%}         & \textcolor{purple}{40,25}                & {24,91\%}                   & \textcolor{purple}{41,5}  & {21,44\%}                   \\
+        {2}                           & {2}     & \textcolor{purple}{33,76}           & {40,67\%}                           & {35,86}                                  & \textcolor{purple}{31,73\%} & \textcolor{purple}{41,79} & {22,14\%}                   \\
+        {2}                           & {10}    & {31,37}                             & {43,27\%}                           & {37,23}                                  & \textcolor{purple}{30,34\%} & \textcolor{purple}{42,53} & \textcolor{purple}{19,75\%} \\
+        {2}                           & {16}    & {31,19}                             & {46,27\%}                           & \textcolor{purple}{39,38}                & {28,21\%}                   & \textcolor{purple}{41,74} & {22,7\%}                    \\
+        {2}                           & {50}    & \textcolor{purple}{33,66}           & {39,28\%}                           & \textcolor{purple}{39,62}                & {25,35\%}                   & \textcolor{purple}{45,63} & \textcolor{purple}{17,93\%} \\
+        {2}                           & {100}   & \textcolor{purple}{35,1}            & \textcolor{purple}{34,81\%}         & \textcolor{purple}{39,86}                & {27,3\%}                    & \textcolor{purple}{45,45} & \textcolor{purple}{19,05\%} \\
+        {3}                           & {2}     & {32,99}                             & {41,15\%}                           & {37,76}                                  & \textcolor{purple}{30,99\%} & \textcolor{purple}{46,45} & {20,92\%}                   \\
+        {3}                           & {10}    & \textcolor{purple}{33,74}           & {38,89\%}                           & \textcolor{purple}{40,71}                & {26,48\%}                   & \textcolor{purple}{44,53} & \textcolor{purple}{19,01\%} \\
+        {3}                           & {16}    & \textcolor{purple}{34,18}           & {38,98\%}                           & \textcolor{purple}{38,52}                & \textcolor{purple}{30,03\%} & \textcolor{purple}{41,22} & {21,53\%}                   \\
+        {3}                           & {50}    & \textcolor{purple}{34,38}           & {38,93\%}                           & \textcolor{purple}{39,52}                & {26,74\%}                   & {38,4}                    & {24,87\%}                   \\
+        {3}                           & {100}   & \textcolor{purple}{33,16}           & {40,19\%}                           & \textcolor{purple}{39,75}                & {27,78\%}                   & \textcolor{purple}{41,29} & {21,83\%}                   \\
+        {4}                           & {2}     & \textcolor{purple}{33,48}           & {40,36\%}                           & \textcolor{purple}{38,28}                & \textcolor{purple}{30,9\%}  & \textcolor{purple}{42,53} & {22,4\%}                    \\
+        {4}                           & {10}    & \textcolor{purple}{35,01}           & {39,97\%}                           & \textcolor{purple}{40,88}                & {26,39\%}                   & \textcolor{purple}{42,33} & {20,44\%}                   \\
+        {4}                           & {16}    & \textcolor{purple}{35,59}           & \textcolor{purple}{35,59\%}         & \textcolor{purple}{37,91}                & \textcolor{purple}{31,86\%} & \textcolor{purple}{41,9}  & {20,27\%}                   \\
+        {4}                           & {50}    & \textcolor{purple}{34,99}           & {36,68\%}                           & \textcolor{purple}{40,89}                & {23,26\%}                   & \textcolor{purple}{43,77} & \textcolor{purple}{18,36\%} \\
+        {4}                           & {100}   & \textcolor{purple}{35,46}           & {37,46\%}                           & \textcolor{purple}{38,53}                & \textcolor{purple}{29,08\%} & {40,6}                    & {21,22\%}                   \\
+        {6}                           & {2}     & \textcolor{purple}{34,97}           & {37,28\%}                           & \textcolor{purple}{39,52}                & \textcolor{purple}{29,69\%} & {40,53}                   & {24,7\%}                    \\
+        {6}                           & {10}    & {32,39}                             & {43,53\%}                           & \textcolor{purple}{38,61}                & {28,26\%}                   & {40,84}                   & {21,74\%}                   \\
+        {6}                           & {16}    & \textcolor{purple}{34,87}           & \textcolor{purple}{34,38\%}         & \textcolor{purple}{39,03}                & {26,82\%}                   & {40,34}                   & {23,31\%}                   \\
+        {6}                           & {50}    & \textcolor{purple}{33,85}           & {37,07\%}                           & \textcolor{purple}{39,39}                & {27,17\%}                   & {40,8}                    & {22,09\%}                   \\
+        {6}                           & {100}   & \textcolor{purple}{34,62}           & {40,02\%}                           & \textcolor{purple}{39,46}                & {25,48\%}                   & \textcolor{purple}{42,53} & \textcolor{purple}{18,4\%}  \\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO: caption}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c}
+        \toprule
+                                      &           & \multicolumn{1}{c}{\textbf{Dale-2}} & \multicolumn{1}{c}{\textbf{Dale-5}} & \multicolumn{1}{c}{\textbf{CLEVR color}} \\  \cmidrule(lr){3-3}\cmidrule(lr){4-4}\cmidrule(lr){5-5}
+        $n$                           & $|V|$     & \textbf{Probability mass}           & \textbf{Probability mass}           & \textbf{Probability mass}                \\\midrule
+        \multicolumn{2}{c|}{baseline} & {62,16\%} & {49,61\%}                           & {41,68\%}                                                                      \\\midrule
+        {1}                           & {2}       & {93,3\%}                            & {65,36\%}                           & \textcolor{purple}{39,22\%}              \\
+        {1}                           & {10}      & {93,85\%}                           & \textcolor{purple}{46,97\%}         & \textcolor{purple}{40,1\%}               \\
+        {1}                           & {16}      & {92,06\%}                           & {76,68\%}                           & \textcolor{purple}{36,43\%}              \\
+        {1}                           & {50}      & {94\%}                              & {76,43\%}                           & \textcolor{purple}{39,97\%}              \\
+        {1}                           & {100}     & {94,03\%}                           & \textcolor{purple}{54,14\%}         & \textcolor{purple}{40,27\%}              \\
+        {2}                           & {2}       & {92,27\%}                           & \textcolor{purple}{52,15\%}         & \textcolor{purple}{33,64\%}              \\
+        {2}                           & {10}      & {96,16\%}                           & {80,26\%}                           & \textcolor{purple}{36,53\%}              \\
+        {2}                           & {16}      & {95,84\%}                           & {84,03\%}                           & \textcolor{purple}{39,65\%}              \\
+        {2}                           & {50}      & {93,78\%}                           & \textcolor{purple}{52,24\%}         & \textcolor{purple}{39,56\%}              \\
+        {2}                           & {100}     & {92,43\%}                           & \textcolor{purple}{53,23\%}         & \textcolor{purple}{37,68\%}              \\
+        {3}                           & {2}       & {94,52\%}                           & \textcolor{purple}{51,97\%}         & \textcolor{purple}{37,09\%}              \\
+        {3}                           & {10}      & {94,9\%}                            & \textcolor{purple}{53,47\%}         & \textcolor{purple}{38,24\%}              \\
+        {3}                           & {16}      & {94,59\%}                           & {81,46\%}                           & {67,88\%}                                \\
+        {3}                           & {50}      & {93,88\%}                           & {79,65\%}                           & \textcolor{purple}{40,36\%}              \\
+        {3}                           & {100}     & {95,25\%}                           & \textcolor{purple}{48,52\%}         & \textcolor{purple}{42,55\%}              \\
+        {4}                           & {2}       & {89,15\%}                           & \textcolor{purple}{51,98\%}         & \textcolor{purple}{39,68\%}              \\
+        {4}                           & {10}      & {96,08\%}                           & \textcolor{purple}{48,03\%}         & {64,31\%}                                \\
+        {4}                           & {16}      & {94,14\%}                           & \textcolor{purple}{49,81\%}         & \textcolor{purple}{40,84\%}              \\
+        {4}                           & {50}      & {96,24\%}                           & \textcolor{purple}{48,79\%}         & \textcolor{purple}{43,61\%}              \\
+        {4}                           & {100}     & {95,55\%}                           & \textcolor{purple}{49,65\%}         & \textcolor{purple}{42,85\%}              \\
+        {6}                           & {2}       & \textcolor{purple}{59,68\%}         & \textcolor{purple}{53,57\%}         & \textcolor{purple}{38,43\%}              \\
+        {6}                           & {10}      & \textcolor{purple}{63,46\%}         & {82,12\%}                           & \textcolor{purple}{40,11\%}              \\
+        {6}                           & {16}      & {95,86\%}                           & \textcolor{purple}{50,71\%}         & \textcolor{purple}{40,61\%}              \\
+        {6}                           & {50}      & {91,27\%}                           & \textcolor{purple}{52,55\%}         & \textcolor{purple}{40,21\%}              \\
+        {6}                           & {100}     & \textcolor{purple}{60,27\%}         & \textcolor{purple}{46,92\%}         & \textcolor{purple}{41,98\%}              \\
+        \bottomrule
+    \end{tabular}
+    \caption{TODO: caption}
+    \label{TODO: label}
+\end{table}
+
 In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.

### 2024-03-26 thesis/sections/6.1-object-identification.tex
results
results reference resolution game

@@
         {100}                         & {1000}    & {4} & {2}   & {94,88\%}                           \\
-        {100}  & {1000} & {6} & {10}  & {63,15\%}                           \\
-        {1000} & {500}  & {2} & {2}   & {62,98\%}                           \\\midrule
-        {-}    & {-}    & {-} & {-}   & {62,72\% (baseline)}                \\\midrule
-        {100}  & {1000} & {3} & {2}   & {61,2\%}                            \\
+        {100}                         & {1000}    & {6} & {10}  & \textcolor{purple}{63,15\%}         \\
+        {1000}                        & {500}     & {2} & {2}   & \textcolor{purple}{62,98\%}         \\\midrule
+        \multicolumn{4}{c|}{baseline} & {62,72\%}                                                     \\\midrule
+        {100}                         & {1000}    & {3} & {2}   & \textcolor{purple}{61,2\%}          \\
         \bottomrule
@@
         {100}                         & {100}     & {3} & {2}   & {74\%}                              \\
-        {500}  & {500}  & {1} & {2}   & {63,41\%}                           \\
-        {500}  & {1000} & {2} & {100} & {34,98\%}                           \\
-        {500}  & {500}  & {3} & {16}  & {34,77\%}                           \\
-        {500}  & {500}  & {6} & {2}   & {34,38\%}                           \\
-        {1000} & {1000} & {4} & {2}   & {35,33\%}                           \\
-        {1000} & {1000} & {6} & {100} & {35,24\%}                           \\
+        {500}                         & {500}     & {1} & {2}   & {63,41\%}                           \\\midrule
+        \multicolumn{4}{c|}{baseline} & {35,07\%}                                                     \\\midrule
+        {500}                         & {1000}    & {2} & {100} & \textcolor{purple}{34,98\%}         \\
+        {500}                         & {500}     & {3} & {16}  & \textcolor{purple}{34,77\%}         \\
+        {500}                         & {500}     & {6} & {2}   & \textcolor{purple}{34,38\%}         \\
+        {1000}                        & {1000}    & {4} & {2}   & \textcolor{purple}{35,33\%}         \\
+        {1000}                        & {1000}    & {6} & {100} & \textcolor{purple}{35,24\%}         \\
         \bottomrule
@@
         {100}                         & {500}     & {3} & {2}   & {51,48\%}                                \\
-        {500}  & {1000} & {3} & {16}  & {24,61\%}                                \\
-        {100}  & {1000} & {3} & {50}  & {24,48\%}                                \\
-        {1000} & {1000} & {4} & {100} & {24\%}                                   \\
-        {500}  & {1000} & {1} & {2}   & {23,96\%}                                \\
-        {1000} & {100}  & {6} & {10}  & {23,87\%}                                \\
-        {500}  & {1000} & {6} & {50}  & {23,87\%}                                \\
-        {500}  & {500}  & {2} & {2}   & {22,35\%}                                \\
+        {500}                         & {1000}    & {3} & {16}  & \textcolor{purple}{24,61\%}              \\
+        {100}                         & {1000}    & {3} & {50}  & \textcolor{purple}{24,48\%}              \\
+        {1000}                        & {1000}    & {4} & {100} & \textcolor{purple}{24\%}                 \\
+        {500}                         & {1000}    & {1} & {2}   & \textcolor{purple}{23,96\%}              \\
+        {1000}                        & {100}     & {6} & {10}  & \textcolor{purple}{23,87\%}              \\
+        {500}                         & {1000}    & {6} & {50}  & \textcolor{purple}{23,87\%}              \\\midrule
+        \multicolumn{4}{c|}{baseline} & {23,74\%}                                                          \\\midrule
+        {500}                         & {500}     & {2} & {2}   & \textcolor{purple}{22,35\%}              \\
         \bottomrule

### 2024-03-26 thesis/sections/6.3-reference-resolution.tex
results reference resolution game

@@
         \multicolumn{2}{c|}{baseline} & {34,49} & {34,3\%}                            & {38,13}                             & {30,31\%}                                & {42,72}                     & {18,28\%}                                               \\\midrule
-        {1}                           & {2}     & \textcolor{purple}{34,91}           & \textcolor{purple}{34,81\%}         & \textcolor{purple}{40,61}                & {25,48\%}                   & {39,13}                   & {25\%}                      \\
-        {1}                           & {10}    & \textcolor{purple}{35,31}           & {40,15\%}                           & \textcolor{purple}{38,73}                & {28,12\%}                   & \textcolor{purple}{44,19} & {20,53\%}                   \\
-        {1}                           & {16}    & \textcolor{purple}{35,4}            & {42,93\%}                           & \textcolor{purple}{39,58}                & {26,17\%}                   & {39,64}                   & {24,18\%}                   \\
-        {1}                           & {50}    & {32,64}                             & {42,01\%}                           & \textcolor{purple}{38,99}                & {28,52\%}                   & \textcolor{purple}{42,61} & {20,79\%}                   \\
-        {1}                           & {100}   & \textcolor{purple}{36,73}           & \textcolor{purple}{31,21\%}         & \textcolor{purple}{40,25}                & {24,91\%}                   & \textcolor{purple}{41,5}  & {21,44\%}                   \\
-        {2}                           & {2}     & \textcolor{purple}{33,76}           & {40,67\%}                           & {35,86}                                  & \textcolor{purple}{31,73\%} & \textcolor{purple}{41,79} & {22,14\%}                   \\
+        {1}                           & {2}     & \textcolor{purple}{34,91}           & \textcolor{purple}{34,81\%}         & \textcolor{purple}{40,61}                & \textcolor{purple}{25,48\%} & {39,13}                   & {25\%}                      \\
+        {1}                           & {10}    & \textcolor{purple}{35,31}           & {40,15\%}                           & \textcolor{purple}{38,73}                & \textcolor{purple}{28,12\%} & \textcolor{purple}{44,19} & {20,53\%}                   \\
+        {1}                           & {16}    & \textcolor{purple}{35,4}            & {42,93\%}                           & \textcolor{purple}{39,58}                & \textcolor{purple}{26,17\%} & {39,64}                   & {24,18\%}                   \\
+        {1}                           & {50}    & {32,64}                             & {42,01\%}                           & \textcolor{purple}{38,99}                & \textcolor{purple}{28,52\%} & \textcolor{purple}{42,61} & {20,79\%}                   \\
+        {1}                           & {100}   & \textcolor{purple}{36,73}           & \textcolor{purple}{31,21\%}         & \textcolor{purple}{40,25}                & \textcolor{purple}{24,91\%} & \textcolor{purple}{41,5}  & {21,44\%}                   \\
+        {2}                           & {2}     & \textcolor{purple}{33,76}           & {40,67\%}                           & {35,86}                                  & {31,73\%}                   & \textcolor{purple}{41,79} & {22,14\%}                   \\
         {2}                           & {10}    & {31,37}                             & {43,27\%}                           & {37,23}                                  & \textcolor{purple}{30,34\%} & \textcolor{purple}{42,53} & \textcolor{purple}{19,75\%} \\
-        {2}                           & {16}    & {31,19}                             & {46,27\%}                           & \textcolor{purple}{39,38}                & {28,21\%}                   & \textcolor{purple}{41,74} & {22,7\%}                    \\
-        {2}                           & {50}    & \textcolor{purple}{33,66}           & {39,28\%}                           & \textcolor{purple}{39,62}                & {25,35\%}                   & \textcolor{purple}{45,63} & \textcolor{purple}{17,93\%} \\
-        {2}                           & {100}   & \textcolor{purple}{35,1}            & \textcolor{purple}{34,81\%}         & \textcolor{purple}{39,86}                & {27,3\%}                    & \textcolor{purple}{45,45} & \textcolor{purple}{19,05\%} \\
+        {2}                           & {16}    & {31,19}                             & {46,27\%}                           & \textcolor{purple}{39,38}                & \textcolor{purple}{28,21\%} & \textcolor{purple}{41,74} & {22,7\%}                    \\
+        {2}                           & {50}    & \textcolor{purple}{33,66}           & {39,28\%}                           & \textcolor{purple}{39,62}                & \textcolor{purple}{25,35\%} & \textcolor{purple}{45,63} & \textcolor{purple}{17,93\%} \\
+        {2}                           & {100}   & \textcolor{purple}{35,1}            & \textcolor{purple}{34,81\%}         & \textcolor{purple}{39,86}                & \textcolor{purple}{27,3\%}  & \textcolor{purple}{45,45} & \textcolor{purple}{19,05\%} \\
         {3}                           & {2}     & {32,99}                             & {41,15\%}                           & {37,76}                                  & \textcolor{purple}{30,99\%} & \textcolor{purple}{46,45} & {20,92\%}                   \\
-        {3}                           & {10}    & \textcolor{purple}{33,74}           & {38,89\%}                           & \textcolor{purple}{40,71}                & {26,48\%}                   & \textcolor{purple}{44,53} & \textcolor{purple}{19,01\%} \\
+        {3}                           & {10}    & \textcolor{purple}{33,74}           & {38,89\%}                           & \textcolor{purple}{40,71}                & \textcolor{purple}{26,48\%} & \textcolor{purple}{44,53} & \textcolor{purple}{19,01\%} \\
         {3}                           & {16}    & \textcolor{purple}{34,18}           & {38,98\%}                           & \textcolor{purple}{38,52}                & \textcolor{purple}{30,03\%} & \textcolor{purple}{41,22} & {21,53\%}                   \\
-        {3}                           & {50}    & \textcolor{purple}{34,38}           & {38,93\%}                           & \textcolor{purple}{39,52}                & {26,74\%}                   & {38,4}                    & {24,87\%}                   \\
-        {3}                           & {100}   & \textcolor{purple}{33,16}           & {40,19\%}                           & \textcolor{purple}{39,75}                & {27,78\%}                   & \textcolor{purple}{41,29} & {21,83\%}                   \\
+        {3}                           & {50}    & \textcolor{purple}{34,38}           & {38,93\%}                           & \textcolor{purple}{39,52}                & \textcolor{purple}{26,74\%} & {38,4}                    & {24,87\%}                   \\
+        {3}                           & {100}   & \textcolor{purple}{33,16}           & {40,19\%}                           & \textcolor{purple}{39,75}                & \textcolor{purple}{27,78\%} & \textcolor{purple}{41,29} & {21,83\%}                   \\
         {4}                           & {2}     & \textcolor{purple}{33,48}           & {40,36\%}                           & \textcolor{purple}{38,28}                & \textcolor{purple}{30,9\%}  & \textcolor{purple}{42,53} & {22,4\%}                    \\
-        {4}                           & {10}    & \textcolor{purple}{35,01}           & {39,97\%}                           & \textcolor{purple}{40,88}                & {26,39\%}                   & \textcolor{purple}{42,33} & {20,44\%}                   \\
-        {4}                           & {16}    & \textcolor{purple}{35,59}           & \textcolor{purple}{35,59\%}         & \textcolor{purple}{37,91}                & \textcolor{purple}{31,86\%} & \textcolor{purple}{41,9}  & {20,27\%}                   \\
-        {4}                           & {50}    & \textcolor{purple}{34,99}           & {36,68\%}                           & \textcolor{purple}{40,89}                & {23,26\%}                   & \textcolor{purple}{43,77} & \textcolor{purple}{18,36\%} \\
+        {4}                           & {10}    & \textcolor{purple}{35,01}           & {39,97\%}                           & \textcolor{purple}{40,88}                & \textcolor{purple}{26,39\%} & \textcolor{purple}{42,33} & {20,44\%}                   \\
+        {4}                           & {16}    & \textcolor{purple}{35,59}           & \textcolor{purple}{35,59\%}         & \textcolor{purple}{37,91}                & {31,86\%}                   & \textcolor{purple}{41,9}  & {20,27\%}                   \\
+        {4}                           & {50}    & \textcolor{purple}{34,99}           & {36,68\%}                           & \textcolor{purple}{40,89}                & \textcolor{purple}{23,26\%} & \textcolor{purple}{43,77} & \textcolor{purple}{18,36\%} \\
         {4}                           & {100}   & \textcolor{purple}{35,46}           & {37,46\%}                           & \textcolor{purple}{38,53}                & \textcolor{purple}{29,08\%} & {40,6}                    & {21,22\%}                   \\
         {6}                           & {2}     & \textcolor{purple}{34,97}           & {37,28\%}                           & \textcolor{purple}{39,52}                & \textcolor{purple}{29,69\%} & {40,53}                   & {24,7\%}                    \\
-        {6}                           & {10}    & {32,39}                             & {43,53\%}                           & \textcolor{purple}{38,61}                & {28,26\%}                   & {40,84}                   & {21,74\%}                   \\
-        {6}                           & {16}    & \textcolor{purple}{34,87}           & \textcolor{purple}{34,38\%}         & \textcolor{purple}{39,03}                & {26,82\%}                   & {40,34}                   & {23,31\%}                   \\
-        {6}                           & {50}    & \textcolor{purple}{33,85}           & {37,07\%}                           & \textcolor{purple}{39,39}                & {27,17\%}                   & {40,8}                    & {22,09\%}                   \\
-        {6}                           & {100}   & \textcolor{purple}{34,62}           & {40,02\%}                           & \textcolor{purple}{39,46}                & {25,48\%}                   & \textcolor{purple}{42,53} & \textcolor{purple}{18,4\%}  \\
+        {6}                           & {10}    & {32,39}                             & {43,53\%}                           & \textcolor{purple}{38,61}                & \textcolor{purple}{28,26\%} & {40,84}                   & {21,74\%}                   \\
+        {6}                           & {16}    & \textcolor{purple}{34,87}           & \textcolor{purple}{34,38\%}         & \textcolor{purple}{39,03}                & \textcolor{purple}{26,82\%} & {40,34}                   & {23,31\%}                   \\
+        {6}                           & {50}    & \textcolor{purple}{33,85}           & {37,07\%}                           & \textcolor{purple}{39,39}                & \textcolor{purple}{27,17\%} & {40,8}                    & {22,09\%}                   \\
+        {6}                           & {100}   & \textcolor{purple}{34,62}           & {40,02\%}                           & \textcolor{purple}{39,46}                & \textcolor{purple}{25,48\%} & \textcolor{purple}{42,53} & \textcolor{purple}{18,4\%}  \\
         \bottomrule
@@
     \caption{TODO: caption}
-    \label{TODO: label}
+    \label{tab:results:coordinate-reference-resolver-game}
 \end{table}
+
+Table \ref{tab:results:coordinate-reference-resolver-game} shows the results of the coordinate reference resolver on different datasets.
+The baselines show the predictions of the receiver without information of the sender, and correspond to the structural bias in the dataset.
+When looking at how the agents fare with each dataset, not much general improvement can be seen.
+Most of the configurations don't beat the baseline, independently of the shown dataset.
+If they do, they can improve the mean distance of the predicted coordinates to the center of the target object only by 2 to 4 pixels.
+In other words, the guesses are still not consistently precise on the target objects and the agents instead learn to point towards the center of the scene, as the single models in Section \ref{sec:referring_expression_generation}.
+However, an influence on the accuracy can be identified.
+The accuracy evaluates how many of the predictions are on the target object.
+An increase in the accuracy with a stable mean distance corresponds shows a behavior where the agents make more correct predictions on the target objects, while the wrong predictions a further away from the target object.
+As a conclusion, this can mean, that the agents' predictions, both correct and incorrect are more confident.
+This seems to happen on the \emph{Dale-2} dataset.
+The mean distance is not improved, but the accuracy reaches in some configurations up to 46,27\%.
+This is an increase of around 13\% compared to the baseline.
+While this value is still lower than a random guess between the objects in the scene, it shows, that some information is exchanged between the agents.
+The meaning of the messages however doesn't seem to correspond to any attribute.
 
@@
         {2}                           & {2}       & {92,27\%}                           & \textcolor{purple}{52,15\%}         & \textcolor{purple}{33,64\%}              \\
-        {2}                           & {10}      & {96,16\%}                           & {80,26\%}                           & \textcolor{purple}{36,53\%}              \\
-        {2}                           & {16}      & {95,84\%}                           & {84,03\%}                           & \textcolor{purple}{39,65\%}              \\
+        {2}                           & {10}      & \textbf{96,16\%}                    & {80,26\%}                           & \textcolor{purple}{36,53\%}              \\
+        {2}                           & {16}      & {95,84\%}                           & \textbf{84,03\%}                    & \textcolor{purple}{39,65\%}              \\
         {2}                           & {50}      & {93,78\%}                           & \textcolor{purple}{52,24\%}         & \textcolor{purple}{39,56\%}              \\
@@
         {3}                           & {10}      & {94,9\%}                            & \textcolor{purple}{53,47\%}         & \textcolor{purple}{38,24\%}              \\
-        {3}                           & {16}      & {94,59\%}                           & {81,46\%}                           & {67,88\%}                                \\
+        {3}                           & {16}      & {94,59\%}                           & \textbf{81,46\%}                    & \textbf{67,88\%}                         \\
         {3}                           & {50}      & {93,88\%}                           & {79,65\%}                           & \textcolor{purple}{40,36\%}              \\
@@
         {4}                           & {2}       & {89,15\%}                           & \textcolor{purple}{51,98\%}         & \textcolor{purple}{39,68\%}              \\
-        {4}                           & {10}      & {96,08\%}                           & \textcolor{purple}{48,03\%}         & {64,31\%}                                \\
+        {4}                           & {10}      & \textbf{96,08\%}                    & \textcolor{purple}{48,03\%}         & \textbf{64,31\%}                         \\
         {4}                           & {16}      & {94,14\%}                           & \textcolor{purple}{49,81\%}         & \textcolor{purple}{40,84\%}              \\
-        {4}                           & {50}      & {96,24\%}                           & \textcolor{purple}{48,79\%}         & \textcolor{purple}{43,61\%}              \\
+        {4}                           & {50}      & \textbf{96,24\%}                    & \textcolor{purple}{48,79\%}         & \textcolor{purple}{43,61\%}              \\
         {4}                           & {100}     & {95,55\%}                           & \textcolor{purple}{49,65\%}         & \textcolor{purple}{42,85\%}              \\
         {6}                           & {2}       & \textcolor{purple}{59,68\%}         & \textcolor{purple}{53,57\%}         & \textcolor{purple}{38,43\%}              \\
-        {6}                           & {10}      & \textcolor{purple}{63,46\%}         & {82,12\%}                           & \textcolor{purple}{40,11\%}              \\
+        {6}                           & {10}      & \textcolor{purple}{63,46\%}         & \textbf{82,12\%}                    & \textcolor{purple}{40,11\%}              \\
         {6}                           & {16}      & {95,86\%}                           & \textcolor{purple}{50,71\%}         & \textcolor{purple}{40,61\%}              \\

### 2024-03-26 thesis/sections/6.1-object-identification.tex
results reference resolver

@@
 Concerning the vocabulary size, a different pattern can be seen.
-A vocabulary size of $|V| = 10$, $|V| = 16$ and $|V| = 50$ gives the best results.
+A vocabulary size of $|V| = \{10,16,50\}$ gives the best results.
 All the configurations with results above 90\% use these vocabulary sizes.
@@
 Different hidden sizes don't seem to have an influence.
-For this dataset, medium message lengths of $n=2$, $n=3$ and $n=4$ provide the best results for the agents.
+For this dataset, medium message lengths of $n \in \{2,3,4\}$ provide the best results for the agents.
 However, the agents can still communicate successfully with $n=1$ and $n=6$, but the accuracy reaches maximum 64\% and mainly stays around 50\%.

### 2024-03-26 thesis/sections/6.2-re-generation.tex
results reference resolver

@@
 Finally, vocabulary sizes of $|V|=10$ and $|V|=16$ consistently perform the best with accuracies of 74\% to 81\%.
-However, a difference in the \emph{F1-score} is visible: configurations with $|V|=16$ reach 67,55\% to 73,66\% with message lengths $n=[2,4]$, where $|V|=10$ only reaches an F1-score of 66,4\% with $n=4$ and otherwise performs worse.
+However, a difference in the \emph{F1-score} is visible: configurations with $|V|=16$ reach 67,55\% to 73,66\% with message lengths $n \in [2,4]$, where $|V|=10$ only reaches an F1-score of 66,4\% with $n=4$ and otherwise performs worse.
 This indicates as before that the agents were able to encode more attributes with the larger vocabulary.

### 2024-03-26 thesis/sections/6.3-reference-resolution.tex
results reference resolver

@@
 While this value is still lower than a random guess between the objects in the scene, it shows, that some information is exchanged between the agents.
-The meaning of the messages however doesn't seem to correspond to any attribute.
+The meaning of the messages however doesn't seem to correspond to any attribute or object to consistently solve the task.
+
+Looking at the results on the \emph{Dale-5}, no improvements for the mean distance nor the accuracy can be seen when a sender is involved.
+However, while the loss stays again constant, the accuracy drops lower in many configurations, up to 7\% points to an accuracy of 23,26\%.
+Again, this is an indicator that messages are transferred and interpreted by the receiver.
+The receiver seems to use this information to focus more on the objects.
+Since there are more distractors present in the \emph{Dale-5} dataset, the receiver more likely chooses one of those, as again no discriminating information about the target object seems to be transferred.
+
+Lastly, the similar pattern on the mean distance is also visible for the \emph{CLEVR color} dataset.
+The accuracy of the baseline is already lower to begin, and a sender can lift this score in almost all configurations.
+But as for the other datasets, the final accuracy is still low and reaches a maximum of only 25\%.
+The greater number of distractors seems to raise the complexity of the task even more.
+
+These numbers align with the results of the single model, discussed in Section \ref{sec:referring_expression_generation}.
+Hereby, the single model was only able to predict correct coordinates for the \emph{Dale-2} dataset with one distractor.
+The higher complexity for the agents that includes more layers in the model as well as the discrete bottleneck raises the bar of the task too high to be solved in this configuration.
+The second setup in this section tries to reduce the complexity, by not predicting precise geometric locations, but instead attending to larger areas.
+The results are summed up in Table \ref{tab:results:attention-reference-resolver-game}.
 
@@
     \caption{TODO: caption}
-    \label{TODO: label}
+    \label{tab:results:attention-reference-resolver-game}
 \end{table}
 
-In the final setup, the agents are tasked with communicating objects with fewer infused human knowledge.
-Table \ref{tab:results_dale_predictor_game} shows the results for the setup, in which the sender is pointed towards the target object with a human description based on the GRE algorithm.
-Hereby, the 'CLEVR single' dataset acts as a baseline, to test if the agents are able to predict coordinates of objects at all.
-In every configuration of the variables, the agents achieve a very high performance.
-The worst average distance across the test dataset is 10 pixels, which still points onto an object.
-Also the accuracy, which evaluates how many guesses of the receiver were pointing onto an object reflects this fact.
-All configuration achieve an accuracy higher than 96,7\%.
-% SD: Just occured to me: how do we measure accuracy if we are predicting numbers (numeric prediction)?
-% DK: same as in the experiments with one model (done)
-This aligns also with the results from the single neural models, where the average distance was similarly low.
-In general, this shows that the agents are able to predict coordinates together.
-However, the question arises if the message by the sender has actually an effect on the receivers' decision, or if the receiver learns to point towards the target coordinates on his own and the message is ignored.
-Having a look at the transferred messages, it in fact shows that the receiver learns to point towards the target object on its own.
-% SD: But how could it do so if the visual data is randomly distribution. It must rely on the message as that is the only source of information.
-% DK: it does it in the same way as the single model above (TODO)
-As in the experiment before, all communicated messages contain the same symbol independent of the input image.
+Opposed to the \emph{coordinate reference resolver}, the baseline can already find and attend to the correct regions in many cases.
+The probability mass, that is the sum of the predicted probabilities of all regions of the target object is higher than a uniform distribution (\approx 4,6\%) and a random guess of an object.
+It reaches 62,16\% on the \emph{Dale-2} dataset, 49,61\% on the \emph{Dale-5} dataset and 41,68\% on the \emph{CLEVR color} dataset.
+Interestingly, including a sender in the setup can improve the result for all dataset.
+Looking at the \emph{Dale-2} dataset, almost all configurations beat the baseline and achieve performances of over 90\%, the best configurations reach even 96\%.
+Only three configurations stay on the level of the baseline.
+When comparing the results, mostly the message length $n$ seems to have an influence on the performance.
+While configurations with $n=6$ can perform well, this is not constant.
+All three configurations that don't pass the baseline are allowed to produce message with $n=6$.
+$n \in \{3,4\}$ seem to help the agents the most to perform consistently well, but the difference to configurations with $n \in \{1,2\}$ is very small.
+In both cases, the target object is unambiguously identified.
+The small number of experiments doesn't allow definite conclusions on the influence of the vocabulary size $|V|$, though $|V|=2$ performs slightly worse than the remaining vocabulary sizes.
+A correlation between $n$ and $|V|$ is not identifiable.
 
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{ccc|ccc|ccc|ccc}
-        \toprule
-              &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){4-6}\cmidrule(lr){7-9}\cmidrule(lr){10-12}
-        $|V|$ & $h_{s}$ & $h_{r}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {10,1}                                    & {98,5\%}                            & {1}                                 & {36,5}         & {19,9\%}      & {1}             & {45,7}         & {14,4\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {6}                                       & {99\%}                              & {0}                                 & {38}           & {20,4\%}      & {1}             & {47,3}         & {10,8\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {9,7}                                     & {96,7\%}                            & {1}                                 & {37,3}         & {21,2\%}      & {1}             & {47,3}         & {11,3\%}      & {0}             \\
-        {100} & {10}    & {10}    & {7,7}                                     & {98,4\%}                            & {1}                                 & {40,4}         & {21,7\%}      & {1}             & {45,4}         & {10,8\%}      & {1}             \\
-        {100} & {100}   & {100}   & {7,5}                                     & {96,9\%}                            & {1}                                 & {40,1}         & {17,8\%}      & {1}             & {44,3}         & {11,8\%}      & {0}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the description coordinate predictor: $|V|$ are different vocabulary sizes and $h$ hidden sizes.}
-    \label{tab:results_dale_predictor_game}
-\end{table}
+On the \emph{Dale-5} dataset, the agents already have bigger problems to beat the baseline.
+Only 8 out of 30 configuration perform better and reach probability masses around 76\% to 84\%.
+However, the increase compared to the baseline is with around 30\% points as high as on the \emph{Dale-2} dataset.
+Smaller message lengths ($n \in \{1,2,3\}$) as well as a medium-sized vocabulary ($|V| \in \{10,16,50\}$) tend to help the agents more, to solve the task successfully.
+As before, no correlation is visible with the few successful games.
+That the agents struggle more with the \emph{Dale-5} dataset is not surprising.
+First, the larger number of distractors makes it more difficult for the receiver to focus, as can be seen already in the baseline performances.
+Additionally, in case the sender communicates the attributes of the target object to the receiver, more information needs to be encoded to uniquely describe the target object.
+This is naturally more complex to learn for the agents.
+Finally, since more objects are present, they are more likely clustered closer together, which can result in the identification of adjacent regions to the target regions.
 
-When the experiments are run on the 'Dale-2' dataset, the results are much worse.
-For the \emph{description coordinate predictor}, the average distance ranges from 36,5 pixels to 40,4 pixels.
-The configuration with a vocabulary size of only 10 symbols fares the best, while a vocabulary of 100 symbols produces the worst results.
-% SD: Remind the reader what is the size of the image so that they can interprert the distance relatively to that.
-% DK: TODO
-Still, the accuracy shows that around 19,9\% to 21,7\% of the guesses are on the target object.
-Here, the configurations with higher vocabulary sizes fare slightly better, but the differences are very small and likely due to other factors.
+The agents struggle the most on the \emph{CLEVR color} dataset.
+Hereby, only two configurations perform better than the baseline and reach a probability mass of around 64\% to 67\%.
+Both utilize a medium message length of $n \in \{3,4\}$ and a medium-sized vocabulary of $|V| \in \{10,16\}$.
+Interestingly, several configurations with short message lengths of $n \in \{1,2\}$ perform worse than the baseline.
+This indicates that there is communication between the agents, but it rather distracts the receiver from the target object towards the distractors.
+The same argument for a more difficult task when more objects are involved can be made for the \emph{CLEVR color} dataset.
+This dataset includes even up to 10 objects, present in the scene which increases the likelihood that the receiver focuses on a wrong object.
 
-The results for 'Dale-5' dataset are even worse, but are comparable with the results with a single neural model.
-Apparently, the agents are not able to communicate the target object, and the predictions by the receiver are general in the middle of the image, which results in an average distance of around 45 to 50 pixels.
-The differences of the mean distances are not very significant in this range, to allow an analysis of the different configurations.
-% SD: I don't understand this
-% DK: TODO
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cccc|ccc|ccc|ccc}
-        \toprule
-              &         &         &         & \multicolumn{3}{c}{\textbf{CLEVR single}} & \multicolumn{3}{c}{\textbf{Dale-2}} & \multicolumn{3}{c}{\textbf{Dale-5}}                                                                                                       \\\cmidrule(lr){5-7}\cmidrule(lr){8-10}\cmidrule(lr){11-13}
-        $|V|$ & $h_{s}$ & $h_{r}$ & $e_{s}$ & \textbf{Dist.}                            & \textbf{Acc.}                       & \textbf{length}                     & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} & \textbf{Dist.} & \textbf{Acc.} & \textbf{length} \\\midrule
-        {10}  & {10}    & {10}    & {1024}  & {10,8}                                    & {93,1\%}                            & {1}                                 & {34,8}         & {24,3\%}      & {0}             & {44,3}         & {11,8\%}      & {1}             \\
-        {10}  & {10}    & {10}    & {512}   & {9,3}                                     & {92\%}                              & {1}                                 & {36,3}         & {19,9\%}      & {0,7}           & {45,9}         & {12,7\%}      & {1}             \\
-        {13}  & {10}    & {10}    & {1024}  & {7,8}                                     & {96,8\%}                            & {1}                                 & {36,3}         & {20,2\%}      & {0}             & {45,4}         & {11,4\%}      & {1}             \\
-        {20}  & {10}    & {10}    & {1024}  & {6,6}                                     & {98,3\%}                            & {1}                                 & {37,8}         & {16,1\%}      & {1}             & {45,2}         & {11\%}        & {1}             \\
-        {100} & {10}    & {10}    & {1024}  & {5,2}                                     & {98,5\%}                            & {1}                                 & {37,4}         & {20,1\%}      & {1}             & {43,6}         & {16,7\%}      & {1}             \\
-        {100} & {100}   & {100}   & {1024}  & {12,5}                                    & {92,1\%}                            & {1}                                 & {36,5}         & {20,7\%}      & {1}             & {44,6}         & {12,7\%}      & {1}             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of the masked coordinate predictor: $|V|$ are different vocabulary sizes, $h$ hidden sizes and $e$ embedding sizes.}
-    \label{tab:results_masked_predictor_game}
-\end{table}
-
-Finally looking at the setup, when using the masked image as an input shows that the results are similarly bad as when using the encoded captions.
-This is easily explainable with the emerged language.
-For both the experiments using the encoded captions and the experiments using masked images as input, no meaningful symbols are transferred.
-% SD: A bit gloomy result. Negative result is okay but then we need an explanation why this is so and have some ideas what to change. What we can learn from these results?
-% DK: TODO
-Following, the receiver needs to solve the task alone and the different setups of the sender don't play any role in the overall success.
 % SD: This identifies that the problem lies in how the sender encodes the messages; what is its policy to generate longer strings or reuse the word.
@@
 % DK: TODO
-
-When comparing these results to the neural models in Section \ref{sec:reference_resolution} that are not part of a language game, it can be seen that the metrics are very similar for all datasets.
-The model was already not able to solve the task without the increased complexity of a language game.
-This therefore indicates that the challenge for the agents doesn't lie in grounding the extracted features in new symbols, but already in extracting the features in the first place.

### 2024-03-27 thesis/sections/6.1-object-identification.tex
learning curves, architectures

@@
 \subsubsection*{Results}
+
+\begin{figure}[ht!]
+    \centering
+    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-2_vocab-size.png}
+        \label{fig:learning-curve_discriminator_dale-2_vocab-size}
+    }
+    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-2_max-len.png}
+        \label{fig:learning-curve_discriminator_dale-2_max-len}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-5_vocab-size.png}
+        \label{fig:learning-curve_discriminator_dale-5_vocab-size}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-5_max-len.png}
+        \label{fig:learning-curve_discriminator_dale-5_max-len}
+    }
+    \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_colour_vocab-size.png}
+        \label{fig:learning-curve_discriminator_colour_vocab-size}
+    }
+    \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_colour_max-len.png}
+        \label{fig:learning-curve_discriminator_colour_max-len}
+    }
+    \caption{Learning curves of all discrimination games on each dataset.
+In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
+The baseline is marked in black.}
+    \label{fig:learning-curves_discriminator}
+\end{figure}
+
 When looking at the results of the experiments, it can be seen that the agents are successfully discriminating the objects across all datasets.
@@
 
-\cmtDK[inline]{baseline}
 \cmtDK[inline]{learning curve}

### 2024-03-27 thesis/sections/6.2-re-generation.tex
learning curves, architectures

@@
 For this the same sender as in the previous section is used.
-Based on the results, $h_s$ is fixed to $500$ and $e_s$ to $100$
+Based on the results, $h_s$ is fixed to $500$ and $e_s$ to $100$.
 However, the task and by that the receiver's architecture is adapted in two different ways.
@@
     \includegraphics[width=.7\linewidth]{figures/arch_caption_generator_game.png}
-    \caption{Simplified architecture of the caption generator game}
+    \caption{Simplified architecture of the referring expression generator game}
     \label{fig:caption_generator_game_architecture}
@@
 In other words, the sender might just learn to effectively repeat the target referring expressions without extracting the necessary features themselves.
-For that reason, in the second setup the receiver is tasked to predict only the attributes of the target object in the shape of a one-hot vector.
+For that reason, in the second setup, the \textbf{attribute generator} the receiver is tasked to predict only the attributes of the target object in the shape of a multi-hot vector (shown in Figure \ref{fig:attribute_generator_game_architecture}).
 By this, no natural language including its salience order is involved and the complexity of the complete game is reduced.
@@
 
-\cmtDK[inline]{figure}
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.55\linewidth]{figures/arch_attribute_generator_game.png}
+    \caption{Simplified architecture of the attribute generator game}
+    \label{fig:attribute_generator_game_architecture}
+\end{figure}
 
@@
 The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results for the first setup.
-When predicting one-hot vectors, an overall \textbf{accuracy} is reported that reflects if all attributes were predicted correctly.
+When predicting multi-hot vectors, an overall \textbf{accuracy} is reported that reflects if all attributes were predicted correctly.
 Additionally, the accuracy for each attribute is calculated separately to show which attributes give bigger challenges for the models.
@@
 \subsubsection*{Results}
+\begin{figure}[ht!]
+    \centering
+    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-2_vocab-size.png}
+        \label{fig:learning-curve_re-generator_dale-2_vocab-size}
+    }
+    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-2_max-len.png}
+        \label{fig:learning-curve_re-generator_dale-2_max-len}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-5_vocab-size.png}
+        \label{fig:learning-curve_re-generator_dale-5_vocab-size}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-5_max-len.png}
+        \label{fig:learning-curve_re-generator_dale-5_max-len}
+    }
+    \caption{Learning curves of all referring expression generation games on the \emph{Dale} datasets.
+In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
+The baseline is marked in black.}
+    \label{fig:learning-curves_re-generator}
+\end{figure}
+
 \begin{table}[ht]
@@
 
+\begin{figure}[ht!]
+    \centering
+    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-2_vocab-size.png}
+        \label{fig:learning-curve_oh-generator_dale-2_vocab-size}
+    }
+    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-2_max-len.png}
+        \label{fig:learning-curve_oh-generator_dale-2_max-len}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-5_vocab-size.png}
+        \label{fig:learning-curve_oh-generator_dale-5_vocab-size}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-5_max-len.png}
+        \label{fig:learning-curve_oh-generator_dale-5_max-len}
+    }
+    \caption{Learning curves of all attribute-predictor games on the \emph{Dale} datasets.
+The graphs show the accuracy of the \emph{shape}, since this highlights the learning effect best.
+In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
+The baseline is marked in black.}
+    \label{fig:learning-curves_oh-generator}
+\end{figure}
+
 \begin{table}[ht]
@@
 
-Finally, Table \ref{tab:results:attribute-predictor-game} shows the results for the games in which the receiver predicts a one-hot vector of the attributes.
+Finally, Table \ref{tab:results:attribute-predictor-game} shows the results for the games in which the receiver predicts a multi-hot vector of the attributes.
 Again, the baselines are games, the receiver plays alone.
@@
 First, the complexity of the receiver is greatly reduced.
-When only predicting one-hot vectors, much fewer layers and weights are involved that need to be trained and hence the models can learn the given task.
+When only predicting multi-hot vectors, much fewer layers and weights are involved that need to be trained and hence the models can learn the given task.
 Secondly, the complexity of the task is reduced as well.
 The generation of referring expressions not only includes the learning of the tokens (and their embeddings), but also their order in the referring expression and their salience for the incremental algorithm.
-Using one-hot vectors, both order and salience are irrelevant and only the attributes themselves are needed to be learned.
+Using multi-hot vectors, both order and salience are irrelevant and only the attributes themselves are needed to be learned.
 The goal of these experiments is now to see if the information of the sender improves the predictions of the receiver.
@@
 A reason why this happens might be the larger number of distractors.
-This number might hinder the receiver to converge and understand the task, by connecting the one-hot vector to the colors of the target object.
+This number might hinder the receiver to converge and understand the task, by connecting the multi-hot vector to the colors of the target object.
 While for the \emph{Dale} datasets the receiver has a choice between only 2 to 5 objects, they need to decide between up to 10 objects on the \emph{CLEVR color} dataset.

### 2024-03-27 thesis/sections/6.3-reference-resolution.tex
learning curves, architectures

@@
     \includegraphics[width=.7\linewidth]{figures/arch_coordinate_predictor_game.png}
-    \caption{Simplified architecture of the masked coordinate predictor game}
+    \caption{Simplified architecture of the coordinate predictor game}
     \label{fig:coordinate_predictor_game_architecture}
@@
 The dot product is calculated between each encoded region of the image and the encoded message, and the $softmax$ function is applied subsequently.
+Figure \ref{fig:attention_predictor_game_architecture} shows the architecture.
 The loss is calculated using \emph{binary cross entropy} between the predictions and ground truth regions.
 
-\cmtDK[inline]{figure}
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.7\linewidth]{figures/arch_attention_predictor_game.png}
+    \caption{Simplified architecture of the attention predictor game}
+    \label{fig:attention_predictor_game_architecture}
+    % SD: Again, why LSTM, are we inceremtally identifying the image while we are processing each input word?
+    % DK: because it's a message sequence of discrete symbols (done)
+\end{figure}
 
@@
 \subsection*{Results}
+\begin{figure}[ht!]
+    \centering
+    \subfigure[\emph{Dale-2} dataset]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_coordinate-predictor_dale-2.png}
+        \label{fig:learning-curve_coordinate-predictor_dale-2}
+    }
+    \subfigure[\emph{Dale-5} dataset]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_coordinate-predictor_dale-5.png}
+        \label{fig:learning-curve_coordinate-predictor_dale-5}
+    }
+    \subfigure[\emph{CELVR color} dataset]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_coordinate-predictor_colour.png}
+        \label{fig:learning-curve_coordinate-predictor_colour}
+    }
+    \caption{Learning curves of all discrimination games on each dataset. As almost no games beat the baseline, $|V|$ and $n$ are not highlighted separately. The baseline is marked in black.}
+    \label{fig:learning-curves_coordinate-predictor}
+\end{figure}
+
 \begin{table}[ht]
@@
 The results are summed up in Table \ref{tab:results:attention-reference-resolver-game}.
+
+\begin{figure}[ht!]
+    \centering
+    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-2_vocab-size.png}
+        \label{fig:learning-curve_attention-predictor_dale-2_vocab-size}
+    }
+    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-2_max-len.png}
+        \label{fig:learning-curve_attention-predictor_dale-2_max-len}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-5_vocab-size.png}
+        \label{fig:learning-curve_attention-predictor_dale-5_vocab-size}
+    }
+    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-5_max-len.png}
+        \label{fig:learning-curve_attention-predictor_dale-5_max-len}
+    }
+    \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_colour_vocab-size.png}
+        \label{fig:learning-curve_attention-predictor_colour_vocab-size}
+    }
+    \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
+        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_colour_max-len.png}
+        \label{fig:learning-curve_attention-predictor_colour_max-len}
+    }
+    \caption{Learning curves of all discrimination games on each dataset.
+In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
+The baseline is marked in black.}
+    \label{fig:learning-curves_attention-predictor}
+\end{figure}
 

### 2024-04-04 thesis/sections/6.1-object-identification.tex
learning curves

@@
 
-The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
+The language game is run on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
 A random guess corresponds to 50\% in the \emph{Dale-2} dataset, 20\% in the \emph{Dale-5} dataset and 13\% in the \emph{CLEVR color} dataset.
@@
     \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-2_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-2_vocab-size.png}
         \label{fig:learning-curve_discriminator_dale-2_vocab-size}
@@
     \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-2_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-2_max-len.png}
         \label{fig:learning-curve_discriminator_dale-2_max-len}
@@
     \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-5_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-5_vocab-size.png}
         \label{fig:learning-curve_discriminator_dale-5_vocab-size}
@@
     \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_dale-5_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-5_max-len.png}
         \label{fig:learning-curve_discriminator_dale-5_max-len}
@@
     \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_colour_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_colour_vocab-size.png}
         \label{fig:learning-curve_discriminator_colour_vocab-size}
@@
     \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_discriminator_colour_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_colour_max-len.png}
         \label{fig:learning-curve_discriminator_colour_max-len}
     }
-    \caption{Learning curves of all discrimination games on each dataset.
+    \caption{Learning curves of all object identification games on each dataset.
 In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
@@
 When looking at the results of the experiments, it can be seen that the agents are successfully discriminating the objects across all datasets.
-Tables \ref{tab:results_discriminator_dale-2} to \ref{tab:results_discriminator_color} give an overview over the accuracy scores of selected configurations.
+Figure \ref{fig:learning-curves_discriminator} gives an overview of the learning process of the agents.
+Each graph shows the learning curve of a configuration.
+In the left column, every color stands for configurations with a certain vocabulary size while in the right column a color corresponds to a message length.
+The rows show the three datasets.
+First of all, it can be seen that there are many configurations that outperform the baseline across all datasets and the agents are able to solve the task together.
+However, there are differences in the speed, the performance and the influence of vocabulary size and message length.
+
+The agents can solve the task on \emph{Dale-2} dataset very fast.
+Most of the configurations reach accuracies close to the final accuracy already after 2000 to 7000 games.
+If the agents learn to discriminate the object, they do so during these first games.
+After they reached these high accuracies, the learning is relatively unstable and can drop several times for up to 20\% points.
+However, the agents can recover after some thousand games to the original performance.
+The more games are played, the more stable the learning gets.
+The influence of the vocabulary size and message length is not as clear, but smaller vocabulary sizes tend to help the agents to learn faster.
+On the \emph{Dale-5} dataset, there is already a wider spread in performances and the learning is slower than on the \emph{Dale-2} dataset.
+The agents learn fast up to an accuracy of around 50\%, but for most configurations it then slows down, and the accuracy improves until the cutoff after 128.000 games.
+When the agents are given a vocabulary size of $|V|=2$, for several runs, the learning even stops after they reach around 50\% accuracy or progresses only very slowly.
+With a vocabulary size of $|V|=100$, the agents reach the 50\% some thousand games after the remaining vocabulary sizes, but the learning goes on afterwards.
+While there is no visible influence of the message length on the learning speed, all configurations that stay at an accuracy between 45\% and 65\% have a message length of $n=1$.
+Interestingly, there is also one configuration that diverges from the baseline very late after almost 80.000 games.
+Finally, on the \emph{CLEVR color} dataset, difference in learning speed are again more apparent.
+Agents tend to learn the fastest with a vocabulary size of $|V|=2$, the accuracy however almost stagnates after the initial boost to around 40\%.
+When agents are given a vocabulary of size $|V| \in \{10,16\}$, the initial learning takes place several thousand games later, but therefore reaches higher final accuracies.
+As seen already for the \emph{Dale-5} dataset, most of the configurations converge in the first 20.000 games, but there are several converge much later.
+An influence of the message length on the speed is again not visible.
+% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
+As David pointed out it may be that convergence is unstable and changes happen in between.
+% DK: done
+
+Tables \ref{tab:results_discriminator_dale-2} to \ref{tab:results_discriminator_color} give an overview over the final accuracy scores after 128.000 games of selected configurations.
 For each dataset there are many configurations that outperform the baseline models by a high margin.
 However, there are still clear differences between the accuracy scores of each dataset.
-
-\cmtDK[inline]{learning curve}
-% SD: As we discussed at Semdial, it would be good to have a learning curves here that would show how fast or slow the languages converge.
-As David pointed out it may be that convergence is unstable and changes happen in between.
-% DK: TODO
 
@@
     \end{tabular}
-    \caption{Selected results of the object identification games on the Dale-2 dataset with $e_s=100$ and $e_r=500$}
+    \caption{Selected results of the object identification games on the Dale-2 dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
     \label{tab:results_discriminator_dale-2}
@@
     \end{tabular}
-    \caption{Selected results of the object identification games on the Dale-5 dataset with $e_s=100$ and $e_r=500$}
+    \caption{Selected results of the object identification games on the Dale-5 dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
     \label{tab:results_discriminator_dale-5}
@@
     \end{tabular}
-    \caption{Selected results of the object identification games on the CLEVR color dataset with $e_s=100$ and $e_r=500$}
+    \caption{Selected results of the object identification games on the CLEVR color dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
     \label{tab:results_discriminator_color}

### 2024-04-04 thesis/sections/6.2-re-generation.tex
learning curves

@@
 In the first setup, the \textbf{referring expression generator}, the receiver is tasked to describe the target object with natural language following the \emph{incremental GRE-algorithm} based on the sender's message (see Figure \ref{fig:caption_generator_game_architecture}).
-This follows the approach of the approach of the \emph{RE generators} described in section \ref{sec:referring_expression_generation}.
+This follows the approach of the approach of the \emph{RE generators} described in Section \ref{sec:referring_expression_generation}.
 The whole scene is encoded using the \emph{image encoder} submodule, projecting it to the image encoding dimension $e_{ri}=100$.
@@
 % DK: TODO
-The same metrics as in the section \ref{sec:referring_expression_generation} are used to evaluate the results for the first setup.
+The same metrics as in the Section \ref{sec:referring_expression_generation} are used to evaluate the results for the first setup.
 When predicting multi-hot vectors, an overall \textbf{accuracy} is reported that reflects if all attributes were predicted correctly.
@@
     \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-2_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-2_vocab-size.png}
         \label{fig:learning-curve_re-generator_dale-2_vocab-size}
@@
     \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-2_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-2_max-len.png}
         \label{fig:learning-curve_re-generator_dale-2_max-len}
@@
     \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-5_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-5_vocab-size.png}
         \label{fig:learning-curve_re-generator_dale-5_vocab-size}
@@
     \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_re-generator_dale-5_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-5_max-len.png}
         \label{fig:learning-curve_re-generator_dale-5_max-len}
@@
     \caption{Learning curves of all referring expression generation games on the \emph{Dale} datasets.
+On the \emph{CLEVR color} dataset, the learning curves don't diverge from the baseline and are therefore left out.
 In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
@@
 \end{figure}
+
+As in the previous section, Figure \ref{fig:learning-curves_re-generator} shows the learning curves of all configurations of the \textbf{referring expression generator}.
+The baselines are shown in black.
+Hereby, the sender is sending random messages, so that the receiver needs to solve the task on their own.
+In all configurations that don't pass this baseline, no meaningful messages are exchanged, and no language emerged.
+If configurations perform better on the other hand means that the agents learned to encode some meaning in their messages.
+While for both \emph{Dale} dataset, several configuration diverge from the baseline, in other words, learning takes place, all learning curves align with the baseline for the \emph{CLEVR color} dataset and are therefore left out.
+Looking at the learning curves for the \emph{Dale-2} dataset, most of the configurations can decrease the loss compared to the baseline faster on to a bigger extend.
+Fast learning takes place in the first 10.000 to 15.000 games and decelerates afterwards.
+Also on this task, there are several configurations that converge much later than the initial fast decrease of the loss.
+Most of these configurations diverge from the baseline between 15.000 to 35.000 games, but some even need up to 70.000 games.
+While the vocabulary size now seems to have a lower impact on the learning speed, configurations that converge later tend to have a message length of $n=6$.
+For the \emph{Dale-2} dataset a clear distinction between configurations that align with the baseline and configurations that diverge from the baseline is visible.
+For the \emph{Dale-5} dataset, however, all learning curves are closer together and success of learning is not instantly apparent.
+Nonetheless, several configurations (with low- to medium-sized vocabularies) start to converge earlier than the baseline.
+Configurations with $|V|=2$ start to do so already after 5.000 games, with $|V|\in\{10,16\}$ after around 7.000 games and with $|V|=50$ after around 12.000 games.
+With $|V|=100$, almost no configuration diverges from the baseline.
+The message length doesn't seem to have an impact.
+Interestingly, there are several configurations that perform worse than the baseline, and communication between the agents seems to hinder the learning.
+This happens across all vocabulary sizes and message lengths.
 
@@
     \end{tabular}
-    \caption{Overall accuracies (Acc.), F1-score (F1) and non-target accuracies (NT) in \% of the RE generator games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
+    \caption{Overall accuracies (Acc.), F1-score (F1) and non-target accuracies (NT) in \% of the RE generator after 128.000 games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
     \label{tab:results:re-generator-game}
@@
 Table \ref{tab:results:re-generator-game} shows the \emph{overall accuracies}, \emph{F1-scores} and \emph{non-target accuracies} across all datasets for different message lengths $n$ and vocabulary sizes $|V|$.
-The baselines are shown in the first row.
-Hereby, the sender is sending random messages, so that the receiver needs to solve the task on their own.
-In all configurations that don't pass this baseline, no meaningful messages are exchanged, and no language emerged.
-If configurations perform better on the other hand means that the agents learned to encode some meaning in their messages.
-
 When first having a look at the results of the \emph{Dale-2} dataset, it can be seen that the baseline already identifies correct referring expressions in 41,8\% of the samples.
@@
 
-Since the single model in section \ref{sec:referring_expression_generation} already struggled with predicting the \emph{color} on the \emph{CLEVR color} dataset, it is no surprise that it performs badly in the more complex setup of language games.
+Since the single model in Section \ref{sec:referring_expression_generation} already struggled with predicting the \emph{color} on the \emph{CLEVR color} dataset, it is no surprise that it performs badly in the more complex setup of language games.
 Hereby, the agents are not able to communicate anything and the receiver only generates two colors (green and purple) for all shown samples.
@@
 While the agents easily learn that the \emph{size} is never part of the referring expression for the \emph{CLEVR color} dataset, they struggle with the \emph{Dale} datasets.
-As in section \ref{sec:referring_expression_generation}, the agents again struggle to predict the correct length of the referring expression.
+As in Section \ref{sec:referring_expression_generation}, the agents again struggle to predict the correct length of the referring expression.
 Especially, the agents predict too short referring expressions as the precision of the \emph{<pad>} token in lower than its recall.
@@
     \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-2_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-2_vocab-size.png}
         \label{fig:learning-curve_oh-generator_dale-2_vocab-size}
@@
     \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-2_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-2_max-len.png}
         \label{fig:learning-curve_oh-generator_dale-2_max-len}
@@
     \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-5_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-5_vocab-size.png}
         \label{fig:learning-curve_oh-generator_dale-5_vocab-size}
@@
     \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_oh-generator_dale-5_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-5_max-len.png}
         \label{fig:learning-curve_oh-generator_dale-5_max-len}
@@
     \end{tabular}
-    \caption{Accuracies for each attribute in \% of the attribute predictor games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
+    \caption{Accuracies for each attribute in \% of the attribute predictor after 128.000 games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
     \label{tab:results:attribute-predictor-game}
@@
 
-Finally, Table \ref{tab:results:attribute-predictor-game} shows the results for the games in which the receiver predicts a multi-hot vector of the attributes.
-Again, the baselines are games, the receiver plays alone.
-Hereby, it can be seen that the receiver can already predict the \emph{shape} and \emph{color}  very well: 88,91\% and 92\% respectively on the \emph{Dale-2} dataset, 80,78\% and 89,38\% respectively on the \emph{Dale-5} dataset, and 100\% and 90,78\% respectively on the \emph{CLEVR color} dataset.
+Finally, Figure \ref{fig:learning-curves_oh-generator} shows the learning curves for the \textbf{attribute generator}.
+The measure of learning is hereby the accuracy of the \emph{shape}.
+The reason for this is that the \emph{size} is not learned at all, while also the \emph{color} diverges only very slightly from baseline.
+Communication is therefore impacting mostly the \emph{shape}, and the influence of the vocabulary size and message length is shown in the figures.
+On the \emph{Dale-2} dataset, configuration that learn the \emph{shape} quickly tend to use smaller vocabularies.
+However, there are configurations with $|V| \in \{16,10\}$ that need 15.000 to 20.000 games, but outperform the configurations which learned faster.
+A similar result is visible for the message length, where shorter messages help the agents to learn faster while longer messages need more games.
+Looking at the \emph{Dale-5} dataset, more configurations are able to learn the \emph{shape}.
+Most of the configurations again outperform the baseline already after 5.000 to 10.000 games, but the learning decelerates shortly after.
+Interestingly, most of the configurations reach a final accuracy of the \emph{shape} after 128.000 games of 85\% to 97\%.
+However, the learning progress is different across the configurations.
+Several configurations progress quickly only to around 65\% with a following relatively steep learning over the remaining 120.000 games.
+A second group of configurations has an initial learning to around 80\% and then progresses slower.
+On configuration even reaches over 95\% after 15.000 games, but then stagnates until the cutoff.
+Looking at the vocabulary size and message length, no consistent influence is visible.
+However, agents seem to learn slower with $|V| \in \{50,100\}$, while $|V| \in \{2,16\}$ tends to increase the speed.
+Opposed, a short message length tends to decrease the speed.
+
+Table \ref{tab:results:attribute-predictor-game} shows the results for the games in which the receiver predicts a multi-hot vector of the attributes.
+In the results for the baseline, it can be seen that the receiver can already predict the \emph{shape} and \emph{color}  very well: 88,91\% and 92\% respectively on the \emph{Dale-2} dataset, 80,78\% and 89,38\% respectively on the \emph{Dale-5} dataset, and 100\% and 90,78\% respectively on the \emph{CLEVR color} dataset.
 The \emph{size} however corresponds to a random guess on the \emph{Dale} dataset and only achieves 76,72\% on the \emph{CLEVR color} dataset.

### 2024-04-04 thesis/sections/6.3-reference-resolution.tex
learning curves

@@
 By this, the agents can't just reuse these human referring expressions, but need to learn to generate them themselves during the language games just based on the visual input.
-This is done similarly to the single model reference resolution task, described in section \ref{sec:reference_resolution}.
+This is done similarly to the single model reference resolution task, described in Section \ref{sec:reference_resolution}.
 As before, the sender has the same architecture as in the previous language games, encoding the bounding boxes of all objects.
@@
 First, in the \textbf{coordinate reference resolver}, the agents are tasked to predict the center coordinates of the target object (see Figure \ref{fig:coordinate_predictor_game_architecture}).
-The receiver is the same in both setups and is based on the \emph{reference resolver} + referring expressions, described in section \ref{sec:reference_resolution}.
+The receiver is the same in both setups and is based on the \emph{reference resolver} + referring expressions, described in Section \ref{sec:reference_resolution}.
 Again, instead of the included LSTM, the receiver uses the \emph{EGG} LSTM with the hidden size $h_r=500$ and $e_r=100$ to encode the message.
@@
     \subfigure[\emph{Dale-2} dataset]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_coordinate-predictor_dale-2.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_coordinate-predictor_dale-2.png}
         \label{fig:learning-curve_coordinate-predictor_dale-2}
@@
     \subfigure[\emph{Dale-5} dataset]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_coordinate-predictor_dale-5.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_coordinate-predictor_dale-5.png}
         \label{fig:learning-curve_coordinate-predictor_dale-5}
@@
     \subfigure[\emph{CELVR color} dataset]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_coordinate-predictor_colour.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_coordinate-predictor_colour.png}
         \label{fig:learning-curve_coordinate-predictor_colour}
     }
-    \caption{Learning curves of all discrimination games on each dataset. As almost no games beat the baseline, $|V|$ and $n$ are not highlighted separately. The baseline is marked in black.}
+    \caption{Learning curves of all coordinate reference resolving games on each dataset.
+As almost no games beat the baseline, $|V|$ and $n$ are not highlighted separately.
+The baseline is marked in black.}
     \label{fig:learning-curves_coordinate-predictor}
@@
     \end{tabular}
-    \caption{TODO: caption}
+    \caption{Mean distance (loss) in pixels and accuracy in \% of the coordinate reference resolver after 128.000 games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
     \label{tab:results:coordinate-reference-resolver-game}
@@
 
-Table \ref{tab:results:coordinate-reference-resolver-game} shows the results of the coordinate reference resolver on different datasets.
+Figure \ref{fig:learning-curves_coordinate-predictor} show the learning curves for the \textbf{coordinate reference resolver}.
+As visible, all configurations perform exactly as the baselines across all datasets.
 The baselines show the predictions of the receiver without information of the sender, and correspond to the structural bias in the dataset.
+This shows, that the agents don't communicate successfully in any of the configurations to solve the task and predict correct coordinates.
+Table \ref{tab:results:coordinate-reference-resolver-game} shows the final results after 128.000 games on different datasets.
 When looking at how the agents fare with each dataset, not much general improvement can be seen.
@@
 
-Looking at the results on the \emph{Dale-5}, no improvements for the mean distance nor the accuracy can be seen when a sender is involved.
+Looking at the results on the \emph{Dale-5} dataset, no improvements for the mean distance nor the accuracy can be seen when a sender is involved.
 However, while the loss stays again constant, the accuracy drops lower in many configurations, up to 7\% points to an accuracy of 23,26\%.
@@
 The higher complexity for the agents that includes more layers in the model as well as the discrete bottleneck raises the bar of the task too high to be solved in this configuration.
-The second setup in this section tries to reduce the complexity, by not predicting precise geometric locations, but instead attending to larger areas.
-The results are summed up in Table \ref{tab:results:attention-reference-resolver-game}.
+The second setup in this section, the \textbf{attention reference resolver}, tries to reduce the complexity, by not predicting precise geometric locations, but instead attending to larger areas.
+The learning curves are shown in Figure \ref{fig:learning-curves_attention-predictor}.
+
+As expected, the task is solver better by the agents, and they communicate successfully across all datasets.
+Compared to the experiments in the previous sections, the learning is still generally slower, and the configurations diverge from the baseline overall later.
+However, when the agents start to learn to communicate, the probability is boosted instantly to a higher level, where it again learns at a slower speed parallel to the baseline.
+On the \emph{Dale-2} dataset, the boost is around 40\% points.
+Most of the learning takes place in the first 40.000 games, but there are also two configurations that increase the performance very late after 70.000 and 105.000 games respectively.
+Hereby, agents tend to learn faster the smaller their vocabulary size is.
+The same applies as well to the message length.
+Using the \emph{Dale-5} dataset, the probability masses are boosted around 30\% points when the agents start to communicate successfully.
+Compared to the \emph{Dale-2} dataset, fewer configurations start to converge, while most achieve performances close to the baseline.
+The smaller number of learning curves makes the analysis more difficult, but the same trend about the vocabulary size and the message length is still visible.
+Interestingly, only one configuration with $|V|=2$ beats the baseline, but behaves relatively unstable over the remaining training.
+On the other hand no configuration with $|V|=100$ is successful.
+This indicates that one symbol too few to encode all meaning, but too many symbols are too difficult to learn.
+This hypothesis is amplified by the results on the \emph{CLEVR color} dataset.
+Only two configurations beat the baseline, both with a medium-sized vocabulary size and message length.
+In both cases, the learning takes place relatively late, after 15.000 and 30.000 games respectively.
 
@@
     \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-2_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-2_vocab-size.png}
         \label{fig:learning-curve_attention-predictor_dale-2_vocab-size}
@@
     \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-2_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-2_max-len.png}
         \label{fig:learning-curve_attention-predictor_dale-2_max-len}
@@
     \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-5_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-5_vocab-size.png}
         \label{fig:learning-curve_attention-predictor_dale-5_vocab-size}
@@
     \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_dale-5_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-5_max-len.png}
         \label{fig:learning-curve_attention-predictor_dale-5_max-len}
@@
     \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_colour_vocab-size.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_colour_vocab-size.png}
         \label{fig:learning-curve_attention-predictor_colour_vocab-size}
@@
     \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
-        \includegraphics[width=0.46\linewidth]{figures/learning-curve_attention-predictor_colour_max-len.png}
+        \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_colour_max-len.png}
         \label{fig:learning-curve_attention-predictor_colour_max-len}
     }
-    \caption{Learning curves of all discrimination games on each dataset.
+    \caption{Learning curves of all attention reference resolver games on each dataset.
 In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
@@
     \end{tabular}
-    \caption{TODO: caption}
+    \caption{Probability masses of the attention reference resolver after 128.000 games: $n$ are different maximum message lengths and $|V|$ are different vocabulary sizes.}
     \label{tab:results:attention-reference-resolver-game}
@@
 
+The final probability masses after 128.000 games are summed up in Table \ref{tab:results:attention-reference-resolver-game}.
 Opposed to the \emph{coordinate reference resolver}, the baseline can already find and attend to the correct regions in many cases.

### 2024-04-05 thesis/sections/5.1-object-identification.tex
rename

@@
 The first trend that is visible is that fewer distractor increase the accuracy of the model.
-The model achieves almost perfect accuracy in all configurations for the \emph{Dale 2} dataset with only two objects.
-With five objects in the \emph{Dale-5} dataset, the accuracy drops to 95\% and with a maximum possible number of 10 objects in the \emph{CLEVR color} dataset the model only achieves 94\% in the best configuration.
+The model achieves almost perfect accuracy in all configurations for the 'Dale 2' dataset with only two objects.
+With five objects in the 'Dale-5' dataset, the accuracy drops to 95\% and with a maximum possible number of 10 objects in the 'CLEVR color' dataset the model only achieves 94\% in the best configuration.
 This is not very surprising, since the model has a higher chance of predicting a wrong object, given the random baselines of 50\%, 20\% and 10\%.

### 2024-04-05 thesis/sections/5.2-re-generation.tex
rename

@@
 Again, each of these datasets increases the complexity of the description.
-While the referring expression for the \emph{Dale-2} datasets are generally shorter, expressions of the \emph{Dale-5} datasets need to be more specific and use more attributes.
+While the referring expression for the 'Dale-2' datasets are generally shorter, expressions of the 'Dale-5' datasets need to be more specific and use more attributes.
 Additionally, the model needs to attend to many more locations in the image at the same time to find discriminating factors between those.
-Furthermore, the models are trained on the \emph{CLEVR color} dataset.
+Furthermore, the models are trained on the 'CLEVR color' dataset.
 In this dataset, the focus lies mainly on only one discriminating attribute, the color.
@@
 \subsubsection*{Results}
-Table \ref{tab:results:bb-re-generator} shows the \emph{overall accuracy}, \emph{F1 scores} for each word and the \emph{non-target accuracies} of the \textbf{bounding box RE generator} when trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
+Table \ref{tab:results:bb-re-generator} shows the \emph{overall accuracy}, \emph{F1 scores} for each word and the \emph{non-target accuracies} of the \textbf{bounding box RE generator} when trained on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
 As can be seen, the overall accuracies, in other words perfect matches of the generated referring expression depend very much on the dataset.
-With the \emph{Dale-2} dataset, the model can achieve perfect matches in 99\% of the cases in its best configuration.
-Also the \emph{CLEVR color} dataset allows the model to predict the correct referring expression in 93\% of the samples.
-Opposed to that the model can only generate perfect referring expressions in 69\% of the samples of the \emph{Dale-5} dataset.
-Looking at the \emph{non-target accuracy}, one of the problems can be identified for the \emph{CLEVR color} dataset.
+With the 'Dale-2' dataset, the model can achieve perfect matches in 99\% of the cases in its best configuration.
+Also the 'CLEVR color' dataset allows the model to predict the correct referring expression in 93\% of the samples.
+Opposed to that the model can only generate perfect referring expressions in 69\% of the samples of the 'Dale-5' dataset.
+Looking at the \emph{non-target accuracy}, one of the problems can be identified for the 'CLEVR color' dataset.
 Summing the \emph{non-target accuracy} and the \emph{overall accuracy}, one gets the accuracy that any of the shown object was described independently if it was the target object.
-For the \emph{CLEVR color} dataset, this score lies at 96\% to 97\% in the best configurations, which means that the model described a shown object for almost all the samples.
+For the 'CLEVR color' dataset, this score lies at 96\% to 97\% in the best configurations, which means that the model described a shown object for almost all the samples.
 In 3\% to 4\% of the cases, the model only picked the wrong object to describe.
-This looks different for the \emph{Dale-5} dataset.
+This looks different for the 'Dale-5' dataset.
 Here, the model describes only in 71\% of the cases one of the shown objects, 69\% the target object and 2\% a distractor.
@@
 
-When looking at the different configurations, in especially different values for $e$, $LSTM_o$ and $LSTM_e$, it can be seen that the results are not very different for the \emph{Dale-2} and \emph{CLEVR color} datasets.
-However, bigger effects can be identified for the \emph{Dale-5} dataset.
+When looking at the different configurations, in especially different values for $e$, $LSTM_o$ and $LSTM_e$, it can be seen that the results are not very different for the 'Dale-2' and 'CLEVR color' datasets.
+However, bigger effects can be identified for the 'Dale-5' dataset.
 Here, a low output dimension of the LSTM $LSTM_o$ tends to give lower scores.
@@
 
-Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results and especially what mistakes the model is making for both the \emph{Dale-5} and \emph{CLEVR color} datasets.
+Tables \ref{tab:results:bb-re-generator_size-shape} and \ref{tab:results:bb-re-generator_color} give a more detailed insight in the results and especially what mistakes the model is making for both the 'Dale-5' and 'CLEVR color' datasets.
 They list \emph{precision} and \emph{recall} metrics for each token for the best configuration of the model with $e=100$, $LSTM_o=500$ and $LSTM_e=30$.
 The tokens are grouped by attribute and also show the metrics averaged over each of the attributes.
-Since the overall results for the \emph{Dale-2} dataset are already close to perfect, the focus for this analysis lies on the remaining two datasets.
+Since the overall results for the 'Dale-2' dataset are already close to perfect, the focus for this analysis lies on the remaining two datasets.
 The metrics of the \emph{<pad>} token indicate if the model produced the correct length of the referring expression, in other words if it was able to determine which attributes are necessary to discriminate the target object from the distractors.
-For the \emph{CLEVR color} dataset, the scores are perfect.
-This is not surprising, because all referring expressions for the \emph{CLEVR color} dataset consist of exactly two attributes, shape and color, and the first generated token will always be the only \emph{<pad>} token in the referring expression (corresponding to the unspecified size).
+For the 'CLEVR color' dataset, the scores are perfect.
+This is not surprising, because all referring expressions for the 'CLEVR color' dataset consist of exactly two attributes, shape and color, and the first generated token will always be the only \emph{<pad>} token in the referring expression (corresponding to the unspecified size).
 The \emph{<pad>} token is therefore easy to learn.
-For the \emph{Dale-5} dataset, the model struggles more to predict the correct length of the referring expression.
+For the 'Dale-5' dataset, the model struggles more to predict the correct length of the referring expression.
 
 When looking at the tokens for the shape, it can be seen that the model is able to identify it very well across all datasets.
-The model predicts the correct shape for all samples using the \emph{CLEVR color} dataset, while both \emph{precision} and \emph{recall} lie around 98,3\% when using the \emph{Dale-5} dataset.
+The model predicts the correct shape for all samples using the 'CLEVR color' dataset, while both \emph{precision} and \emph{recall} lie around 98,3\% when using the 'Dale-5' dataset.
 Even though the score is almost perfect, the slight difference might stem from the fact that all distractors have the same shape in the first case, while distractors can be different in the second case.
@@
 
-For the color attribute, the metrics drop significantly for both \emph{Dale-5} and \emph{CLEVR color} to an average of around 93\%.
+For the color attribute, the metrics drop significantly for both 'Dale-5' and 'CLEVR color' to an average of around 93\%.
 Hereby, no meaningful difference can be seen across the datasets, but there are differences between the colors.
@@
 Finally, the size is the most difficult attribute to predict for the model.
-Apart from the \emph{CLEVR color} dataset, where a size never needs to be predicted and also is never predicted, the metrics for the prediction of size tokens are the lowest across all tokens.
-They are the only mistakes, the model makes, when exposed to the \emph{Dale-2} dataset and the average \emph{precision} lies around 23\% below the average of predictions of the color for the \emph{Dale-5} dataset, while the average \emph{recall} lies around 28,82\% below.
+Apart from the 'CLEVR color' dataset, where a size never needs to be predicted and also is never predicted, the metrics for the prediction of size tokens are the lowest across all tokens.
+They are the only mistakes, the model makes, when exposed to the 'Dale-2' dataset and the average \emph{precision} lies around 23\% below the average of predictions of the color for the 'Dale-5' dataset, while the average \emph{recall} lies around 28,82\% below.
 The reason why the \emph{precision} is higher than the \emph{recall} is the \emph{<pad>} token, which is predicted very often instead of a token specifying the size.
@@
 However, it can't decide which object to describe.
-Most clearly, this is shown with the \emph{Dale-2} dataset where the sum lies around 90\%.
+Most clearly, this is shown with the 'Dale-2' dataset where the sum lies around 90\%.
 In almost all samples, the  model describes an object perfectly, but in 45\% of the cases it is the target object, in 45\% the distractor.
@@
 The algorithm then generates two identical referring expressions which can affect the \emph{non-target accuracy}.
-Indeed, while the average number of distractors per scene in the \emph{CLEVR color} dataset is 7,5, the average number of unique distractors lies at 2,7.
+Indeed, while the average number of distractors per scene in the 'CLEVR color' dataset is 7,5, the average number of unique distractors lies at 2,7.
 As seen later, the model learns to correctly never predict a \emph{size} and always predict the correct \emph{shape}.
 Given this fact, a random guess of a color would predict in $\frac{1}{8} * 2,7 = 33,9\%$ of the cases a unique distractor while in $\frac{1}{8}=12,5\%$ of the cases it would predict the target object.
-This corresponds the \emph{non-target accuracies} and \emph{overall accuracies} the \emph{CLEVR color} dataset.
+This corresponds the \emph{non-target accuracies} and \emph{overall accuracies} the 'CLEVR color' dataset.
 The high \emph{non-target accuracy} is therefore misleading and doesn't represent that the model identified other objects in the scene.
@@
 A first look at the sum of the \emph{overall accuracy} and the \emph{non-target accuracy} shows the model is able to extract the attributes from the shown objects.
-For the \emph{Dale-2} dataset, this number lies at 90\% for the best configuration, at 51\% for the \emph{Dale-5} dataset and at 55\% for the \emph{CLEVR color} dataset.
+For the 'Dale-2' dataset, this number lies at 90\% for the best configuration, at 51\% for the 'Dale-5' dataset and at 55\% for the 'CLEVR color' dataset.
 Those sums correspond more or less to the sums of the model without masking shown in Table \ref{tab:results:basic-re-generator}.
@@
 However, the balance in between the two accuracy scores shifts towards the \emph{overall accuracy}, meaning that the model is now able to identify the correct target object more often.
-While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the \emph{Dale-2} dataset, the model only reaches 36\% for the \emph{Dale-5} dataset and 29\% for the \emph{CLEVR color} dataset.
+While the model still achieves 85\% \emph{overall accuracy} in its best configuration, for the 'Dale-2' dataset, the model only reaches 36\% for the 'Dale-5' dataset and 29\% for the 'CLEVR color' dataset.
 The scores for all datasets are all well above the \emph{basic RE generator} baselines of 45\%, 16\% and 15\% respectively, while in all cases the \emph{non-target accuracies} drop lower.
@@
 Comparing it to the less complex \emph{bounding box RE generator}, the results are much less accurate.
-This applies to all datasets, even though it is most apparent for the \emph{CLEVR color} dataset.
+This applies to all datasets, even though it is most apparent for the 'CLEVR color' dataset.
 Given the highly increased complexity of the task, the model fares well in comparison on the \emph{Dale} datasets.
 
-Moreover, the referring expression generated for the \emph{Dale-2} dataset are also efficient in the sense that they follow the incremental GRE-algorithm and only use necessary discriminative attributes.
+Moreover, the referring expression generated for the 'Dale-2' dataset are also efficient in the sense that they follow the incremental GRE-algorithm and only use necessary discriminative attributes.
 Features of both target object and distractor are therefore extracted and associated with the vocabulary.
 Striking in these results is also the \emph{non-target accuracy}, which is much higher than for the \emph{bounding box RE generator}.
-For the \emph{Dale-2} dataset, it arrives at 11\% in one configuration, for the \emph{Dale-5} dataset, the model describes for at least 12\% of the samples a distractor, for one configuration even in over 25\% of the cases.
-The \emph{non-target accuracy} of the \emph{CLEVR color} is misleading, and is therefore not considered.
+For the 'Dale-2' dataset, it arrives at 11\% in one configuration, for the 'Dale-5' dataset, the model describes for at least 12\% of the samples a distractor, for one configuration even in over 25\% of the cases.
+The \emph{non-target accuracy} of the 'CLEVR color' is misleading, and is therefore not considered.
 This shows that the model still struggles to learn which of the objects in the scene is the target object, even though masking is applied.
 
-Looking at the effects of the LSTM output size $LSTM_o$ and the LSTM embedding size $LSTM_e$, a difference between the \emph{Dale} datasets and the \emph{CLEVR color} dataset can be identified.
+Looking at the effects of the LSTM output size $LSTM_o$ and the LSTM embedding size $LSTM_e$, a difference between the \emph{Dale} datasets and the 'CLEVR color' dataset can be identified.
 A higher $LSTM_o$ tends to increase the performance of the model, when presented with the \emph{Dale} datasets until a size of 1500 to 2000 dimensions.
@@
 With an $LSTM_o$ of around 1500 to 2000, the model seems to fare slightly better with higher $LSTM_e$ of 15 or 30 dimensions.
-For the \emph{CLEVR color} dataset, the results look different.
+For the 'CLEVR color' dataset, the results look different.
 Opposed to the former datasets, a lower $LSTM_o$ helps the model to describe the correct target object.
@@
 Tables \ref{tab:results:masked-re-generator_size-shape} and \ref{tab:results:masked-re-generator_color} shows the precision and recall for each token.
-The metrics on the \emph{<pad>} token show that the model always predicts the correct length of the referring expression when presented with the \emph{CLEVR color} dataset.
-For the \emph{Dale-2} dataset, the model is able to do this in almost all the cases, while the model struggles the most with the \emph{Dale-5} dataset.
+The metrics on the \emph{<pad>} token show that the model always predicts the correct length of the referring expression when presented with the 'CLEVR color' dataset.
+For the 'Dale-2' dataset, the model is able to do this in almost all the cases, while the model struggles the most with the 'Dale-5' dataset.
 While this was similar to the \emph{bounding box RE generator}, the effect here is amplified.
 Furthermore, the easiest attribute to predict is the shape.
-Again, the metrics for the \emph{Dale-2} and \emph{CLEVR color} dataset are almost the same, but the model is much less able to predict the correct shape for images from the \emph{Dale-5} dataset with only around 85\%.
+Again, the metrics for the 'Dale-2' and 'CLEVR color' dataset are almost the same, but the model is much less able to predict the correct shape for images from the 'Dale-5' dataset with only around 85\%.
 Hereby, the precision is the highest for predicting \emph{cubes} and lowest for predicting \emph{cylinders}, but this varies across runs.
@@
 While the metrics all lied above 90\% for the \emph{bounding box RE generator}, now none of datasets passes this limit.
-With the \emph{Dale-2} dataset, the precision and recall lie around 85\%, with the \emph{Dale-5} dataset around 70\% and around 30\% with the \emph{CLEVR color} dataset.
-The drop for the 'Dale' datasets is expected, because of the higher complexity of the task, but the model's performance on the \emph{CLEVR color} dataset is much worse.
+With the 'Dale-2' dataset, the precision and recall lie around 85\%, with the 'Dale-5' dataset around 70\% and around 30\% with the 'CLEVR color' dataset.
+The drop for the 'Dale' datasets is expected, because of the higher complexity of the task, but the model's performance on the 'CLEVR color' dataset is much worse.
 Only every third prediction is correct.
@@
 Finally, the size is again the most difficult attribute to predict.
-As the \emph{bounding box RE generator}, the \emph{masked RE generator} learns correctly to never predict a size for the \emph{CLEVR color} dataset.
+As the \emph{bounding box RE generator}, the \emph{masked RE generator} learns correctly to never predict a size for the 'CLEVR color' dataset.
 Interestingly is the fact that the recall is low which shows that the model likely predicts a \emph{<pad>} token instead.
@@
 The difference lies for both \emph{Dale} datasets at around 8\%.
-However, the precision is still far above a random guess, when used on the \emph{Dale-2} dataset.
-Using the \emph{Dale-5} dataset, the precision only lies at around 50\% to 57\%.
+However, the precision is still far above a random guess, when used on the 'Dale-2' dataset.
+Using the 'Dale-5' dataset, the precision only lies at around 50\% to 57\%.
 
@@
 In conclusion, when the model doesn't need to extract the objects from the scene it performs very well.
-In this case, it is able to achieve almost perfect results on the \emph{Dale-2} dataset and very good results on the remaining two datasets.
+In this case, it is able to achieve almost perfect results on the 'Dale-2' dataset and very good results on the remaining two datasets.
 However, using complete scenes impairs the performance highly.
 The \emph{overall accuracies} decrease and the \emph{non-target accuracies} increase; the model has more difficulties to focus on the correct object.
-The most difficulties the model has with the \emph{CLEVR color} dataset where only the color is a discriminative attribute.
-Looking at the attributes, the last slot (shape) is always the easiest to predict, while the first slot (size for the \emph{Dale} dataset and color for the \emph{CLEVR color} dataset) is the most difficult.
+The most difficulties the model has with the 'CLEVR color' dataset where only the color is a discriminative attribute.
+Looking at the attributes, the last slot (shape) is always the easiest to predict, while the first slot (size for the \emph{Dale} dataset and color for the 'CLEVR color' dataset) is the most difficult.
 This effect is again amplified when the model needs to extract the objects itself instead of being presented with bounding boxes.
@@
 First, as already seen in the previous experiments, a bigger number of distractors confuses the model more where to focus on.
-In the \emph{Dale-2} dataset, there are only two possibilities, while the four distractors in the \emph{Dale-5} dataset and up to nine distractors in the \emph{CLEVR color} give the model a bigger choice.
+In the 'Dale-2' dataset, there are only two possibilities, while the four distractors in the 'Dale-5' dataset and up to nine distractors in the 'CLEVR color' give the model a bigger choice.
 Moreover, having only one attribute to discriminate the target object, poses the model with more difficulties.
@@
 The second explanation lies in the incremental GRE-algorithm.
-For instance using the \emph{Dale-2} dataset, when only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
+For instance using the 'Dale-2' dataset, when only two random objects are placed in a scene, a second (or third) attribute to discriminate the objects is only needed, when the shape is the same.
 Otherwise, the shape is enough, and the RE is only one word long.
@@
 For shape and color being enough, the probability lies at 29,2\% and that all three attributes are necessary is 4,1\%.
-Opposed to that the probabilities with four distractors as in the \emph{Dale-5} dataset are 19,8\%, 47\% and 33,2\% respectively.
+Opposed to that the probabilities with four distractors as in the 'Dale-5' dataset are 19,8\%, 47\% and 33,2\% respectively.
 With four distractors the algorithm is much more likely to produce longer referring expressions.

### 2024-04-05 thesis/sections/6.1-object-identification.tex
rename

@@
 
-The language game is run on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets.
-A random guess corresponds to 50\% in the \emph{Dale-2} dataset, 20\% in the \emph{Dale-5} dataset and 13\% in the \emph{CLEVR color} dataset.
+The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
+A random guess corresponds to 50\% in the 'Dale-2' dataset, 20\% in the 'Dale-5' dataset and 13\% in the 'CLEVR color' dataset.
 
@@
     \centering
-    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-2' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-2_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+    \subfigure['Dale-2' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-2_max-len.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-5' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-5_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+    \subfigure['Dale-5' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_dale-5_max-len.png}
@@
     }
-    \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
+    \subfigure['CLEVR color' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_colour_vocab-size.png}
@@
     }
-    \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
+    \subfigure['CLEVR color' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_discriminator_colour_max-len.png}
@@
 
-The agents can solve the task on \emph{Dale-2} dataset very fast.
+The agents can solve the task on 'Dale-2' dataset very fast.
 Most of the configurations reach accuracies close to the final accuracy already after 2000 to 7000 games.
@@
 The influence of the vocabulary size and message length is not as clear, but smaller vocabulary sizes tend to help the agents to learn faster.
-On the \emph{Dale-5} dataset, there is already a wider spread in performances and the learning is slower than on the \emph{Dale-2} dataset.
+On the 'Dale-5' dataset, there is already a wider spread in performances and the learning is slower than on the 'Dale-2' dataset.
 The agents learn fast up to an accuracy of around 50\%, but for most configurations it then slows down, and the accuracy improves until the cutoff after 128.000 games.
@@
 Interestingly, there is also one configuration that diverges from the baseline very late after almost 80.000 games.
-Finally, on the \emph{CLEVR color} dataset, difference in learning speed are again more apparent.
+Finally, on the 'CLEVR color' dataset, difference in learning speed are again more apparent.
 Agents tend to learn the fastest with a vocabulary size of $|V|=2$, the accuracy however almost stagnates after the initial boost to around 40\%.
 When agents are given a vocabulary of size $|V| \in \{10,16\}$, the initial learning takes place several thousand games later, but therefore reaches higher final accuracies.
-As seen already for the \emph{Dale-5} dataset, most of the configurations converge in the first 20.000 games, but there are several converge much later.
+As seen already for the 'Dale-5' dataset, most of the configurations converge in the first 20.000 games, but there are several converge much later.
 An influence of the message length on the speed is again not visible.
@@
 
-When the agents are tasked to discriminate between the two objects of the \emph{Dale-2} dataset, it succeeds in almost all configurations with an accuracy of over 94\%.
+When the agents are tasked to discriminate between the two objects of the 'Dale-2' dataset, it succeeds in almost all configurations with an accuracy of over 94\%.
 Selected results are listed in Table \ref{tab:results_discriminator_dale-2}.
@@
 
-When the agents are presented with the \emph{Dale-5} dataset, the results are less clear (see Table \ref{tab:results_discriminator_dale-5}).
+When the agents are presented with the 'Dale-5' dataset, the results are less clear (see Table \ref{tab:results_discriminator_dale-5}).
 Again, two groups can be identified: One group of configurations doesn't surpass the baseline of 35\% accuracy, while the other group achieves accuracies in between 62\% and 94\%.
-For this dataset, the group of successful communication has a wider range of results compared to the \emph{Dale-2} dataset.
+For this dataset, the group of successful communication has a wider range of results compared to the 'Dale-2' dataset.
 Only 16 configurations surpass 90\% accuracy, while the majority lies between 80\% and 90\%.
@@
 
-Finally, the results for the \emph{CLEVR color} dataset show a similar structure (see Table \ref{tab:results_discriminator_color}).
+Finally, the results for the 'CLEVR color' dataset show a similar structure (see Table \ref{tab:results_discriminator_color}).
 Two main groups can be identified: Configurations that perform as the baseline with an accuracy of around 23\% and configurations that successfully communicate with accuracies of 40\% to 73\%.
-Compared to the \emph{Dale-5} dataset, the highest accuracy is much lower.
-Furthermore, almost 70\% of the configurations are unsuccessful, which is much bigger compared to the \emph{Dale-5} dataset.
+Compared to the 'Dale-5' dataset, the highest accuracy is much lower.
+Furthermore, almost 70\% of the configurations are unsuccessful, which is much bigger compared to the 'Dale-5' dataset.
 Still, similar conclusions can be drawn.
@@
 The sender is able to encode a description of the target object into the message which the receiver is able to decode.
-Unsurprisingly, this works the best for simplest dataset \emph{Dale-2} with only one distractor.
+Unsurprisingly, this works the best for simplest dataset 'Dale-2' with only one distractor.
 With an increasing number of distractors, fewer games are successful, and successful games perform worse.

### 2024-04-05 thesis/sections/6.2-re-generation.tex
rename

@@
 
-Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} dataset.
-The \emph{Dale-5} and the \emph{CLEVR color} should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
+Since the agents are trained to describe the target object discriminatively based on the described GRE-algorithm, they are trained on the 'Dale-2', 'Dale-5' and 'CLEVR color' dataset.
+The 'Dale-5' and the 'CLEVR color' should be again much harder to learn, since there are more objects that the agents need to discriminate the target object from.
 % SD: So the sender is generating descriptions following the GRE policy?
@@
     \centering
-    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-2' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-2_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+    \subfigure['Dale-2' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-2_max-len.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-5' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-5_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+    \subfigure['Dale-5' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_re-generator_dale-5_max-len.png}
@@
     \caption{Learning curves of all referring expression generation games on the \emph{Dale} datasets.
-On the \emph{CLEVR color} dataset, the learning curves don't diverge from the baseline and are therefore left out.
+On the 'CLEVR color' dataset, the learning curves don't diverge from the baseline and are therefore left out.
 In the left column, the colors correspond to different vocabulary sizes $|V|$ while in the right column, the colors correspond to different message lengths $n$.
@@
 If configurations perform better on the other hand means that the agents learned to encode some meaning in their messages.
-While for both \emph{Dale} dataset, several configuration diverge from the baseline, in other words, learning takes place, all learning curves align with the baseline for the \emph{CLEVR color} dataset and are therefore left out.
-Looking at the learning curves for the \emph{Dale-2} dataset, most of the configurations can decrease the loss compared to the baseline faster on to a bigger extend.
+While for both \emph{Dale} dataset, several configuration diverge from the baseline, in other words, learning takes place, all learning curves align with the baseline for the 'CLEVR color' dataset and are therefore left out.
+Looking at the learning curves for the 'Dale-2' dataset, most of the configurations can decrease the loss compared to the baseline faster on to a bigger extend.
 Fast learning takes place in the first 10.000 to 15.000 games and decelerates afterwards.
@@
 While the vocabulary size now seems to have a lower impact on the learning speed, configurations that converge later tend to have a message length of $n=6$.
-For the \emph{Dale-2} dataset a clear distinction between configurations that align with the baseline and configurations that diverge from the baseline is visible.
-For the \emph{Dale-5} dataset, however, all learning curves are closer together and success of learning is not instantly apparent.
+For the 'Dale-2' dataset a clear distinction between configurations that align with the baseline and configurations that diverge from the baseline is visible.
+For the 'Dale-5' dataset, however, all learning curves are closer together and success of learning is not instantly apparent.
 Nonetheless, several configurations (with low- to medium-sized vocabularies) start to converge earlier than the baseline.
@@
 Table \ref{tab:results:re-generator-game} shows the \emph{overall accuracies}, \emph{F1-scores} and \emph{non-target accuracies} across all datasets for different message lengths $n$ and vocabulary sizes $|V|$.
-When first having a look at the results of the \emph{Dale-2} dataset, it can be seen that the baseline already identifies correct referring expressions in 41,8\% of the samples.
+When first having a look at the results of the 'Dale-2' dataset, it can be seen that the baseline already identifies correct referring expressions in 41,8\% of the samples.
 However, almost all other configurations perform much better.
@@
 
-The results look different when the agents are trained to generate referring expressions for the \emph{Dale-5} dataset.
+The results look different when the agents are trained to generate referring expressions for the 'Dale-5' dataset.
 First, the baseline, in other words the receiver on their own, performs much better.
@@
 
-Finally, the results of the experiments with the \emph{CLEVR color} dataset show that the agents are not able to learn a language to describe the target objects.
+Finally, the results of the experiments with the 'CLEVR color' dataset show that the agents are not able to learn a language to describe the target objects.
 The baseline corresponds to a random guess of the color while the other attributes are identified correctly (since they are identical for target and distractor in the complete scene).
@@
 For the \emph{Dale} dataset, the \emph{shape} is identified well while the agents struggle more with the \emph{color} and the most with the \emph{size}.
-The agents identify the \emph{shape} and \emph{size} perfectly on the \emph{CLEVR color} dataset, but only predict few \emph{colors}.
+The agents identify the \emph{shape} and \emph{size} perfectly on the 'CLEVR color' dataset, but only predict few \emph{colors}.
 
-Looking at the \emph{shape}, the agents perform almost perfectly with the \emph{Dale-2} dataset.
-The precision and recall are slightly lower with the \emph{Dale-5} dataset, and the agents predict the \emph{sphere} too often, while predicting the \emph{cube} too few.
+Looking at the \emph{shape}, the agents perform almost perfectly with the 'Dale-2' dataset.
+The precision and recall are slightly lower with the 'Dale-5' dataset, and the agents predict the \emph{sphere} too often, while predicting the \emph{cube} too few.
 The difference however lies only around 1\%.
 
-Interestingly, the agents can predict the \emph{color} the best, when presented with the \emph{Dale-5} dataset.
-The precision and recall lie around 90\%, while for the \emph{Dale-2} dataset they lie only around 73\%.
+Interestingly, the agents can predict the \emph{color} the best, when presented with the 'Dale-5' dataset.
+The precision and recall lie around 90\%, while for the 'Dale-2' dataset they lie only around 73\%.
 However, there is a bigger difference across the colors.
 Some colors are predicted more often (e.g.
-purple and red for \emph{Dale-5}, and yellow red, cyan and blue for \emph{Dale-2}) than others (e.g.
-brown and green for \emph{Dale-5}, and purple, gray and brown for \emph{Dale-2}).
+purple and red for 'Dale-5', and yellow red, cyan and blue for 'Dale-2') than others (e.g.
+brown and green for 'Dale-5', and purple, gray and brown for 'Dale-2').
 As can be seen, the colors are not consistent and "difficult" or "easy" colors can't be identified.
@@
 The differences between the two \emph{Dale} dataset, might be explainable by the number of scenes in which the color is part of the referring expression.
-As seen in Table \ref{tab:results:re-generator-game}, the \emph{overall accuracy} is higher for the \emph{Dale-2} dataset, even though the prediction of the \emph{color} attribute is worse.
-The reason for this is that many referring expressions in the \emph{Dale-2} dataset only rely on the \emph{shape} and no \emph{color} is needed.
+As seen in Table \ref{tab:results:re-generator-game}, the \emph{overall accuracy} is higher for the 'Dale-2' dataset, even though the prediction of the \emph{color} attribute is worse.
+The reason for this is that many referring expressions in the 'Dale-2' dataset only rely on the \emph{shape} and no \emph{color} is needed.
 The agents therefore can perform well even though predictions of the \emph{color} are wrong.
-On the other side, the \emph{color} is a central part of almost all referring expressions in the \emph{Dale-5} dataset and the agents are pushed to learn this attribute more intensely.
+On the other side, the \emph{color} is a central part of almost all referring expressions in the 'Dale-5' dataset and the agents are pushed to learn this attribute more intensely.
 
-Since the single model in Section \ref{sec:referring_expression_generation} already struggled with predicting the \emph{color} on the \emph{CLEVR color} dataset, it is no surprise that it performs badly in the more complex setup of language games.
+Since the single model in Section \ref{sec:referring_expression_generation} already struggled with predicting the \emph{color} on the 'CLEVR color' dataset, it is no surprise that it performs badly in the more complex setup of language games.
 Hereby, the agents are not able to communicate anything and the receiver only generates two colors (green and purple) for all shown samples.
@@
 The final attribute is the \emph{size}.
-While the agents easily learn that the \emph{size} is never part of the referring expression for the \emph{CLEVR color} dataset, they struggle with the \emph{Dale} datasets.
+While the agents easily learn that the \emph{size} is never part of the referring expression for the 'CLEVR color' dataset, they struggle with the \emph{Dale} datasets.
 As in Section \ref{sec:referring_expression_generation}, the agents again struggle to predict the correct length of the referring expression.
@@
 This is mostly due to the \emph{size} attribute, since a large difference of the precision and recall is visible.
-As for the \emph{color}, the agents perform better on the \emph{Dale-5} dataset.
+As for the \emph{color}, the agents perform better on the 'Dale-5' dataset.
 This reason might be the same.
@@
     \centering
-    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-2' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-2_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+    \subfigure['Dale-2' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-2_max-len.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-5' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-5_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+    \subfigure['Dale-5' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_oh-generator_dale-5_max-len.png}
@@
 Communication is therefore impacting mostly the \emph{shape}, and the influence of the vocabulary size and message length is shown in the figures.
-On the \emph{Dale-2} dataset, configuration that learn the \emph{shape} quickly tend to use smaller vocabularies.
+On the 'Dale-2' dataset, configuration that learn the \emph{shape} quickly tend to use smaller vocabularies.
 However, there are configurations with $|V| \in \{16,10\}$ that need 15.000 to 20.000 games, but outperform the configurations which learned faster.
 A similar result is visible for the message length, where shorter messages help the agents to learn faster while longer messages need more games.
-Looking at the \emph{Dale-5} dataset, more configurations are able to learn the \emph{shape}.
+Looking at the 'Dale-5' dataset, more configurations are able to learn the \emph{shape}.
 Most of the configurations again outperform the baseline already after 5.000 to 10.000 games, but the learning decelerates shortly after.
@@
 Table \ref{tab:results:attribute-predictor-game} shows the results for the games in which the receiver predicts a multi-hot vector of the attributes.
-In the results for the baseline, it can be seen that the receiver can already predict the \emph{shape} and \emph{color}  very well: 88,91\% and 92\% respectively on the \emph{Dale-2} dataset, 80,78\% and 89,38\% respectively on the \emph{Dale-5} dataset, and 100\% and 90,78\% respectively on the \emph{CLEVR color} dataset.
-The \emph{size} however corresponds to a random guess on the \emph{Dale} dataset and only achieves 76,72\% on the \emph{CLEVR color} dataset.
+In the results for the baseline, it can be seen that the receiver can already predict the \emph{shape} and \emph{color}  very well: 88,91\% and 92\% respectively on the 'Dale-2' dataset, 80,78\% and 89,38\% respectively on the 'Dale-5' dataset, and 100\% and 90,78\% respectively on the 'CLEVR color' dataset.
+The \emph{size} however corresponds to a random guess on the \emph{Dale} dataset and only achieves 76,72\% on the 'CLEVR color' dataset.
 This shows that the receiver can already identify attributes of the target object on their own and discriminate it from the distractors.
@@
 As can be seen in the table, the improvement is dependent on the attributes.
-When the agents are trained on the \emph{Dale-2} dataset, almost all configurations improve the prediction of the shape.
+When the agents are trained on the 'Dale-2' dataset, almost all configurations improve the prediction of the shape.
 While 11 out of 30 configurations only improve by a maximum of 2\% points, most can predict the \emph{shape} much better, with an accuracy of up to 98,91\%.
@@
 
-The results are similar when the agents need to describe the \emph{Dale-5} dataset.
-The baseline accuracies of the \emph{shape} and \emph{color} are lower, but can be improved to similar peaks as for the \emph{Dale-2} dataset.
+The results are similar when the agents need to describe the 'Dale-5' dataset.
+The baseline accuracies of the \emph{shape} and \emph{color} are lower, but can be improved to similar peaks as for the 'Dale-2' dataset.
 The highest accuracy for the \emph{shape} lies at 97,18\%, while it lies at 95,23\% for the \emph{color}.
@@
 
-Using the \emph{CLEVR color} dataset, almost no improvements can be seen.
+Using the 'CLEVR color' dataset, almost no improvements can be seen.
 The predictions of the \emph{shape} stay at 100\% accuracy, while the predictions of the \emph{color} get slightly worse for most configurations and if they improve they only improve by maximum 1\% point.
 The predictions of the \emph{size} stay the same.
-The fact that the sender has a smaller influence on the \emph{CLEVR color} dataset than on the \emph{Dale} dataset is surprising since the sender would need to communicate only the \emph{color} instead of additionally the \emph{shape} (and \emph{size}).
+The fact that the sender has a smaller influence on the 'CLEVR color' dataset than on the \emph{Dale} dataset is surprising since the sender would need to communicate only the \emph{color} instead of additionally the \emph{shape} (and \emph{size}).
 A reason why this happens might be the larger number of distractors.
 This number might hinder the receiver to converge and understand the task, by connecting the multi-hot vector to the colors of the target object.
-While for the \emph{Dale} datasets the receiver has a choice between only 2 to 5 objects, they need to decide between up to 10 objects on the \emph{CLEVR color} dataset.
+While for the \emph{Dale} datasets the receiver has a choice between only 2 to 5 objects, they need to decide between up to 10 objects on the 'CLEVR color' dataset.
 That might confuse the model where to focus on and subsequently to connect it to the on-hot vector.

### 2024-04-05 thesis/sections/6.3-reference-resolution.tex
rename

@@
 As in section \ref{sec:reference_resolution}, the agents are trained first on the \emph{CLEVR single} dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the \emph{Dale-2}, \emph{Dale-5} and \emph{CLEVR color} datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
+In a next step, the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
 The same metrics are used to evaluate the results as described in \ref{sec:reference_resolution}.
@@
     \centering
-    \subfigure[\emph{Dale-2} dataset]{
+    \subfigure['Dale-2' dataset]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_coordinate-predictor_dale-2.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset]{
+    \subfigure['Dale-5' dataset]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_coordinate-predictor_dale-5.png}
@@
 As a conclusion, this can mean, that the agents' predictions, both correct and incorrect are more confident.
-This seems to happen on the \emph{Dale-2} dataset.
+This seems to happen on the 'Dale-2' dataset.
 The mean distance is not improved, but the accuracy reaches in some configurations up to 46,27\%.
@@
 
-Looking at the results on the \emph{Dale-5} dataset, no improvements for the mean distance nor the accuracy can be seen when a sender is involved.
+Looking at the results on the 'Dale-5' dataset, no improvements for the mean distance nor the accuracy can be seen when a sender is involved.
 However, while the loss stays again constant, the accuracy drops lower in many configurations, up to 7\% points to an accuracy of 23,26\%.
@@
 The receiver seems to use this information to focus more on the objects.
-Since there are more distractors present in the \emph{Dale-5} dataset, the receiver more likely chooses one of those, as again no discriminating information about the target object seems to be transferred.
+Since there are more distractors present in the 'Dale-5' dataset, the receiver more likely chooses one of those, as again no discriminating information about the target object seems to be transferred.
 
-Lastly, the similar pattern on the mean distance is also visible for the \emph{CLEVR color} dataset.
+Lastly, the similar pattern on the mean distance is also visible for the 'CLEVR color' dataset.
 The accuracy of the baseline is already lower to begin, and a sender can lift this score in almost all configurations.
@@
 These numbers align with the results of the single model, discussed in Section \ref{sec:referring_expression_generation}.
-Hereby, the single model was only able to predict correct coordinates for the \emph{Dale-2} dataset with one distractor.
+Hereby, the single model was only able to predict correct coordinates for the 'Dale-2' dataset with one distractor.
 The higher complexity for the agents that includes more layers in the model as well as the discrete bottleneck raises the bar of the task too high to be solved in this configuration.
@@
 However, when the agents start to learn to communicate, the probability is boosted instantly to a higher level, where it again learns at a slower speed parallel to the baseline.
-On the \emph{Dale-2} dataset, the boost is around 40\% points.
+On the 'Dale-2' dataset, the boost is around 40\% points.
 Most of the learning takes place in the first 40.000 games, but there are also two configurations that increase the performance very late after 70.000 and 105.000 games respectively.
@@
 The same applies as well to the message length.
-Using the \emph{Dale-5} dataset, the probability masses are boosted around 30\% points when the agents start to communicate successfully.
-Compared to the \emph{Dale-2} dataset, fewer configurations start to converge, while most achieve performances close to the baseline.
+Using the 'Dale-5' dataset, the probability masses are boosted around 30\% points when the agents start to communicate successfully.
+Compared to the 'Dale-2' dataset, fewer configurations start to converge, while most achieve performances close to the baseline.
 The smaller number of learning curves makes the analysis more difficult, but the same trend about the vocabulary size and the message length is still visible.
@@
 This indicates that one symbol too few to encode all meaning, but too many symbols are too difficult to learn.
-This hypothesis is amplified by the results on the \emph{CLEVR color} dataset.
+This hypothesis is amplified by the results on the 'CLEVR color' dataset.
 Only two configurations beat the baseline, both with a medium-sized vocabulary size and message length.
@@
     \centering
-    \subfigure[\emph{Dale-2} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-2' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-2_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-2} dataset with different $n$ highlighted]{
+    \subfigure['Dale-2' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-2_max-len.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $|V|$ highlighted]{
+    \subfigure['Dale-5' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-5_vocab-size.png}
@@
     }
-    \subfigure[\emph{Dale-5} dataset with different $n$ highlighted]{
+    \subfigure['Dale-5' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_dale-5_max-len.png}
@@
     }
-    \subfigure[\emph{CLEVR color} dataset with different $|V|$ highlighted]{
+    \subfigure['CLEVR color' dataset with different $|V|$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_colour_vocab-size.png}
@@
     }
-    \subfigure[\emph{CLEVR color} dataset with different $n$ highlighted]{
+    \subfigure['CLEVR color' dataset with different $n$ highlighted]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_attention-predictor_colour_max-len.png}
@@
 The probability mass, that is the sum of the predicted probabilities of all regions of the target object is higher than a uniform distribution (\approx 4,6\%) and a random guess of an object.
-It reaches 62,16\% on the \emph{Dale-2} dataset, 49,61\% on the \emph{Dale-5} dataset and 41,68\% on the \emph{CLEVR color} dataset.
+It reaches 62,16\% on the 'Dale-2' dataset, 49,61\% on the 'Dale-5' dataset and 41,68\% on the 'CLEVR color' dataset.
 Interestingly, including a sender in the setup can improve the result for all dataset.
-Looking at the \emph{Dale-2} dataset, almost all configurations beat the baseline and achieve performances of over 90\%, the best configurations reach even 96\%.
+Looking at the 'Dale-2' dataset, almost all configurations beat the baseline and achieve performances of over 90\%, the best configurations reach even 96\%.
 Only three configurations stay on the level of the baseline.
@@
 
-On the \emph{Dale-5} dataset, the agents already have bigger problems to beat the baseline.
+On the 'Dale-5' dataset, the agents already have bigger problems to beat the baseline.
 Only 8 out of 30 configuration perform better and reach probability masses around 76\% to 84\%.
-However, the increase compared to the baseline is with around 30\% points as high as on the \emph{Dale-2} dataset.
+However, the increase compared to the baseline is with around 30\% points as high as on the 'Dale-2' dataset.
 Smaller message lengths ($n \in \{1,2,3\}$) as well as a medium-sized vocabulary ($|V| \in \{10,16,50\}$) tend to help the agents more, to solve the task successfully.
 As before, no correlation is visible with the few successful games.
-That the agents struggle more with the \emph{Dale-5} dataset is not surprising.
+That the agents struggle more with the 'Dale-5' dataset is not surprising.
 First, the larger number of distractors makes it more difficult for the receiver to focus, as can be seen already in the baseline performances.
@@
 
-The agents struggle the most on the \emph{CLEVR color} dataset.
+The agents struggle the most on the 'CLEVR color' dataset.
 Hereby, only two configurations perform better than the baseline and reach a probability mass of around 64\% to 67\%.
@@
 This indicates that there is communication between the agents, but it rather distracts the receiver from the target object towards the distractors.
-The same argument for a more difficult task when more objects are involved can be made for the \emph{CLEVR color} dataset.
+The same argument for a more difficult task when more objects are involved can be made for the 'CLEVR color' dataset.
 This dataset includes even up to 10 objects, present in the scene which increases the likelihood that the receiver focuses on a wrong object.

### 2024-04-05 thesis/sections/4-datasets.tex
structure and comments
describe experimental setups
attention

@@
-\section{Creation of the dataset}
+\section{Creation of the datasets}
 \label{sec:creation-dataset}
@@
 Two datasets following these rules are created.
-The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
+The 'Dale-2' dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
 Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}.
@@
 More specifically it helps to understand how models learn to discriminate objects from each other, since the model may only need to learn discriminative features of objects and not all features of the whole object.
-
-\cmtDK{probabilities of shared attributes for both Dale-2 and Dale-5}

### 2024-04-05 thesis/sections/5.3-reference-resolution.tex
learning curves, architectures
rename
attention

@@
 More specifically, the image is encoded using the \emph{image encoder} submodule, omitting the last \emph{max pool} layer.
-By this, the resulting matrix has 128\times14\times14 dimension, corresponding to the 14\time14 regions.
+By this, the resulting matrix has 128\times14\times14 dimension, corresponding to the 14\times14 regions.
 For the target object, the incremental RE is passed through an LSTM as described before.
@@
 To calculate the loss, the $softmax$ function is applied over the prediction and compared to the ground truth matrix $A$ using the \emph{binary cross entropy}.
+Figure \ref{fig:attention_predictor_architecture} shows the architecture.
 Hereby, the models converged faster and were trained for 20 epochs with a learning rate of $2\times10^{-4}$.
@@
 Projection sizes $p=\{50,100,300,500,1000\}$ are compared.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=.25\linewidth]{figures/arch_attention_predictor.png}
+    \caption{Simplified architecture of the attention predictor}
+    \label{fig:attention_predictor_architecture}
+\end{figure}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[p]
     \centering
@@
 Here, the predicted coordinates compared to the ground truth coordinates are visualized (Figure \ref{fig:visualizations}), using the best performing model with $e_i=300$, $c=2048$, $LSTM_o=1500$ and $LSTM_e=15$.
-Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of Dale-2.
+Figure \ref{fig:visualizations_dale-2_train} shows random examples of predictions for images in the train dataset of 'Dale-2'.
 The green circle shows the ground truth center coordinates of the target object, while the red circle shows the prediction of the model.
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
     \centering
@@
 % DK: TODO (future work)
+
+A central problem in the models' predictions is therefore the focussing on precise locations in the image and connecting them to the referring expression.
+The \textbf{attention reference resolver} is meant to solve this problem in two ways.
+First, the predictions must not be as precise, as only regions in a 14\times14 grid need to be identified.
+Second, the referring expressions are naturally combined with each region in the image.
+While the \emph{coordinate reference resolver} only combines the complete feature map with the encoded referring expression, the \emph{attention reference resolver} calculates the similarity of each region with the referring expression separately.
+This should connect both encodings easier.
+As Table \ref{tab:results:attention-reference-resolver} shows, this hypothesis seems to be confirmed.
+Across all datasets, the model is able focus on the correct region in the image with high precision.
+On the 'Dale-2' dataset, the model is able to make 96,57\% of its predictions in the correct regions around the target object with a projection size of $p=500$.
+The worst result with $p=300$ still only lies around 3\% points below and therefore identifies the target object almost always correctly.
+These results still hold, when more distractors and more complex referring expressions are introduced.
+On the 'Dale-5' dataset, the model can achieve similar performances, with a probability mass of 93,61\% with $p=300$, while three of the remaining configurations stay above 92\%.
+When the \emph{color} is the only discriminating attribute, the model achieves again similar results to the 'Dale-2' dataset, despite the higher number of distractors of up to 10.
+The best performance of 95,33\% is achieved with $p=100$, but all configurations beat are almost reach 94\%.
+In conclusion, the models are able to easily understand the referring expression and to ground them in the target object in the image.
+While more complex referring expressions challenge the model slightly more, it still can focus on the correct object.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        \toprule
+               & \multicolumn{1}{c}{\textbf{Dale-2}} & \multicolumn{1}{c}{\textbf{Dale-5}} & \multicolumn{1}{c}{\textbf{CLEVR color}} \\  \cmidrule(lr){2-2}\cmidrule(lr){3-3}\cmidrule(lr){4-4}
+        $p$    & \textbf{Probability mass}           & \textbf{Probability mass}           & \textbf{Probability mass}                \\\midrule
+        {50}   & {94,47\%}                           & {89,69\%}                           & {94,36\%}                                \\
+        {100}  & {95,16\%}                           & {92,19\%}                           & {95,33\%}                                \\
+        {300}  & {93,65\%}                           & {93,61\%}                           & {94,15\%}                                \\
+        {500}  & {96,57\%}                           & {92,29\%}                           & {94,15\%}                                \\
+        {1000} & {95,93\%}                           & {92,93\%}                           & {93,82\%}                                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Probability masses of the \emph{attention reference resolver} for different projection sizes $p$ of the model after 20 epochs with $LSTM_e=15$ and $LSTM_o=1500$}
+    \label{tab:results:attention-reference-resolver}
+\end{table}
+
+This conclusion gets even more clear, when having a look at the image, where the model identified the wrong regions.
+Figure \ref{fig:visualizations_attention} shows three images for each dataset, where the model's predicted probability mass was lower than 60\%.
+The errors of the models can be classified into three categories.
+Most of the errors are based on non-visible target objects.
+Either the target object is located outside the preprocessed cropped image, or it is partially hidden behind another object.
+Unsurprisingly, the model has difficulties to focus on these regions.
+However, in several cases, the models can still make use of small hints of the target object.
+In the left image if figure \ref{fig:visualizations_dale-2_attention}, a corner of the target object is visible, which the model uses to 31,9\% of its attention on.
+In the central image of the same figure, only the shadow of the target image is visible.
+Still, the model 44,8\% of its attention onto the correct region, while the remaining probabilities are on adjacent regions.
+This indicates that the model can base its decision on excluding a distractor if it doesn't match the referring expression.
+This behavior is mainly visible on the 'Dale-2' dataset with only one distractor.
+On the 'Dale-5' dataset, the model often focuses one on of the remaining distractors.
+Hereby, the model still focuses on distractors that have attributes mentioned in the referring expressions, as the \emph{sphere} in the left image of Figure \ref{fig:visualizations_dale-5_attention} or the \emph{cyan} and \emph{cylinder} distractors in the central image.
+
+The second category of errors includes errors in which the model confuses a distractor with the target object.
+Most of the time, the model still focuses partially on target object, but the main focus lies somewhere else.
+Again, in these cases, the distractor usually shares an attribute with the target object.
+This is for instance shown in the right image of Figure \ref{fig:visualizations_dale-5_attention}, where the model focuses on the \emph{blue} and \emph{small cube} distractors.
+
+Finally, errors also happen when the target region is not completely placed on the target object.
+In several cases the automatic calculation of the target region around the center point doesn't include the parts of the object.
+The model then doesn't actually make an error and focuses on the correct object, but attends only to adjacent regions.
+These predictions are then evaluated as wrong, even though the model was attending correctly.
+This can be for example seen in the left and central images of Figure \ref{fig:visualizations_colour_attention}.
+
+\begin{figure}[ht]
+    \centering
+    \subfigure['Dale-2']{
+        \includegraphics[width=.839\linewidth]{figures/visualization_dale-2_attention.png}
+        \label{fig:visualizations_dale-2_attention}
+    }
+    \subfigure['Dale-5']{
+        \includegraphics[width=.839\linewidth]{figures/visualization_dale-5_attention.png}
+        \label{fig:visualizations_dale-5_attention}
+    }
+    \subfigure['CLEVR color']{
+        \includegraphics[width=.839\linewidth]{figures/visualization_colour_attention.png}
+        \label{fig:visualizations_colour_attention}
+    }
+    \caption{Examples of the \emph{attention reference resolvers'} predictions with a probability mass lower than 60\% on the 'Dale' and 'CLEVR color' datasets.
+The black rectangle shows the cropped section, the model is actually seeing after the image is preprocessed.
+The green rectangle surrounds the target region that needs to be predicted while the red regions show the actual predictions of the model.
+The more intense the red, the higher is the probability that the model assigned to this region.}
+    \label{fig:visualizations_attention}
+\end{figure}

### 2024-04-05 thesis/sections/6.1-object-identification.tex
attention

@@
     \end{tabular}
-    \caption{Selected results of the object identification games on the Dale-2 dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
+    \caption{Selected results of the object identification games on the 'Dale-2' dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
     \label{tab:results_discriminator_dale-2}
@@
     \end{tabular}
-    \caption{Selected results of the object identification games on the Dale-5 dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
+    \caption{Selected results of the object identification games on the 'Dale-5' dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
     \label{tab:results_discriminator_dale-5}
@@
     \end{tabular}
-    \caption{Selected results of the object identification games on the CLEVR color dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
+    \caption{Selected results of the object identification games on the 'CLEVR color' dataset with $e_s=100$ and $e_r=500$ after 128.000 games}
     \label{tab:results_discriminator_color}

### 2024-04-05 thesis/sections/6.3-reference-resolution.tex
attention

@@
 
-As in section \ref{sec:reference_resolution}, the agents are trained first on the \emph{CLEVR single} dataset to understand if they are capable of predicting locations in an image together.
-In a next step, the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets are used to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
+As in section \ref{sec:reference_resolution}, the agents are trained on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets to test if the agents are able to first communicate a target object and second describe the target object discriminatively with a small vocabulary.
 The same metrics are used to evaluate the results as described in \ref{sec:reference_resolution}.
@@
     }
-    \subfigure[\emph{CELVR color} dataset]{
+    \subfigure['CLEVR color' dataset]{
         \includegraphics[width=0.485\linewidth]{figures/learning-curve_coordinate-predictor_colour.png}

### 2024-04-05 source/mlt/print_results.py

@@ -145,6 +145,9 @@ class Run:
 
 variable_dict = {
     "max_len": "$n$",
+    "encoder_embedding": "$LSTM_e$",
+    "encoder_out": "$LSTM_o$",
+    "projection": "$p$",
     "receiver_embedding": "$e_r$",
     "receiver_hidden": "$h_r$",
     "receiver_projection": "$p$",

### 2024-04-05 thesis/sections/6.3-reference-resolution.tex
visualization

@@
 The same argument for a more difficult task when more objects are involved can be made for the 'CLEVR color' dataset.
-This dataset includes even up to 10 objects, present in the scene which increases the likelihood that the receiver focuses on a wrong object.
+This dataset includes even up to 10 objects present in the scene which increases the likelihood that the receiver focuses on a wrong object.
+
+\begin{figure}[ht]
+    \centering
+    \subfigure['Dale-2']{
+        \includegraphics[width=.839\linewidth]{figures/visualization_games_dale-2_attention.png}
+        \label{fig:visualizations_game_dale-2_attention}
+    }
+    \subfigure['Dale-5']{
+        \includegraphics[width=.839\linewidth]{figures/visualization_games_dale-5_attention.png}
+        \label{fig:visualizations_game_dale-5_attention}
+    }
+    \subfigure['CLEVR color']{
+        \includegraphics[width=.839\linewidth]{figures/visualization_games_colour_attention.png}
+        \label{fig:visualizations_game_colour_attention}
+    }
+    \caption{Examples of the \emph{attention reference resolvers'} predictions in games with successful communication with a probability mass lower than 50\% on the 'Dale' and 'CLEVR color' datasets.
+The black rectangle shows the cropped section, the model is actually seeing after the image is preprocessed.
+The green rectangle surrounds the target region that needs to be predicted while the red regions show the actual predictions of the model.
+The more intense the red, the higher is the probability that the model assigned to this region.}
+    \label{fig:visualizations_game_attention}
+\end{figure}
+
+Figure \ref{fig:visualizations_game_attention} shows examples of the wrongly identified regions on each dataset.
+These are predictions by configurations in which the agents were successfully communicating.
+In other words the sender was not able to transfer information to the receiver, even though a language had already emerged.
+As can be seen on the 'Dale-2' dataset in Figure \ref{fig:visualizations_game_dale-2_attention}, similar problems about non-visible objects occur as for the single models described in Section \ref{sec:reference_resolution}.
+These are in fact the main errors on this dataset.
+However, in several cases (as in the central image), all objects are visible, and the agents still don't attend solely on the target object.
+Rather than choosing one of the distractors, the agents usually attend to both objects relatively equally.
+This indicates that the receiver is uncertain which object the sender is describing.
+Similar errors also occur on both remaining datasets.
+In contrast, the share of errors of the latter type is drastically higher.
+Looking at the results on the 'Dale-5' dataset (see Figure \ref{fig:visualizations_game_dale-5_attention}), the model focusses usually on several objects at once.
+While a general pattern is difficult to identify, the receiver tends to confuse the target object with distractors that share multiple attributes with each other.
+In the central and right image, the wrongly identified distractors share both \emph{size} and \emph{shape} with the target object.
+Finally, on the 'CLEVR color' dataset, no overall reason for the errors can be identified in the qualitative analysis.
 

### 2024-04-05 source/mlt/language_games/map_predictions.py

@@ -14,17 +14,17 @@ dataset_mapping = {
 
 
 def save_csv(run_dir, dataset, split, file_name):
+    interaction_dir = os.path.join(run_dir, "interactions/")
+    if not os.path.exists(interaction_dir):
+        return
+
     with open(
         os.path.join(run_dir, f"{file_name}_outputs.csv"), "w", encoding="utf-8"
     ) as f:
         f.write("image_id,x,y,target_x,target_y\n")
 
     train_interaction = torch.load(
-        glob(
-            os.path.join(
-                run_dir, "interactions/", f"{split}/", "epoch*/", "interaction*"
-            )
-        )[0]
+        glob(os.path.join(interaction_dir, f"{split}/", "epoch*/", "interaction*"))[0]
     )
 
     for image_id, label, receiver_output, length in zip(

### 2024-04-07 thesis/sections/6.1-object-identification.tex
message results

@@
                                       &           &     &       & \multicolumn{1}{c}{\textbf{CLEVR color}} \\  \cmidrule(lr){5-5}
-        $h\_r$                        & $h\_s$    & $n$ & $|V|$ & \textbf{Accuracy}                        \\\midrule
+        $h_r$                         & $h_s$     & $n$ & $|V|$ & \textbf{Accuracy}                        \\\midrule
         {100}                         & {1000}    & {3} & {10}  & {73,65\%}                                \\

### 2024-04-07 thesis/sections/7-language-analysis.tex
message results

@@
 This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{c > sh > s}               & \textbf{c > si > sh}              & \textbf{sh > c > si}     & \textbf{sh > si > c}      & \textbf{si > c > sh}      & \textbf{si > sh > c}      \\\midrule
+        {1} & {2}   & \textcolor{red}{95,17\%}          & \textcolor{red}{97,23\%}          & {87,74\%}                & {82,8\%}                  & \textcolor{red}{94,4\%}   & {88,75\%}                 \\
+        {1} & {10}  & \textcolor{red}{96,04\%}          & \textcolor{red}{94,67\%}          & \textcolor{red}{94,06\%} & {88,67\%}                 & {82,86\%}                 & {79,71\%}                 \\
+        {1} & {16}  & \textcolor{red}{93,82\%}          & \textcolor{red}{95,48\%}          & {85,33\%}                & {79,96\%}                 & \textcolor{red}{90,17\%}  & {84,6\%}                  \\
+        {1} & {50}  & \textcolor{red}{94,47\%}          & \textcolor{red}{96,65\%}          & {86,98\%}                & {81,89\%}                 & \textcolor{red}{90,93\%}  & {86,51\%}                 \\
+        {1} & {100} & \textcolor{red}{95,6\%}           & \textcolor{red}{98,19\%}          & \textcolor{red}{90,41\%} & {86,13\%}                 & \textcolor{red}{95,74\%}  & \textcolor{red}{93,76\%}  \\
+        {2} & {2}   & \textcolor{red}{94,75\%}          & \textcolor{red}{96,65\%}          & {86,51\%}                & {80,84\%}                 & \textcolor{red}{92,99\%}  & {86,3\%}                  \\
+        {2} & {10}  & \textcolor{red}{94,69\%}          & \textcolor{red}{98\%}             & {86,92\%}                & {82,2\%}                  & \textcolor{red}{96,31\%}  & \textcolor{red}{92,55\%}  \\
+        {2} & {16}  & \textcolor{red}{95,04\%}          & \textcolor{red}{96,39\%}          & {87,82\%}                & {82,54\%}                 & \textcolor{red}{90,91\%}  & {87,2\%}                  \\
+        {2} & {50}  & \textcolor{red}{97,18\%}          & \textcolor{red}{93,99\%}          & \textcolor{red}{96,42\%} & \textcolor{red}{92,26\%}  & \textbf{77,97\%}          & {75,58\%}                 \\
+        {2} & {100} & \textcolor{red}{93,71\%}          & \textcolor{red}{94,65\%}          & {85,56\%}                & {78,95\%}                 & {83,77\%}                 & {78,69\%}                 \\
+        {3} & {2}   & \textcolor{red}{95,24\%}          & \textcolor{red}{97,37\%}          & {87,56\%}                & {82,49\%}                 & \textcolor{red}{94,38\%}  & {88,87\%}                 \\
+        {3} & {10}  & \textcolor{red}{96,62\%}          & \textcolor{red}{\textbf{91,61\%}} & \textcolor{red}{94,87\%} & {89\%}                    & \textbf{69,2\%}           & \textbf{65,45\%}          \\
+        {3} & {16}  & \textcolor{red}{93,45\%}          & \textcolor{red}{97,16\%}          & {83,6\%}                 & {77,06\%}                 & \textcolor{red}{94,5\%}   & {89\%}                    \\
+        {3} & {50}  & \textcolor{red}{\textbf{90,97\%}} & \textcolor{red}{95,74\%}          & \textbf{74,49\%}         & \textbf{68,89\%}          & \textcolor{red}{92,37\%}  & {84,98\%}                 \\
+        {3} & {100} & \textcolor{red}{96,74\%}          & \textcolor{red}{\textbf{92,76\%}} & \textcolor{red}{95,51\%} & \textcolor{red}{90,35\%}  & \textbf{71,88\%}          & \textbf{69,69\%}          \\
+        {4} & {2}   & \textcolor{red}{94,94\%}          & \textcolor{red}{96,45\%}          & {89,08\%}                & {83,89\%}                 & \textcolor{red}{92,29\%}  & {85,7\%}                  \\
+        {4} & {10}  & \textcolor{red}{\textbf{91,62\%}} & \textcolor{red}{\textbf{93,29\%}} & {78,53\%}                & {70,85\%}                 & {81,99\%}                 & \textbf{74,83\%}          \\
+        {4} & {16}  & \textcolor{red}{92,86\%}          & \textcolor{red}{96,45\%}          & {79,49\%}                & {73,49\%}                 & \textcolor{red}{94,05\%}  & {86,05\%}                 \\
+        {4} & {50}  & \textcolor{red}{94\%}             & \textcolor{red}{94,78\%}          & {86,17\%}                & {79,61\%}                 & {88,11\%}                 & {82,65\%}                 \\
+        {4} & {100} & \textcolor{red}{92,15\%}          & \textcolor{red}{96,99\%}          & {78,91\%}                & {72,51\%}                 & \textcolor{red}{94,8\%}   & {88,6\%}                  \\
+        {6} & {2}   & \textcolor{red}{99,41\%}          & \textcolor{red}{100,29\%}         & \textcolor{red}{97,91\%} & \textcolor{red}{97\%}     & \textcolor{red}{101,11\%} & \textcolor{red}{98,09\%}  \\
+        {6} & {10}  & \textcolor{red}{100,14\%}         & \textcolor{red}{99,98\%}          & \textcolor{red}{99,94\%} & \textcolor{red}{100,07\%} & \textcolor{red}{99,92\%}  & \textcolor{red}{100,06\%} \\
+        {6} & {16}  & \textcolor{red}{\textbf{90,04\%}} & \textcolor{red}{95,83\%}          & \textbf{69,71\%}         & \textbf{61,75\%}          & \textcolor{red}{91,77\%}  & {81,6\%}                  \\
+        {6} & {50}  & \textcolor{red}{91,9\%}           & \textcolor{red}{96,28\%}          & \textbf{76,04\%}         & \textbf{69,86\%}          & \textcolor{red}{92,81\%}  & {85,2\%}                  \\
+        {6} & {100} & \textcolor{red}{99,95\%}          & \textcolor{red}{100,09\%}         & \textcolor{red}{100\%}   & \textcolor{red}{100,1\%}  & \textcolor{red}{99,9\%}   & \textcolor{red}{100,04\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Dale-2}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{('color', 'shape', 'size')} & \textbf{('color', 'size', 'shape')} & \textbf{('shape', 'color', 'size')} & \textbf{('shape', 'size', 'color')} & \textbf{('size', 'color', 'shape')} & \textbf{('size', 'shape', 'color')} \\\midrule
+        {1} & {1}   & {99,59\%}                           & {100,07\%}                          & {99,22\%}                           & {99,34\%}                           & {99,87\%}                           & {99,38\%}                           \\
+        {1} & {2}   & {95,35\%}                           & {89,87\%}                           & {95,09\%}                           & {87,23\%}                           & {86,01\%}                           & {84,14\%}                           \\
+        {1} & {10}  & {99,96\%}                           & {100\%}                             & {99,96\%}                           & {99,99\%}                           & {99,98\%}                           & {100,01\%}                          \\
+        {1} & {16}  & {92,95\%}                           & {87,03\%}                           & {92,63\%}                           & {83,91\%}                           & {82,77\%}                           & {80,06\%}                           \\
+        {1} & {50}  & {93,35\%}                           & {87,23\%}                           & {93,26\%}                           & {84,65\%}                           & {82,46\%}                           & {80,38\%}                           \\
+        {1} & {100} & {99,94\%}                           & {100,1\%}                           & {99,93\%}                           & {99,94\%}                           & {99,96\%}                           & {100,09\%}                          \\
+        {2} & {2}   & {100,38\%}                          & {99,41\%}                           & {100,25\%}                          & {98,61\%}                           & {99,01\%}                           & {98,53\%}                           \\
+        {2} & {10}  & {88,94\%}                           & {84,78\%}                           & {87,5\%}                            & {77,76\%}                           & {80,24\%}                           & {74,95\%}                           \\
+        {2} & {16}  & {85,71\%}                           & {84,53\%}                           & {82,98\%}                           & {72,85\%}                           & {79,72\%}                           & {71,85\%}                           \\
+        {2} & {50}  & {99,95\%}                           & {99,98\%}                           & {99,99\%}                           & {100,02\%}                          & {99,99\%}                           & {99,93\%}                           \\
+        {2} & {100} & {100,05\%}                          & {99,98\%}                           & {99,96\%}                           & {99,93\%}                           & {99,99\%}                           & {99,97\%}                           \\
+        {3} & {2}   & {100,38\%}                          & {99,51\%}                           & {100,31\%}                          & {98,57\%}                           & {98,96\%}                           & {98,5\%}                            \\
+        {3} & {10}  & {99,93\%}                           & {100,01\%}                          & {99,99\%}                           & {100,15\%}                          & {100,01\%}                          & {100,01\%}                          \\
+        {3} & {16}  & {92,18\%}                           & {86,23\%}                           & {91,93\%}                           & {82,59\%}                           & {81,69\%}                           & {78,72\%}                           \\
+        {3} & {50}  & {92,68\%}                           & {86,53\%}                           & {91,9\%}                            & {83,35\%}                           & {82,45\%}                           & {79,99\%}                           \\
+        {3} & {100} & {100,03\%}                          & {99,98\%}                           & {100,03\%}                          & {99,93\%}                           & {100,02\%}                          & {99,87\%}                           \\
+        {4} & {2}   & {100,37\%}                          & {99,38\%}                           & {100,31\%}                          & {98,66\%}                           & {98,9\%}                            & {98,53\%}                           \\
+        {4} & {10}  & {99,99\%}                           & {99,97\%}                           & {99,96\%}                           & {99,91\%}                           & {99,95\%}                           & {99,99\%}                           \\
+        {4} & {16}  & {100\%}                             & {100,02\%}                          & {99,97\%}                           & {99,94\%}                           & {100\%}                             & {99,94\%}                           \\
+        {4} & {50}  & {99,99\%}                           & {99,95\%}                           & {99,95\%}                           & {100,01\%}                          & {99,99\%}                           & {99,98\%}                           \\
+        {4} & {100} & {99,95\%}                           & {100,01\%}                          & {100\%}                             & {100,02\%}                          & {100,05\%}                          & {99,98\%}                           \\
+        {6} & {2}   & {100,41\%}                          & {99,42\%}                           & {100,29\%}                          & {98,6\%}                            & {98,89\%}                           & {98,55\%}                           \\
+        {6} & {10}  & {93,4\%}                            & {86,51\%}                           & {91,64\%}                           & {84,03\%}                           & {82,51\%}                           & {80,06\%}                           \\
+        {6} & {16}  & {100,05\%}                          & {100,01\%}                          & {99,94\%}                           & {99,91\%}                           & {100,03\%}                          & {100\%}                             \\
+        {6} & {50}  & {99,96\%}                           & {99,96\%}                           & {99,92\%}                           & {99,99\%}                           & {100,02\%}                          & {100,03\%}                          \\
+        {6} & {100} & {99,98\%}                           & {99,99\%}                           & {99,94\%}                           & {100,03\%}                          & {100,01\%}                          & {99,93\%}                           \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Dale-5}
+    \label{TODO: label}
+\end{table}

### 2024-04-07 source/mlt/language_games/message_analyzer.py

@@ -502,7 +502,7 @@ if __name__ == "__main__":
                 with open(
                     os.path.join(
                         out_dir,
-                        f"translate_log_{args.combination if args.combination else combination_index}.txt",
+                        f"translate_log_{args.translator}_{args.combination if args.combination != -1 else combination_index}.txt",
                     ),
                     "w",
                     encoding="utf-8",

### 2024-04-07 source/mlt/language_games/print_message_analysis.py

@@ -1,0 +1,394 @@
+import argparse
+import ast
+import glob
+import json
+import os
+import re
+
+# necessary for input()
+# pylint: disable-next=unused-import
+import readline
+from abc import abstractmethod
+from dataclasses import dataclass
+
+from tabulate import tabulate
+
+
+class Run:
+    def __init__(self, file, dataset, baseline_runs=None, english_runs=None) -> None:
+        self.epochs = []
+        with open(file, "r", encoding="utf-8") as f:
+            for line in f:
+                if line.startswith("('"):
+                    self.combination = ast.literal_eval(line)
+
+                if line.startswith("epoch"):
+                    metrics = {"loss": float(line.split()[-1])}
+
+                    if baseline_runs is not None and english_runs is not None:
+                        baseline = [
+                            baseline_run
+                            for baseline_run in baseline_runs[dataset]
+                            if baseline_run.combination == self.combination
+                        ][0].last_epoch
+                        english = [
+                            english_run
+                            for english_run in english_runs[dataset]
+                            if english_run.combination == self.combination
+                        ][0].last_epoch
+
+                        metrics["normalized_loss"] = (
+                            metrics["loss"] - english["loss"]
+                        ) / (baseline["loss"] - english["loss"])
+
+                    self.epochs.append(metrics)
+
+    @property
+    def last_epoch(self):
+        return self.epochs[-1]
+
+    @property
+    def metrics(self):
+        return set(self.last_epoch.keys())
+
+
+variable_dict = {
+    "max_len": "$n$",
+    "encoder_embedding": "$LSTM_e$",
+    "encoder_out": "$LSTM_o$",
+    "projection": "$p$",
+    "receiver_embedding": "$e_r$",
+    "receiver_hidden": "$h_r$",
+    "receiver_projection": "$p$",
+    "receiver_image_embedding": "$e_{ri}$",
+    "receiver_decoder_embedding": "$LSTM_e$",
+    "receiver_decoder_out": "$LSTM_o$",
+    "receiver_coordinate_classifier_dimension": "$c$",
+    "sender_embedding": "$e_s$",
+    "sender_hidden": "$h_s$",
+    "sender_image_embedding": "$e_{si}$",
+    "vocab_size": "$|V|$",
+}
+
+
+@dataclass
+class Experiment:
+    folder_name: str
+    dataset: str
+    variables: dict
+    run: Run
+
+
+class Experiments:
[…]

### 2024-04-11 thesis/sections/7-language-analysis.tex
results

--- 2023-11-30 thesis/sections/7-language-analysis.tex
@@
 This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{95,5\%}           & \textcolor{red}{97,57\%}          & {86,98\%}                & {81,43\%}                & \textcolor{red}{91,67\%} & {86,45\%}                \\
+        {1} & {10}  & \textcolor{red}{96,11\%}          & \textcolor{red}{96,9\%}           & {89,61\%}                & {85\%}                   & {89,4\%}                 & {84,49\%}                \\
+        {1} & {16}  & \textcolor{red}{95,19\%}          & \textcolor{red}{96,56\%}          & {87,25\%}                & {81,22\%}                & {89,34\%}                & {83,13\%}                \\
+        {1} & {50}  & \textcolor{red}{96,61\%}          & \textcolor{red}{97,07\%}          & \textcolor{red}{91,26\%} & {85,75\%}                & {87,14\%}                & {82,61\%}                \\
+        {1} & {100} & \textcolor{red}{95,91\%}          & \textcolor{red}{96,73\%}          & \textcolor{red}{90,63\%} & {86,01\%}                & {88,95\%}                & {84,17\%}                \\
+        {2} & {2}   & \textcolor{red}{99,67\%}          & \textcolor{red}{99,91\%}          & \textcolor{red}{97,02\%} & \textcolor{red}{95,37\%} & \textcolor{red}{96,97\%} & \textcolor{red}{94,64\%} \\
+        {2} & {10}  & \textcolor{red}{96,36\%}          & \textcolor{red}{96,34\%}          & \textcolor{red}{91,32\%} & {86,75\%}                & {87,69\%}                & {82,87\%}                \\
+        {2} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{97,12\%}          & {88,11\%}                & {83,21\%}                & {89,86\%}                & {85,35\%}                \\
+        {2} & {50}  & \textcolor{red}{95,34\%}          & \textcolor{red}{97,3\%}           & {85,48\%}                & {80,89\%}                & \textcolor{red}{92,22\%} & {86,18\%}                \\
+        {2} & {100} & \textcolor{red}{94,88\%}          & \textcolor{red}{96,32\%}          & {87,37\%}                & {81,17\%}                & {88,35\%}                & {82,61\%}                \\
+        {3} & {2}   & \textcolor{red}{\textbf{94,03\%}} & \textcolor{red}{96,96\%}          & \textbf{79,59\%}         & \textbf{74,54\%}         & \textcolor{red}{92,32\%} & {85,63\%}                \\
+        {3} & {10}  & \textcolor{red}{96,1\%}           & \textcolor{red}{95,97\%}          & \textcolor{red}{90,45\%} & {84,84\%}                & {85,98\%}                & {80,5\%}                 \\
+        {3} & {16}  & \textcolor{red}{95,71\%}          & \textcolor{red}{\textbf{95,1\%}}  & \textcolor{red}{90,97\%} & {85,33\%}                & {81\%}                   & \textbf{75,56\%}         \\
+        {3} & {50}  & \textcolor{red}{96,32\%}          & \textcolor{red}{\textbf{94,55\%}} & \textcolor{red}{92,5\%}  & {86,81\%}                & \textbf{80,48\%}         & {75,73\%}                \\
+        {3} & {100} & \textcolor{red}{95,33\%}          & \textcolor{red}{97,82\%}          & {86,05\%}                & {81,33\%}                & \textcolor{red}{93,91\%} & {88,54\%}                \\
+        {4} & {2}   & \textcolor{red}{96,81\%}          & \textcolor{red}{95,12\%}          & \textcolor{red}{93,03\%} & {87,41\%}                & \textbf{80,73\%}         & \textbf{75,47\%}         \\
+        {4} & {10}  & \textcolor{red}{95,31\%}          & \textcolor{red}{97,26\%}          & {86,9\%}                 & {82,67\%}                & \textcolor{red}{92,39\%} & {86,98\%}                \\
+        {4} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{\textbf{94,65\%}} & \textcolor{red}{91,99\%} & {86,65\%}                & \textbf{79,23\%}         & \textbf{74,54\%}         \\
+        {4} & {50}  & \textcolor{red}{94,76\%}          & \textcolor{red}{96,69\%}          & {86,8\%}                 & {81,33\%}                & {89,65\%}                & {83,92\%}                \\
+        {4} & {100} & \textcolor{red}{96,16\%}          & \textcolor{red}{97,94\%}          & {89,57\%}                & {85,18\%}                & \textcolor{red}{92,56\%} & {87,74\%}                \\
+        {6} & {2}   & \textcolor{red}{95,73\%}          & \textcolor{red}{96,99\%}          & {87,77\%}                & {83,27\%}                & \textcolor{red}{91,77\%} & {86,34\%}                \\
+        {6} & {10}  & \textcolor{red}{\textbf{94,59\%}} & \textcolor{red}{99,89\%}          & {86,24\%}                & {81,46\%}                & \textcolor{red}{91,42\%} & {86,43\%}                \\
+        {6} & {16}  & \textcolor{red}{95,51\%}          & \textcolor{red}{96,3\%}           & {88,89\%}                & {83,24\%}                & {87,61\%}                & {82,71\%}                \\
+        {6} & {50}  & \textcolor{red}{\textbf{94,66\%}} & \textcolor{red}{97,43\%}          & \textbf{84,93\%}         & \textbf{78,61\%}         & \textcolor{red}{91,48\%} & {83,66\%}                \\
+        {6} & {100} & \textcolor{red}{95,37\%}          & \textcolor{red}{97,77\%}          & \textbf{83,75\%}         & \textbf{80,84\%}         & \textcolor{red}{93,96\%} & {89,02\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Discriminator: Dale-2}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
+        {1} & {2}   & \textcolor{red}{94,96\%} & \textcolor{red}{91,27\%} & \textcolor{red}{94,51\%} & \textcolor{red}{90,99\%} & {88,52\%}            & {88,17\%}            \\
+        {1} & {10}  & {84,34\%}                & {84,03\%}                & {80,91\%}                & {73,23\%}                & {79,95\%}            & {72,34\%}            \\
+        {1} & {16}  & {86,27\%}                & {84,96\%}                & {82,76\%}                & {75,35\%}                & {80,72\%}            & {74,04\%}            \\
+        {1} & {50}  & {85,95\%}                & {84,4\%}                 & {82,96\%}                & {74,97\%}                & {80,2\%}             & {73,54\%}            \\
+        {1} & {100} & {86,31\%}                & {85,58\%}                & {82,9\%}                 & {76,03\%}                & {81,64\%}            & {74,62\%}            \\
+        {2} & {10}  & \textbf{82,66\%}         & {83,23\%}                & \textbf{78,51\%}         & {71,01\%}                & {78,89\%}            & {70,61\%}            \\
+        {2} & {16}  & {89,78\%}                & {85,53\%}                & {87,68\%}                & {80,64\%}                & {81,06\%}            & {77,24\%}            \\
+        {2} & {50}  & {84,4\%}                 & {82,65\%}                & {80,62\%}                & {71,42\%}                & {77,53\%}            & {69,46\%}            \\
+        {2} & {100} & {89,65\%}                & {85,3\%}                 & {87,26\%}                & {79,91\%}                & {80,88\%}            & {77,1\%}             \\
+        {3} & {2}   & {87,19\%}                & {86,93\%}                & {83,99\%}                & {78,18\%}                & {83,65\%}            & {77,08\%}            \\
+        {3} & {10}  & {87,15\%}                & {84,67\%}                & {84,01\%}                & {76,24\%}                & {80,41\%}            & {74,11\%}            \\
+        {3} & {50}  & \textbf{82,98\%}         & {82,25\%}                & {78,85\%}                & {70,62\%}                & {77,42\%}            & {69,27\%}            \\
+        {3} & {100} & {83,78\%}                & {83,07\%}                & {80,23\%}                & {72,15\%}                & {78,94\%}            & {71,1\%}             \\
+        {4} & {2}   & {89,23\%}                & {87,43\%}                & {87,09\%}                & {81,45\%}                & {83,83\%}            & {79,11\%}            \\
+        {4} & {10}  & {88,66\%}                & {84,83\%}                & {85,9\%}                 & {78,98\%}                & {80,2\%}             & {75,64\%}            \\
+        {4} & {16}  & {83,14\%}                & \textbf{81,51\%}         & \textbf{78,15\%}         & \textbf{70,21\%}         & \textbf{77,32\%}     & \textbf{68,48\%}     \\
+        {4} & {50}  & {87,62\%}                & {84,38\%}                & {85,25\%}                & {77,29\%}                & {79,94\%}            & {74,03\%}            \\
+        {4} & {100} & \textcolor{red}{90,17\%} & {85,76\%}                & {87,68\%}                & {80,71\%}                & {81,37\%}            & {78,1\%}             \\
+        {6} & {10}  & \textbf{82,03\%}         & \textbf{80,51\%}         & \textbf{77,46\%}         & \textbf{68,63\%}         & \textbf{75,71\%}     & \textbf{67,15\%}     \\
+        {6} & {16}  & {84,73\%}                & \textbf{81,12\%}         & {80,99\%}                & \textbf{70,5\%}          & \textbf{76,06\%}     & \textbf{69,12\%}     \\
+        {6} & {50}  & {86,95\%}                & {85,13\%}                & {84,06\%}                & {76,95\%}                & {80,11\%}            & {74,51\%}            \\
+        {6} & {100} & {86,15\%}                & {82,19\%}                & {83,56\%}                & {75,56\%}                & {77,76\%}            & {72,47\%}            \\
+        \bottomrule
+    \end{tabular}
+    \caption{Discriminator: Dale-5}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{91,62\%} & \textcolor{red}{91,57\%} & \textcolor{red}{94,31\%} & \textcolor{red}{95,13\%} & \textcolor{red}{93,11\%} & \textcolor{red}{94,92\%} \\
+        {1} & {10}  & {79,08\%}                & {79,08\%}                & {85,04\%}                & {86,86\%}                & {82,37\%}                & {86,38\%}                \\
+        {1} & {16}  & {84,02\%}                & {83,96\%}                & {88,67\%}                & {89,65\%}                & {86,36\%}                & {89,36\%}                \\
+        {1} & {50}  & \textcolor{red}{93,88\%} & \textcolor{red}{93,74\%} & \textcolor{red}{95,03\%} & \textcolor{red}{95,49\%} & \textcolor{red}{94,64\%} & \textcolor{red}{95,35\%} \\
+        {2} & {10}  & {88,67\%}                & {88,36\%}                & {89,59\%}                & \textcolor{red}{90,77\%} & \textcolor{red}{90,01\%} & {89,81\%}                \\
+        {2} & {16}  & {81,13\%}                & {81,02\%}                & {86,16\%}                & {87,97\%}                & {83,85\%}                & {87,56\%}                \\
+        {3} & {2}   & {83,38\%}                & {83,46\%}                & {88,21\%}                & {89,71\%}                & {86,05\%}                & {89,37\%}                \\
+        {3} & {10}  & \textbf{77\%}            & \textbf{76,77\%}         & \textbf{83,54\%}         & \textbf{85,57\%}         & \textbf{79,97\%}         & \textbf{84,9\%}          \\
+        {3} & {16}  & \textcolor{red}{93,52\%} & \textcolor{red}{93,6\%}  & \textcolor{red}{94,16\%} & \textcolor{red}{94,14\%} & \textcolor{red}{93,75\%} & \textcolor{red}{93,55\%} \\
+        {4} & {10}  & \textbf{68,04\%}         & \textbf{66,51\%}         & \textbf{76,33\%}         & \textbf{78,56\%}         & \textbf{71,42\%}         & \textbf{78,56\%}         \\
+        {4} & {16}  & \textbf{75,1\%}          & \textbf{74,93\%}         & \textbf{81,2\%}          & \textbf{83,35\%}         & \textbf{77,6\%}          & \textbf{82,76\%}         \\
+        {6} & {2}   & {84,21\%}                & {84,02\%}                & {88,41\%}                & {89,77\%}                & {86,61\%}                & {89,35\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Discriminator: Color}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{95,17\%}          & \textcolor{red}{97,23\%}          & {87,74\%}                & {82,8\%}                 & \textcolor{red}{94,4\%}  & {88,75\%}                \\
+        {1} & {10}  & \textcolor{red}{96,04\%}          & \textcolor{red}{94,67\%}          & \textcolor{red}{94,06\%} & {88,67\%}                & {82,86\%}                & {79,71\%}                \\
+        {1} & {16}  & \textcolor{red}{93,82\%}          & \textcolor{red}{95,48\%}          & {85,33\%}                & {79,96\%}                & \textcolor{red}{90,17\%} & {84,6\%}                 \\
+        {1} & {50}  & \textcolor{red}{94,47\%}          & \textcolor{red}{96,65\%}          & {86,98\%}                & {81,89\%}                & \textcolor{red}{90,93\%} & {86,51\%}                \\
+        {1} & {100} & \textcolor{red}{95,6\%}           & \textcolor{red}{98,19\%}          & \textcolor{red}{90,41\%} & {86,13\%}                & \textcolor{red}{95,74\%} & \textcolor{red}{93,76\%} \\
+        {2} & {2}   & \textcolor{red}{94,75\%}          & \textcolor{red}{96,65\%}          & {86,51\%}                & {80,84\%}                & \textcolor{red}{92,99\%} & {86,3\%}                 \\
+        {2} & {10}  & \textcolor{red}{94,69\%}          & \textcolor{red}{98\%}             & {86,92\%}                & {82,2\%}                 & \textcolor{red}{96,31\%} & \textcolor{red}{92,55\%} \\
+        {2} & {16}  & \textcolor{red}{95,04\%}          & \textcolor{red}{96,39\%}          & {87,82\%}                & {82,54\%}                & \textcolor{red}{90,91\%} & {87,2\%}                 \\
+        {2} & {50}  & \textcolor{red}{97,18\%}          & \textcolor{red}{93,99\%}          & \textcolor{red}{96,42\%} & \textcolor{red}{92,26\%} & \textbf{77,97\%}         & {75,58\%}                \\
+        {2} & {100} & \textcolor{red}{93,71\%}          & \textcolor{red}{94,65\%}          & {85,56\%}                & {78,95\%}                & {83,77\%}                & {78,69\%}                \\
+        {3} & {2}   & \textcolor{red}{95,24\%}          & \textcolor{red}{97,37\%}          & {87,56\%}                & {82,49\%}                & \textcolor{red}{94,38\%} & {88,87\%}                \\
+        {3} & {10}  & \textcolor{red}{96,62\%}          & \textcolor{red}{\textbf{91,61\%}} & \textcolor{red}{94,87\%} & {89\%}                   & \textbf{69,2\%}          & \textbf{65,45\%}         \\
+        {3} & {16}  & \textcolor{red}{93,45\%}          & \textcolor{red}{97,16\%}          & {83,6\%}                 & {77,06\%}                & \textcolor{red}{94,5\%}  & {89\%}                   \\
+        {3} & {50}  & \textcolor{red}{\textbf{90,97\%}} & \textcolor{red}{95,74\%}          & \textbf{74,49\%}         & \textbf{68,89\%}         & \textcolor{red}{92,37\%} & {84,98\%}                \\
+        {3} & {100} & \textcolor{red}{96,74\%}          & \textcolor{red}{\textbf{92,76\%}} & \textcolor{red}{95,51\%} & \textcolor{red}{90,35\%} & \textbf{71,88\%}         & \textbf{69,69\%}         \\
+        {4} & {2}   & \textcolor{red}{94,94\%}          & \textcolor{red}{96,45\%}          & {89,08\%}                & {83,89\%}                & \textcolor{red}{92,29\%} & {85,7\%}                 \\
+        {4} & {10}  & \textcolor{red}{\textbf{91,62\%}} & \textcolor{red}{\textbf{93,29\%}} & {78,53\%}                & {70,85\%}                & {81,99\%}                & \textbf{74,83\%}         \\
+        {4} & {16}  & \textcolor{red}{92,86\%}          & \textcolor{red}{96,45\%}          & {79,49\%}                & {73,49\%}                & \textcolor{red}{94,05\%} & {86,05\%}                \\
+        {4} & {50}  & \textcolor{red}{94\%}             & \textcolor{red}{94,78\%}          & {86,17\%}                & {79,61\%}                & {88,11\%}                & {82,65\%}                \\
+        {4} & {100} & \textcolor{red}{92,15\%}          & \textcolor{red}{96,99\%}          & {78,91\%}                & {72,51\%}                & \textcolor{red}{94,8\%}  & {88,6\%}                 \\
+        {6} & {16}  & \textcolor{red}{\textbf{90,04\%}} & \textcolor{red}{95,83\%}          & \textbf{69,71\%}         & \textbf{61,75\%}         & \textcolor{red}{91,77\%} & {81,6\%}                 \\
+        {6} & {50}  & \textcolor{red}{91,9\%}           & \textcolor{red}{96,28\%}          & \textbf{76,04\%}         & \textbf{69,86\%}         & \textcolor{red}{92,81\%} & {85,2\%}                 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Dale-2}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh} & \textbf{Sh > C > Si}              & \textbf{Sh > Si > C} & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
+        {1} & {2}   & \textcolor{red}{95,35\%}          & {89,87\%}            & \textcolor{red}{95,09\%}          & {87,23\%}            & {86,01\%}            & {84,14\%}            \\
+        {1} & {16}  & \textcolor{red}{92,95\%}          & {87,03\%}            & \textcolor{red}{92,63\%}          & {83,91\%}            & {82,77\%}            & {80,06\%}            \\
+        {1} & {50}  & \textcolor{red}{93,35\%}          & {87,23\%}            & \textcolor{red}{93,26\%}          & {84,65\%}            & {82,46\%}            & {80,38\%}            \\
+        {2} & {10}  & \textbf{88,94\%}                  & \textbf{84,78\%}     & \textbf{87,5\%}                   & \textbf{77,76\%}     & \textbf{80,24\%}     & \textbf{74,95\%}     \\
+        {2} & {16}  & \textbf{85,71\%}                  & \textbf{84,53\%}     & \textbf{82,98\%}                  & \textbf{72,85\%}     & \textbf{79,72\%}     & \textbf{71,85\%}     \\
+        {3} & {16}  & \textcolor{red}{\textbf{92,18\%}} & \textbf{86,23\%}     & \textcolor{red}{91,93\%}          & \textbf{82,59\%}     & \textbf{81,69\%}     & \textbf{78,72\%}     \\
+        {3} & {50}  & \textcolor{red}{92,68\%}          & {86,53\%}            & \textcolor{red}{91,9\%}           & {83,35\%}            & {82,45\%}            & {79,99\%}            \\
+        {6} & {10}  & \textcolor{red}{93,4\%}           & {86,51\%}            & \textcolor{red}{\textbf{91,64\%}} & {84,03\%}            & {82,51\%}            & {80,06\%}            \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Dale-5}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {3} & {16}  & \textcolor{red}{98,48\%} & \textcolor{red}{98,58\%} & \textcolor{red}{97,87\%} & \textcolor{red}{96,11\%} & \textcolor{red}{96,6\%}  & \textcolor{red}{95,97\%} \\
+        {4} & {10}  & \textcolor{red}{98,88\%} & \textcolor{red}{98,75\%} & \textcolor{red}{97,23\%} & \textcolor{red}{96,67\%} & \textcolor{red}{97,18\%} & \textcolor{red}{96,75\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Color}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}      & \textbf{Si > Sh > C}      \\\midrule
+        {1} & {2}   & \textcolor{red}{95,73\%} & \textcolor{red}{94,23\%} & \textcolor{red}{92,96\%} & \textcolor{red}{90,59\%} & {81,59\%}                 & {78,21\%}                 \\
+        {1} & {10}  & \textcolor{red}{91,77\%} & \textcolor{red}{91,19\%} & {73,14\%}                & {63,99\%}                & {67,45\%}                 & {57,55\%}                 \\
+        {1} & {16}  & \textcolor{red}{91,74\%} & \textcolor{red}{90,87\%} & {73,07\%}                & {63,78\%}                & {66,16\%}                 & {55,48\%}                 \\
+        {1} & {50}  & \textcolor{red}{97,78\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,22\%} & {87,45\%}                & {69,23\%}                 & {64,97\%}                 \\
+        {1} & {100} & \textcolor{red}{97,91\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,18\%} & {87,35\%}                & {69,25\%}                 & {64,99\%}                 \\
+        {2} & {2}   & \textcolor{red}{93,17\%} & \textcolor{red}{91,69\%} & {81,96\%}                & {75,97\%}                & {72,05\%}                 & {64,29\%}                 \\
+        {2} & {10}  & \textcolor{red}{94,07\%} & \textcolor{red}{98,04\%} & {75,32\%}                & {71,91\%}                & \textcolor{red}{94,38\%}  & {86,62\%}                 \\
+        {2} & {16}  & \textcolor{red}{97,72\%} & \textcolor{red}{92,9\%}  & \textcolor{red}{94,18\%} & {87,09\%}                & {68,54\%}                 & {64,13\%}                 \\
+        {2} & {50}  & \textcolor{red}{97,87\%} & \textcolor{red}{93,08\%} & \textcolor{red}{94,25\%} & {87,54\%}                & {69,2\%}                  & {65,17\%}                 \\
+        {2} & {100} & {89,6\%}                 & \textbf{89,58\%}         & \textbf{65,65\%}         & \textbf{55,69\%}         & {64,92\%}                 & {53,53\%}                 \\
+        {3} & {2}   & \textcolor{red}{97,73\%} & \textcolor{red}{99,28\%} & \textcolor{red}{95,84\%} & \textcolor{red}{97,72\%} & \textcolor{red}{101,58\%} & \textcolor{red}{100,22\%} \\
+        {3} & {10}  & \textcolor{red}{91,04\%} & \textcolor{red}{90,52\%} & {71,06\%}                & {61,36\%}                & {65,59\%}                 & {54,93\%}                 \\
+        {3} & {16}  & \textcolor{red}{91,09\%} & \textcolor{red}{90,3\%}  & {71,71\%}                & {62,42\%}                & {66,28\%}                 & {56,16\%}                 \\
+        {3} & {50}  & \textcolor{red}{97,07\%} & \textcolor{red}{91,72\%} & \textcolor{red}{92,7\%}  & {85,33\%}                & {66,38\%}                 & {61,26\%}                 \\
+        {3} & {100} & \textcolor{red}{97,81\%} & \textcolor{red}{93,01\%} & \textcolor{red}{94,16\%} & {87,3\%}                 & {68,83\%}                 & {64,6\%}                  \\
+        {4} & {2}   & \textcolor{red}{91,95\%} & \textcolor{red}{96,87\%} & {73,55\%}                & {72,66\%}                & \textcolor{red}{98,37\%}  & \textcolor{red}{90,23\%}  \\
+        {4} & {10}  & \textcolor{red}{97,7\%}  & \textcolor{red}{93,1\%}  & \textcolor{red}{94,16\%} & {87,24\%}                & {69,22\%}                 & {64,81\%}                 \\
+        {4} & {16}  & \textbf{86,75\%}         & \textbf{89\%}            & \textbf{57,02\%}         & \textbf{44,33\%}         & \textbf{63,78\%}          & \textbf{48,21\%}          \\
+        {4} & {50}  & \textcolor{red}{90,86\%} & \textcolor{red}{90,5\%}  & {71,22\%}                & {61,09\%}                & {65,26\%}                 & {54,87\%}                 \\
+        {4} & {100} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,22\%} & \textcolor{red}{92,44\%} & {83,78\%}                & \textbf{59,98\%}          & {54,38\%}                 \\
+        {6} & {2}   & \textcolor{red}{90,06\%} & \textcolor{red}{90,46\%} & {71,62\%}                & {64,56\%}                & {70,67\%}                 & {61,35\%}                 \\
+        {6} & {10}  & \textbf{88,38\%}         & \textcolor{red}{90,24\%} & \textbf{63,68\%}         & \textbf{53,21\%}         & \textbf{64,22\%}          & \textbf{52,19\%}          \\
+        {6} & {16}  & \textbf{89,01\%}         & \textbf{89,74\%}         & {67,03\%}                & {56,73\%}                & {64,44\%}                 & \textbf{52,78\%}          \\
+        {6} & {50}  & \textcolor{red}{97,68\%} & \textcolor{red}{92,77\%} & \textcolor{red}{97,03\%} & {87,24\%}                & {68,51\%}                 & {64,27\%}                 \\
+        {6} & {100} & {89,84\%}                & {89,81\%}                & {69,75\%}                & {60,55\%}                & {65,19\%}                 & {54,93\%}                 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attribute generator: Dale-2}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{91,81\%} & \textcolor{red}{96,22\%} & {89,08\%}                & {88,2\%}                 & \textcolor{red}{95,91\%} & \textcolor{red}{90,3\%}  \\
+        {1} & {10}  & {83,4\%}                 & {82,11\%}                & {79,68\%}                & {67,88\%}                & {77,12\%}                & {66,75\%}                \\
+        {1} & {16}  & {86,76\%}                & {87,06\%}                & {83,08\%}                & {75,03\%}                & {83,79\%}                & {74,72\%}                \\
+        {1} & {50}  & {83,12\%}                & {82\%}                   & {78,81\%}                & {66,76\%}                & {76,95\%}                & {65,95\%}                \\
+        {1} & {100} & \textcolor{red}{93,84\%} & {87,09\%}                & \textcolor{red}{93,01\%} & {83,36\%}                & {82,19\%}                & {79,14\%}                \\
+        {2} & {2}   & {88,86\%}                & {86,82\%}                & {86,62\%}                & {78,45\%}                & {82,59\%}                & {76,49\%}                \\
+        {2} & {10}  & {83,91\%}                & {81,57\%}                & {79,76\%}                & {67,89\%}                & {76,47\%}                & {66,1\%}                 \\
+        {2} & {16}  & \textbf{75,6\%}          & \textbf{77,74\%}         & \textbf{68,78\%}         & \textbf{55,43\%}         & \textbf{72,24\%}         & \textbf{56,24\%}         \\
+        {2} & {50}  & {85,1\%}                 & {83,08\%}                & {81,27\%}                & {70,13\%}                & {77,86\%}                & {68,54\%}                \\
+        {2} & {100} & {83,24\%}                & {81,87\%}                & {79,02\%}                & {67,32\%}                & {77,06\%}                & {66,19\%}                \\
+        {3} & {2}   & \textcolor{red}{97,21\%} & \textcolor{red}{96,08\%} & \textcolor{red}{96,43\%} & \textcolor{red}{95,23\%} & \textcolor{red}{95,18\%} & \textcolor{red}{94,24\%} \\
+        {3} & {10}  & {82,66\%}                & {81,11\%}                & {78,38\%}                & {66,21\%}                & {75,68\%}                & {64,52\%}                \\
+        {3} & {16}  & \textcolor{red}{93,26\%} & {86,09\%}                & \textcolor{red}{92,67\%} & {82,25\%}                & {81,17\%}                & {77,84\%}                \\
+        {3} & {50}  & {82,83\%}                & {81,44\%}                & {78,57\%}                & {66,29\%}                & {76,04\%}                & {64,84\%}                \\
+        {3} & {100} & {82,74\%}                & {81\%}                   & {78,39\%}                & {66,04\%}                & {75,44\%}                & {64,91\%}                \\
+        {4} & {2}   & \textcolor{red}{92,91\%} & {87,03\%}                & \textcolor{red}{92,16\%} & {84,06\%}                & {82,45\%}                & {79,57\%}                \\
+        {4} & {10}  & \textbf{79,3\%}          & \textbf{79,42\%}         & \textbf{74,65\%}         & \textbf{61,71\%}         & \textbf{74,12\%}         & \textbf{60,62\%}         \\
+        {4} & {16}  & {82,55\%}                & {81,31\%}                & {78,61\%}                & {66,51\%}                & {75,61\%}                & {65,05\%}                \\
+        {4} & {50}  & {88,51\%}                & {84,17\%}                & {85,77\%}                & {75,07\%}                & {79\%}                   & {71,93\%}                \\
+        {4} & {100} & {82,57\%}                & {81,61\%}                & {78,06\%}                & {66,33\%}                & {76,32\%}                & {65,54\%}                \\
+        {6} & {2}   & \textcolor{red}{92,36\%} & {88,15\%}                & \textcolor{red}{91,49\%} & {84,38\%}                & {85,28\%}                & {80,85\%}                \\
+        {6} & {10}  & \textbf{78,11\%}         & \textbf{79,19\%}         & \textbf{72,81\%}         & \textbf{59,86\%}         & \textbf{73,95\%}         & \textbf{60,17\%}         \\
+        {6} & {16}  & {82,35\%}                & {81,09\%}                & {78,84\%}                & {66,02\%}                & {75,9\%}                 & {64,79\%}                \\
+        {6} & {50}  & {82,04\%}                & {81,49\%}                & {78,34\%}                & {66,25\%}                & {76,46\%}                & {64,91\%}                \\
+        {6} & {100} & {82,09\%}                & {80,73\%}                & {77,42\%}                & {65,3\%}                 & {75,52\%}                & {64,52\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attribute generator: Dale-5}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si} & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C} \\\midrule
+        {1} & {2}   & \textcolor{red}{99,92\%} & \textcolor{red}{99,88\%} & \textbf{79,83\%}     & \textbf{82,7\%}      & \textcolor{red}{99,71\%} & \textbf{82,14\%}     \\
+        {1} & {10}  & \textbf{70,52\%}         & \textbf{70,36\%}         & \textbf{80,23\%}     & \textbf{82,36\%}     & \textbf{74,49\%}         & \textbf{81,77\%}     \\
+        {1} & {16}  & \textbf{77,25\%}         & \textbf{77,16\%}         & {84,55\%}            & {86,2\%}             & \textbf{80,72\%}         & {85,56\%}            \\
+        {3} & {10}  & \textbf{71,46\%}         & \textbf{71,47\%}         & \textbf{78,81\%}     & \textbf{81,87\%}     & \textbf{75,95\%}         & \textbf{80,93\%}     \\
+        {6} & {2}   & {83,25\%}                & {83,23\%}                & {87,74\%}            & {88,95\%}            & {85,38\%}                & {88,87\%}            \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attribute generator: Color}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{95,07\%} & \textcolor{red}{98,06\%}          & {86,63\%}                & {85,3\%}                 & \textcolor{red}{96,26\%} & \textcolor{red}{93,32\%} \\
+        {1} & {10}  & \textcolor{red}{90,8\%}  & \textcolor{red}{96,4\%}           & {71,88\%}                & {69,29\%}                & \textcolor{red}{94,05\%} & {85,13\%}                \\
+        {1} & {16}  & \textcolor{red}{91,47\%} & \textcolor{red}{96,96\%}          & {71,2\%}                 & {66,82\%}                & \textcolor{red}{93,83\%} & {83,35\%}                \\
+        {1} & {50}  & \textcolor{red}{93,26\%} & \textcolor{red}{96,63\%}          & {77,53\%}                & {71,94\%}                & \textcolor{red}{93,56\%} & {84,62\%}                \\
+        {1} & {100} & \textcolor{red}{93,5\%}  & \textcolor{red}{97,86\%}          & {77,42\%}                & {71,93\%}                & \textcolor{red}{94,68\%} & {85,77\%}                \\
+        {2} & {2}   & \textcolor{red}{93,57\%} & \textcolor{red}{97,72\%}          & {80,89\%}                & {79,8\%}                 & \textcolor{red}{96,08\%} & \textcolor{red}{90,71\%} \\
+        {2} & {10}  & \textbf{84,86\%}         & \textcolor{red}{\textbf{92,07\%}} & {58,8\%}                 & \textbf{52,09\%}         & \textbf{88,36\%}         & {77,01\%}                \\
+        {2} & {16}  & {87,29\%}                & \textcolor{red}{92,88\%}          & {68,52\%}                & {71,5\%}                 & \textcolor{red}{92,56\%} & {83,9\%}                 \\
+        {2} & {50}  & {89,15\%}                & \textcolor{red}{95\%}             & {68,19\%}                & {68,93\%}                & \textcolor{red}{93,11\%} & {83,95\%}                \\
+        {2} & {100} & \textcolor{red}{91,06\%} & \textcolor{red}{96,14\%}          & {71,98\%}                & {67,23\%}                & \textcolor{red}{92,81\%} & {82,71\%}                \\
+        {3} & {2}   & \textcolor{red}{93,05\%} & \textcolor{red}{97,15\%}          & {82,41\%}                & {77,36\%}                & \textcolor{red}{94,43\%} & {88,8\%}                 \\
+        {3} & {10}  & {88,52\%}                & \textcolor{red}{95,41\%}          & {66,61\%}                & {63,63\%}                & \textcolor{red}{92,25\%} & {81,46\%}                \\
+        {3} & {16}  & {85,71\%}                & \textcolor{red}{92,88\%}          & \textbf{56,44\%}         & \textbf{49,22\%}         & \textbf{88,87\%}         & \textbf{74,54\%}         \\
+        {3} & {50}  & \textcolor{red}{90,78\%} & \textcolor{red}{96,28\%}          & {68,17\%}                & {64,84\%}                & \textcolor{red}{94,16\%} & {83,5\%}                 \\
+        {3} & {100} & {86,54\%}                & \textcolor{red}{93,54\%}          & {65,9\%}                 & {66,03\%}                & \textcolor{red}{92,49\%} & {83,28\%}                \\
+        {4} & {10}  & {87,61\%}                & \textcolor{red}{94,63\%}          & {66,83\%}                & {64,85\%}                & \textcolor{red}{92,24\%} & {82,85\%}                \\
+        {4} & {16}  & \textbf{85,7\%}          & \textcolor{red}{93,56\%}          & \textbf{56,9\%}          & \textbf{52,1\%}          & \textbf{88,98\%}         & \textbf{76,15\%}         \\
+        {4} & {50}  & {87,55\%}                & \textcolor{red}{94,24\%}          & {61,03\%}                & {55,82\%}                & \textcolor{red}{91,73\%} & \textbf{76,52\%}         \\
+        {4} & {100} & {88,84\%}                & \textcolor{red}{94,84\%}          & {68,68\%}                & {66,73\%}                & \textcolor{red}{92,57\%} & {81,79\%}                \\
+        {6} & {2}   & \textcolor{red}{93,48\%} & \textcolor{red}{98,14\%}          & {81,36\%}                & {79,08\%}                & \textcolor{red}{96,09\%} & \textcolor{red}{90,95\%} \\
+        {6} & {10}  & \textbf{85\%}            & \textcolor{red}{\textbf{92,44\%}} & \textbf{57,21\%}         & {55,07\%}                & {89,39\%}                & {77,44\%}                \\
+        {6} & {16}  & {88,02\%}                & \textcolor{red}{94,22\%}          & {71,42\%}                & {66,35\%}                & \textcolor{red}{90,86\%} & {84,12\%}                \\
+        {6} & {50}  & \textcolor{red}{98,57\%} & \textcolor{red}{99,45\%}          & \textcolor{red}{95,57\%} & \textcolor{red}{95,06\%} & \textcolor{red}{99,16\%} & \textcolor{red}{96,64\%} \\
+        {6} & {100} & {86,44\%}                & \textcolor{red}{\textbf{92,76\%}} & {65,59\%}                & {65,93\%}                & \textcolor{red}{92,42\%} & {82,16\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{RE generator: Dale-2}
+    \label{TODO: label}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{92,86\%} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,06\%} & {89,5\%}             & \textcolor{red}{95,73\%} & \textcolor{red}{90,93\%} \\
+        {1} & {10}  & {83,9\%}                 & \textcolor{red}{90,9\%}  & {78,63\%}                & {74,74\%}            & {89,6\%}                 & {78,99\%}                \\
+        {1} & {16}  & {84,68\%}                & \textcolor{red}{91,65\%} & {80,11\%}                & {77,03\%}            & \textcolor{red}{90,16\%} & {81,05\%}                \\
+        {1} & {50}  & {84,85\%}                & \textcolor{red}{91,67\%} & {79,95\%}                & {77,11\%}            & \textcolor{red}{90,32\%} & {81,04\%}                \\
+        {2} & {2}   & \textcolor{red}{90,51\%} & \textcolor{red}{95,36\%} & {87,5\%}                 & {86,43\%}            & \textcolor{red}{94,77\%} & {88,35\%}                \\
+        {2} & {10}  & {82,44\%}                & \textcolor{red}{90,32\%} & {77,65\%}                & {73,73\%}            & {88,89\%}                & {78,41\%}                \\
+        {2} & {16}  & {81,85\%}                & \textbf{89,4\%}          & {76,07\%}                & \textbf{71,66\%}     & {87,9\%}                 & \textbf{75,91\%}         \\
+        {2} & {50}  & {83,58\%}                & \textcolor{red}{91,12\%} & {78,67\%}                & {75,03\%}            & {89,94\%}                & {79,06\%}                \\
+        {3} & {2}   & \textcolor{red}{92,51\%} & \textcolor{red}{95,96\%} & {89,96\%}                & {88,7\%}             & \textcolor{red}{95,18\%} & \textcolor{red}{90,43\%} \\
+        {3} & {10}  & {82,11\%}                & {89,54\%}                & {75,9\%}                 & {71,78\%}            & {88,18\%}                & {76,72\%}                \\
+        {3} & {50}  & {83,12\%}                & \textcolor{red}{90,82\%} & {77,62\%}                & {74,08\%}            & {88,88\%}                & {78,74\%}                \\
+        {3} & {100} & {83,74\%}                & {89,59\%}                & {79,29\%}                & {74,48\%}            & \textbf{87,46\%}         & {78,05\%}                \\
+        {4} & {2}   & {88,86\%}                & \textcolor{red}{94\%}    & {84,8\%}                 & {83,19\%}            & \textcolor{red}{93,37\%} & {85,92\%}                \\
+        {4} & {10}  & \textbf{79,93\%}         & \textbf{89,29\%}         & \textbf{74,1\%}          & \textbf{69,5\%}      & \textbf{87,19\%}         & \textbf{74,79\%}         \\
+        {4} & {16}  & \textbf{81,26\%}         & {89,68\%}                & \textbf{75,73\%}         & {71,96\%}            & {88,4\%}                 & {76,93\%}                \\
+        {4} & {50}  & {81,55\%}                & {89,42\%}                & {75,9\%}                 & {71,68\%}            & {88,17\%}                & {76,55\%}                \\
+        {6} & {2}   & \textcolor{red}{91,09\%} & \textcolor{red}{95,26\%} & {87,86\%}                & {86,81\%}            & \textcolor{red}{94,78\%} & {88,82\%}                \\
+        {6} & {50}  & \textbf{80,39\%}         & \textbf{87,85\%}         & \textbf{75,01\%}         & \textbf{68,71\%}     & \textbf{84,54\%}         & \textbf{72,93\%}         \\
+        \bottomrule
+    \end{tabular}
+    \caption{RE generator: Dale-5}
+    \label{TODO: label}
+\end{table}

### 2024-04-11 source/mlt/language_games/print_message_analysis.py

@@ -134,6 +134,19 @@ class Experiments:
 
         return dataset, experiment_variables
 
+    def filter_variables(self, variables):
+        renamed_variables = {
+            variable_dict[key]: float(value) for key, value in variables.items()
+        }
+
+        self.experiments = [
+            experiment
+            for experiment in self.experiments
+            if set(renamed_variables.items()).issubset(
+                set(experiment.variables.items())
+            )
+        ]
+
     def exclude_folders(self):
         excluded_folders = self._get_excluded_folders()
 
@@ -273,7 +286,10 @@ class LatexOutputProcessor(OutputProcessor):
         number_combinations = len(all_combinations)
 
         combination_header = " & ".join(
-            [rf"\textbf{{{combination}}}" for combination in all_combinations]
+            [
+                rf"\textbf{{{self.format_combination(combination)}}}"
+                for combination in all_combinations
+            ]
         )
 
         result_lines = []
@@ -321,6 +337,10 @@ class LatexOutputProcessor(OutputProcessor):
 
         print(latex_stub)
 
+    def format_combination(self, combination):
+        abbr = {"color": "C", "shape": "Sh", "size": "Si"}
+        return f"{' > '.join([abbr[attribute] for attribute in combination])}"
+
 
 METRIC_TYPES = {
     "%": OutputProcessor.to_percent,
@@ -363,8 +383,13 @@ if __name__ == "__main__":
         help="how the output should be printed",
     )
 
-    args = parser.parse_args()
+    args, additional_args = parser.parse_known_args()
     print(args)
+    additional_args = {
+        key.removeprefix("--"): value
+        for key, value in zip(additional_args[::2], additional_args[1::2])
+    }
+    print(additional_args)
 
     baseline_runs = {}
     english_runs = {}
@@ -387,6 +412,7 @@ if __name__ == "__main__":
     all_experiments = Experiments(
         args.folder, args.datasets, baseline_runs, english_runs
     )
+    all_experiments.filter_variables(additional_args)
     all_experiments.exclude_folders()
     all_experiments.sort_variables()
 

### 2024-04-30 thesis/sections/7-language-analysis.tex
language analysis

@@
 The languages that emerge are analyzed in two ways.
+First, for each language that emerges through successful communication, a score is calculated that represents the similarity of the messages to valid English referring expressions.
+In a second step, several languages are analyzed qualitatively, to determine what certain symbols mean and what they might refer to in the visual scenes.
+\cmtDK[inline]{talks about why and which few languages}
+
+To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
+Probing can be used to analyze and interpret the hidden representations in neural networks.
+Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
+If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
+If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
+In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
+Recall, that this incremental is based on an order in salience for the attributes.
+In English, \citet{Dale1995} propose the order of \emph{shape} > \emph{color} > \emph{size}.
+However, there is no policy for the agents, that enforces specifically this order or indeed any salience order.
+For this reason, the emerged languages are compared to all possible salience orders, given the three attributes \emph{shape}, \emph{color} and \emph{size}.
+This will show if first, the languages follow any salience order and secondly, which attributes are more important for the agents to communicate.
+% SD: The probing test is to show how well the emerged language can be trasnlated into English descriptions consistting of labels of scene features using Dale and Reiter idea of incrementality.
+However, we would need to know now what was the policy of the sender to generate message.
+Was this similar to Dale and Reiter to enforce longer descriptions?
+% DK: done
+
+The probing model consists of an encoder LSTM, one linear layer and a decoder LSTM.
+The encoder LSTM encodes the messages of the emerged language.
+Its final hidden state is used as a representation of the meaning of the complete message and passed through the linear layer, which learns an abstract representation of the message.
+The dimension of the linear layer is kept small to prevent memorizing the dataset and force the model to find abstract connections between the input and the target.
+The resulting vector is used as the initial state of the decoder LSTM.
+The decoder LSTM is then trained to produce the English referring expression.
+Both LSTMs have an embedding dimension of $LSTM_e=10$ and a hidden size of $LSTM_{out}=10$, while the linear layer has a size of $l=10$.
+During decoding, teacher forcing is applied.
+The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
+Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
+The models are trained for 200 epochs with a learning rate of $lr=0,004$ and a batch size of 32 samples per batch.
+
+The translation model will learn two characteristics with this setup.
+First, it can learn to find correlations between the emerged language and the English referring expressions, which is the goal of the probing.
+But secondly, it can also learn underlying patterns in the English referring expressions that are independent of the emerged language.
+For instance, it can learn that the referring expressions are likely to be two symbols long, or that for instance 'cube' is the most common shape.
+This second characteristic will lead to a low loss during training, even though there are no connections to the emerged language.
+In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
+For that reason, two reference baselines are calculated for each salience order of the attributes, to eliminate the influence of the second characteristic.
+% SD: Baselines?
+% DK: (done)
+The first baseline $L_{baseline}$ uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
+By this, the model is trained to learn the structural patterns in the English referring expressions.
+The resulting loss is the highest possible loss the model can achieve, independent of the input.
+Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
+If not, the loss will be as high as this baseline.
+On the other side, the lowest possible loss $L_{English}$ is calculated, by using the English referring expressions as both input and target.
+This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since the input is perfectly correlated to the target).
+These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
+A loss close to $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
+% SD: Is this encessary? But when and where is this loss using 0 inouts calculated?
+% DK: TODO, QUESTION
+By doing this, all configurations can be compared directly to each other.
+The lower $L_{norm}$, the higher the correlation between the emerged language and English referring expressions.
+
+Tables \ref{tab:probing:discriminator:dale-2} to \ref{tab:probing:attention-predictor:colour} show the normalized losses $L_{norm}$ of the probing for the different tasks on all three datasets.
+Hereby, only languages are listed for experiments in which the agents communicated successfully, that is $L_{norm}$ is lower than 100\%.
+The first thing that can be noticed is that across almost all languages emerged through successful communication, $L_{norm}$ is lower than 100\% with at least one salience order, which shows that they make use of the attributes in some way.
+On the other side, no language reaches an $L_{norm}$ close to 0, that is no language correlates perfectly with English referring expressions.
+The closest is a language with $L_{norm}$ just above 40\%.
+This shows that while many emerged languages have similarities with English, they make use of either other additional visual characteristics in the scenes or use different linguistic structures.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{95,5\%}           & \textcolor{red}{97,57\%}          & {86,98\%}                & {81,43\%}                & \textcolor{red}{91,67\%} & {86,45\%}                \\
+        {1} & {10}  & \textcolor{red}{96,11\%}          & \textcolor{red}{96,9\%}           & {89,61\%}                & {85\%}                   & {89,4\%}                 & {84,49\%}                \\
+        {1} & {16}  & \textcolor{red}{95,19\%}          & \textcolor{red}{96,56\%}          & {87,25\%}                & {81,22\%}                & {89,34\%}                & {83,13\%}                \\
+        {1} & {50}  & \textcolor{red}{96,61\%}          & \textcolor{red}{97,07\%}          & \textcolor{red}{91,26\%} & {85,75\%}                & {87,14\%}                & {82,61\%}                \\
+        {1} & {100} & \textcolor{red}{95,91\%}          & \textcolor{red}{96,73\%}          & \textcolor{red}{90,63\%} & {86,01\%}                & {88,95\%}                & {84,17\%}                \\
+        {2} & {2}   & \textcolor{red}{99,67\%}          & \textcolor{red}{99,91\%}          & \textcolor{red}{97,02\%} & \textcolor{red}{95,37\%} & \textcolor{red}{96,97\%} & \textcolor{red}{94,64\%} \\
+        {2} & {10}  & \textcolor{red}{96,36\%}          & \textcolor{red}{96,34\%}          & \textcolor{red}{91,32\%} & {86,75\%}                & {87,69\%}                & {82,87\%}                \\
+        {2} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{97,12\%}          & {88,11\%}                & {83,21\%}                & {89,86\%}                & {85,35\%}                \\
+        {2} & {50}  & \textcolor{red}{95,34\%}          & \textcolor{red}{97,3\%}           & {85,48\%}                & {80,89\%}                & \textcolor{red}{92,22\%} & {86,18\%}                \\
+        {2} & {100} & \textcolor{red}{94,88\%}          & \textcolor{red}{96,32\%}          & {87,37\%}                & {81,17\%}                & {88,35\%}                & {82,61\%}                \\
+        {3} & {2}   & \textcolor{red}{\textbf{94,03\%}} & \textcolor{red}{96,96\%}          & \textbf{79,59\%}         & \textbf{74,54\%}         & \textcolor{red}{92,32\%} & {85,63\%}                \\
+        {3} & {10}  & \textcolor{red}{96,1\%}           & \textcolor{red}{95,97\%}          & \textcolor{red}{90,45\%} & {84,84\%}                & {85,98\%}                & {80,5\%}                 \\
+        {3} & {16}  & \textcolor{red}{95,71\%}          & \textcolor{red}{\textbf{95,1\%}}  & \textcolor{red}{90,97\%} & {85,33\%}                & {81\%}                   & \textbf{75,56\%}         \\
+        {3} & {50}  & \textcolor{red}{96,32\%}          & \textcolor{red}{\textbf{94,55\%}} & \textcolor{red}{92,5\%}  & {86,81\%}                & \textbf{80,48\%}         & {75,73\%}                \\
+        {3} & {100} & \textcolor{red}{95,33\%}          & \textcolor{red}{97,82\%}          & {86,05\%}                & {81,33\%}                & \textcolor{red}{93,91\%} & {88,54\%}                \\
+        {4} & {2}   & \textcolor{red}{96,81\%}          & \textcolor{red}{95,12\%}          & \textcolor{red}{93,03\%} & {87,41\%}                & \textbf{80,73\%}         & \textbf{75,47\%}         \\
+        {4} & {10}  & \textcolor{red}{95,31\%}          & \textcolor{red}{97,26\%}          & {86,9\%}                 & {82,67\%}                & \textcolor{red}{92,39\%} & {86,98\%}                \\
+        {4} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{\textbf{94,65\%}} & \textcolor{red}{91,99\%} & {86,65\%}                & \textbf{79,23\%}         & \textbf{74,54\%}         \\
+        {4} & {50}  & \textcolor{red}{94,76\%}          & \textcolor{red}{96,69\%}          & {86,8\%}                 & {81,33\%}                & {89,65\%}                & {83,92\%}                \\
+        {4} & {100} & \textcolor{red}{96,16\%}          & \textcolor{red}{97,94\%}          & {89,57\%}                & {85,18\%}                & \textcolor{red}{92,56\%} & {87,74\%}                \\
+        {6} & {2}   & \textcolor{red}{95,73\%}          & \textcolor{red}{96,99\%}          & {87,77\%}                & {83,27\%}                & \textcolor{red}{91,77\%} & {86,34\%}                \\
+        {6} & {10}  & \textcolor{red}{\textbf{94,59\%}} & \textcolor{red}{99,89\%}          & {86,24\%}                & {81,46\%}                & \textcolor{red}{91,42\%} & {86,43\%}                \\
+        {6} & {16}  & \textcolor{red}{95,51\%}          & \textcolor{red}{96,3\%}           & {88,89\%}                & {83,24\%}                & {87,61\%}                & {82,71\%}                \\
+        {6} & {50}  & \textcolor{red}{\textbf{94,66\%}} & \textcolor{red}{97,43\%}          & \textbf{84,93\%}         & \textbf{78,61\%}         & \textcolor{red}{91,48\%} & {83,66\%}                \\
+        {6} & {100} & \textcolor{red}{95,37\%}          & \textcolor{red}{97,77\%}          & \textbf{83,75\%}         & \textbf{80,84\%}         & \textcolor{red}{93,96\%} & {89,02\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Discriminator: Dale-2}
+    \label{tab:probing:discriminator:dale-2}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
+        {1} & {2}   & \textcolor{red}{94,96\%} & \textcolor{red}{91,27\%} & \textcolor{red}{94,51\%} & \textcolor{red}{90,99\%} & {88,52\%}            & {88,17\%}            \\
+        {1} & {10}  & {84,34\%}                & {84,03\%}                & {80,91\%}                & {73,23\%}                & {79,95\%}            & {72,34\%}            \\
+        {1} & {16}  & {86,27\%}                & {84,96\%}                & {82,76\%}                & {75,35\%}                & {80,72\%}            & {74,04\%}            \\
+        {1} & {50}  & {85,95\%}                & {84,4\%}                 & {82,96\%}                & {74,97\%}                & {80,2\%}             & {73,54\%}            \\
+        {1} & {100} & {86,31\%}                & {85,58\%}                & {82,9\%}                 & {76,03\%}                & {81,64\%}            & {74,62\%}            \\
+        {2} & {10}  & \textbf{82,66\%}         & {83,23\%}                & \textbf{78,51\%}         & {71,01\%}                & {78,89\%}            & {70,61\%}            \\
+        {2} & {16}  & {89,78\%}                & {85,53\%}                & {87,68\%}                & {80,64\%}                & {81,06\%}            & {77,24\%}            \\
+        {2} & {50}  & {84,4\%}                 & {82,65\%}                & {80,62\%}                & {71,42\%}                & {77,53\%}            & {69,46\%}            \\
+        {2} & {100} & {89,65\%}                & {85,3\%}                 & {87,26\%}                & {79,91\%}                & {80,88\%}            & {77,1\%}             \\
+        {3} & {2}   & {87,19\%}                & {86,93\%}                & {83,99\%}                & {78,18\%}                & {83,65\%}            & {77,08\%}            \\
+        {3} & {10}  & {87,15\%}                & {84,67\%}                & {84,01\%}                & {76,24\%}                & {80,41\%}            & {74,11\%}            \\
+        {3} & {50}  & \textbf{82,98\%}         & {82,25\%}                & {78,85\%}                & {70,62\%}                & {77,42\%}            & {69,27\%}            \\
+        {3} & {100} & {83,78\%}                & {83,07\%}                & {80,23\%}                & {72,15\%}                & {78,94\%}            & {71,1\%}             \\
+        {4} & {2}   & {89,23\%}                & {87,43\%}                & {87,09\%}                & {81,45\%}                & {83,83\%}            & {79,11\%}            \\
+        {4} & {10}  & {88,66\%}                & {84,83\%}                & {85,9\%}                 & {78,98\%}                & {80,2\%}             & {75,64\%}            \\
+        {4} & {16}  & {83,14\%}                & \textbf{81,51\%}         & \textbf{78,15\%}         & \textbf{70,21\%}         & \textbf{77,32\%}     & \textbf{68,48\%}     \\
+        {4} & {50}  & {87,62\%}                & {84,38\%}                & {85,25\%}                & {77,29\%}                & {79,94\%}            & {74,03\%}            \\
+        {4} & {100} & \textcolor{red}{90,17\%} & {85,76\%}                & {87,68\%}                & {80,71\%}                & {81,37\%}            & {78,1\%}             \\
+        {6} & {10}  & \textbf{82,03\%}         & \textbf{80,51\%}         & \textbf{77,46\%}         & \textbf{68,63\%}         & \textbf{75,71\%}     & \textbf{67,15\%}     \\
+        {6} & {16}  & {84,73\%}                & \textbf{81,12\%}         & {80,99\%}                & \textbf{70,5\%}          & \textbf{76,06\%}     & \textbf{69,12\%}     \\
+        {6} & {50}  & {86,95\%}                & {85,13\%}                & {84,06\%}                & {76,95\%}                & {80,11\%}            & {74,51\%}            \\
+        {6} & {100} & {86,15\%}                & {82,19\%}                & {83,56\%}                & {75,56\%}                & {77,76\%}            & {72,47\%}            \\
+        \bottomrule
+    \end{tabular}
+    \caption{Discriminator: Dale-5}
+    \label{tab:probing:discriminator:dale-5}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{91,62\%} & \textcolor{red}{91,57\%} & \textcolor{red}{94,31\%} & \textcolor{red}{95,13\%} & \textcolor{red}{93,11\%} & \textcolor{red}{94,92\%} \\
+        {1} & {10}  & {79,08\%}                & {79,08\%}                & {85,04\%}                & {86,86\%}                & {82,37\%}                & {86,38\%}                \\
+        {1} & {16}  & {84,02\%}                & {83,96\%}                & {88,67\%}                & {89,65\%}                & {86,36\%}                & {89,36\%}                \\
+        {1} & {50}  & \textcolor{red}{93,88\%} & \textcolor{red}{93,74\%} & \textcolor{red}{95,03\%} & \textcolor{red}{95,49\%} & \textcolor{red}{94,64\%} & \textcolor{red}{95,35\%} \\
+        {2} & {10}  & {88,67\%}                & {88,36\%}                & {89,59\%}                & \textcolor{red}{90,77\%} & \textcolor{red}{90,01\%} & {89,81\%}                \\
+        {2} & {16}  & {81,13\%}                & {81,02\%}                & {86,16\%}                & {87,97\%}                & {83,85\%}                & {87,56\%}                \\
+        {3} & {2}   & {83,38\%}                & {83,46\%}                & {88,21\%}                & {89,71\%}                & {86,05\%}                & {89,37\%}                \\
+        {3} & {10}  & \textbf{77\%}            & \textbf{76,77\%}         & \textbf{83,54\%}         & \textbf{85,57\%}         & \textbf{79,97\%}         & \textbf{84,9\%}          \\
+        {3} & {16}  & \textcolor{red}{93,52\%} & \textcolor{red}{93,6\%}  & \textcolor{red}{94,16\%} & \textcolor{red}{94,14\%} & \textcolor{red}{93,75\%} & \textcolor{red}{93,55\%} \\
+        {4} & {10}  & \textbf{68,04\%}         & \textbf{66,51\%}         & \textbf{76,33\%}         & \textbf{78,56\%}         & \textbf{71,42\%}         & \textbf{78,56\%}         \\
+        {4} & {16}  & \textbf{75,1\%}          & \textbf{74,93\%}         & \textbf{81,2\%}          & \textbf{83,35\%}         & \textbf{77,6\%}          & \textbf{82,76\%}         \\
+        {6} & {2}   & {84,21\%}                & {84,02\%}                & {88,41\%}                & {89,77\%}                & {86,61\%}                & {89,35\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Discriminator: Color}
+    \label{tab:probing:discriminator:colour}
+\end{table}
+
+Looking at the results for the discriminator, one can see that the results are very different for each dataset.
+On the 'Dale-2' dataset
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{95,07\%} & \textcolor{red}{98,06\%}          & {86,63\%}                & {85,3\%}                 & \textcolor{red}{96,26\%} & \textcolor{red}{93,32\%} \\
+        {1} & {10}  & \textcolor{red}{90,8\%}  & \textcolor{red}{96,4\%}           & {71,88\%}                & {69,29\%}                & \textcolor{red}{94,05\%} & {85,13\%}                \\
+        {1} & {16}  & \textcolor{red}{91,47\%} & \textcolor{red}{96,96\%}          & {71,2\%}                 & {66,82\%}                & \textcolor{red}{93,83\%} & {83,35\%}                \\
+        {1} & {50}  & \textcolor{red}{93,26\%} & \textcolor{red}{96,63\%}          & {77,53\%}                & {71,94\%}                & \textcolor{red}{93,56\%} & {84,62\%}                \\
+        {1} & {100} & \textcolor{red}{93,5\%}  & \textcolor{red}{97,86\%}          & {77,42\%}                & {71,93\%}                & \textcolor{red}{94,68\%} & {85,77\%}                \\
+        {2} & {2}   & \textcolor{red}{93,57\%} & \textcolor{red}{97,72\%}          & {80,89\%}                & {79,8\%}                 & \textcolor{red}{96,08\%} & \textcolor{red}{90,71\%} \\
+        {2} & {10}  & \textbf{84,86\%}         & \textcolor{red}{\textbf{92,07\%}} & {58,8\%}                 & \textbf{52,09\%}         & \textbf{88,36\%}         & {77,01\%}                \\
+        {2} & {16}  & {87,29\%}                & \textcolor{red}{92,88\%}          & {68,52\%}                & {71,5\%}                 & \textcolor{red}{92,56\%} & {83,9\%}                 \\
+        {2} & {50}  & {89,15\%}                & \textcolor{red}{95\%}             & {68,19\%}                & {68,93\%}                & \textcolor{red}{93,11\%} & {83,95\%}                \\
+        {2} & {100} & \textcolor{red}{91,06\%} & \textcolor{red}{96,14\%}          & {71,98\%}                & {67,23\%}                & \textcolor{red}{92,81\%} & {82,71\%}                \\
+        {3} & {2}   & \textcolor{red}{93,05\%} & \textcolor{red}{97,15\%}          & {82,41\%}                & {77,36\%}                & \textcolor{red}{94,43\%} & {88,8\%}                 \\
+        {3} & {10}  & {88,52\%}                & \textcolor{red}{95,41\%}          & {66,61\%}                & {63,63\%}                & \textcolor{red}{92,25\%} & {81,46\%}                \\
+        {3} & {16}  & {85,71\%}                & \textcolor{red}{92,88\%}          & \textbf{56,44\%}         & \textbf{49,22\%}         & \textbf{88,87\%}         & \textbf{74,54\%}         \\
+        {3} & {50}  & \textcolor{red}{90,78\%} & \textcolor{red}{96,28\%}          & {68,17\%}                & {64,84\%}                & \textcolor{red}{94,16\%} & {83,5\%}                 \\
+        {3} & {100} & {86,54\%}                & \textcolor{red}{93,54\%}          & {65,9\%}                 & {66,03\%}                & \textcolor{red}{92,49\%} & {83,28\%}                \\
+        {4} & {10}  & {87,61\%}                & \textcolor{red}{94,63\%}          & {66,83\%}                & {64,85\%}                & \textcolor{red}{92,24\%} & {82,85\%}                \\
+        {4} & {16}  & \textbf{85,7\%}          & \textcolor{red}{93,56\%}          & \textbf{56,9\%}          & \textbf{52,1\%}          & \textbf{88,98\%}         & \textbf{76,15\%}         \\
+        {4} & {50}  & {87,55\%}                & \textcolor{red}{94,24\%}          & {61,03\%}                & {55,82\%}                & \textcolor{red}{91,73\%} & \textbf{76,52\%}         \\
+        {4} & {100} & {88,84\%}                & \textcolor{red}{94,84\%}          & {68,68\%}                & {66,73\%}                & \textcolor{red}{92,57\%} & {81,79\%}                \\
+        {6} & {2}   & \textcolor{red}{93,48\%} & \textcolor{red}{98,14\%}          & {81,36\%}                & {79,08\%}                & \textcolor{red}{96,09\%} & \textcolor{red}{90,95\%} \\
+        {6} & {10}  & \textbf{85\%}            & \textcolor{red}{\textbf{92,44\%}} & \textbf{57,21\%}         & {55,07\%}                & {89,39\%}                & {77,44\%}                \\
+        {6} & {16}  & {88,02\%}                & \textcolor{red}{94,22\%}          & {71,42\%}                & {66,35\%}                & \textcolor{red}{90,86\%} & {84,12\%}                \\
+        {6} & {50}  & \textcolor{red}{98,57\%} & \textcolor{red}{99,45\%}          & \textcolor{red}{95,57\%} & \textcolor{red}{95,06\%} & \textcolor{red}{99,16\%} & \textcolor{red}{96,64\%} \\
+        {6} & {100} & {86,44\%}                & \textcolor{red}{\textbf{92,76\%}} & {65,59\%}                & {65,93\%}                & \textcolor{red}{92,42\%} & {82,16\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{RE generator: Dale-2}
+    \label{tab:probing:re-generator:dale-2}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{92,86\%} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,06\%} & {89,5\%}             & \textcolor{red}{95,73\%} & \textcolor{red}{90,93\%} \\
+        {1} & {10}  & {83,9\%}                 & \textcolor{red}{90,9\%}  & {78,63\%}                & {74,74\%}            & {89,6\%}                 & {78,99\%}                \\
+        {1} & {16}  & {84,68\%}                & \textcolor{red}{91,65\%} & {80,11\%}                & {77,03\%}            & \textcolor{red}{90,16\%} & {81,05\%}                \\
+        {1} & {50}  & {84,85\%}                & \textcolor{red}{91,67\%} & {79,95\%}                & {77,11\%}            & \textcolor{red}{90,32\%} & {81,04\%}                \\
+        {2} & {2}   & \textcolor{red}{90,51\%} & \textcolor{red}{95,36\%} & {87,5\%}                 & {86,43\%}            & \textcolor{red}{94,77\%} & {88,35\%}                \\
+        {2} & {10}  & {82,44\%}                & \textcolor{red}{90,32\%} & {77,65\%}                & {73,73\%}            & {88,89\%}                & {78,41\%}                \\
+        {2} & {16}  & {81,85\%}                & \textbf{89,4\%}          & {76,07\%}                & \textbf{71,66\%}     & {87,9\%}                 & \textbf{75,91\%}         \\
+        {2} & {50}  & {83,58\%}                & \textcolor{red}{91,12\%} & {78,67\%}                & {75,03\%}            & {89,94\%}                & {79,06\%}                \\
+        {3} & {2}   & \textcolor{red}{92,51\%} & \textcolor{red}{95,96\%} & {89,96\%}                & {88,7\%}             & \textcolor{red}{95,18\%} & \textcolor{red}{90,43\%} \\
+        {3} & {10}  & {82,11\%}                & {89,54\%}                & {75,9\%}                 & {71,78\%}            & {88,18\%}                & {76,72\%}                \\
+        {3} & {50}  & {83,12\%}                & \textcolor{red}{90,82\%} & {77,62\%}                & {74,08\%}            & {88,88\%}                & {78,74\%}                \\
+        {3} & {100} & {83,74\%}                & {89,59\%}                & {79,29\%}                & {74,48\%}            & \textbf{87,46\%}         & {78,05\%}                \\
+        {4} & {2}   & {88,86\%}                & \textcolor{red}{94\%}    & {84,8\%}                 & {83,19\%}            & \textcolor{red}{93,37\%} & {85,92\%}                \\
+        {4} & {10}  & \textbf{79,93\%}         & \textbf{89,29\%}         & \textbf{74,1\%}          & \textbf{69,5\%}      & \textbf{87,19\%}         & \textbf{74,79\%}         \\
+        {4} & {16}  & \textbf{81,26\%}         & {89,68\%}                & \textbf{75,73\%}         & {71,96\%}            & {88,4\%}                 & {76,93\%}                \\
+        {4} & {50}  & {81,55\%}                & {89,42\%}                & {75,9\%}                 & {71,68\%}            & {88,17\%}                & {76,55\%}                \\
+        {6} & {2}   & \textcolor{red}{91,09\%} & \textcolor{red}{95,26\%} & {87,86\%}                & {86,81\%}            & \textcolor{red}{94,78\%} & {88,82\%}                \\
+        {6} & {50}  & \textbf{80,39\%}         & \textbf{87,85\%}         & \textbf{75,01\%}         & \textbf{68,71\%}     & \textbf{84,54\%}         & \textbf{72,93\%}         \\
+        \bottomrule
+    \end{tabular}
+    \caption{RE generator: Dale-5}
+    \label{tab:probing:re-generator:dale-5}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}      & \textbf{Si > Sh > C}      \\\midrule
+        {1} & {2}   & \textcolor{red}{95,73\%} & \textcolor{red}{94,23\%} & \textcolor{red}{92,96\%} & \textcolor{red}{90,59\%} & {81,59\%}                 & {78,21\%}                 \\
+        {1} & {10}  & \textcolor{red}{91,77\%} & \textcolor{red}{91,19\%} & {73,14\%}                & {63,99\%}                & {67,45\%}                 & {57,55\%}                 \\
+        {1} & {16}  & \textcolor{red}{91,74\%} & \textcolor{red}{90,87\%} & {73,07\%}                & {63,78\%}                & {66,16\%}                 & {55,48\%}                 \\
+        {1} & {50}  & \textcolor{red}{97,78\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,22\%} & {87,45\%}                & {69,23\%}                 & {64,97\%}                 \\
+        {1} & {100} & \textcolor{red}{97,91\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,18\%} & {87,35\%}                & {69,25\%}                 & {64,99\%}                 \\
+        {2} & {2}   & \textcolor{red}{93,17\%} & \textcolor{red}{91,69\%} & {81,96\%}                & {75,97\%}                & {72,05\%}                 & {64,29\%}                 \\
+        {2} & {10}  & \textcolor{red}{94,07\%} & \textcolor{red}{98,04\%} & {75,32\%}                & {71,91\%}                & \textcolor{red}{94,38\%}  & {86,62\%}                 \\
+        {2} & {16}  & \textcolor{red}{97,72\%} & \textcolor{red}{92,9\%}  & \textcolor{red}{94,18\%} & {87,09\%}                & {68,54\%}                 & {64,13\%}                 \\
+        {2} & {50}  & \textcolor{red}{97,87\%} & \textcolor{red}{93,08\%} & \textcolor{red}{94,25\%} & {87,54\%}                & {69,2\%}                  & {65,17\%}                 \\
+        {2} & {100} & {89,6\%}                 & \textbf{89,58\%}         & \textbf{65,65\%}         & \textbf{55,69\%}         & {64,92\%}                 & {53,53\%}                 \\
+        {3} & {2}   & \textcolor{red}{97,73\%} & \textcolor{red}{99,28\%} & \textcolor{red}{95,84\%} & \textcolor{red}{97,72\%} & \textcolor{red}{101,58\%} & \textcolor{red}{100,22\%} \\
+        {3} & {10}  & \textcolor{red}{91,04\%} & \textcolor{red}{90,52\%} & {71,06\%}                & {61,36\%}                & {65,59\%}                 & {54,93\%}                 \\
+        {3} & {16}  & \textcolor{red}{91,09\%} & \textcolor{red}{90,3\%}  & {71,71\%}                & {62,42\%}                & {66,28\%}                 & {56,16\%}                 \\
+        {3} & {50}  & \textcolor{red}{97,07\%} & \textcolor{red}{91,72\%} & \textcolor{red}{92,7\%}  & {85,33\%}                & {66,38\%}                 & {61,26\%}                 \\
+        {3} & {100} & \textcolor{red}{97,81\%} & \textcolor{red}{93,01\%} & \textcolor{red}{94,16\%} & {87,3\%}                 & {68,83\%}                 & {64,6\%}                  \\
+        {4} & {2}   & \textcolor{red}{91,95\%} & \textcolor{red}{96,87\%} & {73,55\%}                & {72,66\%}                & \textcolor{red}{98,37\%}  & \textcolor{red}{90,23\%}  \\
+        {4} & {10}  & \textcolor{red}{97,7\%}  & \textcolor{red}{93,1\%}  & \textcolor{red}{94,16\%} & {87,24\%}                & {69,22\%}                 & {64,81\%}                 \\
+        {4} & {16}  & \textbf{86,75\%}         & \textbf{89\%}            & \textbf{57,02\%}         & \textbf{44,33\%}         & \textbf{63,78\%}          & \textbf{48,21\%}          \\
+        {4} & {50}  & \textcolor{red}{90,86\%} & \textcolor{red}{90,5\%}  & {71,22\%}                & {61,09\%}                & {65,26\%}                 & {54,87\%}                 \\
+        {4} & {100} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,22\%} & \textcolor{red}{92,44\%} & {83,78\%}                & \textbf{59,98\%}          & {54,38\%}                 \\
+        {6} & {2}   & \textcolor{red}{90,06\%} & \textcolor{red}{90,46\%} & {71,62\%}                & {64,56\%}                & {70,67\%}                 & {61,35\%}                 \\
+        {6} & {10}  & \textbf{88,38\%}         & \textcolor{red}{90,24\%} & \textbf{63,68\%}         & \textbf{53,21\%}         & \textbf{64,22\%}          & \textbf{52,19\%}          \\
+        {6} & {16}  & \textbf{89,01\%}         & \textbf{89,74\%}         & {67,03\%}                & {56,73\%}                & {64,44\%}                 & \textbf{52,78\%}          \\
+        {6} & {50}  & \textcolor{red}{97,68\%} & \textcolor{red}{92,77\%} & \textcolor{red}{97,03\%} & {87,24\%}                & {68,51\%}                 & {64,27\%}                 \\
+        {6} & {100} & {89,84\%}                & {89,81\%}                & {69,75\%}                & {60,55\%}                & {65,19\%}                 & {54,93\%}                 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attribute generator: Dale-2}
+    \label{tab:probing:attribute-generator:dale-2}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{91,81\%} & \textcolor{red}{96,22\%} & {89,08\%}                & {88,2\%}                 & \textcolor{red}{95,91\%} & \textcolor{red}{90,3\%}  \\
+        {1} & {10}  & {83,4\%}                 & {82,11\%}                & {79,68\%}                & {67,88\%}                & {77,12\%}                & {66,75\%}                \\
+        {1} & {16}  & {86,76\%}                & {87,06\%}                & {83,08\%}                & {75,03\%}                & {83,79\%}                & {74,72\%}                \\
+        {1} & {50}  & {83,12\%}                & {82\%}                   & {78,81\%}                & {66,76\%}                & {76,95\%}                & {65,95\%}                \\
+        {1} & {100} & \textcolor{red}{93,84\%} & {87,09\%}                & \textcolor{red}{93,01\%} & {83,36\%}                & {82,19\%}                & {79,14\%}                \\
+        {2} & {2}   & {88,86\%}                & {86,82\%}                & {86,62\%}                & {78,45\%}                & {82,59\%}                & {76,49\%}                \\
+        {2} & {10}  & {83,91\%}                & {81,57\%}                & {79,76\%}                & {67,89\%}                & {76,47\%}                & {66,1\%}                 \\
+        {2} & {16}  & \textbf{75,6\%}          & \textbf{77,74\%}         & \textbf{68,78\%}         & \textbf{55,43\%}         & \textbf{72,24\%}         & \textbf{56,24\%}         \\
+        {2} & {50}  & {85,1\%}                 & {83,08\%}                & {81,27\%}                & {70,13\%}                & {77,86\%}                & {68,54\%}                \\
+        {2} & {100} & {83,24\%}                & {81,87\%}                & {79,02\%}                & {67,32\%}                & {77,06\%}                & {66,19\%}                \\
+        {3} & {2}   & \textcolor{red}{97,21\%} & \textcolor{red}{96,08\%} & \textcolor{red}{96,43\%} & \textcolor{red}{95,23\%} & \textcolor{red}{95,18\%} & \textcolor{red}{94,24\%} \\
+        {3} & {10}  & {82,66\%}                & {81,11\%}                & {78,38\%}                & {66,21\%}                & {75,68\%}                & {64,52\%}                \\
+        {3} & {16}  & \textcolor{red}{93,26\%} & {86,09\%}                & \textcolor{red}{92,67\%} & {82,25\%}                & {81,17\%}                & {77,84\%}                \\
+        {3} & {50}  & {82,83\%}                & {81,44\%}                & {78,57\%}                & {66,29\%}                & {76,04\%}                & {64,84\%}                \\
+        {3} & {100} & {82,74\%}                & {81\%}                   & {78,39\%}                & {66,04\%}                & {75,44\%}                & {64,91\%}                \\
+        {4} & {2}   & \textcolor{red}{92,91\%} & {87,03\%}                & \textcolor{red}{92,16\%} & {84,06\%}                & {82,45\%}                & {79,57\%}                \\
+        {4} & {10}  & \textbf{79,3\%}          & \textbf{79,42\%}         & \textbf{74,65\%}         & \textbf{61,71\%}         & \textbf{74,12\%}         & \textbf{60,62\%}         \\
+        {4} & {16}  & {82,55\%}                & {81,31\%}                & {78,61\%}                & {66,51\%}                & {75,61\%}                & {65,05\%}                \\
+        {4} & {50}  & {88,51\%}                & {84,17\%}                & {85,77\%}                & {75,07\%}                & {79\%}                   & {71,93\%}                \\
+        {4} & {100} & {82,57\%}                & {81,61\%}                & {78,06\%}                & {66,33\%}                & {76,32\%}                & {65,54\%}                \\
+        {6} & {2}   & \textcolor{red}{92,36\%} & {88,15\%}                & \textcolor{red}{91,49\%} & {84,38\%}                & {85,28\%}                & {80,85\%}                \\
+        {6} & {10}  & \textbf{78,11\%}         & \textbf{79,19\%}         & \textbf{72,81\%}         & \textbf{59,86\%}         & \textbf{73,95\%}         & \textbf{60,17\%}         \\
+        {6} & {16}  & {82,35\%}                & {81,09\%}                & {78,84\%}                & {66,02\%}                & {75,9\%}                 & {64,79\%}                \\
+        {6} & {50}  & {82,04\%}                & {81,49\%}                & {78,34\%}                & {66,25\%}                & {76,46\%}                & {64,91\%}                \\
+        {6} & {100} & {82,09\%}                & {80,73\%}                & {77,42\%}                & {65,3\%}                 & {75,52\%}                & {64,52\%}                \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attribute generator: Dale-5}
+    \label{tab:probing:attribute-generator:dale-5}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si} & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C} \\\midrule
+        {1} & {2}   & \textcolor{red}{99,92\%} & \textcolor{red}{99,88\%} & \textbf{79,83\%}     & \textbf{82,7\%}      & \textcolor{red}{99,71\%} & \textbf{82,14\%}     \\
+        {1} & {10}  & \textbf{70,52\%}         & \textbf{70,36\%}         & \textbf{80,23\%}     & \textbf{82,36\%}     & \textbf{74,49\%}         & \textbf{81,77\%}     \\
+        {1} & {16}  & \textbf{77,25\%}         & \textbf{77,16\%}         & {84,55\%}            & {86,2\%}             & \textbf{80,72\%}         & {85,56\%}            \\
+        {3} & {10}  & \textbf{71,46\%}         & \textbf{71,47\%}         & \textbf{78,81\%}     & \textbf{81,87\%}     & \textbf{75,95\%}         & \textbf{80,93\%}     \\
+        {6} & {2}   & {83,25\%}                & {83,23\%}                & {87,74\%}            & {88,95\%}            & {85,38\%}                & {88,87\%}            \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attribute generator: Color}
+    \label{tab:probing:attribute-generator:colour}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {1} & {2}   & \textcolor{red}{95,17\%}          & \textcolor{red}{97,23\%}          & {87,74\%}                & {82,8\%}                 & \textcolor{red}{94,4\%}  & {88,75\%}                \\
+        {1} & {10}  & \textcolor{red}{96,04\%}          & \textcolor{red}{94,67\%}          & \textcolor{red}{94,06\%} & {88,67\%}                & {82,86\%}                & {79,71\%}                \\
+        {1} & {16}  & \textcolor{red}{93,82\%}          & \textcolor{red}{95,48\%}          & {85,33\%}                & {79,96\%}                & \textcolor{red}{90,17\%} & {84,6\%}                 \\
+        {1} & {50}  & \textcolor{red}{94,47\%}          & \textcolor{red}{96,65\%}          & {86,98\%}                & {81,89\%}                & \textcolor{red}{90,93\%} & {86,51\%}                \\
+        {1} & {100} & \textcolor{red}{95,6\%}           & \textcolor{red}{98,19\%}          & \textcolor{red}{90,41\%} & {86,13\%}                & \textcolor{red}{95,74\%} & \textcolor{red}{93,76\%} \\
+        {2} & {2}   & \textcolor{red}{94,75\%}          & \textcolor{red}{96,65\%}          & {86,51\%}                & {80,84\%}                & \textcolor{red}{92,99\%} & {86,3\%}                 \\
+        {2} & {10}  & \textcolor{red}{94,69\%}          & \textcolor{red}{98\%}             & {86,92\%}                & {82,2\%}                 & \textcolor{red}{96,31\%} & \textcolor{red}{92,55\%} \\
+        {2} & {16}  & \textcolor{red}{95,04\%}          & \textcolor{red}{96,39\%}          & {87,82\%}                & {82,54\%}                & \textcolor{red}{90,91\%} & {87,2\%}                 \\
+        {2} & {50}  & \textcolor{red}{97,18\%}          & \textcolor{red}{93,99\%}          & \textcolor{red}{96,42\%} & \textcolor{red}{92,26\%} & \textbf{77,97\%}         & {75,58\%}                \\
+        {2} & {100} & \textcolor{red}{93,71\%}          & \textcolor{red}{94,65\%}          & {85,56\%}                & {78,95\%}                & {83,77\%}                & {78,69\%}                \\
+        {3} & {2}   & \textcolor{red}{95,24\%}          & \textcolor{red}{97,37\%}          & {87,56\%}                & {82,49\%}                & \textcolor{red}{94,38\%} & {88,87\%}                \\
+        {3} & {10}  & \textcolor{red}{96,62\%}          & \textcolor{red}{\textbf{91,61\%}} & \textcolor{red}{94,87\%} & {89\%}                   & \textbf{69,2\%}          & \textbf{65,45\%}         \\
+        {3} & {16}  & \textcolor{red}{93,45\%}          & \textcolor{red}{97,16\%}          & {83,6\%}                 & {77,06\%}                & \textcolor{red}{94,5\%}  & {89\%}                   \\
+        {3} & {50}  & \textcolor{red}{\textbf{90,97\%}} & \textcolor{red}{95,74\%}          & \textbf{74,49\%}         & \textbf{68,89\%}         & \textcolor{red}{92,37\%} & {84,98\%}                \\
+        {3} & {100} & \textcolor{red}{96,74\%}          & \textcolor{red}{\textbf{92,76\%}} & \textcolor{red}{95,51\%} & \textcolor{red}{90,35\%} & \textbf{71,88\%}         & \textbf{69,69\%}         \\
+        {4} & {2}   & \textcolor{red}{94,94\%}          & \textcolor{red}{96,45\%}          & {89,08\%}                & {83,89\%}                & \textcolor{red}{92,29\%} & {85,7\%}                 \\
+        {4} & {10}  & \textcolor{red}{\textbf{91,62\%}} & \textcolor{red}{\textbf{93,29\%}} & {78,53\%}                & {70,85\%}                & {81,99\%}                & \textbf{74,83\%}         \\
+        {4} & {16}  & \textcolor{red}{92,86\%}          & \textcolor{red}{96,45\%}          & {79,49\%}                & {73,49\%}                & \textcolor{red}{94,05\%} & {86,05\%}                \\
+        {4} & {50}  & \textcolor{red}{94\%}             & \textcolor{red}{94,78\%}          & {86,17\%}                & {79,61\%}                & {88,11\%}                & {82,65\%}                \\
+        {4} & {100} & \textcolor{red}{92,15\%}          & \textcolor{red}{96,99\%}          & {78,91\%}                & {72,51\%}                & \textcolor{red}{94,8\%}  & {88,6\%}                 \\
+        {6} & {16}  & \textcolor{red}{\textbf{90,04\%}} & \textcolor{red}{95,83\%}          & \textbf{69,71\%}         & \textbf{61,75\%}         & \textcolor{red}{91,77\%} & {81,6\%}                 \\
+        {6} & {50}  & \textcolor{red}{91,9\%}           & \textcolor{red}{96,28\%}          & \textbf{76,04\%}         & \textbf{69,86\%}         & \textcolor{red}{92,81\%} & {85,2\%}                 \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Dale-2}
+    \label{tab:probing:attention-predictor:dale-2}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh} & \textbf{Sh > C > Si}              & \textbf{Sh > Si > C} & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
+        {1} & {2}   & \textcolor{red}{95,35\%}          & {89,87\%}            & \textcolor{red}{95,09\%}          & {87,23\%}            & {86,01\%}            & {84,14\%}            \\
+        {1} & {16}  & \textcolor{red}{92,95\%}          & {87,03\%}            & \textcolor{red}{92,63\%}          & {83,91\%}            & {82,77\%}            & {80,06\%}            \\
+        {1} & {50}  & \textcolor{red}{93,35\%}          & {87,23\%}            & \textcolor{red}{93,26\%}          & {84,65\%}            & {82,46\%}            & {80,38\%}            \\
+        {2} & {10}  & \textbf{88,94\%}                  & \textbf{84,78\%}     & \textbf{87,5\%}                   & \textbf{77,76\%}     & \textbf{80,24\%}     & \textbf{74,95\%}     \\
+        {2} & {16}  & \textbf{85,71\%}                  & \textbf{84,53\%}     & \textbf{82,98\%}                  & \textbf{72,85\%}     & \textbf{79,72\%}     & \textbf{71,85\%}     \\
+        {3} & {16}  & \textcolor{red}{\textbf{92,18\%}} & \textbf{86,23\%}     & \textcolor{red}{91,93\%}          & \textbf{82,59\%}     & \textbf{81,69\%}     & \textbf{78,72\%}     \\
+        {3} & {50}  & \textcolor{red}{92,68\%}          & {86,53\%}            & \textcolor{red}{91,9\%}           & {83,35\%}            & {82,45\%}            & {79,99\%}            \\
+        {6} & {10}  & \textcolor{red}{93,4\%}           & {86,51\%}            & \textcolor{red}{\textbf{91,64\%}} & {84,03\%}            & {82,51\%}            & {80,06\%}            \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Dale-5}
+    \label{tab:probing:attention-predictor:dale-5}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{cc|c|c|c|c|c|c}
+        \toprule
+        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
+        {3} & {16}  & \textcolor{red}{98,48\%} & \textcolor{red}{98,58\%} & \textcolor{red}{97,87\%} & \textcolor{red}{96,11\%} & \textcolor{red}{96,6\%}  & \textcolor{red}{95,97\%} \\
+        {4} & {10}  & \textcolor{red}{98,88\%} & \textcolor{red}{98,75\%} & \textcolor{red}{97,23\%} & \textcolor{red}{96,67\%} & \textcolor{red}{97,18\%} & \textcolor{red}{96,75\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Attention predictor: Color}
+    \label{tab:probing:attention-predictor:colour}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{c|cc|cc|cc|cc}
+        \toprule
+        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
+        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
+        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
+        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
+        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
+        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
+        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
+    % SD: I have a trouble of understanding and interpreting these results.
+First the loss should be estimated on a test set, after the system has been trained.
+I'm not sure how you estaimate the baseline loss.
+    % DK: TODO, why a test set?
+Isn't the purpose to learn how well the generated captions map to English and not how well they generalize.
+Using only the train set exactly shows if a mapping is possible
+    \label{tab:probing_discrimnator_language}
+\end{table}
+
+Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
+First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
+Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
+This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
+Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
+% SD: I think at this point a readerwill be confused since we are discussing Dale and Reiter bith for generation of scenes and now for the generation of strings.
+We should have clarified the difference earlier.
+Also, rather than referring to it as Dale and Reiter we should call it "the incremental algorithm" and add a reference at first mention.
+% DK: TODO
+
+Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
+The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
+With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
+Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
+The attribute \emph{color} seems to be less important.
+This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
+
 An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
@@
 % DK: TODO
-
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
-Probing can be used to analyze and interpret the hidden representations in neural networks.
-Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
-If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
-If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
-In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
-% SD: The probing test is to show how well the emerged language can be trasnlated into English descriptions consistting of labels of scene features using Dale and Reiter idea of incrementality.
-However, we would need to know now what was the policy of the sender to generate message.
-Was this similar to Dale and Reiter to enforce longer descriptions?
-% DK: TODO
-Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
-The encoder LSTM encodes the messages of the emerged language.
-The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
-The resulting vector is used as the initial state of the decoder LSTM.
-The decoder LSTM is then trained to produce the English referring expression.
-While decoding, teacher forcing is applied.
-The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
-Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.
-
-In the sections above, the attributes are ordered by importance in the way, it is usually used in English: shape > color > size.
-Since the agents do not necessarily need to follow this, all possible orders of importance for the attributes are compared to each other.
-% SD: Move to the earlier paragraph when you introduce Dale and Reiter.
-% DK: TODO
-
-The translation model with this setup can learn two characteristics.
-First, it can learn to find correlations between the emerged language and the English referring expressions.
-Secondly, it can learn patterns in the English referring expressions that are independent of the emerged language.
-For instance, it can learn that the referring expressions are likely to be two symbols long, or that 'cube' is the most common shape.
-This second characteristic can lead to a low loss of the model, even though there are no connections to the emerged language.
-In this test, only the first characteristic is interesting and would show the correlation between the emerged language and English referring expressions.
-For that reason, two reference baselines are calculated for each of the attribute importance order.
-% SD: Baselines?
-% DK: (done)
-The first reference uses a non-informative input for the encoder LSTM, more specifically it always receives a vector of zeros and therefore can't learn any meaningful representation in the linear layer; the input for the decoder LSTM is the same for every sample.
-By this, the model is trained to learn the patterns in the English referring expressions.
-The resulting loss is the highest possible loss the model can achieve, independent of the input.
-Resulting, if the emergent language is connected to English, the loss will be lower than this baseline.
-If not, the loss will be as high as this baseline.
-On the other side, we calculated the lowest loss possible, by using the English referring expressions as both input and target.
-This verifies that the model is able to learn an abstract representation from the input and the resulting loss should be close to zero (since input is perfectly correlated to the target).
-These two references are used, to normalize the results of the actual emergent languages $L_{em}$, using the formula $L_{norm} = \frac{L_{em}-L_{English}}{L_{baseline} - L_{English}}$.
-A loss close as $L_{baseline}$ will lead to 100\%, while a loss as $L_{English}$ will lead to 0\%.
-% SD: Is this encessary? But when and where is this loss using 0 inouts calculated?
-% DK: TODO, QUESTION
-By doing this, all configurations can be compared directly to each other.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    % SD: I have a trouble of understanding and interpreting these results.
-First the loss should be estimated on a test set, after the system has been trained.
-I'm not sure how you estaimate the baseline loss.
-    % DK: TODO, why a test set?
-Isn't the purpose to learn how well the generated captions map to English and not how well they generalize.
-Using only the train set exactly shows if a mapping is possible
-    \label{tab:probing_discrimnator_language}
-\end{table}
-
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-% SD: I think at this point a readerwill be confused since we are discussing Dale and Reiter bith for generation of scenes and now for the generation of strings.
-We should have clarified the difference earlier.
-Also, rather than referring to it as Dale and Reiter we should call it "the incremental algorithm" and add a reference at first mention.
-% DK: TODO
-
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{95,5\%}           & \textcolor{red}{97,57\%}          & {86,98\%}                & {81,43\%}                & \textcolor{red}{91,67\%} & {86,45\%}                \\
-        {1} & {10}  & \textcolor{red}{96,11\%}          & \textcolor{red}{96,9\%}           & {89,61\%}                & {85\%}                   & {89,4\%}                 & {84,49\%}                \\
-        {1} & {16}  & \textcolor{red}{95,19\%}          & \textcolor{red}{96,56\%}          & {87,25\%}                & {81,22\%}                & {89,34\%}                & {83,13\%}                \\
-        {1} & {50}  & \textcolor{red}{96,61\%}          & \textcolor{red}{97,07\%}          & \textcolor{red}{91,26\%} & {85,75\%}                & {87,14\%}                & {82,61\%}                \\
-        {1} & {100} & \textcolor{red}{95,91\%}          & \textcolor{red}{96,73\%}          & \textcolor{red}{90,63\%} & {86,01\%}                & {88,95\%}                & {84,17\%}                \\
-        {2} & {2}   & \textcolor{red}{99,67\%}          & \textcolor{red}{99,91\%}          & \textcolor{red}{97,02\%} & \textcolor{red}{95,37\%} & \textcolor{red}{96,97\%} & \textcolor{red}{94,64\%} \\
-        {2} & {10}  & \textcolor{red}{96,36\%}          & \textcolor{red}{96,34\%}          & \textcolor{red}{91,32\%} & {86,75\%}                & {87,69\%}                & {82,87\%}                \\
-        {2} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{97,12\%}          & {88,11\%}                & {83,21\%}                & {89,86\%}                & {85,35\%}                \\
-        {2} & {50}  & \textcolor{red}{95,34\%}          & \textcolor{red}{97,3\%}           & {85,48\%}                & {80,89\%}                & \textcolor{red}{92,22\%} & {86,18\%}                \\
-        {2} & {100} & \textcolor{red}{94,88\%}          & \textcolor{red}{96,32\%}          & {87,37\%}                & {81,17\%}                & {88,35\%}                & {82,61\%}                \\
-        {3} & {2}   & \textcolor{red}{\textbf{94,03\%}} & \textcolor{red}{96,96\%}          & \textbf{79,59\%}         & \textbf{74,54\%}         & \textcolor{red}{92,32\%} & {85,63\%}                \\
-        {3} & {10}  & \textcolor{red}{96,1\%}           & \textcolor{red}{95,97\%}          & \textcolor{red}{90,45\%} & {84,84\%}                & {85,98\%}                & {80,5\%}                 \\
-        {3} & {16}  & \textcolor{red}{95,71\%}          & \textcolor{red}{\textbf{95,1\%}}  & \textcolor{red}{90,97\%} & {85,33\%}                & {81\%}                   & \textbf{75,56\%}         \\
-        {3} & {50}  & \textcolor{red}{96,32\%}          & \textcolor{red}{\textbf{94,55\%}} & \textcolor{red}{92,5\%}  & {86,81\%}                & \textbf{80,48\%}         & {75,73\%}                \\
-        {3} & {100} & \textcolor{red}{95,33\%}          & \textcolor{red}{97,82\%}          & {86,05\%}                & {81,33\%}                & \textcolor{red}{93,91\%} & {88,54\%}                \\
-        {4} & {2}   & \textcolor{red}{96,81\%}          & \textcolor{red}{95,12\%}          & \textcolor{red}{93,03\%} & {87,41\%}                & \textbf{80,73\%}         & \textbf{75,47\%}         \\
-        {4} & {10}  & \textcolor{red}{95,31\%}          & \textcolor{red}{97,26\%}          & {86,9\%}                 & {82,67\%}                & \textcolor{red}{92,39\%} & {86,98\%}                \\
-        {4} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{\textbf{94,65\%}} & \textcolor{red}{91,99\%} & {86,65\%}                & \textbf{79,23\%}         & \textbf{74,54\%}         \\
-        {4} & {50}  & \textcolor{red}{94,76\%}          & \textcolor{red}{96,69\%}          & {86,8\%}                 & {81,33\%}                & {89,65\%}                & {83,92\%}                \\
-        {4} & {100} & \textcolor{red}{96,16\%}          & \textcolor{red}{97,94\%}          & {89,57\%}                & {85,18\%}                & \textcolor{red}{92,56\%} & {87,74\%}                \\
-        {6} & {2}   & \textcolor{red}{95,73\%}          & \textcolor{red}{96,99\%}          & {87,77\%}                & {83,27\%}                & \textcolor{red}{91,77\%} & {86,34\%}                \\
-        {6} & {10}  & \textcolor{red}{\textbf{94,59\%}} & \textcolor{red}{99,89\%}          & {86,24\%}                & {81,46\%}                & \textcolor{red}{91,42\%} & {86,43\%}                \\
-        {6} & {16}  & \textcolor{red}{95,51\%}          & \textcolor{red}{96,3\%}           & {88,89\%}                & {83,24\%}                & {87,61\%}                & {82,71\%}                \\
-        {6} & {50}  & \textcolor{red}{\textbf{94,66\%}} & \textcolor{red}{97,43\%}          & \textbf{84,93\%}         & \textbf{78,61\%}         & \textcolor{red}{91,48\%} & {83,66\%}                \\
-        {6} & {100} & \textcolor{red}{95,37\%}          & \textcolor{red}{97,77\%}          & \textbf{83,75\%}         & \textbf{80,84\%}         & \textcolor{red}{93,96\%} & {89,02\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{Discriminator: Dale-2}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
-        {1} & {2}   & \textcolor{red}{94,96\%} & \textcolor{red}{91,27\%} & \textcolor{red}{94,51\%} & \textcolor{red}{90,99\%} & {88,52\%}            & {88,17\%}            \\
-        {1} & {10}  & {84,34\%}                & {84,03\%}                & {80,91\%}                & {73,23\%}                & {79,95\%}            & {72,34\%}            \\
-        {1} & {16}  & {86,27\%}                & {84,96\%}                & {82,76\%}                & {75,35\%}                & {80,72\%}            & {74,04\%}            \\
-        {1} & {50}  & {85,95\%}                & {84,4\%}                 & {82,96\%}                & {74,97\%}                & {80,2\%}             & {73,54\%}            \\
-        {1} & {100} & {86,31\%}                & {85,58\%}                & {82,9\%}                 & {76,03\%}                & {81,64\%}            & {74,62\%}            \\
-        {2} & {10}  & \textbf{82,66\%}         & {83,23\%}                & \textbf{78,51\%}         & {71,01\%}                & {78,89\%}            & {70,61\%}            \\
-        {2} & {16}  & {89,78\%}                & {85,53\%}                & {87,68\%}                & {80,64\%}                & {81,06\%}            & {77,24\%}            \\
-        {2} & {50}  & {84,4\%}                 & {82,65\%}                & {80,62\%}                & {71,42\%}                & {77,53\%}            & {69,46\%}            \\
-        {2} & {100} & {89,65\%}                & {85,3\%}                 & {87,26\%}                & {79,91\%}                & {80,88\%}            & {77,1\%}             \\
-        {3} & {2}   & {87,19\%}                & {86,93\%}                & {83,99\%}                & {78,18\%}                & {83,65\%}            & {77,08\%}            \\
-        {3} & {10}  & {87,15\%}                & {84,67\%}                & {84,01\%}                & {76,24\%}                & {80,41\%}            & {74,11\%}            \\
-        {3} & {50}  & \textbf{82,98\%}         & {82,25\%}                & {78,85\%}                & {70,62\%}                & {77,42\%}            & {69,27\%}            \\
-        {3} & {100} & {83,78\%}                & {83,07\%}                & {80,23\%}                & {72,15\%}                & {78,94\%}            & {71,1\%}             \\
-        {4} & {2}   & {89,23\%}                & {87,43\%}                & {87,09\%}                & {81,45\%}                & {83,83\%}            & {79,11\%}            \\
-        {4} & {10}  & {88,66\%}                & {84,83\%}                & {85,9\%}                 & {78,98\%}                & {80,2\%}             & {75,64\%}            \\
-        {4} & {16}  & {83,14\%}                & \textbf{81,51\%}         & \textbf{78,15\%}         & \textbf{70,21\%}         & \textbf{77,32\%}     & \textbf{68,48\%}     \\
-        {4} & {50}  & {87,62\%}                & {84,38\%}                & {85,25\%}                & {77,29\%}                & {79,94\%}            & {74,03\%}            \\
-        {4} & {100} & \textcolor{red}{90,17\%} & {85,76\%}                & {87,68\%}                & {80,71\%}                & {81,37\%}            & {78,1\%}             \\
-        {6} & {10}  & \textbf{82,03\%}         & \textbf{80,51\%}         & \textbf{77,46\%}         & \textbf{68,63\%}         & \textbf{75,71\%}     & \textbf{67,15\%}     \\
-        {6} & {16}  & {84,73\%}                & \textbf{81,12\%}         & {80,99\%}                & \textbf{70,5\%}          & \textbf{76,06\%}     & \textbf{69,12\%}     \\
-        {6} & {50}  & {86,95\%}                & {85,13\%}                & {84,06\%}                & {76,95\%}                & {80,11\%}            & {74,51\%}            \\
-        {6} & {100} & {86,15\%}                & {82,19\%}                & {83,56\%}                & {75,56\%}                & {77,76\%}            & {72,47\%}            \\
-        \bottomrule
-    \end{tabular}
-    \caption{Discriminator: Dale-5}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{91,62\%} & \textcolor{red}{91,57\%} & \textcolor{red}{94,31\%} & \textcolor{red}{95,13\%} & \textcolor{red}{93,11\%} & \textcolor{red}{94,92\%} \\
-        {1} & {10}  & {79,08\%}                & {79,08\%}                & {85,04\%}                & {86,86\%}                & {82,37\%}                & {86,38\%}                \\
-        {1} & {16}  & {84,02\%}                & {83,96\%}                & {88,67\%}                & {89,65\%}                & {86,36\%}                & {89,36\%}                \\
-        {1} & {50}  & \textcolor{red}{93,88\%} & \textcolor{red}{93,74\%} & \textcolor{red}{95,03\%} & \textcolor{red}{95,49\%} & \textcolor{red}{94,64\%} & \textcolor{red}{95,35\%} \\
-        {2} & {10}  & {88,67\%}                & {88,36\%}                & {89,59\%}                & \textcolor{red}{90,77\%} & \textcolor{red}{90,01\%} & {89,81\%}                \\
-        {2} & {16}  & {81,13\%}                & {81,02\%}                & {86,16\%}                & {87,97\%}                & {83,85\%}                & {87,56\%}                \\
-        {3} & {2}   & {83,38\%}                & {83,46\%}                & {88,21\%}                & {89,71\%}                & {86,05\%}                & {89,37\%}                \\
-        {3} & {10}  & \textbf{77\%}            & \textbf{76,77\%}         & \textbf{83,54\%}         & \textbf{85,57\%}         & \textbf{79,97\%}         & \textbf{84,9\%}          \\
-        {3} & {16}  & \textcolor{red}{93,52\%} & \textcolor{red}{93,6\%}  & \textcolor{red}{94,16\%} & \textcolor{red}{94,14\%} & \textcolor{red}{93,75\%} & \textcolor{red}{93,55\%} \\
-        {4} & {10}  & \textbf{68,04\%}         & \textbf{66,51\%}         & \textbf{76,33\%}         & \textbf{78,56\%}         & \textbf{71,42\%}         & \textbf{78,56\%}         \\
-        {4} & {16}  & \textbf{75,1\%}          & \textbf{74,93\%}         & \textbf{81,2\%}          & \textbf{83,35\%}         & \textbf{77,6\%}          & \textbf{82,76\%}         \\
-        {6} & {2}   & {84,21\%}                & {84,02\%}                & {88,41\%}                & {89,77\%}                & {86,61\%}                & {89,35\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{Discriminator: Color}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{95,17\%}          & \textcolor{red}{97,23\%}          & {87,74\%}                & {82,8\%}                 & \textcolor{red}{94,4\%}  & {88,75\%}                \\
-        {1} & {10}  & \textcolor{red}{96,04\%}          & \textcolor{red}{94,67\%}          & \textcolor{red}{94,06\%} & {88,67\%}                & {82,86\%}                & {79,71\%}                \\
-        {1} & {16}  & \textcolor{red}{93,82\%}          & \textcolor{red}{95,48\%}          & {85,33\%}                & {79,96\%}                & \textcolor{red}{90,17\%} & {84,6\%}                 \\
-        {1} & {50}  & \textcolor{red}{94,47\%}          & \textcolor{red}{96,65\%}          & {86,98\%}                & {81,89\%}                & \textcolor{red}{90,93\%} & {86,51\%}                \\
-        {1} & {100} & \textcolor{red}{95,6\%}           & \textcolor{red}{98,19\%}          & \textcolor{red}{90,41\%} & {86,13\%}                & \textcolor{red}{95,74\%} & \textcolor{red}{93,76\%} \\
-        {2} & {2}   & \textcolor{red}{94,75\%}          & \textcolor{red}{96,65\%}          & {86,51\%}                & {80,84\%}                & \textcolor{red}{92,99\%} & {86,3\%}                 \\
-        {2} & {10}  & \textcolor{red}{94,69\%}          & \textcolor{red}{98\%}             & {86,92\%}                & {82,2\%}                 & \textcolor{red}{96,31\%} & \textcolor{red}{92,55\%} \\
-        {2} & {16}  & \textcolor{red}{95,04\%}          & \textcolor{red}{96,39\%}          & {87,82\%}                & {82,54\%}                & \textcolor{red}{90,91\%} & {87,2\%}                 \\
-        {2} & {50}  & \textcolor{red}{97,18\%}          & \textcolor{red}{93,99\%}          & \textcolor{red}{96,42\%} & \textcolor{red}{92,26\%} & \textbf{77,97\%}         & {75,58\%}                \\
-        {2} & {100} & \textcolor{red}{93,71\%}          & \textcolor{red}{94,65\%}          & {85,56\%}                & {78,95\%}                & {83,77\%}                & {78,69\%}                \\
-        {3} & {2}   & \textcolor{red}{95,24\%}          & \textcolor{red}{97,37\%}          & {87,56\%}                & {82,49\%}                & \textcolor{red}{94,38\%} & {88,87\%}                \\
-        {3} & {10}  & \textcolor{red}{96,62\%}          & \textcolor{red}{\textbf{91,61\%}} & \textcolor{red}{94,87\%} & {89\%}                   & \textbf{69,2\%}          & \textbf{65,45\%}         \\
-        {3} & {16}  & \textcolor{red}{93,45\%}          & \textcolor{red}{97,16\%}          & {83,6\%}                 & {77,06\%}                & \textcolor{red}{94,5\%}  & {89\%}                   \\
-        {3} & {50}  & \textcolor{red}{\textbf{90,97\%}} & \textcolor{red}{95,74\%}          & \textbf{74,49\%}         & \textbf{68,89\%}         & \textcolor{red}{92,37\%} & {84,98\%}                \\
-        {3} & {100} & \textcolor{red}{96,74\%}          & \textcolor{red}{\textbf{92,76\%}} & \textcolor{red}{95,51\%} & \textcolor{red}{90,35\%} & \textbf{71,88\%}         & \textbf{69,69\%}         \\
-        {4} & {2}   & \textcolor{red}{94,94\%}          & \textcolor{red}{96,45\%}          & {89,08\%}                & {83,89\%}                & \textcolor{red}{92,29\%} & {85,7\%}                 \\
-        {4} & {10}  & \textcolor{red}{\textbf{91,62\%}} & \textcolor{red}{\textbf{93,29\%}} & {78,53\%}                & {70,85\%}                & {81,99\%}                & \textbf{74,83\%}         \\
-        {4} & {16}  & \textcolor{red}{92,86\%}          & \textcolor{red}{96,45\%}          & {79,49\%}                & {73,49\%}                & \textcolor{red}{94,05\%} & {86,05\%}                \\
-        {4} & {50}  & \textcolor{red}{94\%}             & \textcolor{red}{94,78\%}          & {86,17\%}                & {79,61\%}                & {88,11\%}                & {82,65\%}                \\
-        {4} & {100} & \textcolor{red}{92,15\%}          & \textcolor{red}{96,99\%}          & {78,91\%}                & {72,51\%}                & \textcolor{red}{94,8\%}  & {88,6\%}                 \\
-        {6} & {16}  & \textcolor{red}{\textbf{90,04\%}} & \textcolor{red}{95,83\%}          & \textbf{69,71\%}         & \textbf{61,75\%}         & \textcolor{red}{91,77\%} & {81,6\%}                 \\
-        {6} & {50}  & \textcolor{red}{91,9\%}           & \textcolor{red}{96,28\%}          & \textbf{76,04\%}         & \textbf{69,86\%}         & \textcolor{red}{92,81\%} & {85,2\%}                 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attention predictor: Dale-2}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh} & \textbf{Sh > C > Si}              & \textbf{Sh > Si > C} & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
-        {1} & {2}   & \textcolor{red}{95,35\%}          & {89,87\%}            & \textcolor{red}{95,09\%}          & {87,23\%}            & {86,01\%}            & {84,14\%}            \\
-        {1} & {16}  & \textcolor{red}{92,95\%}          & {87,03\%}            & \textcolor{red}{92,63\%}          & {83,91\%}            & {82,77\%}            & {80,06\%}            \\
-        {1} & {50}  & \textcolor{red}{93,35\%}          & {87,23\%}            & \textcolor{red}{93,26\%}          & {84,65\%}            & {82,46\%}            & {80,38\%}            \\
-        {2} & {10}  & \textbf{88,94\%}                  & \textbf{84,78\%}     & \textbf{87,5\%}                   & \textbf{77,76\%}     & \textbf{80,24\%}     & \textbf{74,95\%}     \\
-        {2} & {16}  & \textbf{85,71\%}                  & \textbf{84,53\%}     & \textbf{82,98\%}                  & \textbf{72,85\%}     & \textbf{79,72\%}     & \textbf{71,85\%}     \\
-        {3} & {16}  & \textcolor{red}{\textbf{92,18\%}} & \textbf{86,23\%}     & \textcolor{red}{91,93\%}          & \textbf{82,59\%}     & \textbf{81,69\%}     & \textbf{78,72\%}     \\
-        {3} & {50}  & \textcolor{red}{92,68\%}          & {86,53\%}            & \textcolor{red}{91,9\%}           & {83,35\%}            & {82,45\%}            & {79,99\%}            \\
-        {6} & {10}  & \textcolor{red}{93,4\%}           & {86,51\%}            & \textcolor{red}{\textbf{91,64\%}} & {84,03\%}            & {82,51\%}            & {80,06\%}            \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attention predictor: Dale-5}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {3} & {16}  & \textcolor{red}{98,48\%} & \textcolor{red}{98,58\%} & \textcolor{red}{97,87\%} & \textcolor{red}{96,11\%} & \textcolor{red}{96,6\%}  & \textcolor{red}{95,97\%} \\
-        {4} & {10}  & \textcolor{red}{98,88\%} & \textcolor{red}{98,75\%} & \textcolor{red}{97,23\%} & \textcolor{red}{96,67\%} & \textcolor{red}{97,18\%} & \textcolor{red}{96,75\%} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attention predictor: Color}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}      & \textbf{Si > Sh > C}      \\\midrule
-        {1} & {2}   & \textcolor{red}{95,73\%} & \textcolor{red}{94,23\%} & \textcolor{red}{92,96\%} & \textcolor{red}{90,59\%} & {81,59\%}                 & {78,21\%}                 \\
-        {1} & {10}  & \textcolor{red}{91,77\%} & \textcolor{red}{91,19\%} & {73,14\%}                & {63,99\%}                & {67,45\%}                 & {57,55\%}                 \\
-        {1} & {16}  & \textcolor{red}{91,74\%} & \textcolor{red}{90,87\%} & {73,07\%}                & {63,78\%}                & {66,16\%}                 & {55,48\%}                 \\
-        {1} & {50}  & \textcolor{red}{97,78\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,22\%} & {87,45\%}                & {69,23\%}                 & {64,97\%}                 \\
-        {1} & {100} & \textcolor{red}{97,91\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,18\%} & {87,35\%}                & {69,25\%}                 & {64,99\%}                 \\
-        {2} & {2}   & \textcolor{red}{93,17\%} & \textcolor{red}{91,69\%} & {81,96\%}                & {75,97\%}                & {72,05\%}                 & {64,29\%}                 \\
-        {2} & {10}  & \textcolor{red}{94,07\%} & \textcolor{red}{98,04\%} & {75,32\%}                & {71,91\%}                & \textcolor{red}{94,38\%}  & {86,62\%}                 \\
-        {2} & {16}  & \textcolor{red}{97,72\%} & \textcolor{red}{92,9\%}  & \textcolor{red}{94,18\%} & {87,09\%}                & {68,54\%}                 & {64,13\%}                 \\
-        {2} & {50}  & \textcolor{red}{97,87\%} & \textcolor{red}{93,08\%} & \textcolor{red}{94,25\%} & {87,54\%}                & {69,2\%}                  & {65,17\%}                 \\
-        {2} & {100} & {89,6\%}                 & \textbf{89,58\%}         & \textbf{65,65\%}         & \textbf{55,69\%}         & {64,92\%}                 & {53,53\%}                 \\
-        {3} & {2}   & \textcolor{red}{97,73\%} & \textcolor{red}{99,28\%} & \textcolor{red}{95,84\%} & \textcolor{red}{97,72\%} & \textcolor{red}{101,58\%} & \textcolor{red}{100,22\%} \\
-        {3} & {10}  & \textcolor{red}{91,04\%} & \textcolor{red}{90,52\%} & {71,06\%}                & {61,36\%}                & {65,59\%}                 & {54,93\%}                 \\
-        {3} & {16}  & \textcolor{red}{91,09\%} & \textcolor{red}{90,3\%}  & {71,71\%}                & {62,42\%}                & {66,28\%}                 & {56,16\%}                 \\
-        {3} & {50}  & \textcolor{red}{97,07\%} & \textcolor{red}{91,72\%} & \textcolor{red}{92,7\%}  & {85,33\%}                & {66,38\%}                 & {61,26\%}                 \\
-        {3} & {100} & \textcolor{red}{97,81\%} & \textcolor{red}{93,01\%} & \textcolor{red}{94,16\%} & {87,3\%}                 & {68,83\%}                 & {64,6\%}                  \\
-        {4} & {2}   & \textcolor{red}{91,95\%} & \textcolor{red}{96,87\%} & {73,55\%}                & {72,66\%}                & \textcolor{red}{98,37\%}  & \textcolor{red}{90,23\%}  \\
-        {4} & {10}  & \textcolor{red}{97,7\%}  & \textcolor{red}{93,1\%}  & \textcolor{red}{94,16\%} & {87,24\%}                & {69,22\%}                 & {64,81\%}                 \\
-        {4} & {16}  & \textbf{86,75\%}         & \textbf{89\%}            & \textbf{57,02\%}         & \textbf{44,33\%}         & \textbf{63,78\%}          & \textbf{48,21\%}          \\
-        {4} & {50}  & \textcolor{red}{90,86\%} & \textcolor{red}{90,5\%}  & {71,22\%}                & {61,09\%}                & {65,26\%}                 & {54,87\%}                 \\
-        {4} & {100} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,22\%} & \textcolor{red}{92,44\%} & {83,78\%}                & \textbf{59,98\%}          & {54,38\%}                 \\
-        {6} & {2}   & \textcolor{red}{90,06\%} & \textcolor{red}{90,46\%} & {71,62\%}                & {64,56\%}                & {70,67\%}                 & {61,35\%}                 \\
-        {6} & {10}  & \textbf{88,38\%}         & \textcolor{red}{90,24\%} & \textbf{63,68\%}         & \textbf{53,21\%}         & \textbf{64,22\%}          & \textbf{52,19\%}          \\
-        {6} & {16}  & \textbf{89,01\%}         & \textbf{89,74\%}         & {67,03\%}                & {56,73\%}                & {64,44\%}                 & \textbf{52,78\%}          \\
-        {6} & {50}  & \textcolor{red}{97,68\%} & \textcolor{red}{92,77\%} & \textcolor{red}{97,03\%} & {87,24\%}                & {68,51\%}                 & {64,27\%}                 \\
-        {6} & {100} & {89,84\%}                & {89,81\%}                & {69,75\%}                & {60,55\%}                & {65,19\%}                 & {54,93\%}                 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attribute generator: Dale-2}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{91,81\%} & \textcolor{red}{96,22\%} & {89,08\%}                & {88,2\%}                 & \textcolor{red}{95,91\%} & \textcolor{red}{90,3\%}  \\
-        {1} & {10}  & {83,4\%}                 & {82,11\%}                & {79,68\%}                & {67,88\%}                & {77,12\%}                & {66,75\%}                \\
-        {1} & {16}  & {86,76\%}                & {87,06\%}                & {83,08\%}                & {75,03\%}                & {83,79\%}                & {74,72\%}                \\
-        {1} & {50}  & {83,12\%}                & {82\%}                   & {78,81\%}                & {66,76\%}                & {76,95\%}                & {65,95\%}                \\
-        {1} & {100} & \textcolor{red}{93,84\%} & {87,09\%}                & \textcolor{red}{93,01\%} & {83,36\%}                & {82,19\%}                & {79,14\%}                \\
-        {2} & {2}   & {88,86\%}                & {86,82\%}                & {86,62\%}                & {78,45\%}                & {82,59\%}                & {76,49\%}                \\
-        {2} & {10}  & {83,91\%}                & {81,57\%}                & {79,76\%}                & {67,89\%}                & {76,47\%}                & {66,1\%}                 \\
-        {2} & {16}  & \textbf{75,6\%}          & \textbf{77,74\%}         & \textbf{68,78\%}         & \textbf{55,43\%}         & \textbf{72,24\%}         & \textbf{56,24\%}         \\
-        {2} & {50}  & {85,1\%}                 & {83,08\%}                & {81,27\%}                & {70,13\%}                & {77,86\%}                & {68,54\%}                \\
-        {2} & {100} & {83,24\%}                & {81,87\%}                & {79,02\%}                & {67,32\%}                & {77,06\%}                & {66,19\%}                \\
-        {3} & {2}   & \textcolor{red}{97,21\%} & \textcolor{red}{96,08\%} & \textcolor{red}{96,43\%} & \textcolor{red}{95,23\%} & \textcolor{red}{95,18\%} & \textcolor{red}{94,24\%} \\
-        {3} & {10}  & {82,66\%}                & {81,11\%}                & {78,38\%}                & {66,21\%}                & {75,68\%}                & {64,52\%}                \\
-        {3} & {16}  & \textcolor{red}{93,26\%} & {86,09\%}                & \textcolor{red}{92,67\%} & {82,25\%}                & {81,17\%}                & {77,84\%}                \\
-        {3} & {50}  & {82,83\%}                & {81,44\%}                & {78,57\%}                & {66,29\%}                & {76,04\%}                & {64,84\%}                \\
-        {3} & {100} & {82,74\%}                & {81\%}                   & {78,39\%}                & {66,04\%}                & {75,44\%}                & {64,91\%}                \\
-        {4} & {2}   & \textcolor{red}{92,91\%} & {87,03\%}                & \textcolor{red}{92,16\%} & {84,06\%}                & {82,45\%}                & {79,57\%}                \\
-        {4} & {10}  & \textbf{79,3\%}          & \textbf{79,42\%}         & \textbf{74,65\%}         & \textbf{61,71\%}         & \textbf{74,12\%}         & \textbf{60,62\%}         \\
-        {4} & {16}  & {82,55\%}                & {81,31\%}                & {78,61\%}                & {66,51\%}                & {75,61\%}                & {65,05\%}                \\
-        {4} & {50}  & {88,51\%}                & {84,17\%}                & {85,77\%}                & {75,07\%}                & {79\%}                   & {71,93\%}                \\
-        {4} & {100} & {82,57\%}                & {81,61\%}                & {78,06\%}                & {66,33\%}                & {76,32\%}                & {65,54\%}                \\
-        {6} & {2}   & \textcolor{red}{92,36\%} & {88,15\%}                & \textcolor{red}{91,49\%} & {84,38\%}                & {85,28\%}                & {80,85\%}                \\
-        {6} & {10}  & \textbf{78,11\%}         & \textbf{79,19\%}         & \textbf{72,81\%}         & \textbf{59,86\%}         & \textbf{73,95\%}         & \textbf{60,17\%}         \\
-        {6} & {16}  & {82,35\%}                & {81,09\%}                & {78,84\%}                & {66,02\%}                & {75,9\%}                 & {64,79\%}                \\
-        {6} & {50}  & {82,04\%}                & {81,49\%}                & {78,34\%}                & {66,25\%}                & {76,46\%}                & {64,91\%}                \\
-        {6} & {100} & {82,09\%}                & {80,73\%}                & {77,42\%}                & {65,3\%}                 & {75,52\%}                & {64,52\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attribute generator: Dale-5}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si} & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C} \\\midrule
-        {1} & {2}   & \textcolor{red}{99,92\%} & \textcolor{red}{99,88\%} & \textbf{79,83\%}     & \textbf{82,7\%}      & \textcolor{red}{99,71\%} & \textbf{82,14\%}     \\
-        {1} & {10}  & \textbf{70,52\%}         & \textbf{70,36\%}         & \textbf{80,23\%}     & \textbf{82,36\%}     & \textbf{74,49\%}         & \textbf{81,77\%}     \\
-        {1} & {16}  & \textbf{77,25\%}         & \textbf{77,16\%}         & {84,55\%}            & {86,2\%}             & \textbf{80,72\%}         & {85,56\%}            \\
-        {3} & {10}  & \textbf{71,46\%}         & \textbf{71,47\%}         & \textbf{78,81\%}     & \textbf{81,87\%}     & \textbf{75,95\%}         & \textbf{80,93\%}     \\
-        {6} & {2}   & {83,25\%}                & {83,23\%}                & {87,74\%}            & {88,95\%}            & {85,38\%}                & {88,87\%}            \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attribute generator: Color}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{95,07\%} & \textcolor{red}{98,06\%}          & {86,63\%}                & {85,3\%}                 & \textcolor{red}{96,26\%} & \textcolor{red}{93,32\%} \\
-        {1} & {10}  & \textcolor{red}{90,8\%}  & \textcolor{red}{96,4\%}           & {71,88\%}                & {69,29\%}                & \textcolor{red}{94,05\%} & {85,13\%}                \\
-        {1} & {16}  & \textcolor{red}{91,47\%} & \textcolor{red}{96,96\%}          & {71,2\%}                 & {66,82\%}                & \textcolor{red}{93,83\%} & {83,35\%}                \\
-        {1} & {50}  & \textcolor{red}{93,26\%} & \textcolor{red}{96,63\%}          & {77,53\%}                & {71,94\%}                & \textcolor{red}{93,56\%} & {84,62\%}                \\
-        {1} & {100} & \textcolor{red}{93,5\%}  & \textcolor{red}{97,86\%}          & {77,42\%}                & {71,93\%}                & \textcolor{red}{94,68\%} & {85,77\%}                \\
-        {2} & {2}   & \textcolor{red}{93,57\%} & \textcolor{red}{97,72\%}          & {80,89\%}                & {79,8\%}                 & \textcolor{red}{96,08\%} & \textcolor{red}{90,71\%} \\
-        {2} & {10}  & \textbf{84,86\%}         & \textcolor{red}{\textbf{92,07\%}} & {58,8\%}                 & \textbf{52,09\%}         & \textbf{88,36\%}         & {77,01\%}                \\
-        {2} & {16}  & {87,29\%}                & \textcolor{red}{92,88\%}          & {68,52\%}                & {71,5\%}                 & \textcolor{red}{92,56\%} & {83,9\%}                 \\
-        {2} & {50}  & {89,15\%}                & \textcolor{red}{95\%}             & {68,19\%}                & {68,93\%}                & \textcolor{red}{93,11\%} & {83,95\%}                \\
-        {2} & {100} & \textcolor{red}{91,06\%} & \textcolor{red}{96,14\%}          & {71,98\%}                & {67,23\%}                & \textcolor{red}{92,81\%} & {82,71\%}                \\
-        {3} & {2}   & \textcolor{red}{93,05\%} & \textcolor{red}{97,15\%}          & {82,41\%}                & {77,36\%}                & \textcolor{red}{94,43\%} & {88,8\%}                 \\
-        {3} & {10}  & {88,52\%}                & \textcolor{red}{95,41\%}          & {66,61\%}                & {63,63\%}                & \textcolor{red}{92,25\%} & {81,46\%}                \\
-        {3} & {16}  & {85,71\%}                & \textcolor{red}{92,88\%}          & \textbf{56,44\%}         & \textbf{49,22\%}         & \textbf{88,87\%}         & \textbf{74,54\%}         \\
-        {3} & {50}  & \textcolor{red}{90,78\%} & \textcolor{red}{96,28\%}          & {68,17\%}                & {64,84\%}                & \textcolor{red}{94,16\%} & {83,5\%}                 \\
-        {3} & {100} & {86,54\%}                & \textcolor{red}{93,54\%}          & {65,9\%}                 & {66,03\%}                & \textcolor{red}{92,49\%} & {83,28\%}                \\
-        {4} & {10}  & {87,61\%}                & \textcolor{red}{94,63\%}          & {66,83\%}                & {64,85\%}                & \textcolor{red}{92,24\%} & {82,85\%}                \\
-        {4} & {16}  & \textbf{85,7\%}          & \textcolor{red}{93,56\%}          & \textbf{56,9\%}          & \textbf{52,1\%}          & \textbf{88,98\%}         & \textbf{76,15\%}         \\
-        {4} & {50}  & {87,55\%}                & \textcolor{red}{94,24\%}          & {61,03\%}                & {55,82\%}                & \textcolor{red}{91,73\%} & \textbf{76,52\%}         \\
-        {4} & {100} & {88,84\%}                & \textcolor{red}{94,84\%}          & {68,68\%}                & {66,73\%}                & \textcolor{red}{92,57\%} & {81,79\%}                \\
-        {6} & {2}   & \textcolor{red}{93,48\%} & \textcolor{red}{98,14\%}          & {81,36\%}                & {79,08\%}                & \textcolor{red}{96,09\%} & \textcolor{red}{90,95\%} \\
-        {6} & {10}  & \textbf{85\%}            & \textcolor{red}{\textbf{92,44\%}} & \textbf{57,21\%}         & {55,07\%}                & {89,39\%}                & {77,44\%}                \\
-        {6} & {16}  & {88,02\%}                & \textcolor{red}{94,22\%}          & {71,42\%}                & {66,35\%}                & \textcolor{red}{90,86\%} & {84,12\%}                \\
-        {6} & {50}  & \textcolor{red}{98,57\%} & \textcolor{red}{99,45\%}          & \textcolor{red}{95,57\%} & \textcolor{red}{95,06\%} & \textcolor{red}{99,16\%} & \textcolor{red}{96,64\%} \\
-        {6} & {100} & {86,44\%}                & \textcolor{red}{\textbf{92,76\%}} & {65,59\%}                & {65,93\%}                & \textcolor{red}{92,42\%} & {82,16\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{RE generator: Dale-2}
-    \label{TODO: label}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{92,86\%} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,06\%} & {89,5\%}             & \textcolor{red}{95,73\%} & \textcolor{red}{90,93\%} \\
-        {1} & {10}  & {83,9\%}                 & \textcolor{red}{90,9\%}  & {78,63\%}                & {74,74\%}            & {89,6\%}                 & {78,99\%}                \\
-        {1} & {16}  & {84,68\%}                & \textcolor{red}{91,65\%} & {80,11\%}                & {77,03\%}            & \textcolor{red}{90,16\%} & {81,05\%}                \\
-        {1} & {50}  & {84,85\%}                & \textcolor{red}{91,67\%} & {79,95\%}                & {77,11\%}            & \textcolor{red}{90,32\%} & {81,04\%}                \\
-        {2} & {2}   & \textcolor{red}{90,51\%} & \textcolor{red}{95,36\%} & {87,5\%}                 & {86,43\%}            & \textcolor{red}{94,77\%} & {88,35\%}                \\
-        {2} & {10}  & {82,44\%}                & \textcolor{red}{90,32\%} & {77,65\%}                & {73,73\%}            & {88,89\%}                & {78,41\%}                \\
-        {2} & {16}  & {81,85\%}                & \textbf{89,4\%}          & {76,07\%}                & \textbf{71,66\%}     & {87,9\%}                 & \textbf{75,91\%}         \\
-        {2} & {50}  & {83,58\%}                & \textcolor{red}{91,12\%} & {78,67\%}                & {75,03\%}            & {89,94\%}                & {79,06\%}                \\
-        {3} & {2}   & \textcolor{red}{92,51\%} & \textcolor{red}{95,96\%} & {89,96\%}                & {88,7\%}             & \textcolor{red}{95,18\%} & \textcolor{red}{90,43\%} \\
-        {3} & {10}  & {82,11\%}                & {89,54\%}                & {75,9\%}                 & {71,78\%}            & {88,18\%}                & {76,72\%}                \\
-        {3} & {50}  & {83,12\%}                & \textcolor{red}{90,82\%} & {77,62\%}                & {74,08\%}            & {88,88\%}                & {78,74\%}                \\
-        {3} & {100} & {83,74\%}                & {89,59\%}                & {79,29\%}                & {74,48\%}            & \textbf{87,46\%}         & {78,05\%}                \\
-        {4} & {2}   & {88,86\%}                & \textcolor{red}{94\%}    & {84,8\%}                 & {83,19\%}            & \textcolor{red}{93,37\%} & {85,92\%}                \\
-        {4} & {10}  & \textbf{79,93\%}         & \textbf{89,29\%}         & \textbf{74,1\%}          & \textbf{69,5\%}      & \textbf{87,19\%}         & \textbf{74,79\%}         \\
-        {4} & {16}  & \textbf{81,26\%}         & {89,68\%}                & \textbf{75,73\%}         & {71,96\%}            & {88,4\%}                 & {76,93\%}                \\
-        {4} & {50}  & {81,55\%}                & {89,42\%}                & {75,9\%}                 & {71,68\%}            & {88,17\%}                & {76,55\%}                \\
-        {6} & {2}   & \textcolor{red}{91,09\%} & \textcolor{red}{95,26\%} & {87,86\%}                & {86,81\%}            & \textcolor{red}{94,78\%} & {88,82\%}                \\
-        {6} & {50}  & \textbf{80,39\%}         & \textbf{87,85\%}         & \textbf{75,01\%}         & \textbf{68,71\%}     & \textbf{84,54\%}         & \textbf{72,93\%}         \\
-        \bottomrule
-    \end{tabular}
-    \caption{RE generator: Dale-5}
-    \label{TODO: label}
-\end{table}

### 2024-05-07 thesis/sections/7-language-analysis.tex
language analysis

@@
 
+In the following paragraphs, languages will be referred to by their message length $n$ and their vocabulary size $|V|$ as $Lang_{n,|V|}$.
+
 \begin{table}[ht]
@@
     \end{tabular}
-    \caption{Discriminator: Dale-2}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{discrimination games} on the 'Dale-2' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:discriminator:dale-2}
 \end{table}
+
+Looking at the results for the discriminator, one can see that the results are very different for each dataset.
+On the 'Dale-2' dataset $L_{norm}$ is relatively high for all salience orders (see Table \ref{tab:probing:discriminator:dale-2}).
+Yet, the correlation to English referring expressions with the \emph{color} as most important attribute is the lowest.
+When the \emph{color} attribute is in the second position, the correlation becomes higher, but the highest correlation is achieved when the \emph{color} has the least importance.
+This shows that the agents don't use the \emph{color} in the communication, but rather rely on \emph{shape} and \emph{size}.
+The order of importance of these attributes however is not as clear.
+For some combinations of the message length $n$ and vocabulary size $|V|$, the \emph{size} seems to be the dominant attribute, while for others the \emph{shape} seems more important in the communication.
+Interestingly, one can see that for many languages only one salience order has a low $L_{norm}$, while the $L_{norm}$ for the remaining salience orders with a different attribute in the first position is higher.
+This applies for example to $Lang_{3,2}$, where the loss for 'Sh > Si > C' is 74,54\% while the second lowest is only at 85,63\% for 'Si > Sh > C'.
+The same can be seen for instance for $Lang_{3,16}$ in the other direction.
+This difference indicates that the agents are relying mostly on the most important attribute of this salience order to discriminate the target object.
+If the $L_{norm}$ for instance were similar for the salience orders 'Sh > Si > C' and 'Si > Sh > C', this would mean that the emerged language could be connected to both salience orders in the same way.
+Following, the agents would use the \emph{shape} as often as \emph{size} to discriminate the objects.
+This is however only the case on the 'Dale-2' dataset when the loss is higher (for instance languages $Lang_{2,16}$, $Lang_{2,100}$ or $Lang_{6,16}$).
+Since the lowest $L_{norm}$ with 74,54\% is still very high compared to a complete correlation (0\%), the agents seem to encode additionally other information in their messages.
+
+Looking at the influence of $n$ and $|V|$, one can see that across almost all values, there is at least some correlation to one salience order.
+The correlation seems to be highest with $n \in \{3,4\}$, where the loss can reach down 74,54\%.
+Hereby, the most important attribute in the messages seems to be the \emph{size}.
+For the remaining message lengths $n$, the loss stays mostly between 80\% and 85\%.
+For $n \in \{2,6\}$ the \emph{shape} seems to be the most dominant attribute, while for $n=1$ the \emph{size} and \emph{shape} seem to have a similar importance.
+The influence of the vocabulary size $|V|$ is not as clear.
+Small to medium-sized vocabularies seem to promote higher correlations with English referring expressions, but also languages with $|V|=100$ can reach losses of around 80\%.
 
@@
     \end{tabular}
-    \caption{Discriminator: Dale-5}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{discrimination games} on the 'Dale-5' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:discriminator:dale-5}
 \end{table}
+
+The emerged languages on the 'Dale-5' dataset have generally a higher correlation to English referring expressions (see Table \ref{tab:probing:discriminator:dale-5}).
+For almost all salience orders and emerged languages, the loss is lower than 90\%.
+In the best cases, it even reaches down to 67,15\%.
+Again, the \emph{size} and the \emph{shape} are mostly communicated between the agents, but the \emph{color} now plays a bigger role in the messages.
+The agents seem to need to rely on more attributes to discriminate the target object.
+This aligns also with English referring expressions, where the average length of the referring expressions across the whole 'Dale-5' dataset is longer than for the 'Dale-2' dataset.
+However, opposed to the 'Dale-2' dataset, there are almost no languages, where there is one dominant attribute.
+More often the loss is similar for two salience orders with different primary attributes.
+For instance the salience orders 'Sh > Si > C' and 'Si > Sh > C' both provide losses around 70\% for $Lang_{6,16}$.
+Only the languages $Lang_{2,16}$, $Lang_{4,10}$, $Lang_{4,50}$ and $Lang_{6,100}$ have differences of more than 3\% points for these salience orders.
+This indicates, that the agents make use of both \emph{shape} and \emph{size} similarly often, while the \emph{color} is used for much fewer images.
+
+When the agents are able to produce longer messages, their languages seem to have more similarities with possible English languages.
+With $n>1$ the agents produce languages that can have losses lower than 70\% for the salience order 'Si > Sh > C'.
+Yet, languages with $n=1$ are not far off and have mostly losses between 72\% and 75\%.
+Looking at the vocabulary size, languages with $|V|=2$ are less likely to emerge and if they emerge they have a lower correlation with English.
+The best results are produces with $|V| \in \{10,16\}$ for longer messages (for instance $Lang_{4,16}$, $Lang_{6,10}$ and $Lang_{6,16}$), while shorter messages fare generally better with a larger vocabulary of $|V| \in \{50,100\}$ ($Lang_{2,50}$, $Lang_{3,50}$ and $Lang_{3,100}$).
 
@@
     \end{tabular}
-    \caption{Discriminator: Color}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{discrimination games} on the 'CLEVR color' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:discriminator:colour}
@@
 
-Looking at the results for the discriminator, one can see that the results are very different for each dataset.
-On the 'Dale-2' dataset
+Finally on the 'CLEVR color' dataset, much fewer languages emerge (see Table \ref{tab:probing:discriminator:colour}).
+Those languages that emerge have very different levels of correlation to English
+While three languages stay above 90\% for all salience orders, most of them stay between 80\% and 90\%.
+Four languages even get below 80\% for the salience orders where the \emph{color} is the most important attribute.
+One of those even reaches 66,51\%, the highest correlation across all emerged languages on the discrimination task.
+Unsurprisingly, the \emph{color} is now the most important attribute.
+The loss with the remaining salience orders is still relatively low, but this is likely due to the fact, that the color occurs in any referring expression as it is the only discriminating attribute.
+The probing model can therefore easily connect this pattern to the emerged language.
+As expected, the losses for 'Sh > Si > C' and 'Si > Sh > C' are almost the same, which indicates that the agents make (no) use of both \emph{shape} and \emph{size} in the same way.
+
+Even though only one attribute would need to be communicated, languages with $n \in \{3,4\}$ tend to be closer to English referring expressions.
+Hereby, vocabulary sizes of $|V| \in \{10,16\}$ produce the most correlating languages, while languages with only 2 symbols have relatively high losses.
+The agents were not able to communicate with vocabulary sizes larger than 16.
 
@@
     \end{tabular}
-    \caption{RE generator: Dale-2}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{referring expression generation games} on the 'Dale-2' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:re-generator:dale-2}
 \end{table}
+
+Tables \ref{tab:probing:re-generator:dale-2} and \ref{tab:probing:re-generator:dale-5} show the results of the probing on the emerged languages of the \emph{referring expression generation games}.
+This is the only task, where the final task for the agents is to generate English referring expressions with the salience order 'Sh > C > Si'.
+English referring expressions are therefore explicitly passed to the agents as correct labels during training.
+One can therefore expect the agents to align their own language with the target referring expressions and a high correlation with the salience order 'Sh > C > Si' is expected.
+Looking at the results, this is partially confirmed for the 'Dale-2' dataset.
+Some emerged languages reach with a normalized loss of 49,22\% one of the highest correlations to English referring expressions across all emerged languages in this thesis.
+However, the salience order with the highest correlation is 'Sh > Si > C', and values the \emph{size} more than the \emph{color}, instead of the reversed way as in the training data of the agents.
+The color seems, as in the previous task, not as important to communicate as the other attributes.
+While the \emph{size} plays a bigger role, the \emph{shape} is the central attribute in the communication.
+This fact is interesting for two reasons:
+First, the \emph{shape} being the most important attribute in the messages aligns with the receiver's generated referring expressions where the \emph{shape} was predicted almost perfectly.
+The receiver therefore seems to make use of the sender's ability to extract the \emph{shape} from the images and reproduces what the sender communicated.
+However, without a deeper analysis, it can't be ruled out that the receiver extracts the \emph{shape} themselves from the image, and uses the sender's message only to determine the correct target object in the scene.
+Secondly, this does not apply to the \emph{size} and the \emph{color}.
+While the size seems to be communicated, it is not generated correctly by the receiver.
+The opposite applies to the \emph{color}.
+The \emph{color} is not important in the communication, but the receiver however is able to produce correct colors in the generated referring expression.
+This shows that the agents actually don't align their own language with the target referring expression.
+The communication of the target object between the agents seems to be a separate step from the receiver's generation of the referring expression.
+
+Looking at the influence of the message length, one can see that the language is the least correlated to the salience order 'Sh > Si > C' with $n=1$.
+Still, even with these short messages, the loss can go down to 66,82\%.
+With $n>1$, almost all languages have a loss of below 70\%, while at least one configuration is even lower than 60\%.
+Hereby, languages with $n \in \{3,4\}$ have the highest correlation to English.
+Languages with the highest correlation also mostly make use of a medium-sized vocabulary of $|V| \in \{10,16\}$ as for example $Lang_{2,10}$, $Lang_{3,16}$, $Lang_{4,16}$ and $Lang_{6,10}$.
+While larger vocabularies, can still correlate relatively much with English referring expressions, languages with $|V|=2$ are much less similar.
 
@@
     \end{tabular}
-    \caption{RE generator: Dale-5}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{referring expression generation games} on the 'Dale-5' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:re-generator:dale-5}
 \end{table}
+
+On the 'Dale-5' dataset, the emerged languages again resemble the English referring expressions less.
+While the dominant salience order is 'Sh > Si > C' as on the 'Dale-2' dataset, the \emph{color} plays a bigger role.
+This effect is the same as for the discrimination games, and can be explained by the higher number or distractors.
+While the loss for the salience order 'Sh > Si > C' is lower than 70\% for two languages, it stays between 70\% and 80\% for most.
 
@@
     \end{tabular}
-    \caption{Attribute generator: Dale-2}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attribute generation games} on the 'Dale-2' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:attribute-generator:dale-2}
@@
     \end{tabular}
-    \caption{Attribute generator: Dale-5}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attribute generation games} on the 'Dale-5' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:attribute-generator:dale-5}
@@
     \end{tabular}
-    \caption{Attribute generator: Color}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attribute generation games} on the 'CLEVR color' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:attribute-generator:colour}
@@
     \end{tabular}
-    \caption{Attention predictor: Dale-2}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attention prediction games} on the 'Dale-2' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:attention-predictor:dale-2}
@@
     \end{tabular}
-    \caption{Attention predictor: Dale-5}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attention prediction games} on the 'Dale-5' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:attention-predictor:dale-5}
@@
     \end{tabular}
-    \caption{Attention predictor: Color}
+    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attention prediction games} on the 'CLEVR color' dataset.
+The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}.
+The table only includes languages, with which the agents could successfully solve the task.}
     \label{tab:probing:attention-predictor:colour}
@@
 
-\cmtDK[inline]{differnces in different languages correspond to differences in color prediction of single model caption generation}
+\cmtDK[inline]{differences in different languages correspond to differences in color prediction of single model caption generation}
 

### 2024-05-08 thesis/sections/6.1-object-identification.tex
language analysis

@@
         {100}                         & {1000}    & {4} & {2}   & {94,88\%}                           \\
-        {100}                         & {1000}    & {6} & {10}  & \textcolor{purple}{63,15\%}         \\
-        {1000}                        & {500}     & {2} & {2}   & \textcolor{purple}{62,98\%}         \\\midrule
+        {100}                         & {1000}    & {6} & {10}  & \textcolor{red}{63,15\%}            \\
+        {1000}                        & {500}     & {2} & {2}   & \textcolor{red}{62,98\%}            \\\midrule
         \multicolumn{4}{c|}{baseline} & {62,72\%}                                                     \\\midrule
-        {100}                         & {1000}    & {3} & {2}   & \textcolor{purple}{61,2\%}          \\
+        {100}                         & {1000}    & {3} & {2}   & \textcolor{red}{61,2\%}             \\
         \bottomrule
@@
         \multicolumn{4}{c|}{baseline} & {35,07\%}                                                     \\\midrule
-        {500}                         & {1000}    & {2} & {100} & \textcolor{purple}{34,98\%}         \\
-        {500}                         & {500}     & {3} & {16}  & \textcolor{purple}{34,77\%}         \\
-        {500}                         & {500}     & {6} & {2}   & \textcolor{purple}{34,38\%}         \\
-        {1000}                        & {1000}    & {4} & {2}   & \textcolor{purple}{35,33\%}         \\
-        {1000}                        & {1000}    & {6} & {100} & \textcolor{purple}{35,24\%}         \\
+        {500}                         & {1000}    & {2} & {100} & \textcolor{red}{34,98\%}            \\
+        {500}                         & {500}     & {3} & {16}  & \textcolor{red}{34,77\%}            \\
+        {500}                         & {500}     & {6} & {2}   & \textcolor{red}{34,38\%}            \\
+        {1000}                        & {1000}    & {4} & {2}   & \textcolor{red}{35,33\%}            \\
+        {1000}                        & {1000}    & {6} & {100} & \textcolor{red}{35,24\%}            \\
         \bottomrule
@@
         {100}                         & {500}     & {3} & {2}   & {51,48\%}                                \\
-        {500}                         & {1000}    & {3} & {16}  & \textcolor{purple}{24,61\%}              \\
-        {100}                         & {1000}    & {3} & {50}  & \textcolor{purple}{24,48\%}              \\
-        {1000}                        & {1000}    & {4} & {100} & \textcolor{purple}{24\%}                 \\
-        {500}                         & {1000}    & {1} & {2}   & \textcolor{purple}{23,96\%}              \\
-        {1000}                        & {100}     & {6} & {10}  & \textcolor{purple}{23,87\%}              \\
-        {500}                         & {1000}    & {6} & {50}  & \textcolor{purple}{23,87\%}              \\\midrule
+        {500}                         & {1000}    & {3} & {16}  & \textcolor{red}{24,61\%}                 \\
+        {100}                         & {1000}    & {3} & {50}  & \textcolor{red}{24,48\%}                 \\
+        {1000}                        & {1000}    & {4} & {100} & \textcolor{red}{24\%}                    \\
+        {500}                         & {1000}    & {1} & {2}   & \textcolor{red}{23,96\%}                 \\
+        {1000}                        & {100}     & {6} & {10}  & \textcolor{red}{23,87\%}                 \\
+        {500}                         & {1000}    & {6} & {50}  & \textcolor{red}{23,87\%}                 \\\midrule
         \multicolumn{4}{c|}{baseline} & {23,74\%}                                                          \\\midrule
-        {500}                         & {500}     & {2} & {2}   & \textcolor{purple}{22,35\%}              \\
+        {500}                         & {500}     & {2} & {2}   & \textcolor{red}{22,35\%}                 \\
         \bottomrule

### 2024-05-08 thesis/sections/6.2-re-generation.tex
language analysis

@@
         \multicolumn{2}{c|}{baseline} & {41,8} & {42,36}                             & {37,11}                             & {62,66}                                  & {81,16}                & {18,36}                & {13,52}                & {33,08}                & {34,53}                                         \\\midrule
-        {1}                           & {2}    & {70,49}                             & {54,73}                             & {7,29}                                   & {70,16}                   & {85,06}                   & {13,06}                   & \textcolor{purple}{11,68} & \textcolor{purple}{32,25} & \textcolor{purple}{34,03} \\
-        {1}                           & {10}   & {78,26}                             & {63,33}                             & {3,56}                                   & {68,97}                   & {85,01}                   & {10,85}                   & \textcolor{purple}{12,72} & \textcolor{purple}{32,77} & \textcolor{purple}{33,85} \\
-        {1}                           & {16}   & {74,18}                             & {52,21}                             & {5,38}                                   & {71,57}                   & {86,87}                   & {10,72}                   & \textcolor{purple}{13,19} & \textcolor{purple}{33,64} & \textcolor{purple}{33,33} \\
-        {1}                           & {50}   & {70,83}                             & \textcolor{purple}{43,56}           & {4,21}                                   & {65,23}                   & {83,11}                   & {12,54}                   & \textcolor{purple}{11,81} & \textcolor{purple}{33,14} & \textcolor{purple}{34,72} \\
-        {1}                           & {100}  & {73,18}                             & {52,14}                             & {3,86}                                   & \textcolor{purple}{60,94} & \textcolor{purple}{80,85} & \textcolor{purple}{20,36} & \textcolor{purple}{13,67} & \textcolor{purple}{33,71} & \textcolor{purple}{33,64} \\
-        {2}                           & {2}    & {72,31}                             & {50,42}                             & {5,52}                                   & {66,2}                    & {83,39}                   & {14,42}                   & \textcolor{purple}{13,32} & \textcolor{purple}{33,62} & \textcolor{purple}{32,73} \\
-        {2}                           & {10}   & {75,22}                             & {58,4}                              & {3,99}                                   & {73,22}                   & {87,68}                   & {9,03}                    & \textcolor{purple}{12,37} & \textcolor{purple}{33,3}  & \textcolor{purple}{34,11} \\
-        {2}                           & {16}   & {\textbf{81,64}}                    & {\textbf{73,66}}                    & {\textbf{2,86}}                          & {73,44}                   & {88,21}                   & {8,85}                    & \textcolor{purple}{12,98} & \textcolor{purple}{33,94} & \textcolor{purple}{34,24} \\
-        {2}                           & {50}   & {74,39}                             & {48,53}                             & {3,17}                                   & {\textbf{75,35}}          & \textbf{88,77}            & \textbf{8,07}             & \textcolor{purple}{14,19} & \textcolor{purple}{33,04} & \textcolor{purple}{33,46} \\
-        {2}                           & {100}  & {74,09}                             & {54,88}                             & {3,65}                                   & \textcolor{purple}{54,56} & \textcolor{purple}{76,95} & \textcolor{purple}{22,57} & \textcolor{purple}{13,28} & \textcolor{purple}{32,8}  & \textcolor{purple}{33,38} \\
-        {3}                           & {2}    & {71,03}                             & {50,76}                             & {4,62}                                   & {65,18}                   & \textcolor{purple}{82,68} & {14,23}                   & \textcolor{purple}{12,76} & \textcolor{purple}{33,58} & \textcolor{purple}{33,64} \\
-        {3}                           & {10}   & {76,91}                             & {59,42}                             & {2,6}                                    & {71,66}                   & {87,14}                   & {9,9}                     & \textcolor{purple}{12,54} & \textcolor{purple}{32,75} & \textcolor{purple}{33,85} \\
-        {3}                           & {16}   & {79,95}                             & {67,55}                             & {3,17}                                   & \textcolor{purple}{50,13} & \textcolor{purple}{73,24} & {24,22}                   & \textcolor{purple}{12,93} & \textcolor{purple}{33,79} & \textcolor{purple}{34,46} \\
-        {3}                           & {50}   & {62,93}                             & {48,03}                             & {16,58}                                  & {\textbf{76,09}}          & {\textbf{88,5}}           & \textbf{9,2}              & \textcolor{purple}{13,67} & \textcolor{purple}{33,73} & \textcolor{purple}{34,29} \\
-        {3}                           & {100}  & {75,74}                             & {57,51}                             & {2,43}                                   & {69,7}                    & {86,14}                   & {9,38}                    & \textcolor{purple}{13,02} & \textcolor{purple}{33,64} & \textcolor{purple}{34,2}  \\
-        {4}                           & {2}    & \textcolor{purple}{41,23}           & \textcolor{purple}{42,53}           & \textcolor{purple}{35,33}                & {63,93}                   & \textcolor{purple}{82,34} & {15,3}                    & \textcolor{purple}{12,2}  & \textcolor{purple}{33,25} & \textcolor{purple}{33,64} \\
-        {4}                           & {10}   & {78,99}                             & {66,4}                              & {2,08}                                   & {73,22}                   & {88,01}                   & {10,85}                   & \textcolor{purple}{12,5}  & \textcolor{purple}{33,55} & \textcolor{purple}{33,16} \\
-        {4}                           & {16}   & {\textbf{80,08}}                    & {\textbf{70,48}}                    & {\textbf{3,47}}                          & {69,4}                    & {86,09}                   & {10,63}                   & \textcolor{purple}{11,72} & \textcolor{purple}{33,1}  & \textcolor{purple}{34,03} \\
-        {4}                           & {50}   & {75,48}                             & {54,44}                             & {3,34}                                   & {71,35}                   & {86,61}                   & {10,68}                   & \textcolor{purple}{11,81} & \textcolor{purple}{33,18} & \textcolor{purple}{34,33} \\
-        {4}                           & {100}  & {74,78}                             & {57,09}                             & {3,86}                                   & \textcolor{purple}{59,2}  & \textcolor{purple}{80,14} & \textcolor{purple}{20,57} & \textcolor{purple}{12,67} & \textcolor{purple}{32,3}  & \textcolor{purple}{34,55} \\
-        {6}                           & {2}    & {72,87}                             & {51,44}                             & {2,78}                                   & {60,82}                   & \textcolor{purple}{80,42} & {17,07}                   & \textcolor{purple}{13,45} & \textcolor{purple}{33,73} & \textcolor{purple}{33,64} \\
-        {6}                           & {10}   & {73,61}                             & {48,73}                             & {2,82}                                   & \textcolor{purple}{32,64} & \textcolor{purple}{58,42} & \textcolor{purple}{33,12} & \textcolor{purple}{11,89} & \textcolor{purple}{33,15} & \textcolor{purple}{35,03} \\
-        {6}                           & {16}   & {73,52}                             & {49,67}                             & {2,43}                                   & \textcolor{purple}{49,39} & \textcolor{purple}{73,19} & \textcolor{purple}{23,78} & \textcolor{purple}{13,06} & \textcolor{purple}{32,31} & \textcolor{purple}{34,64} \\
-        {6}                           & {50}   & \textcolor{purple}{44,31}           & {45,07}                             & \textcolor{purple}{33,29}                & {\textbf{78,82}}          & {\textbf{90,22}}          & {\textbf{7,29}}           & \textcolor{purple}{12,72} & \textcolor{purple}{33,29} & \textcolor{purple}{32,99} \\
-        {6}                           & {100}  & {\textbf{80,95}}                    & {\textbf{69,88}}                    & {\textbf{3,34}}                          & {62,2}                    & \textcolor{purple}{80,62} & \textcolor{purple}{19,88} & \textcolor{purple}{14,24} & \textcolor{purple}{32,36} & \textcolor{purple}{32,94} \\
+        {1}                           & {2}    & {70,49}                             & {54,73}                             & {7,29}                                   & {70,16}                & {85,06}                & {13,06}                & \textcolor{red}{11,68} & \textcolor{red}{32,25} & \textcolor{red}{34,03} \\
+        {1}                           & {10}   & {78,26}                             & {63,33}                             & {3,56}                                   & {68,97}                & {85,01}                & {10,85}                & \textcolor{red}{12,72} & \textcolor{red}{32,77} & \textcolor{red}{33,85} \\
+        {1}                           & {16}   & {74,18}                             & {52,21}                             & {5,38}                                   & {71,57}                & {86,87}                & {10,72}                & \textcolor{red}{13,19} & \textcolor{red}{33,64} & \textcolor{red}{33,33} \\
+        {1}                           & {50}   & {70,83}                             & \textcolor{red}{43,56}              & {4,21}                                   & {65,23}                & {83,11}                & {12,54}                & \textcolor{red}{11,81} & \textcolor{red}{33,14} & \textcolor{red}{34,72} \\
+        {1}                           & {100}  & {73,18}                             & {52,14}                             & {3,86}                                   & \textcolor{red}{60,94} & \textcolor{red}{80,85} & \textcolor{red}{20,36} & \textcolor{red}{13,67} & \textcolor{red}{33,71} & \textcolor{red}{33,64} \\
+        {2}                           & {2}    & {72,31}                             & {50,42}                             & {5,52}                                   & {66,2}                 & {83,39}                & {14,42}                & \textcolor{red}{13,32} & \textcolor{red}{33,62} & \textcolor{red}{32,73} \\
+        {2}                           & {10}   & {75,22}                             & {58,4}                              & {3,99}                                   & {73,22}                & {87,68}                & {9,03}                 & \textcolor{red}{12,37} & \textcolor{red}{33,3}  & \textcolor{red}{34,11} \\
+        {2}                           & {16}   & {\textbf{81,64}}                    & {\textbf{73,66}}                    & {\textbf{2,86}}                          & {73,44}                & {88,21}                & {8,85}                 & \textcolor{red}{12,98} & \textcolor{red}{33,94} & \textcolor{red}{34,24} \\
+        {2}                           & {50}   & {74,39}                             & {48,53}                             & {3,17}                                   & {\textbf{75,35}}       & \textbf{88,77}         & \textbf{8,07}          & \textcolor{red}{14,19} & \textcolor{red}{33,04} & \textcolor{red}{33,46} \\
+        {2}                           & {100}  & {74,09}                             & {54,88}                             & {3,65}                                   & \textcolor{red}{54,56} & \textcolor{red}{76,95} & \textcolor{red}{22,57} & \textcolor{red}{13,28} & \textcolor{red}{32,8}  & \textcolor{red}{33,38} \\
+        {3}                           & {2}    & {71,03}                             & {50,76}                             & {4,62}                                   & {65,18}                & \textcolor{red}{82,68} & {14,23}                & \textcolor{red}{12,76} & \textcolor{red}{33,58} & \textcolor{red}{33,64} \\
+        {3}                           & {10}   & {76,91}                             & {59,42}                             & {2,6}                                    & {71,66}                & {87,14}                & {9,9}                  & \textcolor{red}{12,54} & \textcolor{red}{32,75} & \textcolor{red}{33,85} \\
+        {3}                           & {16}   & {79,95}                             & {67,55}                             & {3,17}                                   & \textcolor{red}{50,13} & \textcolor{red}{73,24} & {24,22}                & \textcolor{red}{12,93} & \textcolor{red}{33,79} & \textcolor{red}{34,46} \\
+        {3}                           & {50}   & {62,93}                             & {48,03}                             & {16,58}                                  & {\textbf{76,09}}       & {\textbf{88,5}}        & \textbf{9,2}           & \textcolor{red}{13,67} & \textcolor{red}{33,73} & \textcolor{red}{34,29} \\
+        {3}                           & {100}  & {75,74}                             & {57,51}                             & {2,43}                                   & {69,7}                 & {86,14}                & {9,38}                 & \textcolor{red}{13,02} & \textcolor{red}{33,64} & \textcolor{red}{34,2}  \\
+        {4}                           & {2}    & \textcolor{red}{41,23}              & \textcolor{red}{42,53}              & \textcolor{red}{35,33}                   & {63,93}                & \textcolor{red}{82,34} & {15,3}                 & \textcolor{red}{12,2}  & \textcolor{red}{33,25} & \textcolor{red}{33,64} \\
+        {4}                           & {10}   & {78,99}                             & {66,4}                              & {2,08}                                   & {73,22}                & {88,01}                & {10,85}                & \textcolor{red}{12,5}  & \textcolor{red}{33,55} & \textcolor{red}{33,16} \\
+        {4}                           & {16}   & {\textbf{80,08}}                    & {\textbf{70,48}}                    & {\textbf{3,47}}                          & {69,4}                 & {86,09}                & {10,63}                & \textcolor{red}{11,72} & \textcolor{red}{33,1}  & \textcolor{red}{34,03} \\
+        {4}                           & {50}   & {75,48}                             & {54,44}                             & {3,34}                                   & {71,35}                & {86,61}                & {10,68}                & \textcolor{red}{11,81} & \textcolor{red}{33,18} & \textcolor{red}{34,33} \\
+        {4}                           & {100}  & {74,78}                             & {57,09}                             & {3,86}                                   & \textcolor{red}{59,2}  & \textcolor{red}{80,14} & \textcolor{red}{20,57} & \textcolor{red}{12,67} & \textcolor{red}{32,3}  & \textcolor{red}{34,55} \\
+        {6}                           & {2}    & {72,87}                             & {51,44}                             & {2,78}                                   & {60,82}                & \textcolor{red}{80,42} & {17,07}                & \textcolor{red}{13,45} & \textcolor{red}{33,73} & \textcolor{red}{33,64} \\
+        {6}                           & {10}   & {73,61}                             & {48,73}                             & {2,82}                                   & \textcolor{red}{32,64} & \textcolor{red}{58,42} & \textcolor{red}{33,12} & \textcolor{red}{11,89} & \textcolor{red}{33,15} & \textcolor{red}{35,03} \\
+        {6}                           & {16}   & {73,52}                             & {49,67}                             & {2,43}                                   & \textcolor{red}{49,39} & \textcolor{red}{73,19} & \textcolor{red}{23,78} & \textcolor{red}{13,06} & \textcolor{red}{32,31} & \textcolor{red}{34,64} \\
+        {6}                           & {50}   & \textcolor{red}{44,31}              & {45,07}                             & \textcolor{red}{33,29}                   & {\textbf{78,82}}       & {\textbf{90,22}}       & {\textbf{7,29}}        & \textcolor{red}{12,72} & \textcolor{red}{33,29} & \textcolor{red}{32,99} \\
+        {6}                           & {100}  & {\textbf{80,95}}                    & {\textbf{69,88}}                    & {\textbf{3,34}}                          & {62,2}                 & \textcolor{red}{80,62} & \textcolor{red}{19,88} & \textcolor{red}{14,24} & \textcolor{red}{32,36} & \textcolor{red}{32,94} \\
         \bottomrule
@@
         \multicolumn{2}{c|}{baseline} & {88,91} & {92,03}                             & {50,55}                             & {80,78}                                  & {89,38}                         & {49,92}                         & {100}                          & {90,78}                         & {76,72}                                                          \\\midrule
-        {1}                           & {2}     & \textcolor{purple}{89,69}           & {93,68}                             & \textcolor{purple}{52,04}                & {88,09}                         & \textcolor{purple}{90,12}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{87,37}       & \textcolor{purple}{76,56}      \\
-        {1}                           & {10}    & {96,31}                             & {94,21}                             & \textcolor{purple}{52,78}                & {90,6}                          & \textcolor{purple}{90,15}       & \textcolor{purple}{49,74}      & {100}                           & \textbf{91,32}                  & \textcolor{purple}{76,52}      \\
-        {1}                           & {16}    & {94,97}                             & \textcolor{purple}{92,8}            & \textcolor{purple}{52,78}                & {91,97}                         & {93,45}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{91,03}       & \textcolor{purple}{76,52}      \\
-        {1}                           & {50}    & \textcolor{purple}{90,1}            & \textcolor{purple}{91,58}           & \textcolor{purple}{52,78}                & {94,66}                         & {93,39}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,41}       & \textcolor{purple}{76,52}      \\
-        {1}                           & {100}   & {90,41}                             & \textcolor{purple}{92,66}           & \textcolor{purple}{52,78}                & \textcolor{purple}{83,03}       & \textcolor{purple}{90,23}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,72}       & \textcolor{purple}{76,52}      \\
-        {2}                           & {2}     & {92,45}                             & \textcolor{purple}{92,45}           & \textcolor{purple}{52,04}                & {88,72}                         & \textbf{94,38}                  & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,89}       & \textcolor{purple}{76,56}      \\
-        {2}                           & {10}    & {96,09}                             & \textbf{94,44}                      & \textcolor{purple}{52,78}                & {91,54}                         & \textcolor{purple}{90,71}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,54}       & \textcolor{purple}{76,52}      \\
-        {2}                           & {16}    & {90,58}                             & \textcolor{purple}{92,66}           & \textcolor{purple}{52,78}                & \textbf{97,18}                  & {94,14}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,98}       & \textcolor{purple}{76,52}      \\
-        {2}                           & {50}    & {89,28}                             & \textcolor{purple}{92,45}           & \textcolor{purple}{52,78}                & {92,01}                         & {93,01}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,5}        & \textcolor{purple}{76,52}      \\
-        {2}                           & {100}   & \textbf{98,91}                      & \textcolor{purple}{92,97}           & \textcolor{purple}{52,78}                & {92,66}                         & {92,49}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,68}       & \textcolor{purple}{76,52}      \\
-        {3}                           & {2}     & {90,97}                             & {93,32}                             & \textcolor{purple}{52,04}                & {87,88}                         & {92,53}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{90,58}       & \textcolor{purple}{76,56}      \\
-        {3}                           & {10}    & {97,01}                             & \textbf{{94,79}}                    & \textcolor{purple}{52,78}                & {94,4}                          & {92,8}                          & \textcolor{purple}{49,74}      & {100}                           & \textbf{91,17}                  & \textcolor{purple}{76,52}      \\
-        {3}                           & {16}    & {95,36}                             & {93,1}                              & \textcolor{purple}{52,78}                & {86,72}                         & {91,88}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,06}       & \textcolor{purple}{76,52}      \\
-        {3}                           & {50}    & {90,93}                             & \textcolor{purple}{90,84}           & \textcolor{purple}{52,78}                & {94,92}                         & \textbf{94,18}                  & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,59}       & \textcolor{purple}{76,52}      \\
-        {3}                           & {100}   & \textcolor{purple}{89,37}           & \textcolor{purple}{90,76}           & \textcolor{purple}{52,78}                & {94,1}                          & {92,8}                          & \textcolor{purple}{49,74}      & {100}                           & \textbf{91,84}                  & \textcolor{purple}{76,52}      \\
-        {4}                           & {2}     & {94,86}                             & \textcolor{purple}{92,11}           & \textcolor{purple}{52,04}                & {88,56}                         & {93,86}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{87,11}       & \textcolor{purple}{76,56}      \\
-        {4}                           & {10}    & {90,58}                             & \textcolor{purple}{92,75}           & \textcolor{purple}{52,78}                & {95,05}                         & {93,71}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,89}       & \textcolor{purple}{76,52}      \\
-        {4}                           & {16}    & \textbf{98,57}                      & {94,14}                             & \textcolor{purple}{52,78}                & \textbf{96,14}                  & \textbf{95,23}                  & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,06}       & \textcolor{purple}{76,52}      \\
-        {4}                           & {50}    & {95,66}                             & {93,32}                             & \textcolor{purple}{52,78}                & {91,01}                         & \textcolor{purple}{90,45}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,85}       & \textcolor{purple}{76,52}      \\
-        {4}                           & {100}   & \textcolor{purple}{89,41}           & {93,58}                             & \textcolor{purple}{52,78}                & {94,66}                         & {93,19}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{86,98}       & \textcolor{purple}{76,52}      \\
-        {6}                           & {2}     & {93,39}                             & {92,94}                             & \textcolor{purple}{52,04}                & \textcolor{purple}{82,18}       & {92,16}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{90,05}       & \textcolor{purple}{76,56}      \\
-        {6}                           & {10}    & \textbf{98,78}                      & \textbf{94,23}                      & \textcolor{purple}{52,78}                & \textbf{96,79}                  & {92,75}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{87,67}       & \textcolor{purple}{76,52}      \\
-        {6}                           & {16}    & {96,96}                             & \textcolor{purple}{91,32}           & \textcolor{purple}{52,78}                & {95,57}                         & {93,75}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,37}       & \textcolor{purple}{76,52}      \\
-        {6}                           & {50}    & \textcolor{purple}{88,28}           & \textcolor{purple}{91,45}           & \textcolor{purple}{52,78}                & {91,71}                         & \textcolor{purple}{86,24}       & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{88,32}       & \textcolor{purple}{76,52}      \\
-        {6}                           & {100}   & {94,92}                             & \textcolor{purple}{92,45}           & \textcolor{purple}{52,78}                & {95,3}                          & {93,62}                         & \textcolor{purple}{49,74}      & {100}                           & \textcolor{purple}{89,19}       & \textcolor{purple}{76,52}      \\
+        {1}                           & {2}     & \textcolor{red}{89,69}              & {93,68}                             & \textcolor{red}{52,04}                   & {88,09}                         & \textcolor{red}{90,12}          & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{87,37}          & \textcolor{red}{76,56}         \\
+        {1}                           & {10}    & {96,31}                             & {94,21}                             & \textcolor{red}{52,78}                   & {90,6}                          & \textcolor{red}{90,15}          & \textcolor{red}{49,74}         & {100}                           & \textbf{91,32}                  & \textcolor{red}{76,52}         \\
+        {1}                           & {16}    & {94,97}                             & \textcolor{red}{92,8}               & \textcolor{red}{52,78}                   & {91,97}                         & {93,45}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{91,03}          & \textcolor{red}{76,52}         \\
+        {1}                           & {50}    & \textcolor{red}{90,1}               & \textcolor{red}{91,58}              & \textcolor{red}{52,78}                   & {94,66}                         & {93,39}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{89,41}          & \textcolor{red}{76,52}         \\
+        {1}                           & {100}   & {90,41}                             & \textcolor{red}{92,66}              & \textcolor{red}{52,78}                   & \textcolor{red}{83,03}          & \textcolor{red}{90,23}          & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{86,72}          & \textcolor{red}{76,52}         \\
+        {2}                           & {2}     & {92,45}                             & \textcolor{red}{92,45}              & \textcolor{red}{52,04}                   & {88,72}                         & \textbf{94,38}                  & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{88,89}          & \textcolor{red}{76,56}         \\
+        {2}                           & {10}    & {96,09}                             & \textbf{94,44}                      & \textcolor{red}{52,78}                   & {91,54}                         & \textcolor{red}{90,71}          & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{89,54}          & \textcolor{red}{76,52}         \\
+        {2}                           & {16}    & {90,58}                             & \textcolor{red}{92,66}              & \textcolor{red}{52,78}                   & \textbf{97,18}                  & {94,14}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{86,98}          & \textcolor{red}{76,52}         \\
+        {2}                           & {50}    & {89,28}                             & \textcolor{red}{92,45}              & \textcolor{red}{52,78}                   & {92,01}                         & {93,01}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{86,5}           & \textcolor{red}{76,52}         \\
+        {2}                           & {100}   & \textbf{98,91}                      & \textcolor{red}{92,97}              & \textcolor{red}{52,78}                   & {92,66}                         & {92,49}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{86,68}          & \textcolor{red}{76,52}         \\
+        {3}                           & {2}     & {90,97}                             & {93,32}                             & \textcolor{red}{52,04}                   & {87,88}                         & {92,53}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{90,58}          & \textcolor{red}{76,56}         \\
+        {3}                           & {10}    & {97,01}                             & \textbf{{94,79}}                    & \textcolor{red}{52,78}                   & {94,4}                          & {92,8}                          & \textcolor{red}{49,74}         & {100}                           & \textbf{91,17}                  & \textcolor{red}{76,52}         \\
+        {3}                           & {16}    & {95,36}                             & {93,1}                              & \textcolor{red}{52,78}                   & {86,72}                         & {91,88}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{88,06}          & \textcolor{red}{76,52}         \\
+        {3}                           & {50}    & {90,93}                             & \textcolor{red}{90,84}              & \textcolor{red}{52,78}                   & {94,92}                         & \textbf{94,18}                  & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{88,59}          & \textcolor{red}{76,52}         \\
+        {3}                           & {100}   & \textcolor{red}{89,37}              & \textcolor{red}{90,76}              & \textcolor{red}{52,78}                   & {94,1}                          & {92,8}                          & \textcolor{red}{49,74}         & {100}                           & \textbf{91,84}                  & \textcolor{red}{76,52}         \\
+        {4}                           & {2}     & {94,86}                             & \textcolor{red}{92,11}              & \textcolor{red}{52,04}                   & {88,56}                         & {93,86}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{87,11}          & \textcolor{red}{76,56}         \\
+        {4}                           & {10}    & {90,58}                             & \textcolor{red}{92,75}              & \textcolor{red}{52,78}                   & {95,05}                         & {93,71}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{89,89}          & \textcolor{red}{76,52}         \\
+        {4}                           & {16}    & \textbf{98,57}                      & {94,14}                             & \textcolor{red}{52,78}                   & \textbf{96,14}                  & \textbf{95,23}                  & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{88,06}          & \textcolor{red}{76,52}         \\
+        {4}                           & {50}    & {95,66}                             & {93,32}                             & \textcolor{red}{52,78}                   & {91,01}                         & \textcolor{red}{90,45}          & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{88,85}          & \textcolor{red}{76,52}         \\
+        {4}                           & {100}   & \textcolor{red}{89,41}              & {93,58}                             & \textcolor{red}{52,78}                   & {94,66}                         & {93,19}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{86,98}          & \textcolor{red}{76,52}         \\
+        {6}                           & {2}     & {93,39}                             & {92,94}                             & \textcolor{red}{52,04}                   & \textcolor{red}{82,18}          & {92,16}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{90,05}          & \textcolor{red}{76,56}         \\
+        {6}                           & {10}    & \textbf{98,78}                      & \textbf{94,23}                      & \textcolor{red}{52,78}                   & \textbf{96,79}                  & {92,75}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{87,67}          & \textcolor{red}{76,52}         \\
+        {6}                           & {16}    & {96,96}                             & \textcolor{red}{91,32}              & \textcolor{red}{52,78}                   & {95,57}                         & {93,75}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{89,37}          & \textcolor{red}{76,52}         \\
+        {6}                           & {50}    & \textcolor{red}{88,28}              & \textcolor{red}{91,45}              & \textcolor{red}{52,78}                   & {91,71}                         & \textcolor{red}{86,24}          & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{88,32}          & \textcolor{red}{76,52}         \\
+        {6}                           & {100}   & {94,92}                             & \textcolor{red}{92,45}              & \textcolor{red}{52,78}                   & {95,3}                          & {93,62}                         & \textcolor{red}{49,74}         & {100}                           & \textcolor{red}{89,19}          & \textcolor{red}{76,52}         \\
         \bottomrule

### 2024-05-08 thesis/sections/6.3-reference-resolution.tex
language analysis

@@
         \multicolumn{2}{c|}{baseline} & {34,49} & {34,3\%}                            & {38,13}                             & {30,31\%}                                & {42,72}                  & {18,28\%}                                         \\\midrule
-        {1}                           & {2}     & \textcolor{purple}{34,91}           & \textcolor{purple}{34,81\%}         & \textcolor{purple}{40,61}                & \textcolor{purple}{25,48\%} & {39,13}                   & {25\%}                      \\
-        {1}                           & {10}    & \textcolor{purple}{35,31}           & {40,15\%}                           & \textcolor{purple}{38,73}                & \textcolor{purple}{28,12\%} & \textcolor{purple}{44,19} & {20,53\%}                   \\
-        {1}                           & {16}    & \textcolor{purple}{35,4}            & {42,93\%}                           & \textcolor{purple}{39,58}                & \textcolor{purple}{26,17\%} & {39,64}                   & {24,18\%}                   \\
-        {1}                           & {50}    & {32,64}                             & {42,01\%}                           & \textcolor{purple}{38,99}                & \textcolor{purple}{28,52\%} & \textcolor{purple}{42,61} & {20,79\%}                   \\
-        {1}                           & {100}   & \textcolor{purple}{36,73}           & \textcolor{purple}{31,21\%}         & \textcolor{purple}{40,25}                & \textcolor{purple}{24,91\%} & \textcolor{purple}{41,5}  & {21,44\%}                   \\
-        {2}                           & {2}     & \textcolor{purple}{33,76}           & {40,67\%}                           & {35,86}                                  & {31,73\%}                   & \textcolor{purple}{41,79} & {22,14\%}                   \\
-        {2}                           & {10}    & {31,37}                             & {43,27\%}                           & {37,23}                                  & \textcolor{purple}{30,34\%} & \textcolor{purple}{42,53} & \textcolor{purple}{19,75\%} \\
-        {2}                           & {16}    & {31,19}                             & {46,27\%}                           & \textcolor{purple}{39,38}                & \textcolor{purple}{28,21\%} & \textcolor{purple}{41,74} & {22,7\%}                    \\
-        {2}                           & {50}    & \textcolor{purple}{33,66}           & {39,28\%}                           & \textcolor{purple}{39,62}                & \textcolor{purple}{25,35\%} & \textcolor{purple}{45,63} & \textcolor{purple}{17,93\%} \\
-        {2}                           & {100}   & \textcolor{purple}{35,1}            & \textcolor{purple}{34,81\%}         & \textcolor{purple}{39,86}                & \textcolor{purple}{27,3\%}  & \textcolor{purple}{45,45} & \textcolor{purple}{19,05\%} \\
-        {3}                           & {2}     & {32,99}                             & {41,15\%}                           & {37,76}                                  & \textcolor{purple}{30,99\%} & \textcolor{purple}{46,45} & {20,92\%}                   \\
-        {3}                           & {10}    & \textcolor{purple}{33,74}           & {38,89\%}                           & \textcolor{purple}{40,71}                & \textcolor{purple}{26,48\%} & \textcolor{purple}{44,53} & \textcolor{purple}{19,01\%} \\
-        {3}                           & {16}    & \textcolor{purple}{34,18}           & {38,98\%}                           & \textcolor{purple}{38,52}                & \textcolor{purple}{30,03\%} & \textcolor{purple}{41,22} & {21,53\%}                   \\
-        {3}                           & {50}    & \textcolor{purple}{34,38}           & {38,93\%}                           & \textcolor{purple}{39,52}                & \textcolor{purple}{26,74\%} & {38,4}                    & {24,87\%}                   \\
-        {3}                           & {100}   & \textcolor{purple}{33,16}           & {40,19\%}                           & \textcolor{purple}{39,75}                & \textcolor{purple}{27,78\%} & \textcolor{purple}{41,29} & {21,83\%}                   \\
-        {4}                           & {2}     & \textcolor{purple}{33,48}           & {40,36\%}                           & \textcolor{purple}{38,28}                & \textcolor{purple}{30,9\%}  & \textcolor{purple}{42,53} & {22,4\%}                    \\
-        {4}                           & {10}    & \textcolor{purple}{35,01}           & {39,97\%}                           & \textcolor{purple}{40,88}                & \textcolor{purple}{26,39\%} & \textcolor{purple}{42,33} & {20,44\%}                   \\
-        {4}                           & {16}    & \textcolor{purple}{35,59}           & \textcolor{purple}{35,59\%}         & \textcolor{purple}{37,91}                & {31,86\%}                   & \textcolor{purple}{41,9}  & {20,27\%}                   \\
-        {4}                           & {50}    & \textcolor{purple}{34,99}           & {36,68\%}                           & \textcolor{purple}{40,89}                & \textcolor{purple}{23,26\%} & \textcolor{purple}{43,77} & \textcolor{purple}{18,36\%} \\
-        {4}                           & {100}   & \textcolor{purple}{35,46}           & {37,46\%}                           & \textcolor{purple}{38,53}                & \textcolor{purple}{29,08\%} & {40,6}                    & {21,22\%}                   \\
-        {6}                           & {2}     & \textcolor{purple}{34,97}           & {37,28\%}                           & \textcolor{purple}{39,52}                & \textcolor{purple}{29,69\%} & {40,53}                   & {24,7\%}                    \\
-        {6}                           & {10}    & {32,39}                             & {43,53\%}                           & \textcolor{purple}{38,61}                & \textcolor{purple}{28,26\%} & {40,84}                   & {21,74\%}                   \\
-        {6}                           & {16}    & \textcolor{purple}{34,87}           & \textcolor{purple}{34,38\%}         & \textcolor{purple}{39,03}                & \textcolor{purple}{26,82\%} & {40,34}                   & {23,31\%}                   \\
-        {6}                           & {50}    & \textcolor{purple}{33,85}           & {37,07\%}                           & \textcolor{purple}{39,39}                & \textcolor{purple}{27,17\%} & {40,8}                    & {22,09\%}                   \\
-        {6}                           & {100}   & \textcolor{purple}{34,62}           & {40,02\%}                           & \textcolor{purple}{39,46}                & \textcolor{purple}{25,48\%} & \textcolor{purple}{42,53} & \textcolor{purple}{18,4\%}  \\
+        {1}                           & {2}     & \textcolor{red}{34,91}              & \textcolor{red}{34,81\%}            & \textcolor{red}{40,61}                   & \textcolor{red}{25,48\%} & {39,13}                & {25\%}                   \\
+        {1}                           & {10}    & \textcolor{red}{35,31}              & {40,15\%}                           & \textcolor{red}{38,73}                   & \textcolor{red}{28,12\%} & \textcolor{red}{44,19} & {20,53\%}                \\
+        {1}                           & {16}    & \textcolor{red}{35,4}               & {42,93\%}                           & \textcolor{red}{39,58}                   & \textcolor{red}{26,17\%} & {39,64}                & {24,18\%}                \\
+        {1}                           & {50}    & {32,64}                             & {42,01\%}                           & \textcolor{red}{38,99}                   & \textcolor{red}{28,52\%} & \textcolor{red}{42,61} & {20,79\%}                \\
+        {1}                           & {100}   & \textcolor{red}{36,73}              & \textcolor{red}{31,21\%}            & \textcolor{red}{40,25}                   & \textcolor{red}{24,91\%} & \textcolor{red}{41,5}  & {21,44\%}                \\
+        {2}                           & {2}     & \textcolor{red}{33,76}              & {40,67\%}                           & {35,86}                                  & {31,73\%}                & \textcolor{red}{41,79} & {22,14\%}                \\
+        {2}                           & {10}    & {31,37}                             & {43,27\%}                           & {37,23}                                  & \textcolor{red}{30,34\%} & \textcolor{red}{42,53} & \textcolor{red}{19,75\%} \\
+        {2}                           & {16}    & {31,19}                             & {46,27\%}                           & \textcolor{red}{39,38}                   & \textcolor{red}{28,21\%} & \textcolor{red}{41,74} & {22,7\%}                 \\
+        {2}                           & {50}    & \textcolor{red}{33,66}              & {39,28\%}                           & \textcolor{red}{39,62}                   & \textcolor{red}{25,35\%} & \textcolor{red}{45,63} & \textcolor{red}{17,93\%} \\
+        {2}                           & {100}   & \textcolor{red}{35,1}               & \textcolor{red}{34,81\%}            & \textcolor{red}{39,86}                   & \textcolor{red}{27,3\%}  & \textcolor{red}{45,45} & \textcolor{red}{19,05\%} \\
+        {3}                           & {2}     & {32,99}                             & {41,15\%}                           & {37,76}                                  & \textcolor{red}{30,99\%} & \textcolor{red}{46,45} & {20,92\%}                \\
+        {3}                           & {10}    & \textcolor{red}{33,74}              & {38,89\%}                           & \textcolor{red}{40,71}                   & \textcolor{red}{26,48\%} & \textcolor{red}{44,53} & \textcolor{red}{19,01\%} \\
+        {3}                           & {16}    & \textcolor{red}{34,18}              & {38,98\%}                           & \textcolor{red}{38,52}                   & \textcolor{red}{30,03\%} & \textcolor{red}{41,22} & {21,53\%}                \\
+        {3}                           & {50}    & \textcolor{red}{34,38}              & {38,93\%}                           & \textcolor{red}{39,52}                   & \textcolor{red}{26,74\%} & {38,4}                 & {24,87\%}                \\
+        {3}                           & {100}   & \textcolor{red}{33,16}              & {40,19\%}                           & \textcolor{red}{39,75}                   & \textcolor{red}{27,78\%} & \textcolor{red}{41,29} & {21,83\%}                \\
+        {4}                           & {2}     & \textcolor{red}{33,48}              & {40,36\%}                           & \textcolor{red}{38,28}                   & \textcolor{red}{30,9\%}  & \textcolor{red}{42,53} & {22,4\%}                 \\
+        {4}                           & {10}    & \textcolor{red}{35,01}              & {39,97\%}                           & \textcolor{red}{40,88}                   & \textcolor{red}{26,39\%} & \textcolor{red}{42,33} & {20,44\%}                \\
+        {4}                           & {16}    & \textcolor{red}{35,59}              & \textcolor{red}{35,59\%}            & \textcolor{red}{37,91}                   & {31,86\%}                & \textcolor{red}{41,9}  & {20,27\%}                \\
+        {4}                           & {50}    & \textcolor{red}{34,99}              & {36,68\%}                           & \textcolor{red}{40,89}                   & \textcolor{red}{23,26\%} & \textcolor{red}{43,77} & \textcolor{red}{18,36\%} \\
+        {4}                           & {100}   & \textcolor{red}{35,46}              & {37,46\%}                           & \textcolor{red}{38,53}                   & \textcolor{red}{29,08\%} & {40,6}                 & {21,22\%}                \\
+        {6}                           & {2}     & \textcolor{red}{34,97}              & {37,28\%}                           & \textcolor{red}{39,52}                   & \textcolor{red}{29,69\%} & {40,53}                & {24,7\%}                 \\
+        {6}                           & {10}    & {32,39}                             & {43,53\%}                           & \textcolor{red}{38,61}                   & \textcolor{red}{28,26\%} & {40,84}                & {21,74\%}                \\
+        {6}                           & {16}    & \textcolor{red}{34,87}              & \textcolor{red}{34,38\%}            & \textcolor{red}{39,03}                   & \textcolor{red}{26,82\%} & {40,34}                & {23,31\%}                \\
+        {6}                           & {50}    & \textcolor{red}{33,85}              & {37,07\%}                           & \textcolor{red}{39,39}                   & \textcolor{red}{27,17\%} & {40,8}                 & {22,09\%}                \\
+        {6}                           & {100}   & \textcolor{red}{34,62}              & {40,02\%}                           & \textcolor{red}{39,46}                   & \textcolor{red}{25,48\%} & \textcolor{red}{42,53} & \textcolor{red}{18,4\%}  \\
         \bottomrule
@@
         \multicolumn{2}{c|}{baseline} & {62,16\%} & {49,61\%}                           & {41,68\%}                                                                      \\\midrule
-        {1}                           & {2}       & {93,3\%}                            & {65,36\%}                           & \textcolor{purple}{39,22\%}              \\
-        {1}                           & {10}      & {93,85\%}                           & \textcolor{purple}{46,97\%}         & \textcolor{purple}{40,1\%}               \\
-        {1}                           & {16}      & {92,06\%}                           & {76,68\%}                           & \textcolor{purple}{36,43\%}              \\
-        {1}                           & {50}      & {94\%}                              & {76,43\%}                           & \textcolor{purple}{39,97\%}              \\
-        {1}                           & {100}     & {94,03\%}                           & \textcolor{purple}{54,14\%}         & \textcolor{purple}{40,27\%}              \\
-        {2}                           & {2}       & {92,27\%}                           & \textcolor{purple}{52,15\%}         & \textcolor{purple}{33,64\%}              \\
-        {2}                           & {10}      & \textbf{96,16\%}                    & {80,26\%}                           & \textcolor{purple}{36,53\%}              \\
-        {2}                           & {16}      & {95,84\%}                           & \textbf{84,03\%}                    & \textcolor{purple}{39,65\%}              \\
-        {2}                           & {50}      & {93,78\%}                           & \textcolor{purple}{52,24\%}         & \textcolor{purple}{39,56\%}              \\
-        {2}                           & {100}     & {92,43\%}                           & \textcolor{purple}{53,23\%}         & \textcolor{purple}{37,68\%}              \\
-        {3}                           & {2}       & {94,52\%}                           & \textcolor{purple}{51,97\%}         & \textcolor{purple}{37,09\%}              \\
-        {3}                           & {10}      & {94,9\%}                            & \textcolor{purple}{53,47\%}         & \textcolor{purple}{38,24\%}              \\
+        {1}                           & {2}       & {93,3\%}                            & {65,36\%}                           & \textcolor{red}{39,22\%}                 \\
+        {1}                           & {10}      & {93,85\%}                           & \textcolor{red}{46,97\%}            & \textcolor{red}{40,1\%}                  \\
+        {1}                           & {16}      & {92,06\%}                           & {76,68\%}                           & \textcolor{red}{36,43\%}                 \\
+        {1}                           & {50}      & {94\%}                              & {76,43\%}                           & \textcolor{red}{39,97\%}                 \\
+        {1}                           & {100}     & {94,03\%}                           & \textcolor{red}{54,14\%}            & \textcolor{red}{40,27\%}                 \\
+        {2}                           & {2}       & {92,27\%}                           & \textcolor{red}{52,15\%}            & \textcolor{red}{33,64\%}                 \\
+        {2}                           & {10}      & \textbf{96,16\%}                    & {80,26\%}                           & \textcolor{red}{36,53\%}                 \\
+        {2}                           & {16}      & {95,84\%}                           & \textbf{84,03\%}                    & \textcolor{red}{39,65\%}                 \\
+        {2}                           & {50}      & {93,78\%}                           & \textcolor{red}{52,24\%}            & \textcolor{red}{39,56\%}                 \\
+        {2}                           & {100}     & {92,43\%}                           & \textcolor{red}{53,23\%}            & \textcolor{red}{37,68\%}                 \\
+        {3}                           & {2}       & {94,52\%}                           & \textcolor{red}{51,97\%}            & \textcolor{red}{37,09\%}                 \\
+        {3}                           & {10}      & {94,9\%}                            & \textcolor{red}{53,47\%}            & \textcolor{red}{38,24\%}                 \\
         {3}                           & {16}      & {94,59\%}                           & \textbf{81,46\%}                    & \textbf{67,88\%}                         \\
-        {3}                           & {50}      & {93,88\%}                           & {79,65\%}                           & \textcolor{purple}{40,36\%}              \\
-        {3}                           & {100}     & {95,25\%}                           & \textcolor{purple}{48,52\%}         & \textcolor{purple}{42,55\%}              \\
-        {4}                           & {2}       & {89,15\%}                           & \textcolor{purple}{51,98\%}         & \textcolor{purple}{39,68\%}              \\
-        {4}                           & {10}      & \textbf{96,08\%}                    & \textcolor{purple}{48,03\%}         & \textbf{64,31\%}                         \\
-        {4}                           & {16}      & {94,14\%}                           & \textcolor{purple}{49,81\%}         & \textcolor{purple}{40,84\%}              \\
-        {4}                           & {50}      & \textbf{96,24\%}                    & \textcolor{purple}{48,79\%}         & \textcolor{purple}{43,61\%}              \\
-        {4}                           & {100}     & {95,55\%}                           & \textcolor{purple}{49,65\%}         & \textcolor{purple}{42,85\%}              \\
-        {6}                           & {2}       & \textcolor{purple}{59,68\%}         & \textcolor{purple}{53,57\%}         & \textcolor{purple}{38,43\%}              \\
-        {6}                           & {10}      & \textcolor{purple}{63,46\%}         & \textbf{82,12\%}                    & \textcolor{purple}{40,11\%}              \\
-        {6}                           & {16}      & {95,86\%}                           & \textcolor{purple}{50,71\%}         & \textcolor{purple}{40,61\%}              \\
-        {6}                           & {50}      & {91,27\%}                           & \textcolor{purple}{52,55\%}         & \textcolor{purple}{40,21\%}              \\
-        {6}                           & {100}     & \textcolor{purple}{60,27\%}         & \textcolor{purple}{46,92\%}         & \textcolor{purple}{41,98\%}              \\
+        {3}                           & {50}      & {93,88\%}                           & {79,65\%}                           & \textcolor{red}{40,36\%}                 \\
+        {3}                           & {100}     & {95,25\%}                           & \textcolor{red}{48,52\%}            & \textcolor{red}{42,55\%}                 \\
+        {4}                           & {2}       & {89,15\%}                           & \textcolor{red}{51,98\%}            & \textcolor{red}{39,68\%}                 \\
+        {4}                           & {10}      & \textbf{96,08\%}                    & \textcolor{red}{48,03\%}            & \textbf{64,31\%}                         \\
+        {4}                           & {16}      & {94,14\%}                           & \textcolor{red}{49,81\%}            & \textcolor{red}{40,84\%}                 \\
+        {4}                           & {50}      & \textbf{96,24\%}                    & \textcolor{red}{48,79\%}            & \textcolor{red}{43,61\%}                 \\
+        {4}                           & {100}     & {95,55\%}                           & \textcolor{red}{49,65\%}            & \textcolor{red}{42,85\%}                 \\
+        {6}                           & {2}       & \textcolor{red}{59,68\%}            & \textcolor{red}{53,57\%}            & \textcolor{red}{38,43\%}                 \\
+        {6}                           & {10}      & \textcolor{red}{63,46\%}            & \textbf{82,12\%}                    & \textcolor{red}{40,11\%}                 \\
+        {6}                           & {16}      & {95,86\%}                           & \textcolor{red}{50,71\%}            & \textcolor{red}{40,61\%}                 \\
+        {6}                           & {50}      & {91,27\%}                           & \textcolor{red}{52,55\%}            & \textcolor{red}{40,21\%}                 \\
+        {6}                           & {100}     & \textcolor{red}{60,27\%}            & \textcolor{red}{46,92\%}            & \textcolor{red}{41,98\%}                 \\
         \bottomrule

### 2024-05-08 thesis/sections/7-language-analysis.tex
language analysis

@@
 \cmtDK[inline]{talks about why and which few languages}
+\cmtDK[inline]{chosen salience order might be due to number of possibilities (more efficient to exclude larger number of distractors instantly)}
 
@@
 However, without a deeper analysis, it can't be ruled out that the receiver extracts the \emph{shape} themselves from the image, and uses the sender's message only to determine the correct target object in the scene.
-Secondly, this does not apply to the \emph{size} and the \emph{color}.
-While the size seems to be communicated, it is not generated correctly by the receiver.
+Secondly, this does not apply directly to the \emph{size} and the \emph{color}.
+The \emph{size} seems to be communicated often by the sender to discriminate the target object.
+On the other hand, as the target referring expression uses the salience order 'Sh > C > Si', the shape is only needed to be generated in $\frac{1}{3} * \frac{1}{8} \approx 4,2\%$ of the samples.
 The opposite applies to the \emph{color}.
-The \emph{color} is not important in the communication, but the receiver however is able to produce correct colors in the generated referring expression.
+The \emph{color} is not important in the agents' communication, but the receiver however needs to produce colors in the generated referring expression in $\frac{1}{3} \approx 33,3\%$ of the samples.
 This shows that the agents actually don't align their own language with the target referring expression.
@@
 While the dominant salience order is 'Sh > Si > C' as on the 'Dale-2' dataset, the \emph{color} plays a bigger role.
-This effect is the same as for the discrimination games, and can be explained by the higher number or distractors.
+This effect is the same as for the discrimination games, and can be explained by the higher number of distractors.
 While the loss for the salience order 'Sh > Si > C' is lower than 70\% for two languages, it stays between 70\% and 80\% for most.
+Interesting now is that all attributes are communicated by the sender, while the receiver is also generating the correct attributes in the referring expression.
+Opposed to the 'Dale-2' dataset, the need for \emph{size} and \emph{color} is much bigger to discriminate the target object in any salience order, due to the larger number of distractors.
+
+$Lang_{6,50}$ is the emerged language, with which the agents can achieve the best results on the task.
+While the receiver achieves high precision and recall scores on each attribute (recall Tables \ref{tab:results:re-generator-game_size-shape} and \ref{tab:results:re-generator-game_color}), each attribute also plays a bigger role in the message be the sender.
+Indeed, one can see a correlation between the predictions of the receiver and the usage of attributes in emerged language.
+A more frequent usage of an attribute in the emerged language leads to a more accurate generation of this attribute in the final referring expression.
 

### 2024-05-12 thesis/sections/7-language-analysis.tex
language analysis
quantitative language analysis

--- 2024-04-30 thesis/sections/7-language-analysis.tex
@@
 \cmtDK[inline]{talks about why and which few languages}
+\cmtDK[inline]{chosen salience order might be due to number of possibilities (more efficient to exclude larger number of distractors instantly)}
 
-To compare the emerged language to English in a \textbf{quantitative} way, a probing approach is used.
+To compare the emerged language to English \textbf{quantitatively}, a probing approach is used.
 Probing can be used to analyze and interpret the hidden representations in neural networks.
@@
 First, it can learn to find correlations between the emerged language and the English referring expressions, which is the goal of the probing.
-But secondly, it can also learn underlying patterns in the English referring expressions that are independent of the emerged language.
+But secondly, it can as well learn underlying patterns in the English referring expressions that are independent of the emerged language.
 For instance, it can learn that the referring expressions are likely to be two symbols long, or that for instance 'cube' is the most common shape.
@@
 
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{95,5\%}           & \textcolor{red}{97,57\%}          & {86,98\%}                & {81,43\%}                & \textcolor{red}{91,67\%} & {86,45\%}                \\
-        {1} & {10}  & \textcolor{red}{96,11\%}          & \textcolor{red}{96,9\%}           & {89,61\%}                & {85\%}                   & {89,4\%}                 & {84,49\%}                \\
-        {1} & {16}  & \textcolor{red}{95,19\%}          & \textcolor{red}{96,56\%}          & {87,25\%}                & {81,22\%}                & {89,34\%}                & {83,13\%}                \\
-        {1} & {50}  & \textcolor{red}{96,61\%}          & \textcolor{red}{97,07\%}          & \textcolor{red}{91,26\%} & {85,75\%}                & {87,14\%}                & {82,61\%}                \\
-        {1} & {100} & \textcolor{red}{95,91\%}          & \textcolor{red}{96,73\%}          & \textcolor{red}{90,63\%} & {86,01\%}                & {88,95\%}                & {84,17\%}                \\
-        {2} & {2}   & \textcolor{red}{99,67\%}          & \textcolor{red}{99,91\%}          & \textcolor{red}{97,02\%} & \textcolor{red}{95,37\%} & \textcolor{red}{96,97\%} & \textcolor{red}{94,64\%} \\
-        {2} & {10}  & \textcolor{red}{96,36\%}          & \textcolor{red}{96,34\%}          & \textcolor{red}{91,32\%} & {86,75\%}                & {87,69\%}                & {82,87\%}                \\
-        {2} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{97,12\%}          & {88,11\%}                & {83,21\%}                & {89,86\%}                & {85,35\%}                \\
-        {2} & {50}  & \textcolor{red}{95,34\%}          & \textcolor{red}{97,3\%}           & {85,48\%}                & {80,89\%}                & \textcolor{red}{92,22\%} & {86,18\%}                \\
-        {2} & {100} & \textcolor{red}{94,88\%}          & \textcolor{red}{96,32\%}          & {87,37\%}                & {81,17\%}                & {88,35\%}                & {82,61\%}                \\
-        {3} & {2}   & \textcolor{red}{\textbf{94,03\%}} & \textcolor{red}{96,96\%}          & \textbf{79,59\%}         & \textbf{74,54\%}         & \textcolor{red}{92,32\%} & {85,63\%}                \\
-        {3} & {10}  & \textcolor{red}{96,1\%}           & \textcolor{red}{95,97\%}          & \textcolor{red}{90,45\%} & {84,84\%}                & {85,98\%}                & {80,5\%}                 \\
-        {3} & {16}  & \textcolor{red}{95,71\%}          & \textcolor{red}{\textbf{95,1\%}}  & \textcolor{red}{90,97\%} & {85,33\%}                & {81\%}                   & \textbf{75,56\%}         \\
-        {3} & {50}  & \textcolor{red}{96,32\%}          & \textcolor{red}{\textbf{94,55\%}} & \textcolor{red}{92,5\%}  & {86,81\%}                & \textbf{80,48\%}         & {75,73\%}                \\
-        {3} & {100} & \textcolor{red}{95,33\%}          & \textcolor{red}{97,82\%}          & {86,05\%}                & {81,33\%}                & \textcolor{red}{93,91\%} & {88,54\%}                \\
-        {4} & {2}   & \textcolor{red}{96,81\%}          & \textcolor{red}{95,12\%}          & \textcolor{red}{93,03\%} & {87,41\%}                & \textbf{80,73\%}         & \textbf{75,47\%}         \\
-        {4} & {10}  & \textcolor{red}{95,31\%}          & \textcolor{red}{97,26\%}          & {86,9\%}                 & {82,67\%}                & \textcolor{red}{92,39\%} & {86,98\%}                \\
-        {4} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{\textbf{94,65\%}} & \textcolor{red}{91,99\%} & {86,65\%}                & \textbf{79,23\%}         & \textbf{74,54\%}         \\
-        {4} & {50}  & \textcolor{red}{94,76\%}          & \textcolor{red}{96,69\%}          & {86,8\%}                 & {81,33\%}                & {89,65\%}                & {83,92\%}                \\
-        {4} & {100} & \textcolor{red}{96,16\%}          & \textcolor{red}{97,94\%}          & {89,57\%}                & {85,18\%}                & \textcolor{red}{92,56\%} & {87,74\%}                \\
-        {6} & {2}   & \textcolor{red}{95,73\%}          & \textcolor{red}{96,99\%}          & {87,77\%}                & {83,27\%}                & \textcolor{red}{91,77\%} & {86,34\%}                \\
-        {6} & {10}  & \textcolor{red}{\textbf{94,59\%}} & \textcolor{red}{99,89\%}          & {86,24\%}                & {81,46\%}                & \textcolor{red}{91,42\%} & {86,43\%}                \\
-        {6} & {16}  & \textcolor{red}{95,51\%}          & \textcolor{red}{96,3\%}           & {88,89\%}                & {83,24\%}                & {87,61\%}                & {82,71\%}                \\
-        {6} & {50}  & \textcolor{red}{\textbf{94,66\%}} & \textcolor{red}{97,43\%}          & \textbf{84,93\%}         & \textbf{78,61\%}         & \textcolor{red}{91,48\%} & {83,66\%}                \\
-        {6} & {100} & \textcolor{red}{95,37\%}          & \textcolor{red}{97,77\%}          & \textbf{83,75\%}         & \textbf{80,84\%}         & \textcolor{red}{93,96\%} & {89,02\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{Discriminator: Dale-2}
-    \label{tab:probing:discriminator:dale-2}
-\end{table}
+In the following paragraphs, languages will be referred to by their message length $n$ and their vocabulary size $|V|$ as $Lang_{n,|V|}$.
 
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
-        {1} & {2}   & \textcolor{red}{94,96\%} & \textcolor{red}{91,27\%} & \textcolor{red}{94,51\%} & \textcolor{red}{90,99\%} & {88,52\%}            & {88,17\%}            \\
-        {1} & {10}  & {84,34\%}                & {84,03\%}                & {80,91\%}                & {73,23\%}                & {79,95\%}            & {72,34\%}            \\
-        {1} & {16}  & {86,27\%}                & {84,96\%}                & {82,76\%}                & {75,35\%}                & {80,72\%}            & {74,04\%}            \\
-        {1} & {50}  & {85,95\%}                & {84,4\%}                 & {82,96\%}                & {74,97\%}                & {80,2\%}             & {73,54\%}            \\
-        {1} & {100} & {86,31\%}                & {85,58\%}                & {82,9\%}                 & {76,03\%}                & {81,64\%}            & {74,62\%}            \\
-        {2} & {10}  & \textbf{82,66\%}         & {83,23\%}                & \textbf{78,51\%}         & {71,01\%}                & {78,89\%}            & {70,61\%}            \\
-        {2} & {16}  & {89,78\%}                & {85,53\%}                & {87,68\%}                & {80,64\%}                & {81,06\%}            & {77,24\%}            \\
-        {2} & {50}  & {84,4\%}                 & {82,65\%}                & {80,62\%}                & {71,42\%}                & {77,53\%}            & {69,46\%}            \\
-        {2} & {100} & {89,65\%}                & {85,3\%}                 & {87,26\%}                & {79,91\%}                & {80,88\%}            & {77,1\%}             \\
-        {3} & {2}   & {87,19\%}                & {86,93\%}                & {83,99\%}                & {78,18\%}                & {83,65\%}            & {77,08\%}            \\
-        {3} & {10}  & {87,15\%}                & {84,67\%}                & {84,01\%}                & {76,24\%}                & {80,41\%}            & {74,11\%}            \\
-        {3} & {50}  & \textbf{82,98\%}         & {82,25\%}                & {78,85\%}                & {70,62\%}                & {77,42\%}            & {69,27\%}            \\
-        {3} & {100} & {83,78\%}                & {83,07\%}                & {80,23\%}                & {72,15\%}                & {78,94\%}            & {71,1\%}             \\
-        {4} & {2}   & {89,23\%}                & {87,43\%}                & {87,09\%}                & {81,45\%}                & {83,83\%}            & {79,11\%}            \\
-        {4} & {10}  & {88,66\%}                & {84,83\%}                & {85,9\%}                 & {78,98\%}                & {80,2\%}             & {75,64\%}            \\
-        {4} & {16}  & {83,14\%}                & \textbf{81,51\%}         & \textbf{78,15\%}         & \textbf{70,21\%}         & \textbf{77,32\%}     & \textbf{68,48\%}     \\
-        {4} & {50}  & {87,62\%}                & {84,38\%}                & {85,25\%}                & {77,29\%}                & {79,94\%}            & {74,03\%}            \\
-        {4} & {100} & \textcolor{red}{90,17\%} & {85,76\%}                & {87,68\%}                & {80,71\%}                & {81,37\%}            & {78,1\%}             \\
-        {6} & {10}  & \textbf{82,03\%}         & \textbf{80,51\%}         & \textbf{77,46\%}         & \textbf{68,63\%}         & \textbf{75,71\%}     & \textbf{67,15\%}     \\
-        {6} & {16}  & {84,73\%}                & \textbf{81,12\%}         & {80,99\%}                & \textbf{70,5\%}          & \textbf{76,06\%}     & \textbf{69,12\%}     \\
-        {6} & {50}  & {86,95\%}                & {85,13\%}                & {84,06\%}                & {76,95\%}                & {80,11\%}            & {74,51\%}            \\
-        {6} & {100} & {86,15\%}                & {82,19\%}                & {83,56\%}                & {75,56\%}                & {77,76\%}            & {72,47\%}            \\
-        \bottomrule
-    \end{tabular}
-    \caption{Discriminator: Dale-5}
-    \label{tab:probing:discriminator:dale-5}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{91,62\%} & \textcolor{red}{91,57\%} & \textcolor{red}{94,31\%} & \textcolor{red}{95,13\%} & \textcolor{red}{93,11\%} & \textcolor{red}{94,92\%} \\
-        {1} & {10}  & {79,08\%}                & {79,08\%}                & {85,04\%}                & {86,86\%}                & {82,37\%}                & {86,38\%}                \\
-        {1} & {16}  & {84,02\%}                & {83,96\%}                & {88,67\%}                & {89,65\%}                & {86,36\%}                & {89,36\%}                \\
-        {1} & {50}  & \textcolor{red}{93,88\%} & \textcolor{red}{93,74\%} & \textcolor{red}{95,03\%} & \textcolor{red}{95,49\%} & \textcolor{red}{94,64\%} & \textcolor{red}{95,35\%} \\
-        {2} & {10}  & {88,67\%}                & {88,36\%}                & {89,59\%}                & \textcolor{red}{90,77\%} & \textcolor{red}{90,01\%} & {89,81\%}                \\
-        {2} & {16}  & {81,13\%}                & {81,02\%}                & {86,16\%}                & {87,97\%}                & {83,85\%}                & {87,56\%}                \\
-        {3} & {2}   & {83,38\%}                & {83,46\%}                & {88,21\%}                & {89,71\%}                & {86,05\%}                & {89,37\%}                \\
-        {3} & {10}  & \textbf{77\%}            & \textbf{76,77\%}         & \textbf{83,54\%}         & \textbf{85,57\%}         & \textbf{79,97\%}         & \textbf{84,9\%}          \\
-        {3} & {16}  & \textcolor{red}{93,52\%} & \textcolor{red}{93,6\%}  & \textcolor{red}{94,16\%} & \textcolor{red}{94,14\%} & \textcolor{red}{93,75\%} & \textcolor{red}{93,55\%} \\
-        {4} & {10}  & \textbf{68,04\%}         & \textbf{66,51\%}         & \textbf{76,33\%}         & \textbf{78,56\%}         & \textbf{71,42\%}         & \textbf{78,56\%}         \\
-        {4} & {16}  & \textbf{75,1\%}          & \textbf{74,93\%}         & \textbf{81,2\%}          & \textbf{83,35\%}         & \textbf{77,6\%}          & \textbf{82,76\%}         \\
-        {6} & {2}   & {84,21\%}                & {84,02\%}                & {88,41\%}                & {89,77\%}                & {86,61\%}                & {89,35\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{Discriminator: Color}
-    \label{tab:probing:discriminator:colour}
-\end{table}
-
-Looking at the results for the discriminator, one can see that the results are very different for each dataset.
-On the 'Dale-2' dataset
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{95,07\%} & \textcolor{red}{98,06\%}          & {86,63\%}                & {85,3\%}                 & \textcolor{red}{96,26\%} & \textcolor{red}{93,32\%} \\
-        {1} & {10}  & \textcolor{red}{90,8\%}  & \textcolor{red}{96,4\%}           & {71,88\%}                & {69,29\%}                & \textcolor{red}{94,05\%} & {85,13\%}                \\
-        {1} & {16}  & \textcolor{red}{91,47\%} & \textcolor{red}{96,96\%}          & {71,2\%}                 & {66,82\%}                & \textcolor{red}{93,83\%} & {83,35\%}                \\
-        {1} & {50}  & \textcolor{red}{93,26\%} & \textcolor{red}{96,63\%}          & {77,53\%}                & {71,94\%}                & \textcolor{red}{93,56\%} & {84,62\%}                \\
-        {1} & {100} & \textcolor{red}{93,5\%}  & \textcolor{red}{97,86\%}          & {77,42\%}                & {71,93\%}                & \textcolor{red}{94,68\%} & {85,77\%}                \\
-        {2} & {2}   & \textcolor{red}{93,57\%} & \textcolor{red}{97,72\%}          & {80,89\%}                & {79,8\%}                 & \textcolor{red}{96,08\%} & \textcolor{red}{90,71\%} \\
-        {2} & {10}  & \textbf{84,86\%}         & \textcolor{red}{\textbf{92,07\%}} & {58,8\%}                 & \textbf{52,09\%}         & \textbf{88,36\%}         & {77,01\%}                \\
-        {2} & {16}  & {87,29\%}                & \textcolor{red}{92,88\%}          & {68,52\%}                & {71,5\%}                 & \textcolor{red}{92,56\%} & {83,9\%}                 \\
-        {2} & {50}  & {89,15\%}                & \textcolor{red}{95\%}             & {68,19\%}                & {68,93\%}                & \textcolor{red}{93,11\%} & {83,95\%}                \\
-        {2} & {100} & \textcolor{red}{91,06\%} & \textcolor{red}{96,14\%}          & {71,98\%}                & {67,23\%}                & \textcolor{red}{92,81\%} & {82,71\%}                \\
-        {3} & {2}   & \textcolor{red}{93,05\%} & \textcolor{red}{97,15\%}          & {82,41\%}                & {77,36\%}                & \textcolor{red}{94,43\%} & {88,8\%}                 \\
-        {3} & {10}  & {88,52\%}                & \textcolor{red}{95,41\%}          & {66,61\%}                & {63,63\%}                & \textcolor{red}{92,25\%} & {81,46\%}                \\
-        {3} & {16}  & {85,71\%}                & \textcolor{red}{92,88\%}          & \textbf{56,44\%}         & \textbf{49,22\%}         & \textbf{88,87\%}         & \textbf{74,54\%}         \\
-        {3} & {50}  & \textcolor{red}{90,78\%} & \textcolor{red}{96,28\%}          & {68,17\%}                & {64,84\%}                & \textcolor{red}{94,16\%} & {83,5\%}                 \\
-        {3} & {100} & {86,54\%}                & \textcolor{red}{93,54\%}          & {65,9\%}                 & {66,03\%}                & \textcolor{red}{92,49\%} & {83,28\%}                \\
-        {4} & {10}  & {87,61\%}                & \textcolor{red}{94,63\%}          & {66,83\%}                & {64,85\%}                & \textcolor{red}{92,24\%} & {82,85\%}                \\
-        {4} & {16}  & \textbf{85,7\%}          & \textcolor{red}{93,56\%}          & \textbf{56,9\%}          & \textbf{52,1\%}          & \textbf{88,98\%}         & \textbf{76,15\%}         \\
-        {4} & {50}  & {87,55\%}                & \textcolor{red}{94,24\%}          & {61,03\%}                & {55,82\%}                & \textcolor{red}{91,73\%} & \textbf{76,52\%}         \\
-        {4} & {100} & {88,84\%}                & \textcolor{red}{94,84\%}          & {68,68\%}                & {66,73\%}                & \textcolor{red}{92,57\%} & {81,79\%}                \\
-        {6} & {2}   & \textcolor{red}{93,48\%} & \textcolor{red}{98,14\%}          & {81,36\%}                & {79,08\%}                & \textcolor{red}{96,09\%} & \textcolor{red}{90,95\%} \\
-        {6} & {10}  & \textbf{85\%}            & \textcolor{red}{\textbf{92,44\%}} & \textbf{57,21\%}         & {55,07\%}                & {89,39\%}                & {77,44\%}                \\
-        {6} & {16}  & {88,02\%}                & \textcolor{red}{94,22\%}          & {71,42\%}                & {66,35\%}                & \textcolor{red}{90,86\%} & {84,12\%}                \\
-        {6} & {50}  & \textcolor{red}{98,57\%} & \textcolor{red}{99,45\%}          & \textcolor{red}{95,57\%} & \textcolor{red}{95,06\%} & \textcolor{red}{99,16\%} & \textcolor{red}{96,64\%} \\
-        {6} & {100} & {86,44\%}                & \textcolor{red}{\textbf{92,76\%}} & {65,59\%}                & {65,93\%}                & \textcolor{red}{92,42\%} & {82,16\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{RE generator: Dale-2}
-    \label{tab:probing:re-generator:dale-2}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{92,86\%} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,06\%} & {89,5\%}             & \textcolor{red}{95,73\%} & \textcolor{red}{90,93\%} \\
-        {1} & {10}  & {83,9\%}                 & \textcolor{red}{90,9\%}  & {78,63\%}                & {74,74\%}            & {89,6\%}                 & {78,99\%}                \\
-        {1} & {16}  & {84,68\%}                & \textcolor{red}{91,65\%} & {80,11\%}                & {77,03\%}            & \textcolor{red}{90,16\%} & {81,05\%}                \\
-        {1} & {50}  & {84,85\%}                & \textcolor{red}{91,67\%} & {79,95\%}                & {77,11\%}            & \textcolor{red}{90,32\%} & {81,04\%}                \\
-        {2} & {2}   & \textcolor{red}{90,51\%} & \textcolor{red}{95,36\%} & {87,5\%}                 & {86,43\%}            & \textcolor{red}{94,77\%} & {88,35\%}                \\
-        {2} & {10}  & {82,44\%}                & \textcolor{red}{90,32\%} & {77,65\%}                & {73,73\%}            & {88,89\%}                & {78,41\%}                \\
-        {2} & {16}  & {81,85\%}                & \textbf{89,4\%}          & {76,07\%}                & \textbf{71,66\%}     & {87,9\%}                 & \textbf{75,91\%}         \\
-        {2} & {50}  & {83,58\%}                & \textcolor{red}{91,12\%} & {78,67\%}                & {75,03\%}            & {89,94\%}                & {79,06\%}                \\
-        {3} & {2}   & \textcolor{red}{92,51\%} & \textcolor{red}{95,96\%} & {89,96\%}                & {88,7\%}             & \textcolor{red}{95,18\%} & \textcolor{red}{90,43\%} \\
-        {3} & {10}  & {82,11\%}                & {89,54\%}                & {75,9\%}                 & {71,78\%}            & {88,18\%}                & {76,72\%}                \\
-        {3} & {50}  & {83,12\%}                & \textcolor{red}{90,82\%} & {77,62\%}                & {74,08\%}            & {88,88\%}                & {78,74\%}                \\
-        {3} & {100} & {83,74\%}                & {89,59\%}                & {79,29\%}                & {74,48\%}            & \textbf{87,46\%}         & {78,05\%}                \\
-        {4} & {2}   & {88,86\%}                & \textcolor{red}{94\%}    & {84,8\%}                 & {83,19\%}            & \textcolor{red}{93,37\%} & {85,92\%}                \\
-        {4} & {10}  & \textbf{79,93\%}         & \textbf{89,29\%}         & \textbf{74,1\%}          & \textbf{69,5\%}      & \textbf{87,19\%}         & \textbf{74,79\%}         \\
-        {4} & {16}  & \textbf{81,26\%}         & {89,68\%}                & \textbf{75,73\%}         & {71,96\%}            & {88,4\%}                 & {76,93\%}                \\
-        {4} & {50}  & {81,55\%}                & {89,42\%}                & {75,9\%}                 & {71,68\%}            & {88,17\%}                & {76,55\%}                \\
-        {6} & {2}   & \textcolor{red}{91,09\%} & \textcolor{red}{95,26\%} & {87,86\%}                & {86,81\%}            & \textcolor{red}{94,78\%} & {88,82\%}                \\
-        {6} & {50}  & \textbf{80,39\%}         & \textbf{87,85\%}         & \textbf{75,01\%}         & \textbf{68,71\%}     & \textbf{84,54\%}         & \textbf{72,93\%}         \\
-        \bottomrule
-    \end{tabular}
-    \caption{RE generator: Dale-5}
-    \label{tab:probing:re-generator:dale-5}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}      & \textbf{Si > Sh > C}      \\\midrule
-        {1} & {2}   & \textcolor{red}{95,73\%} & \textcolor{red}{94,23\%} & \textcolor{red}{92,96\%} & \textcolor{red}{90,59\%} & {81,59\%}                 & {78,21\%}                 \\
-        {1} & {10}  & \textcolor{red}{91,77\%} & \textcolor{red}{91,19\%} & {73,14\%}                & {63,99\%}                & {67,45\%}                 & {57,55\%}                 \\
-        {1} & {16}  & \textcolor{red}{91,74\%} & \textcolor{red}{90,87\%} & {73,07\%}                & {63,78\%}                & {66,16\%}                 & {55,48\%}                 \\
-        {1} & {50}  & \textcolor{red}{97,78\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,22\%} & {87,45\%}                & {69,23\%}                 & {64,97\%}                 \\
-        {1} & {100} & \textcolor{red}{97,91\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,18\%} & {87,35\%}                & {69,25\%}                 & {64,99\%}                 \\
-        {2} & {2}   & \textcolor{red}{93,17\%} & \textcolor{red}{91,69\%} & {81,96\%}                & {75,97\%}                & {72,05\%}                 & {64,29\%}                 \\
-        {2} & {10}  & \textcolor{red}{94,07\%} & \textcolor{red}{98,04\%} & {75,32\%}                & {71,91\%}                & \textcolor{red}{94,38\%}  & {86,62\%}                 \\
-        {2} & {16}  & \textcolor{red}{97,72\%} & \textcolor{red}{92,9\%}  & \textcolor{red}{94,18\%} & {87,09\%}                & {68,54\%}                 & {64,13\%}                 \\
-        {2} & {50}  & \textcolor{red}{97,87\%} & \textcolor{red}{93,08\%} & \textcolor{red}{94,25\%} & {87,54\%}                & {69,2\%}                  & {65,17\%}                 \\
-        {2} & {100} & {89,6\%}                 & \textbf{89,58\%}         & \textbf{65,65\%}         & \textbf{55,69\%}         & {64,92\%}                 & {53,53\%}                 \\
-        {3} & {2}   & \textcolor{red}{97,73\%} & \textcolor{red}{99,28\%} & \textcolor{red}{95,84\%} & \textcolor{red}{97,72\%} & \textcolor{red}{101,58\%} & \textcolor{red}{100,22\%} \\
-        {3} & {10}  & \textcolor{red}{91,04\%} & \textcolor{red}{90,52\%} & {71,06\%}                & {61,36\%}                & {65,59\%}                 & {54,93\%}                 \\
-        {3} & {16}  & \textcolor{red}{91,09\%} & \textcolor{red}{90,3\%}  & {71,71\%}                & {62,42\%}                & {66,28\%}                 & {56,16\%}                 \\
-        {3} & {50}  & \textcolor{red}{97,07\%} & \textcolor{red}{91,72\%} & \textcolor{red}{92,7\%}  & {85,33\%}                & {66,38\%}                 & {61,26\%}                 \\
-        {3} & {100} & \textcolor{red}{97,81\%} & \textcolor{red}{93,01\%} & \textcolor{red}{94,16\%} & {87,3\%}                 & {68,83\%}                 & {64,6\%}                  \\
-        {4} & {2}   & \textcolor{red}{91,95\%} & \textcolor{red}{96,87\%} & {73,55\%}                & {72,66\%}                & \textcolor{red}{98,37\%}  & \textcolor{red}{90,23\%}  \\
-        {4} & {10}  & \textcolor{red}{97,7\%}  & \textcolor{red}{93,1\%}  & \textcolor{red}{94,16\%} & {87,24\%}                & {69,22\%}                 & {64,81\%}                 \\
-        {4} & {16}  & \textbf{86,75\%}         & \textbf{89\%}            & \textbf{57,02\%}         & \textbf{44,33\%}         & \textbf{63,78\%}          & \textbf{48,21\%}          \\
-        {4} & {50}  & \textcolor{red}{90,86\%} & \textcolor{red}{90,5\%}  & {71,22\%}                & {61,09\%}                & {65,26\%}                 & {54,87\%}                 \\
-        {4} & {100} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,22\%} & \textcolor{red}{92,44\%} & {83,78\%}                & \textbf{59,98\%}          & {54,38\%}                 \\
-        {6} & {2}   & \textcolor{red}{90,06\%} & \textcolor{red}{90,46\%} & {71,62\%}                & {64,56\%}                & {70,67\%}                 & {61,35\%}                 \\
-        {6} & {10}  & \textbf{88,38\%}         & \textcolor{red}{90,24\%} & \textbf{63,68\%}         & \textbf{53,21\%}         & \textbf{64,22\%}          & \textbf{52,19\%}          \\
-        {6} & {16}  & \textbf{89,01\%}         & \textbf{89,74\%}         & {67,03\%}                & {56,73\%}                & {64,44\%}                 & \textbf{52,78\%}          \\
-        {6} & {50}  & \textcolor{red}{97,68\%} & \textcolor{red}{92,77\%} & \textcolor{red}{97,03\%} & {87,24\%}                & {68,51\%}                 & {64,27\%}                 \\
-        {6} & {100} & {89,84\%}                & {89,81\%}                & {69,75\%}                & {60,55\%}                & {65,19\%}                 & {54,93\%}                 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attribute generator: Dale-2}
-    \label{tab:probing:attribute-generator:dale-2}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{91,81\%} & \textcolor{red}{96,22\%} & {89,08\%}                & {88,2\%}                 & \textcolor{red}{95,91\%} & \textcolor{red}{90,3\%}  \\
-        {1} & {10}  & {83,4\%}                 & {82,11\%}                & {79,68\%}                & {67,88\%}                & {77,12\%}                & {66,75\%}                \\
-        {1} & {16}  & {86,76\%}                & {87,06\%}                & {83,08\%}                & {75,03\%}                & {83,79\%}                & {74,72\%}                \\
-        {1} & {50}  & {83,12\%}                & {82\%}                   & {78,81\%}                & {66,76\%}                & {76,95\%}                & {65,95\%}                \\
-        {1} & {100} & \textcolor{red}{93,84\%} & {87,09\%}                & \textcolor{red}{93,01\%} & {83,36\%}                & {82,19\%}                & {79,14\%}                \\
-        {2} & {2}   & {88,86\%}                & {86,82\%}                & {86,62\%}                & {78,45\%}                & {82,59\%}                & {76,49\%}                \\
-        {2} & {10}  & {83,91\%}                & {81,57\%}                & {79,76\%}                & {67,89\%}                & {76,47\%}                & {66,1\%}                 \\
-        {2} & {16}  & \textbf{75,6\%}          & \textbf{77,74\%}         & \textbf{68,78\%}         & \textbf{55,43\%}         & \textbf{72,24\%}         & \textbf{56,24\%}         \\
-        {2} & {50}  & {85,1\%}                 & {83,08\%}                & {81,27\%}                & {70,13\%}                & {77,86\%}                & {68,54\%}                \\
-        {2} & {100} & {83,24\%}                & {81,87\%}                & {79,02\%}                & {67,32\%}                & {77,06\%}                & {66,19\%}                \\
-        {3} & {2}   & \textcolor{red}{97,21\%} & \textcolor{red}{96,08\%} & \textcolor{red}{96,43\%} & \textcolor{red}{95,23\%} & \textcolor{red}{95,18\%} & \textcolor{red}{94,24\%} \\
-        {3} & {10}  & {82,66\%}                & {81,11\%}                & {78,38\%}                & {66,21\%}                & {75,68\%}                & {64,52\%}                \\
-        {3} & {16}  & \textcolor{red}{93,26\%} & {86,09\%}                & \textcolor{red}{92,67\%} & {82,25\%}                & {81,17\%}                & {77,84\%}                \\
-        {3} & {50}  & {82,83\%}                & {81,44\%}                & {78,57\%}                & {66,29\%}                & {76,04\%}                & {64,84\%}                \\
-        {3} & {100} & {82,74\%}                & {81\%}                   & {78,39\%}                & {66,04\%}                & {75,44\%}                & {64,91\%}                \\
-        {4} & {2}   & \textcolor{red}{92,91\%} & {87,03\%}                & \textcolor{red}{92,16\%} & {84,06\%}                & {82,45\%}                & {79,57\%}                \\
-        {4} & {10}  & \textbf{79,3\%}          & \textbf{79,42\%}         & \textbf{74,65\%}         & \textbf{61,71\%}         & \textbf{74,12\%}         & \textbf{60,62\%}         \\
-        {4} & {16}  & {82,55\%}                & {81,31\%}                & {78,61\%}                & {66,51\%}                & {75,61\%}                & {65,05\%}                \\
-        {4} & {50}  & {88,51\%}                & {84,17\%}                & {85,77\%}                & {75,07\%}                & {79\%}                   & {71,93\%}                \\
-        {4} & {100} & {82,57\%}                & {81,61\%}                & {78,06\%}                & {66,33\%}                & {76,32\%}                & {65,54\%}                \\
-        {6} & {2}   & \textcolor{red}{92,36\%} & {88,15\%}                & \textcolor{red}{91,49\%} & {84,38\%}                & {85,28\%}                & {80,85\%}                \\
-        {6} & {10}  & \textbf{78,11\%}         & \textbf{79,19\%}         & \textbf{72,81\%}         & \textbf{59,86\%}         & \textbf{73,95\%}         & \textbf{60,17\%}         \\
-        {6} & {16}  & {82,35\%}                & {81,09\%}                & {78,84\%}                & {66,02\%}                & {75,9\%}                 & {64,79\%}                \\
-        {6} & {50}  & {82,04\%}                & {81,49\%}                & {78,34\%}                & {66,25\%}                & {76,46\%}                & {64,91\%}                \\
-        {6} & {100} & {82,09\%}                & {80,73\%}                & {77,42\%}                & {65,3\%}                 & {75,52\%}                & {64,52\%}                \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attribute generator: Dale-5}
-    \label{tab:probing:attribute-generator:dale-5}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si} & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C} \\\midrule
-        {1} & {2}   & \textcolor{red}{99,92\%} & \textcolor{red}{99,88\%} & \textbf{79,83\%}     & \textbf{82,7\%}      & \textcolor{red}{99,71\%} & \textbf{82,14\%}     \\
-        {1} & {10}  & \textbf{70,52\%}         & \textbf{70,36\%}         & \textbf{80,23\%}     & \textbf{82,36\%}     & \textbf{74,49\%}         & \textbf{81,77\%}     \\
-        {1} & {16}  & \textbf{77,25\%}         & \textbf{77,16\%}         & {84,55\%}            & {86,2\%}             & \textbf{80,72\%}         & {85,56\%}            \\
-        {3} & {10}  & \textbf{71,46\%}         & \textbf{71,47\%}         & \textbf{78,81\%}     & \textbf{81,87\%}     & \textbf{75,95\%}         & \textbf{80,93\%}     \\
-        {6} & {2}   & {83,25\%}                & {83,23\%}                & {87,74\%}            & {88,95\%}            & {85,38\%}                & {88,87\%}            \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attribute generator: Color}
-    \label{tab:probing:attribute-generator:colour}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {1} & {2}   & \textcolor{red}{95,17\%}          & \textcolor{red}{97,23\%}          & {87,74\%}                & {82,8\%}                 & \textcolor{red}{94,4\%}  & {88,75\%}                \\
-        {1} & {10}  & \textcolor{red}{96,04\%}          & \textcolor{red}{94,67\%}          & \textcolor{red}{94,06\%} & {88,67\%}                & {82,86\%}                & {79,71\%}                \\
-        {1} & {16}  & \textcolor{red}{93,82\%}          & \textcolor{red}{95,48\%}          & {85,33\%}                & {79,96\%}                & \textcolor{red}{90,17\%} & {84,6\%}                 \\
-        {1} & {50}  & \textcolor{red}{94,47\%}          & \textcolor{red}{96,65\%}          & {86,98\%}                & {81,89\%}                & \textcolor{red}{90,93\%} & {86,51\%}                \\
-        {1} & {100} & \textcolor{red}{95,6\%}           & \textcolor{red}{98,19\%}          & \textcolor{red}{90,41\%} & {86,13\%}                & \textcolor{red}{95,74\%} & \textcolor{red}{93,76\%} \\
-        {2} & {2}   & \textcolor{red}{94,75\%}          & \textcolor{red}{96,65\%}          & {86,51\%}                & {80,84\%}                & \textcolor{red}{92,99\%} & {86,3\%}                 \\
-        {2} & {10}  & \textcolor{red}{94,69\%}          & \textcolor{red}{98\%}             & {86,92\%}                & {82,2\%}                 & \textcolor{red}{96,31\%} & \textcolor{red}{92,55\%} \\
-        {2} & {16}  & \textcolor{red}{95,04\%}          & \textcolor{red}{96,39\%}          & {87,82\%}                & {82,54\%}                & \textcolor{red}{90,91\%} & {87,2\%}                 \\
-        {2} & {50}  & \textcolor{red}{97,18\%}          & \textcolor{red}{93,99\%}          & \textcolor{red}{96,42\%} & \textcolor{red}{92,26\%} & \textbf{77,97\%}         & {75,58\%}                \\
-        {2} & {100} & \textcolor{red}{93,71\%}          & \textcolor{red}{94,65\%}          & {85,56\%}                & {78,95\%}                & {83,77\%}                & {78,69\%}                \\
-        {3} & {2}   & \textcolor{red}{95,24\%}          & \textcolor{red}{97,37\%}          & {87,56\%}                & {82,49\%}                & \textcolor{red}{94,38\%} & {88,87\%}                \\
-        {3} & {10}  & \textcolor{red}{96,62\%}          & \textcolor{red}{\textbf{91,61\%}} & \textcolor{red}{94,87\%} & {89\%}                   & \textbf{69,2\%}          & \textbf{65,45\%}         \\
-        {3} & {16}  & \textcolor{red}{93,45\%}          & \textcolor{red}{97,16\%}          & {83,6\%}                 & {77,06\%}                & \textcolor{red}{94,5\%}  & {89\%}                   \\
-        {3} & {50}  & \textcolor{red}{\textbf{90,97\%}} & \textcolor{red}{95,74\%}          & \textbf{74,49\%}         & \textbf{68,89\%}         & \textcolor{red}{92,37\%} & {84,98\%}                \\
-        {3} & {100} & \textcolor{red}{96,74\%}          & \textcolor{red}{\textbf{92,76\%}} & \textcolor{red}{95,51\%} & \textcolor{red}{90,35\%} & \textbf{71,88\%}         & \textbf{69,69\%}         \\
-        {4} & {2}   & \textcolor{red}{94,94\%}          & \textcolor{red}{96,45\%}          & {89,08\%}                & {83,89\%}                & \textcolor{red}{92,29\%} & {85,7\%}                 \\
-        {4} & {10}  & \textcolor{red}{\textbf{91,62\%}} & \textcolor{red}{\textbf{93,29\%}} & {78,53\%}                & {70,85\%}                & {81,99\%}                & \textbf{74,83\%}         \\
-        {4} & {16}  & \textcolor{red}{92,86\%}          & \textcolor{red}{96,45\%}          & {79,49\%}                & {73,49\%}                & \textcolor{red}{94,05\%} & {86,05\%}                \\
-        {4} & {50}  & \textcolor{red}{94\%}             & \textcolor{red}{94,78\%}          & {86,17\%}                & {79,61\%}                & {88,11\%}                & {82,65\%}                \\
-        {4} & {100} & \textcolor{red}{92,15\%}          & \textcolor{red}{96,99\%}          & {78,91\%}                & {72,51\%}                & \textcolor{red}{94,8\%}  & {88,6\%}                 \\
-        {6} & {16}  & \textcolor{red}{\textbf{90,04\%}} & \textcolor{red}{95,83\%}          & \textbf{69,71\%}         & \textbf{61,75\%}         & \textcolor{red}{91,77\%} & {81,6\%}                 \\
-        {6} & {50}  & \textcolor{red}{91,9\%}           & \textcolor{red}{96,28\%}          & \textbf{76,04\%}         & \textbf{69,86\%}         & \textcolor{red}{92,81\%} & {85,2\%}                 \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attention predictor: Dale-2}
-    \label{tab:probing:attention-predictor:dale-2}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh} & \textbf{Sh > C > Si}              & \textbf{Sh > Si > C} & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
-        {1} & {2}   & \textcolor{red}{95,35\%}          & {89,87\%}            & \textcolor{red}{95,09\%}          & {87,23\%}            & {86,01\%}            & {84,14\%}            \\
-        {1} & {16}  & \textcolor{red}{92,95\%}          & {87,03\%}            & \textcolor{red}{92,63\%}          & {83,91\%}            & {82,77\%}            & {80,06\%}            \\
-        {1} & {50}  & \textcolor{red}{93,35\%}          & {87,23\%}            & \textcolor{red}{93,26\%}          & {84,65\%}            & {82,46\%}            & {80,38\%}            \\
-        {2} & {10}  & \textbf{88,94\%}                  & \textbf{84,78\%}     & \textbf{87,5\%}                   & \textbf{77,76\%}     & \textbf{80,24\%}     & \textbf{74,95\%}     \\
-        {2} & {16}  & \textbf{85,71\%}                  & \textbf{84,53\%}     & \textbf{82,98\%}                  & \textbf{72,85\%}     & \textbf{79,72\%}     & \textbf{71,85\%}     \\
-        {3} & {16}  & \textcolor{red}{\textbf{92,18\%}} & \textbf{86,23\%}     & \textcolor{red}{91,93\%}          & \textbf{82,59\%}     & \textbf{81,69\%}     & \textbf{78,72\%}     \\
-        {3} & {50}  & \textcolor{red}{92,68\%}          & {86,53\%}            & \textcolor{red}{91,9\%}           & {83,35\%}            & {82,45\%}            & {79,99\%}            \\
-        {6} & {10}  & \textcolor{red}{93,4\%}           & {86,51\%}            & \textcolor{red}{\textbf{91,64\%}} & {84,03\%}            & {82,51\%}            & {80,06\%}            \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attention predictor: Dale-5}
-    \label{tab:probing:attention-predictor:dale-5}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{cc|c|c|c|c|c|c}
-        \toprule
-        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
-        {3} & {16}  & \textcolor{red}{98,48\%} & \textcolor{red}{98,58\%} & \textcolor{red}{97,87\%} & \textcolor{red}{96,11\%} & \textcolor{red}{96,6\%}  & \textcolor{red}{95,97\%} \\
-        {4} & {10}  & \textcolor{red}{98,88\%} & \textcolor{red}{98,75\%} & \textcolor{red}{97,23\%} & \textcolor{red}{96,67\%} & \textcolor{red}{97,18\%} & \textcolor{red}{96,75\%} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Attention predictor: Color}
-    \label{tab:probing:attention-predictor:colour}
-\end{table}
-
-\begin{table}[ht]
-    \centering
-    \begin{tabular}{c|cc|cc|cc|cc}
-        \toprule
-        \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-                                &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
-        {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
-        {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
-        {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
-        {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
-        {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
-        {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
-        \bottomrule
-    \end{tabular}
-    \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
-    % SD: I have a trouble of understanding and interpreting these results.
+% \begin{table}[ht]
+%     \centering
+%     \begin{tabular}{c|cc|cc|cc|cc}
+%         \toprule
+%         \textbf{Order}          & $L_{baseline}$ & $L_{English}$ & \multicolumn{2}{c}{$|V| = 10$} & \multicolumn{2}{c}{$|V| = 13$} & \multicolumn{2}{c}{$|V| = 100$}                                                      \\\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+%                                 &                &               & $L_{em}$                       & $L_{norm}$                     & $L_{em}$                        & $L_{norm}$         & $L_{em}$ & $L_{norm}$         \\\midrule
+%         {shape > color > size}  & {0,659}        & {0,0001}      & {0,569}                        & \textbf{86,19\%}               & {0,622}                         & {94,24\%}          & {0,597}  & {90,4\%}           \\
+%         {shape > size > color}  & {0,589}        & {0,0002}      & {0,489}                        & \textbf{83,09\%}               & {0,531}                         & {90,21\%}          & {0,47}   & {\textbf{79,78\%}} \\
+%         {color > shape > size}  & {0,849}        & {0,0}         & {0,802}                        & 94,49\%                        & {0,801}                         & {94,36\%}          & {0,791}  & {93,15\%}          \\
+%         {color  > size > shape} & {0,836}        & {0,0}         & {0,819}                        & 98,01\%                        & {0,772}                         & {92,35\%}          & {0,786}  & {94\%}             \\
+%         {size > shape > color}  & {0,532}        & {0,0052}      & {0,492}                        & 92,26\%                        & {0,437}                         & {\textbf{81,95\%}} & {0,457}  & {\textbf{85,61\%}} \\
+%         {size > color > shape}  & {0,599}        & {0,0001}      & {0,573}                        & 95,71\%                        & {0,495}                         & {\textbf{82,67\%}} & {0,538}  & {\textbf{89,87\%}} \\
+%         \bottomrule
+%     \end{tabular}
+%     \caption{Cross entropy losses while probing with emerged languages successful on 'Dale-2'}
+%     % SD: I have a trouble of understanding and interpreting these results.
 First the loss should be estimated on a test set, after the system has been trained.
 I'm not sure how you estaimate the baseline loss.
-    % DK: TODO, why a test set?
+%     % DK: TODO, why a test set?
 Isn't the purpose to learn how well the generated captions map to English and not how well they generalize.
 Using only the train set exactly shows if a mapping is possible
-    \label{tab:probing_discrimnator_language}
-\end{table}
+%     \label{tab:probing_discrimnator_language}
+% \end{table}
 
-Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
-First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
-Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
-This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
-Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
-% SD: I think at this point a readerwill be confused since we are discussing Dale and Reiter bith for generation of scenes and now for the generation of strings.
+% Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
+% First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
+% Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
+% This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
+% Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
+% % SD: I think at this point a readerwill be confused since we are discussing Dale and Reiter bith for generation of scenes and now for the generation of strings.
 We should have clarified the difference earlier.
 Also, rather than referring to it as Dale and Reiter we should call it "the incremental algorithm" and add a reference at first mention.
-% DK: TODO
+% % DK: TODO
 
-Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
-The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
-With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
-Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
-The attribute \emph{color} seems to be less important.
-This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
+% Secondly, it can be seen that the correlation between the emerged language and the natural language referring expressions differ for each of the languages.
+% The vocabulary consisting of 10 symbols is the closest related to the orders \emph{shape > size > color} and \emph{shape > color > size}, the vocabulary based on 13 symbols on the other hand to the orders \emph{size > shape > color} and \emph{size > color > shape}.
+% With 100 symbols, the vocabulary resembles mostly \emph{shape > size > color} and \emph{size > shape > color}.
+% Striking here is that even though the related orders are different, the most important attributes are either the \emph{size} or \emph{shape} across all three emerged languages.
+% The attribute \emph{color} seems to be less important.
+% This is reinforced by the fact that the losses are closer to the baseline, when the \emph{color} is the most or second most important attribute in the order.
 
-An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
-In a second step, the language is compared to several referring expressions in natural language.
-By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
-% from prev chapter
+% An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
+% In a second step, the language is compared to several referring expressions in natural language.
+% By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.
+% % from prev chapter
 
-In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
-They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
-In this section, these three emerged languages is analyzed in more detail.
-This is done in two parts.
-The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
-In the second part, the emerged languages are compared to English.
-More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
+% In the above experiments, only three configurations lead to success, where a language emerged that the agents used to exchange information: languages with vocabulary sizes of $|V| = 10$, $|V| = 13$ and $|V| = 100$.
+% They will be referred to as $Lang_{10}$, $Lang_{13}$ and $Lang_{100}$ respectively.
+% In this section, these three emerged languages is analyzed in more detail.
+% This is done in two parts.
+% The first part is a qualitative analysis, to understand, which symbols are used to transfer which information.
+% In the second part, the emerged languages are compared to English.
+% More specifically, it is tested if the messages of the sender align with English referring expressions of the target object.
 
-\begin{figure}[ht]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
-        \label{fig:relative_frequencies_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
-        \label{fig:relative_frequencies_v13}
-    }
-    \subfigure[$Lang_{100}$, (13 most frequent messages)]{
-        \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
-        \label{fig:relative_frequencies_v100}
-    }
-    \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
-    \label{fig:relative_frequencies_vocabularies}
-\end{figure}
+% \begin{figure}[ht]
+%     \centering
+%     \subfigure[$Lang_{10}$]{
+%         \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v10.png}
+%         \label{fig:relative_frequencies_v10}
+%     }
+%     \subfigure[$Lang_{13}$]{
+%         \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v13.png}
+%         \label{fig:relative_frequencies_v13}
+%     }
+%     \subfigure[$Lang_{100}$, (13 most frequent messages)]{
+%         \includegraphics[width=0.31\linewidth]{figures/relative_frequencies_v100.png}
+%         \label{fig:relative_frequencies_v100}
+%     }
+%     \caption{Relative frequencies of messages (ordered by frequency in test dataset)}
+%     \label{fig:relative_frequencies_vocabularies}
+% \end{figure}
 
-When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
-Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
-% SD: I don't understand how test and training set are used in these experiments.
+% When looking at the emerged vocabulary \textbf{qualitatively}, a few properties can be seen.
+% Figure \ref{fig:relative_frequencies_vocabularies} shows an overview of the frequencies of messages in all three emerged languages for the training and the test split.
+% % SD: I don't understand how test and training set are used in these experiments.
 In language games one only has a training set and success is measured through the success in the game.
 It is not possible to have a test set as learning is continuous.
-% DK: TODO
-Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
-By this the languages are easier comparable across different runs and vocabulary sizes.
-% SD: I don't understand this.
-% DK: TODO
-Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
-The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
+% % DK: TODO
+% Since the tokens themselves are arbitrary, they are ordered by the relative frequency in the messages for the test set and indices for the tokens are added from index 1 for the most frequent token and index $|V|$ for the least frequent token.
+% By this the languages are easier comparable across different runs and vocabulary sizes.
+% % SD: I don't understand this.
+% % DK: TODO
+% Figure \ref{fig:relative_frequencies_v100} shows only the 13 most frequent message to provide a better overview.
+% The lesser frequent messages are never used for the test data and each only used one or two times in the train data.
 
-The first property is that when a message is transferred, it consists of only one symbol.
-% SD: Is this controlled by the sender.
+% The first property is that when a message is transferred, it consists of only one symbol.
+% % SD: Is this controlled by the sender.
 The sender must have some policy how it is assigning messages.
@@
 If it is not controlled - how could we implement some policies for message length?
-% DK: what do you mean by policy? The sender is only controled by the success of the receiver. If a longer message leads to a better success, it learns to prodcue longer messages. TODO
-In some rare cases, also an empty message is communicated.
-The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
-Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
-More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
-Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
-This is especially visible for $Lang_{100}$.
-Symbols that are used for 16,5\% of the training sample are not used at all in test split.
-Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
-% SD: A table or a graph with these figures?
-% DK: is part of the existing figure. The sum of the first two columns (QUESTION)
-% SD: Differences between training and testing configuration; what happens during testing, there is no learning involved but the systems see new images but they have to use the same vocabulary?
-% DK: exactly
+% % DK: what do you mean by policy? The sender is only controled by the success of the receiver. If a longer message leads to a better success, it learns to prodcue longer messages. TODO
+% In some rare cases, also an empty message is communicated.
+% The models therefore don't learn any compositionality by combining symbols to create new meaning, but rather encode everything in separate symbols.
+% Secondly, in all three languages, only very few symbols occur with a high frequency, while most of the symbols are used very rarely.
+% More specifically, two symbols are used in 95\% of the images with $Lang_{10}$ and $Lang_{13}$, while three symbols are used with $Lang_{100}$.
+% Thirdly, the agents make use of fewer symbols, when presented with unseen test images compared to when communicating about images in the training split.
+% This is especially visible for $Lang_{100}$.
+% Symbols that are used for 16,5\% of the training sample are not used at all in test split.
+% Furthermore, the frequencies in the test split is much more focused on the two most frequent symbols, while it is more distributed around 5 symbols in the train split.
+% % SD: A table or a graph with these figures?
+% % DK: is part of the existing figure. The sum of the first two columns (QUESTION)
+% % SD: Differences between training and testing configuration; what happens during testing, there is no learning involved but the systems see new images but they have to use the same vocabulary?
+% % DK: exactly
 
-These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
-However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
-% SD: Are they?
+% These findings indicate that referring expressions do emerge in each of the newly emerged languages since the agents are able to communicate the correct object.
+% However, the agents converge towards very few different referring expressions that are made up differently than in English and likely don't rely on the high level attributes \emph{shape}, \emph{color} and \emph{size}.
+% % SD: Are they?
 We haven't tested how humans would describe these scene.
 I see what you mean, but they you have to explain that you count as an English expression a description containing sequences of colour, size and type and then following the Dale and Reiter's algorithm.
-% DK: TODO
-The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
-% SD: All this has to do with the sender's policty how to generate messages.
+% % DK: TODO
+% The similar frequencies across all three languages suggest that a greater vocabulary size $|V|$ doesn't necessarily lead to different referring expressions, but the languages still converge towards two main expressions.
+% % SD: All this has to do with the sender's policty how to generate messages.
 We should have varied this policy.
@@
 How is the system motivated to generate more than one symbol?
-% DK: TODO
+% % DK: TODO
 
-\begin{figure}[ht]
-    \centering
-    \subfigure[$Lang_{10}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
-        \label{fig:language_analysis_v10}
-    }
-    \subfigure[$Lang_{13}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
-        \label{fig:language_analysis_v13}
-    }
-    \subfigure[$Lang_{100}$]{
-        \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
-        \label{fig:language_analysis_v100}
-    }
-    \caption{Relative share of the described target object's attributes for the top three messages}
+% \begin{figure}[ht]
+%     \centering
+%     \subfigure[$Lang_{10}$]{
+%         \includegraphics[width=0.31\linewidth]{figures/language_analysis_v10.png}
+%         \label{fig:language_analysis_v10}
+%     }
+%     \subfigure[$Lang_{13}$]{
+%         \includegraphics[width=0.31\linewidth]{figures/language_analysis_v13.png}
+%         \label{fig:language_analysis_v13}
+%     }
+%     \subfigure[$Lang_{100}$]{
+%         \includegraphics[width=0.31\linewidth]{figures/language_analysis_v100.png}
+%         \label{fig:language_analysis_v100}
+%     }
+%     \caption{Relative share of the described target object's attributes for the top three messages}
 
-    \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
-    \textbf{Size:} brown: small, blue: large
-    \label{fig:language_analysis_vocabularies}
-\end{figure}
+%     \textbf{Shape:} brown: sphere, blue: cube, red: cylinder \\
+%     \textbf{Size:} brown: small, blue: large
+%     \label{fig:language_analysis_vocabularies}
+% \end{figure}
 
-\cmtDK[inline]{differnces in different languages correspond to differences in color prediction of single model caption generation}
+% \cmtDK[inline]{differences in different languages correspond to differences in color prediction of single model caption generation}
 
-Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
-Hereby, the relative share each value of all three attributes are displayed.
-For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
-In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
+% Figure \ref{fig:language_analysis_vocabularies} shows the attributes of the target object that are described by each message.
+% Hereby, the relative share each value of all three attributes are displayed.
+% For instance the first bar in Figure \ref{fig:language_analysis_v10} shows that 16\% of the target objects that are described with symbol (1) in the language $Lang_{10}$ are spheres.
+% In the figures, only the three most frequently used messages are included, which make up over 95\% of all messages.
 
-The first thing that can be seen is that different symbols convey different values of attributes.
-In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
-There is not much difference for target objects that are cylinders.
-The distribution for colors is almost constant across symbols (1) and (2).
-Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
-This different distribution may also be caused by the much lower absolute usage of symbol (3).
-The different sizes are encoded by all symbols in the same way.
+% The first thing that can be seen is that different symbols convey different values of attributes.
+% In the language $Lang_{10}$, symbol (1) is mostly used for target objects that are cubes, while symbol (2) is mostly used for spheres.
+% There is not much difference for target objects that are cylinders.
+% The distribution for colors is almost constant across symbols (1) and (2).
+% Symbol (3) is more frequently used for blue, green and gray objects, while being less used for red, cyan and yellow objects.
+% This different distribution may also be caused by the much lower absolute usage of symbol (3).
+% The different sizes are encoded by all symbols in the same way.
 
-Looking at $Lang_{13}$ the symbol usage differs.
-The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
-Again, the colors look similar with small deviations for brown, green and yellow objects.
-Most striking however, is the difference for the size.
-Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
+% Looking at $Lang_{13}$ the symbol usage differs.
+% The frequencies for the shape look similar, but (the much less used) symbol (3) encodes most of the time spheres instead of cubes.
+% Again, the colors look similar with small deviations for brown, green and yellow objects.
+% Most striking however, is the difference for the size.
+% Symbol (1) is used in 77\% of the cases for large objects, while symbol (2) and (3) are used in 90\% and 71\% of the messages for small objects.
 
-Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
-The usage for the color is similar to $Lang_{13}$.
-Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
+% Language $Lang_{100}$ uses symbols its symbols to discriminate cubes and cylinders, while the frequencies for sphere remain constant across all symbols.
+% The usage for the color is similar to $Lang_{13}$.
+% Looking at the size, as for $Lang_{13}$, symbol (1) is mostly used to encode small objects, while symbols (2) and (3) are mostly used to encode large objects.
 
-An interesting observation is that symbols are not used for the same attributes across the languages.
-In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
-The same applies to the values of attributes, especially to the shapes.
-Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
-Which shape are encoded, differs from language to language.
+% An interesting observation is that symbols are not used for the same attributes across the languages.
+% In some languages, an attribute is not captured at all by the symbols, while another language heavily relies on it.
+% The same applies to the values of attributes, especially to the shapes.
+% Only two of the shapes are distinguished by the usage of symbols, while the third is not captured.
+% Which shape are encoded, differs from language to language.
 
-Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
-For instance $Lang_{10}$ only encodes the shape in its symbols.
-This would not be enough to distinguish the target object from the distractor consistently.
-Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
-% SD: But this is slightly problematic, since we strcutured the world this way to emphasise these attributes visually.
+% Furthermore, these numbers confirm even more that the agents don't rely solely on the human defined attributes.
+% For instance $Lang_{10}$ only encodes the shape in its symbols.
+% This would not be enough to distinguish the target object from the distractor consistently.
+% Following, the agents also encode some additional underlying attributes and patterns to the three above defined attributes.
+% % SD: But this is slightly problematic, since we strcutured the world this way to emphasise these attributes visually.
 Hence, a mapping would indicate that the symbols have good semantics if they can discriminate the colours.
 But this could have to do with the compisitionality issue, perhaps the system was biased to use single word expressions and hence this interfered with the grounding.
-% DK: TODO
+% % DK: TODO
+
+\cmtDK[inline]{qualitative analysis}

### 2024-05-12 thesis/sections/7.1-object-identification.tex
quantitative language analysis

\subsection{Object identification}
\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {1} & {2}   & \textcolor{red}{95,5\%}           & \textcolor{red}{97,57\%}          & {86,98\%}                & {81,43\%}                & \textcolor{red}{91,67\%} & {86,45\%}                \\
        {1} & {10}  & \textcolor{red}{96,11\%}          & \textcolor{red}{96,9\%}           & {89,61\%}                & {85\%}                   & {89,4\%}                 & {84,49\%}                \\
        {1} & {16}  & \textcolor{red}{95,19\%}          & \textcolor{red}{96,56\%}          & {87,25\%}                & {81,22\%}                & {89,34\%}                & {83,13\%}                \\
        {1} & {50}  & \textcolor{red}{96,61\%}          & \textcolor{red}{97,07\%}          & \textcolor{red}{91,26\%} & {85,75\%}                & {87,14\%}                & {82,61\%}                \\
        {1} & {100} & \textcolor{red}{95,91\%}          & \textcolor{red}{96,73\%}          & \textcolor{red}{90,63\%} & {86,01\%}                & {88,95\%}                & {84,17\%}                \\
        {2} & {2}   & \textcolor{red}{99,67\%}          & \textcolor{red}{99,91\%}          & \textcolor{red}{97,02\%} & \textcolor{red}{95,37\%} & \textcolor{red}{96,97\%} & \textcolor{red}{94,64\%} \\
        {2} & {10}  & \textcolor{red}{96,36\%}          & \textcolor{red}{96,34\%}          & \textcolor{red}{91,32\%} & {86,75\%}                & {87,69\%}                & {82,87\%}                \\
        {2} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{97,12\%}          & {88,11\%}                & {83,21\%}                & {89,86\%}                & {85,35\%}                \\
        {2} & {50}  & \textcolor{red}{95,34\%}          & \textcolor{red}{97,3\%}           & {85,48\%}                & {80,89\%}                & \textcolor{red}{92,22\%} & {86,18\%}                \\
        {2} & {100} & \textcolor{red}{94,88\%}          & \textcolor{red}{96,32\%}          & {87,37\%}                & {81,17\%}                & {88,35\%}                & {82,61\%}                \\
        {3} & {2}   & \textcolor{red}{\textbf{94,03\%}} & \textcolor{red}{96,96\%}          & \textbf{79,59\%}         & \textbf{74,54\%}         & \textcolor{red}{92,32\%} & {85,63\%}                \\
        {3} & {10}  & \textcolor{red}{96,1\%}           & \textcolor{red}{95,97\%}          & \textcolor{red}{90,45\%} & {84,84\%}                & {85,98\%}                & {80,5\%}                 \\
        {3} & {16}  & \textcolor{red}{95,71\%}          & \textcolor{red}{\textbf{95,1\%}}  & \textcolor{red}{90,97\%} & {85,33\%}                & {81\%}                   & \textbf{75,56\%}         \\
        {3} & {50}  & \textcolor{red}{96,32\%}          & \textcolor{red}{\textbf{94,55\%}} & \textcolor{red}{92,5\%}  & {86,81\%}                & \textbf{80,48\%}         & {75,73\%}                \\
        {3} & {100} & \textcolor{red}{95,33\%}          & \textcolor{red}{97,82\%}          & {86,05\%}                & {81,33\%}                & \textcolor{red}{93,91\%} & {88,54\%}                \\
        {4} & {2}   & \textcolor{red}{96,81\%}          & \textcolor{red}{95,12\%}          & \textcolor{red}{93,03\%} & {87,41\%}                & \textbf{80,73\%}         & \textbf{75,47\%}         \\
        {4} & {10}  & \textcolor{red}{95,31\%}          & \textcolor{red}{97,26\%}          & {86,9\%}                 & {82,67\%}                & \textcolor{red}{92,39\%} & {86,98\%}                \\
        {4} & {16}  & \textcolor{red}{95,63\%}          & \textcolor{red}{\textbf{94,65\%}} & \textcolor{red}{91,99\%} & {86,65\%}                & \textbf{79,23\%}         & \textbf{74,54\%}         \\
        {4} & {50}  & \textcolor{red}{94,76\%}          & \textcolor{red}{96,69\%}          & {86,8\%}                 & {81,33\%}                & {89,65\%}                & {83,92\%}                \\
        {4} & {100} & \textcolor{red}{96,16\%}          & \textcolor{red}{97,94\%}          & {89,57\%}                & {85,18\%}                & \textcolor{red}{92,56\%} & {87,74\%}                \\
        {6} & {2}   & \textcolor{red}{95,73\%}          & \textcolor{red}{96,99\%}          & {87,77\%}                & {83,27\%}                & \textcolor{red}{91,77\%} & {86,34\%}                \\
        {6} & {10}  & \textcolor{red}{\textbf{94,59\%}} & \textcolor{red}{99,89\%}          & {86,24\%}                & {81,46\%}                & \textcolor{red}{91,42\%} & {86,43\%}                \\
        {6} & {16}  & \textcolor{red}{95,51\%}          & \textcolor{red}{96,3\%}           & {88,89\%}                & {83,24\%}                & {87,61\%}                & {82,71\%}                \\
        {6} & {50}  & \textcolor{red}{\textbf{94,66\%}} & \textcolor{red}{97,43\%}          & \textbf{84,93\%}         & \textbf{78,61\%}         & \textcolor{red}{91,48\%} & {83,66\%}                \\
        {6} & {100} & \textcolor{red}{95,37\%}          & \textcolor{red}{97,77\%}          & \textbf{83,75\%}         & \textbf{80,84\%}         & \textcolor{red}{93,96\%} & {89,02\%}                \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{discrimination games} on the 'Dale-2' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:discriminator:dale-2}
\end{table}

Looking at the results for the discriminator, one can see that the results are very different for each dataset.
On the 'Dale-2' dataset $L_{norm}$ is relatively high for all salience orders (see Table \ref{tab:probing:discriminator:dale-2}).
Yet, the correlation to English referring expressions with the \emph{color} as most important attribute is the lowest.
When the \emph{color} attribute is in the second position, the correlation becomes higher, but the highest correlation is achieved when the \emph{color} has the least importance.
This shows that the agents don't use the \emph{color} in the communication, but rather rely on \emph{shape} and \emph{size}.
The order of importance of these attributes however is not as clear.
For some combinations of the message length $n$ and vocabulary size $|V|$, the \emph{size} seems to be the dominant attribute, while for others the \emph{shape} seems more important in the communication.
Interestingly, one can see that for many languages only one salience order has a low $L_{norm}$, while the $L_{norm}$ for the remaining salience orders with a different attribute in the first position is higher.
This applies for example to $Lang_{3,2}$, where the loss for 'Sh > Si > C' is 74,54\% while the second lowest is only at 85,63\% for 'Si > Sh > C'.
The same can be seen for instance for $Lang_{3,16}$ in the other direction.
This difference indicates that the agents are relying mostly on the most important attribute of this salience order to discriminate the target object.
If the $L_{norm}$ for instance were similar for the salience orders 'Sh > Si > C' and 'Si > Sh > C', this would mean that the emerged language could be connected to both salience orders in the same way.
Following, the agents would use the \emph{shape} as often as \emph{size} to discriminate the objects.
This is however only the case on the 'Dale-2' dataset when the loss is higher (for instance languages $Lang_{2,16}$, $Lang_{2,100}$ or $Lang_{6,16}$).
Since the lowest $L_{norm}$ with 74,54\% is still very high compared to a complete correlation (0\%), the agents seem to encode additionally other information in their messages.

Looking at the influence of $n$ and $|V|$, one can see that across almost all values, there is at least some correlation to one salience order.
The correlation seems to be highest with $n \in \{3,4\}$, where the loss can reach down 74,54\%.
Hereby, the most important attribute in the messages seems to be the \emph{size}.
For the remaining message lengths $n$, the loss stays mostly between 80\% and 85\%.
For $n \in \{2,6\}$ the \emph{shape} seems to be the most dominant attribute, while for $n=1$ the \emph{size} and \emph{shape} seem to have a similar importance.
The influence of the vocabulary size $|V|$ is not as clear.
Small to medium-sized vocabularies seem to promote higher correlations with English referring expressions, but also languages with $|V|=100$ can reach losses of around 80\%.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
        {1} & {2}   & \textcolor{red}{94,96\%} & \textcolor{red}{91,27\%} & \textcolor{red}{94,51\%} & \textcolor{red}{90,99\%} & {88,52\%}            & {88,17\%}            \\
        {1} & {10}  & {84,34\%}                & {84,03\%}                & {80,91\%}                & {73,23\%}                & {79,95\%}            & {72,34\%}            \\
        {1} & {16}  & {86,27\%}                & {84,96\%}                & {82,76\%}                & {75,35\%}                & {80,72\%}            & {74,04\%}            \\
        {1} & {50}  & {85,95\%}                & {84,4\%}                 & {82,96\%}                & {74,97\%}                & {80,2\%}             & {73,54\%}            \\
        {1} & {100} & {86,31\%}                & {85,58\%}                & {82,9\%}                 & {76,03\%}                & {81,64\%}            & {74,62\%}            \\
        {2} & {10}  & \textbf{82,66\%}         & {83,23\%}                & \textbf{78,51\%}         & {71,01\%}                & {78,89\%}            & {70,61\%}            \\
        {2} & {16}  & {89,78\%}                & {85,53\%}                & {87,68\%}                & {80,64\%}                & {81,06\%}            & {77,24\%}            \\
        {2} & {50}  & {84,4\%}                 & {82,65\%}                & {80,62\%}                & {71,42\%}                & {77,53\%}            & {69,46\%}            \\
        {2} & {100} & {89,65\%}                & {85,3\%}                 & {87,26\%}                & {79,91\%}                & {80,88\%}            & {77,1\%}             \\
        {3} & {2}   & {87,19\%}                & {86,93\%}                & {83,99\%}                & {78,18\%}                & {83,65\%}            & {77,08\%}            \\
        {3} & {10}  & {87,15\%}                & {84,67\%}                & {84,01\%}                & {76,24\%}                & {80,41\%}            & {74,11\%}            \\
        {3} & {50}  & \textbf{82,98\%}         & {82,25\%}                & {78,85\%}                & {70,62\%}                & {77,42\%}            & {69,27\%}            \\
        {3} & {100} & {83,78\%}                & {83,07\%}                & {80,23\%}                & {72,15\%}                & {78,94\%}            & {71,1\%}             \\
        {4} & {2}   & {89,23\%}                & {87,43\%}                & {87,09\%}                & {81,45\%}                & {83,83\%}            & {79,11\%}            \\
        {4} & {10}  & {88,66\%}                & {84,83\%}                & {85,9\%}                 & {78,98\%}                & {80,2\%}             & {75,64\%}            \\
        {4} & {16}  & {83,14\%}                & \textbf{81,51\%}         & \textbf{78,15\%}         & \textbf{70,21\%}         & \textbf{77,32\%}     & \textbf{68,48\%}     \\
        {4} & {50}  & {87,62\%}                & {84,38\%}                & {85,25\%}                & {77,29\%}                & {79,94\%}            & {74,03\%}            \\
        {4} & {100} & \textcolor{red}{90,17\%} & {85,76\%}                & {87,68\%}                & {80,71\%}                & {81,37\%}            & {78,1\%}             \\
        {6} & {10}  & \textbf{82,03\%}         & \textbf{80,51\%}         & \textbf{77,46\%}         & \textbf{68,63\%}         & \textbf{75,71\%}     & \textbf{67,15\%}     \\
        {6} & {16}  & {84,73\%}                & \textbf{81,12\%}         & {80,99\%}                & \textbf{70,5\%}          & \textbf{76,06\%}     & \textbf{69,12\%}     \\
        {6} & {50}  & {86,95\%}                & {85,13\%}                & {84,06\%}                & {76,95\%}                & {80,11\%}            & {74,51\%}            \\
        {6} & {100} & {86,15\%}                & {82,19\%}                & {83,56\%}                & {75,56\%}                & {77,76\%}            & {72,47\%}            \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{discrimination games} on the 'Dale-5' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:discriminator:dale-5}
\end{table}

The emerged languages on the 'Dale-5' dataset have generally a higher correlation to English referring expressions (see Table \ref{tab:probing:discriminator:dale-5}).
For almost all salience orders and emerged languages, the loss is lower than 90\%.
In the best cases, it even reaches down to 67,15\%.
Again, the \emph{size} and the \emph{shape} are mostly communicated between the agents, but the \emph{color} now plays a bigger role in the messages.
The agents seem to need to rely on more attributes to discriminate the target object.
This aligns also with English referring expressions, where the average length of the referring expressions across the whole 'Dale-5' dataset is longer than for the 'Dale-2' dataset.
However, opposed to the 'Dale-2' dataset, there are almost no languages, where there is one dominant attribute.
More often the loss is similar for two salience orders with different primary attributes.
For instance the salience orders 'Sh > Si > C' and 'Si > Sh > C' both provide losses around 70\% for $Lang_{6,16}$.
Only the languages $Lang_{2,16}$, $Lang_{4,10}$, $Lang_{4,50}$ and $Lang_{6,100}$ have differences of more than 3\% points for these salience orders.
This indicates, that the agents make use of both \emph{shape} and \emph{size} similarly often, while the \emph{color} is used for much fewer images.

When the agents are able to produce longer messages, their languages seem to have more similarities with possible English languages.
With $n>1$ the agents produce languages that can have losses lower than 70\% for the salience order 'Si > Sh > C'.
Yet, languages with $n=1$ are not far off and have mostly losses between 72\% and 75\%.
Looking at the vocabulary size, languages with $|V|=2$ are less likely to emerge and if they emerge they have a lower correlation with English.
The best results are produces with $|V| \in \{10,16\}$ for longer messages (for instance $Lang_{4,16}$, $Lang_{6,10}$ and $Lang_{6,16}$), while shorter messages fare generally better with a larger vocabulary of $|V| \in \{50,100\}$ ($Lang_{2,50}$, $Lang_{3,50}$ and $Lang_{3,100}$).

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {1} & {2}   & \textcolor{red}{91,62\%} & \textcolor{red}{91,57\%} & \textcolor{red}{94,31\%} & \textcolor{red}{95,13\%} & \textcolor{red}{93,11\%} & \textcolor{red}{94,92\%} \\
        {1} & {10}  & {79,08\%}                & {79,08\%}                & {85,04\%}                & {86,86\%}                & {82,37\%}                & {86,38\%}                \\
        {1} & {16}  & {84,02\%}                & {83,96\%}                & {88,67\%}                & {89,65\%}                & {86,36\%}                & {89,36\%}                \\
        {1} & {50}  & \textcolor{red}{93,88\%} & \textcolor{red}{93,74\%} & \textcolor{red}{95,03\%} & \textcolor{red}{95,49\%} & \textcolor{red}{94,64\%} & \textcolor{red}{95,35\%} \\
        {2} & {10}  & {88,67\%}                & {88,36\%}                & {89,59\%}                & \textcolor{red}{90,77\%} & \textcolor{red}{90,01\%} & {89,81\%}                \\
        {2} & {16}  & {81,13\%}                & {81,02\%}                & {86,16\%}                & {87,97\%}                & {83,85\%}                & {87,56\%}                \\
        {3} & {2}   & {83,38\%}                & {83,46\%}                & {88,21\%}                & {89,71\%}                & {86,05\%}                & {89,37\%}                \\
        {3} & {10}  & \textbf{77\%}            & \textbf{76,77\%}         & \textbf{83,54\%}         & \textbf{85,57\%}         & \textbf{79,97\%}         & \textbf{84,9\%}          \\
        {3} & {16}  & \textcolor{red}{93,52\%} & \textcolor{red}{93,6\%}  & \textcolor{red}{94,16\%} & \textcolor{red}{94,14\%} & \textcolor{red}{93,75\%} & \textcolor{red}{93,55\%} \\
        {4} & {10}  & \textbf{68,04\%}         & \textbf{66,51\%}         & \textbf{76,33\%}         & \textbf{78,56\%}         & \textbf{71,42\%}         & \textbf{78,56\%}         \\
        {4} & {16}  & \textbf{75,1\%}          & \textbf{74,93\%}         & \textbf{81,2\%}          & \textbf{83,35\%}         & \textbf{77,6\%}          & \textbf{82,76\%}         \\
        {6} & {2}   & {84,21\%}                & {84,02\%}                & {88,41\%}                & {89,77\%}                & {86,61\%}                & {89,35\%}                \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{discrimination games} on the 'CLEVR color' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:discriminator:colour}
\end{table}

Finally on the 'CLEVR color' dataset, much fewer languages emerge (see Table \ref{tab:probing:discriminator:colour}).
Those languages that emerge have very different levels of correlation to English
While three languages stay above 90\% for all salience orders, most of them stay between 80\% and 90\%.
Four languages even get below 80\% for the salience orders where the \emph{color} is the most important attribute.
One of those even reaches 66,51\%, the highest correlation across all emerged languages on the discrimination task.
Unsurprisingly, the \emph{color} is now the most important attribute.
The loss with the remaining salience orders is still relatively low, but this is likely due to the fact, that the color occurs in any referring expression as it is the only discriminating attribute.
The probing model can therefore easily connect this pattern to the emerged language.
As expected, the losses for 'Sh > Si > C' and 'Si > Sh > C' are almost the same, which indicates that the agents make (no) use of both \emph{shape} and \emph{size} in the same way.

Even though only one attribute would need to be communicated, languages with $n \in \{3,4\}$ tend to be closer to English referring expressions.
Hereby, vocabulary sizes of $|V| \in \{10,16\}$ produce the most correlating languages, while languages with only 2 symbols have relatively high losses.
The agents were not able to communicate with vocabulary sizes larger than 16.

### 2024-05-12 thesis/sections/7.2-re-generation.tex
quantitative language analysis

\subsection{Referring expression generation}

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {1} & {2}   & \textcolor{red}{95,07\%} & \textcolor{red}{98,06\%}          & {86,63\%}                & {85,3\%}                 & \textcolor{red}{96,26\%} & \textcolor{red}{93,32\%} \\
        {1} & {10}  & \textcolor{red}{90,8\%}  & \textcolor{red}{96,4\%}           & {71,88\%}                & {69,29\%}                & \textcolor{red}{94,05\%} & {85,13\%}                \\
        {1} & {16}  & \textcolor{red}{91,47\%} & \textcolor{red}{96,96\%}          & {71,2\%}                 & {66,82\%}                & \textcolor{red}{93,83\%} & {83,35\%}                \\
        {1} & {50}  & \textcolor{red}{93,26\%} & \textcolor{red}{96,63\%}          & {77,53\%}                & {71,94\%}                & \textcolor{red}{93,56\%} & {84,62\%}                \\
        {1} & {100} & \textcolor{red}{93,5\%}  & \textcolor{red}{97,86\%}          & {77,42\%}                & {71,93\%}                & \textcolor{red}{94,68\%} & {85,77\%}                \\
        {2} & {2}   & \textcolor{red}{93,57\%} & \textcolor{red}{97,72\%}          & {80,89\%}                & {79,8\%}                 & \textcolor{red}{96,08\%} & \textcolor{red}{90,71\%} \\
        {2} & {10}  & \textbf{84,86\%}         & \textcolor{red}{\textbf{92,07\%}} & {58,8\%}                 & \textbf{52,09\%}         & \textbf{88,36\%}         & {77,01\%}                \\
        {2} & {16}  & {87,29\%}                & \textcolor{red}{92,88\%}          & {68,52\%}                & {71,5\%}                 & \textcolor{red}{92,56\%} & {83,9\%}                 \\
        {2} & {50}  & {89,15\%}                & \textcolor{red}{95\%}             & {68,19\%}                & {68,93\%}                & \textcolor{red}{93,11\%} & {83,95\%}                \\
        {2} & {100} & \textcolor{red}{91,06\%} & \textcolor{red}{96,14\%}          & {71,98\%}                & {67,23\%}                & \textcolor{red}{92,81\%} & {82,71\%}                \\
        {3} & {2}   & \textcolor{red}{93,05\%} & \textcolor{red}{97,15\%}          & {82,41\%}                & {77,36\%}                & \textcolor{red}{94,43\%} & {88,8\%}                 \\
        {3} & {10}  & {88,52\%}                & \textcolor{red}{95,41\%}          & {66,61\%}                & {63,63\%}                & \textcolor{red}{92,25\%} & {81,46\%}                \\
        {3} & {16}  & {85,71\%}                & \textcolor{red}{92,88\%}          & \textbf{56,44\%}         & \textbf{49,22\%}         & \textbf{88,87\%}         & \textbf{74,54\%}         \\
        {3} & {50}  & \textcolor{red}{90,78\%} & \textcolor{red}{96,28\%}          & {68,17\%}                & {64,84\%}                & \textcolor{red}{94,16\%} & {83,5\%}                 \\
        {3} & {100} & {86,54\%}                & \textcolor{red}{93,54\%}          & {65,9\%}                 & {66,03\%}                & \textcolor{red}{92,49\%} & {83,28\%}                \\
        {4} & {10}  & {87,61\%}                & \textcolor{red}{94,63\%}          & {66,83\%}                & {64,85\%}                & \textcolor{red}{92,24\%} & {82,85\%}                \\
        {4} & {16}  & \textbf{85,7\%}          & \textcolor{red}{93,56\%}          & \textbf{56,9\%}          & \textbf{52,1\%}          & \textbf{88,98\%}         & \textbf{76,15\%}         \\
        {4} & {50}  & {87,55\%}                & \textcolor{red}{94,24\%}          & {61,03\%}                & {55,82\%}                & \textcolor{red}{91,73\%} & \textbf{76,52\%}         \\
        {4} & {100} & {88,84\%}                & \textcolor{red}{94,84\%}          & {68,68\%}                & {66,73\%}                & \textcolor{red}{92,57\%} & {81,79\%}                \\
        {6} & {2}   & \textcolor{red}{93,48\%} & \textcolor{red}{98,14\%}          & {81,36\%}                & {79,08\%}                & \textcolor{red}{96,09\%} & \textcolor{red}{90,95\%} \\
        {6} & {10}  & \textbf{85\%}            & \textcolor{red}{\textbf{92,44\%}} & \textbf{57,21\%}         & {55,07\%}                & {89,39\%}                & {77,44\%}                \\
        {6} & {16}  & {88,02\%}                & \textcolor{red}{94,22\%}          & {71,42\%}                & {66,35\%}                & \textcolor{red}{90,86\%} & {84,12\%}                \\
        {6} & {50}  & \textcolor{red}{98,57\%} & \textcolor{red}{99,45\%}          & \textcolor{red}{95,57\%} & \textcolor{red}{95,06\%} & \textcolor{red}{99,16\%} & \textcolor{red}{96,64\%} \\
        {6} & {100} & {86,44\%}                & \textcolor{red}{\textbf{92,76\%}} & {65,59\%}                & {65,93\%}                & \textcolor{red}{92,42\%} & {82,16\%}                \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{referring expression generation games} on the 'Dale-2' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:re-generator:dale-2}
\end{table}

Tables \ref{tab:probing:re-generator:dale-2} and \ref{tab:probing:re-generator:dale-5} show the results of the probing on the emerged languages of the \emph{referring expression generation games}.
This is the only task, where the final task for the agents is to generate English referring expressions with the salience order 'Sh > C > Si'.
English referring expressions are therefore explicitly passed to the agents as correct labels during training.
One can therefore expect the agents to align their own language with the target referring expressions and a high correlation with the salience order 'Sh > C > Si' is expected.
Looking at the results, this is partially confirmed for the 'Dale-2' dataset.
Some emerged languages reach with a normalized loss of 49,22\% one of the highest correlations to English referring expressions across all emerged languages in this thesis.
However, the salience order with the highest correlation is 'Sh > Si > C', and values the \emph{size} more than the \emph{color}, instead of the reversed way as in the training data of the agents.
The color seems, as in the previous task, not as important to communicate as the other attributes.
While the \emph{size} plays a bigger role, the \emph{shape} is the central attribute in the communication.
This fact is interesting for two reasons:
First, the \emph{shape} being the most important attribute in the messages aligns with the receiver's generated referring expressions where the \emph{shape} was predicted almost perfectly.
The receiver therefore seems to make use of the sender's ability to extract the \emph{shape} from the images and reproduces what the sender communicated.
However, without a deeper analysis, it can't be ruled out that the receiver extracts the \emph{shape} themselves from the image, and uses the sender's message only to determine the correct target object in the scene.
Secondly, this does not apply directly to the \emph{size} and the \emph{color}.
The \emph{size} seems to be communicated often by the sender to discriminate the target object.
On the other hand, as the target referring expression uses the salience order 'Sh > C > Si', the shape is only needed to be generated in $\frac{1}{3} * \frac{1}{8} \approx 4,2\%$ of the samples.
The opposite applies to the \emph{color}.
The \emph{color} is not important in the agents' communication, but the receiver however needs to produce colors in the generated referring expression in $\frac{1}{3} \approx 33,3\%$ of the samples.
This shows that the agents actually don't align their own language with the target referring expression.
The communication of the target object between the agents seems to be a separate step from the receiver's generation of the referring expression.

Looking at the influence of the message length, one can see that the language is the least correlated to the salience order 'Sh > Si > C' with $n=1$.
Still, even with these short messages, the loss can go down to 66,82\%.
With $n>1$, almost all languages have a loss of below 70\%, while at least one configuration is even lower than 60\%.
Hereby, languages with $n \in \{3,4\}$ have the highest correlation to English.
Languages with the highest correlation also mostly make use of a medium-sized vocabulary of $|V| \in \{10,16\}$ as for example $Lang_{2,10}$, $Lang_{3,16}$, $Lang_{4,16}$ and $Lang_{6,10}$.
While larger vocabularies, can still correlate relatively much with English referring expressions, languages with $|V|=2$ are much less similar.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {1} & {2}   & \textcolor{red}{92,86\%} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,06\%} & {89,5\%}             & \textcolor{red}{95,73\%} & \textcolor{red}{90,93\%} \\
        {1} & {10}  & {83,9\%}                 & \textcolor{red}{90,9\%}  & {78,63\%}                & {74,74\%}            & {89,6\%}                 & {78,99\%}                \\
        {1} & {16}  & {84,68\%}                & \textcolor{red}{91,65\%} & {80,11\%}                & {77,03\%}            & \textcolor{red}{90,16\%} & {81,05\%}                \\
        {1} & {50}  & {84,85\%}                & \textcolor{red}{91,67\%} & {79,95\%}                & {77,11\%}            & \textcolor{red}{90,32\%} & {81,04\%}                \\
        {2} & {2}   & \textcolor{red}{90,51\%} & \textcolor{red}{95,36\%} & {87,5\%}                 & {86,43\%}            & \textcolor{red}{94,77\%} & {88,35\%}                \\
        {2} & {10}  & {82,44\%}                & \textcolor{red}{90,32\%} & {77,65\%}                & {73,73\%}            & {88,89\%}                & {78,41\%}                \\
        {2} & {16}  & {81,85\%}                & \textbf{89,4\%}          & {76,07\%}                & \textbf{71,66\%}     & {87,9\%}                 & \textbf{75,91\%}         \\
        {2} & {50}  & {83,58\%}                & \textcolor{red}{91,12\%} & {78,67\%}                & {75,03\%}            & {89,94\%}                & {79,06\%}                \\
        {3} & {2}   & \textcolor{red}{92,51\%} & \textcolor{red}{95,96\%} & {89,96\%}                & {88,7\%}             & \textcolor{red}{95,18\%} & \textcolor{red}{90,43\%} \\
        {3} & {10}  & {82,11\%}                & {89,54\%}                & {75,9\%}                 & {71,78\%}            & {88,18\%}                & {76,72\%}                \\
        {3} & {50}  & {83,12\%}                & \textcolor{red}{90,82\%} & {77,62\%}                & {74,08\%}            & {88,88\%}                & {78,74\%}                \\
        {3} & {100} & {83,74\%}                & {89,59\%}                & {79,29\%}                & {74,48\%}            & \textbf{87,46\%}         & {78,05\%}                \\
        {4} & {2}   & {88,86\%}                & \textcolor{red}{94\%}    & {84,8\%}                 & {83,19\%}            & \textcolor{red}{93,37\%} & {85,92\%}                \\
        {4} & {10}  & \textbf{79,93\%}         & \textbf{89,29\%}         & \textbf{74,1\%}          & \textbf{69,5\%}      & \textbf{87,19\%}         & \textbf{74,79\%}         \\
        {4} & {16}  & \textbf{81,26\%}         & {89,68\%}                & \textbf{75,73\%}         & {71,96\%}            & {88,4\%}                 & {76,93\%}                \\
        {4} & {50}  & {81,55\%}                & {89,42\%}                & {75,9\%}                 & {71,68\%}            & {88,17\%}                & {76,55\%}                \\
        {6} & {2}   & \textcolor{red}{91,09\%} & \textcolor{red}{95,26\%} & {87,86\%}                & {86,81\%}            & \textcolor{red}{94,78\%} & {88,82\%}                \\
        {6} & {50}  & \textbf{80,39\%}         & \textbf{87,85\%}         & \textbf{75,01\%}         & \textbf{68,71\%}     & \textbf{84,54\%}         & \textbf{72,93\%}         \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{referring expression generation games} on the 'Dale-5' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:re-generator:dale-5}
\end{table}

On the 'Dale-5' dataset, the emerged languages again resemble the English referring expressions less.
While the dominant salience order is 'Sh > Si > C' as on the 'Dale-2' dataset, the \emph{color} plays a bigger role.
This effect is the same as for the discrimination games, and can be explained by the higher number of distractors.
While the loss for the salience order 'Sh > Si > C' is lower than 70\% for two languages, it stays between 70\% and 80\% for most.
Interesting now is that all attributes are communicated by the sender, while the receiver is also generating the correct attributes in the referring expression.
Opposed to the 'Dale-2' dataset, the need for \emph{size} and \emph{color} is much bigger to discriminate the target object in any salience order, due to the larger number of distractors.

$Lang_{6,50}$ is the emerged language, with which the agents can achieve the best results on the task.
While the receiver achieves high precision and recall scores on each attribute (recall Tables \ref{tab:results:re-generator-game_size-shape} and \ref{tab:results:re-generator-game_color}), each attribute also plays a bigger role in the message be the sender.
Indeed, one can see a correlation between the predictions of the receiver and the usage of attributes in emerged language.
A more frequent usage of an attribute in the emerged language leads to a more accurate generation of this attribute in the final referring expression.

As for the previous analyses, a vocabulary size of $|V|=2$ seems to be too small for the agents to use similar referring expressions like in English.
Additionally to $|V| \in \{10,16\}$, also a vocabulary size of $|V|=50$ now helps the agents to create to most resembling languages.
Only the language $L_{3,100}$ emerges, that utilizes a vocabulary of $|V|=100$ and it correlates similarly to English as $L_{3,50}$ and slightly worse than $L_{3,10}$.

Looking at the message length, a larger $n$ brings the emerged languages close to English referring expressions.
With $n=1$ the normalized loss remains above 74,74\%.
With growing $n$, the loss decreases, and even gets lower than 70\% for $n \in \{4,6\}$.
However, while $Lang_{6,50}$ is the language with the highest correlation, it is one of only two languages that emerged with $n=6$.
This indicates that agents struggle with so few contraint, but if they do converge to a language, it makes use of similar structures as English.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}      & \textbf{Si > Sh > C}      \\\midrule
        {1} & {2}   & \textcolor{red}{95,73\%} & \textcolor{red}{94,23\%} & \textcolor{red}{92,96\%} & \textcolor{red}{90,59\%} & {81,59\%}                 & {78,21\%}                 \\
        {1} & {10}  & \textcolor{red}{91,77\%} & \textcolor{red}{91,19\%} & {73,14\%}                & {63,99\%}                & {67,45\%}                 & {57,55\%}                 \\
        {1} & {16}  & \textcolor{red}{91,74\%} & \textcolor{red}{90,87\%} & {73,07\%}                & {63,78\%}                & {66,16\%}                 & {55,48\%}                 \\
        {1} & {50}  & \textcolor{red}{97,78\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,22\%} & {87,45\%}                & {69,23\%}                 & {64,97\%}                 \\
        {1} & {100} & \textcolor{red}{97,91\%} & \textcolor{red}{93,1\%}  & \textcolor{red}{94,18\%} & {87,35\%}                & {69,25\%}                 & {64,99\%}                 \\
        {2} & {2}   & \textcolor{red}{93,17\%} & \textcolor{red}{91,69\%} & {81,96\%}                & {75,97\%}                & {72,05\%}                 & {64,29\%}                 \\
        {2} & {10}  & \textcolor{red}{94,07\%} & \textcolor{red}{98,04\%} & {75,32\%}                & {71,91\%}                & \textcolor{red}{94,38\%}  & {86,62\%}                 \\
        {2} & {16}  & \textcolor{red}{97,72\%} & \textcolor{red}{92,9\%}  & \textcolor{red}{94,18\%} & {87,09\%}                & {68,54\%}                 & {64,13\%}                 \\
        {2} & {50}  & \textcolor{red}{97,87\%} & \textcolor{red}{93,08\%} & \textcolor{red}{94,25\%} & {87,54\%}                & {69,2\%}                  & {65,17\%}                 \\
        {2} & {100} & {89,6\%}                 & \textbf{89,58\%}         & \textbf{65,65\%}         & \textbf{55,69\%}         & {64,92\%}                 & {53,53\%}                 \\
        {3} & {2}   & \textcolor{red}{97,73\%} & \textcolor{red}{99,28\%} & \textcolor{red}{95,84\%} & \textcolor{red}{97,72\%} & \textcolor{red}{101,58\%} & \textcolor{red}{100,22\%} \\
        {3} & {10}  & \textcolor{red}{91,04\%} & \textcolor{red}{90,52\%} & {71,06\%}                & {61,36\%}                & {65,59\%}                 & {54,93\%}                 \\
        {3} & {16}  & \textcolor{red}{91,09\%} & \textcolor{red}{90,3\%}  & {71,71\%}                & {62,42\%}                & {66,28\%}                 & {56,16\%}                 \\
        {3} & {50}  & \textcolor{red}{97,07\%} & \textcolor{red}{91,72\%} & \textcolor{red}{92,7\%}  & {85,33\%}                & {66,38\%}                 & {61,26\%}                 \\
        {3} & {100} & \textcolor{red}{97,81\%} & \textcolor{red}{93,01\%} & \textcolor{red}{94,16\%} & {87,3\%}                 & {68,83\%}                 & {64,6\%}                  \\
        {4} & {2}   & \textcolor{red}{91,95\%} & \textcolor{red}{96,87\%} & {73,55\%}                & {72,66\%}                & \textcolor{red}{98,37\%}  & \textcolor{red}{90,23\%}  \\
        {4} & {10}  & \textcolor{red}{97,7\%}  & \textcolor{red}{93,1\%}  & \textcolor{red}{94,16\%} & {87,24\%}                & {69,22\%}                 & {64,81\%}                 \\
        {4} & {16}  & \textbf{86,75\%}         & \textbf{89\%}            & \textbf{57,02\%}         & \textbf{44,33\%}         & \textbf{63,78\%}          & \textbf{48,21\%}          \\
        {4} & {50}  & \textcolor{red}{90,86\%} & \textcolor{red}{90,5\%}  & {71,22\%}                & {61,09\%}                & {65,26\%}                 & {54,87\%}                 \\
        {4} & {100} & \textcolor{red}{96,4\%}  & \textcolor{red}{90,22\%} & \textcolor{red}{92,44\%} & {83,78\%}                & \textbf{59,98\%}          & {54,38\%}                 \\
        {6} & {2}   & \textcolor{red}{90,06\%} & \textcolor{red}{90,46\%} & {71,62\%}                & {64,56\%}                & {70,67\%}                 & {61,35\%}                 \\
        {6} & {10}  & \textbf{88,38\%}         & \textcolor{red}{90,24\%} & \textbf{63,68\%}         & \textbf{53,21\%}         & \textbf{64,22\%}          & \textbf{52,19\%}          \\
        {6} & {16}  & \textbf{89,01\%}         & \textbf{89,74\%}         & {67,03\%}                & {56,73\%}                & {64,44\%}                 & \textbf{52,78\%}          \\
        {6} & {50}  & \textcolor{red}{97,68\%} & \textcolor{red}{92,77\%} & \textcolor{red}{97,03\%} & {87,24\%}                & {68,51\%}                 & {64,27\%}                 \\
        {6} & {100} & {89,84\%}                & {89,81\%}                & {69,75\%}                & {60,55\%}                & {65,19\%}                 & {54,93\%}                 \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attribute generation games} on the 'Dale-2' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:attribute-generator:dale-2}
\end{table}

Tables \ref{tab:probing:attribute-generator:dale-2} to \ref{tab:probing:attribute-generator:colour} show the results for the probing of the emerged languages in the \emph{attribute generation games}.
The first detail that can be seen is the difference of the results between the 'Dale' datasets and the 'CLEVR color' dataset.
On the 'Dale' dataset, the \emph{color} plays no or only a small role in the languages.
While most of the emerged languages correlate to the salience order 'Si > Sh > C', languages with the lowest loss correlate to the salience order 'Sh > Si > C'.
These indeed achieve the lowest loss across all experiments and tasks in this thesis.

When looking closer on the results for the 'Dale-2' dataset, the influence of $n$ and $|V|$ is less clear.
While high correlations to English can be achieved with any $n$, the normalized loss remains consistently low only with $n \geq 2$.
Hereby, a medium vocabulary size of $|V| \in \{10,16\}$ provides the best results (for instance $Lang_{3,10}$, $Lang_{3,16}$, $Lang_{4,16}$, $Lang_{6,10}$ and $Lang_{6,16}$).
However, also large vocabularies of $|V| \in \{50,100\}$ can give low losses as seen in $Lang_{2,100}$, $Lang_{4,50}$, $Lang_{4,100}$ and $Lang_{6,100}$.
This is nonetheless less consistent.
When given a vocabulary with $|V|=2$, the emerged languages have almost always a relatively low correlation to English.

A similar detail to the emerged languages of the \emph{referring expression games} is as well visible here.
Some languages as for instance $Lang_{2,100}$, $Lang_{6,2}$ and $Lang_{6,10}$ achieve similar losses for the salience orders 'Sh > Si > C' and 'Si > Sh > C' which indicates that the \emph{size} and the \emph{shape} are used equally often in the messages.
Interestingly, this doesn't have any effect on the final predictions of the receiver.
No big difference can be identified for games where the agents make use of these languages or of languages with only one dominant attribute.
Indeed, the parsing of the sender's message seems not to be directly connected to the final prediction of the attributes by the receiver.
While the \emph{size} is almost always communicated as the most important attribute, it is never predicted correctly by the receiver, as the accuracy of the \emph{size} lies aroudnd 52\%.
The opposite applies to the \emph{color}.
This indicates that the sender's message is only used to extract the target object from the image, but the extraction of the attributes for the final prediction is a separate step.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {1} & {2}   & \textcolor{red}{91,81\%} & \textcolor{red}{96,22\%} & {89,08\%}                & {88,2\%}                 & \textcolor{red}{95,91\%} & \textcolor{red}{90,3\%}  \\
        {1} & {10}  & {83,4\%}                 & {82,11\%}                & {79,68\%}                & {67,88\%}                & {77,12\%}                & {66,75\%}                \\
        {1} & {16}  & {86,76\%}                & {87,06\%}                & {83,08\%}                & {75,03\%}                & {83,79\%}                & {74,72\%}                \\
        {1} & {50}  & {83,12\%}                & {82\%}                   & {78,81\%}                & {66,76\%}                & {76,95\%}                & {65,95\%}                \\
        {1} & {100} & \textcolor{red}{93,84\%} & {87,09\%}                & \textcolor{red}{93,01\%} & {83,36\%}                & {82,19\%}                & {79,14\%}                \\
        {2} & {2}   & {88,86\%}                & {86,82\%}                & {86,62\%}                & {78,45\%}                & {82,59\%}                & {76,49\%}                \\
        {2} & {10}  & {83,91\%}                & {81,57\%}                & {79,76\%}                & {67,89\%}                & {76,47\%}                & {66,1\%}                 \\
        {2} & {16}  & \textbf{75,6\%}          & \textbf{77,74\%}         & \textbf{68,78\%}         & \textbf{55,43\%}         & \textbf{72,24\%}         & \textbf{56,24\%}         \\
        {2} & {50}  & {85,1\%}                 & {83,08\%}                & {81,27\%}                & {70,13\%}                & {77,86\%}                & {68,54\%}                \\
        {2} & {100} & {83,24\%}                & {81,87\%}                & {79,02\%}                & {67,32\%}                & {77,06\%}                & {66,19\%}                \\
        {3} & {2}   & \textcolor{red}{97,21\%} & \textcolor{red}{96,08\%} & \textcolor{red}{96,43\%} & \textcolor{red}{95,23\%} & \textcolor{red}{95,18\%} & \textcolor{red}{94,24\%} \\
        {3} & {10}  & {82,66\%}                & {81,11\%}                & {78,38\%}                & {66,21\%}                & {75,68\%}                & {64,52\%}                \\
        {3} & {16}  & \textcolor{red}{93,26\%} & {86,09\%}                & \textcolor{red}{92,67\%} & {82,25\%}                & {81,17\%}                & {77,84\%}                \\
        {3} & {50}  & {82,83\%}                & {81,44\%}                & {78,57\%}                & {66,29\%}                & {76,04\%}                & {64,84\%}                \\
        {3} & {100} & {82,74\%}                & {81\%}                   & {78,39\%}                & {66,04\%}                & {75,44\%}                & {64,91\%}                \\
        {4} & {2}   & \textcolor{red}{92,91\%} & {87,03\%}                & \textcolor{red}{92,16\%} & {84,06\%}                & {82,45\%}                & {79,57\%}                \\
        {4} & {10}  & \textbf{79,3\%}          & \textbf{79,42\%}         & \textbf{74,65\%}         & \textbf{61,71\%}         & \textbf{74,12\%}         & \textbf{60,62\%}         \\
        {4} & {16}  & {82,55\%}                & {81,31\%}                & {78,61\%}                & {66,51\%}                & {75,61\%}                & {65,05\%}                \\
        {4} & {50}  & {88,51\%}                & {84,17\%}                & {85,77\%}                & {75,07\%}                & {79\%}                   & {71,93\%}                \\
        {4} & {100} & {82,57\%}                & {81,61\%}                & {78,06\%}                & {66,33\%}                & {76,32\%}                & {65,54\%}                \\
        {6} & {2}   & \textcolor{red}{92,36\%} & {88,15\%}                & \textcolor{red}{91,49\%} & {84,38\%}                & {85,28\%}                & {80,85\%}                \\
        {6} & {10}  & \textbf{78,11\%}         & \textbf{79,19\%}         & \textbf{72,81\%}         & \textbf{59,86\%}         & \textbf{73,95\%}         & \textbf{60,17\%}         \\
        {6} & {16}  & {82,35\%}                & {81,09\%}                & {78,84\%}                & {66,02\%}                & {75,9\%}                 & {64,79\%}                \\
        {6} & {50}  & {82,04\%}                & {81,49\%}                & {78,34\%}                & {66,25\%}                & {76,46\%}                & {64,91\%}                \\
        {6} & {100} & {82,09\%}                & {80,73\%}                & {77,42\%}                & {65,3\%}                 & {75,52\%}                & {64,52\%}                \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attribute generation games} on the 'Dale-5' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:attribute-generator:dale-5}
\end{table}

As expected, Table \ref{tab:probing:attribute-generator:dale-5} shows that the \emph{color} plays a bigger role for the 'Dale-5' dataset.
Still, the dominant salience order is 'Si > Sh > C', in some cases 'Sh > Si > C'.
As opposed to the 'Dale-2' dataset, both salience orders have almost always a similarly low loss which indicates that the \emph{size} and \emph{shape} are equally important.

Looking at the influence of the message length $n$, one can see, that languages with $n=1$ tend to have the worst correlation to English referring expressions.
While there are few outliers, languages with $n \geq 2$ perform similarly.
The same applies as well to the vocabulary size $|V|$.
When the agents are given a very small vocabulary with $|V|=2$, the resulting languages look very different from English.
An increase to $|V|=10$, lets the agents produce the most similar languages.
Increasing it further tends to lower the correlation slightly, but stays mostly still around 65\%.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si} & \textbf{Sh > Si > C} & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C} \\\midrule
        {1} & {2}   & \textcolor{red}{99,92\%} & \textcolor{red}{99,88\%} & \textbf{79,83\%}     & \textbf{82,7\%}      & \textcolor{red}{99,71\%} & \textbf{82,14\%}     \\
        {1} & {10}  & \textbf{70,52\%}         & \textbf{70,36\%}         & \textbf{80,23\%}     & \textbf{82,36\%}     & \textbf{74,49\%}         & \textbf{81,77\%}     \\
        {1} & {16}  & \textbf{77,25\%}         & \textbf{77,16\%}         & {84,55\%}            & {86,2\%}             & \textbf{80,72\%}         & {85,56\%}            \\
        {3} & {10}  & \textbf{71,46\%}         & \textbf{71,47\%}         & \textbf{78,81\%}     & \textbf{81,87\%}     & \textbf{75,95\%}         & \textbf{80,93\%}     \\
        {6} & {2}   & {83,25\%}                & {83,23\%}                & {87,74\%}            & {88,95\%}            & {85,38\%}                & {88,87\%}            \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attribute generation games} on the 'CLEVR color' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:attribute-generator:colour}
\end{table}

Finally on the 'CLEVR color' dataset, only very few languages emerge.
As expected, the most of them correlate with salience orders that have the \emph{color} as its most important attribute.
The one exception is $Lang_{1,2}$ with the salience order 'Sh > C > Si' that doesn't rely on the \emph{color} at all.
However, looking at the results of the receiver's predictions in Table \ref{tab:results:attribute-predictor-game}, one can see that this configuration doesn't beat the baseline and the communication doesn't seem to help to solve the task.

Looking at the remaining languages, the losses for the salience orders 'C > Sh > Si' and 'C > Si > Sh' are almost equal, which indicates, that the \emph{size} and \emph{color} are (not) used equally often.
However, that correlation in between 70\% to 83\% is relatively low compared to the correlations of down to below 50\% on the 'Dale' datasets.
This can be confirmed by the accuracies of the \emph{color} attribute of the receivers' predictions.
Configurations, where a language between the agents emerges perform only around 1\% higher than the baseline of 90\% with no communication.
The receiver can solve the task almost on their own and the messages of the sender only help in few cases.
The sender seems to be only needed to communicate very specific information instead of always communicating the \emph{color}.
This could for instance be specific \emph{colors}, the receiver struggles with, or other underlying structural patterns.
This results in a different language and the lower correlation to English referring expressions.

The languages with the lowest correlation have a medium-sized vocabulary of $|V| \in \{10,16\}$ and are relatively short with $n \in \{1,3\}$.
While the language $Lang_{6,2}$ emerges that uses a small vocabulary with long messages, the normalized loss is much higher.

### 2024-05-12 thesis/sections/7.3-reference-resolution.tex
quantitative language analysis

\subsection{Reference resolution}

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh}              & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {1} & {2}   & \textcolor{red}{95,17\%}          & \textcolor{red}{97,23\%}          & {87,74\%}                & {82,8\%}                 & \textcolor{red}{94,4\%}  & {88,75\%}                \\
        {1} & {10}  & \textcolor{red}{96,04\%}          & \textcolor{red}{94,67\%}          & \textcolor{red}{94,06\%} & {88,67\%}                & {82,86\%}                & {79,71\%}                \\
        {1} & {16}  & \textcolor{red}{93,82\%}          & \textcolor{red}{95,48\%}          & {85,33\%}                & {79,96\%}                & \textcolor{red}{90,17\%} & {84,6\%}                 \\
        {1} & {50}  & \textcolor{red}{94,47\%}          & \textcolor{red}{96,65\%}          & {86,98\%}                & {81,89\%}                & \textcolor{red}{90,93\%} & {86,51\%}                \\
        {1} & {100} & \textcolor{red}{95,6\%}           & \textcolor{red}{98,19\%}          & \textcolor{red}{90,41\%} & {86,13\%}                & \textcolor{red}{95,74\%} & \textcolor{red}{93,76\%} \\
        {2} & {2}   & \textcolor{red}{94,75\%}          & \textcolor{red}{96,65\%}          & {86,51\%}                & {80,84\%}                & \textcolor{red}{92,99\%} & {86,3\%}                 \\
        {2} & {10}  & \textcolor{red}{94,69\%}          & \textcolor{red}{98\%}             & {86,92\%}                & {82,2\%}                 & \textcolor{red}{96,31\%} & \textcolor{red}{92,55\%} \\
        {2} & {16}  & \textcolor{red}{95,04\%}          & \textcolor{red}{96,39\%}          & {87,82\%}                & {82,54\%}                & \textcolor{red}{90,91\%} & {87,2\%}                 \\
        {2} & {50}  & \textcolor{red}{97,18\%}          & \textcolor{red}{93,99\%}          & \textcolor{red}{96,42\%} & \textcolor{red}{92,26\%} & \textbf{77,97\%}         & {75,58\%}                \\
        {2} & {100} & \textcolor{red}{93,71\%}          & \textcolor{red}{94,65\%}          & {85,56\%}                & {78,95\%}                & {83,77\%}                & {78,69\%}                \\
        {3} & {2}   & \textcolor{red}{95,24\%}          & \textcolor{red}{97,37\%}          & {87,56\%}                & {82,49\%}                & \textcolor{red}{94,38\%} & {88,87\%}                \\
        {3} & {10}  & \textcolor{red}{96,62\%}          & \textcolor{red}{\textbf{91,61\%}} & \textcolor{red}{94,87\%} & {89\%}                   & \textbf{69,2\%}          & \textbf{65,45\%}         \\
        {3} & {16}  & \textcolor{red}{93,45\%}          & \textcolor{red}{97,16\%}          & {83,6\%}                 & {77,06\%}                & \textcolor{red}{94,5\%}  & {89\%}                   \\
        {3} & {50}  & \textcolor{red}{\textbf{90,97\%}} & \textcolor{red}{95,74\%}          & \textbf{74,49\%}         & \textbf{68,89\%}         & \textcolor{red}{92,37\%} & {84,98\%}                \\
        {3} & {100} & \textcolor{red}{96,74\%}          & \textcolor{red}{\textbf{92,76\%}} & \textcolor{red}{95,51\%} & \textcolor{red}{90,35\%} & \textbf{71,88\%}         & \textbf{69,69\%}         \\
        {4} & {2}   & \textcolor{red}{94,94\%}          & \textcolor{red}{96,45\%}          & {89,08\%}                & {83,89\%}                & \textcolor{red}{92,29\%} & {85,7\%}                 \\
        {4} & {10}  & \textcolor{red}{\textbf{91,62\%}} & \textcolor{red}{\textbf{93,29\%}} & {78,53\%}                & {70,85\%}                & {81,99\%}                & \textbf{74,83\%}         \\
        {4} & {16}  & \textcolor{red}{92,86\%}          & \textcolor{red}{96,45\%}          & {79,49\%}                & {73,49\%}                & \textcolor{red}{94,05\%} & {86,05\%}                \\
        {4} & {50}  & \textcolor{red}{94\%}             & \textcolor{red}{94,78\%}          & {86,17\%}                & {79,61\%}                & {88,11\%}                & {82,65\%}                \\
        {4} & {100} & \textcolor{red}{92,15\%}          & \textcolor{red}{96,99\%}          & {78,91\%}                & {72,51\%}                & \textcolor{red}{94,8\%}  & {88,6\%}                 \\
        {6} & {16}  & \textcolor{red}{\textbf{90,04\%}} & \textcolor{red}{95,83\%}          & \textbf{69,71\%}         & \textbf{61,75\%}         & \textcolor{red}{91,77\%} & {81,6\%}                 \\
        {6} & {50}  & \textcolor{red}{91,9\%}           & \textcolor{red}{96,28\%}          & \textbf{76,04\%}         & \textbf{69,86\%}         & \textcolor{red}{92,81\%} & {85,2\%}                 \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attention prediction games} on the 'Dale-2' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:attention-predictor:dale-2}
\end{table}

Finally, Tables \ref{tab:probing:attention-predictor:dale-2} to \ref{tab:probing:attention-predictor:colour} show the similarities between the emerged languages on the \emph{attention prediction games} and English referring expressions.
On the 'Dale-2' dataset, one can see that most of the emerged languages correlate to English in some way.
The least correlating language is $Lang_{1,100}$ with 86,13\% towards the salience order 'Sh > Si > C'.
Hereby, a strong influence of the message length $n$ and vocabulary size $|V|$ can be extracted.
Languages with short messages of $n \in \{1,2\}$ are less connected to English referring expressions.
With $n=2$, larger vocabulary sizes can bring the loss down below 80\% as seen for the languages $Lang_{2,50}$ and $Lang_{2,100}$.
Languages with $n \geq 3$ are correlated more, whereas languages with $n=3$ perform the best.
Looking at the vocabulary size, one can see that again a vocabulary size of $|V|=2$ will make the emerged language different from English.
A vocabulary size of $|V|=10$ seems to provide the most similar languages ($Lang_{3,10}$, $Lang_{4,10}$) for larger $n$, but larger vocabularies of $|V| \in \{50,100\}$ are only few percent points higher.
When given languages with $n=6$, the agents can utilize them only in few cases with medium-sized vocabularies, but if they do, the emerged languages are very similar.

Most of the languages are based on the salience order 'Sh > Si > C', and only the languages $Lang_{2,50}$, $Lang_{3,10}$ and $Lang_{3,100}$ use the salience order 'Si > Sh > C'.
The \emph{color} plays almost no role in the communication.
In almost all the languages, one attribute is dominant while the remaining two attributes are used much less frequently.
This can be seen in the large margin between the losses of the salience orders where the first two attributes are swapped (for instance 'Sh > Si > C' and 'Si > Sh > C').

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}              & \textbf{C > Si > Sh} & \textbf{Sh > C > Si}              & \textbf{Sh > Si > C} & \textbf{Si > C > Sh} & \textbf{Si > Sh > C} \\\midrule
        {1} & {2}   & \textcolor{red}{95,35\%}          & {89,87\%}            & \textcolor{red}{95,09\%}          & {87,23\%}            & {86,01\%}            & {84,14\%}            \\
        {1} & {16}  & \textcolor{red}{92,95\%}          & {87,03\%}            & \textcolor{red}{92,63\%}          & {83,91\%}            & {82,77\%}            & {80,06\%}            \\
        {1} & {50}  & \textcolor{red}{93,35\%}          & {87,23\%}            & \textcolor{red}{93,26\%}          & {84,65\%}            & {82,46\%}            & {80,38\%}            \\
        {2} & {10}  & \textbf{88,94\%}                  & \textbf{84,78\%}     & \textbf{87,5\%}                   & \textbf{77,76\%}     & \textbf{80,24\%}     & \textbf{74,95\%}     \\
        {2} & {16}  & \textbf{85,71\%}                  & \textbf{84,53\%}     & \textbf{82,98\%}                  & \textbf{72,85\%}     & \textbf{79,72\%}     & \textbf{71,85\%}     \\
        {3} & {16}  & \textcolor{red}{\textbf{92,18\%}} & \textbf{86,23\%}     & \textcolor{red}{91,93\%}          & \textbf{82,59\%}     & \textbf{81,69\%}     & \textbf{78,72\%}     \\
        {3} & {50}  & \textcolor{red}{92,68\%}          & {86,53\%}            & \textcolor{red}{91,9\%}           & {83,35\%}            & {82,45\%}            & {79,99\%}            \\
        {6} & {10}  & \textcolor{red}{93,4\%}           & {86,51\%}            & \textcolor{red}{\textbf{91,64\%}} & {84,03\%}            & {82,51\%}            & {80,06\%}            \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attention prediction games} on the 'Dale-5' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:attention-predictor:dale-5}
\end{table}

On the 'Dale-5' dataset, mach fewer languages emerge and those that do emerge have a much lower correlation to English referring expressions than languages on the 'Dale-2' dataset.
The losses for salience orders 'Sh > Si > C' and 'Si > Sh > C' are much closer, meaning both \emph{size} and \emph{shape} are used similarly often in the messages.
The \emph{size} is hereby slightly more important with a lower loss of 1\% to 4\% points.
The color plays a larger role than on the 'Dale-2' dataset, but is still the least important attribute.

The emerged languages are making use of shorter messages.
The most similar languages use a message length of $n=2$ followed by $n=3$
While one language emerges with $n=6$, it is less similar to English.
Looking at the vocabulary size, a medium-sized vocabulary with $|V| \in \{10,16,50\}$ provides the most similar languages.
However, no conclusive argument about this dependency can be made, as too few languages emerged to compare the effects of the vocabulary thoroughly.

\begin{table}[ht]
    \centering
    \begin{tabular}{cc|c|c|c|c|c|c}
        \toprule
        $n$ & $|V|$ & \textbf{C > Sh > Si}     & \textbf{C > Si > Sh}     & \textbf{Sh > C > Si}     & \textbf{Sh > Si > C}     & \textbf{Si > C > Sh}     & \textbf{Si > Sh > C}     \\\midrule
        {3} & {16}  & \textcolor{red}{98,48\%} & \textcolor{red}{98,58\%} & \textcolor{red}{97,87\%} & \textcolor{red}{96,11\%} & \textcolor{red}{96,6\%}  & \textcolor{red}{95,97\%} \\
        {4} & {10}  & \textcolor{red}{98,88\%} & \textcolor{red}{98,75\%} & \textcolor{red}{97,23\%} & \textcolor{red}{96,67\%} & \textcolor{red}{97,18\%} & \textcolor{red}{96,75\%} \\
        \bottomrule
    \end{tabular}
    \caption{Normalized losses $L_{norm}$ of the probing model on emerged languages of the \emph{attention prediction games} on the 'CLEVR color' dataset. The emerged languages are probed with different salience orders where 'C' corresponds to the \emph{color}, 'Sh' to the \emph{shape} and 'Si' to the \emph{size}. The table only includes languages, with which the agents could successfully solve the task.}
    \label{tab:probing:attention-predictor:colour}
\end{table}

Finally, Table \ref{tab:probing:attention-predictor:colour} shows the similarities of the emerged languages on the 'CLEVR color' dataset.
Only two languages emerge that help the receiver to increase its predicted probability mass on the target object.
While the results are not as high as on the 'Dale' dataset, the agents still fare better than the baseline model without any communication.
This indicates that the sender is communicating meaningful information to the sender to solve the task.
This is however not visible in the results of the probing model.
None of the two emerged languages has high similarities with any salience order in English.
Interestingly, the lowest (but still very high) normalized losses of 95\% to 96\% are not based on the only discriminating attribute \emph{color}, but rather on the non-discriminating attributes \emph{shape} and \emph{size}.
That suggests that the agents make use of other underlying structural patterns than the defined attributes to communicate and solve the task.
However, these patterns seem to be difficult to exploit as only two configurations achieve it and both don't solve the task with high accuracy.
